### 2020-01-12 ChapterAppendices/AppendixExperimentalValidation.tex
Add files via upload
Initial set of documents

%!TEX root = ..\Thesis.tex
\chapter{AppendixExperimentalValidation} \label{chap:AppendixExperimentalValidation}

\section{Appendix - Experimental validation}

Proof of Lemma \ref{lem:lemma1}.

In the first case considered in the lemma,
the elements of the expected value $\mathbf{A} = \mathbb{E} \left\{ \mathbf{E} \mathbf{H} \mathbf{E} \right\}$ are
\[ a_{ij} = \mathbb{E} \left\{ \mathbf{E} \mathbf{H} \mathbf{E} \right\}_{ij} = \mathbb{E} \left\{ \mathbf{E}_i^\top \mathbf{H} \mathbf{E}_j \right\} = \mathrm{tr} \left( \mathbf{H} \ \mathbb{E} \left\{ \mathbf{E}_j \mathbf{E}_i^\top \right\} \right) , \]
where $\mathbf{E}_i^\top$ is the $i\text{-th}$ row and $\mathbf{E}_j$ is the $j\text{-th}$ column of $\mathbf{E}$, for $i = 1, \ldots, T-n$, and $j = 2, \ldots, n+1$.
The matrix $\mathbb{E} \left\{ \mathbf{E}_j \mathbf{E}_i^\top \right\}$ is the product of $\sigma_{\epsilon}^2$ times a matrix whose elements are 0 in the first column, 2 in the $(j-i-1) \text{-th}$ diagonal, and -1 in the $(j-i-2) \text{-th}$ and $(j-i) \text{-th}$ diagonals, and zeros elsewhere.
Using the notation of the second differential operator we express
\[ \mathbb{E} \left\{ \mathbf{E}_j \mathbf{E}_i^\top \right\} = \sigma_{\epsilon}^2 \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{2, j-i} \end{bmatrix}. \]

The other cases in the Lemma can be proven similarly.

### 2020-01-12 ChapterAppendices/AppendixRampInput.tex
Add files via upload
Initial set of documents

%!TEX root = ..\Thesis.tex
\chapter{AppendixRampInput} \label{chap:AppendixRampInput}

\section{Appendix - Ramp input}

The entries of the Jacobian matrix are the first order partial derivatives of the residual error $\mathbf{r}$ with respect to the optimization variables.
The state space representation of the weighing model allows to find the analytical expression of the Jacobian.
The partial derivative of the residual error $\mathbf{r}$ with respect to the optimization variable $a$ is 
\begin{equation} \mathbf{J}_a=\dfrac{\partial \mathbf{r}}{\partial a} = \dfrac{\partial \widehat{y}}{\partial a} = \begin{bmatrix} 1 & 0  \end{bmatrix} \dfrac{\partial \mathbf{x}}{\partial a} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x}_a \end{equation}
where we use $\mathbf{x}_a = \partial \mathbf{x}/ \partial a$ to simplify the notation. 
Now, from the derivative of the state equation, we have
\begin{equation} \begin{aligned} 
    & \dot{\mathbf{x}}_a = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_a 
    + \begin{bmatrix} 0 & 0 \\ \frac{k_{\mathrm{s}} t}{(a t + b + m)^2} & \frac{k_{\mathrm{d}} t - b - m}{(a t + b + m)^2} \end{bmatrix} \mathbf{x} 
    - \frac{t \ \delta(t)}{(a t + b + m)^2} \mathbf{x}_{\text{ini}}   . 
\end{aligned} \end{equation}
Then, the partial derivative of the error $\mathbf{r}$ with respect to the optimization variable $a$ results in an additional dynamic system.

By repeating the procedure, we obtain the partial derivatives with respect to $b$ as follows: 
\begin{equation} \begin{aligned} 
    & \dot{\mathbf{x}}_b = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_b 
    + \begin{bmatrix} 0 & 0 \\ \frac{k_{\mathrm{s}}}{(a t + b + m)^2} & \frac{a+k_{\mathrm{d}}}{(a t + b + m)^2} \end{bmatrix} \mathbf{x} 
 - \frac{\delta(t) }{(a t + b + m)^2}  \mathbf{x}_{\text{ini}} ,
 \\ & \mathbf{J}_b = \begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_b . 
\end{aligned} \end{equation}
The partial derivatives of the error $\mathbf{r}$ with respect to the initial conditions yield the following two Jacobians
\begin{equation} \begin{aligned}
    & \dot{\mathbf{x}}_{\mathbf{x}_{\text{ini,1}}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,1}}} + \begin{bmatrix} \frac{\delta(t)}{a t + b + m} \\ 0 \end{bmatrix} , \\
    & \mathbf{J}_{\mathbf{x}_{\text{ini,1}}} = \begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,1}}}.
\end{aligned} \end{equation}
and
\begin{equation} \begin{aligned}
    & \dot{\mathbf{x}}_{\mathbf{x}_{\text{ini,2}}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a+k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,2}}} + \begin{bmatrix} 0 \\ \frac{\delta(t)}{a t + b + m} \end{bmatrix} ,  \\
    & \mathbf{J}_{\mathbf{x}_{\text{ini,2}}}=\begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,2}}}.
\end{aligned} \end{equation}

The Jacobian matrix is constructed using the responses of the additional dynamic systems  
\begin{equation} \begin{aligned} \mathbf{J} &= \begin{bmatrix} \mathbf{J}_a & \mathbf{J}_b & \mathbf{J}_{\mathbf{x}_{\text{ini,1}}} & \mathbf{J}_{\mathbf{x}_{\text{ini,2}}} \end{bmatrix} \end{aligned} \end{equation}

### 2020-01-12 ChapterAppendices/AppendixStatisticalAnalysis.tex
Add files via upload
Initial set of documents

%!TEX root = ..\Thesis.tex
\chapter{AppendixStatisticalAnalysis} \label{chap:AppendixStatisticalAnalysis}

\section{Appendix - Statistical analysis}

\section*{Appendix 1. Derivation of bias and covariance expressions.}

The bias and covariance of the least-squares (LS) estimate (\ref{eqn:xhatexp}) are obtained using the mathematical expectation in the definitions (\ref{eqn:biasdef}) and (\ref{eqn:covdef}).
For an unstructured and uncorrelated EIV problem, the expected value, and the covariance of $\widehat{\mathbf{x}}$ are approximated by
\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E}  \right\} \mathbf{x}, \quad \text{and}  \\ 
\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{x} \mathbf{x}^\top + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \bm{\epsilon} \bm{\epsilon}^\top \mathbf{K} + \mathbf{K}^\top \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} \right\} \mathbf{Q}^{-1} \\ 
& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{x} \mathbf{x}^\top \\
& + \mathbf{x} \mathbf{x}^\top \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{Q}^{-1} - \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}^\top , \label{eqn:EC_Eu} \end{aligned} \end{equation} 
where we have considered the second order Taylor series approximation (\ref{eqn:TseriesExp}), and we have removed the terms of zero expected value, and the terms of order higher than 2.
After an elementwise evaluation of the corresponding expected values in (\ref{eqn:EC_Eu}), the expressions result in 
\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  + \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) = \mathbf{x}  +  \sigma_{\mathbf{E}}^2 \mathbf{Q}^{-1} \left( 2\mathbf{I} + 2n\mathbf{I} - T \mathbf{I} \right) \mathbf{x} , \quad \text{and} \\ 
\mathbf{C}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \mathbf{x} \mathbf{x}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \ \mathbf{Q}^{-1} \mathbf{x} \mathbf{x}^\top \mathbf{Q}^{-1} , \end{aligned} \end{equation}
from where equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) are obtained.

On the other hand, due to the correlation, the expressions that approximate the expected value of the LS estimate of the structured and EIV problem (\ref{eqn:min_ls}) have a different form:
\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  +  \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E}  \right\} \mathbf{x} \\ 
& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{E}^\top \bm{\epsilon} - \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} - \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\} , \quad \text{and}  \\ 
\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{x} \mathbf{x}^\top + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \bm{\epsilon} \bm{\epsilon}^\top \mathbf{K} + \mathbf{K}^\top \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} - \mathbf{K}^\top \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \mathbf{K} - \mathbf{K}^\top \bm{\epsilon} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} \right\} \mathbf{Q}^{-1}  \\
& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{x} \mathbf{x}^\top \\
& + \mathbf{x} \mathbf{x}^\top \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{Q}^{-1} - \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}^\top \\ 
& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{E}^\top \bm{\epsilon} - \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} - \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\} \mathbf{x}^\top \\
& + \mathbf{x} \mathbb{E} \left\{ \bm{\epsilon}^\top \mathbf{E} - \bm{\epsilon}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} - \bm{\epsilon}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} \right\} \mathbf{Q}^{-1}. \label{eqn:EC_E} \end{aligned} \end{equation} 
These expressions have the non zero expected value terms, up to the second order. 
We have then
\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & = \mathbf{x} + \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \approx \mathbf{x}  +  \mathbf{Q}^{-1} \mathbf{K}^\top \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} \right\}}_{\mathbf{B}_1} - \mathbf{Q}^{-1} \underbrace{\mathbb{E} \left\{ \mathbf{E}^\top \left( \mathbf{I} - \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \right) \mathbf{E} \right\}}_{\mathbf{B}_2} \mathbf{x} \\ 
& - \mathbf{Q}^{-1} \mathbf{K}^\top \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\}}_{\mathbf{B}_3} + \mathbf{Q}^{-1} \underbrace{\mathbb{E} \left\{ \mathbf{E}^\top \left( \mathbf{I} - \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \right) \bm{\epsilon} \right\}}_{\mathbf{B}_4} , \quad \text{and} \\ 
\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{Q}^{-1} \mathbf{K}^\top \left( \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \bm{\epsilon}^\top \right\}}_{\sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n}}  + \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_1}  - \underbrace{\mathbb{E} \left\{  \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \right\}}_{\mathbf{C}_2} - \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \mathbf{x}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_2^\top} \right) \mathbf{K} \mathbf{Q}^{-1}  - \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\mathbf{x}} \right). \end{aligned} \end{equation} 
from where the expressions (\ref{eqn:biasE}) and (\ref{eqn:varE}) are obtained.

\section*{Appendix 2. Proof of Lemma \ref{lem:lemma1}}

\begin{pf}
In the first case considered in the lemma, the elements of the expected value $\mathbf{Z} = \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \mathbf{E} \right\}$ are
\begin{equation} z_{ij} = \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \mathbf{E} \right\}_{ij} = \mathbb{E} \left\{ \mathbf{e}_i \mathbf{A} \mathbf{E}_j \right\} = \mathrm{tr} \left( \mathbf{A} \ \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} \right) , \end{equation}
where $\mathbf{e}_i$, and $\mathbf{E}_j$ are the $i\text{-th}$ row, and the $j\text{-th}$ column of $\mathbf{E}$, for $i = 1, \ldots, T-n$, and $j = 2, \ldots, n+1$.
The matrix $\mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\}$ is the product of $\sigma_{\bm{\epsilon}}^2$ times a matrix whose elements are 0 in the first column, 2 in the $(j-i-1) \text{-th}$ diagonal, and -1 in the $(j-i-2) \text{-th}$, and $(j-i) \text{-th}$ diagonals, with zeros elsewhere.
By using the definition of the second differential operator, we express
\begin{equation} \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} = \sigma_{\bm{\epsilon}}^2 \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{2, j-i} \end{bmatrix}. \end{equation}

The proof of the other cases in the Lemma is similar. 
For the second case, the elements of the expected value $\mathbf{Z} = \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \mathbf{E} \right\}$ are
\begin{equation} z_{ij} = \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \mathbf{E} \right\}_{ij} = \mathbb{E} \left\{ \mathbf{e}_i \mathbf{A} \mathbf{E}_j \right\} = \mathrm{tr} \left( \mathbf{A} \ \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} \right) , \end{equation}
where now $\mathbf{e}_i$ is the $i\text{-th}$ row of $\mathbf{E}^\top$, and $\mathbf{E}_j$ is the $j\text{-th}$ column of $\mathbf{E}$, for $i = 2, \ldots, n+1$, and $j = 2, \ldots, n+1$.
The matrix $\mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\}$ is $\sigma_{\bm{\epsilon}}^2$ times a matrix whose elements are 2 in the $(j-i) \text{-th}$ diagonal, and -1 in the $(j-i-1) \text{-th}$ and $(j-i+1) \text{-th}$ diagonals, with zeros elsewhere.
Therefore we have
\begin{equation} \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} = \sigma_{\bm{\epsilon}}^2 \mathbf{D}_{T-n \times T-n}^{2,j-i+1}. \end{equation}

The expected values that involve the vector $\bm{\epsilon}$ are especial cases of the previous cases. 
The vector $\bm{\epsilon}$ in the expected values $\mathbb{E} \left\{ \mathbf{E} \mathbf{A} \epsilon \right\}$, $\mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \epsilon \right\}$, and $\mathbb{E} \left\{ \mathbf{E} \mathbf{A} \bm{\epsilon}^\top \right\}$ is
\begin{equation} \bm{\epsilon} = \begin{bmatrix} \epsilon(n+1) & \epsilon(n+2) & \cdots & \epsilon(T) \end{bmatrix}^\top, \end{equation}
as it is imposed by the input estimation method formulation.
\end{pf}

### 2020-01-12 ChapterAppendices/ChapterAppendices.tex
Add files via upload
Initial set of documents

--- 2020-01-12 ChapterAppendices/AppendixStatisticalAnalysis.tex
@@
 %!TEX root = ..\Thesis.tex
-\chapter{AppendixStatisticalAnalysis} \label{chap:AppendixStatisticalAnalysis}
+\glsresetall
 
-\section{Appendix - Statistical analysis}
+\chapter{Appendices }\label{chap:Appendices}
+% \begin{quote}
+% \emph{The results of this chapter are based on the \cite{VanNechel2018}.}\vfill{}
+% \end{quote}
+\vfill{}
 
-\section*{Appendix 1. Derivation of bias and covariance expressions.}
+\section{Appendix - Jacobians in Ramp input ML estimation}
+
+The entries of the Jacobian matrix are the first order partial derivatives of the residual error $\mathbf{r}$ with respect to the optimization variables.
+The state space representation of the weighing model allows to find the analytical expression of the Jacobian.
+The partial derivative of the residual error $\mathbf{r}$ with respect to the optimization variable $a$ is 
+\begin{equation} \mathbf{J}_a=\dfrac{\partial \mathbf{r}}{\partial a} = \dfrac{\partial \widehat{y}}{\partial a} = \begin{bmatrix} 1 & 0  \end{bmatrix} \dfrac{\partial \mathbf{x}}{\partial a} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x}_a \end{equation}
+where we use $\mathbf{x}_a = \partial \mathbf{x}/ \partial a$ to simplify the notation. 
+Now, from the derivative of the state equation, we have
+\begin{equation} \begin{aligned} 
+    & \dot{\mathbf{x}}_a = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_a 
+    + \begin{bmatrix} 0 & 0 \\ \frac{k_{\mathrm{s}} t}{(a t + b + m)^2} & \frac{k_{\mathrm{d}} t - b - m}{(a t + b + m)^2} \end{bmatrix} \mathbf{x} 
+    - \frac{t \ \delta(t)}{(a t + b + m)^2} \mathbf{x}_{\text{ini}}   . 
+\end{aligned} \end{equation}
+Then, the partial derivative of the error $\mathbf{r}$ with respect to the optimization variable $a$ results in an additional dynamic system.
+
+By repeating the procedure, we obtain the partial derivatives with respect to $b$ as follows: 
+\begin{equation} \begin{aligned} 
+    & \dot{\mathbf{x}}_b = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_b 
+    + \begin{bmatrix} 0 & 0 \\ \frac{k_{\mathrm{s}}}{(a t + b + m)^2} & \frac{a+k_{\mathrm{d}}}{(a t + b + m)^2} \end{bmatrix} \mathbf{x} 
+ - \frac{\delta(t) }{(a t + b + m)^2}  \mathbf{x}_{\text{ini}} ,
+ \\ & \mathbf{J}_b = \begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_b . 
+\end{aligned} \end{equation}
+The partial derivatives of the error $\mathbf{r}$ with respect to the initial conditions yield the following two Jacobians
+\begin{equation} \begin{aligned}
+    & \dot{\mathbf{x}}_{\mathbf{x}_{\text{ini,1}}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,1}}} + \begin{bmatrix} \frac{\delta(t)}{a t + b + m} \\ 0 \end{bmatrix} , \\
+    & \mathbf{J}_{\mathbf{x}_{\text{ini,1}}} = \begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,1}}}.
+\end{aligned} \end{equation}
+and
+\begin{equation} \begin{aligned}
+    & \dot{\mathbf{x}}_{\mathbf{x}_{\text{ini,2}}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a+k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,2}}} + \begin{bmatrix} 0 \\ \frac{\delta(t)}{a t + b + m} \end{bmatrix} ,  \\
+    & \mathbf{J}_{\mathbf{x}_{\text{ini,2}}}=\begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,2}}}.
+\end{aligned} \end{equation}
+
+The Jacobian matrix is constructed using the responses of the additional dynamic systems  
+\begin{equation} \begin{aligned} \mathbf{J} &= \begin{bmatrix} \mathbf{J}_a & \mathbf{J}_b & \mathbf{J}_{\mathbf{x}_{\text{ini,1}}} & \mathbf{J}_{\mathbf{x}_{\text{ini,2}}} \end{bmatrix} \end{aligned} \end{equation}
+
+\section{Appendix - Derivation of bias and covariance expressions.}
 
@@
 
-\section*{Appendix 2. Proof of Lemma \ref{lem:lemma1}}
+\section{Appendix - Proof of Lemma \ref{lem:lemma1}}
 
@@
 \end{pf}
+
+\newpage

### 2020-01-12 ChapterConclusion/ChapterConclusion.tex
Add files via upload
Initial set of documents

%!TEX root = ..\Thesis.tex
\glsresetall

\chapter[Conclusions and future work]{Conclusions and \linebreak future work} \label{chap:concl}

\vfill{}

In this final chapter, we first summarize the contributions made to both parts of the thesis: \textbf{Part \ref{part:Filter}} and \textbf{Part \ref{part:PA}}. Then, we propose
some possible further improvements to extend this work in future research.

\newpage
\section*{Conclusions}

\section{Conclusions - Ramp input}

An adaptive subspace method was proposed for the estimation of affine input parameters given the measurement of the caused sensor transient response. 
The subspace estimation method is a recursive method that allows online implementation.
This method tracks the input of a system, using exponential forgetting, to process the system response.
The subspace method is model-free and estimates directly the input parameters without identifying a sensor model.
Therefore, it can be applied to the measurement of different physical magnitudes.
In the specific weighing example described in the manuscript, the input is an affine function.
The method is also applicable when the sensor is time-varying.
The subspace method is computationally cheap, simple and suitable for implementation on digital signal processor of low computational power.

A maximum-likelihood estimator based on local optimization was designed to obtain a comparative reference for the other methods.
The maximum-likelihood method estimates the affine input parameters and also model parameters and the sensor's initial conditions.
This method simulates, in a receding horizon scheme, the response of a sensor model to estimate the input and minimizes the sum of the squares of the residual between the measured and the estimated responses.
The main drawback of the maximum-likelihood method is its computational cost and efficient implementation of the method is left for future work.

A linear time-invariant weighing system is used as an test example for the estimation methods.
The weighing system becomes time-varying when an affine input excites the system.
The estimation methods are compared in a simulation study where the time-varying sensor response is perturbed by measurement noise, that is assumed white of zero mean and known variance.
The subspace method results are also compared to those of an existing digital time-varying filter.
The coefficients of the time-varying filter require offline optimization.
The estimation results obtained with the subspace method converges two times faster and is one order of magnitude smaller than those obtained with the time-varying filter.
The empirical mean squared errors of the subspace method estimation is two orders of magnitude larger than the theoretical minimum given by the Cr\'amer-Rao Lower bound.

Future work of this research is the practical implementation of the subspace method for real-time measurements.

\section{Conclusions - Statistical analysis}

 We conducted a statistical analysis of a structured errors-in-variables (EIV) estimation problem with correlation to find the first and second moments of its least-squares solution.
This estimation problem occurs in metrology when we estimate the value of a measurand directly from the sensor transient response.
The data-driven estimation of the physical quantity is formulated as a structured EIV problem with correlation that uses the observed transient response to construct both the regression matrix and the regressor.
The real-time implementation of the method uses a recursive least squares algorithm that is simple and has low computational complexity.
The assessment of the uncertainty is done using the estimate bias and variance.

The conducted statistical analysis produced expressions that predict the estimate bias and variance for given sample size and perturbation level of the observed response.
The Monte Carlo simulation validated the predictions.
We compared the results of solving an unstructured and uncorrelated EIV problem with a structured and correlated EIV problem to understand how the structure and the correlation impacts in the estimation.
We found that the predictions in the structured case are more susceptible to perturbations.
This is due to the two approximations involved, a second-order Taylor series expansion of the estimate, and the substitution of perturbed data on the prediction expressions.
The relative error results indicate that the estimate bias, and variance are predicted using the derived expressions, and the observed data.
The mean squared error of the estimate is close to the results of the maximum likelihood estimate given by the Cram\'er-Rao lower bound.

The bias and variance can be accurately predicted, provided that the Taylor series expansion is valid.
This constraint has to be taken into account to ensure the effectiveness of the method in practical applications.
In the example, it was observed that when the SNR lies outside the validity region, the bias and variance estimation was at most three times larger than the empirical values.

The methodology presented in this paper can be applied to estimate the uncertainty of the solutions to other structured EIV problems.
The bias and variance expressions obtained after the statistical analysis depend on each specific structure.

\section{Conclusions - Experimental validation}

In this paper we investigated the statistical properties of a data-driven step input estimation method in a real-life application.
The step input estimation method is a structured and correlated errors-in-variables problem that is solved with recursive least squares.
A statistical analysis was conducted using the ordinary least squares condensed notation. 
This statistical analysis of the input estimate provides expressions that approximate the estimation bias and variance assuming that the measurement noise is Gaussian white noise. 
The variance approximation is useful to assess the uncertainty of the input estimate.
In simulation we observed that the mean squared error of the input estimate is close to the theoretical minimum that uses the Cr\'amer-Rao lower bound for biased estimators.
Since the data-driven step input estimation method is not statistically efficient, there is room for improvement. This is a topic for future research.
In the practical experiments, the measurement noise is not white.
The noise variance obtained from the sensor steady state response underestimates the measurement noise variance, that was observed 5 dB larger in the power spectrum due to nonlinearities of the sensor.
Considering this difference in the measurement noise variance, we introduced a conservative bound of the measurement noise variance so that the first and second moments of the input estimate are more accurately predicted.
Using the variance approximation, we can assess the uncertainty of the input estimate with respect to the number of samples processed by the data-driven step input estimation method.
The step input estimation method is useful in practical applications where the whiteness assumption of the measurement noise is not fulfilled.

\section*{Future work}
text

### 2020-01-12 ChapterExperimentalValidation/ChapterExperimentalValidation.tex
Add files via upload
Initial set of documents

%!TEX root = ..\Thesis.tex
\glsresetall

\chapter{Experimental validation of the step input estimation method }\label{chap:ExperimentalValidation}
% \begin{quote}
% \emph{The results of this chapter are based on the \cite{VanNechel2018}.}\vfill{}
% \end{quote}
\vfill{}

 \section{Introduction  - Experimental validation}

 In this paper we consider that a measurement is a dynamic process, where an input excites a dynamic system, the sensor, and causes a dynamic transient response that also depends on the initial conditions of the sensor.
The to-be-measured quantity is an unknown input that excites the sensor.
The consequent transient response is further processed to estimate quickly the measurand value.
The steady-state response of the sensor, that exists after the stabilization of dynamic effects, gives easy access to the measurand value but this approach is mainly exploited for calibration purposes.

A compensator is an additional dynamic system that acts on the transient response aiming to reduce the sensor transient time.
The compensation is motivated by the need of inverting the sensor dynamic effects to recreate the input.
The convolution of the compensator impulse response with the sensor transient response yields the input estimate.
Therefore, the design of a compensator is based on the sensor model and requires a deconvolution \cite{Eichstadt10}.
Examples of input estimation using compensation of the sensor transient response include a recursive estimation of the compensator parameters \cite{Shu93}, 
finite impulse response (FIR) \cite{Elster07, Niedzwiecki16b} filters and 
infinite impulse response (IIR) filters \cite{Pintelon90, Elster08}.
The filters in these works estimate in real-time the unknown input value.

An alternative to the compensation approach is to use digital signal processing methods that are independent of the sensor model.
A data-driven method that estimates the unknown level of step inputs by processing the sensor step response was introduced in \cite{Markovsky15ieee}. 
This data-driven input estimation method avoids the sensor modeling stage and estimates directly the input.
This method reduces the estimation time compared to a conventional compensator.
The step input estimation method performance was demonstrated by simulations and experiments on a digital signal processor (DSP) of low cost \cite{Markovsky15cep}.
The uncertainty of the step input estimation method has not been assessed before.

To validate the input estimation methods it is necessary to assess the uncertainty associated with their estimates \cite{daSilva12, Ferrero06}.
There are uncertainty propagation studies for model-based compensators such as the FIR and IIR filters for acceleration measurements where the uncertainty is computed in real time \cite{Elster07, Elster08, Link09}.
In these works, the uncertainty expression is based on the transfer function or state space representations of the LTI sensor and filter systems.
Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \cite{Pietrzak14} for mass and in \cite{Ogorevc16} for temperature sensors.
A deconvolution method is implemented to estimate the input waveform in \cite{Hale09} and the uncertainty is obtained from the input estimate covariance. 
The impact that the signal processing data-driven dynamic error correction has on the uncertainty is investigated in \cite{Saggin01}. 
A statistical analysis of the data-driven step input estimation method \cite{Markovsky15cep} was investigated in \cite{Quintana19} and the method uncertainty was obtained with a Monte Carlo simulation study.

This paper provides an uncertainty assessment of the data-driven step input estimation method in a real-life application.
The measurements were conducted in a weighing system based on a load cell sensor.
We observed that even when the whiteness assumptions of the measurement noise are not fulfilled, the step input estimation method still is able to provide a good estimation. 
We found that the mean squared error of the input estimate is near the Cram\'er-Rao lower bound of the EIV problem.
A confidence interval is provided for the input estimate in terms of the number of samples required to satisfy the accuracy specifications of the user.

The novelty of the paper is threefold.
First, using the results of \cite{Quintana19}, that describes a statistical analysis of structured errors in variables (EIV) problems, in this paper we describe the statistical properties of a the data-driven step input estimation method in both simulation and a real-life experiments.
Second, this manuscript also presents the Cram\'er-Rao lower bound for unbiased estimators of the structured and correlated EIV problem that the step input estimation method formulates.
Using this bound we have the minimum mean-squared error (MSE) for this estimation problem that we use compare with the MSE computed from the predictions obtained after the statistical analysis.
The third novelty in this manuscript is the model order selection for the step input estimation method where we use the MSE to select the order that provides the smaller MSE with the lowest computational complexity.

\section{SIMULATION RESULTS}

A Monte Carlo (MC) simulation was conducted to test the bias and covariance expressions (\ref{eqn:biasST}) and (\ref{eqn:varST}).
The MC simulation performed $N_{MC} = 10^4$ runs of the data-driven step input estimation with different realizations of the measurement noise $\bm{\epsilon}$.
We are interested in the first element of $\widehat{\mathbf{x}}$, which is the input estimate $\widehat{u}$.
The measurement noise variance was selected to have a signal-to-noise ratio (SNR) in the interval [30 dB, 80 dB], according to
\begin{equation} \mathrm{SNR} = 20 \log_{10}{ \dfrac{ \sqrt{ \dfrac{1}{T} \int\limits_{0}^{T}{ y(\tau)^2  \mathrm{d} \tau } } }{ \sigma_{ \epsilon }} } \end{equation}

The MC simulation was conducted processing $T = 5000$ samples of a simulated transient step response $\widehat{\mathbf{y}}$ generated by a stable LTI system of order $n = 5$, with a sampling frequency of $f_s=4$ kHz.
This system is a state space model obtained with the System Identification Toolbox using the measured step response of the actual sensor described in the Practical Implementation Section.
The steady state response of the system is practically reached after 500 samples because from there on the relative error between the transient response and the steady-state response is smaller than 0.2\%, see Fig. \ref{fig:uh_sim}.

A sensor is a dynamic system, therefore, a fast measurement process must necessarily cope with the system’s transient response. In that respect we must distinguish between the transient response of the system under test, and the transient phase of the measurement process (i.e. before the process has settled on a final measurement outcome). Notice that the transient phase of the measurement process is considerably smaller than the settling time of the system under test, and this is a major advantage of the step input estimation method.

\begin{figure}[!htb]
\centering
\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_1.pdf} 
\caption{ \label{fig:uh_sim} 
Above: example of a simulated response $\widehat{\mathbf{y}}$ and its step input estimation $\widehat{u}$ assuming measurement Gaussian noise with 50 dB of SNR. 
Below: the relative error $|\widehat{u} - u| / u$ is below 1\% after 100 samples. 
We take the estimate at 500 samples because there the relative error is smaller than 0.2\%.  }
\end{figure}

The difference between the sample mean $\widehat{\mu}_u$ of the step input estimates and the true value $u$ is the empirical bias $b_\mathrm{e}$.
\begin{equation} {b}_\mathrm{e} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \widehat{u}_i - u } = \widehat{\mu}_u - u.  \end{equation}

The sample variance $\widehat{\sigma}_u^2$ of the step input estimates is used to obtain the standard error of the MC simulation $\sigma_\mathrm{e}$, which decreases with respect to the square root of the number of MC runs $N_{MC}$. 
\begin{equation}  \sigma_\mathrm{e} = \frac{\widehat{\sigma}_u}{\sqrt{N_{MC}}}, \ \mathrm{where} \ \ \widehat{\sigma}_u^2 = \frac{1}{N_{MC}-1} \sum_{i=1}^{N_{MC}}{ \left( \widehat{u}_i - \widehat{\mu}_u \right)^2 } . \end{equation}

In each of the $N_{MC}$ runs, we compute the predictions of the step input estimation bias and variance from measured data using Equations (\ref{eqn:biasST}) and (\ref{eqn:varST}). 
The step input bias and variance predictions from observed data $b_{\widetilde{\mathrm{p}}}$ and $v_{\widetilde{\mathrm{p}}}$, and the associated standard error $\sigma_{\widetilde{\mathrm{p}}}$, are obtained from
\begin{equation} \begin{aligned} & b_{\widetilde{\mathrm{p}}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathbf{b}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\mathbf{x}} \right) \big|_{\left[1\right]} }, \ v_{\widetilde{\mathrm{p}}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\mathbf{x}} \right) \big|_{\left[1,1\right]} },  \\ & \mathrm{and} \quad \sigma_{\widetilde{\mathrm{p}}} = \sqrt{   \frac{ \sum_{i=1}^{N_{MC}}{ \left( \mathbf{b}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\mathbf{x}} \right) \big|_{\left[1\right]} - b_{\widetilde{\mathrm{p}}} \right)^2 } }{N_{MC}\left( N_{MC}-1 \right)}  } , \end{aligned} \end{equation}
where $\widetilde{\mathbf{b}}_{\mathrm{p}}^i \left( \widehat{\mathbf{x}} \right) \big|_{\left[1\right]}$ is the first element in the bias vector and $\widetilde{\mathbf{C}}_{\mathrm{p}}^i \left( \widehat{\mathbf{x}} \right) \big|_{\left[1,1\right]}$ is the first element in the covariance matrix obtained in the $i-\mathrm{th}$ approximations.
The predicted bias and variance from exact data are obtained with one evaluation of the expressions
(\ref{eqn:biasE}) and (\ref{eqn:varE})
\begin{equation} b_{\mathrm{p}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \big|_{\left[1\right]} }, \ \mathrm{and} \ v_{\mathrm{p}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathrm{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \big|_{\left[1,1\right]} } . \end{equation}
The uncertainty of the step input estimate is defined as the spread of the estimates that is given by the predicted variance $v_{\mathrm{p}}$.

Fig. \ref{bias_sigma_sim_unstr_str_n2} shows the empirical bias, the bias predictions and the standard errors of the MC simulation.
It can be seen that the empirical bias $b_\mathrm{e}$ and the predicted bias $b_{\widetilde{\mathrm{p}}}$ are proportional to the perturbation noise variance while the standard errors $\sigma_\mathrm{e}$ and $\sigma_{\widetilde{\mathrm{p}}}$ are proportional to the perturbation noise standard deviation.
For SNR below 40 dB there is a difference of a small order of magnitude between the empirical bias $b_\mathrm{e}$ and the bias prediction $b_{\widetilde{\mathrm{p}}}$.

The standard errors of the MC simulation $\sigma_\mathrm{e}$ and $\sigma_{\widetilde{\mathrm{p}}}$ are smaller than  $b_\mathrm{e}$ and $b_{\widetilde{\mathrm{p}}}$.
The estimates are spread near the sample mean and the uncertainty is smaller than the bias. 
Therefore, the empirical bias of the MC simulation is meaningful.

\begin{figure}[!htb]
\centering
\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_2.pdf} 
\caption{ \label{bias_sigma_sim_unstr_str_n2} 
The results of the Monte Carlo simulation of the step input estimation method are the empirical bias $b_{\mathrm{e}}$, the predicted bias using exact data $b_{\mathrm{p}}$, the predicted bias using measured data $b_{\widetilde{\mathrm{p}}}$, the empirical standard error $\sigma_{\mathrm{e}}$, and the predicted standard error $\sigma_{\widetilde{\mathrm{p}}}$. The estimation biases are proportional to the perturbation variance and the estimation standard errors are proportional to the perturbation standard deviation. Since the standard errors are smaller than the biases, the MC simulation is meaningful.  }
\end{figure}

The mean squared error (MSE) of the step input estimate, defined as
\begin{equation} \mathrm{MSE} = b^2 + v,   \end{equation}
where $b$ and $v$ are the bias and the variance of the step input estimate, can be applied to the obtained empirical and predicted results and can be compared to the Cram\'er-Rao lower bound (CRB).
Fig. \ref{fig:MSE_CRB} shows that $\mathrm{MSE}_\mathrm{e} = b_\mathrm{e}^2 + v_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}  = b_{\widetilde{\mathrm{p}}}^2 + v_{\widetilde{\mathrm{p}}}$ have the same proportionality with respect to the measurement noise variance as the bound for an unbiased estimator $\mathrm{CRB_{ub}}$.
For SNR above 35 dB, $\mathrm{MSE}_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ are equivalent, and below 35 dB the difference between them is of less than a factor of 10.

We obtained an approximation of the $\mathrm{CRB}_{\mathrm{b}}$ for our biased estimator using the partial derivative of the bias in expression (\ref{eqn:biasST}).
Fig. \label{fig:MSE_CRB} shows that the bounds for the unbiased and biased estimators are almost equal because the partial derivatives of the bias are negligible w.r.t. 1 in Equation (\ref{eqn:CRB_EIV}).
By adding the square of the predicted bias to the biased estimator bound $\mathrm{CRB_{b}}$ we obtain an approximation of the minimum MSE that the biased estimator can achieve.
This minimum MSE is close to the CRBs for large SNR but the square of the bias causes an increase of the MSE around 35 dB.
The differences between the CRBs and $\mathrm{MSE}_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ are of one order of magnitude for large SNR and become small for SNR lower than 40 dB.
This difference is the cost of solving a structured EIV problem with a simple LS method.
%Therefore, for large measurement noise variance, the step input estimation method produces results that are comparable to maximum-likelihood estimation, since $\mathrm{MSE}_{\widetilde{\mathrm{p}}} \approx \mathrm{CRB_{b}} + b_{\widetilde{\mathrm{p}}}^2$.

\begin{figure}[!htb]
\centering
\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_3.pdf}
\caption{\label{fig:MSE_CRB} The observation instant is fixed at 500 samples.
The bias and the MSEs decrease for large SNR. 
The empirical $\mathrm{MSE_e}$ and the predicted $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ of the step input estimation are one order of magnitude larger than the Cram\'er-Rao lower bound.
Adding the CRB for a biased estimator with the predicted bias squared we have the minimum MSE, that grows in the interval [25 dB, 45 dB].
Below 35 dB the difference between $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ and the CRB is less than a factor of 10.}
\end{figure}

There are two features of the system step response that make the CRB small.
One is the measurement noise variance and the second is sample size.
The estimation from step response perturbed with small noise variance has lower uncertainty.
Also, using larger sample size to perform the estimation boils down to smaller estimation uncertainty.

In order to get more insight into the step input estimation method we conducted another simulation study. 
The step input estimation method assumes the order $n$ is given in (\ref{eqn:matrixK}).
In this simulation, the step input estimation method processed the step response generated by a $5-th$ order system using different values of $n$ in the interval from 2 to 100.

The step response is perturbed with Gaussian white noise with SNR values in the interval [20 dB, 80 dB].
For each order $n$ and SNR value, 100 step input estimations are performed from independent noise realizations. 
Fig. \ref{fig:msee_vs_n_sim} shows the average of the squared biases and the variances, and the MSEs of the input estimate using the first 500 samples.
It is evident that the estimation variance and MSE depend on the SNR.

Increasing the order $n$ is equivalent to adding more regressors in the regression problem.
It is well known that increasing the order $n$ causes a monotonic decrement of the estimation bias and increment of the variance.
This is the asymptotic behavior of the estimation statistical moments with respect to the number of regressors.
Nevertheless, the simulation results presented in Fig. \ref{fig:msee_vs_n_sim} show that the variance first increases for small values of $n$, followed by a decrement and finally after $n \approx 40$ the variances exhibit a slow and steady increment. 
This apparent contradiction does not prove the invalidity of the estimation method since the results presented correspond to a finite sample size and the asymptotic results cannot be applied.
The theoretical explanation of the estimation statistics for finite sample sizes is out of the scope of this paper.

There is a bias-variance tradeoff and the MSEs exhibit local minima with respect to $n$.
The principal contribution to the MSE is the squared bias for the smaller values of $n$ and the variance for the larger values of $n$. 
However, the higher orders do not produce overfitting since the MSEs do not grow fast and remain close to the minimum values.

The optimum value of $n$ is not necessarily equal to the order of the generating system and varies for each SNR.
According to the plots in Fig. \ref{fig:msee_vs_n_sim}, there are orders that provide local minima of the step input estimation MSEs.
From the right hand side of Fig. \ref{fig:msee_vs_n_sim}, the orders that give the first two MSE minima were identified and those values are listed in Table \ref{Tbl:orders}.
For each SNR, there is a first minimum at a low order and a second minimum at a high order.
For SNR of 30 and 40 dB, it is recommended to use the order that gives the first minimum since the MSE at the second minimum is less than one order of magnitude smaller than at the first minimum.
Depending on the requirements, the user can choose between the simplicity of an estimation with a low order or an estimation with higher computational complexity and a smaller MSE. 
In a calibration stage, during the setup of the estimation method, the user can search and set the order that enables the estimation method to provide a required MSE.

\begin{table}[!ht]
\caption{ Orders $n$ that provide local minima for the MSE of the step input estimate. It is recommended to use the order that gives the first minimum when there is a difference of small order of magnitude with respect to the MSE at the second minimum.} 
\label{Tbl:orders} 
\centering
\begin{tabular}{ |c|c c c c| } 
\hline
SNR [dB] & 30 & 40 & 50 & 60 \\
\hline
order at first minimum & 11 & 7 & 4 & 3 \\ 
order at second minimum & 40 & 35 & 40 & 31 \\
\hline
\end{tabular}
\end{table}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterExperimentalValidation/fig/Fig_4.pdf} 
\caption{ \label{fig:msee_vs_n_sim} 
The simulated step responses of a $\mathrm{5-}th$ order system are processed by the step input estimation method using different orders $n$ and independent noise realizations. 
The perturbation of the step responses is Gaussian white noise with SNRs of 30, 40, 50, and 60 dB.
The square of the empirical bias (solid) and the empirical variance (dashed) are shown on the left hand side and the MSE is shown on the right hand side, for $n$ between 2 and 100.
These results suggest that, during the setup of the estimation method, we have to search the order that gives the minimum MSE without increasing unnecessarily the complexity of the estimation method.  }
\end{figure}

% \begin{figure}
% \centering
% \includegraphics[width=0.5\columnwidth]{./msee_vs_n_sim.pdf}

%\end{figure}

\section{PRACTICAL IMPLEMENTATION} 
An experimental setup was constructed to test the step input estimation method.
The implementation is a weighing system that uses a load cell Tedea Huntleigh 1004 \cite{tedea1004}.
The maximum rating of the load cell is 600 g.
A cylindrical aluminium object of 138.32 g of mass was used to excite the load cell.
This value was found by calibration using a balance KERN PCB 200-2 that has an uncertainty of 0.01 g.
The step input excitation was provided by a magnet that holds and releases a mass from above the load cell.
The magnet is located sufficiently far from the load cell to avoid magnetic interference in the sensor response.

\ctikzset{bipoles/resistor/height=0.075}
\ctikzset{bipoles/resistor/width=0.25}
\begin{figure}[!htb]
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
\tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
\tikzstyle{damper}=[thick,decoration={markings,  
  mark connection node=dmp,
  mark=at position 0.5 with 
  {
    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
  }
}, decorate]
\tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]

% load cell
\node (M1) [minimum width=1.6cm,minimum height=0.4cm, xshift=-3cm, yshift=-1cm] {};

\node (ground1) at (M1.east) [ground,yshift=0.0cm,rotate=90,xshift=0.0cm,anchor=north] {};
\draw (ground1.north east) -- (ground1.north west);

\node at (M1.north) [fill=none, xshift=-0.75cm, yshift=0.750cm] (2.0cm, 1.50cm) {\small Mass};
\draw (M1.north) [-latex,thick]  ++(-0.75cm,0.5cm) -- +(0cm,-0.5cm);
\draw (-3.3750cm,-0.935cm) arc(30:330:1.25mm);
\draw (-2.625cm,-1.05cm) arc(210:510:1.25mm);
\draw (M1.west) [thick]  ++(0.4cm, 0.0750cm) -- +(0.80cm,0.0cm);
\draw (M1.west) [thick]  ++(0.4cm,-0.0750cm) -- +(0.80cm,0.0cm);

\node at (M1.south) [align=center, draw=none,fill=none, xshift=0.0cm, yshift=-0.5cm] (2.0cm, 1.50cm) {\small load cell \\ \small Tedea 1004};

% amplifier
\draw (-1.6,-0.0) -- (-1.6,-2.0) -- (1.2,-1) -- cycle;
    \draw (-1.5,-1) to[R] (-1.1,-0.6) to[R] (-0.7,-1);
    \draw (-1.5,-1) to[R] (-1.1,-1.4) to[R] (-0.7,-1);
\draw [-latex,thick]  ++(1.2,-1) -- +(0.8,0);

\node (filter) [minimum width=0.50cm,minimum height=0.50cm, xshift=-0.0cm, yshift=-1cm] {};
\draw [thick]  ++(-0.275,-0.9) -- +(0.275,0.0);
\draw [thick]  ++(0,-0.9) -- +(0.25,-0.250);
\node at (M1.east) [draw=none, fill=none, xshift=4.0cm, yshift=0.25cm] (2.0cm, 1.50cm) {$\widetilde{y}$};
% wire
\draw plot [smooth] coordinates {(-2.3,-0.8) (-2.2,-0.5) (-1.7,-0.5) (-1.65,-0.9) (-1.6,-1)};

\node at (M1.south) [align=center, draw=none,fill=none, xshift=3.75cm, yshift=-0.750cm] (2.0cm, 1.50cm) {\small amplifier and filter};

\node (ADC) [align=center, minimum width=1.2cm,minimum height=0.9cm, xshift=2.6cm, yshift=-1cm] {\small ADC \\ \small 16 bits};

\end{tikzpicture}
\caption{Diagram of the load cell and the conditioning amplifier that provide the sensor response.} \label{lcm}
\end{figure}

A two-stage linear conditioning amplifier performs amplification and filtering of the load cell signal.
The first stage is a precision instrumentation amplifier INA114 that has high common mode rejection ratio.
The second stage is a third-order low-pass Butterworth filter with cut-off frequency of 100 Hz.
The low-pass filter prevents the aliasing noise in the measured transient response.
The signal obtained from the conditioning amplifier is considered to be the response of the sensor.
The sensor responses to step excitations were sampled with a frequency of $f_s=4$ kHz, and therefore the Nyquist frequency is 2 kHz.
The step responses were collected and stored as datasets for further analysis.
The number of samples collected for each step response is $N=20000$.
For practical purposes, we consider that the last 10000 samples correspond to the steady state response.
%The datasets are available in the following repository: %[[https://www.dropbox.com/sh/0dlam0rqqiynyiq/AAAq0H9rFxpkQON4zkmPXPXXa?dl=0][dropbox folder]]. 
\begin{figure}[!htb]
\centering
\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_6.pdf} 
\caption{ \label{fig:uh_exp} 
Above: a typical measured sensor transient response $\widetilde{y}$ takes more than 1.25 s (5000 samples, $f_s = 4$ kHz) to converge to the steady state response. 
Below: the relative error of the input estimate $\widehat{u}$ is smaller than 0.2\% from 300 samples.
We consider that at $500$ samples the relative error of the estimate $\widehat{u}$ is small enough to consider that $\widehat{u}$ is close to its expected value $u$.  }
\end{figure}

The step input estimation method processed 100 measured sensor step responses, assuming the sensor is of $\mathrm{7-}th$ order.
Fig. \ref{fig:uh_exp} shows a typical measured transient response $\widetilde{y}$ and an example of the estimated input $\widehat{u}$.

The empirical bias $b_{\mathrm{e}}$ is the difference between the average of the 100 estimates $\widehat{u}$ and the mass calibration value \linebreak $u = 138.32$ g, at each instant of time, and the standard error $\sigma_\mathrm{e}$ is the standard deviation of the mean estimate of the responses processed, i.e.,
\begin{equation} \begin{aligned} \widehat{\mu}_{e} &= \frac{1}{100} \sum_{i=1}^{100}{ \widehat{u}_{i} }, \quad {b}_{\mathrm{e}} = \widehat{\mu}_{e} - u, \quad \mathrm{and} \\  \sigma_{\mathrm{e}} &= \frac{\widehat{\sigma}}{\sqrt{100}}, \quad \mathrm{where} \ \widehat{\sigma}^2 = \frac{1}{99} \sum_{i=1}^{100}{ \left( \widehat{u}_i - \widehat{\mu}_{e} \right)^2 } . \end{aligned} \end{equation}

The bias $\widetilde{b}_{\mathrm{p}}$ and variance $\widetilde{v}_{\mathrm{p}}$ predictions from the measured data were obtained by processing off-line the 100 measured sensor transient step responses with expressions (\ref{eqn:biasST}) and (\ref{eqn:varST}).
These expressions require the measurement noise variance $\sigma_\epsilon^2$ to obtain the bias and variance prediction.
One way to estimate the measurement noise variance is computing the variance of each sensor steady state response, see Fig. \ref{fig:y_ss}.
Later in this section we will explore another way to estimate the measurement noise variance.
Computing the noise variance from the steady state response we observed that the SNR of the measured step responses is 55 dB in average.

\begin{figure}[!htb]
\centering
\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_7.pdf}
\caption{ \label{fig:y_ss} 
From the sensor steady-state response an estimation of the measurement noise variance is obtained.}
\end{figure}

Fig. \ref{fig:b_sigma_exp_str} shows the empirical bias $b_{\mathrm{e}}$ and the standard error $\sigma_{\mathrm{e}}$ that result after processing the first $T=500$ samples of the 100 measured step responses $\widetilde{\mathbf{y}}$.
The standard error is smaller than the bias.
As it was observed in the MC simulation, this is the uncertainty of the estimation method.
The oscillations observed in the bias are mainly due to the transient response and not to the measurement noise.
The measurement noise effects are partially removed since we averaged the 100 transient responses, 
which is a small number compared with the $N_{MC}$ runs averaged in the simulation section.

It is expected that the empirical bias is large when a small number of samples is processed. 
The data-driven input estimation method is recursive and it is implemented in real-time.
The estimation errors decrease as more data is processed.

\begin{figure}[!htb]
\centering
\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_8.pdf} 
\caption{ \label{fig:b_sigma_exp_str} 
The results of estimating the step input level after processing 100 measured step responses are the empirical bias $b_{\mathrm{e}}$ and the empirical standard error $\sigma_{\mathrm{e}}$. 
The estimation bias and the estimation standard error decrease as more samples are processed. 
The estimation bias is affected by the transient effects of the sensor response.  
The values of $b_{\mathrm{e}}$ and $\sigma_{\mathrm{e}}$ provide the estimate accuracy and uncertainty for a given sample size.
}
\end{figure}

The measurement noise is not white since there is evidence of frequency components in the sensor steady state response that are observed as oscillations in Fig \ref{fig:y_ss}.
To get insight into the properties of the measured sensor response $\widetilde{\mathbf{y}}$, a $\mathrm{7-}th$ order model was identified from input-output data assuming that the input is a step of level $u$.
A response $\widehat{\mathbf{y}}$ was simulated from the identified model and the residual $\mathbf{r} = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}$ was obtained. 
We can observe these signals in the frequency domain using the discrete Fourier transform, that for the signal $\widetilde{\mathbf{y}}$ is defined as 
\begin{equation} \widetilde{Y}(f) = \dfrac{1}{\sqrt{N}} \sum_{k=0}^{N-1} \widetilde{y} \left( k \right) e^{-j2\pi k f / N} \label{eqn:FFT} \end{equation}
where $f = 1, \ldots, N/2$ are the frequency lines and $N$ is the total number of samples.
The power spectrum of the signal $\widetilde{\mathbf{y}}$ is given in decibels by $\widetilde{\mathbf{Y}}_\mathrm{dB} = 20 \log_{10}{ |\widetilde{\mathbf{Y}}| } $.
Figure \ref{fig:YYhR_n7} shows the corresponding power spectra of the sensor response $\mathbf{Y}_\mathrm{dB}$, the simulated response $\widehat{\mathbf{Y}}_\mathrm{dB}$, and the residual $\mathbf{R}_\mathrm{dB}$. 
There are frequency components near the main resonance peak in the magnitude spectrum of the residual.
The presence of frequency components near the main resonance peak is commonly found in mechanical devices.
The vibrations captured from the environment explain the accumulation of energy near the main resonance modes.

\begin{figure}[!htb]
\centering
\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_9.pdf}
\caption{ \label{fig:YYhR_n7} 
The power spectra of a measured response $\widetilde{\mathbf{Y}}_\mathrm{dB}$, a simulated response $\widehat{\mathbf{Y}}_\mathrm{dB}$, and the residual $\mathbf{R}_\mathrm{dB}$ is not flat and then the measurement noise is not white. 
The average of the residual power spectrum provides a conservative estimate of the measurement noise variance $\widehat{\sigma}_\epsilon^2$, represented with the dotted line.}
\end{figure}

Even when the residual $\mathbf{r}$ is not white, it provides an alternative way to estimate the measurement noise variance. 
The average of the residual power spectrum approximates the measurement noise variance as follows
\begin{equation} \widehat{\sigma}_\epsilon^2 \approx \dfrac{2}{N} \sum_{f = 1}^{N/2} \left| R \left( f \right) \right|^2 . \label{eqn:P_R} \end{equation}
The dotted line in Figure \ref{fig:YYhR_n7} indicates the $10 \log_{10} \left( \widehat{\sigma}_\epsilon^2 \right)$ level of the measurement noise variance estimated from the residual.
This level is higher than the mean value of the residual power spectrum $\mathbf{R}_\mathrm{dB}$ in the frequencies above 120 Hz.

Using the residual power spectra that correspond to the measured step responses, we obtained the measurement noise variance and the SNR for each experiment.
Figure \ref{fig:meas_SNR} shows the estimated SNRs from the residual power spectra. 
The SNR mean value is 50 dB.
Therefore, we assume that the SNR of the measured transient responses is 50 dB instead of 55 dB, as it was estimated from the steady state response.

\begin{figure}[!htb]
\centering
\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_10.pdf}
\caption{ \label{fig:meas_SNR} 
The mean value of the signal-to-noise ratios estimated from the residual power spectra is 50 dB.
We consider that this is the estimated SNR of the measured step responses.}
\end{figure}

The 5 dB difference provides a conservative bound since the bias and variance computed from 50 dB of SNR are higher than those obtained using the variance estimation from the steady-state response. 
Fig. \ref{fig:b_sigma_exp_str2} shows a comparison of the results obtained with both measurement noise variance estimations after processing the first $T=500$ samples of the step response $\mathbf{y}$.
Using expression (\ref{eqn:biasST}), the bias prediction  $b_{\widetilde{\mathrm{p}}2}$ obtained using an SNR of 50 dB approximates more closely the empirical bias than $b_{\widetilde{\mathrm{p}}1}$ obtained using an SNR of 55dB.
In accordance, the standard error of the bias predictions $\sigma_{\widetilde{\mathrm{p}}2}$ is larger than $\sigma_{\widetilde{\mathrm{p}}1}$ and is a conservative measure of the input estimation uncertainty.
In conclusion, using the noise variance estimated from the residual prevents underestimating the step input estimation uncertainty.

\begin{figure}[!htb]
\centering
\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_11.pdf} 
\caption{ \label{fig:b_sigma_exp_str2} 
Comparative view of the bias prediction using two different noise variance estimations. 
Estimating the variance from the step response residual gives a bias prediction $b_{\widetilde{\mathrm{p}}2}$ and a standard error $\sigma_{\widetilde{\mathrm{p}}2}$ that are slightly higher than using the noise variance estimated from the steady state response $\sigma_{\widetilde{\mathrm{p}}1}$ and $\sigma_{\widetilde{\mathrm{p}}1}$. 
The bias prediction $b_{\widetilde{\mathrm{p}}2}$ approximates better the empirical bias. 
The standard error $\sigma_{\widetilde{\mathrm{p}}2}$ provides a conservative value of the input estimation uncertainty. }
\end{figure}

We investigated another aspect of the step input estimation method performance when processing measured step responses.
The step input estimation method requires an assumption of the generating system order in the formulation of the estimation problem (\ref{eqn:min_ls}).
The estimation method performance is assessed under different assumptions of the values of $n$ in the interval from 2 to 10. 
For each value of $n$, 100 step input estimations are computed from measured transient responses and the empirical MSEs are compared.
Fig. \ref{fig:msee_vs_n_exp} shows that, similar to the observations made in the simulation study, the MSEs have two local minima at $n=7$ and $n=48$.
It is recommended to use $n = 7$ in the estimation method to provide a small estimation MSE without the higher computational complexity that $n=48$ implies.

\begin{figure}[!htbp]
\centering
\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_12.pdf} 
\caption{ \label{fig:msee_vs_n_exp} 
The hundred measured sensor step responses are processed by the step input estimation method for different values of the order $n$. 
The empirical squared bias (solid) and the empirical variance (dashed) are shown on the left and the empirical MSE on the right.
The MSE has a local minimum at $n=7$ and another at $n=48$.
It is recommended to use the estimation method with $n=7$ because at $n=48$ the decrement of the MSE is not significant.  }
\end{figure}

\section{Conclusions}

In this paper we investigated the statistical properties of a data-driven step input estimation method in a real-life application.
The step input estimation method is a structured and correlated errors-in-variables problem that is solved with recursive least squares.
A statistical analysis was conducted using the ordinary least squares condensed notation. 
This statistical analysis of the input estimate provides expressions that approximate the estimation bias and variance assuming that the measurement noise is Gaussian white noise. 
The variance approximation is useful to assess the uncertainty of the input estimate.
In simulation we observed that the mean squared error of the input estimate is close to the theoretical minimum that uses the Cr\'amer-Rao lower bound for biased estimators.
Since the data-driven step input estimation method is not statistically efficient, there is room for improvement. This is a topic for future research.
In the practical experiments, the measurement noise is not white.
The noise variance obtained from the sensor steady state response underestimates the measurement noise variance, that was observed 5 dB larger in the power spectrum due to nonlinearities of the sensor.
Considering this difference in the measurement noise variance, we introduced a conservative bound of the measurement noise variance so that the first and second moments of the input estimate are more accurately predicted.
Using the variance approximation, we can assess the uncertainty of the input estimate with respect to the number of samples processed by the data-driven step input estimation method.
The step input estimation method is useful in practical applications where the whiteness assumption of the measurement noise is not fulfilled.

\newpage

### 2020-01-12 ChapterIntroduction/ChapterIntroduction.tex
Add files via upload
Initial set of documents

%!TEX root = ..\Thesis.tex
\chapter{Introduction} \label{chap:Introduction}

% Context:Metrology One of the objectives in metrology is the estimation of the true value of physical quantities.
Measurements are dynamical processes.
Sensors are dynamical systems that interact with physical quantities and deliver electrical signals.
This interaction is the input to the sensor and causes energy transferences that modify the sensor state.
The sensor response is the output that depends on both the input applied and the sensor initial conditions.

% Context:Tradeoff speed vs accuracy
There is a trade-off between speed and accuracy in a measurement.
The input excitation drives a linear time invariant sensor into a transient state, that is followed by a steady state.
During the sensor transient state, the response does not represent directly the input, but 
in steady state, the sensor response is proportional to the input.
The input can be estimated accurately from the sensor steady state response using the sensor static gain.
However, waiting for the steady state is not advisable for practical applications that need fast measurements.
In these applications the input must be estimated during the sensor transient state.

% Context:Compensation systems, # Context:Digital signal processing
One approach to estimate the input is filtering the sensor transient response with another dynamical system that inverts the dynamics of the sensor.
The filter output is an input estimate that compensates the measurement time.
The transient duration of the compensation filter is smaller than that of the sensor.
The compensator is designed to deconvolve the sensor response and is based on a sensor model.
Another approach is using digital signal processors (DSP).
DSPs are versatile since they allow the implementation of methods that do not necessarily recreate dynamic systems, such as digital filters.
A suitable model-independent, data-driven, method in a DSP can provide faster input estimations than with the model-based compensators.

% Context:Data-driven step input estimation method
An example of a data-driven method is the direct estimation of the step input level from the sensor step response \cite{Markovsky15cep}.
This method formulates a Hankel structured errors-in-variables (EIV) problem with correlation.
The regression matrix has a block-Hankel structure.
The regression matrix and the regressor are constructed from the transient response perturbed by measurement noise of zero mean and given variance.
The method is implemented in real-time using a recursive least-squares (RLS) solution of the structured EIV problem.
% Avoiding the model identification from input-output data and directly estimating the input from the transient response reduces the input estimation time and makes data-driven input estimation methods suitable for real-time

% Context:Data-driven methodology
% The philosophy behind the estimation method is avoiding the explicit identification of a system model from input-output data and estimating directly the input from the transient response.
The main advantage of the data-driven input estimation method is that it does not identify the sensor model but instead it directly estimates the input. 
This differs from the classic two-stage methodology where the sensor model is first identified before the input is estimated.
In this method, the output-error (OE) problem is converted into an EIV problem that is harder to solve, but the RLS solution is easy to compute.
The range of application of the data-driven input estimation method is wide because it is independent of the sensor model.
The main disadvantage of the data-driven input estimation method is that its stochastic properties are not straightforwardly evident. 
It is more complex to find the stochastic properties of EIV problems when they have structure and correlation.
% Moreover, the online uncertainty assessment may not be feasible and we have to rely on confidence bounds.

% Problem under study: Dynamic measurements
% DM:Uncertainty assessment of data-driven input estimation methods
% DM:Experimental validation of the data-driven step input estimation method
The estimate uncertainty must be assessed to validate the estimation methods for metrology applications.
The uncertainty of the data-driven step input estimation method \cite{Markovsky15cep} is unknown.
The uncertainty can be defined in terms of the estimate bias and variance \cite{Pintelon12Book}, and these can be obtained by conducting an elementwise statistical analysis.
The validation of the step input estimation method requires also to demonstrate its effectiveness on real-life measurements.
In real-life experiments, the experimens were conducted with sensors of temperature and mass.
One challenge of using real-life data is that the noise may not fulfill the whiteness assumptions considered in the estimation problem formulation and in the statistical analysis.
The simulation and experimental results permit to compare the performance of the step input estimation method.

% DM:Estimation of affine input parameters
The ideas behind the step input estimation method raise the curiosity towards the design of estimation methods for other input models.
One of this input models is the ramp.
The ramp input estimation method is, in addition, motivated by the dynamic weighing that is performed in conveyor systems.
The dynamic weighing estimates the mass of materials or products during their transportation.
Ideally, when the conveyor belt transports the materials at a constant speed, i.e., the weighing sensor is excited with a ramp profile. 
The ramp is an affine input model that consists of two parameters, the slope and the interception.
The slope depends on the applied mass and can be used to estimate it from the transient response.
An adaptation of the step input estimation method can estimate the parameters of affine inputs.

\subsection{State of the art}
% StArt: Load cell sensors 
% experimental application is weighing
Weighing has been basic for the development of scientific and trade activites. 
The load cell is now a standard transducer for weight determination and also for the improvement of measurement techniques, such as the geometric approach to processing of load cell responses \cite{Kesilmis16}, the design of new conveyor machinery \cite{Yamani18}, and electronic truck scales \cite{Guo18}.

In safety studies, a six axis load cell is devised to quantify accelerations and impact forces exerted on a dummy \cite{Ballo16}.
In alternative energy developments the load cells are useful to measure the forces on the arms of a vertical axis wind turbine \cite{Rossander15}.
An academic study of the load that a structure withstands is conducted with strain gage load cells that confirms the numerical results and facilitates the design of complex shaped structures \cite{Olmi16}.
In sports, the performance of new instrumented crank mechanisms is fostered by the utilization of load cells in the characterization, analysis and validation design stages \cite{Casas16}.

The versatility of the load cells permits the physiological signal monitoring of the the heart and breathing rates \cite{Lee16}, clinical analysis of sleep quality \cite{Zahradka18} and the classification of the movement intensity of people while they are sleeping \cite{Alaziz17}. 
All of these experimental studies have load cells installed on bed setups.

% StArt:Compensation filters
The use of compensation filters for load cell sensor responses has been reported 
in \cite{Shu93}, where an adaptive digital filter was first explored,
in \cite{Jafaripanah05}, where an analog filter alternative was suggested,
in \cite{Hernandez06} where a recursive LS lattice adaptive filter improves tension forces measurement, 
in \cite{Boschetti13} where models of the weighing machine, load cells and accelerometers are exploited to remove the environmental vibrations effects, 
in \cite{Dienstfrey14} where the negative effects of the finite bandwidth of the measurement system response are diminished by regularized deconvolution, and
in \cite{Huang16}, where the noises that affect an electromagnetic-force-compensated load cell are removed with a set of filters.
Some guidelines to implement a compensation filter by deconvolution are described in \cite{Eichstadt10}, and
the synthesis of filters by mapping the sensor model parameters into the filter parameters and time reversed filtering is proposed in \cite{Hessling08a}.

% StArt:Digital signal processing estimation methods
% The use of a DSP in metrology motivated the data-driven step input estimation method \cite{Markovsky15cep}.
A short list of the DSP measurement methods for different physical quantities include 
the impact that the signal processing data-driven dynamic error correction has on the temperature uncertainty \cite{Saggin01},
the development of an electronic nose using bio-inspired signal processing techniques \cite{Jing16},
the impedance measurements for material damage estimation using cross-correlation signal processing techniques \cite{deCastro19},
the modulation quality measurement of microwave access systems is measured using digital signal processing \cite{Angrisani10}, and
the real-time rotational speed estimation using correlation signal processing techniques \cite{Wang14}.

% StArt:Uncertainty assessment 
% Monte Carlo sampling
% Deterministic sampling
In metrology, the result of a measurement is a random variable.
The measurement is inherently perturbed by noise and, therefore, the input estimation is represented with the two first statistical moments \cite{Ferrero06}.
There is a guide to assess the uncertainty of the quantity estimation \cite{GUM08}, but there is still a need to study of the uncertainty of estimation methods and many approaches have been proposed \cite{Esward09, Hessling10}.
Some of the measurement uncertainty analysis are reviewed in \cite{daSilva12}.
The Monte Carlo method has been shown to be an effective uncertainty evaluation tool in \cite{Cox06},
and supports the usefulness of simulation for quantifying measurement uncertainties by software \cite{Esward16}.
Another example of the Monte Carlo method is the dynamic measurement uncertainty evaluation of clinical thermometers described in \cite{Ogorevc16}.

% 
It is recommended to consider the uncertainties of all the measurement chain components \cite{Diniz17}, and to avoid the direct uncertainty propagation from the calibration towards the to-be-measured quantity.  
Methods for evaluating the uncertainty associated with the output of compensation filters have been investigated, such as for
a discrete-time infinite-response filter in \cite{Link09},
a discrete-time finite-response filter in \cite{Elster07, Elster08}, and
the Kalman filter in \cite{Eichstadt16b}.
All these works propagate the uncertanity through the filter but it is also necessary to upward the propagation up to the sensor model to include all systematic error contributions \cite{Hessling11}.

%
An extension of the methodology proposed for the data-driven step input estimation was formulated to estimate the parameters of an affine input that changes at a constant rate,  by signal processing of the sensor transient response.
This type of ramp inputs is observed in the measurement of mass during the transportation of products on conveyor systems, ranging from few grams \cite{Burmen09} to almost hundreds of kilograms \cite{Tasaki07}.
The data-driven affine input estimation method is proposed as an alternative to existing compensation filters, such as 
the time-variant low-pass filters introduced in \cite{Piskorowski08, Pietrzak14}, and
the combination of filters in cascade proposed in \cite{Niedzwiecki16a}.

\subsection{Original contributions}
% OrigContrib:Affine input estimation
% OrigContrib:Statistical analysis of structured EIV problems
% OrigContrib:Experimental validation of the step input estimation method 
In this thesis are proposed two methods for estimating affine input parameters.
The first method is an adaptation of the data-driven input estimation method and the second is a maximum likelihood method based on local optimization.
The adaptation consists of the use of exponential weighing for the recursive least squares solution of the structured errors-in-variables problem.
The exponential weight gives preference to recent samples over the older samples.
This forgetting factor considers that the newer samples are more relevant for the parameters estimation.
The maximum-likelihood method method simulates the response of a sensor model to an affine input. 
The to-be-minimized cost function is the sum of the squared differences between the actual and the simulated sensor responses. 
The maximum-likelihood method needs more computational resources, and can simultaneously estimate parameters of the sensor model, but in it current formulation cannot run in real time.

In this document is presented a statistical analysis of the least squares solution of the structured errors-in-variables problem that the data-driven step input estimation method formulates.
The statistical analysis provides expressions that predict the bias and variance of the step input estimate for given sample size and measurement noise variance.
The second order Taylor series expansion of the LS solution obtain was derived to find expressions for the bias and variance of the estimate.
The Crámer-Rao lower bound of the structured EIV problem was obtained and it was compared to the empirical mean-squared error of the estimate.
It was observed that the input estimation is biased but with small variance, and the difference between the empirical MSE and the theoretical minimum and was determined quantitatively.

In this work a series of experiments was conducted to validate the step input estimation method in real-life applications.
The experiments were realizations of the step input excitation using temperature and mass sensors.
The weighing setup was constructed to ensure repeatability and reproducibility of the experimental realizations.
The step responses were stored in sets of 100 elements, and the empirical statistical moment of the input estimates were computed from the sample mean and variance of the estimations.
The step input estimation method showed robustness when the measurement noise is not Gaussian and white as it was assumed in the theoretical analyses.
The results of the estimation method with respect to different sensor model order were compared and we found that increasing the order does not necessarily benefits the input estimation uncertainty.

\subsection{Intro of Chapter ramp input}

Measurements estimate the unknown value of a physical quantity, namely the measurand.
The to-be-measured physical quantity is applied as an input signal to a sensor. 
The sensor is a dynamical system and its output changes as a consequence of the input excitation and the sensor initial conditions.
The goal of a measurement is to estimate accurately the measurand value using the sensor transient response.
The transient response of a stable sensor decays to a steady state response.
In steady state, the most accurate estimation of the input true value is simply found using the sensor static gain.
However, the steady state is reached in theory after an infinite period of time and in practice we require fast estimations.
The trade-off between accuracy and speed exists in all measurements.

The measurand can be assumed to be constant or variable during the measurement.
A dynamic measurement is present when the fluctuations of the measurand impact on the input estimation.
A typical example of a dynamic measurement problem is a low-bandwidth sensor excited with a fast changing input.
Some characteristics of the input, like the minimum or maximum or the effects of the environment on the measured quantity, occur in small periods of time.  
The detection of the input characteristics is needed in several scientific and industrial applications such as measurements of temperature \cite{Saggin01}, pressure \cite{Matthews14}, acceleration \cite{Link07}, force \cite{Vlajic16, Hessling08a} and mass \cite{Shu93, Boschetti13}.

The solution to dynamic measurement problems is non-trivial.
An approach is to add a dynamical system to compensate the sensor transient response, inverting the effects of the sensor dynamics.
The purpose of such a compensator is to reduce the transient time.
The sensor dynamics are considered in the design of finite and infinite impulse response compensation digital filters based on deconvolution \cite{Eichstadt10} or synthesized to correct dynamic errors \cite{Hessling08a}. 
The model-based deconvolution design of compensators implies that the measurand true value should be known a-priori for certain applications, such as mass determinations \cite{Boschetti13, Niedzwiecki16b}.
In the literature most of the measurements systems are assumed linear time-invariant, but the compensation digital filters can be linear \cite{Tasaki07},  nonlinear \cite{Shu93} or time-varying \cite{Pietrzak14}.

The digital signal processors enable a different approach where the input estimation can be obtained with algorithms that do not necessarily recreate the dynamics of a system.
One of the authors of this paper proposed a data-driven signal processing method that estimates the measurand true value using subspace techniques \cite{Markovsky15cep, Markovsky15ieee}.
The subspace estimation method bypasses the model identification step to estimate the unknown input directly from the response data.
This method was developed to estimate inputs modeled as step functions of unknown scaling level.
We extended the subspace input estimation method to estimate the parameters of inputs that vary at a constant rate.

The inputs that vary at a constant rate are found in applications where the measurand activates the sensor gradually. 
An example of this activation is the measurement of mass while the to-be-weighted object is transported by a conveyor belt, and the profile of the input is a saturated ramp.
Current solutions to the weighing in motion are low pass filters that estimate the mass using the saturated ramp \cite{Tasaki07, Pietrzak14}.
The signal processing affine input estimation methods are motivated by the need to obtain the mass of the object from the ramp before it reaches saturation.
The ramp is parameterized as a straight line model where the slope and the interception are the parameters of interest.

This paper describes a subspace method for the estimation of the affine input parameters.
This method is a recursive algorithm that can be implemented in real-time since it has low computational cost.
The subspace method is independent of the sensor model and, therefore, it is suitable for a variety of applications.
The dynamic weighing is one of the applications and was chosen as an implementation example.
The effectiveness of the method is evaluated in a simulation study.
The performance of the proposed method is compared to that of a maximum-likelihood (ML) estimation method based on local-optimization and a time-varying compensation filter.
%We found that the subspace method obtains the slope estimate from the sensor transient response using the same assumptions of the data-driven step input estimation method.

The ML method resembles the model predictive control approach in the sense that a cost function is minimized iteratively to optimize the parameters of a sensor model using the observed sensor response in a receding time horizon \cite{Mayne14}.
The difference is that the ML method aims to estimate the unknown value of the affine input parameters instead of identifying a model and controlling the dynamic system.
The ML method is more appropriate for off-line processing of the sensor transient response.
Nevertheless, the ML method can estimate the parameters of the affine input, the parameters of a sensor model, and the initial conditions of the sensor.

The uncertainty of the subspace method is assessed using a Taylor expansion of the estimate and Monte Carlo random sampling approach \cite{Quintana19}.
The Monte Carlo approach requires a large set of generated random samples, and for simple systems it is the recommended method.
There exists a deterministic sampling approach to study the uncertainty propagation of complex systems \cite{Hessling13a, Hessling13b}. 
Deterministic sampling aims to represent the minimal statistical information that is relevant to the uncertainty estimation in a finite set.
The uncertainty of the ML method is assessed using the derivatives of the residual error that constructs the to-be-minimized cost function.
The covariance of the optimization method estimate is found using the inverse of the Hessian matrix \cite{Pintelon12Book}.

\subsection{Intro of Chapter statistical analysis}

Errors-in-variables (EIV) are linear estimation problems in which the regression matrix and the regressor are perturbed \cite{VanHuffel91Book}, \cite{Markovsky07overview}.
In structured EIV problems, the regression matrix has a given structure that depends on the problem formulation.
Hankel, Toeplitz, or other application-specific matrices appear in problems of metrology \cite{Markovsky15cep}, system identification \cite{Soderstrom07}, image restoration \cite{Feiz17}, nuclear magnetic resonance spectroscopy \cite{Cai16}, direction-of-arrival estimation \cite{Pan18}, and time-of-arrival estimation \cite{Jia18}.

In metrology, the direct estimation of the input from the sensor transient response is formulated as a structured EIV problem.
The only observed signal is the sensor output.
To estimate a step input, the regression matrix, and the regressor are built from the step response observations \cite{Markovsky15cep}, and
the structure in the regression matrix is block-Hankel.
This data-driven estimation methodology reduces the estimation time of the classical two-stage approach where a sensor model is first identified, and later the input is estimated using the sensor model \cite{Azam15, Niedzwiecki16a}.
%The problem is EIV because the perturbation of the sensor response gets into the regression matrix.

Total least-squares (TLS) is the typical estimator for unstructured EIV problems and is consistent when the perturbations have zero mean, and the covariance is a given positive definite matrix.
When the perturbations are i.i.d. normally distributed, the solution of the TLS is equivalent to that of the maximum likelihood estimator (ML) \cite{Markovsky07overview}. 
For structured EIV problems, the TLS estimator does not give general results since each specific structure requires a particular treatment \cite{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where finding the global optimum is not guaranteed \cite{Rhode14recursive}.
Moreover, the computational complexity of TLS and ML inhibits real-time implementation to solve structured EIV problems.

The least-squares (LS) estimator is a suboptimal but simple solution to structured EIV problems that admits a recursive form for easy real-time implementations.
Some of the reported works that propose LS estimators for structured EIV problems include 
the design of a fast algorithm for matrices with small displacement rank \cite{Mastronardi07fast},
the study of the estimator consistency \cite{Palanthandalam10parameter},
the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models \cite{Kiviet12high} \cite{Kiviet14improved}, and
a discussion of the causes of bias and inconsistency in homogeneous estimators \cite{Yeredor04homogeneous}.

 In measurement applications, it is highly relevant to assess the uncertainty of the input estimate.
The uncertainty of the reported LS estimators for structured EIV problems has not been addressed, and then, remains unknown.
The estimator uncertainty is expressed using the estimation bias and covariance \cite{Pintelon12Book}.
To know the LS estimator uncertainty, we quantified the bias and the covariance of the LS solution of EIV problems in the unstructured and structured cases. 
This extends the perturbation analysis of the LS estimator of unstructured and uncorrelated problems that was investigated in \cite{Stewart90SPT} and in \cite{Vaccaro94}.
This paper presents a study of the statistics of the LS estimate of unstructured and structured EIV problems. 
 We provide a discussion of the unstructured case as a reference, to get an insight of the impact that the structure of the regression matrix and the correlation between the regression matrix and the regressor perturbations have on the uncertainty of the LS estimate.
The structured case is motivated by the metrology estimation problem \cite{Markovsky15cep}, where the EIV problem has a Hankel structure, and the perturbations are correlated.
The study of the metrology input estimate illustrates a methodology to conduct statistical analysis for any structured EIV problem.
The mathematical expectation of the second-order Taylor series expansion of the LS estimate boils down to expressions that quantify the first and second-order moments of the LS estimate.
Via Monte Carlo simulations, we validated the accuracy of the bias, and the covariance approximations.
The derived approximations predict the LS estimate bias, and the covariance, for given sample size and perturbation level.
The predicted variance gives the uncertainty of the LS estimate.
We observed that, for the step input estimation problem, the mean squared error of the LS estimate is near to the minimum variance limit given by the Cram\'er-Rao lower bound of the structured EIV problem.
 By following this methodology, the bias and variance of the solutions of EIV problems with other structures is determined, and therefore, the uncertainty of the estimate.

\subsection{Intro of Chapter experimental validation}

 In this paper we consider that a measurement is a dynamic process, where an input excites a dynamic system, the sensor, and causes a dynamic transient response that also depends on the initial conditions of the sensor.
The to-be-measured quantity is an unknown input that excites the sensor.
The consequent transient response is further processed to estimate quickly the measurand value.
The steady-state response of the sensor, that exists after the stabilization of dynamic effects, gives easy access to the measurand value but this approach is mainly exploited for calibration purposes.

A compensator is an additional dynamic system that acts on the transient response aiming to reduce the sensor transient time.
The compensation is motivated by the need of inverting the sensor dynamic effects to recreate the input.
The convolution of the compensator impulse response with the sensor transient response yields the input estimate.
Therefore, the design of a compensator is based on the sensor model and requires a deconvolution \cite{Eichstadt10}.
Examples of input estimation using compensation of the sensor transient response include a recursive estimation of the compensator parameters \cite{Shu93}, 
finite impulse response (FIR) \cite{Elster07, Niedzwiecki16b} filters and 
infinite impulse response (IIR) filters \cite{Pintelon90, Elster08}.
The filters in these works estimate in real-time the unknown input value.

An alternative to the compensation approach is to use digital signal processing methods that are independent of the sensor model.
A data-driven method that estimates the unknown level of step inputs by processing the sensor step response was introduced in \cite{Markovsky15ieee}. 
This data-driven input estimation method avoids the sensor modeling stage and estimates directly the input.
This method reduces the estimation time compared to a conventional compensator.
The step input estimation method performance was demonstrated by simulations and experiments on a digital signal processor (DSP) of low cost \cite{Markovsky15cep}.
The uncertainty of the step input estimation method has not been assessed before.

To validate the input estimation methods it is necessary to assess the uncertainty associated with their estimates \cite{daSilva12, Ferrero06}.
There are uncertainty propagation studies for model-based compensators such as the FIR and IIR filters for acceleration measurements where the uncertainty is computed in real time \cite{Elster07, Elster08, Link09}.
In these works, the uncertainty expression is based on the transfer function or state space representations of the LTI sensor and filter systems.
Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \cite{Pietrzak14} for mass and in \cite{Ogorevc16} for temperature sensors.
A deconvolution method is implemented to estimate the input waveform in \cite{Hale09} and the uncertainty is obtained from the input estimate covariance. 
The impact that the signal processing data-driven dynamic error correction has on the uncertainty is investigated in \cite{Saggin01}. 
A statistical analysis of the data-driven step input estimation method \cite{Markovsky15cep} was investigated in \cite{Quintana19} and the method uncertainty was obtained with a Monte Carlo simulation study.

This paper provides an uncertainty assessment of the data-driven step input estimation method in a real-life application.
The measurements were conducted in a weighing system based on a load cell sensor.
We observed that even when the whiteness assumptions of the measurement noise are not fulfilled, the step input estimation method still is able to provide a good estimation. 
We found that the mean squared error of the input estimate is near the Cram\'er-Rao lower bound of the EIV problem.
A confidence interval is provided for the input estimate in terms of the number of samples required to satisfy the accuracy specifications of the user.

The novelty of the paper is threefold.
First, using the results of \cite{Quintana19}, that describes a statistical analysis of structured errors in variables (EIV) problems, in this paper we describe the statistical properties of a the data-driven step input estimation method in both simulation and a real-life experiments.
Second, this manuscript also presents the Cram\'er-Rao lower bound for unbiased estimators of the structured and correlated EIV problem that the step input estimation method formulates.
Using this bound we have the minimum mean-squared error (MSE) for this estimation problem that we use compare with the MSE computed from the predictions obtained after the statistical analysis.
The third novelty in this manuscript is the model order selection for the step input estimation method where we use the MSE to select the order that provides the smaller MSE with the lowest computational complexity.

\newpage

### 2020-01-12 ChapterPreliminaries/ChapterPreliminaries.tex
Add files via upload
Initial set of documents

%!TEX root = ..\Thesis.tex
\glsresetall

\chapter{Preliminaries} \label{chap:Preliminaries}

\vfill{}

\section{preliminaries}

In metrology, we use the concepts of linear systems theory to estimate the value of an unknown quantity.
A sensor is considered to be a causal linear time-invariant (LTI) system.
The unknown quantity is the input $\mathbf{u}$ of the sensor, and the consequences of this excitation are a change in the sensor state $\mathbf{x}$, from the initial conditions $\mathbf{x}_{\text{ini}}$, and a transient response in the sensor output $\mathbf{y}$, see Figure \ref{fig:sysLTI}.  
A measurement estimates the input value using the sensor response.

\begin{figure}[htb!]
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]

 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {Sensor};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini}}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=-0.5cm,yshift=0.25cm] {$\mathbf{y}$};
 \draw [-latex,thick] (NB1.west) +(-1.0,0) -- +(0.0cm,0);

 \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {$\mathbf{u}$};

 \end{tikzpicture}
 \caption{Block diagram of an LTI sensor. The input $\mathbf{u}$ excites the sensor and generates the sensor response $\mathbf{y}$. To estimate the input value it is necessary to process the response.} \label{fig:sysLTI}
 \end{figure}

The discrete-time state-space representation of an LTI system is
\begin{equation} \begin{aligned} \mathbf{x}(k+1) &= \mathbf{A} \mathbf{x}(k) + \mathbf{B} \mathbf{u}(k), \quad \text{with} \quad \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\ 
\mathbf{y}(k) &= \mathbf{C} \mathbf{x}(k) + \mathbf{D} \mathbf{u}(k) + \mathbf{\nu}(k),  \label{eqn:dtsslti} \end{aligned} \end{equation}
where $\mathbf{A} \in \mathbb{R}^{n \times n}$, $\mathbf{B} \in \mathbb{R}^{n \times m}$, $\mathbf{C} \in \mathbb{R}^{p \times n}$, and $\mathbf{D} \in \mathbb{R}^{p \times m}$, are the model matrices, $\nu$ is the measurement noise, $n$ is the system order, $m$ is the number of inputs, and $p$ is the number of outputs.
Although we may think that most sensors are single-input single output (SISO) like temperature sensors, there are sensors with single input and multiple outputs such as gas sensors \cite{Munther19}, and sensors with multiple inputs and multiple outputs like the three-axis accelerometers \cite{DEmilia16}, the radio-frequency intruder-detection sensors \cite{Ushiki13}, and the radar sensor \cite{Kueppers17}.

The discrete-time state-space representation suggests that when the model, and the initial conditions are known, and in absence of measurement noise, it is sufficient to solve the system of equations 
\begin{equation} 
\underbrace{ \begin{bmatrix} y(0) \\ y(1) \\ y(2) \\ \vdots \\ y(T) \end{bmatrix} }_{\mathbf{y}}}
 = \underbrace{ \begin{bmatrix} \mathbf{C} \\ \mathbf{C} \mathbf{A} \\ \mathbf{C} \mathbf{A}^2 \\ \vdots \\ \mathbf{C} \mathbf{A}^T \end{bmatrix} }_{\mathbfcal{O}}} \mathbf{x}(0) +
 \underbrace{ \begin{bmatrix} \mathbf{D} \\ \mathbf{C} \mathbf{B} & \mathbf{D} \\ \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \\ \vdots & \ddots \\ \mathbf{C} \mathbf{A}^{T-1} \mathbf{B} & \cdots  &  \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \end{bmatrix} }_{\mathbfcal{T}}}
\underbrace{ \begin{bmatrix} u(0) \\ u(1) \\ u(2) \\ \vdots \\ u(T) \end{bmatrix} }_{\mathbf{u}}} ,
 \label{eqn:knownmodel} \end{equation}
to find the input. 
This system of equations is constructed from the recursions of (\ref{eqn:dtsslti}), where $\mathbfcal{O}$ is the observability matrix of the system, and $\mathbfcal{T}$ is a Toeplitz matrix of the Markov parameters of the system.
# formulated as the input estimation problem from the response of a system.

The input estimation problem is more complex when one or more of these assumptions are not fulfilled.
If the initial conditions and the measurement noise are unknown, the typical approach is to feed the output $\mathbf{y}$ to an additional system.
This additional system is modeled to invert the dynamics of the sensor by doing an operation that is equivalent to deconvolution, aiming to estimate the input $\mathbf{u}$ with the output $\widehat{\mathbf{u}}$ of the additional system.
This system is called compensator because the transient time of the estimated input $\widehat{\mathbf{u}}$ is smaller than the transient time of $\mathbf{y}$, see Figure \ref{fig:compensator}.

\begin{figure}[htb!]
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]

 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {Sensor};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini}}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=-0.5cm,yshift=0.25cm] {$\mathbf{y}$};
 \draw [-latex,thick] (NB1.west) +(-1.0,0) -- +(0.0cm,0);

 \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {$\mathbf{u}$};

\node (NB2) [minimum width=1.5cm, minimum height=0.75cm, xshift=2.9cm, yshift=0.0cm] {Compensator}; 
 \draw [-latex,thick] (NB2.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB2,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini,c}}$};
 \draw [-latex,thick] (NB2.east) ++(0,0) -- +(1.0cm,0);
 \node[draw=none,fill=none] [right=of NB2,xshift=-0.5cm,yshift=0.25cm] {$\widehat{\mathbf{u}}$};

 \end{tikzpicture}
 \caption{Input estimation using a compensator that processes the sensor response $\mathbf{y}$. The compensator is an additional system that reverts the dynamics of the sensor.} \label{fig:compensator}
 \end{figure}

If a model of the sensor is not available, the first idea may be to identify a model using the inputs and outputs, and later to estimate the input using the identified model.
However, the input is unknown and the model must be identified using only the sensor output.
Modeling the input as a multiple of the unit step function allows an analytical treatment of the input estimation problem.
This chapter describes step input estimation methods derived from linear systems theory. 
The description starts with the case in which the model is known, continues with the unknown sensor model case that performs model identification, and ends with a direct estimation of the input without model identification.
# The suitability of data-driven methods on-line implementation is higher because there is no need for high computational power to identify the system.

Step input estimation with system model

With a step input $\mathbf{u} = \bar{\mathbf{u}} s$, where $s(t) = 1$, for $t \geq 0$, and $s(t) = 0$, for $t < 0$, the discrete-time state space representation of the LTI sensor is equivalent to an augmented autonomous system  
\begin{equation} \begin{aligned} \mathbf{x}_\text{a}(k+1) &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_\text{a}(k) , \quad \mathbf{y}(k) = \underbrace{ \begin{bmatrix} \mathbf{C} & \mathbf{D} \end{bmatrix} }_{\mathbf{C}_\text{a}} \mathbf{x}_\text{a}(k), \quad \text{where} \ \ \mathbf{x}_\text{a}(k) = \begin{bmatrix} \mathbf{x}(k) \\ \mathbf{u}(k) \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}(0) . \label{eqn:augmented} \end{aligned} \end{equation}
The eigenvalues $\lambda$ of the augmented autonomous system are found using
\[ \left| \lambda \mathbf{I} - \mathbf{A}_\text{a} \right| = \left| \lambda \mathbf{I} - \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} \right| = \left| \lambda \mathbf{I} - \mathbf{A} \right| \left( \lambda - 1 \right) = 0.\]
Therefore, the eigenvalues, poles, of the augmented autonomous system (\ref{eqn:augmented}) are the eigenvalues, poles, of the LTI system, with the additional eigenvalue $\lambda = 1$, pole at $(1,0)$.

Since the input $\mathbf{u}(t+1) = \mathbf{u}(t)$, for $t \geq 0$, is an augmented state of the autonomous system, and the system is known, a state estimator is sufficient to estimate the input.
In these conditions, the Kalman filter estimates recursively the input value.

Step input estimation without system model

Reduction to autonomous equivalent model

A first method, when the model of the LTI sensor is unknown and we have exact data, identifies a model of the sensor from the step response, and the estimates the input using the identified sensor model.
The model identification consists in the estimation of the matrices $\widehat{\mathbf{A}}_\text{a}$ and $\widehat{\mathbf{C}}_\text{a}$, and the initial conditions $\widehat{\mathbf{x}}(0)$.
To do this, consider that a Hankel matrix $\mathbfcal{H}(\mathbf{y}) \in \mathbb{R}^{n \times n}$, constructed from any linearly independent $n$ autonomous responses from the sensor initial conditions, is full column rank.
Thus, we can express 
\[ \mathbfcal{H}(\mathbf{y}) = \begin{bmatrix} y(1) & y(2) & \cdots & y(n) \\ y(2) & y(3) & \cdots & y(n+1) \\ \iddots & \iddots & \iddots \\ y(n) & y(n+1) & \cdots & y(2n-1) \end{bmatrix} = \begin{bmatrix} \widehat{\mathbf{C}}_\text{a} \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a} \\ \vdots \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a}^{n} \end{bmatrix}  \begin{bmatrix} \mathbf{x}_\text{a}(0) & \mathbf{x}_\text{a}(1) & \cdots & \mathbf{x}_\text{a}(n) \end{bmatrix} = \mathbfcal{O}_\text{a} \mathbf{X}_\text{ini}, \]
where $\mathbf{X}_\text{ini}$ is a matrix with the initial conditions of the $n$ free responses. 
A singular value decomposition of $\mathbfcal{H}(\mathbf{y})$ 
\[ \mathbf{U} \bm{\Sigma} \mathbf{V} = \mathbfcal{H}(\mathbf{y}), \]
permits the estimation of the observability matrix $\mathbfcal{O}_\text{a}$ and the initial conditions $\mathbf{X}_\text{ini}$, for example, by choosing
\[ \widehat{\mathbfcal{O}}_\text{a} = \mathbf{U} \sqrt{\bm{\Sigma}}, \quad \text{and} \quad \widehat{\mathbf{X}}_\text{ini} = \sqrt{\bm{\Sigma}} \mathbf{V} . \]
The matrices $\widehat{\mathbf{A}}_\text{a}$ and $\widehat{\mathbf{C}}_\text{a}$ can be estimated, from $\widehat{\mathbfcal{O}}_\text{a}$, by solving a system of equations.
In order to estimate the input, it is necessary to find a minimal representation of the autonomous system, by doing a linear transformation that removes the pole at $(1,0)$, and recovers the matrices $\widehat{\mathbf{A}}, \widehat{\mathbf{B}}, \widehat{\mathbf{C}}$, and $\widehat{\mathbf{D}}$.

Identification of the model followed by input estimation

A second method is proposed to estimate the step input when the model of the LTI system is unknown but its steady-state gain $\mathbf{G}$ is known.   
Using the steady-state gain $\mathbf{G}$, we can express $\bar{\mathbf{y}} = \mathbf{G} \bar{\mathbf{u}}$, where $\bar{\mathbf{u}}$ is the input exact value and $\bar{\mathbf{y}}$ is the sensor steady state response.
The total response of the system is the sum of the transient and the steady-state responses.
Thus, considering the augmented autonomous model, we can write
\[ \mathbf{y} = \mathbf{G} \ \bar{\mathbf{u}} \ + \ \mathbfcal{O}_\text{a} \ \mathbf{x}(0) , \]
that in matrix form is
\[ \underbrace{ \begin{bmatrix}y(0) \\ y(1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \mathbf{C}_\text{a} \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a} \\ \vdots & \vdots \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a}^T \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{\mathbf{u}} \\ \mathbf{x}(0) \end{bmatrix} .\]
It is necessary to estimate the observability matrix of the augmented system, followed by the estimation of the input  $\mathbf{u}$, and the initial conditions $\mathbf{x}_{\text{ini}}$, using least squares
\[ \begin{bmatrix} \widehat{\mathbf{u}} \\ \widehat{\mathbf{x}}_{\text{ini}} \end{bmatrix} = \left( \mathbf{K}^\top \mathbf{K} \right)^{-1} \mathbf{K}^\top \mathbf{y}. \]

Data-driven estimation of an unknown step input given system response

A third method directly estimates the input value from the step response, without an identification of the sensor model.
Applying the first difference operator, $\Delta = \sigma - 1$, to the system state space representation (\ref{eqn:dtsslti}), we have
\begin{equation} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta \mathbf{y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
where $\sigma$ is the shift operator, defined as $(\sigma^\tau y) (t) = y(t + \tau)$, and
$\Delta \mathbf{u}(k) = \mathbf{0}$, for $k \geq 0$, because we have a step input, and
$\Delta \mathbf{x}(0) = (\mathbf{A} - \mathbf{I}) \mathbf{x}(0) + \mathbf{B} \bar{\mathbf{u}}$.
The resulting system (\ref{eqn:ssalti}) is autonomous.
When the response $\Delta \mathbf{y}$ is persistingly exciting of order $L$, i.e., when the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta \mathbf{y})$ of $L+1$ block rows constructed from $\Delta \mathbf{y}$ satisfies
\[\mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta \mathbf{y} \right) \right) \leq L,  \]
the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta \mathbf{y})$ provides a linear map to the free responses of the augmented autonomous system (\ref{eqn:augmented}).  
In this case, the total response of the system is given by 
\[ \mathbf{y} = \mathbf{G} \ \bar{\mathbf{u}} + \mathbfcal{H}(\Delta \mathbf{y}) \ \bm\ell} \]
that is equivalent to
\[ \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \mathbf{G} & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ \mathbf{G} & \Delta y(T-n) & \Delta y(T-n+1) & \cdots & \Delta y(T-1) \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{\mathbf{u}} \\ \bm{\ell} \end{bmatrix} .\]

The step response observations ${y}(t) \in \mathbb{R}^{}$ for $t = 1,\ldots,T$, where $T$ is the number of samples, is perturbed by normally distributed measurement noise $\epsilon$ with zero mean and given variance $\sigma_{\epsilon}^2$.
 We have \begin{equation} y = \bar{y} + \epsilon, \quad \mathrm{and} \quad K = \bar{K} + E, \label{eqn:y0plusnoise} \end{equation} 
 where $\bar{y}$ is the exact system response and $E$ is the matrix that whose elements are the corresponding noise perturbations to the exact data elements in $\bar{K}$, and the vector ${\ell}$ is a linear transformation of the system initial conditions.  
 The observed response of the system is a step-invariant discretization of the continuous-time response.

 The system of equations ${y} = {K} x$ is an errors-in-variables (EIV) minimization problem with block Hankel structure.
 The perturbation noise present in ${K}$ is correlated to the perturbation of the elements in ${y}$.
 This EIV problem admits a least-squares (LS) solution and the recursive LS enables the online implementation for real-time step input estimation applications.
 The step input estimation method was described in \cite{Markovsky15cep}.

 The uncertainty assessment is compulsory in metrology

 Pending Rik comments :noexport:

** 20190701
*** Introduction - Data-driven methodology:
Moreover, the online uncertainty assessment may not be feasible and we have to rely on confidence bounds.

Consequently, avoiding the explicit model identification from input-output data and estimating directly the input from the transient response reduces the input estimation time and makes data-driven input estimation methods suitable for real-time metrology applications, where confidence bounds describe the estimation uncertainty.
**** Rik: Obained in a calibration step? Please clarify 
*** Original contributions - Statistical analysis of structured EIV problems
It was observed that the input estimation is biased but with small variance, and that the difference between theoretical minimium and the empirical variance is not large.

**** Rik: order of magnitude of RMSE for SNRs larger than XXX  --- (smaller than XXX for SNRs larger than YYY)

*** Original contributions - Experimental validation of the step input estimation method
The results of the estimation method with respect to different sensor model order assumptions were compared and we concluded that increasing the order does not necessarily benefits the input estimation uncertainty.

**** Rik: On the contrary: the important conclusion here is that using an order that is larger than the true model order results in a smaller mean squared error.

***  Estimation of a step input
y in R^p  u in R^m
**** Rik: Relationship between "m" and "p"? For sensors p=m is the generic situation. Do you know sensors where p<>m? If so, discuss these cases.

***  Estimation of a step input -  Step input estimation without system model  -  Data-driven estimation of an unknown step input given system response

The data-driven method is motivated by the persistency of excitation lemma of linear systems theory. Persistency of excitation of the input is a necessary identifiability condition in exact system identification.

**** Rik: persistence of excitations stems from identification theory, not system theory

\section{Step input estimation method - Ramp input}

This section describes the step input estimation method and presents a formulation of the affine input estimation problem.
The step input estimation method does not require a sensor model to estimate the unknown level of the step input by processing the sensor step response.
The affine input estimation problem is formulated as a signal processing estimation problem where the parameters of the affine input can be obtained from the sensor response.

\subsection{Step input estimation method}
The step input estimation method estimates the unknown value $\widebar{u} \in {\rm I\!R}^{}$, that is the level of a step input $u = \widebar{u} s$, where $s$ is the unit step function.
The step input $u$ is applied to a stable linear time-invariant sensor of order $n$ and static gain $\gamma \in {\rm I\!R}^{}$.
The input estimate $\widehat{u}$ is obtained by processing the measured sequence of output observations $\big( {y}(0), \ldots, {y}(T) \big)$, where ${y}(t) \in {\rm I\!R}^{}$ for $t = 1,\ldots,T$, where $T$ is the sample size, and
\begin{equation} \mathbf{y} = \widebar{\mathbf{y}} + \bm{\epsilon} . \label{eqn:y0plusnoise} \end{equation} 
The exact sensor response $\widebar{\mathbf{y}}$ is perturbed by additive measurement noise $\bm{\epsilon}$.
The measurement noise $\epsilon$ is assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.

The measured sensor response is a zero-order hold discretization of the continuous-time response.
The discrete-time sensor response is considered to be piecewise constant.

The step input can be estimated by solving the minimization problem
\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert  \mathbf{y} - \mathbf{K} \mathbf{x} \right\Vert^2_2 . \label{eqn:min_ls} \end{equation}
where $\mathbf{y} = \begin{bmatrix} {y}(n+1) & \ldots & {y}(T) \end{bmatrix}^\top$, 
$\widehat{\mathbf{x}} = \begin{bmatrix} \widehat{u} & \widehat{\bm{\ell}}^\top \end{bmatrix}^\top$, is a vector whose first element is the input estimation $\widehat{u}$ and the $n\text{-vector} \ \widehat{\bm{\ell}}$ is a linear transformation of the sensor initial conditions. The matrix
\begin{equation} \mathbf{K} = \begin{bmatrix} \gamma & \Delta {y}(1) & \Delta {y}(2) & \cdots & \Delta {y}(n) \\ \gamma & \Delta {y}(2) & \Delta {y}(3) & \cdots & \Delta {y}(n+1) \\ \vdots & \iddots & \iddots & \iddots & \\ \gamma & \Delta {y}(T-n) & \Delta {y}(T-n+1) & \cdots & \Delta {y}(T-1) \label{eqn:matrix} \end{bmatrix} \end{equation}
is a Hankel matrix of $(T - n)-\text{block rows}$, constructed from the consecutive differences $\Delta {y}(t) = {y}(t) - {y}(t-1)$ of the measured transient response, augmented in the left with a $(T - n)\text{-vector}$ of elements equal to the sensor static gain $\gamma$.

The minimization problem (\ref{eqn:min_ls}) is a structured errors-in-variables (EIV) problem.
The structure of the EIV problem is due to the presence of the Hankel matrix in $\mathbf{K}$.
The measurement noise $\bm{\epsilon}$ enters in the regression matrix $\mathbf{K}$ and we can express 
\begin{equation} \mathbf{K} = \widebar{\mathbf{K}} + \mathbf{E}, \label{eqn:K0plusnoise} \end{equation} 
where $\widebar{\mathbf{K}}$ is exact data information and $\mathbf{E}$ is the corresponding perturbation noise.
Details of the method formulation are described in \cite{Markovsky15cep}.

The recursive least squares (RLS) algorithm provides a solution to the structured EIV problem and enables real-time implementations of the estimation method.
The RLS has an exponential forgetting factor that selects the observations of the transient response to perform the estimation.
With the exponential  forgetting, the subspace estimation method tracks the evolution of any time-varying applied input.
The affine input is one case in which the input evolves proportionally to time.

\section{Step input estimation method - Statistical analysis}

The step input estimation method estimates the unknown value $u \in {\rm I\!R}^{}$ of the input $\mathbf{u} = u \mathbf{s}$, where $\mathbf{s}$ is the unit step function ($s(t)=0$ if $t<0$ and 1 elsewhere), applied to a bounded-input bounded-output stable linear time-invariant system of order $n$ and given dc-gain $g \in {\rm I\!R}^{}$.
The method processes the sequence of step response observations $\big( \widetilde{y}(0), \ldots, \widetilde{y}(T) \big)^\top$, where $\widetilde{y}(t) \in {\rm I\!R}^{}$ for $t = 0,\ldots,T$, where $T$ is the number of samples, and
\begin{equation} \widetilde{y}(t) = y(t) + \epsilon(t) . \label{eqn:y0plusnoise} \end{equation} 
The exact system response $\mathbf{y}$ is affected by additive i.i.d. normally distributed perturbation $\bm{\epsilon}$ with zero mean and given variance $\sigma_{\epsilon}^2$. 
The observed response of the system is a step-invariant discretization of the continuous-time response.

The step input level estimation is formulated as the minimization problem 
\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert  \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x} \right\Vert^2_2 , \label{eqn:min_ls} \end{equation}
where $\widetilde{\mathbf{y}} = \Big( \widetilde{y}(n+1), \ \ldots, \ \widetilde{y}(T) \Big)^\top$, 
the first element of the to-be-estimated vector
$\mathbf{x} = \Big( u, \ \bm{\ell}^\top \Big)^\top$ is the step input level, 
the vector $\bm{\ell}$ is linked to the system initial conditions,
and the matrix
\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} g & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ g & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ g & \Delta \widetilde{y}(T-n) & \Delta \widetilde{y}(T-n+1) & \cdots & \Delta \widetilde{y}(T-1) \end{bmatrix} , \label{eqn:matrix} \end{equation}
is a Hankel matrix of $(T - n)-\text{block rows}$, constructed from consecutive differences $\Delta \widetilde{y}(t) = \widetilde{y}(t) - \widetilde{y}(t-1)$ of the observed transient response, augmented in the left side with a $(T - n)\text{-vector}$ of elements equal to the known dc-gain $g$, (see \cite{Markovsky15cep}).
The perturbations $\bm{\epsilon}$ enter in matrix $\widetilde{\mathbf{K}}$ and we can express 
\begin{equation} \widetilde{\mathbf{K}} = \mathbf{K} + \mathbf{E}, \label{eqn:K0plusnoise} \end{equation} 
where $K$ is exact data information and $E$ is the additive perturbation noise given as
\begin{equation} \mathbf{E} = \begin{bmatrix} 0 & \Delta \epsilon(1) & \Delta \epsilon(2) & \cdots & \Delta \epsilon(n) \\ 0 & \Delta \epsilon(2) & \Delta \epsilon(3) & \cdots & \Delta \epsilon(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ 0 & \Delta \epsilon(T-n) & \Delta \epsilon(T-n+1) & \cdots & \Delta \epsilon(T-1) \end{bmatrix} . \label{eqn:matrixE} \end{equation}

The underlying system of equations $\widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \mathbf{x}$  in the minimization problem (\ref{eqn:min_ls}) is an errors-in-variables (EIV) problem with Hankel structure.
For metrology applications, the least-squares (LS) approximate solution of this system of equations offers a simple alternative, in its recursive form, to implement the estimation method in real-time.
The LS solution is examined even when it may have some bias because the perturbation errors in $\widetilde{\mathbf{K}}$ are correlated to the perturbations in $\widetilde{\mathbf{y}}$.
The classical LS results for the bias and the covariance cannot be invoked because LS assumes that the additive perturbation only affects the regressor, and that there is no correlation between the regressor and the regression matrix.
The recursive LS method allows for a real-time implementation of the step input estimation method.
Details of the step input estimation method are described in \cite{Markovsky15cep}.

\section{Step input estimation method - Experimental validation}

The step input estimation method is formulated as a signal processing method where the true value of the input is estimated from the sensor response.
The objective of the statistical analysis is to obtain the bias and the covariance of the input estimate.

\subsection{STEP INPUT ESTIMATION METHOD}

The step input estimation method estimates the unknown value $u \in {\rm I\!R}^{}$ of the input $\mathbf{u} = u \mathbf{s}$, where $\mathbf{s}$ is the unit step function ($s(t)=0$ if $t<0$, and $s(t)=1$ elsewhere), applied to a bounded-input bounded-output stable linear time-invariant sensor of order $n$ and given exact dc-gain $g \in {\rm I\!R}^{}$.
The method processes the sequence of step response observations $\widetilde{\mathbf{y}} = \big( \widetilde{y}(0), \ldots, \widetilde{y}(T) \big)^\top$, where 
\begin{equation} \widetilde{y}(t) = y(t) + \epsilon(t) \in {\rm I\!R}^{} \quad \mathrm{for} \quad t = 0,\ldots,T, \label{eqn:y0plusnoise} \end{equation} 
and $T$ is the number of samples.  
The exact sensor response $\mathbf{y}$ is affected by additive Gaussian white measurement noise $\bm{\epsilon}$ with zero mean and given variance $\sigma_{\bm{\epsilon}}^2$.  
The response of the sensor is a step-invariant discretization of the continuous-time response.

The estimation of the step input level is obtained as the solution of the minimization problem \cite{Markovsky15ieee, Markovsky15cep} 
\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert  \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x} \right\Vert^2_2 \label{eqn:min_ls} \end{equation}
where $\widetilde{\mathbf{y}} = \begin{bmatrix} \widetilde{y}(n+1) & \ldots & \widetilde{y}(T) \end{bmatrix}^\top$, 
the first element of the vector
 $\widehat{\mathbf{x}} = \begin{bmatrix} \widehat{u} & \widehat{\mathbf{\ell}}^\top \end{bmatrix}^\top$ is the estimated step input level, the vector $\widehat{\mathbf{\ell}}$ is linked to the sensor initial conditions,
and the matrix
\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} g & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ g & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ g & \Delta \widetilde{y}(T-n) & \Delta \widetilde{y}(T-n+1) & \cdots & \Delta \widetilde{y}(T-1)  \end{bmatrix} \label{eqn:matrixK} \end{equation}
is a Hankel matrix of $(T - n)-\text{block rows}$, constructed from consecutive differences 
\begin{equation*} \Delta \widetilde{y}(t) = \widetilde{y}(t) - \widetilde{y}(t-1)\end{equation*} 
of the measured transient response, augmented in the left side with a $(T - n)\text{-vector}$ of repeated elements equal to the dc-gain $g$.
The measurement noise enters in the matrix $\widetilde{\mathbf{K}}$ 
\begin{equation} \widetilde{\mathbf{K}} = \mathbf{K} + \mathbf{E}, \label{eqn:K0plusnoise} \end{equation} 
where $\mathbf{K}$ contains exact data information and $\mathbf{E}$ is given as
\begin{equation} \mathbf{E} = \begin{bmatrix} 0 & \Delta \epsilon(1) & \Delta \epsilon(2) & \cdots & \Delta \epsilon(n) \\ 0 & \Delta \epsilon(2) & \Delta \epsilon(3) & \cdots & \Delta \epsilon(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ 0 & \Delta \epsilon(T-n) & \Delta \epsilon(T-n+1) & \cdots & \Delta \epsilon(T-1) \end{bmatrix} . \label{eqn:matrixE} \end{equation}

The data-driven step input estimation method builds a system of equations $\widetilde{\mathbf{y}} \approx \widetilde{\mathbf{K}} \mathbf{x}$, that is solved using least squares.
The least squares admit a recursive implementation that avoids the inversion of matrix $\widetilde{\mathbf{K}}$ that increases its size with respect to the sample size $T$.
Instead, the recursive least-squares (RLS) updates the solution of the system of equations considering the previous value of the estimation.
However, to conduct the statistical analysis of the step input estimation method, it is more convenient to use the standard least-squares terminology.
The statistical analysis results obtained from LS treatment are fully compatible with the RLS estimation results.
For the interested reader, the details of the RLS implementation of the step input estimation method are described in [8].

The structured measurement noise in $\widetilde{\mathbf{K}}$ is correlated with the measurement noise in $\widetilde{\mathbf{y}}$.
It is necessary to study the bias and covariance of the LS solution to express the uncertainty of the step input estimate.
The uncertainty assessment of the input estimate is crucial for metrology applications.

The data-driven step input estimation method converts the output-error simultaneous model identification and input estimation problem into an errors-in-variables (EIV) input estimation problem.
The cost of avoiding the parametric sensor modeling is to deal with a more difficult stochastic framework.

\newpage

### 2020-01-12 ChapterRampInput/ChapterRampInput.tex
Add files via upload
Initial set of documents

\glsresetall

\chapter{Ramp input estimation method}\label{chap:RampInput}

\begin{quote}
\emph{The results of this chapter ...}\vfill{}
\end{quote}

\section{Introduction  - Ramp input}

Measurements estimate the unknown value of a physical quantity, namely the measurand.
The to-be-measured physical quantity is applied as an input signal to a sensor. 
The sensor is a dynamical system and its output changes as a consequence of the input excitation and the sensor initial conditions.
The goal of a measurement is to estimate accurately the measurand value using the sensor transient response.
The transient response of a stable sensor decays to a steady state response.
In steady state, the most accurate estimation of the input true value is simply found using the sensor static gain.
However, the steady state is reached in theory after an infinite period of time and in practice we require fast estimations.
The trade-off between accuracy and speed exists in all measurements.

The measurand can be assumed to be constant or variable during the measurement.
A dynamic measurement is present when the fluctuations of the measurand impact on the input estimation.
A typical example of a dynamic measurement problem is a low-bandwidth sensor excited with a fast changing input.
Some characteristics of the input, like the minimum or maximum or the effects of the environment on the measured quantity, occur in small periods of time.  
The detection of the input characteristics is needed in several scientific and industrial applications such as measurements of temperature \cite{Saggin01}, pressure \cite{Matthews14}, acceleration \cite{Link07}, force \cite{Vlajic16, Hessling08a} and mass \cite{Shu93, Boschetti13}.

The solution to dynamic measurement problems is non-trivial.
An approach is to add a dynamical system to compensate the sensor transient response, inverting the effects of the sensor dynamics.
The purpose of such a compensator is to reduce the transient time.
The sensor dynamics are considered in the design of finite and infinite impulse response compensation digital filters based on deconvolution \cite{Eichstadt10} or synthesized to correct dynamic errors \cite{Hessling08a}. 
The model-based deconvolution design of compensators implies that the measurand true value should be known a-priori for certain applications, such as mass determinations \cite{Boschetti13, Niedzwiecki16b}.
In the literature most of the measurements systems are assumed linear time-invariant, but the compensation digital filters can be linear \cite{Tasaki07},  nonlinear \cite{Shu93} or time-varying \cite{Pietrzak14}.

The digital signal processors enable a different approach where the input estimation can be obtained with algorithms that do not necessarily recreate the dynamics of a system.
One of the authors of this paper proposed a data-driven signal processing method that estimates the measurand true value using subspace techniques \cite{Markovsky15cep, Markovsky15ieee}.
The subspace estimation method bypasses the model identification step to estimate the unknown input directly from the response data.
This method was developed to estimate inputs modeled as step functions of unknown scaling level.
We extended the subspace input estimation method to estimate the parameters of inputs that vary at a constant rate.

The inputs that vary at a constant rate are found in applications where the measurand activates the sensor gradually. 
An example of this activation is the measurement of mass while the to-be-weighted object is transported by a conveyor belt, and the profile of the input is a saturated ramp.
Current solutions to the weighing in motion are low pass filters that estimate the mass using the saturated ramp \cite{Tasaki07, Pietrzak14}.
The signal processing affine input estimation methods are motivated by the need to obtain the mass of the object from the ramp before it reaches saturation.
The ramp is parameterized as a straight line model where the slope and the interception are the parameters of interest.

This paper describes a subspace method for the estimation of the affine input parameters.
This method is a recursive algorithm that can be implemented in real-time since it has low computational cost.
The subspace method is independent of the sensor model and, therefore, it is suitable for a variety of applications.
The dynamic weighing is one of the applications and was chosen as an implementation example.
The effectiveness of the method is evaluated in a simulation study.
The performance of the proposed method is compared to that of a maximum-likelihood (ML) estimation method based on local-optimization and a time-varying compensation filter.
%We found that the subspace method obtains the slope estimate from the sensor transient response using the same assumptions of the data-driven step input estimation method.

The ML method resembles the model predictive control approach in the sense that a cost function is minimized iteratively to optimize the parameters of a sensor model using the observed sensor response in a receding time horizon \cite{Mayne14}.
The difference is that the ML method aims to estimate the unknown value of the affine input parameters instead of identifying a model and controlling the dynamic system.
The ML method is more appropriate for off-line processing of the sensor transient response.
Nevertheless, the ML method can estimate the parameters of the affine input, the parameters of a sensor model, and the initial conditions of the sensor.

The uncertainty of the subspace method is assessed using a Taylor expansion of the estimate and Monte Carlo random sampling approach \cite{Quintana19}.
The Monte Carlo approach requires a large set of generated random samples, and for simple systems it is the recommended method.
There exists a deterministic sampling approach to study the uncertainty propagation of complex systems \cite{Hessling13a, Hessling13b}. 
Deterministic sampling aims to represent the minimal statistical information that is relevant to the uncertainty estimation in a finite set.
The uncertainty of the ML method is assessed using the derivatives of the residual error that constructs the to-be-minimized cost function.
The covariance of the optimization method estimate is found using the inverse of the Hessian matrix \cite{Pintelon12Book}.

\subsection{Affine input estimation problem}

The affine input is modeled as a straight line $u = {a} t + {b}$ with parameters the slope $a$ and the interception $b$.
The affine input estimation problem is formulated as a signal processing problem as follows.

\paragraph{Problem} 
Given the sequence of measured output observations $\mathbf{y} = \left( y(1),\ldots,y(T)\right), y(t) \in {\rm I\!R}$, of a stable linear time-invariant system of order $n$, and static gain $\gamma$, generated by an affine input $u = \widebar{a} t + \widebar{b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u} = \widehat{a} t + \widehat{b}$ approximates $u$.
The measured observations $\mathbf{y} = \widebar{\mathbf{y}} + \bm{\epsilon}$ are exact sensor responses $\widebar{\mathbf{y}}$ perturbed by additive noise  $\bm{\epsilon}$ assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.

\paragraph{Motivating example}
Dynamic weighing is an application example where the affine input can be observed.
The weighing of objects in a conveyor belt gives the sensor input an ideal straight line profile when the conveyor belt moves at a constant speed.
The straight line represents the mass coming gradually into the weighing scale sensor in the conveyor belt.
The mass can be estimated from the slope $a$ of the straight line model. 
The mechanical vibrations of the conveyor belt perturb the input and the sensor response is affected by measurement noise.
The interest is to estimate the mass of the object using the sensor response observations.

Consider the weighing scale modeled as a second order mass-spring-damper system, such as the one shown in the diagram of Figure \ref{fig:msd_system}.
The application of an affine input turns the linear time-invariant system into a linear time-varying system, whose
dynamics depends on the input $u(t) = \widebar{a} t + \widebar{b}$, as it is described by the differential equation:
\begin{equation} \dfrac{d}{dt} \left( \left( \widebar{a} t + \widebar{b} + m \right) \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = \left( \widebar{a} t + \widebar{b} + m \right) g \end{equation}
where $m$ is the mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.

The weighing system admits a state space representation where the states $x_1=y$ and $x_2=\dot{y}$ are the position and the speed of the weighing scale: 
\[ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{\widebar{a} t + \widebar{b} + m} & \dfrac{-(k_{\mathrm{d}} + \widebar{a})}{\widebar{a} t + \widebar{b} + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 \\ g \end{bmatrix},  \quad y = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} . \]

\begin{figure}[htb!]
\centering

\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
\tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
\tikzstyle{damper}=[thick,decoration={markings,  
  mark connection node=dmp,
  mark=at position 0.5 with 
  {
    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
  }
}, decorate]
\tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]

\node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
\node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u(t)$};

\node (ground1) at (M.south) [ground,yshift=-1.5cm,xshift=-0.625cm,anchor=north] {};
\draw (ground1.north west) -- (ground1.north east);
\draw [spring] (ground1.north) -- ($(M.south east)!(ground1.north)!(M.south west)$);

\node (groundc) at (M.south) [ground,yshift=-1.5cm,anchor=north] {}; 
\draw (groundc.north west) -- (groundc.north east);

\node (ground2) at (M.south) [ground,yshift=-1.5cm,xshift=0.625cm,anchor=north] {};
\draw (ground2.north west) -- (ground2.north east);
\draw [damper] (ground2.north) -- ($(M.south east)!(ground2.north)!(M.south west)$);

\node[draw=none,fill=none] at (-0.9cm,-1cm) {$k_{\mathrm{s}}$};
\node[draw=none,fill=none] at (0.15cm,-1cm) {$k_{\mathrm{d}}$};
\node[draw=none,fill=none] at (2.0cm,1.0cm) {$y$};
\draw [-latex,thick]  ++(2.2cm,-1cm) -- +(0cm,2.25cm);

\draw [-latex,thick] (M.east) ++(0,0) -- +(1cm,0);
\draw [line width=0.25mm] (2.2cm,-1cm) -- (2.2cm,1cm);
\draw [line width=0.25mm] (2.1cm,-1cm) -- (2.3cm,-1cm);
\draw [line width=0.25mm] (2.1cm,1cm) -- (2.3cm,1cm);
\draw [line width=0.25mm] (2.1cm,-0.5cm) -- (2.3cm,-0.5cm);
\draw [line width=0.25mm] (2.1cm,0.5cm) -- (2.3cm,0.5cm);
\draw [line width=0.25mm] (2.15cm,-0.25cm) -- (2.25cm,-0.25cm);
\draw [line width=0.25mm] (2.15cm,0.25cm) -- (2.25cm,0.25cm);
\draw [line width=0.25mm] (2.15cm,-0.75cm) -- (2.25cm,-0.75cm);
\draw [line width=0.25mm] (2.15cm,0.75cm) -- (2.25cm,0.75cm);
\draw [line width=0.25mm] (2.1cm,0cm) -- (2.3cm,0cm);

\end{tikzpicture}

\caption{\label{fig:msd_system} A second order mass-spring-damper model represents the dynamic weighing system. The dynamics of the system depend on the affine input. The weighing system is time-varying when the applied input changes with respect to time.} 
\end{figure}

In this paper we use the dynamic weighing example to illustrate the implementation of the affine input estimation methods.

\section{Solution methods}

In this section we describe the proposed subspace method to solve the affine input estimation problem.
The method is motivated by the step input estimation method that is formulated as a structured errors-in-variables (EIV) problem and is solved using recursive least squares (RLS).
The exponentially weighted recursive least squares (EWRLS) is a generalization of RLS that allows the extension of the estimation method to reconstruct the affine input.

A maximum-likelihood (ML) method that performs simultaneous system identification and input estimation is described and its results are used as reference to compare the subspace method results.

An example of the subspace method is illustrated with a weighing system.
An existing time-varying (TV) compensation filter that was designed for weighing applications is described briefly.
This TV filter is also used to compare the results of the proposed subspace method.

\subsection{Subspace method}

RLS is a special case of the EWRLS that can solve the minimization problem (\ref{eqn:min_ls}).
The minimization problem is a structured EIV problem and its EWRLS solution is equivalent to 
\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert \mathbf{W}^{1/2} \left( \mathbf{y} - \mathbf{K} \mathbf{x} \right) \right\Vert^2_2 . \label{eqn:min_ewrls} \end{equation}
where $\mathbf{y}$, $\widehat{\mathbf{x}}$, and $\mathbf{K}$ are defined in (\ref{eqn:min_ls}) and (\ref{eqn:matrix}) and
$\mathbf{W} \in \mathbb{R}^{(T-n) \times (T-n)}$ is a diagonal matrix of descending powers of the weight $\lambda \in [0, 1)$, \textit{i.e.}, $W = \mathrm{diag}\left(\lambda^{T-n}, \ldots, \lambda^2, \lambda^1 \right)$.
The weight $\lambda$ is a data selection forgetting factor since it enables to apply different weights to the residuals $\mathbf{y} - \mathbf{K} \mathbf{x}$.
When $\lambda=1$, we have the same solution as RLS.
When $\lambda<1$, the older residuals are weighted with lower values than the residuals of recent observations.
In this way, the solution of the minimization problem depends more on new data.

The profile of an affine input excitation can be reconstructed by solving the structured errors-in-variables problem  (\ref{eqn:min_ewrls}) where $y$ is the corresponding response of a stable dynamic system, $\mathbf{K}$ is constructed from the output observations, and $\lambda<1$.

After the affine input $\widehat{u}$ has been estimated, the affine input parameters $a$ and $b$ are estimated by fitting $\widehat{u}$ to the straight line $\widehat{u} = a t + b$ using linear regression, as follows
\begin{equation} 
  \begin{bmatrix} 1 & 1  \\ \vdots & \vdots \\ T-\tau+1 & 1 \end{bmatrix} 
  \begin{bmatrix} \widehat{a} \\ \widehat{b}  \end{bmatrix} = 
  \begin{bmatrix} \widehat{u}(\tau) \\ \vdots \\ \widehat{u}(T) \end{bmatrix} \label{eqn:LS}
\end{equation}
where $\tau$ is a number of samples used as a tuning parameter that counteracts the time delay that exists when a LTI system is excited with an affine input $u$.
The subspace method can process online the sensor response to the affine excitation.
For each new observation $y(t)$, the estimation $\widehat{u}$ is updated followed by the update of the slope $\widehat{a}$ and the interception $\widehat{b}$.
The values of the tuning parameters $\lambda$ and $\tau$ can be obtained in the calibration of the method using the response of the sensor.

The subspace method estimates the input applied to a dynamic system directly from the caused transient response.
This is a recursive method that can be implemented in real time to estimate the input  using low cost digital signal processors.
The method is a model-free approach and can be used in a variety of physical measurements.
The method  tracks any arbitrary time-varying input and can estimate the parameters of the input when it is associated to a particular input model.

\subsection{Maximum-likelihood method}
Using a model of the sensor, the ML method estimates the sensor initial conditions and the parameters of the applied affine input.
The affine input parameters $a$ and $b$ are the slope and the interception of the straight line model $u = at + b$.

Algorithm \ref{Alg1} lists the steps of the proposed ML method.
This is an iterative minimization method that uses the Jacobian of the residual error function $\mathbf{r} = \widehat{\mathbf{y}} - \mathbf{y}$ to search the direction that minimizes the difference between the measured sensor response $\mathbf{y}$ and the simulated response $\widehat{\mathbf{y}}$.

\begin{algorithm}
\caption{ML Affine input estimation.}\label{Alg1}
\begin{algorithmic}
  \REQUIRE{$\mathbf{y}$, and sensor model parameters}{}
  \STATE {Initialize $\mathbf{\theta} = (a, b, \mathbf{x}_{\text{ini}})$}{}
  \FOR{each $N$ observations of $\mathbf{y}$}{}
  \STATE {Simulate model response $\widehat{\mathbf{y}}(\mathbf{\theta})$}{}
  \STATE {Compute error $\mathbf{r}(\mathbf{\theta}) = \mathbf{y} - \widehat{\mathbf{y}}(\mathbf{\theta})$}{}
  \STATE {Minimize $\mathbf{r}^\top \mathbf{r}$ over $\mathbf{\theta}$}{}
  \STATE {\hspace{0.5cm}  using analytic Jacobian $\partial \mathbf{r} / \partial \mathbf{\theta}$}{}
  \STATE {Update $\mathbf{\theta}$}{}
\ENDFOR
\ENSURE {Optimized parameters $\widehat{a}, \widehat{b}$, and $\widehat{\mathbf{x}}_{\text{ini}}$}{}

\end{algorithmic}
\end{algorithm}

To initialize the optimization variables $a$, $b$, and $\mathbf{x}_{\text{ini}}$,
we use the subspace estimation method using at least the first $2n+2$ samples.
With the initial affine input parameters we simulate a sensor excitation and, since we are using few samples, we have an approximation of the initial conditions.
The optimization variables updates are computed every $N$ new observations.
The minimization can be done using the Levenberg-Marquardt algorithm \cite{Nocedal06}.

\subsubsection{Covariance of the ML method estimates}

The ML method simulates a dynamic system, and computes the Jacobian of the residual error in each iteration.
The analytic formulation of the Jacobian benefits the estimation method in two ways: it speeds up the minimization and gives direct access to the variance of the estimates.
The covariance matrix of the ML estimates can be expressed as \cite{Pintelon12Book}
\begin{equation} \mathbf{Cov} \left( \widehat{x} \right) = \left( \left( \dfrac{\partial \mathbf{r} }{ \partial \mathbf{\theta} } \right)^\top \left( \dfrac{\partial \mathbf{r} }{ \partial \mathbf{\theta} } \right) \right)^{-1}. \label{eqn:covOpt} \end{equation}

\subsubsection{ML affine input estimation example}
We use the previously described dynamic weighing system to illustrate the ML method.
In this case, the formulation of the problem is: 
 \begin{equation} \begin{aligned}
     & \text{Minimize} \quad \text{over} \ a, b, \mathbf{x}_{\text{ini}} \quad \mathbf{r}^T \mathbf{r} \text{, subject to:} \\ & \ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 & \frac{x_{\text{ini,1}}}{a t + b + m} \\ g & \frac{x_{\text{ini,2}}}{a t + b + m}  \end{bmatrix} \begin{bmatrix} 1 \\ \delta(t)  \end{bmatrix}, \\ & \ \widehat{y} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .  
 \end{aligned} \end{equation}
where $a$, $b$ are the affine input parameters, $m$, $k_{\mathrm{d}}$, and $k_{\mathrm{s}}$ are the model parameters, the model states $\mathbf{x}$ are the position and the speed of the weighing scale, the difference between the sensor response $\mathbf{y}$ and the simulated response $\widehat{\mathbf{y}}$ is the residual $\mathbf{r} = \widehat{\mathbf{y}} - \mathbf{y}$.  
The model initial conditions $\mathbf{x}_{\text{ini}}$ are considered optimization variables
and appear in the augmented column of the input matrix.

The analytic Jacobian for the weighing system example is described in the appendix.

\subsection{Time-varying compensation filter}
The time-varying (TV) filter described in \cite{Pietrzak14} was designed to compensate the measured responses of a conveyor weighing system, considering they are modeled as a saturated ramp.
The TV filter consists of three low-pass infinite impulse response (IIR) filters in cascade, where the $i-\mathrm{th}$ IIR filter is given by
\begin{equation} \widehat{y}_i(t) + k_1(t) \widehat{y}_i(t-1) = k_2(t) \left( \widehat{y}_{i-1}(t) + \widehat{y}_{i-1}(t-1) \right) \end{equation}
for $i = 1,\ldots,3$ and $t=0,\ldots,T$.
The sensor response is fed to the filter, then $\widehat{y}_0(t) = y(t)$, and the output of the TV filter $\widehat{u}_\mathrm{ltv}(t) = \widehat{y}_3(t)$ is an estimation of the affine input.
Since in our case we are processing only the ramp without the saturation, the estimates $\widehat{a}_\mathrm{ltv}$ and $\widehat{b}_\mathrm{ltv}$ of the input parameters are obtained by fitting a straight line to the estimated input $\widehat{u}_\mathrm{ltv}$ using linear regression.

The time-varying coefficients $k_1(t)$ and $k_2(t)$ are computed from
\begin{equation} k_1(t) = \frac{f_c(t) - \frac{k_3}{\pi T_s}}{f_c(t) + \frac{k_3}{\pi T_s}}, \quad k_2(t) = \frac{1 + k_1(t)}{2} , \quad k_3 = \sqrt{ \sqrt[3]{2}  - 1} \end{equation}
where $T_s$ is the sampling time and $f_c(t)$ is a heuristic "cutoff" frequency
\begin{equation} f_c(t) = f_u + \left( f_l - f_u \right) \beta^{\left( t-1 \right) /  \alpha \left( T-1 \right) } \end{equation}
that changes between the lower $f_l$ and upper $f_u$ limits, where the coefficient $\beta$ is lower than one, and $\alpha$ is the decay rate. 
The lower frequency value $f_l$ and the coefficient $\beta$ are fixed and the variables $f_u$ and $\alpha$ are optimized off-line by solving the minimization problem 
\begin{equation} \mathrm{minimize} \ \quad \ \mathrm{over} \ f_u, \alpha \ \quad \ \mathrm{max} \left( \dfrac{\mu_{\widetilde{a}_{\mathrm{ltv}}}}{\mu_{\mathrm{spec}}}, \dfrac{\mu_{\widetilde{b}_{\mathrm{ltv}}}}{\mu_{\mathrm{spec}}}, \dfrac{\sigma_{\widetilde{a}_{\mathrm{ltv}}}}{\sigma_{\mathrm{spec}}}, \dfrac{\sigma_{\widetilde{b}_{\mathrm{ltv}}}}{\sigma_{\mathrm{spec}}} \right) + \mathrm{max} \left( \dfrac{\eta_{i}}{T} \right) \label{eqn:tv_optim} \end{equation}
where $\mu_{\widetilde{a}_\mathrm{ltv}}$, $\mu_{\widetilde{b}_\mathrm{ltv}}$, $\sigma_{\widetilde{a}_\mathrm{ltv}}$ and $\sigma_{\widetilde{b}_\mathrm{ltv}}$ are the mean values and the standard deviations of the estimation errors $\widetilde{a}_\mathrm{ltv} = \widehat{a}_\mathrm{ltv} - \widebar{a}$, and $\widetilde{a}_\mathrm{ltv} = \widehat{b}_\mathrm{ltv} - \widebar{b}$, and where $\widebar{a}$ and $\widebar{b}$ are the true values of the input parameters.
The values $\mu_{\mathrm{spec}}$ and $\sigma_{\mathrm{spec}}$ are specified in the OIML recommendation R51 \cite{OIML_R51_1} for mass measurements that use a conveyor belt.

\section{Simulation results}
The results of the affine input parameters estimation are discussed in this section.
We performed a simulation study using the weighing system presented as an example.
We compared the performance of the proposed subspace method to a conventional time-varying (TV) filter, that was conceived for weighing applications, and to the maximum-likelihood (ML) method.

A second order weighing system was excited with an affine input to get the transient response.
The parameters of the weighing system are $m = 15$ g, $d = 5.5$ Ns/m and $k = 10250$ N/m.
The applied affine input $u = 100 t + 10$ represents a mass that changes from $10$ g to $110$ g, at a constant rate, in a time interval of $0.1$ s.
This change of mass represents one example of the weighing input in a conveyor weighing system when an object of 100 g is measured while it is moving at constant speed.
The sensor response is acquired using sampling time $T_s = 0.1$ ms.
In Figure \ref{fig:sensor_weight} the input $u$ is represented with the dotted line, the oscillatory curve is the corresponding sensor ramp response $y$, and $\widehat{u}_{\mathrm{}}$ is a typical input estimate obtained with the subspace method.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_2.pdf} 
\caption{ \label{fig:sensor_weight} The sensor transient response $y$ to an affine input excitation $u = at+b$ is processed by the estimation methods to estimate the parameters $a$ and $b$. In the figure we observe an example of the input estimate $\widehat{u}$ obtained with the subspace method. The input parameters are calculated from $\widehat{u}$ using linear regression.}
\end{figure}

In each simulation, the sensor response was perturbed with an independent realization of additive normally distributed measurement noise.
The added perturbation noise has signal-to-noise ratios (SNR) in the interval [20 dB, 60 dB], values that are realistic in practical applications.
The SNR is defined as the ratio of signal power to the noise power, that is equivalent to the root-mean-square value of the true signal to the variance of the perturbation noise, and in dB is given as

\begin{equation} \mathrm{SNR} = 20 \log_{10}{\dfrac{\dfrac{1}{T}\int\limits_0^T{\widebar{y}(t)^2 d t}}{\sigma_{\bm{\epsilon}}}} \end{equation}

\subsection{Results of the subspace method}

The subspace method processed online the sensor transient response.
The first estimation was obtained with $2n+1$ samples.
The method updated recursively the value of the estimated parameters for each new collected sample, using the forgetting factor $\lambda$ listed in Table \ref{table:lambdas}.
In Figure \ref{fig:rele_dd_40dB_s1} we observe the relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ obtained when SNR = 40 dB. 
The relative errors are smaller than 5 \% after 400 and 500 samples are processed, i.e., 0.04 s and 0.05 s, respectively.
As more samples are collected, the parameter estimation improves.
Figure \ref{fig:rele_SNR_dd_10000} shows the final value of the relative errors, found at $t=0.1$ s, for the different SNR values considered.
The relative errors are smaller than 2 \% regardless of the measurement noise level.

\begin{table}[h!]
\centering
\caption{The values of the forgetting factor $\lambda$ and the samples shift $\tau$ that configure the subspace method for the different values of SNR. These values were obtained after calibration of the method and fixed during the simulation study.}
\begin{tabular}{|c c c c c c|} 
 \hline
 SNR \ [dB] & 20 & 30 & 40 & 50 & 60 \\ [0.5ex] 
 \hline
 $\lambda$ & 0.939 & 0.940 & 0.955 & 0.959 & 0.959 \\ % forgetting factor
 $\tau$ \  & 15 & 15 & 17 & 14 & 20 \\ [0.5ex] % [\mathrm{samples}]
 \hline
\end{tabular}
\label{table:lambdas}
\end{table}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_3.pdf} 
\caption{ \label{fig:rele_dd_40dB_s1} The relative errors of the affine input parameters estimation decrease as the subspace method processes more samples. The relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ are smaller than 5 \% after 400 and 500 samples, respectively ($T_s = 0.1$ ms). }
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_4.pdf} 
\caption{ \label{fig:rele_SNR_dd_10000} The minimum value of the estimation relative errors obtained with the subspace method is less than 2  \% regardless of the SNR between 20 dB and 60 dB. }
\end{figure}

The Cram\'er-Rao lower bound (CRLB) of the errors-in-variables problem formulated by the subspace method was numerically computed for different sample size using Equation (\ref{eqn:FIM}).
The CRLB is the minimum variance that the estimates $\widehat{a}$ and $\widehat{b}$ can have.
The average of $10^4$ runs with independent noise realizations allows to find the empirical mean squared error (MSE) of the estimates, defined as
\begin{equation} \mathrm{MSE}_{\widehat{a}} = \left(b_{\mathrm{p}}\left( \widehat{a} \right) \right)^2 + v_{\mathrm{p}} \left( \widehat{a} \right), \quad \text{and} \quad  \mathrm{MSE}_{\widehat{b}} = ( b_{\mathrm{p}} ( \widehat{b} ) )^2 + v_{\mathrm{p}} ( \widehat{b} ), \end{equation}
where $b_{\mathrm{p}} \left( \widehat{a} \right)$ and $b_{\mathrm{p}} ( \widehat{b} )$ are the bias, and $v_{\mathrm{p}} \left( \widehat{a} \right)$ and $v_{\mathrm{p}} ( \widehat{b} )$ are the variances of the input parameters.
Figure \ref{fig:CRLB_MSE_ab_dd_40dB_MC_10000} shows that the mean squared errors $\mathrm{MSE}_{\hat{a}}$ and $\mathrm{MSE}_{\hat{b}}$ are near to their theoretical minimum $\mathrm{CRLB}_{a}$ and $\mathrm{CRLB}_{b}$ within two orders of magnitude, when SNR = 40 dB.
Figure \ref{fig:CRLB_MSE_SNR_ab_dd_MC_10000} shows the final value of the Cram\'er-Rao lower bounds and the empirical mean-squared errors, found at $t=0.1$ s, for the different SNR values considered.
Both $\mathrm{MSE}_{\hat{a}}$ and $\mathrm{MSE}_{\hat{b}}$ are less than one order of magnitude near to $\mathrm{CRLB}_a$ and $\mathrm{CRLB}_{b}$, respectively, for SNR $\leq 30$ dB.
The difference increases for larger SNR but the maximum is two orders of magnitude for SNR = 60 dB.

Table \ref{table:differentmasses} shows a comparative view of the estimation mean-squared-errors maximum values when the ramp that excites the sensor corresponds to different masses and time durations. 
For each mass and duration, the sensor responses were perturbed with measurement noise of SNR in the interval [20 dB, 60 dB].
The sensor parameters and sampling frequency are fixed and are the same described in the first paragraph of this section.
The maximum values of the MSE are mainly found at low SNR values between 20 and 40 dB.
The higher levels of noise increase the uncertainty of the estimation defined in terms of the MSE.
For fast ramp excitations, the MSE's increase considerably.
The used sampling frequency constrains the estimation method effectiveness for the ramp input duration of 0.05 s or shorter and there it is recommended to use  a higher sampling frequency that will reduce the estimation MSE.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_5.pdf} 
\caption{ \label{fig:CRLB_MSE_ab_dd_40dB_MC_10000} When the SNR of the sensor response is 40 dB, the mean squared errors of the slope estimate $\widehat{a}$ and the interception estimate $\widehat{b}$, obtained by the subspace method, are two orders of magnitude above the theoretical minimum variance given by the Cram\'er-Rao lower bound.}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_6.pdf} 
\caption{ \label{fig:CRLB_MSE_SNR_ab_dd_MC_10000} The Cram\'er-Rao lower bounds of the estimates $\mathrm{CRLB}_a$ and $\mathrm{CRLB}_b$ determine the minimum uncertainty that can be achieved and increases with the measurement noise. 
The empirical mean squared errors $\mathrm{MSE}_{\hat{a}}$ and $\mathrm{MSE}_{\hat{b}}$ are near to the Cram\'er-Rao lower bounds within one order of magnitude for SNR smaller than 30 dB, and within two orders of magnitude for SNR between 40 dB and 60 dB. }
\end{figure}

\begin{table}[h!]
\centering
\caption{ The maximum values of the estimation mean squared errors observed when the subspace method processed the sensor transient responses caused by ramp excitations of masses 0.1, 0.3, 0.5, and 1.0 kg, that last 0.05, 0.1 and 0.5 s, with signal to noise ratios in the interval [20 dB, 60 dB] occur mainly at 40 dB and for lower SNR. There is an increment in the MSE values when the ramp excitation is faster.}

\begin{tabular}{|L|L L L L L|} 
 \hline
 \mathrm{Time \ [s]} & \mathrm{Mass \ [kg]} & 0.1 & 0.3 & 0.5 & 1.0 \\ [0.5ex] 
 \hline
 0.05 & \hspace{7.5mm} \mathrm{MSE}_{\hat{a}}: & 3.55\mathrm{x}10^{0} (@40 \ \mathrm{dB}) & 13.5\mathrm{x}10^{0} (@50 \ \mathrm{dB}) & 3.80\mathrm{x}10^{0} (@20 \ \mathrm{dB}) & 8.29\mathrm{x}10^{0}  (@50 \ \mathrm{dB}) \\ 
 & \hspace{7.5mm} \mathrm{MSE}_{\hat{b}}: & 3.10\mathrm{x}10^{-3} (@40 \ \mathrm{dB})  & 1.11\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) & 1.07\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) & 1.92\mathrm{x}10^{-2} (@60  \ \mathrm{dB})\\
 0.1 & \hspace{7.5mm} \mathrm{MSE}_{\hat{a}}: & 3.08\mathrm{x}10^{-2} (@30 \ \mathrm{dB}) & 1.16\mathrm{x}10^{0} (@40  \ \mathrm{dB}) & 8.34\mathrm{x}10^{-1} (@60 \ \mathrm{dB}) & 1.27\mathrm{x}10^{0} (@50  \ \mathrm{dB}) \\
 & \hspace{7.5mm} \mathrm{MSE}_{\hat{b}}: & 5.99\mathrm{x}10^{-5} (@30 \ \mathrm{dB}) & 1.17\mathrm{x}10^{-1} (@50  \ \mathrm{dB}) & 2.10\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) & 3.76\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) \\
 0.5 & \hspace{7.5mm} \mathrm{MSE}_{\hat{a}}: & 3.03\mathrm{x}10^{-2} (@20 \ \mathrm{dB}) & 2.98\mathrm{x}10^{-1} (@20  \ \mathrm{dB}) & 3.25\mathrm{x}10^{0} (@20 \ \mathrm{dB}) & 8.60\mathrm{x}10^{-2} (@20  \ \mathrm{dB}) \\
 & \hspace{7.5mm} \mathrm{MSE}_{\hat{b}}: & 3.10\mathrm{x}10^{-5} (@50 \ \mathrm{dB}) & 9.86\mathrm{x}10^{-4} (@40 \ \mathrm{dB}) & 2.06\mathrm{x}10^{-5} (@50 \ \mathrm{dB}) & 1.74\mathrm{x}10^{-5} (@20 \ \mathrm{dB}) \\ [0.5ex] % [\mathrm{samples}]
 \hline
\end{tabular}
\
\label{table:differentmasses}
\end{table}

A numerical sensitivity analysis of the subspace method was conducted by adding uncertainty to ramp input generation and looking into the estimation results. 
The uncertainty $\sigma_{\mathrm{s}}$ of the speed in which the ramp increases, and the uncertainties of the input parameters, represented by $\sigma_{a,b}$, were selected to be 0\%, 5\% and 10\% of their true values.
A Monte Carlo simulation with $10^4$ runs was performed for each SNR and the maximum values of the estimation uncertainty are shown in Table \ref{table:dd_sensitivity}. 
According to these results, the input parameters uncertainties $\sigma_{a,b}$ affect more the uncertainty of the estimation than the speed uncertainty $\sigma_{\mathrm{s}}$. 
The parameter that is more affected by the input parameters uncertainty is the interception $\widehat{b}$, since the uncertainty of the slope $\widehat{a}$ is smaller.

\begin{table}[h!]
\centering
\caption{ A sensitivity analysis of the subspace method was conducted by adding uncertainty to the ramp input. The speed $\sigma_{\mathrm{s}}$, and the input parameters $\sigma_{\mathrm{s}}$ uncertainties are 0\%, 5\%, and 10\% of their true values. The table shows the maximum values of the estimates uncertainty. The speed uncertainty causes a smaller spread of the estimates than the input parameters uncertainty.}

\begin{tabular}{|C| C C C C|} 
\hline  
 &  & \sigma_{\mathrm{s}}: \hspace{4mm} 0\% & \hspace{8.1mm} 5\% & \hspace{8.1mm} 10\% \\ [0.5ex] 
 \hline
 \sigma_{{a},{b}}: \hspace{1.6mm} 0\% & \widehat{a} \ [\mathrm{kg/s}] & 0.990 \pm 9.6 \% \ (@50 \ \mathrm{dB}) &  0.999 \pm 8.2 \% \ (@20 \ \mathrm{dB}) & 0.996 \pm 8.2 \% \ (@50 \ \mathrm{dB}) \\ 
  & \widehat{b} \ \mathrm{[g]} & \hspace{-1.75mm} 10.0 \pm 10.3 \% \ (@20 \ \mathrm{dB}) & 9.91 \pm 14.2 \% \ (@30 \ \mathrm{dB}) &  9.99 \pm 21.7 \% \ (@60 \ \mathrm{dB}) \\ 
 \hspace{6.65mm} 5\% & \widehat{a} \ [\mathrm{kg/s}] & 0.994 \pm 9.7 \% \ (@50 \ \mathrm{dB}) & 0.993 \pm 11.3 \% \ (@20 \ \mathrm{dB}) & 0.997 \pm 8.0 \% \ (@50 \ \mathrm{dB}) \\  
  & \widehat{b} \ \mathrm{[g]} & 9.98 \pm 22.4 \% \ (@50 \ \mathrm{dB}) & \hspace{-1.75mm} 10.02 \pm 33.2 \% \ (@30 \ \mathrm{dB}) & 9.94 \pm 16.6 \% \ (@50 \ \mathrm{dB}) \\ 
\hspace{4.88mm} 10\% & \widehat{a} \ [\mathrm{kg/s}] & 1.00 \pm 20.3 \% \ (@40 \ \mathrm{dB}) & 1.00 \pm 22.0 \% \ (@50 \ \mathrm{dB}) & 0.996 \pm 17.7 \% \ (@40 \ \mathrm{dB}) \\    
 & \widehat{b} \ \mathrm{[g]} & 9.97 \pm 46.8 \% \ (@40 \ \mathrm{dB}) & 9.78 \pm 58.8 \% \ (@50 \ \mathrm{dB}) & 9.91 \pm 41.6 \% \ (@40 \ \mathrm{dB}) \\ [0.5ex] 
\hline
\end{tabular}
\
\label{table:dd_sensitivity}
\end{table}

\subsection{Results of the maximum-likelihood method}
The maximum-likelihood (ML) method processed off-line the sensor transient response.
The ML method used the first 50 samples to initialize the optimization variables and updated the variables every $N = 1$ sample.
In Figure \ref{fig:rele_dd_40dB_s1} are shown the relative errors of the estimates $\widehat{a}$, $\widehat{b}$, $\widehat{x}_{\mathrm{ini,1}}$, and $\widehat{x}_{\mathrm{ini,2}}$.
The convergence of the ML estimates gives relative errors below 5 \% after three iterations.
The largest relative error observed is in the scale velocity $\widehat{x}_{\mathrm{ini,2}}$ estimate, which is more sensitive than the other optimization variables.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_7.pdf} 
\caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor initial conditions are estimated with the ML method. After three iterations the relative errors of the estimates are smaller than 5 \%.  }
\end{figure}

Using the analytic Jacobian and Equation (\ref{eqn:covOpt}), we computed the covariance of the estimate.
In Figure \ref{fig:cov_lo_40dB_s1} are shown the variances of the optimization variables taken from the diagonal of the covariance matrix. 
We can see that the variances of $\widehat{a}$ and $\widehat{b}$ decrease faster than the variances of $\widehat{x}_{\mathrm{ini,1}}$ and $\widehat{x}_{\mathrm{ini,2}}$ as more samples are processed.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_8.pdf} 
\caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian. The variances of the affine input parameters estimates decrease faster than the variances of the initial conditions estimates.  }
\end{figure}

The ML method is  computationally more expensive than the subspace method because the ML method simulates the response of a sensor model to optimize the input parameters and the sensor initial conditions.

A typical run of the ML method takes 30 s to complete.
With this execution time, the ML estimation can only be performed offline.
Nevertheless, the ML method objectives are to give the best estimation possible and to serve as a reference to assess the results of the other methods.
An efficient implementation of the ML method to make it feasible for real-time implementation is not trivial and requires additional research that is considered a topic for future research.

A numerical sensitivity analysis of the ML method was conducted by adding uncertainty to the ramp input, and to the parameters of the time-varying model. 
The ramp input was perturbed with uncertainty of the speed in which the ramp increases $\sigma_{\mathrm{s}}$, and with uncertainty on the input parameters $\sigma_{a,b}$. 
The perturbation uncertainty of the model parameters $m$, $d$, and $k$ is represented by $\sigma_{m,d,k}$. 
The perturbation uncertainty was simulated by adding normally distributed random noise with standard deviation equal to 0\%, 5\% and 10\% of the corresponding true values of the perturbed parameters.
A Monte Carlo simulation with $10^3$ runs was conducted for each SNR in the SNR innterval of interest, and in Table \ref{table:ml_sensitivity} are shown the maximum values of the observed estimation uncertainties. 
The results show that the speed uncertainty $\sigma_{\mathrm{s}}$ has a small impact on the estimation uncertainty.
On the contrary, the input parameters uncertainties $\sigma_{a,b}$, and the uncertainties of the model parameters $\sigma_{m,d,k}$ cause a large increment in the uncertainty of the estimation.

\begin{comment}

\begin{table}[h!]
\centering
\caption{ A sensitivity analysis of the ML method was conducted by adding uncertainty to ramp input, and to the model parameters. The perturbation uncertainty was selected with standard deviation of 0\%, 5\%, and 10\% of the parameters true values. The observed maximum estimation uncertainties are shown in the table. The speed uncertainty $\sigma_{\mathrm{s}}$ affects less the input estimation, but the uncertainties of the input parameters $\sigma_{a,b}$, and the model parameters $\sigma_{m,d,k}$ cause an increase of the estimation parameters spread around their mean values.}

\begin{tabular}{|C| C C C C|} 
\hline  
 &  & \sigma_{\mathrm{s}} \hspace{10mm} 0\% & \hspace{15mm} 5\% & \hspace{14mm} 10\% \\ [0.5ex] 
\hline
\sigma_{{a},{b}} \hspace{5.6mm} 0\% & \widehat{a} \ [\mathrm{kg/s}] & 0.999 \pm 0.7 \% \ (@20 \ \mathrm{dB}) & 1.0 \pm 0.7 \% \ (@20 \ \mathrm{dB}) & 0.999 \pm 0.4 \% \ (@30 \ \mathrm{dB}) \\ 
& \widehat{b} \ \mathrm{[g]} & \hspace{-1.75mm} 10.0 \pm 1.1 \% \ (@20 \ \mathrm{dB}) & 9.99 \pm 1.1 \% \ (@20 \ \mathrm{dB}) & 10.0 \pm 0.5 \% \ (@30 \ \mathrm{dB}) \\
& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.6 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) &  0.1 \pm 0.2 \% \ (@30 \ \mathrm{dB}) \\
& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.1 \pm 122 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 130 \% \ (@40 \ \mathrm{dB}) &  0.09 \pm 38 \% \ (@50 \ \mathrm{dB}) \\

\hspace{11.375mm} 5\% & \widehat{a} \ [\mathrm{kg/s}] & 1.0 \pm 9.0 \% \ (@50 \ \mathrm{dB}) & 0.997 \pm 5.2 \% \ (@40 \ \mathrm{dB}) & 1.0 \pm 5.6 \% \ (@40 \ \mathrm{dB}) \\  
& \widehat{b} \ \mathrm{[g]} & 9.88 \pm 25.4 \% \ (@50 \ \mathrm{dB}) & 9.97 \pm 5.2 \% \ (@20 \ \mathrm{dB}) & 10.01 \pm 4.7 \% \ (@40 \ \mathrm{dB}) \\ 
& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 0.57 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 6.24 \% \ (@30 \ \mathrm{dB}) \\
& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.1 \pm 116 \% \ (@40 \ \mathrm{dB}) & 0.09 \pm 128 \% \ (@40 \ \mathrm{dB}) & 0.11 \pm 115 \% \ (@40 \ \mathrm{dB}) \\

\hspace{9.595mm} 10\% & \widehat{a} \ [\mathrm{kg/s}] & 1.00 \pm 10.4 \% \ (@30 \ \mathrm{dB}) & 0.995 \pm 10.3 \% \ (@20 \ \mathrm{dB}) & 0.996 \pm 10.3 \% \ (@20 \ \mathrm{dB}) \\    
& \widehat{b} \ \mathrm{[g]} & 9.94 \pm 10.5 \% \ (@20 \ \mathrm{dB}) & 9.96 \pm 10.4 \% \ (@20 \ \mathrm{dB}) & 10.0 \pm 10.5 \% \ (@20 \ \mathrm{dB}) \\
& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) \\
& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.11 \pm 108 \% \ (@50 \ \mathrm{dB}) & 0.11 \pm 110 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 137 \% \ (@50 \ \mathrm{dB}) \\

\sigma_{{m},{d},{k}}  \hspace{2mm} 0\% & \widehat{a} \ [\mathrm{kg/s}] & 0.999 \pm 0.7 \% \ (@20 \ \mathrm{dB}) & 1.0 \pm 0.69 \% \ (@20 \ \mathrm{dB}) & 1.0 \pm 0.33 \% \ (@30 \ \mathrm{dB}) \\ 
& \widehat{b} \ \mathrm{[g]} & \hspace{-1.75mm} 10.0 \pm 1.1 \% \ (@20 \ \mathrm{dB}) & 10.0 \pm 1.13 \% \ (@20 \ \mathrm{dB}) &  10.0 \pm 0.46 \% \ (@30 \ \mathrm{dB}) \\
& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.6 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 0.19 \% \ (@30 \ \mathrm{dB}) &  0.1 \pm 0.18 \% \ (@30 \ \mathrm{dB}) \\
& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.1 \pm 122 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 130 \% \ (@40 \ \mathrm{dB}) & 0.09 \pm 38 \% \ (@50 \ \mathrm{dB}) \\

\hspace{11.375mm} 5\% & \widehat{a} \ [\mathrm{kg/s}] & 0.994 \pm 15.3 \% \ (@20g \ \mathrm{dB}) & 1.0 \pm 0.7 \% \ (@20 \ \mathrm{dB}) & 1.01 \pm 21.5 \% \ (@20 \ \mathrm{dB}) \\  
& \widehat{b} \ \mathrm{[g]} & 10.3 \pm 79.4 \% \ (@20 \ \mathrm{dB}) & 10.0 \pm 1.2 \% \ (@20 \ \mathrm{dB}) & 9.91 \pm 19.0 \% \ (@20 \ \mathrm{dB}) \\ 
& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.2 \% \ (@30 \ \mathrm{dB}) & 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 0.19 \% \ (@20 \ \mathrm{dB}) \\
& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.10 \pm 113 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 119 \% \ (@40 \ \mathrm{dB}) & 0.11 \pm 323 \% \ (@30 \ \mathrm{dB}) \\

\hspace{9.595mm} 10\% & \widehat{a} \ [\mathrm{kg/s}] & 1.01 \pm 22 \% \ (@30 \ \mathrm{dB}) & 1.01 \pm 16.0 \% \ (@30 \ \mathrm{dB}) & 0.987 \pm 28 \% \ (@50 \ \mathrm{dB}) \\    
& \widehat{b} \ \mathrm{[g]} & 0.1 \pm 26 \% \ (@30 \ \mathrm{dB}) & 9.93 \pm 14.8 \% \ (@30 \ \mathrm{dB}) & 10.4 \pm 93 \% \ (@50 \ \mathrm{dB}) \\ 
& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) &  0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) \\
& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.1 \pm 121 \% \ (@40 \ \mathrm{dB}) & 0.09 \pm 136 \% \ (@40 \ \mathrm{dB}) & 0.11 \pm 111 \% \ (@40 \ \mathrm{dB}) \\
[0.5ex] 
\hline
\end{tabular}
\
\label{table:ml_sensitivity}
\end{table}

\end{comment}

\subsection{Results of the time-varying filter}

We fixed the frequency lower value $f_l=0.01$ Hz and the base $\beta = 0.01$. 
The upper value $f_u$ and the decay rate $\alpha$ were found using optimization (\ref{eqn:tv_optim}). 
We chose the values $\mu_{\mathrm{spec}}=0.5$ and $\sigma_{\mathrm{spec}}=0.24$ as they are specified in the OIML recommendation \cite{OIML_R51_1} for a mass of 100 g measured in a conveyor belt.
The optimized values of the frequency upper value and the decay rate, using a dataset of 100 transient responses, were $f_u = 26.94$ Hz and $\alpha = 5.71$.

Figure \ref{fig:rele_tv_40dB_s1} shows the relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ computed with the TV filter after processing the sensor transient response.
The relative error of the slope estimate is below 5 \% after 300 samples but the relative error of the interception estimate is near 10 \%.
The convergence rate of the estimate $\widehat{a}$ was similar to that of the subspace method.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_9.pdf} 
\caption{ \label{fig:rele_tv_40dB_s1} The relative errors of the time-varying filter estimation converge slower than with the subspace method. 
The relative errors of $\widehat{a}$ and $\widehat{b}$ are smaller than 5 \% only after 800 and 950 samples, respectively.
With the subspace method the relative errors are below 5 \% after 400 and 500 samples. }
\end{figure}

\subsection{Discussion of the observed results}

The subspace method obtains an estimation of the affine input parameters with a recursive least squares solution of a structured errors-in-variables problem.
Updating the parameter estimates without matrix inversion simplifies the method implementation on digital signal processors of low cost.
The price we pay by computing the least squares solution of an errors-in-variables problem is an increase in the bias of the estimates.
Nevertheless, the empirical mean squared errors of the estimates are at most two orders of magnitude larger than the Cram\'er-Rao lower bound, meaning that the estimates uncertainty is low, even when the SNR is lower than 40 dB.

The proposed subspace method is a general method that can be used in different applications, with realistic signal-to-noise ratios. 
It is suitable not only for mass measurements.
The weighing example shows that the subspace method can be used even when the measurement system is linear time-varying.

It was shown that the time-varying (TV) compensation filter can be modified to estimate the mass only from the increasing section of the saturated ramp, without the need of processing the saturation part.
The modified TV filter can be implemented in real-time as the subspace method after a previous off-line coefficients optimization stage with sensor measured data.
Nevertheless, the estimation results of the subspace method are better than the TV filter since they are twice as fast and one order of magnitude more acurate.

The subspace method can estimate the affine input parameters from the sensor response using few parameters, the sensor order $n$, the sensor static gain $\gamma$, and the RLS forgetting factor $\lambda$.
The subspace method does not necessarily requires optimization of $\lambda$ using a dataset of measured sensor responses.
It is required to tune $\lambda$ online during the calibration of the system and later $\lambda$ remains fixed during the measurements.

The results of the sensitivity analysis show how the uncertainty of the subspace method estimates is affected when the ramp input is subject to perturbation.
The impact on the uncertainty on the slope $\widehat{a}$ and the interception $\widehat{b}$ parameters is different.
The ramp speed uncertainty $\sigma_{\mathrm{s}}$ is added to the uncertainty of the parameter $\widehat{b}$, but does not contribute to the uncertainty of the parameter $\widehat{a}$.
On the other hand, the ramp parameters uncertainty ramp speed uncertainty $\sigma_{a,b}$ is added to the uncertainty of the estimates of both parameters $\widehat{a}$ and $\widehat{b}$.
This is not surprising since the estimation parameters are linked to the ramp input parameters.

The maximum-likelihood (ML) method is an approach that requires larger computational resources.
This is an iterative method and in each iteration computes a simulation of a dynamic system followed by the evaluation of the residual error Jacobian matrix.
The advantage of the ML method is that we can estimate simultaneously sensor parameters and the initial conditions of the sensor.
In the weighing case presented as an illustrative example it was not possible to incorporate other parameters of the sensor because they are not identifiable.
According to the estimation relative errors, that are lower than 0.01 after 100 samples, from there on the ML method estimates are near to the true values and we may not require to run the method along all the measurement period.
With only the first 100 samples we have an accurate parameter estimation and variance assessment.

However, the main drawback of the ML method that prevents online implementations is the required computational power to iteratively simulate the response of a sensor model.
It takes an average of 30 s to complete an estimation with the ML method, and this time is too large for fast changing inputs.
The development of an efficient ML method, suitable for real-time implementation, is not straightforward, and is proposed for future research.

The results of the sensitivity analysis of the ML method show that the uncertainty of the ramp input speed $\sigma_{\mathrm{s}}$ does not have an impact on the estimates uncertainty.
Similar to the subspace case, the uncertainty of the ramp input parameters $\sigma_{a,b}$ is additive to the uncertainty of the estimated parameters $\widehat{a}$ and $\widehat{b}$, but does not contribute to the uncertainty of the first element of the initial conditions.
On the other hand, the estimates uncertainty is affected by the perturbation on the model parameters $\sigma_{m,d,k}$.
It is observed that the uncertainty of the estimated parameters $\widehat{a}$ and $\widehat{b}$ increases two and three times the uncertainty of the model parameters.
This implies that we need to have an accurate model of the dynamic system to have a small uncertainty on the estimated input parameters.
Unfortunately, the uncertainty of the second element of the inital conditions is always very high and this is not because of the perturbation of the ramp speed or the model parameters. 
This issue requires more investigation to see if it is due the identifiability of the parameter in the particular example we have.

\section{Conclusions}

An adaptive subspace method was proposed for the estimation of affine input parameters given the measurement of the caused sensor transient response. 
The subspace estimation method is a recursive method that allows online implementation.
This method tracks the input of a system, using exponential forgetting, to process the system response.
The subspace method is model-free and estimates directly the input parameters without identifying a sensor model.
Therefore, it can be applied to the measurement of different physical magnitudes.
In the specific weighing example described in the manuscript, the input is an affine function.
The method is also applicable when the sensor is time-varying.
The subspace method is computationally cheap, simple and suitable for implementation on digital signal processor of low computational power.

A maximum-likelihood estimator based on local optimization was designed to obtain a comparative reference for the other methods.
The maximum-likelihood method estimates the affine input parameters and also model parameters and the sensor's initial conditions.
This method simulates, in a receding horizon scheme, the response of a sensor model to estimate the input and minimizes the sum of the squares of the residual between the measured and the estimated responses.
The main drawback of the maximum-likelihood method is its computational cost and efficient implementation of the method is left for future work.

A linear time-invariant weighing system is used as an test example for the estimation methods.
The weighing system becomes time-varying when an affine input excites the system.
The estimation methods are compared in a simulation study where the time-varying sensor response is perturbed by measurement noise, that is assumed white of zero mean and known variance.
The subspace method results are also compared to those of an existing digital time-varying filter.
The coefficients of the time-varying filter require offline optimization.
The estimation results obtained with the subspace method converges two times faster and is one order of magnitude smaller than those obtained with the time-varying filter.
The empirical mean squared errors of the subspace method estimation is two orders of magnitude larger than the theoretical minimum given by the Cr\'amer-Rao Lower bound.

Future work of this research is the practical implementation of the subspace method for real-time measurements.

\newpage

### 2020-01-12 ChapterStatisticalAnalysis/ChapterStatisticalAnalysis.tex
Add files via upload
Initial set of documents

%!TEX root = ..\Thesis.tex

\glsresetall

\chapter{Statistical Analysis}\label{chap:StatisticalAnalysis}
\begin{quote}
\emph{The results of this chapter ...}\vfill{}
\end{quote}

\section{Introduction  - Statistical analysis}

Errors-in-variables (EIV) are linear estimation problems in which the regression matrix and the regressor are perturbed \cite{VanHuffel91Book}, \cite{Markovsky07overview}.
In structured EIV problems, the regression matrix has a given structure that depends on the problem formulation.
Hankel, Toeplitz, or other application-specific matrices appear in problems of metrology \cite{Markovsky15cep}, system identification \cite{Soderstrom07}, image restoration \cite{Feiz17}, nuclear magnetic resonance spectroscopy \cite{Cai16}, direction-of-arrival estimation \cite{Pan18}, and time-of-arrival estimation \cite{Jia18}.

In metrology, the direct estimation of the input from the sensor transient response is formulated as a structured EIV problem.
The only observed signal is the sensor output.
To estimate a step input, the regression matrix, and the regressor are built from the step response observations \cite{Markovsky15cep}, and
the structure in the regression matrix is block-Hankel.
This data-driven estimation methodology reduces the estimation time of the classical two-stage approach where a sensor model is first identified, and later the input is estimated using the sensor model \cite{Azam15, Niedzwiecki16a}.
%The problem is EIV because the perturbation of the sensor response gets into the regression matrix.

Total least-squares (TLS) is the typical estimator for unstructured EIV problems and is consistent when the perturbations have zero mean, and the covariance is a given positive definite matrix.
When the perturbations are i.i.d. normally distributed, the solution of the TLS is equivalent to that of the maximum likelihood estimator (ML) \cite{Markovsky07overview}. 
For structured EIV problems, the TLS estimator does not give general results since each specific structure requires a particular treatment \cite{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where finding the global optimum is not guaranteed \cite{Rhode14recursive}.
Moreover, the computational complexity of TLS and ML inhibits real-time implementation to solve structured EIV problems.

The least-squares (LS) estimator is a suboptimal but simple solution to structured EIV problems that admits a recursive form for easy real-time implementations.
Some of the reported works that propose LS estimators for structured EIV problems include 
the design of a fast algorithm for matrices with small displacement rank \cite{Mastronardi07fast},
the study of the estimator consistency \cite{Palanthandalam10parameter},
the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models \cite{Kiviet12high} \cite{Kiviet14improved}, and
a discussion of the causes of bias and inconsistency in homogeneous estimators \cite{Yeredor04homogeneous}.

 In measurement applications, it is highly relevant to assess the uncertainty of the input estimate.
The uncertainty of the reported LS estimators for structured EIV problems has not been addressed, and then, remains unknown.
The estimator uncertainty is expressed using the estimation bias and covariance \cite{Pintelon12Book}.
To know the LS estimator uncertainty, we quantified the bias and the covariance of the LS solution of EIV problems in the unstructured and structured cases. 
This extends the perturbation analysis of the LS estimator of unstructured and uncorrelated problems that was investigated in \cite{Stewart90SPT} and in \cite{Vaccaro94}.
This paper presents a study of the statistics of the LS estimate of unstructured and structured EIV problems. 
 We provide a discussion of the unstructured case as a reference, to get an insight of the impact that the structure of the regression matrix and the correlation between the regression matrix and the regressor perturbations have on the uncertainty of the LS estimate.
The structured case is motivated by the metrology estimation problem \cite{Markovsky15cep}, where the EIV problem has a Hankel structure, and the perturbations are correlated.
The study of the metrology input estimate illustrates a methodology to conduct statistical analysis for any structured EIV problem.
The mathematical expectation of the second-order Taylor series expansion of the LS estimate boils down to expressions that quantify the first and second-order moments of the LS estimate.
Via Monte Carlo simulations, we validated the accuracy of the bias, and the covariance approximations.
The derived approximations predict the LS estimate bias, and the covariance, for given sample size and perturbation level.
The predicted variance gives the uncertainty of the LS estimate.
We observed that, for the step input estimation problem, the mean squared error of the LS estimate is near to the minimum variance limit given by the Cram\'er-Rao lower bound of the structured EIV problem.
 By following this methodology, the bias and variance of the solutions of EIV problems with other structures is determined, and therefore, the uncertainty of the estimate.

\section{Statistical analysis}

For an overdetermined system of linear equations, the least-squares (LS) solution is given by
\begin{equation} \widehat{\mathbf{x}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
where $\widetilde{\mathbf{K}}^\dagger$ is the pseudo-inverse of $\widetilde{\mathbf{K}}$.
The objective of the statistical analysis is to obtain the bias and the covariance of the solution $\widehat{\mathbf{x}}$.
The bias and the covariance are computed using the mathematical expectation operator $\mathbb{E}\{\cdot\}$.
First we substitute (\ref{eqn:y0plusnoise}) and (\ref{eqn:K0plusnoise}) in (\ref{eqn:xhat}) obtaining
\begin{equation*} \widehat{\mathbf{x}} = \left( (\mathbf{K}+\mathbf{E})^\top (\mathbf{K}+\mathbf{E})  \right)^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}), \end{equation*} 
which is equivalent to
\begin{equation} \widehat{\mathbf{x}} = \left( \mathbf{I} + \mathbf{M} \right)^{-1} \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}), \label{eqn:1steq} \end{equation} 
where 
\begin{equation} \mathbf{Q} = \mathbf{K}^\top \mathbf{K}, \quad \text{and} \quad \mathbf{M} = \mathbf{Q}^{-1} ( \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{E} ). \end{equation} 
Applying a second-order Taylor expansion of the inverse matrix
\begin{equation} (\mathbf{I} + \mathbf{M})^{-1} \approx \mathbf{I} - \mathbf{M} + \mathbf{M}^2, \label{eqn:TseriesExp} \end{equation} 
that is valid when the SNR is sufficiently high, and therefore $\mathbf{E}$ and $\mathbf{M}$ are small, satisfying the constraint on the spectral radius $\| \mathbf{M} \| < 1$. 
The neglected term in the Taylor series expansion is of the order $O(\Vert M \Vert^3)$.
We can express the estimate as
\begin{equation} \widehat{\mathbf{x}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}). \label{eqn:xhatexp} \end{equation}

Now that the perturbation variables $\bm{\epsilon}$ and $\mathbf{E}$ are no longer inside an inverse matrix, we can compute the mathematical expectation of expressions derived from the Taylor series approximation (\ref{eqn:xhatexp}) of $\widehat{\mathbf{x}}$. 
The bias of the estimate $\widehat{\mathbf{x}}$ is obtained from
\begin{equation} \mathbf{b} \left(\widehat{\mathbf{x}} \right) = \bm{\mu}\left(\widehat{\mathbf{x}} \right) - \mathbf{x}, \label{eqn:biasdef} \end{equation}
where $\bm{\mu} \left(\widehat{\mathbf{x}} \right) = \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}$, and $\mathbf{x} = \mathbf{K}^\dagger \mathbf{y} = \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{y}$ is the true value.
The covariance of the estimate $\widehat{\mathbf{x}}$ is obtained from
\begin{equation} \begin{aligned} \mathbf{C} \left( \widehat{\mathbf{x}} \right) & = \mathbb{E} \left\{ \left( \widehat{\mathbf{x}} - \bm{\mu} \right) \left( \widehat{\mathbf{x}} - \bm{\mu} \right)^\top \right\} . \end{aligned} \label{eqn:covdef} \end{equation}

The terms derived from (\ref{eqn:xhatexp}) that do not contribute to the bias and to the covariance are filtered out by the mathematical expectation operator considering the following general rules that are valid regardless of the existence of structure in the regression problem: 
\begin{itemize}
	\item the expected values $\mathbb{E} \{ \mathbf{E} \} = \mathbf{0}$, and $\mathbb{E} \{ \bm{\epsilon} \} = \mathbf{0}$, since $\mathbf{E}$ and $\bm{\epsilon}$ are zero-mean random variables, and
	\item the expected values of odd order moments, such as $\mathbb{E} \{ \mathbf{E}^\top \mathbf{E} \mathbf{E}^\top \}$, are zero.
\end{itemize}
Moreover, the second-order approximation disregards moments of order four and higher.

After removing the terms with negligible expected value, we have expressions that are approximations of the LS estimation bias and covariance. 
These expressions are different depending on the type of EIV problem considered.
Subsections 3.1 and 3.2 describe the resulting expressions for the unstructured and structured EIV problems, respectively.
The perturbations in the considered unstructured EIV problem are independent.
Comparing the expressions that result from the statistical analysis, we get an insight of what is the impact that the structure and the correlation have on the LS solution of the structured EIV problem.

\subsection{Bias and covariance of the LS estimator for an unstructured EIV problem with uncorrelated noise}
% We consider the unstructured matrix $\mathbf{K} \in \mathbb{R}^{\left(T-n\right) \times \left(n+1\right)}$ with the same dimensions as the structured EIV case.

First, the statistics of the least squares estimator (\ref{eqn:xhat}) of an unstructured EIV problem is discussed, 
provided that the perturbations of the regression matrix and the regressor are i.i.d. normally distributed with zero mean, and variances $\sigma_{\mathbf{E}}^2$ and $\sigma_{\bm{\epsilon}}^2$, respectively.
Therefore, the terms in the Taylor series expansion (\ref{eqn:xhatexp}) that contain products of $\mathbf{E}$ and $\bm{\epsilon}$ have zero expected value.
 After removing the terms without contribution to the bias, and to the covariance, with the mathematical expectation operator,
the analytic approximation of the bias (\ref{eqn:biasdef}) is given by 
\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - T \right) \mathbf{Q}^{-1} \mathbf{x} , \end{aligned} \label{eqn:biasEu} \end{equation}
and the covariance (\ref{eqn:covdef}) is approximated by 
\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \mathbf{x} \mathbf{x}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \ \mathbf{Q}^{-1} \mathbf{x} \mathbf{x}^\top \mathbf{Q}^{-1} , \end{aligned} \label{eqn:varEu} \end{equation}
where the subscript $\mathrm{p}$ stands for prediction of the bias and covariance.
The derivation of equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) is described in Appendix 1.
We use the results described in \cite{Vaccaro94} \S 3 and \cite{Stewart90SPT} \S 2.1 for the expected values of products of unstructured matrices with perturbations.
Equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) depend on the unobservable true values $\mathbf{x}$, $\mathbf{K}$, and on the variance of the perturbations.
The observed variables are $\widetilde{\mathbf{y}}$, $\widetilde{\mathbf{K}}$, and from them we compute $\widehat{\mathbf{x}}$.
It is proposed to directly substitute the observed variables in the analytic expressions.
The substitution gives an approximation of the estimate bias and covariance using the observed data.
We have then
\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - T \right) \widetilde{\mathbf{Q}}^{-1} \widehat{\mathbf{x}} , \end{aligned} \label{eqn:biasEuST} \end{equation}
\begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_\epsilon^2 \ \widetilde{\mathbf{Q}}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \right) \widetilde{\mathbf{Q}}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \widetilde{\mathbf{Q}}^{-1} \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \widetilde{\mathbf{Q}}^{-1} . \end{aligned} \label{eqn:varEuST} \end{equation}
In order to have a prediction of the estimate bias and covariance, the variances $\sigma_{\mathbf{E}}^2$ and $\sigma_{\bm{\epsilon}}^2$ and the observed variables $\widetilde{\mathbf{y}}$, $\widetilde{\mathbf{K}}$, and $\widehat{\mathbf{x}}$ are needed.

\subsection{Bias and covariance of the LS estimator for a structured EIV problem with noise correlation}

This subsection describes the statistics of a structured EIV problem with correlation between the perturbations of the regression matrix and the regressor. 
The structured EIV problem is given by the step input estimation method (\ref{eqn:min_ls}).
The correlation is a consequence of the construction of the block-Hankel matrix in the regression matrix $\widetilde{\mathbf{K}}$ with the first difference of the elements in the regressor $\widetilde{\mathbf{y}}$.

The mathematical expectation operator is applied to the Taylor series expansion of the LS estimate (\ref{eqn:xhatexp}).
After removing the terms with negligible expected value, and considering the structure of matrix $\mathbf{K}$, the estimation bias (\ref{eqn:biasdef}) is approximated by 
\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{Q}^{-1} \left( \left( \mathbf{K}^\top \mathbf{B}_1 - \mathbf{B}_2 \right) x - \left( \mathbf{K}^\top \mathbf{B}_3 - \mathbf{B}_4 \right) \right) , \end{aligned} \label{eqn:biasE} \end{equation}
whereas, the estimation covariance (\ref{eqn:covdef}) is approximated by 
\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{K}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\mathbf{x}} \right) , \end{aligned} \label{eqn:varE} \end{equation}
where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \Big\}$, $\mathbf{P} = \mathbf{K} \mathbf{K}^\dagger$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{P}$. 
The derivation of equations (\ref{eqn:biasE}) and (\ref{eqn:varE}) is described in Appendix 1.
The expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$ and $\mathbf{C}_2$ are obtained using the results of Lemma \ref{lem:lemma1}.

\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newproof{pf}{Proof}

\small
% <how to set font size here to 10 px ? />

\begin{lem}{\small}\label{lem:lemma1}
Let $\mathbf{E} \in {\rm I\!R}^{\left(T-n\right) \times \left(n+1\right)}$ be the partitioned matrix 
\[ \mathbf{E} = \begin{bmatrix} \mathbf{0}_{T-n \times 1} & \bm{\mathcal{H}}(\bm{\epsilon}) \mathbf{D}_{n+1 \times n}^{1,0} \end{bmatrix}, \]
 where $\bm{\mathcal{H}}(\bm{\epsilon}) \in {\rm I\!R}^{\left(T-n\right) \times \left(n+1\right)}$ is the block-Hankel matrix of $T-n$ rows and $n$ columns
\[ \bm{\mathcal{H}}(\bm{\epsilon}) = \begin{bmatrix} \epsilon(0) & \epsilon(1) & \cdots & \epsilon(n) \\ \epsilon(1) & \epsilon(2) & \cdots & \epsilon(n+1) \\ \vdots & \vdots & & \vdots \\ \epsilon(T-n-1) & \epsilon(T-n) & \cdots & \epsilon(T-1) \end{bmatrix} , \]
constructed from samples of the i.i.d. normally distributed random variable $\bm{\epsilon} \sim \mathcal{N}(0, \sigma_{\bm{\epsilon}}^2)$.
Let $\mathbf{D}_{r \times c}^{1,k}$ and $\mathbf{D}_{r \times c}^{2,k}$ be the first and second-order finite differences matricial operators of dimensions $r \times c$ starting from the subdiagonal $k$, for example, 
\[ \mathbf{D}_{4 \times 3}^{1,-1} = \begin{bmatrix} 0 & 0 & 0 \\ -1 & 0 & 0 \\ 1 & -1 & 0 \\ 0 & 1 & -1 \end{bmatrix}, \ \mathbf{D}_{4 \times 3}^{2,0} = \begin{bmatrix}-1 & 0 & 0 \\ 2 & -1 & 0 \\ - 1 & 2 & -1 \\ 0 & -1 & 2 \end{bmatrix} . \]
For a compatible deterministic matrix $\mathbf{A}$, or vector $\mathbf{a}$, the following expected values hold.
\begin{equation*} \begin{aligned} 
& \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \mathbf{E} \right\} = \mathbf{Z}, \
\text{where} \ \mathbf{z}_{i1} = \mathbf{0}, \ \text{and} \ 
z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{2, j-i} \end{bmatrix} \right), \
\text{for} \ i = 1, \cdots, T-n, \ \text{and} \ j = 2, \cdots, n+1. \\
& \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \mathbf{E} \right\} = \mathbf{Z}, \ 
\text{where} \ \mathbf{z}_{1j} = \mathbf{0}, \ \mathbf{z}_{i1} = \mathbf{0}, \ \text{and} \ 
z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \ \mathbf{D}_{T-n \times T-n}^{2, j-i+1} \right) , \
\text{for} \ i = 2, \cdots, n+1, \ \text{and} \ j=2, \cdots, n+1 . \\ 
& \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \mathbf{E}^\top \right\} = \mathbf{Z}, \
\text{where} \ z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} 0 & \mathbf{0}_{n}^\top \\ \mathbf{0}_{n} & \mathbf{D}_{n \times n}^{2, j-i+1} \end{bmatrix} \right), \
\text{for} \ i = 1,\cdots,T-n, \ \text{and} \ j=1,\cdots,T-n. \\ 
& \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \bm{\epsilon} \right\} = \mathbf{z}, \
\text{where} \ z_i = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{1,n+1-i} \end{bmatrix} \right) , \
\text{for} \ i = 1,\cdots,T-n . \\
& \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \bm{\epsilon} \right\} = \mathbf{z}, \
\text{where} \ z_1 = 0, \ \text{and} \ z_i = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \ \mathbf{D}_{T-n \times T-n}^{1, n+2-i} \right), \
\text{for} \ i = 2,\cdots,n+1 . \\
& \mathbb{E} \left\{ \mathbf{E} \mathbf{a} \bm{\epsilon}^\top \right\} = \mathbf{Z}, \
\text{where each column} \ \mathbf{Z}_{j} = - \sigma_{\bm{\epsilon}}^2 \ \mathbf{D}_{T-n \times n+1}^{1,-j} \mathbf{R}_{n+1} \ \mathbf{a} , \
\text{for} \ j = 1,\cdots,T-n , \ \text{with} \ \mathbf{R}_{n+1} = \begin{bmatrix} & \mathbf{R}_n \\ 0 \end{bmatrix} , \\ & \quad
\text{where} \ \mathbf{R}_n \ \text{is a reversal matrix.}
\end{aligned} \end{equation*} 
\end{lem}

\normalsize

The proof of the lemma is given in Appendix 2.

The matrices $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$ and $\mathbf{C}_2$ are considered in the different cases of Lemma \ref{lem:lemma1}.
Each expected value in the lemma is a matrix, or a vector, whose elements are found following the indicated operations.
These operations mainly compute the trace of a product of the corresponding deterministic matrix $\mathbf{A}$, and a matrix constructed from the finite differences matricial operator, which can be of first order $\mathbf{D}^{1}$ or of second order $\mathbf{D}^{2}$.
The total number of operations in the computation of the bias and the covariance is $O\left( T^3 + n^2 \right)$, but this order can be reduced since the $\mathbf{D}$ matrices are sparse.

The formulas for the bias and covariance (\ref{eqn:biasE}) and (\ref{eqn:varE}) depend on the perturbation variance and on the unobservable variables $\mathbf{x}$ and $\mathbf{K}$.
The substitution of the observed variables in the expression gives an approximation of the estimate bias and covariance based on the observed system response.
We have then
\begin{equation} \begin{aligned} \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{Q}}^{-1} \left( \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\mathbf{x}} - \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasST} \end{equation}
and
\begin{equation} \begin{aligned} \mathbf{C}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \widetilde{\mathbf{K}}^{\dagger \top} - \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right) \mathbf{b}_{\widetilde{\mathrm{p}}}^\top \left( \widehat{\mathbf{x}} \right) , \end{aligned} \label{eqn:varST} \end{equation}
where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \bm{\epsilon}^\top \Big\}$, $\widetilde{\mathbf{C}} = \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}}$, $\widetilde{\mathbf{P}} = \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{P}}$.

\subsection{Cram\'er-Rao lower bound of the structured errors-in-variables problem}

The Cram\'er-Rao lower bound (CRLB) provides the lower limit on the variance of the estimate
\begin{equation} \mathrm{CRLB}(\mathbf{x}) = \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\mathbf{x}}) }{\partial (\widehat{\mathbf{x}})} \right)^\top \mathbf{Fi}^{-1}(\mathbf{x}) \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\mathbf{x}}) }{\partial (\widehat{\mathbf{x}})} \right), \label{eqn:CRLB} \end{equation}
where $\mathbf{b} (\widehat{\mathbf{x}})$ is the bias of the estimate and $\mathbf{Fi}(\widehat{\mathbf{x}})$ is the Fisher information matrix \cite{Pintelon12Book}.
The Fisher information matrix is defined as the expected value of the Hessian of the negative likelihood function 
\begin{equation} \mathbf{Fi}(x) = - \mathbb{E} \left\{ \frac{\partial ^2 l (\widehat{\mathbf{x}}) }{\partial \widehat{\mathbf{x}}^2 } \right\}, \end{equation}
where the partial derivatives are evaluated in $\widehat{\mathbf{x}} = \mathbf{x}$.

The minimization problem (\ref{eqn:min_ls}) is equivalent to a structured EIV problem that can be expressed as a linear in the measurements problem \cite{Pintelon12Book}
\begin{equation} e (\widehat{\mathbf{x}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\mathbf{x}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\mathbf{x}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 . \label{eqn:M1vecyK} \end{equation}
where $\widetilde{\mathbf{z}} = \mathbf{z} + \epsilon_{\mathbf{z}}$.
The CRB requires that the true model $\mathbf{M}_1( \mathbf{x} ) \ \mathbf{z} = 0$ exists.
Under the assumption of the measurement perturbation $\epsilon_{\mathbf{z}}$ being normally distributed with covariance matrix $\mathbf{C}_{\mathbf{z}}$, the loglikelihood function of the structured EIV problem is
\begin{equation} \ln{ l(\widetilde{\mathbf{z}}, \widehat{\mathbf{z}}, \widehat{\mathbf{x}}) } = - \frac{1}{2} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right)^\top \mathbf{C}_{\mathbf{z}}^{-1} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right) + \mathrm{constant}, \end{equation}
where $\widehat{\mathbf{z}}$ are parameters of the measurements $\widetilde{\mathbf{z}}$ that have to be estimated and satisfy $\mathbf{M}_1( \widehat{\mathbf{x}} ) \ \widehat{\mathbf{z}} = 0$.
The size of the Fisher information matrix $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
Moreover, in Chapter 19 of \cite{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ can be obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ after doing inversion by parts, giving
\begin{equation} \mathbf{Fi}(\mathbf{x}) = \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)^\top \left( \mathbf{M}_1( \mathbf{x} ) \ \mathbf{C}_{\mathbf{z}} \ \mathbf{M}_1^\top( \mathbf{x} ) \right)^{-1} \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right) 
 , \label{eqn:FIM} \end{equation} 
where the partial derivatives are evaluated at the true values $\mathbf{x}$, and the covariance matrix of the measurements is
\begin{equation} \mathbf{C}_{\mathbf{z}} = \sigma_{\epsilon}^2 \begin{bmatrix} \mathbf{I}_{T-n} & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{1,n} & \mathbf{D}_{T-n \times T-n}^{1,n-1} & \cdots & \mathbf{D}_{T-n \times T-n}^{1,1} \\ \mathbf{0}_{T-n} & \mathbf{0}_{T-n} & \mathbf{0}_{T-n} & \mathbf{0}_{T-n} & \cdots & \mathbf{0}_{T-n} \\ \left( \mathbf{D}_{T-n \times T-n}^{1,n} \right)^\top & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{2,1} & \mathbf{D}_{T-n \times T-n}^{2,0} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,2-n} \\ \left( \mathbf{D}_{T-n \times T-n}^{1,n-1} \right)^\top & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{2,2} & \mathbf{D}_{T-n \times T-n}^{2,1} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,3-n} \\ \vdots & \vdots & \vdots & \vdots & & \vdots \\ \left( \mathbf{D}_{T-n \times T-n}^{1,1} \right)^\top & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{2,n} & \mathbf{D}_{T-n \times T-n}^{2,n-1} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,1} \end{bmatrix} .
 \label{eqn:Cz} \end{equation}

The Cram\'er-Rao lower bound for an biased estimator of the minimization problem (\ref{eqn:min_ls}) is given by
\begin{equation} \mathrm{CRB}_{\mathrm{b}}(\mathbf{x}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right)^\top \ \mathbf{Fi}^{-1}(\mathbf{x}) \ \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right), \label{eqn:CRB_EIV} \end{equation} 
and for an unbiased estimator it is $\mathrm{CRB}_{\mathrm{ub}}(\mathbf{x}) = \mathbf{Fi}^{-1}(\mathbf{x})$.

\section{Simulation results}

Two Monte Carlo (MC) simulation studies were conducted to test the obtained bias and variance formulas.
One MC simulation was devoted to the unstructured EIV problem and the other to the structured EIV problem that corresponds to the step input estimation method (\ref{eqn:min_ls}).
The MC simulations performed $N_{MC} = 10^6$ runs of the LS estimator with different realizations of the perturbation $[\begin{matrix} \mathbf{E} & \bm{\epsilon}] \end{matrix}$.
In the structured EIV problem the perturbation $\mathbf{E}$ is correlated with $\bm{\epsilon}$.
The perturbations variance was selected to have a signal-to-noise ratio (SNR) in the interval [30 dB, 80 dB], according to 
\begin{equation} \mathrm{SNR} = 20 \log_{10}{ \dfrac{ \sqrt{ \dfrac{1}{T} \sum\limits_{t=1}^{T}{ y(t)^2 } } }{ \sigma_{ \bm{\epsilon}}} } . \end{equation} 
For high enough SNR, the constraint $\| \mathbf{M} \| < 1$ is valid and the Taylor series expansion  (\ref{eqn:TseriesExp}) holds. 
In this case, the derived expressions predict the LS estimate bias and variance.
Thus, it is relevant to monitor the evolution of the largest $\mathbf{M}$ eigenvalue to detect the lower limit of the SNR  that allows the validity of the predictions.

The MC simulations provide empirical values of the bias and variance, through the sample mean and sample variance of the LS estimate $\widehat{\mathbf{x}}$.
The results presented in this Section focus the interest in the first element of $\widehat{\mathbf{x}}$, because in the structured EIV problem, the step input estimate is $\widehat{u} = \widehat{\mathbf{x}}_{\left[1\right]}$.

\subsection{Monte Carlo simulation results for an unstructured EIV problem with uncorrelated perturbations}

The MC simulation of the unstructured EIV problem solution was conducted using the following settings.
The exact data elements in the matrix $\mathbf{K} \in {\rm I\!R}^{\left(T-n\right) \times \left(n+1\right)}$  are normally distributed with zero mean and variance one, for $T = 200$ and $n = 2$.
The to-be-estimated exact data is the vector $x = \begin{bmatrix} 1 & 2 & 3 \end{bmatrix}^\top$.
The perturbations $\mathbf{E}$ and $\bm{\epsilon}$ are normally distributed with zero mean and variances subject to $\sigma_{\mathbf{E}}^2 = 2 \sigma_{\bm{\epsilon}}^2$.
These settings are similar to those of the structured case, aiming to have comparable situations.
%In the following results we focus our interest in the first element of the estimate $\widehat{\mathbf{x}}$.

The difference between the sample mean of the estimate $\widehat{\mathbf{x}}_{\left[1\right]}$ and its true value $\mathbf{x}_{\left[1\right]}$ is the empirical bias $b_\mathrm{e}$.
\begin{equation} {b}_\mathrm{e} = \frac{1}{N_{\mathrm{MC}}} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( \widehat{\mathbf{x}}_{\left[1\right]} \right)_i -  \mathbf{x}_{\left[1\right]} } \approx \mu \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) - \mathbf{x}_{\left[1\right]} . \end{equation}
The standard deviation of the estimate $\widehat{\mathbf{x}}_{\left[1\right]}$ is used to obtain the standard error $\sigma_\mathrm{e}$ of the MC simulation, which decreases with respect to the square root of the number of MC runs $N_{\mathrm{MC}}$ \cite{Hammersley75}: 
\begin{equation} \sigma_\mathrm{e} = \frac{ \sigma \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) }{\sqrt{N_{\mathrm{MC}}}}, \quad \mathrm{where} \quad \sigma^2 \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) = \frac{1}{N_{\mathrm{MC}}-1} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( \left( \widehat{\mathbf{x}}_{\left[1\right]} \right)_i - \mu \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) \right)^2 } . \label{eqn:stderr} \end{equation}

In each of the $N_{\mathrm{MC}}$ runs we also obtain the aproximations of the estimation bias $\mathbf{b}_{\widetilde{\mathrm{p}} \left[1\right]}$ and variance $\mathbf{C}_{\widetilde{\mathrm{p}} \left[1,1\right]}$  from observed data, using Equations (\ref{eqn:biasEuST}), and (\ref{eqn:varEuST}).
Similarly as before, we get the sample mean of the aproximations to have bias and variance predictions from observed data 
 \begin{equation} b_{\widetilde{\mathrm{p}}} = \frac{1}{N_{\mathrm{MC}}} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( \mathbf{b}_{\widetilde{\mathrm{p}} \left[1\right]} \right)_i } , \quad \mathrm{and} 
 \quad v_{\widetilde{\mathrm{p}}}  = \frac{1}{N_{\mathrm{MC}}} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( \mathbf{C}_{\widetilde{\mathrm{p}} \left[1,1\right]} \right)_i } . \end{equation} 
The standard error of the bias prediction $b_{\widetilde{\mathrm{p}}}$ is given by
\begin{equation} \sigma_{\widetilde{\mathrm{p}}} = \sqrt{ \frac{1}{N_{\mathrm{MC}} \left( N_{\mathrm{MC}}-1 \right)} \sum_{i=1}^{N_{\mathrm{MC}}} { \left( \left( \mathbf{b}_{\widetilde{\mathrm{p}} \left[1\right]} \right)_i - b_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) \right)^2 } } . \end{equation}

The predicted bias $b_{\mathrm{p}} = \mathbf{b}_{\mathrm{p} \left[1\right]}$ and variance $v_{\mathrm{p}} = \mathbf{C}_{\mathrm{p} \left[1,1\right]}$ from exact data are obtained with one evaluation of the expressions (\ref{eqn:biasEu}) and (\ref{eqn:varEu}). %for unstructured EIV problems, or expressions (\ref{eqn:biasST}) and (\ref{eqn:varST}) for the structured EIV problem.

Figure \ref{bias_sigma_NMC_unstr_str_n2} shows the empirical bias and the bias predictions, with their corresponding standard errors, for the unstructured (plots in the left) and structured (plots in the right) EIV problems.
The simulation settings for the structured and correlated EIV problem are described in the following subsection.
In the figure, it can be observed that the empirical bias $b_\mathrm{e}$ is proportional to the perturbation noise variance while the standard error $\sigma_\mathrm{e}$ is proportional to the perturbation noise standard deviation.
The bias predictions $b_\mathrm{p}$ and $b_{\widetilde{\mathrm{p}}}$ are accurate since they coincide with the empirical bias $b_\mathrm{e}$ in all the SNR interval for unstructured EIV problems, 
In both unstructured and structured EIV cases, the standard errors of the MC estimates $\sigma_{\mathrm{e}}$ and $\sigma_{\widetilde{\mathrm{p}}}$ are smaller than the bias estimates $b_{\mathrm{e}}$ and $b_{\widetilde{\mathrm{p}}}$.
The bias estimates are spread near their sample means. 
The uncertainty is smaller than the bias, therefore, the MC simulation is meaningful.

\begin{figure}[htb!]
    \centering
%    \begin{minipage}{0.45\textwidth}
%        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_1l.pdf} 
 %   \end{minipage}
 %   \begin{minipage}{0.45\textwidth}
  %      \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_1r.pdf} 
   % \end{minipage}
  \caption{ \label{bias_sigma_NMC_unstr_str_n2} We observe the results of the LS solutions for the unstructured (left) and the structured (right) EIV problems. These results are the empirical bias $b_{\mathrm{e}}$, the predicted bias from exact data $b_{\mathrm{p}}$, the predicted bias from observed data $b_{\widetilde{\mathrm{p}}}$, the empirical standard error $\sigma_{\mathrm{e}}$, and the standard error from the estimations using observed data $\sigma_{\widetilde{\mathrm{p}}}$. The estimation biases are proportional to the perturbation variance and the estimation standard errors are proportional to the perturbation standard deviation. Since the standard errors are smaller than the biases, the MC simulation is meaningful. }
\end{figure}

The absolute and relative errors between the predicted and the empirical bias are shown in Figure \ref{fig:b_bt_abse_rele_unstr_e7}. 
The absolute errors decrease with respect to the perturbation variance.
The relative errors are lower than 5\% for SNR between 30 dB and 70 dB. 
There is an increment in the relative errors for SNR above 55 dB. 
 As the SNR increases, the empirical and the predicted bias decrease, as well as the bias error between them.
In order to reveal the bias error, more Monte Carlo runs are needed to reduce the uncertainty of the Monte Carlo simulation that depends on the square root of $N_{\mathrm{MC}}$, see Equation (\ref{eqn:stderr}).
If $N_{\mathrm{MC}}$ is insufficient, the uncertainty of the Monte Carlo simulation hides the bias error and we see this increasing effect of the relative errors.

\begin{figure}[htb!]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_2l.pdf} 
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_2r.pdf} 
    \end{minipage}
  \caption{ \label{fig:b_bt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (left) between the predicted bias and the empirical bias are proportional to the perturbation noise variance as it is expected, and the relative errors (right) are smaller than 5\% for SNR below 70 dB. The bias prediction computed from exact data ${b}_\mathrm{p}$ is very similar to that computed using observed data $b_{\widetilde{\mathrm{p}}}$. } 
\end{figure}

The errors between the predicted and the empirical variance are shown in Figure \ref{fig:v_vt_abse_rele_unstr_e7}.
The absolute and relative errors decrease with respect to the perturbation variance.
The relative errors are lower than 5\% for SNR above 40 dB.

\begin{figure}[htb!]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_3l.pdf} 
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_3r.pdf} 
    \end{minipage}
  \caption{ \label{fig:v_vt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (left) between the predicted variances and the empirical variance are proportional to the perturbation noise variance, and the relative errors (right) are smaller than 5\% for SNR higher that 40 dB. The variance prediction computed using exact data ${v}_\mathrm{p}$ is very similar to that computed from observed data $v_{\widetilde{\mathrm{p}}}$. } 
\end{figure}

Figure \ref{fig:bv_btvt_abse_rele_unstr_e7} shows the absolute and the relative errors between the predictions computed from observed data and those computed from exact data.
The absolute errors between both predictions are proportional to the perturbation noise variance.
The bias and variance predictions, from either exact data or observed data, are equivalent for SNR above 35 dB since the relative errors are lower than 5\%.
The substitution of observed data on the prediction formulas is a valid procedure that allows the prediction of the estimate statistics.

\begin{figure}[htb!]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_4l.pdf} 
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_4r.pdf} 
    \end{minipage}
  \caption{ \label{fig:bv_btvt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (left) between the prediction with observed data and the prediction with exact data are proportional to the perturbation noise variance. The use of observed data instead of exact data in the prediction formulas is valid when the SNR is above 35 dB since the relative errors (right) are smaller than 5\%. } 
\end{figure}

\subsection{Monte Carlo simulation results for a structured EIV problem with correlated perturbations}

The MC simulation of the structured EIV problem (\ref{eqn:min_ls}) solution was conducted processing $T = 200$ samples of the transient response $\widetilde{\mathbf{y}}$ generated by a linear time invariant system of order $n = 2$, after a step input excitation with $u = 1 \ \mathrm{units}$, where the units represent any physical quantity. 
The processed step response is shown in Figure \ref{fig:y}.
The steady state response of the system is reached after 400 samples because from there on the relative error between the instantaneous values of the transient response and the steady-state response value is smaller than 2\%.
Processing 200 samples ensures that the step input estimation is computed from transient data only.
%In the following results we focus our interest in the first element of the estimate $\widehat{\mathbf{x}}_{\left[1\right]}$, where the step input level estimate $\widehat{u}$ is located.

\begin{figure}[htb!]
\centering
\includegraphics[width=0.5\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_5.pdf} 
\caption{ \label{fig:y} The structured EIV problem is constructed with 200 samples of the step response to ensure that only transient data is used. The relative errors between the values of $\mathbf{y}$ and the steady state value are smaller than 2\% after 400 samples, as it is indicated with dashed lines. } 
\end{figure}

The empirical bias is the sample mean of the $N_{\mathrm{MC}}$ estimates minus the true value
 \begin{equation} {b}_\mathrm{e} = \frac{1}{N_{\mathrm{MC}}} \sum_{i=1}^{N_{\mathrm{MC}}}{ \widehat{u}_i - u } \approx \mu \left( \widehat{u} \right) - u. \end{equation}
The standard error associated to this empirical bias estimation \cite{Hammersley75} is defined as 
\begin{equation} \sigma_\mathrm{e} = \frac{\sigma \left( \widehat{u} \right) }{\sqrt{N_{\mathrm{MC}}}}, \quad \mathrm{where} \quad \sigma^2 \left( \widehat{u} \right) = \frac{1}{N_{\mathrm{MC}}-1} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( { \widehat{u}}_i - \mu \left( \widehat{u} \right) \right)^2 } . \end{equation}

The plots on the right side of Figure \ref{bias_sigma_NMC_unstr_str_n2} show the empirical bias, the bias predictions (\ref{eqn:biasE}) and (\ref{eqn:biasST}), and their corresponding standard errors, for the structured EIV problem.
The empirical bias $b_\mathrm{e}$ is proportional to the perturbation noise variance, and
the bias predictions $b_\mathrm{p}$ and $b_{\widetilde{\mathrm{p}}}$ coincide with the empirical bias $b_\mathrm{e}$ only for SNR above 40 dB.
This indicates that the SNR drops to a point where the constraint $\| \mathbf{M} \| < 1$ is no longer satisfied.
At an SNR of 30 dB the perturbation noise affects the bias prediction from observed data and it is three times smaller than the empirical bias.

The absolute and relative errors between the predicted and the empirical bias are shown in Figure \ref{fig:b_bt_abse_rele_str_e7}.
It can be seen that the absolute errors are proportional to the perturbation variance, and the relative errors are lower than 5\% for SNR higher than 40 dB. 
%Nevertheless, the perturbation SNR in practical applications is not high and is near 40 dB

\begin{figure}[htb!]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_6l.pdf} 
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_6r.pdf} 
    \end{minipage}
  \caption{ \label{fig:b_bt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (left) between the predicted bias and the empirical bias $b_{\mathrm{e}}$ are proportional to the perturbation noise variance, and the relative errors (right) are smaller than 5\% only for SNR above 40 dB.}
\end{figure}

The absolute and relative errors between the empirical and the predicted variance, equations (\ref{eqn:varE}) and (\ref{eqn:varST}), are shown in Figure \ref{fig:v_vt_abse_rele_str_e7}.
These absolute errors are proportional to the perturbation noise variance, whereas
the relative errors are lower than 5\% for SNR higher than 45 dB.

The absolute and relative errors between the two predictions from observed data and from exact data, are shown in Figure \ref{fig:bv_btvt_abse_rele_str_e7}.
These absolute errors are also proportional to the perturbation noise variance, and
the relative errors show that the bias and variance predictions from either of the two alternatives are equivalent for SNR higher than 45 dB.

\begin{figure}[htb!]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_7l.pdf} 
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_7r.pdf} 
    \end{minipage}
  \caption{ \label{fig:v_vt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (left) between the predicted and the empirical variance are proportional to the perturbation variance, and the relative errors (right) between the predicted and the empirical variance are smaller than 5\% for SNR above 45 dB. }
\end{figure}

\begin{figure}[htb!]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_8l.pdf} 
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_8r.pdf} 
    \end{minipage}
  \caption{ \label{fig:bv_btvt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (left) between the predictions computed from observed data and those from exact data are proportional to the perturbation noise variance. According to the relative errors (right), the substitution is valid for SNR higher than 45 dB. } 
\end{figure}

The simulation results show that the LS solution of the structured and correlated EIV problem is more sensitive to the perturbation.
This represents a low limit in the SNR interval imposed by the noise level.
In practical applications it is common to have SNRs of 40 dB and the user needs to be aware of the prediction error that the method has.
We measure this prediction error with the mean squared error (MSE), defined as
\begin{equation} \mathrm{MSE} = \sigma^2 + b^2. \end{equation}
 By comparing the different MSEs to the CRLB of the structured EIV problem,
Figure \ref{fig:MSE_CRLB} shows that the MSEs has the same proportionality as the CRLB with respect to the disturbing noise variance.
The MSEs are three times larger than the CRLB.
Since the difference between $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ and the CRLB is lower than one order of magnitude, 
the LS estimation of the structured EIV problem produces results that are comparable to the ML estimation.
The $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ computed from observed data approaches the CRLB for SNR below 40 dB.
This is due to the constraint violation of the Taylor series expansion.

\begin{figure}[!htpb]
\centering
\includegraphics[width=0.5\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_9.pdf}
\caption{\label{fig:MSE_CRLB}
The mean squared errors of the LS estimate are close to the Cram\'er-Rao lower bound. This is a positive indication of the goodness of the LS estimator for the structured EIV problem. The mean squared error of the empirical estimates is represented by $\mathrm{MSE}_{\mathrm{e}}$, and those of the predictions are $\mathrm{MSE}_{\mathrm{p}}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$. The $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ is smaller than CRLB below 40 dB of SNR because of the introduced bias error, but the difference between $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ and the CRLB is lower than one order of magnitude.}
\end{figure}

\section{Conclusions}

 We conducted a statistical analysis of a structured errors-in-variables (EIV) estimation problem with correlation to find the first and second moments of its least-squares solution.
This estimation problem occurs in metrology when we estimate the value of a measurand directly from the sensor transient response.
The data-driven estimation of the physical quantity is formulated as a structured EIV problem with correlation that uses the observed transient response to construct both the regression matrix and the regressor.
The real-time implementation of the method uses a recursive least squares algorithm that is simple and has low computational complexity.
The assessment of the uncertainty is done using the estimate bias and variance.

The conducted statistical analysis produced expressions that predict the estimate bias and variance for given sample size and perturbation level of the observed response.
The Monte Carlo simulation validated the predictions.
We compared the results of solving an unstructured and uncorrelated EIV problem with a structured and correlated EIV problem to understand how the structure and the correlation impacts in the estimation.
We found that the predictions in the structured case are more susceptible to perturbations.
This is due to the two approximations involved, a second-order Taylor series expansion of the estimate, and the substitution of perturbed data on the prediction expressions.
The relative error results indicate that the estimate bias, and variance are predicted using the derived expressions, and the observed data.
The mean squared error of the estimate is close to the results of the maximum likelihood estimate given by the Cram\'er-Rao lower bound.

The bias and variance can be accurately predicted, provided that the Taylor series expansion is valid.
This constraint has to be taken into account to ensure the effectiveness of the method in practical applications.
In the example, it was observed that when the SNR lies outside the validity region, the bias and variance estimation was at most three times larger than the empirical values.

The methodology presented in this paper can be applied to estimate the uncertainty of the solutions to other structured EIV problems.
The bias and variance expressions obtained after the statistical analysis depend on each specific structure.

\section{STATISTICAL ANALYSIS - Ramp input}

\subsubsection{Statistical analysis of the subspace method}

To obtain the first and second moments of the step input estimate $\widehat{u}$, we need to study the solution
\begin{equation} \widehat{\mathbf{x}} = ( \mathbf{K}^\top \mathbf{W} \mathbf{K}  )^{-1} \mathbf{K}^\top \mathbf{W} \mathbf{y} , \label{eqn:xhat} \end{equation}
of the overdetermined structured errors-in-variables EIV problem (\ref{eqn:min_ewrls}).
Using a second order Taylor series expansion of the inverse matrix we can approximate the LS solution as
\begin{equation} \widehat{\mathbf{x}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{C}^{-1} (\widebar{\mathbf{K}} + \mathbf{E})^\top \mathbf{W} (\widebar{\mathbf{y}} + \bm{\epsilon}). \label{eqn:xhatexp} \end{equation} 
where 
\begin{equation} \widebar{\mathbf{C}} = \widebar{\mathbf{K}}^\top \mathbf{W} \widebar{\mathbf{K}}, \quad \text{and} \quad \mathbf{M} = \widebar{\mathbf{C}}^{-1} ( \widebar{\mathbf{K}}^\top \mathbf{W} \mathbf{E} + \mathbf{E}^\top \mathbf{W} \widebar{\mathbf{K}} + \mathbf{E}^\top \mathbf{W} \mathbf{E} ). \end{equation}

The Taylor series approximation of $\widehat{\mathbf{x}}$ enables the calculation of the estimate bias and covariance since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion. 
The bias and the covariance of the estimate $\widehat{\mathbf{x}}$ are obtained from
\begin{equation}  \mathbf{b} \left(\widehat{\mathbf{x}} \right) = \mathbf{\mu} - \widebar{\mathbf{x}}, \label{eqn:biasdef} \end{equation}
\begin{equation} \begin{aligned} \mathrm{\mathbf{Cov}} \left( \widehat{\mathbf{x}} \right) & = \mathbb{E} \left\{ \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)  \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)^\top \right\} . \end{aligned} \label{eqn:covdef} \end{equation} 
where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}$ is the expected value of the estimate and $\widebar{\mathbf{x}}$ is the true value. 
Considering the structure of the EIV problem, the bias and the covariance of the estimate approximation (\ref{eqn:xhatexp}) can be expressed as
\begin{equation} \begin{aligned} \widebar{\mathbf{b}}_{\mathrm{p}} \left( \widehat{x} \right) & \approx \widebar{\mathbf{C}}^{-1} \left(  \left( \widebar{\mathbf{K}}^\top \mathbf{W} \widebar{\mathbf{B}}_1 - \widebar{\mathbf{B}}_2 \right) \mathbf{x} - \left( \widebar{\mathbf{K}}^\top \mathbf{W} \widebar{\mathbf{B}}_3 - \widebar{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasE} \end{equation}
\begin{equation} \begin{aligned} \widebar{\mathrm{\mathbf{Cov}}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \widebar{\mathbf{K}}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \widebar{\mathbf{C}}_1 - \widebar{\mathbf{C}}_2 - \widebar{\mathbf{C}}_2^\top \right) \mathbf{W} \widebar{\mathbf{K}}^{\dagger \top}  - \widebar{\mathbf{b}}_{\mathrm{p}} \left( \widehat{x} \right) \widebar{\mathbf{b}}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varE} \end{equation}
where $\widebar{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widebar{\mathbf{K}}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\widebar{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widebar{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widebar{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widebar{\mathbf{K}}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\widebar{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widebar{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widebar{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \Big\}$, \linebreak $\widebar{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \Big\}$, $\widebar{\mathbf{P}}_\perp = \mathbf{I} - \widebar{\mathbf{K}} \widebar{\mathbf{K}}^\dagger \mathbf{W}$, and $\mathbf{K}^\dagger$ is the pseudo-inverse matrix of $\mathbf{K}$.

The bias and covariance given by expressions (\ref{eqn:biasE}) and (\ref{eqn:varE}) depend on the unobservable true values $\widebar{\mathbf{x}}$ and $\widebar{\mathbf{K}}$.
The measured observations are in the sensor step response $\mathbf{y}$, and from its observations we construct $\mathbf{K}$ and compute $\widehat{\mathbf{x}}$.
The substitution of the measured data in the expressions gives an approximation of the estimate bias and covariance. 
We have then
\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{C}^{-1} \left(  \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_1 - \mathbf{B}_2 \right) \widehat{\mathbf{x}} - \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasST} \end{equation}
  \begin{equation} \begin{aligned} {\mathrm{\mathbf{Cov}}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{K}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{W} \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{x} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varST} \end{equation}
  where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \bm{\epsilon}^\top \Big\}$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger \mathbf{W}$.

  The results of the expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$, were described by the authors of this paper in \cite{Quintana19}.
The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \cite{Vaccaro94, Stewart90SPT}.
It was shown that the bias and variance expressions (\ref{eqn:biasST}) and (\ref{eqn:varST}) are valid predictions of the first and second moments of the LS estimate of a Hankel structured EIV problem.
The problem formulated by the step input estimation method belongs to this type of structured EIV problems and can use the derived expressions to find the bias and variance of the input estimate $\widehat{u}$.
The bias of the estimate $\widehat{u}$ is the fist element of $\mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right)$ and the variance of $\widehat{u}$ is the first element in the main diagonal of $\mathrm{\mathbf{Cov}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right)$.

\subsubsection{Cram\'er-Rao lower bound of the structured EIV problem}

To find the Cram\'er-Rao lower bound (CRLB) of the structured EIV estimation problem (\ref{eqn:min_ls}), we consider that this structured and correlated estimation problem can be expressed as a linear in the measurements problem \cite{Pintelon12Book}
\begin{equation} e (\widehat{\mathbf{x}}, \mathbf{z}) = \mathbf{M}_1( \widehat{\mathbf{x}} ) \ \mathbf{z} = \underbrace{\begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\mathbf{x}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix}}_{\mathbf{M}_1( \widehat{\mathbf{x}} )} \underbrace{\begin{bmatrix} \mathbf{y} \\ \mathrm{vec} ( \mathbf{K} ) \end{bmatrix}}_{\mathbf{z}} = 0 . \label{eqn:M1vecyK} \end{equation}
where $\mathbf{z} = \widebar{\mathbf{z}} + \bm{\epsilon}_{\mathbf{z}}$.
To have the CRLB, it is necessary the existence of the true model $\mathbf{M}_1( \mathbf{x} ) \ \mathbf{z} = 0$.
The measurement perturbation $\bm{\epsilon}_{\mathbf{z}}$ is assumed to be normally distributed with covariance matrix $\mathbf{C}_{\mathbf{z}}$, and then, the loglikelihood function of the structured and correlated EIV problem is
\begin{equation} \ln{ l(\mathbf{z}, \widehat{\mathbf{z}}, \widehat{\mathbf{x}}) } = - \frac{1}{2} \left( \mathbf{z} - \widehat{\mathbf{z}} \right)^\top \mathbf{C}_{\mathbf{z}}^{-1} \left( \mathbf{z} - \widehat{\mathbf{z}} \right) + \mathrm{constant}, \end{equation}
where the elements of $\widehat{\mathbf{z}}$ are the estimated parameters of the measurements $\mathbf{z}$,  that satisfy $\mathbf{M}_1(\widehat{\mathbf{x}}) \ \widehat{\mathbf{z}} = 0$.
The size of the Fisher information matrix $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
However, the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ is obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$, by doing inversion by parts \cite{Pintelon12Book} \S 19, and can be expressed as  
\begin{equation} \mathbf{Fi}(\mathbf{x}) = \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)^\top \left( \mathbf{M}_1( \mathbf{x} ) \mathbf{C}_{\mathbf{z}}  \mathbf{M}_1^\top( \mathbf{x} ) \right)^{-1}  \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right) .
 \label{eqn:FIM}   \end{equation} 
The partial derivatives are evaluated at the true values $\mathbf{x}$. 
Thus, the covariance matrix of the measurements is

\begin{equation} \mathbf{C}_{\mathbf{z}} = \sigma_{\epsilon}^2 \begin{bmatrix} \mathbf{I}_{T-n} & \mathbf{0}_{T-n} & \mathbf{D}_1 \\ \mathbf{0}_{T-n} & \mathbf{0}_{T-n} & \mathbf{0}_{T-n \times n \left( T-n \right)}  \\  \mathbf{D}_1^\top & \mathbf{0}_{n \left( T-n \right) \times T-n} & \mathbf{D}_2  \end{bmatrix}  \label{eqn:Cz} \end{equation} 
where
\begin{equation*} \begin{aligned} \mathbf{D}_1 &= \begin{bmatrix}\mathbf{D}_{T-n \times T-n}^{1,n} & \mathbf{D}_{T-n \times T-n}^{1,n-1} & \cdots  & \mathbf{D}_{T-n \times T-n}^{1,1}\end{bmatrix}, \\ 
 \mathbf{D}_2 &= \begin{bmatrix} \mathbf{D}_{T-n \times T-n}^{2,1} & \mathbf{D}_{T-n \times T-n}^{2,0} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,2-n} \\ \mathbf{D}_{T-n \times T-n}^{2,2} & \mathbf{D}_{T-n \times T-n}^{2,1} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,3-n} \\ \vdots & \vdots & & \vdots \\ \mathbf{D}_{T-n \times T-n}^{2,n} & \mathbf{D}_{T-n \times T-n}^{2,n-1} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,1} \end{bmatrix} , \end{aligned}  \end{equation*} 
and the matrices
$\mathbf{D}_{r \times c}^{1,k}$ and $\mathbf{D}_{r \times c}^{2,k}$  are the first and second order finite differences matricial operators of dimensions $r \times c$  starting from the subdiagonal $k$, for example \begin{equation*} 
\mathbf{D}_{2 \times 3}^{1,1} = \begin{bmatrix} 1 & -1 & 0 \\ 0 & 1 & -1 \end{bmatrix}, \ \mathrm{and} \quad \mathbf{D}_{3 \times 3}^{2,0} = \begin{bmatrix}-1 & 0 & 0 \\ 2 & -1 & 0 \\ - 1 & 2 & -1  \end{bmatrix} . \end{equation*}

The Cram\'er-Rao lower bound for a biased estimator of the minimization problem (\ref{eqn:min_ls}) is 
\begin{equation} \mathrm{CRLB}_{\mathrm{b}}(\mathbf{x}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right)^\top \mathbf{Fi}^{-1}(\mathbf{x}) \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right), \label{eqn:CRB_EIV} \end{equation} 
whereas for an unbiased estimator, it is $\mathrm{CRLB}_{\mathrm{ub}}(\mathbf{x}) = \mathbf{Fi}^{-1}(\mathbf{x})$.

\section{STATISTICAL ANALYSIS - Experimental validation}

To obtain the first and second moments of the step input estimate $\widehat{u}$, we need to study the least-squares (LS) solution 
\begin{equation} \widehat{\mathbf{x}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}}  )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
of the overdetermined structured errors-in-variables EIV problem (\ref{eqn:min_ls}), 
where $\widetilde{\mathbf{K}}^\dagger$ is the pseudo-inverse matrix of $\widetilde{\mathbf{K}}$.
Using a second order Taylor series expansion of the inverse matrix we can approximate the LS solution as
\begin{equation} \widehat{\mathbf{x}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}). \label{eqn:xhatexp} \end{equation} 
where 
\begin{equation} \mathbf{Q} = \mathbf{K}^\top \mathbf{K}, \quad \text{and} \quad \mathbf{M} = \mathbf{Q}^{-1} ( \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{E} ). \end{equation}

The Taylor series approximation of $\widehat{\mathbf{x}}$ enables the calculation of the bias and covariance of $\widehat{\mathbf{x}}$ since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion. 
The bias and the covariance of the estimate $\widehat{\mathbf{x}}$ are obtained from
\begin{equation}  \mathbf{b} \left(\widehat{\mathbf{x}} \right) = \mathbf{\mu} - \mathbf{x}, \label{eqn:biasdef} \end{equation}
\begin{equation} \begin{aligned} \mathrm{\mathbf{C}} \left( \widehat{\mathbf{\mathbf{x}}} \right) & = \mathbb{E} \left\{ \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)  \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)^\top \right\} . \end{aligned} \label{eqn:covdef} \end{equation} 
where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}$, and $\mathbf{x} = \mathbf{K}^\dagger \mathbf{y}$ is the true value. 
Considering the structure of the EIV problem, the bias and the covariance of the approximation (\ref{eqn:xhatexp}) can be expressed as
\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{Q}^{-1} \left(  \left( \mathbf{K}^\top \mathbf{B}_1 - \mathbf{B}_2 \right) \mathbf{x} - \left( \mathbf{K}^\top \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasE} \end{equation}
\begin{equation} \begin{aligned} \mathrm{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{K}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{K}^{\dagger \top}, \end{aligned} \label{eqn:varE} \end{equation}
where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \Big\}$, \linebreak $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \Big\}$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger$.

The expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$, were studied in \cite{Quintana19} and their results are described in the following Lemma:

\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}

\normalsize % \small
%   <how to set font size here to 10 px ? />

\begin{lem}{\small}\label{lem:lemma1}
Let $\mathbf{E}$ be the matrix defined in (\ref{eqn:matrixE}),
constructed from samples of the i.i.d. normally distributed random variable $\bm{\epsilon} \sim \mathcal{N}(0, \sigma_\epsilon^2)$.
For a compatible deterministic matrix $\mathbf{H}$, or vector $\mathbf{h}$, we have
\begin{equation*} \begin{aligned} 
& \mathbb{E} \left\{ \mathbf{E} \mathbf{H} \mathbf{E} \right\} = \sigma_{\bm{\epsilon}}^2 \mathbf{A}, \
\text{where} \ a_{ij} = \mathrm{tr} \left( \mathbf{H} \begin{bmatrix} 0_{T-n} & \mathbf{D}_{T-n \times n}^{2, j-i} \end{bmatrix} \right), \\
& \ \ \text{for}  \ i = 1, \cdots, T-n, \ \text{and} \ j = 2, \cdots, n+1, \ \text{and} \ a_{i1} = 0. \\
& \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{H} \mathbf{E} \right\} = \sigma_{\bm{\epsilon}}^2 \mathbf{A}, \ 
\text{where} \ a_{ij} = \mathrm{tr} \left( \mathbf{H} \ \mathbf{D}_{T-n \times T-n}^{2, j-i+1} \right)  , \\
& \ \ \text{for} \ i = 2, \cdots, n+1, \ \text{and} \ j=2, \cdots, n+1 , \ \text{and} \   a_{1j} = a_{i1} = 0,  \\ 
& \mathbb{E} \left\{ \mathbf{E} \mathbf{H} \mathbf{E}^\top \right\} = \sigma_{\bm{\epsilon}}^2 \mathbf{A}, \
\text{where} \ a_{ij} = \mathrm{tr} \left( \mathbf{H} \begin{bmatrix} 0 & 0_{n}^\top \\ 0_{n} & \mathbf{D}_{n \times n}^{2, j-i+1} \end{bmatrix} \right), \\
& \ \ \text{for} \ i = 1,\cdots,T-n, \ \text{and} \ j=1,\cdots,T-n. \\ 
& \mathbb{E} \left\{ \mathbf{E} \mathbf{H} \bm{\epsilon} \right\} = \sigma_{\bm{\epsilon}}^2 \mathbf{a} , \
\text{where} \ a_i = \mathrm{tr} \left( \mathbf{H} \begin{bmatrix} 0_{T-n} & \mathbf{D}_{T-n \times n}^{1,n+1-i} \end{bmatrix} \right) , \\
& \ \ \text{for} \ i = 1,\cdots,T-n . \\
& \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{H} \bm{\epsilon} \right\} = \sigma_{\bm{\epsilon}}^2 \mathbf{a}, \
\text{where} \ a_i = \ \mathrm{tr} \left( \mathbf{H} \ \mathbf{D}_{T-n \times T-n}^{1, n+2-i} \right), \\ & \ \ \text{for} \ i = 2,\cdots,n+1  , \ \text{and} \ a_1 = 0,   \\
& \mathbb{E} \left\{ \mathbf{E} \mathbf{h} \bm{\epsilon}^\top \right\} = \sigma_{\bm{\epsilon}}^2 \mathbf{Z}, \
\text{where} \ \mathbf{Z}_{j} = -\mathbf{D}_{T-n \times n+1}^{1,-j} \begin{bmatrix} & \mathbf{R}_n \\ 0 \end{bmatrix} \mathbf{h} ,  \\ 
& \ \ \text{for} \ j = 1,\cdots,T-n ,\\
\end{aligned} \end{equation*} 
where $\mathbf{R}_n$ is a reversal matrix, and the matrices
$\mathbf{D}_{r \times c}^{1,k}$ and $\mathbf{D}_{r \times c}^{2,k}$  are the first and second order finite differences matricial operators of dimensions $r \times c$  starting from the subdiagonal $k$, for example \begin{equation*} 
\mathbf{D}_{2 \times 3}^{1,1} = \begin{bmatrix} 1 & -1 & 0 \\ 0 & 1 & -1 \end{bmatrix}, \ \text{and} \quad \mathbf{D}_{3 \times 3}^{2,0} = \begin{bmatrix}-1 & 0 & 0 \\ 2 & -1 & 0 \\ - 1 & 2 & -1  \end{bmatrix} .
\end{equation*} 
\end{lem}

\normalsize

A proof of the lemma is given in the Appendix.

The bias and covariance given by expressions (\ref{eqn:biasE}) and (\ref{eqn:varE}) depend on the unobservable true values $\mathbf{x}$, $\mathbf{K}$.
The measured variable is the sensor step response $\widetilde{\mathbf{y}}$, and from its observations we construct $\widetilde{\mathbf{K}}$ and compute $\widehat{\mathbf{x}}$.
The substitution of the measured data in the expressions gives an approximation of the bias and covariance estimation.
We have then
\begin{equation} \begin{aligned} \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{Q}}^{-1} \left(  \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\mathbf{x}} - \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasST} \end{equation}
\begin{equation} \begin{aligned} \mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \widetilde{\mathbf{K}}^{\dagger \top}, \end{aligned} \label{eqn:varST} \end{equation}
where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{\mathbf{C}}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \bm{\epsilon}^\top \Big\}$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger$.

The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \cite{Vaccaro94, Stewart90SPT}.
It was shown that the bias and variance expressions (\ref{eqn:biasST}) and (\ref{eqn:varST}) are good predictions of the first and second moments of the LS estimate of a Hankel structured EIV problem.
The problem formulated by the step input estimation method belongs to this type of structured EIV problems and the derived expressions can be used to find the bias and variance of the input estimate $\widehat{u}$.
The bias and variance predictions approximate the empirical bias and variance.

The assessment of the uncertainty of the estimate $\widehat{\mathbf{x}}$ is obtained from the covariance matrix $\mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right)$. In this way, the uncertainty of the step input estimate $\widehat{u}$ is described by the variance in the first element on the diagonal of $\mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right)$.

In the rest of this section we describe the Cram\'er-Rao lower bound (CRB) of the structured EIV estimation problem (\ref{eqn:min_ls}), that can be expressed as a linear in the measurements problem \cite{Pintelon12Book}
\begin{equation} e (\widehat{\mathbf{x}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\mathbf{x}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\mathbf{x}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 . \label{eqn:M1vecyK} \end{equation}
where $\widetilde{\mathbf{z}} = \mathbf{z} + \epsilon_{\mathbf{z}}$.
The CRB requires that the true model $\mathbf{M}_1( \mathbf{x} ) \ \mathbf{z} = 0$ exists.
Under the assumption of the measurement perturbation $\epsilon_{\mathbf{z}}$ being normally distributed with covariance matrix $\mathbf{C}_{\mathbf{z}}$, the loglikelihood function of the structured EIV problem is
\begin{equation} \ln{ l(\widetilde{\mathbf{z}}, \widehat{\mathbf{z}}, \widehat{\mathbf{x}}) } = - \frac{1}{2} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right)^\top \mathbf{C}_{\mathbf{z}}^{-1} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right) + \mathrm{constant}, \end{equation}
where $\widehat{\mathbf{z}}$ are parameters of the measurements $\widetilde{\mathbf{z}}$ that have to be estimated and satisfy $\mathbf{M}_1( \widehat{\mathbf{x}} ) \ \widehat{\mathbf{z}} = 0$.
The size of the Fisher information matrix $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
Moreover, in Chapter 19 of \cite{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ can be obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ after doing inversion by parts, giving
\begin{equation} \mathbf{Fi}(\mathbf{x}) = \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)^\top \left( \mathbf{M}_1( \mathbf{x} ) \mathbf{C}_{\mathbf{z}}  \mathbf{M}_1^\top( \mathbf{x} ) \right)^{-1}  \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right) 
 \label{eqn:FIM}   \end{equation} 
where the partial derivatives are evaluated at the true values $\mathbf{x}$, and the covariance matrix of the measurements is

\begin{equation} \mathbf{C}_{\mathbf{z}} = \sigma_{\epsilon}^2 \begin{bmatrix} \mathbf{I}_{T-n} & \mathbf{0}_{T-n} & \mathbf{D}_1 \\ \mathbf{0}_{T-n} & \mathbf{0}_{T-n} & \mathbf{0}_{T-n \times n \left( T-n \right)}  \\  \mathbf{D}_1^\top & \mathbf{0}_{n \left( T-n \right) \times T-n} & \mathbf{D}_2  \end{bmatrix}  \label{eqn:Cz} \end{equation} 
where
\begin{equation*} \begin{aligned} \mathbf{D}_1 &= \begin{bmatrix}\mathbf{D}_{T-n \times T-n}^{1,n} & \mathbf{D}_{T-n \times T-n}^{1,n-1} & \cdots  & \mathbf{D}_{T-n \times T-n}^{1,1}\end{bmatrix}, \ \mathrm{
and} \\ 
 \mathbf{D}_2 &= \begin{bmatrix} \mathbf{D}_{T-n \times T-n}^{2,1} & \mathbf{D}_{T-n \times T-n}^{2,0} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,2-n} \\ \mathbf{D}_{T-n \times T-n}^{2,2} & \mathbf{D}_{T-n \times T-n}^{2,1} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,3-n} \\ \vdots & \vdots & & \vdots \\ \mathbf{D}_{T-n \times T-n}^{2,n} & \mathbf{D}_{T-n \times T-n}^{2,n-1} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,1} \end{bmatrix} . \end{aligned}  \end{equation*}

The Cram\'er-Rao lower bound for an biased estimator of the minimization problem (\ref{eqn:min_ls}) is given by
\begin{equation} \mathrm{CRB}_{\mathrm{b}}(\mathbf{x}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right)^\top \mathbf{Fi}^{-1}(\mathbf{x}) \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right), \label{eqn:CRB_EIV} \end{equation} 
and for an unbiased estimator it is $\mathrm{CRB}_{\mathrm{ub}}(\mathbf{x}) = \mathbf{Fi}^{-1}(\mathbf{x})$

\newpage

### 2020-01-12 Extras/ListofPublications.tex
Add files via upload
Initial set of documents

%!TEX root = ..\Thesis.tex

\addchap{List of Publications}

\subsubsection*{Journal publication2}

\noindent G. Quintana-Carapia, I. Markovsky, R. Pintelon, P. Z. Csurcsia and D. Verbeke, "Experimental validation of a data-driven step input estimation method for dynamic measurements," \textit{IEEE Transactions on Instrumentation and Measurement}, Vol. x, No. x, pp. xx-xx, 2019, doi: 10.1109/TIM.2019.2951865

\vspace{0.2cm}

\noindent G. Quintana-Carapia, I. Markovsky, R. Pintelon, P. Z. Csurcsia and D. Verbeke, "Bias and covariance of the least squares estimate in a structured errors-in-variables problem," \textit{Computational Statistics & Data Analysis}, Vol. 144, No. 106893, 2020, ISSN 0167-9473, doi: 10.1016/j.csda.2019.106893.

\vspace{0.2cm}
\noindent G. Quintana-Carapia, I. Markovsky, "Input parameters estimation from time-varying measurements," \textit{Measurement}, Vol. 153, No. 1, pp. 107418, 2020, ISSN 0263-2241, doi: 10.1016/j.measurement.2019.107418.

\vspace{0.2cm}
\subsubsection*{Conference publications}

\noindent G. Quintana-Carapia, I. Markovsky, "Data driven dynamic measurements",
In: \textit{9th International Workshop on the Analysis of Dynamic Measurements}", Berlin, Germany, 2016.

\vspace{0.2cm}

\noindent G. Quintana-Carapia, I. Markovsky, "Data driven dynamic measurements",
In: \textit{35th Benelux Meeting on Systems and Control}", Soesterberg, The Netherlands, 2016.

### 2020-01-12 Gus-thesis.tex
Add files via upload
Initial set of documents

%%
%% This code was generated from the thesis template of Adam Cooman--Evi Van Nechel.
%% It has been adapted slightly by Gustavo Quintana.
%% Compile with: pdflatex --shell-escape -synctex=1 -interaction=nonstopmode Gus-thesis.tex 
%%
\documentclass[english,british,BCOR=5mm,DIV=12,fontsize=10pt]{scrbook}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[paperwidth=17cm,paperheight=24cm]{geometry}
\geometry{verbose,tmargin=2.5cm,bmargin=3cm,lmargin=2.5cm,rmargin=2cm}
\setcounter{secnumdepth}{1}
\setcounter{tocdepth}{1}
\setlength{\parskip}{\medskipamount}
\setlength{\parindent}{0pt}
\synctex=-1
\usepackage{color}
\definecolor{note_fontcolor}{rgb}{0.800781, 0.800781, 0.800781}
\usepackage{babel, array, verbatim, float, booktabs, units, amsmath, amssymb, graphicx, tikz}
\usepackage{mathdots, amsthm, textcomp, times, mathptmx, mathtools, bm, amsbsy}
\usepackage[unicode=true,pdfusetitle, bookmarks=true, bookmarksnumbered=false, bookmarksopen=false, breaklinks=false,pdfborder={0 0 0}, pdfborderstyle={}, backref=false, colorlinks=false] {hyperref}
\usepackage[outdir=./]{epstopdf}
\usetikzlibrary{shapes,arrows,calc,patterns,decorations.pathmorphing,decorations.markings,positioning}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% The greyedout annotation environment
\newenvironment{lyxgreyedout}
  {\textcolor{note_fontcolor}\bgroup\ignorespaces}
  {\ignorespacesafterend\egroup}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\input{config/preamble.tex}

% to stop the compiler from complaining about the \it command in the references, I override it with the thing KOMA proposes.
\renewcommand{\it}{\normalfont\itshape}

% lyxgreyedout defined in latex
\renewenvironment{lyxgreyedout}
{\textcolor{red}\bgroup}{\egroup}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Start Document
\makeatother

\begin{document}
{\footnotesize{}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Front cover
%\includepdf[pages={1}]{Cover}
\newpage{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% VUB title page

\titlepage{\fontfamily{pag}\selectfont % I use the ugly VUB font only for this title page
	\hspace{-0.1cm}\includegraphics[width=0.4\columnwidth]{Extras/VUB_logo.pdf}
	% \vspace{-1.7cm}
	\hfill{}%
	\begin{minipage}[t]{0.4\columnwidth}% 
	\vspace{-1.4cm}
		\begin{flushright}
			{\footnotesize{}Department of }\\
			{\footnotesize{}Fundamental Electricity}\\
			{\footnotesize{}and Instrumentation (ELEC)}
		\par\end{flushright}%
	\end{minipage}

	\vspace{1.5cm}
	% I use purple and orange here for the color. In the preamble I have overwritten these colors to be VUB orange and VUB blue respectively
	\textcolor{purple}{\footnotesize{}Thesis submitted in fulfillment of the requirements for the degree of}\\
	\textcolor{purple}{\footnotesize{}Doctor in Engineering (Doctor in de Ingenieurswetenschappen) by}{\footnotesize\par}
	\textcolor{orange}{\Large{}ir. Gustavo Quintana Carapia}{\Large\par}

	\vspace{1cm}

	\begin{flushleft}
		\textcolor{purple}{\LARGE\noindent%\textcolor{purple}
		\textbf{Statistical analysis and experimental validation of data-driven dynamic measurement methods}}\\
	\par\end{flushleft}

	\vfill{}

	\begin{tabular}{@{}lll@{}}
	\textbf{\small PROMOTERS}		& &	{\small prof. dr. ir. Ivan Markovsky}				\\
				& &		{\footnotesize{}Vrije Universiteit Brussel, Belgium}\\[0.3em]
				& &	\small prof. dr. ir. Rik Pintelon				\\
				& &		{\footnotesize{}ELEC Department, Vrije Universiteit Brussel, Belgium.}	\\[0.3em]
				\\
	\textbf{\small MEMBERS OF THE JURY}	&	&\small prof. dr. ir. Patrick Guillaume (President)		\\
				&	&{\footnotesize{}MECH Department, Vrije Universiteit Brussel, Belgium.}	\\[0.3em]
				% &								\\
				&	&\small prof. dr. ir. Roger Vounckx (Vice-President)				\\
				&	&{\footnotesize{}ETRO Department, Vrije Universiteit Brussel, Belgium.}	\\[0.3em]
				% &								\\
				&	&\small dr. ir. Philippe Dreesen (Secretary)					\\
				&	&{\footnotesize{}ELEC Department, Vrije Universiteit Brussel, Belgium.}				\\[0.3em]
				% &								\\
				&	&\small dr. ir. Nikolaos Deligiannis \\
				&	&{\footnotesize{}ETRO Department, Vrije Universiteit Brussel, Belgium.} 			\\[0.3em]
				% &								\\
				&	&\small prof. dr. ir. Lyudmila Mihaylova 		\\
				&	&{\footnotesize{}ACSE Department, University of Sheffield, United Kingdom.} \\
								% &								\\
				&	&\small prof. dr. ir. Stephane Chretien  		\\
				&	&{\footnotesize{}ERIC Laboratory, University Lyon 2, France.} \\
								% &								\\
				&	&\small prof. dr. ir. Guillaume  Merc\`ere 		\\
				&	&{\footnotesize{}LIAS Laboratory, Universit\'e de Poitiers, France.} \\
\end{tabular}

}

\newpage{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Book details

\pagenumbering{gobble}
~\vfill{}

% {\footnotesize{}ISBN: } \\% You need to apply online for an ISBN, this takes a couple of days
{\footnotesize{}DOI: 10.13140/RG.2.2.27482.29129}\\ % I uploaded my thesis to researchgate to generate a DOI: if you remove it from ResearchGate, the DOI will no longer exist!
{\footnotesize{}~}{\footnotesize\par}

{\footnotesize{}University Press BVBA}\\
{\footnotesize{}Rechtstro 2/001, 9185 Wachtebeke, Belgium }\\
{\footnotesize{}https://www.universitypress.be/}\\
{\footnotesize{}~}{\footnotesize\par}

{\footnotesize{}Vrije Universiteit Brussel, dept. ELEC}\\
{\footnotesize{}Pleinlaan 2, 1050 Brussel, Belgium}\\
{\footnotesize{}http://vubirelec.be/}\\
{\footnotesize{}~}{\footnotesize\par}

{\footnotesize{}Cover design by ...}\\
{\footnotesize{}Background image by ...}{\footnotesize\par}
{\footnotesize{}~}{\footnotesize\par}

{\footnotesize{}$\copyright$ February 2019 Gustavo Quintana Carapia.}\\
{\footnotesize{}~}{\footnotesize\par}

{\footnotesize{}All rights reserved. No parts of this document may be reproduced or transmitted in any form or by any means without the prior written permission of the authors.}{\footnotesize\par}

{\footnotesize\par}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Acknowledgements
\newpage{}
% \shipout\null
\pagenumbering{roman}

\include{Extras/Acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  TOC and abbreviations list
\tableofcontents{}

% table of glossaries didn't work so I made a list of abbreviations manually
\include{Extras/Abbreviations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Thesis
% CustomCommands.lyx contains my custom lyx math macros.
\include{config/CustomCommands}

\pagenumbering{arabic}

\include{Summary/Summary}

\include{ChapterIntroduction/ChapterIntroduction}

%\part{Model driven filter design}\label{part:Filter}

\include{ChapterPreliminaries/ChapterPreliminaries}

\include{ChapterRampInput/ChapterRampInput}

\include{ChapterStatisticalAnalysis/ChapterStatisticalAnalysis}

\include{ChapterExperimentalValidation/ChapterExperimentalValidation}

%\part{Figures of merit for~characterizing nonlinear~devices}\label{part:PA}

\addtocontents{toc}{\vspace{2em}} \include{ChapterConclusion/ChapterConclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Start appendix

\appendix
\addtocontents{toc}{\vspace{0em}} \include{ChapterAppendices/ChapterAppendices}

\include{Extras/ListofPublications}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Bibliography

\addcontentsline{toc}{chapter}{Bibliography}

\bibliographystyle{apalike} % wmaainf
\bibliography{Gus-thesis}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Back cover
\newpage{}
\shipout\null
%\includepdf[pages={2}]{Cover}

\end{document}

### 2020-01-12 Summary/Summary.tex
Add files via upload
Initial set of documents

%!TEX root = ..\Thesis.tex
\chapter{Summary} \label{chap:Summary}

\section{Summary - Ramp input}

A measurement is a dynamical process that aims to estimate the true value of a measurand.
The measurand is the input that excites a sensor, and, as a consequence, the sensor output is a transient response. 
The main approach to estimate the input is applying the sensor transient response to another dynamical system. 
This dynamical system is designed by deconvolution to invert the sensor dynamics and compensate the sensor response. 
Digital signal processors enable an alternative approach to estimate the unknown input.
There exists a data-driven subspace-based signal processing method that estimates a measurand, assuming it is constant during the measurement.
To estimate the parameters of a measurand that varies at a constant rate, 
we extended the data-driven input estimation method to make it adaptive to the affine input.
In this paper, we describe the proposed subspace signal processing method for the measurement of an affine measurand and compare its performance to a maximum-likelihood input estimation method and to an existing time-varying compensation filter.
The subspace method is recursive and allows real-time implementations since it directly estimates the input without identifying a sensor model.
The maximum-likelihood method is model-based and requires very high computational effort.
In this form, the maximum-likelihood method cannot be implemented in real-time, however, we used it as a reference to evaluate the subspace method and the time-varying compensation filter results.
The effectiveness of the subspace method is validated in a simulation study with a time-varying sensor.
The results show that the subspace method estimation has relative errors that are one order of magnitude smaller and converges two times faster than the compensation filter.

\section{Summary - Statistical analysis}

A structured errors-in-variables (EIV) problem arising in metrology is studied.
The observations of a sensor response are subject to perturbation. 
The input estimation from the transient response leads to a structured EIV problem.
Total least squares (TLS) is a typical estimation method to solve EIV problems.
The TLS estimator of an EIV problem is consistent, and can be computed efficiently when the perturbations have zero mean, and are independently and identically distributed (i.i.d).
If the perturbation is additionally Gaussian, the TLS solution coincides with maximum-likelihood (ML).
However, the computational complexity of structured TLS and total ML prevents their real-time implementation.
The least-squares (LS) estimator offers a suboptimal but simple recursive solution to structured EIV problems with correlation, but the statistical properties of the LS estimator are unknown.
To know the LS estimate uncertainty in EIV problems, either structured or not, to provide confidence bounds for the estimation uncertainty, and to find the difference from the optimal solutions, the bias and variance of the LS estimates should be quantified.
Expressions to predict the bias and variance of LS estimators applied to unstructured and structured EIV problems are derived.
The predicted bias and variance quantify the statistical properties of the LS estimate and give an approximation of the uncertainty and the mean squared error for comparison to the Cram\'er-Rao lower bound of the structured EIV problem.

\section{Summary - Experimental validation}

Simultaneous fast and accurate measurement is still a challenging and active problem in metrology.
A sensor is a dynamic system that produces a transient response.
For fast measurements, the unknown input needs to be estimated using the sensor transient response.
When a model of the sensor exists, standard compensation filter methods can be used to estimate the input.
If a model is not available, either an adaptive filter is used or a sensor model is identified before the input estimation. 
Recently, a signal processing method was proposed to avoid the identification stage and estimate directly the value of a step input from the sensor response.
This data-driven step input estimation method requires only the order of the sensor dynamics and the sensor static gain.
To validate the data-driven step input estimation method, in this paper, the uncertainty of the input estimate is studied and illustrated on simulation and real-life weighing measurements.
It was found that the predicted mean-squared error of the input estimate is close to an approximate Cram\'er-Rao lower bound for biased estimators.

### 2020-01-12 config/acronyms.tex
Add files via upload
Initial set of documents

\usepackage[toc,acronym,section=section,nonumberlist,xindy]{glossaries}
\renewcommand*{\acronymname}{List of Abbreviations}

\renewcommand{\newacronym}[4][]{\newglossaryentry{#2}{type=\acronymtype,
name={#3},
short={#3},
long={#4},
shortplural={#3\glspluralsuffix},
longplural={#4\glspluralsuffix},
description={#4},
text={#3},
first={#4~(#3)},
plural={#3\glspluralsuffix},
firstplural={#4\glspluralsuffix~(#3\glspluralsuffix)},
#1}}

\newacronym{BLA}{BLA}{Best Linear Approximation}
\newacronym{MIMO}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{SISO}{SISO}{Single-Input Single-Output}
\newacronym{PISPO}{PISPO}{Periodic-In Same Period-Out}
\newacronym{FRF}{FRF}{Frequency Response Function}
\newacronym{ADS}{ADS}{Advanced Design System}
\newacronym{pdf}{pdf}{probability density function}
\newacronym{PSD}{PSD}{Power Spectral Density}
\newacronym{RMS}{RMS}{Root Mean Square}
\newacronym{LTI}{LTI}{Linear Time-Invariant}
\newacronym{RF}{RF}{Radio Frequency}
\newacronym{WLAN}{WLAN}{Wireless Local Area Network}
\newacronym{EM}{EM}{electromagnetic}
\newacronym{DGS}{DGS}{Defected Ground Structure}
\newacronym{RMSE}{RMSE}{Root Mean Square Error}
\newacronym{TL}{TL}{transmission line}
\newacronym{MC}{MC}{Monte Carlo}
\newacronym{SC}{SC}{Stochastic Collocation}
\newacronym{VF}{VF}{Vector Fitting}
\newacronym{LHS}{LHS}{Latin Hypercube Sampling}
\newacronym{QMC}{QMC}{Quasi Monte Carlo}
\newacronym{RBF}{RBF}{Radial Basis Functions}
\newacronym{PCBs}{PCBs}{Printed Circuit Boards}
\newacronym{PDNs}{PDNs}{Power Distribution Networks}
\newacronym{LDS}{LDS}{Low Discrepancy Sampling}
\newacronym{LBTLS}{LBTLS}{Local Bootstrapped Total Least Squares}
\newacronym{VNA}{VNA}{Vector Network Analyzer}
\newacronym{FOMs}{FOMs}{Figures of Merit}
\newacronym{FOM}{FOM}{Figure of Merit}
\newacronym{SNR}{SNR}{Signal-to-Noise Ratio}
\newacronym{SINAD}{SINAD}{Signal-to-Noise and Distortion ratio}
\newacronym{ACPR}{ACPR}{Adjacent Channel Power Ratio}
\newacronym{EVM}{EVM}{Error Vector Magnitude}
\newacronym{NPR}{NPR}{Noise Power Ratio}
\newacronym{DUT}{DUT}{Device Under Test}
\newacronym{TOI}{TOI}{Third Order Intercept point}
\newacronym{OFDM}{OFDM}{Orthogonal Frequency Division Multiplexing}
\newacronym{LTE}{LTE}{Long Term Evolution}

% \makeglossaries
\makenoidxglossaries

### 2020-01-27 ChapterAppendices/ChapterAppendices.tex
Story 182: Intro complete

@@
 
-\section{Appendix - Jacobians in Ramp input ML estimation}
+\section{Appendix - Derivation of bias and covariance expressions.  \label{appendix:biascovSEIV}}
+
+The bias and covariance of the least-squares (LS) estimate (\ref{eqn:xhatexp}) are obtained using the mathematical expectation in the definitions (\ref{eqn:biasdef}) and (\ref{eqn:covdef}).
+For an unstructured and uncorrelated EIV problem, the expected value, and the covariance of $\widehat{\mathbf{x}}$ are approximated by
+\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E}  \right\} \mathbf{x}, \quad \text{and}  \\
+\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{x} \mathbf{x}^\top + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \bm{\epsilon} \bm{\epsilon}^\top \mathbf{K} + \mathbf{K}^\top \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} \right\} \mathbf{Q}^{-1} \\
+& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{x} \mathbf{x}^\top \\
+& + \mathbf{x} \mathbf{x}^\top \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{Q}^{-1} - \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}^\top , \label{eqn:EC_Eu} \end{aligned} \end{equation}
+where we have considered the second order Taylor series approximation (\ref{eqn:TseriesExp}), and we have removed the terms of zero expected value, and the terms of order higher than 2.
+After an elementwise evaluation of the corresponding expected values in (\ref{eqn:EC_Eu}), the expressions result in 
+\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  + \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) = \mathbf{x}  +  \sigma_{\mathbf{E}}^2 \mathbf{Q}^{-1} \left( 2\mathbf{I} + 2n\mathbf{I} - T \mathbf{I} \right) \mathbf{x} , \quad \text{and} \\
+\mathbf{C}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \mathbf{x} \mathbf{x}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \ \mathbf{Q}^{-1} \mathbf{x} \mathbf{x}^\top \mathbf{Q}^{-1} , \end{aligned} \end{equation}
+from where equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) are obtained.
+
+On the other hand, due to the correlation, the expressions that approximate the expected value of the LS estimate of the structured and EIV problem (\ref{eqn:min_ls}) have a different form:
+\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  +  \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E}  \right\} \mathbf{x} \\
+& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{E}^\top \bm{\epsilon} - \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} - \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\} , \quad \text{and}  \\
+\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{x} \mathbf{x}^\top + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \bm{\epsilon} \bm{\epsilon}^\top \mathbf{K} + \mathbf{K}^\top \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} - \mathbf{K}^\top \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \mathbf{K} - \mathbf{K}^\top \bm{\epsilon} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} \right\} \mathbf{Q}^{-1}  \\
+& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{x} \mathbf{x}^\top \\
+& + \mathbf{x} \mathbf{x}^\top \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{Q}^{-1} - \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}^\top \\
+& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{E}^\top \bm{\epsilon} - \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} - \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\} \mathbf{x}^\top \\
+& + \mathbf{x} \mathbb{E} \left\{ \bm{\epsilon}^\top \mathbf{E} - \bm{\epsilon}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} - \bm{\epsilon}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} \right\} \mathbf{Q}^{-1}.
+\label{eqn:EC_E} \end{aligned} \end{equation}
+These expressions have the non zero expected value terms, up to the second order. 
+We have then
+\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & = \mathbf{x} + \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \approx \mathbf{x}  +  \mathbf{Q}^{-1} \mathbf{K}^\top \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} \right\}}_{\mathbf{B}_1} - \mathbf{Q}^{-1} \underbrace{\mathbb{E} \left\{ \mathbf{E}^\top \left( \mathbf{I} - \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \right) \mathbf{E} \right\}}_{\mathbf{B}_2} \mathbf{x} \\
+& - \mathbf{Q}^{-1} \mathbf{K}^\top \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\}}_{\mathbf{B}_3} + \mathbf{Q}^{-1} \underbrace{\mathbb{E} \left\{ \mathbf{E}^\top \left( \mathbf{I} - \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \right) \bm{\epsilon} \right\}}_{\mathbf{B}_4} , \quad \text{and} \\
+\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{Q}^{-1} \mathbf{K}^\top \left( \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \bm{\epsilon}^\top \right\}}_{\sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n}}  + \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_1}  - \underbrace{\mathbb{E} \left\{  \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \right\}}_{\mathbf{C}_2} - \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \mathbf{x}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_2^\top} \right) \mathbf{K} \mathbf{Q}^{-1}  - \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\mathbf{x}} \right).
+\end{aligned} \end{equation}
+from where the expressions (\ref{eqn:biasE}) and (\ref{eqn:varE}) are obtained.
+
+\section{Appendix - Proof of Lemma \ref{lem:lemma1} \label{appendix:lemma1}}
+
+\begin{pf}
+In the first case considered in the lemma, the elements of the expected value $\mathbf{Z} = \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \mathbf{E} \right\}$ are
+\begin{equation} z_{ij} = \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \mathbf{E} \right\}_{ij} = \mathbb{E} \left\{ \mathbf{e}_i \mathbf{A} \mathbf{E}_j \right\} = \mathrm{tr} \left( \mathbf{A} \ \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} \right) , \end{equation}
+where $\mathbf{e}_i$, and $\mathbf{E}_j$ are the $i\text{-th}$ row, and the $j\text{-th}$ column of $\mathbf{E}$, for $i = 1, \ldots, T-n$, and $j = 2, \ldots, n+1$.
+The matrix $\mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\}$ is the product of $\sigma_{\bm{\epsilon}}^2$ times a matrix whose elements are 0 in the first column, 2 in the $(j-i-1) \text{-th}$ diagonal, and -1 in the $(j-i-2) \text{-th}$, and $(j-i) \text{-th}$ diagonals, with zeros elsewhere.
+By using the definition of the second differential operator, we express
+\begin{equation} \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} = \sigma_{\bm{\epsilon}}^2 \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{2, j-i} \end{bmatrix}. \end{equation}
+
+The proof of the other cases in the Lemma is similar. 
+For the second case, the elements of the expected value $\mathbf{Z} = \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \mathbf{E} \right\}$ are
+\begin{equation} z_{ij} = \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \mathbf{E} \right\}_{ij} = \mathbb{E} \left\{ \mathbf{e}_i \mathbf{A} \mathbf{E}_j \right\} = \mathrm{tr} \left( \mathbf{A} \ \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} \right) , \end{equation}
+where now $\mathbf{e}_i$ is the $i\text{-th}$ row of $\mathbf{E}^\top$, and $\mathbf{E}_j$ is the $j\text{-th}$ column of $\mathbf{E}$, for $i = 2, \ldots, n+1$, and $j = 2, \ldots, n+1$.
+The matrix $\mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\}$ is $\sigma_{\bm{\epsilon}}^2$ times a matrix whose elements are 2 in the $(j-i) \text{-th}$ diagonal, and -1 in the $(j-i-1) \text{-th}$ and $(j-i+1) \text{-th}$ diagonals, with zeros elsewhere.
+Therefore we have
+\begin{equation} \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} = \sigma_{\bm{\epsilon}}^2 \mathbf{D}_{T-n \times T-n}^{2,j-i+1}. \end{equation}
+
+The expected values that involve the vector $\bm{\epsilon}$ are especial cases of the previous cases. 
+The vector $\bm{\epsilon}$ in the expected values $\mathbb{E} \left\{ \mathbf{E} \mathbf{A} \epsilon \right\}$, $\mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \epsilon \right\}$, and $\mathbb{E} \left\{ \mathbf{E} \mathbf{A} \bm{\epsilon}^\top \right\}$ is
+\begin{equation} \bm{\epsilon} = \begin{bmatrix} \epsilon(n+1) & \epsilon(n+2) & \cdots & \epsilon(T) \end{bmatrix}^\top, \end{equation}
+as it is imposed by the input estimation method formulation.
+\qed \end{pf}
+
+\section{Appendix - Jacobians in Ramp input ML estimation  \label{appendix:Jacobians}}
 
@@
 
-\section{Appendix - Derivation of bias and covariance expressions.}
-
-The bias and covariance of the least-squares (LS) estimate (\ref{eqn:xhatexp}) are obtained using the mathematical expectation in the definitions (\ref{eqn:biasdef}) and (\ref{eqn:covdef}).
-For an unstructured and uncorrelated EIV problem, the expected value, and the covariance of $\widehat{\mathbf{x}}$ are approximated by
-\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E}  \right\} \mathbf{x}, \quad \text{and}  \\
-\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{x} \mathbf{x}^\top + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \bm{\epsilon} \bm{\epsilon}^\top \mathbf{K} + \mathbf{K}^\top \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} \right\} \mathbf{Q}^{-1} \\
-& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{x} \mathbf{x}^\top \\
-& + \mathbf{x} \mathbf{x}^\top \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{Q}^{-1} - \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}^\top , \label{eqn:EC_Eu} \end{aligned} \end{equation}
-where we have considered the second order Taylor series approximation (\ref{eqn:TseriesExp}), and we have removed the terms of zero expected value, and the terms of order higher than 2.
-After an elementwise evaluation of the corresponding expected values in (\ref{eqn:EC_Eu}), the expressions result in 
-\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  + \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) = \mathbf{x}  +  \sigma_{\mathbf{E}}^2 \mathbf{Q}^{-1} \left( 2\mathbf{I} + 2n\mathbf{I} - T \mathbf{I} \right) \mathbf{x} , \quad \text{and} \\
-\mathbf{C}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \mathbf{x} \mathbf{x}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \ \mathbf{Q}^{-1} \mathbf{x} \mathbf{x}^\top \mathbf{Q}^{-1} , \end{aligned} \end{equation}
-from where equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) are obtained.
-
-On the other hand, due to the correlation, the expressions that approximate the expected value of the LS estimate of the structured and EIV problem (\ref{eqn:min_ls}) have a different form:
-\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  +  \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E}  \right\} \mathbf{x} \\
-& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{E}^\top \bm{\epsilon} - \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} - \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\} , \quad \text{and}  \\
-\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{x} \mathbf{x}^\top + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \bm{\epsilon} \bm{\epsilon}^\top \mathbf{K} + \mathbf{K}^\top \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} - \mathbf{K}^\top \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \mathbf{K} - \mathbf{K}^\top \bm{\epsilon} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} \right\} \mathbf{Q}^{-1}  \\
-& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{x} \mathbf{x}^\top \\
-& + \mathbf{x} \mathbf{x}^\top \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{Q}^{-1} - \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}^\top \\
-& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{E}^\top \bm{\epsilon} - \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} - \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\} \mathbf{x}^\top \\
-& + \mathbf{x} \mathbb{E} \left\{ \bm{\epsilon}^\top \mathbf{E} - \bm{\epsilon}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} - \bm{\epsilon}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} \right\} \mathbf{Q}^{-1}.
-\label{eqn:EC_E} \end{aligned} \end{equation}
-These expressions have the non zero expected value terms, up to the second order. 
-We have then
-\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & = \mathbf{x} + \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \approx \mathbf{x}  +  \mathbf{Q}^{-1} \mathbf{K}^\top \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} \right\}}_{\mathbf{B}_1} - \mathbf{Q}^{-1} \underbrace{\mathbb{E} \left\{ \mathbf{E}^\top \left( \mathbf{I} - \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \right) \mathbf{E} \right\}}_{\mathbf{B}_2} \mathbf{x} \\
-& - \mathbf{Q}^{-1} \mathbf{K}^\top \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\}}_{\mathbf{B}_3} + \mathbf{Q}^{-1} \underbrace{\mathbb{E} \left\{ \mathbf{E}^\top \left( \mathbf{I} - \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \right) \bm{\epsilon} \right\}}_{\mathbf{B}_4} , \quad \text{and} \\
-\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{Q}^{-1} \mathbf{K}^\top \left( \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \bm{\epsilon}^\top \right\}}_{\sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n}}  + \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_1}  - \underbrace{\mathbb{E} \left\{  \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \right\}}_{\mathbf{C}_2} - \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \mathbf{x}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_2^\top} \right) \mathbf{K} \mathbf{Q}^{-1}  - \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\mathbf{x}} \right).
-\end{aligned} \end{equation}
-from where the expressions (\ref{eqn:biasE}) and (\ref{eqn:varE}) are obtained.
-
-\section{Appendix - Proof of Lemma \ref{lem:lemma1}}
-
-\begin{pf}
-In the first case considered in the lemma, the elements of the expected value $\mathbf{Z} = \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \mathbf{E} \right\}$ are
-\begin{equation} z_{ij} = \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \mathbf{E} \right\}_{ij} = \mathbb{E} \left\{ \mathbf{e}_i \mathbf{A} \mathbf{E}_j \right\} = \mathrm{tr} \left( \mathbf{A} \ \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} \right) , \end{equation}
-where $\mathbf{e}_i$, and $\mathbf{E}_j$ are the $i\text{-th}$ row, and the $j\text{-th}$ column of $\mathbf{E}$, for $i = 1, \ldots, T-n$, and $j = 2, \ldots, n+1$.
-The matrix $\mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\}$ is the product of $\sigma_{\bm{\epsilon}}^2$ times a matrix whose elements are 0 in the first column, 2 in the $(j-i-1) \text{-th}$ diagonal, and -1 in the $(j-i-2) \text{-th}$, and $(j-i) \text{-th}$ diagonals, with zeros elsewhere.
-By using the definition of the second differential operator, we express
-\begin{equation} \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} = \sigma_{\bm{\epsilon}}^2 \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{2, j-i} \end{bmatrix}. \end{equation}
-
-The proof of the other cases in the Lemma is similar. 
-For the second case, the elements of the expected value $\mathbf{Z} = \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \mathbf{E} \right\}$ are
-\begin{equation} z_{ij} = \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \mathbf{E} \right\}_{ij} = \mathbb{E} \left\{ \mathbf{e}_i \mathbf{A} \mathbf{E}_j \right\} = \mathrm{tr} \left( \mathbf{A} \ \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} \right) , \end{equation}
-where now $\mathbf{e}_i$ is the $i\text{-th}$ row of $\mathbf{E}^\top$, and $\mathbf{E}_j$ is the $j\text{-th}$ column of $\mathbf{E}$, for $i = 2, \ldots, n+1$, and $j = 2, \ldots, n+1$.
-The matrix $\mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\}$ is $\sigma_{\bm{\epsilon}}^2$ times a matrix whose elements are 2 in the $(j-i) \text{-th}$ diagonal, and -1 in the $(j-i-1) \text{-th}$ and $(j-i+1) \text{-th}$ diagonals, with zeros elsewhere.
-Therefore we have
-\begin{equation} \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} = \sigma_{\bm{\epsilon}}^2 \mathbf{D}_{T-n \times T-n}^{2,j-i+1}. \end{equation}
-
-The expected values that involve the vector $\bm{\epsilon}$ are especial cases of the previous cases. 
-The vector $\bm{\epsilon}$ in the expected values $\mathbb{E} \left\{ \mathbf{E} \mathbf{A} \epsilon \right\}$, $\mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \epsilon \right\}$, and $\mathbb{E} \left\{ \mathbf{E} \mathbf{A} \bm{\epsilon}^\top \right\}$ is
-\begin{equation} \bm{\epsilon} = \begin{bmatrix} \epsilon(n+1) & \epsilon(n+2) & \cdots & \epsilon(T) \end{bmatrix}^\top, \end{equation}
-as it is imposed by the input estimation method formulation.
-\end{pf}
-
 \newpage

### 2020-01-27 ChapterConclusion/ChapterConclusion.tex
Story 182: Intro complete

@@
 
+Conclusions - Statistical analysis
+
+ We conducted a statistical analysis of a structured errors-in-variables (EIV) estimation problem with correlation to find the first and second moments of its least-squares solution.
+This estimation problem occurs in metrology when we estimate the value of a measurand directly from the sensor transient response.
+The data-driven estimation of the physical quantity is formulated as a structured EIV problem with correlation that uses the observed transient response to construct both the regression matrix and the regressor.
+The real-time implementation of the method uses a recursive least squares algorithm that is simple and has low computational complexity.
+The assessment of the uncertainty is done using the estimate bias and variance.
+
+The conducted statistical analysis produced expressions that predict the estimate bias and variance for given sample size and perturbation level of the observed response.
+The Monte Carlo simulation validated the predictions.
+We compared the results of solving an unstructured and uncorrelated EIV problem with a structured and correlated EIV problem to understand how the structure and the correlation impacts in the estimation.
+We found that the predictions in the structured case are more susceptible to perturbations.
+This is due to the two approximations involved, a second-order Taylor series expansion of the estimate, and the substitution of perturbed data on the prediction expressions.
+The relative error results indicate that the estimate bias, and variance are predicted using the derived expressions, and the observed data.
+The mean squared error of the estimate is close to the results of the maximum likelihood estimate given by the Cram\'er-Rao lower bound.
+
+The bias and variance can be accurately predicted, provided that the Taylor series expansion is valid.
+This constraint has to be taken into account to ensure the effectiveness of the method in practical applications.
+In the example, it was observed that when the SNR lies outside the validity region, the bias and variance estimation was at most three times larger than the empirical values.
+
+The methodology presented in this paper can be applied to estimate the uncertainty of the solutions to other structured EIV problems.
+The bias and variance expressions obtained after the statistical analysis depend on each specific structure.
+
+Conclusions - Experimental validation
+
+In this paper we investigated the statistical properties of a data-driven step input estimation method in a real-life application.
+The step input estimation method is a structured and correlated errors-in-variables problem that is solved with recursive least squares.
+A statistical analysis was conducted using the ordinary least squares condensed notation. 
+This statistical analysis of the input estimate provides expressions that approximate the estimation bias and variance assuming that the measurement noise is Gaussian white noise. 
+The variance approximation is useful to assess the uncertainty of the input estimate.
+In simulation we observed that the mean squared error of the input estimate is close to the theoretical minimum that uses the Cram\'er-Rao lower bound for biased estimators.
+Since the data-driven step input estimation method is not statistically efficient, there is room for improvement. This is a topic for future research.
+In the practical experiments, the measurement noise is not white.
+The noise variance obtained from the sensor steady state response underestimates the measurement noise variance, that was observed 5 dB larger in the power spectrum due to nonlinearities of the sensor.
+Considering this difference in the measurement noise variance, we introduced a conservative bound of the measurement noise variance so that the first and second moments of the input estimate are more accurately predicted.
+Using the variance approximation, we can assess the uncertainty of the input estimate with respect to the number of samples processed by the data-driven step input estimation method.
+The step input estimation method is useful in practical applications where the whiteness assumption of the measurement noise is not fulfilled.
+
+Conclusions - Ramp input
+
+An adaptive subspace method was proposed for the estimation of affine input parameters given the measurement of the caused sensor transient response. 
+The subspace estimation method is a recursive method that allows online implementation.
+This method tracks the input of a system, using exponential forgetting, to process the system response.
+The subspace method is model-free and estimates directly the input parameters without identifying a sensor model.
+Therefore, it can be applied to the measurement of different physical magnitudes.
+In the specific weighing example described in the manuscript, the input is an affine function.
+The method is also applicable when the sensor is time-varying.
+The subspace method is computationally cheap, simple and suitable for implementation on digital signal processor of low computational power.
+
+A maximum-likelihood estimator based on local optimization was designed to obtain a comparative reference for the other methods.
+The maximum-likelihood method estimates the affine input parameters and also model parameters and the sensor's initial conditions.
+This method simulates, in a receding horizon scheme, the response of a sensor model to estimate the input and minimizes the sum of the squares of the residual between the measured and the estimated responses.
+The main drawback of the maximum-likelihood method is its computational cost and efficient implementation of the method is left for future work.
+
+A linear time-invariant weighing system is used as an test example for the estimation methods.
+The weighing system becomes time-varying when an affine input excites the system.
+The estimation methods are compared in a simulation study where the time-varying sensor response is perturbed by measurement noise, that is assumed white of zero mean and known variance.
+The subspace method results are also compared to those of an existing digital time-varying filter.
+The coefficients of the time-varying filter require offline optimization.
+The estimation results obtained with the subspace method converges two times faster and is one order of magnitude smaller than those obtained with the time-varying filter.
+The empirical mean squared errors of the subspace method estimation is two orders of magnitude larger than the theoretical minimum given by the Cram\'er-Rao Lower bound.
+
+Future work of this research is the practical implementation of the subspace method for real-time measurements.
+
+\begin{comment}
 \section{Conclusions - Ramp input}
@@
 The estimation results obtained with the subspace method converges two times faster and is one order of magnitude smaller than those obtained with the time-varying filter.
-The empirical mean squared errors of the subspace method estimation is two orders of magnitude larger than the theoretical minimum given by the Cr\'amer-Rao Lower bound.
+The empirical mean squared errors of the subspace method estimation is two orders of magnitude larger than the theoretical minimum given by the Cram\'er-Rao Lower bound.
 
@@
 The variance approximation is useful to assess the uncertainty of the input estimate.
-In simulation we observed that the mean squared error of the input estimate is close to the theoretical minimum that uses the Cr\'amer-Rao lower bound for biased estimators.
+In simulation we observed that the mean squared error of the input estimate is close to the theoretical minimum that uses the Cram\'er-Rao lower bound for biased estimators.
 Since the data-driven step input estimation method is not statistically efficient, there is room for improvement. This is a topic for future research.
@@
 \section*{Future work}
+\end{comment}
+
 text
+
+\newpage

### 2020-01-27 ChapterExperimentalValidation/ChapterExperimentalValidation.tex
Story 182: Intro complete

@@
 \chapter{Experimental validation of the step input estimation method }\label{chap:ExperimentalValidation}
-% \begin{quote}
-% \emph{The results of this chapter are based on the \cite{VanNechel2018}.}\vfill{}
-% \end{quote}
-\vfill{}
 
- \section{Introduction  - Experimental validation}
+\begin{quote}
+\emph{The results of this chapter ...}\vfill{}
+\end{quote}
 
+ \section{Introduction}
+
+A measurement is a dynamic process. 
+The sensor is a dynamic system that experiments changes after an input excitation.
+The sensor output is first a transient state response and followed by a steady state response.
+
+The influence of the input is observed in the sensor output, and thus it makes sense to estimate the input using the sensor response.
+When the sensor reaches its steady-state, the input is a straightforwardly estimated from the sensor output making use of the sensor static gain.
+To obtain a fast estimation, the sensor transient response must be considered instead of the steady state response.
+
+A common approach is to filter the sensor transient response with another dynamic system to reduce the sensor transient time by inverting the sensor dynamics.
+To design such a compensator system it is necessary a sensor model.
+A different approach uses digital signal processing methods that are independent of the sensor model.
+One of these methods is the data-driven method that estimates the unknown level of step inputs by processing the sensor step response. 
+To assess the uncertainty of the data-driven step input estimation method, a experimental validation of this method was conducted in a real-life metrology application. 
+This chapter describes the measurement experiments with a weighing sensor system and discuss the results observed.
+The results are put in perspective to simulations and consider the measurement noise that is additive to the response of the sensor.
+
+The experimental results validate the data-driven step input estimation method, and indicate in which intervals of SNR and sample size the effectiveness of the method is guaranteed.
+
+\begin{comment}
  In this paper we consider that a measurement is a dynamic process, where an input excites a dynamic system, the sensor, and causes a dynamic transient response that also depends on the initial conditions of the sensor.
@@
 The convolution of the compensator impulse response with the sensor transient response yields the input estimate.
-Therefore, the design of a compensator is based on the sensor model and requires a deconvolution \cite{Eichstadt10}.
-Examples of input estimation using compensation of the sensor transient response include a recursive estimation of the compensator parameters \cite{Shu93}, 
-finite impulse response (FIR) \cite{Elster07, Niedzwiecki16b} filters and 
-infinite impulse response (IIR) filters \cite{Pintelon90, Elster08}.
+Therefore, the design of a compensator is based on the sensor model and requires a deconvolution \citep{Eichstadt10}.
+Examples of input estimation using compensation of the sensor transient response include a recursive estimation of the compensator parameters \citep{Shu93}, 
+finite impulse response (FIR) \citep{Elster07, Niedzwiecki16b} filters and 
+infinite impulse response (IIR) filters \citep{Pintelon90, Elster08}.
 The filters in these works estimate in real-time the unknown input value.
@@
 An alternative to the compensation approach is to use digital signal processing methods that are independent of the sensor model.
-A data-driven method that estimates the unknown level of step inputs by processing the sensor step response was introduced in \cite{Markovsky15ieee}. 
+A data-driven method that estimates the unknown level of step inputs by processing the sensor step response was introduced in \citep{Markovsky15ieee}. 
 This data-driven input estimation method avoids the sensor modeling stage and estimates directly the input.
 This method reduces the estimation time compared to a conventional compensator.
-The step input estimation method performance was demonstrated by simulations and experiments on a digital signal processor (DSP) of low cost \cite{Markovsky15cep}.
+The step input estimation method performance was demonstrated by simulations and experiments on a digital signal processor (DSP) of low cost \citep{Markovsky15cep}.
 The uncertainty of the step input estimation method has not been assessed before.
 
-To validate the input estimation methods it is necessary to assess the uncertainty associated with their estimates \cite{daSilva12, Ferrero06}.
-There are uncertainty propagation studies for model-based compensators such as the FIR and IIR filters for acceleration measurements where the uncertainty is computed in real time \cite{Elster07, Elster08, Link09}.
+To validate the input estimation methods it is necessary to assess the uncertainty associated with their estimates \citep{daSilva12, Ferrero06}.
+There are uncertainty propagation studies for model-based compensators such as the FIR and IIR filters for acceleration measurements where the uncertainty is computed in real time \citep{Elster07, Elster08, Link09}.
 In these works, the uncertainty expression is based on the transfer function or state space representations of the LTI sensor and filter systems.
-Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \cite{Pietrzak14} for mass and in \cite{Ogorevc16} for temperature sensors.
-A deconvolution method is implemented to estimate the input waveform in \cite{Hale09} and the uncertainty is obtained from the input estimate covariance. 
-The impact that the signal processing data-driven dynamic error correction has on the uncertainty is investigated in \cite{Saggin01}. 
-A statistical analysis of the data-driven step input estimation method \cite{Markovsky15cep} was investigated in \cite{Quintana19} and the method uncertainty was obtained with a Monte Carlo simulation study.
+Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \citep{Pietrzak14} for mass and in \citep{Ogorevc16} for temperature sensors.
+A deconvolution method is implemented to estimate the input waveform in \citep{Hale09} and the uncertainty is obtained from the input estimate covariance. 
+The impact that the signal processing data-driven dynamic error correction has on the uncertainty is investigated in \citep{Saggin01}. 
+A statistical analysis of the data-driven step input estimation method \citep{Markovsky15cep} was investigated in \citep{Quintana19} and the method uncertainty was obtained with a Monte Carlo simulation study.
 
@@
 The novelty of the paper is threefold.
-First, using the results of \cite{Quintana19}, that describes a statistical analysis of structured errors in variables (EIV) problems, in this paper we describe the statistical properties of a the data-driven step input estimation method in both simulation and a real-life experiments.
+First, using the results of \citep{Quintana19}, that describes a statistical analysis of structured errors in variables (EIV) problems, in this paper we describe the statistical properties of a the data-driven step input estimation method in both simulation and a real-life experiments.
 Second, this manuscript also presents the Cram\'er-Rao lower bound for unbiased estimators of the structured and correlated EIV problem that the step input estimation method formulates.
@@
 The third novelty in this manuscript is the model order selection for the step input estimation method where we use the MSE to select the order that provides the smaller MSE with the lowest computational complexity.
+\end{comment}
 
-\section{SIMULATION RESULTS}
+\section{Simulation results}
 
@@
 \centering
-\begin{tabular}{ |c|c c c c| } 
+\begin{tabular}{c|c c c c} 
 \hline
@@
 
-\section{PRACTICAL IMPLEMENTATION} 
+\section{Practical implementation} 
 An experimental setup was constructed to test the step input estimation method.
-The implementation is a weighing system that uses a load cell Tedea Huntleigh 1004 \cite{tedea1004}.
+The implementation is a weighing system that uses a load cell Tedea Huntleigh 1004 \citep{tedea1004}.
 The maximum rating of the load cell is 600 g.

### 2020-01-27 ChapterIntroduction/ChapterIntroduction.tex
Story 182: Intro complete

@@
 % Context:Metrology One of the objectives in metrology is the estimation of the true value of physical quantities.
-Measurements are dynamical processes.
-Sensors are dynamical systems that interact with physical quantities and deliver electrical signals.
-This interaction is the input to the sensor and causes energy transferences that modify the sensor state.
-The sensor response is the output that depends on both the input applied and the sensor initial conditions.
+A measurement is a dynamical process and a sensor is a dynamical system.
+The physical quantities of interest are the inputs of the system, and the outputs are the electrical signals collected from the sensor.
+The inputs interact with the sensor, and there are energy transferences between them that modify the sensor state.
+The sensor output response depends on the applied inputs, and on the sensor initial conditions.
 
 % Context:Tradeoff speed vs accuracy
-There is a trade-off between speed and accuracy in a measurement.
-The input excitation drives a linear time invariant sensor into a transient state, that is followed by a steady state.
+There is a trade-off between speed and accuracy in a measurement with a linear time invariant sensor.
+The input excitation drives the sensor into a transient state, that is followed by a steady state.
 During the sensor transient state, the response does not represent directly the input, but 
@@
 However, waiting for the steady state is not advisable for practical applications that need fast measurements.
-In these applications the input must be estimated during the sensor transient state.
+In these practical applications, the input must be estimated during the sensor transient state.
 
-% Context:Compensation systems, # Context:Digital signal processing
-One approach to estimate the input is filtering the sensor transient response with another dynamical system that inverts the dynamics of the sensor.
+% Context:Compensation systems, 
+One approach to get a fast input estimation is filtering the sensor transient response with another dynamical system that inverts the dynamics of the sensor.
 The filter output is an input estimate that compensates the measurement time.
@@
 The compensator is designed to deconvolve the sensor response and is based on a sensor model.
-Another approach is using digital signal processors (DSP).
-DSPs are versatile since they allow the implementation of methods that do not necessarily recreate dynamic systems, such as digital filters.
-A suitable model-independent, data-driven, method in a DSP can provide faster input estimations than with the model-based compensators.
+
+% Context:Digital signal processing
+Another approach relies in the use of digital signal processors (DSP) to estimate the input value from the sensor transient response.
+DSPs offer an extra versatility level since they allow the implementation of methods that do not necessarily recreate dynamic systems, such as digital filters.
+A suitable data-driven method in a DSP can provide faster input estimations than with the model-based compensators.
 
 % Context:Data-driven step input estimation method
-An example of a data-driven method is the direct estimation of the step input level from the sensor step response \cite{Markovsky15cep}.
+An example of a data-driven method devised for a DSP is the direct step input level estimation from the sensor step response \citep{Markovsky15cep}.
 This method formulates a Hankel structured errors-in-variables (EIV) problem with correlation.
 The regression matrix has a block-Hankel structure.
-The regression matrix and the regressor are constructed from the transient response perturbed by measurement noise of zero mean and given variance.
+The correlation exists because both the regression matrix and the regressor are constructed from the transient response perturbed by measurement noise.
+The measurement noise is assumed to have zero mean and given variance.
 The method is implemented in real-time using a recursive least-squares (RLS) solution of the structured EIV problem.
@@
 The estimate uncertainty must be assessed to validate the estimation methods for metrology applications.
-The uncertainty of the data-driven step input estimation method \cite{Markovsky15cep} is unknown.
-The uncertainty can be defined in terms of the estimate bias and variance \cite{Pintelon12Book}, and these can be obtained by conducting an elementwise statistical analysis.
+The uncertainty of the data-driven step input estimation method \citep{Markovsky15ieee} is unknown.
+The uncertainty can be defined in terms of the estimate bias and variance \citep{Pintelon12Book}, and these can be obtained by conducting an elementwise statistical analysis.
 The validation of the step input estimation method requires also to demonstrate its effectiveness on real-life measurements.
@@
 
-\subsection{State of the art}
-% StArt: Load cell sensors 
-% experimental application is weighing
+% general preview of the thesis contents
+This thesis describes three works iniciated after the data-driven step input estimation method.
+The first work is a statistical analysis to elucidate the bias and the covariance of the estimate that this method provides.
+With these statistical moments, the estimate uncertainty is built and we can assess the effectiveness of the method.
+
+\section{State of the art}
+
+Recent studies in metrology deal with the dynamic process effects that exist in measurements.
+Processing the response of linear time invariant sensors to get fast and accurate input estimations motivates the need to estimate inherent dynamical errors \citep{Hessling06} and to develop methods for dynamic correction in time domain \citep{Hessling08a}.
+The industrial requeriments impose new challenges by increasing the number of sensors distributed in the production lines that provide feedback for quick decision taking to control autonomous activation \citep{Esward09}.
+
+To implement the input estimation methods in the practice, the compensation filters based on deconvolution \citep{Eichstadt10} are the preferred methodology to develop 
+analog filters \citep{Jafaripanah05}, 
+adaptive digital filters \citep{Shu93}, 
+lattice adaptive filters \citep{Hernandez06}, 
+compensators for simultaneous responses of different sensors \citep{Boschetti13}, 
+regularized deconvolution compensators \citep{Dienstfrey14}, 
+and a multiple choice set of filters \citep{Huang16}. 
+All these methods have in common the use of the measurement system model to build the compensator, for instance, after the model parameters.
+
+% Digital signal processing estimation methods
+A list of measurement methods based on digital signal processors (DSP) include 
+a study of the data-driven dynamic error correction and its impact on the measured temperature \citep{Saggin01},
+a modulation quality measurement of microwave access systems \citep{Angrisani10}, 
+a real-time rotational speed estimation using correlation \citep{Wang14},
+a development of an biology-inspired electronic nose \citep{Jing16},
+and impedance measurements for material damage estimation using cross-correlation \citep{deCastro19}.
+The data-driven step input estimation method \citep{Markovsky15cep} is added to this list, highlighting the fact that can be implemented for the measurement of different physical magnitudes bacause is model independent.
+
+% Uncertainty assessment 
+In metrology, each measurement is an estimation that can be modeled as a random variable.
+Every measurement is inherently perturbed by noise and, therefore, the input estimation is commonly expressed with its two first statistical moments \citep{Ferrero06}.
+In an effort to standardize the uncertainty assessment of the input estimation, a guideline \citep{GUM08}has been prepared and has a overall acceptance in the metrolgy community.
+A review of measurement uncertainty analysis of common use is presented in \citep{daSilva12}.
+The Monte Carlo method has been shown to be an effective uncertainty evaluation tool in \citep{Cox06},
+and supports the usefulness of simulation for quantifying measurement uncertainties by software \citep{Esward16}.
+Another example of the Monte Carlo method is the dynamic measurement uncertainty evaluation of clinical thermometers described in \citep{Ogorevc16}.
+
+Nevertheless, researchers still pinpoint the need to study more the uncertainty assessment methods \citep{Esward09, Hessling10}.
+It is recommended to consider the uncertainties of all the measurement chain components \citep{Diniz17}, and to avoid the direct uncertainty propagation from the calibration towards the to-be-measured quantity.
+Methods for evaluating the uncertainty associated with the output of compensation filters have been investigated, such as for
+a discrete-time infinite-response filter in \citep{Link09},
+a discrete-time finite-response filter in \citep{Elster07, Elster08}, and
+the Kalman filter in \citep{Eichstadt16b}.
+All these works propagate the uncertanity through the filter but it is also necessary to upward the propagation up to the sensor model to include all systematic error contributions \citep{Hessling11}.
+
+\section{Original contributions}
+
+% Preview of the thesis contents and location of the developed topics in relation to the state of the art
+
+This thesis describes research work iniciated after the data-driven step input estimation method.
+This method was proposed for metrology applications but it lacked of an uncertainty assessment, and, therefore, its appropriateness was questionable.
+The first research work conducted is a statistical analysis to elucidate the bias and the covariance of the estimate that the data-driven step input estimation method provides.
+After obtaining these statistical moments, the estimate uncertainty is built, and the effectiveness of the method is evaluated.
+The second research work conducted is a series of experiments that aims to validate the data-driven step input estimation method in real-life applications.
+The experiments were realizations of the step input excitation using temperature and mass sensors.
+The step input estimation method showed robustness when the measurement noise is not Gaussian and white as it was assumed in the theoretical analyses.
+The third research work conducted is the estimation of affine input parameters following the data-driven approach. 
+The data-driven method is proposed for real-time applications, and uses exponential weighing in the recursive least squares solution of a structured errors-in-variables problem, to give preference to recent samples over the older samples.
+The data-driven affine input estimation method methodology and performance is compared those of a maximum likelihood method based.
+
+\subsection{Statistical analysis}
+
+The data-driven step input estimation method formulates a structured errors-in-variables (EIV) problem.
+In linear estimation EIV problems, the regression matrix and the regressor are perturbed by measurement noise \citep{VanHuffel91Book}, \citep{Markovsky07overview}.
+The regression matrix, in structured EIV problems, has a structure that depends on the problem formulation.
+The Hankel and Toeplitz matrices appear in problems of metrology \citep{Markovsky15cep}, system identification \citep{Soderstrom07}, image restoration \citep{Feiz17}, nuclear magnetic resonance spectroscopy \citep{Cai16}, direction-of-arrival estimation \citep{Pan18}, and time-of-arrival estimation \citep{Jia18}.
+
+The data-driven step input estimation method directly estimates the input from the sensor transient response. 
+The perturbations in the EIV problem of interest come from the sensor output, that is the only observed signal \citep{Markovsky15cep}, and is used to construct both the regression matrix and the regressor.
+The structure in this regression matrix is block-Hankel.
+The direct estimation is a data-driven methodology that reduces the input estimation time of the classical two-stage approach where a sensor model is first identified, and later the input is estimated using the sensor model \citep{Azam15, Niedzwiecki16a}.
+
+Instead of using a total least-squares (TLS) solution of the structured EIV problem, the recursive least squares solution (RLS) is proposed to get the step input estimate. 
+The TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
+the TLS solution is equivalent to the maximum likelihood (ML) solution when the perturbations of the EIV problems are i.i.d. normally distributed \citep{Markovsky07overview}. 
+However, regarding structured EIV problems, the TLS estimator does not give general results since each structure requires a specific treatment \citep{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where the global optimum is not guaranteed \citep{Rhode14recursive}.
+Moreover, the computational complexity of TLS and ML inhibits their real-time implementation.
+The least-squares (LS) estimator is a suboptimal but simple solution to structured EIV problems that admits a recursive form (RLS) suitable for real-time implementation.
+
+The previously published works that propose LS estimators to solve structured EIV problems do not study the required statistical moments to know the uncertainty of the data-driven step input estimation method.
+In the literature we find 
+the design of a fast algorithm for matrices with small displacement rank \citep{Mastronardi07fast}, 
+the study of the LS estimator consistency \citep{Palanthandalam10parameter},
+the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models \citep{Kiviet12high}, \citep{Kiviet14improved}, and
+a discussion of the causes of bias and inconsistency in homogeneous estimators \citep{Yeredor04homogeneous}.
+The literature does not address the uncertainty of the reported LS estimators for structured EIV problems.
+
+The uncertainty of an estimation method is expressed using the bias and covariance of the estimate \citep{Pintelon12Book}.
+To know the uncertainty of the data-driven step input estimation method, I quantified the bias and the covariance of the LS solution of EIV problems, for unstructured and structured cases. 
+This quantification extends the perturbation analysis of the LS estimator of unstructured and uncorrelated problems that was investigated in \citep{Stewart90SPT} and in \citep{Vaccaro94}.
+The analysis of the unstructured EIV case provides an insight of the impact that the structure of the regression matrix and the correlation between the regression matrix and the regressor perturbations have on the uncertainty of the LS estimate.
+The study presented in this thesis illustrates a methodology to conduct statistical analysis for any structured EIV problem.
+
+I derived expresions that quantify the bias and covariance by obtaining the mathematical expectation of the LS estimate approximated by the second-order Taylor series expansion.
+Using Monte Carlo simulations, I validated the accuracy of the bias and covariance expressions.
+These expressions estimate the bias and covariance that the data-driven step input estimation method will exhibit for given sample size and perturbation level.
+I compared the mean squared error of the LS estimate to the minimum variance specified by the Cram\'er-Rao lower bound of the structured EIV problem, to determine the conditions in which the data-driven step input estimation method is appropriate for practical applications.
+
+\subsection{Experimental validation}
+
+The validation of the data-driven step input estimation method in a practical application was necessary to demonstrate the usefulness of the method.
+The method performance was illustrated by simulations and temperature experiments on a digital signal processor (DSP) of low cost \citep{Markovsky15cep}.
+The method estimates the unknown level of step inputs by processing the sensor step response, and 
+avoiding the sensor modeling stage.
+The formulation of the estimation method is a correlated errors-in-variables (EIV) problem with block-Hankel structure.
+
+Other methods for input estimation mainly compensate the sensor transient response, for example, by 
+recursive estimation of the compensator parameters \citep{Shu93}, 
+finite impulse response (FIR) filtering \citep{Elster07}, \citep{Niedzwiecki16b} filters and 
+infinite impulse response (IIR) filtering \citep{Pintelon90}, \citep{Elster08}.
+The uncertainty propagation for these model-based compensators is based on the transfer function or state space representations of the LTI sensor and filter systems \citep{Link09}, \citep{Hale09}.
+Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \citep{Pietrzak14} for mass and in \citep{Ogorevc16} for temperature sensors.
+
+To validate the data-driven step input estimation method I built a weighing system setup based on a load cell sensor.
+The weighing setup was constructed to ensure repeatability and reproducibility along different experimental realizations.
+Load cell sensors are versatil devices that are found in 
+heart and breathing physiological signal monitoring \citep{Lee16},
+clinical analysis of sleep quality \citep{Zahradka18},
+automobile safety studies \citep{Ballo16},
+wind turbine design \citep{Rossander15}, 
+civil engineering \citep{Olmi16}, and 
+sport bicycle design \citep{Casas16}, to name a few.
+
+\begin{comment}
+% Load cell sensors 
 Weighing has been basic for the development of scientific and trade activites. 
-The load cell is now a standard transducer for weight determination and also for the improvement of measurement techniques, such as the geometric approach to processing of load cell responses \cite{Kesilmis16}, the design of new conveyor machinery \cite{Yamani18}, and electronic truck scales \cite{Guo18}.
+The load cell is now a standard transducer for weight determination and also for the improvement of measurement techniques, such as the geometric approach to processing of load cell responses \citep{Kesilmis16}, the design of new conveyor machinery \citep{Yamani18}, and electronic truck scales \citep{Guo18}.
 
-In safety studies, a six axis load cell is devised to quantify accelerations and impact forces exerted on a dummy \cite{Ballo16}.
-In alternative energy developments the load cells are useful to measure the forces on the arms of a vertical axis wind turbine \cite{Rossander15}.
-An academic study of the load that a structure withstands is conducted with strain gage load cells that confirms the numerical results and facilitates the design of complex shaped structures \cite{Olmi16}.
-In sports, the performance of new instrumented crank mechanisms is fostered by the utilization of load cells in the characterization, analysis and validation design stages \cite{Casas16}.
+In safety studies, a six axis load cell is devised to quantify accelerations and impact forces exerted on a dummy \citep{Ballo16}.
+In alternative energy developments the load cells are useful to measure the forces on the arms of a vertical axis wind turbine \citep{Rossander15}.
+An academic study of the load that a structure withstands is conducted with strain gage load cells that confirms the numerical results and facilitates the design of complex shaped structures \citep{Olmi16}.
+In sports, the performance of new instrumented crank mechanisms is fostered by the utilization of load cells in the characterization, analysis and validation design stages \citep{Casas16}.
 
-The versatility of the load cells permits the physiological signal monitoring of the the heart and breathing rates \cite{Lee16}, clinical analysis of sleep quality \cite{Zahradka18} and the classification of the movement intensity of people while they are sleeping \cite{Alaziz17}.
+The versatility of the load cells permits the physiological signal monitoring of the the heart and breathing rates \citep{Lee16}, clinical analysis of sleep quality \citep{Zahradka18} and the classification of the movement intensity of people while they are sleeping \citep{Alaziz17}.
 All of these experimental studies have load cells installed on bed setups.
-
-% StArt:Compensation filters
-The use of compensation filters for load cell sensor responses has been reported 
-in \cite{Shu93}, where an adaptive digital filter was first explored,
-in \cite{Jafaripanah05}, where an analog filter alternative was suggested,
-in \cite{Hernandez06} where a recursive LS lattice adaptive filter improves tension forces measurement, 
-in \cite{Boschetti13} where models of the weighing machine, load cells and accelerometers are exploited to remove the environmental vibrations effects, 
-in \cite{Dienstfrey14} where the negative effects of the finite bandwidth of the measurement system response are diminished by regularized deconvolution, and
-in \cite{Huang16}, where the noises that affect an electromagnetic-force-compensated load cell are removed with a set of filters.
-Some guidelines to implement a compensation filter by deconvolution are described in \cite{Eichstadt10}, and
-the synthesis of filters by mapping the sensor model parameters into the filter parameters and time reversed filtering is proposed in \cite{Hessling08a}.
-
-% StArt:Digital signal processing estimation methods
-% The use of a DSP in metrology motivated the data-driven step input estimation method \cite{Markovsky15cep}.
-A short list of the DSP measurement methods for different physical quantities include 
-the impact that the signal processing data-driven dynamic error correction has on the temperature uncertainty \cite{Saggin01},
-the development of an electronic nose using bio-inspired signal processing techniques \cite{Jing16},
-the impedance measurements for material damage estimation using cross-correlation signal processing techniques \cite{deCastro19},
-the modulation quality measurement of microwave access systems is measured using digital signal processing \cite{Angrisani10}, and
-the real-time rotational speed estimation using correlation signal processing techniques \cite{Wang14}.
-
-% StArt:Uncertainty assessment 
-% Monte Carlo sampling
-% Deterministic sampling
-In metrology, the result of a measurement is a random variable.
-The measurement is inherently perturbed by noise and, therefore, the input estimation is represented with the two first statistical moments \cite{Ferrero06}.
-There is a guide to assess the uncertainty of the quantity estimation \cite{GUM08}, but there is still a need to study of the uncertainty of estimation methods and many approaches have been proposed \cite{Esward09, Hessling10}.
-Some of the measurement uncertainty analysis are reviewed in \cite{daSilva12}.
-The Monte Carlo method has been shown to be an effective uncertainty evaluation tool in \cite{Cox06},
-and supports the usefulness of simulation for quantifying measurement uncertainties by software \cite{Esward16}.
-Another example of the Monte Carlo method is the dynamic measurement uncertainty evaluation of clinical thermometers described in \cite{Ogorevc16}.
-
-% 
-It is recommended to consider the uncertainties of all the measurement chain components \cite{Diniz17}, and to avoid the direct uncertainty propagation from the calibration towards the to-be-measured quantity.
-Methods for evaluating the uncertainty associated with the output of compensation filters have been investigated, such as for
-a discrete-time infinite-response filter in \cite{Link09},
-a discrete-time finite-response filter in \cite{Elster07, Elster08}, and
-the Kalman filter in \cite{Eichstadt16b}.
-All these works propagate the uncertanity through the filter but it is also necessary to upward the propagation up to the sensor model to include all systematic error contributions \cite{Hessling11}.
 
@@
 An extension of the methodology proposed for the data-driven step input estimation was formulated to estimate the parameters of an affine input that changes at a constant rate,  by signal processing of the sensor transient response.
-This type of ramp inputs is observed in the measurement of mass during the transportation of products on conveyor systems, ranging from few grams \cite{Burmen09} to almost hundreds of kilograms \cite{Tasaki07}.
+This type of ramp inputs is observed in the measurement of mass during the transportation of products on conveyor systems, ranging from few grams \citep{Burmen09} to almost hundreds of kilograms \citep{Tasaki07}.
 The data-driven affine input estimation method is proposed as an alternative to existing compensation filters, such as 
-the time-variant low-pass filters introduced in \cite{Piskorowski08, Pietrzak14}, and
-the combination of filters in cascade proposed in \cite{Niedzwiecki16a}.
+the time-variant low-pass filters introduced in \citep{Piskorowski08, Pietrzak14}, and
+the combination of filters in cascade proposed in \citep{Niedzwiecki16a}.
+\end{comment}
 
-\subsection{Original contributions}
-% OrigContrib:Affine input estimation
-% OrigContrib:Statistical analysis of structured EIV problems
-% OrigContrib:Experimental validation of the step input estimation method 
-In this thesis are proposed two methods for estimating affine input parameters.
-The first method is an adaptation of the data-driven input estimation method and the second is a maximum likelihood method based on local optimization.
-The adaptation consists of the use of exponential weighing for the recursive least squares solution of the structured errors-in-variables problem.
-The exponential weight gives preference to recent samples over the older samples.
-This forgetting factor considers that the newer samples are more relevant for the parameters estimation.
-The maximum-likelihood method method simulates the response of a sensor model to an affine input. 
-The to-be-minimized cost function is the sum of the squared differences between the actual and the simulated sensor responses. 
-The maximum-likelihood method needs more computational resources, and can simultaneously estimate parameters of the sensor model, but in it current formulation cannot run in real time.
+The signal from the load cell sensor in the weighing system is conditioned using operational amplifiers.
+The signal conditioning amplifier has a low-pass filter to prevent aliasing noise.
+The observed measurement noise has non-white noise properties, mainly due to the characteristics of the load cell sensor.
+One of the assumptions of the step input estimation method formulation is that the measurement noise is Gaussian white. 
+Nevertheless, the estimation results show that the method is still able to provide useful input estimations.
 
-In this document is presented a statistical analysis of the least squares solution of the structured errors-in-variables problem that the data-driven step input estimation method formulates.
-The statistical analysis provides expressions that predict the bias and variance of the step input estimate for given sample size and measurement noise variance.
-The second order Taylor series expansion of the LS solution obtain was derived to find expressions for the bias and variance of the estimate.
-The Crámer-Rao lower bound of the structured EIV problem was obtained and it was compared to the empirical mean-squared error of the estimate.
-It was observed that the input estimation is biased but with small variance, and the difference between the empirical MSE and the theoretical minimum and was determined quantitatively.
+The empirical bias an covariance obtained after repeating the weighing experiment was compared to the bias and covariance estimations obtained in the previous research work \citep{Quintana19}, and to the minimal variance given by the  Cram\'er-Rao lower bound (CRLB) of the structured EIV problem.
+I found that the mean squared error (MSE) of the step input estimate is near the CRLB, and the distance between the MSE and the CRLB provides a confidence interval for the input estimate with respect to the level of the measurement noise.
 
-In this work a series of experiments was conducted to validate the step input estimation method in real-life applications.
-The experiments were realizations of the step input excitation using temperature and mass sensors.
-The weighing setup was constructed to ensure repeatability and reproducibility of the experimental realizations.
-The step responses were stored in sets of 100 elements, and the empirical statistical moment of the input estimates were computed from the sample mean and variance of the estimations.
-The step input estimation method showed robustness when the measurement noise is not Gaussian and white as it was assumed in the theoretical analyses.
-The results of the estimation method with respect to different sensor model order were compared and we found that increasing the order does not necessarily benefits the input estimation uncertainty.
+\subsection{Affine input estimation}
 
-\subsection{Intro of Chapter ramp input}
+A dynamic measurement is present when the fluctuations of the measurand impact on the input estimation, such as when a low-bandwidth sensor is excited with a fast changing input.
+The detection of input characteristics is of interest in several scientific and industrial applications for temperature \citep{Saggin01}, pressure \citep{Matthews14}, acceleration \citep{Link07}, force \citep{Vlajic16}, \citep{Hessling08a}, and mass \citep{Shu93}, \citep{Boschetti13}.
 
-Measurements estimate the unknown value of a physical quantity, namely the measurand.
-The to-be-measured physical quantity is applied as an input signal to a sensor. 
-The sensor is a dynamical system and its output changes as a consequence of the input excitation and the sensor initial conditions.
-The goal of a measurement is to estimate accurately the measurand value using the sensor transient response.
-The transient response of a stable sensor decays to a steady state response.
-In steady state, the most accurate estimation of the input true value is simply found using the sensor static gain.
-However, the steady state is reached in theory after an infinite period of time and in practice we require fast estimations.
-The trade-off between accuracy and speed exists in all measurements.
+I worked on a method to estimate the parameters of inputs that vary at a constant rate, influenced by the data-driven signal processing method that estimates the step level value using subspace techniques \citep{Markovsky15cep}, \citep{Markovsky15ieee}.
+An input varies at a constant rate in applications where the magnitude of interest activates the sensor gradually. 
+An example of this affine activation is the measurement of mass while the to-be-weighted object is transported by a conveyor belt, and the profile of the input is a saturated ramp.
+Current solutions to the weighing in motion are low pass filters that estimate the mass using the saturated part of the input \citep{Tasaki07, Pietrzak14}.
+The signal processing affine input estimation methods are motivated by the need to obtain the mass of the object from the ramp before it reaches saturation.
 
-The measurand can be assumed to be constant or variable during the measurement.
-A dynamic measurement is present when the fluctuations of the measurand impact on the input estimation.
-A typical example of a dynamic measurement problem is a low-bandwidth sensor excited with a fast changing input.
-Some characteristics of the input, like the minimum or maximum or the effects of the environment on the measured quantity, occur in small periods of time.  
-The detection of the input characteristics is needed in several scientific and industrial applications such as measurements of temperature \cite{Saggin01}, pressure \cite{Matthews14}, acceleration \cite{Link07}, force \cite{Vlajic16, Hessling08a} and mass \cite{Shu93, Boschetti13}.
+The proposed data-driven method for the estimation of the affine input parameters.
+The affine input is parameterized as a straight line model where the slope and the interception are the parameters of interest.
+The data-driven affine input estimation method formulates a structured errors-invariables (EIV) problem, similar to the one fourmulated for the step input estimation.
+An exponential weight is added to the recursive least squares.
+This is a forgetting factor that considers that the newer samples are more relevant for the input parameters estimation.
+The data-driven affine input estimation method is a recursive algorithm that can be implemented in real-time since it has low computational cost.
 
-The solution to dynamic measurement problems is non-trivial.
-An approach is to add a dynamical system to compensate the sensor transient response, inverting the effects of the sensor dynamics.
-The purpose of such a compensator is to reduce the transient time.
-The sensor dynamics are considered in the design of finite and infinite impulse response compensation digital filters based on deconvolution \cite{Eichstadt10} or synthesized to correct dynamic errors \cite{Hessling08a}.
-The model-based deconvolution design of compensators implies that the measurand true value should be known a-priori for certain applications, such as mass determinations \cite{Boschetti13, Niedzwiecki16b}.
-In the literature most of the measurements systems are assumed linear time-invariant, but the compensation digital filters can be linear \cite{Tasaki07},  nonlinear \cite{Shu93} or time-varying \cite{Pietrzak14}.
+The performance of the proposed method is compared to that of a maximum-likelihood (ML) estimation method based on local-optimization and a time-varying compensation filter.
+The ML method simulates the response of a sensor model to an affine input, and minimizes a cost function that is the sum of the squared differences between the actual and the simulated sensor responses.
+The ML method resembles the model predictive control approach in the sense that a cost function is minimized iteratively to optimize the parameters of a sensor model using the observed sensor response in a receding time horizon \citep{Mayne14}.
+The difference is that the ML method aims to estimate the unknown value of the affine input parameters instead of identifying a model and controlling the dynamic system.
 
-The digital signal processors enable a different approach where the input estimation can be obtained with algorithms that do not necessarily recreate the dynamics of a system.
-One of the authors of this paper proposed a data-driven signal processing method that estimates the measurand true value using subspace techniques \cite{Markovsky15cep, Markovsky15ieee}.
-The subspace estimation method bypasses the model identification step to estimate the unknown input directly from the response data.
-This method was developed to estimate inputs modeled as step functions of unknown scaling level.
-We extended the subspace input estimation method to estimate the parameters of inputs that vary at a constant rate.
-
-The inputs that vary at a constant rate are found in applications where the measurand activates the sensor gradually. 
-An example of this activation is the measurement of mass while the to-be-weighted object is transported by a conveyor belt, and the profile of the input is a saturated ramp.
-Current solutions to the weighing in motion are low pass filters that estimate the mass using the saturated ramp \cite{Tasaki07, Pietrzak14}.
-The signal processing affine input estimation methods are motivated by the need to obtain the mass of the object from the ramp before it reaches saturation.
-The ramp is parameterized as a straight line model where the slope and the interception are the parameters of interest.
-
-This paper describes a subspace method for the estimation of the affine input parameters.
-This method is a recursive algorithm that can be implemented in real-time since it has low computational cost.
-The subspace method is independent of the sensor model and, therefore, it is suitable for a variety of applications.
-The dynamic weighing is one of the applications and was chosen as an implementation example.
-The effectiveness of the method is evaluated in a simulation study.
-The performance of the proposed method is compared to that of a maximum-likelihood (ML) estimation method based on local-optimization and a time-varying compensation filter.
-%We found that the subspace method obtains the slope estimate from the sensor transient response using the same assumptions of the data-driven step input estimation method.
-
-The ML method resembles the model predictive control approach in the sense that a cost function is minimized iteratively to optimize the parameters of a sensor model using the observed sensor response in a receding time horizon \cite{Mayne14}.
-The difference is that the ML method aims to estimate the unknown value of the affine input parameters instead of identifying a model and controlling the dynamic system.
-The ML method is more appropriate for off-line processing of the sensor transient response.
-Nevertheless, the ML method can estimate the parameters of the affine input, the parameters of a sensor model, and the initial conditions of the sensor.
-
-The uncertainty of the subspace method is assessed using a Taylor expansion of the estimate and Monte Carlo random sampling approach \cite{Quintana19}.
-The Monte Carlo approach requires a large set of generated random samples, and for simple systems it is the recommended method.
-There exists a deterministic sampling approach to study the uncertainty propagation of complex systems \cite{Hessling13a, Hessling13b}. 
-Deterministic sampling aims to represent the minimal statistical information that is relevant to the uncertainty estimation in a finite set.
-The uncertainty of the ML method is assessed using the derivatives of the residual error that constructs the to-be-minimized cost function.
-The covariance of the optimization method estimate is found using the inverse of the Hessian matrix \cite{Pintelon12Book}.
-
-\subsection{Intro of Chapter statistical analysis}
-
-Errors-in-variables (EIV) are linear estimation problems in which the regression matrix and the regressor are perturbed \cite{VanHuffel91Book}, \cite{Markovsky07overview}.
-In structured EIV problems, the regression matrix has a given structure that depends on the problem formulation.
-Hankel, Toeplitz, or other application-specific matrices appear in problems of metrology \cite{Markovsky15cep}, system identification \cite{Soderstrom07}, image restoration \cite{Feiz17}, nuclear magnetic resonance spectroscopy \cite{Cai16}, direction-of-arrival estimation \cite{Pan18}, and time-of-arrival estimation \cite{Jia18}.
-
-In metrology, the direct estimation of the input from the sensor transient response is formulated as a structured EIV problem.
-The only observed signal is the sensor output.
-To estimate a step input, the regression matrix, and the regressor are built from the step response observations \cite{Markovsky15cep}, and
-the structure in the regression matrix is block-Hankel.
-This data-driven estimation methodology reduces the estimation time of the classical two-stage approach where a sensor model is first identified, and later the input is estimated using the sensor model \cite{Azam15, Niedzwiecki16a}.
-%The problem is EIV because the perturbation of the sensor response gets into the regression matrix.
-
-Total least-squares (TLS) is the typical estimator for unstructured EIV problems and is consistent when the perturbations have zero mean, and the covariance is a given positive definite matrix.
-When the perturbations are i.i.d. normally distributed, the solution of the TLS is equivalent to that of the maximum likelihood estimator (ML) \cite{Markovsky07overview}. 
-For structured EIV problems, the TLS estimator does not give general results since each specific structure requires a particular treatment \cite{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where finding the global optimum is not guaranteed \cite{Rhode14recursive}.
-Moreover, the computational complexity of TLS and ML inhibits real-time implementation to solve structured EIV problems.
-
-The least-squares (LS) estimator is a suboptimal but simple solution to structured EIV problems that admits a recursive form for easy real-time implementations.
-Some of the reported works that propose LS estimators for structured EIV problems include 
-the design of a fast algorithm for matrices with small displacement rank \cite{Mastronardi07fast},
-the study of the estimator consistency \cite{Palanthandalam10parameter},
-the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models \cite{Kiviet12high} \cite{Kiviet14improved}, and
-a discussion of the causes of bias and inconsistency in homogeneous estimators \cite{Yeredor04homogeneous}.
-
- In measurement applications, it is highly relevant to assess the uncertainty of the input estimate.
-The uncertainty of the reported LS estimators for structured EIV problems has not been addressed, and then, remains unknown.
-The estimator uncertainty is expressed using the estimation bias and covariance \cite{Pintelon12Book}.
-To know the LS estimator uncertainty, we quantified the bias and the covariance of the LS solution of EIV problems in the unstructured and structured cases. 
-This extends the perturbation analysis of the LS estimator of unstructured and uncorrelated problems that was investigated in \cite{Stewart90SPT} and in \cite{Vaccaro94}.
-This paper presents a study of the statistics of the LS estimate of unstructured and structured EIV problems. 
- We provide a discussion of the unstructured case as a reference, to get an insight of the impact that the structure of the regression matrix and the correlation between the regression matrix and the regressor perturbations have on the uncertainty of the LS estimate.
-The structured case is motivated by the metrology estimation problem \cite{Markovsky15cep}, where the EIV problem has a Hankel structure, and the perturbations are correlated.
-The study of the metrology input estimate illustrates a methodology to conduct statistical analysis for any structured EIV problem.
-The mathematical expectation of the second-order Taylor series expansion of the LS estimate boils down to expressions that quantify the first and second-order moments of the LS estimate.
-Via Monte Carlo simulations, we validated the accuracy of the bias, and the covariance approximations.
-The derived approximations predict the LS estimate bias, and the covariance, for given sample size and perturbation level.
-The predicted variance gives the uncertainty of the LS estimate.
-We observed that, for the step input estimation problem, the mean squared error of the LS estimate is near to the minimum variance limit given by the Cram\'er-Rao lower bound of the structured EIV problem.
- By following this methodology, the bias and variance of the solutions of EIV problems with other structures is determined, and therefore, the uncertainty of the estimate.
-
-\subsection{Intro of Chapter experimental validation}
-
- In this paper we consider that a measurement is a dynamic process, where an input excites a dynamic system, the sensor, and causes a dynamic transient response that also depends on the initial conditions of the sensor.
-The to-be-measured quantity is an unknown input that excites the sensor.
-The consequent transient response is further processed to estimate quickly the measurand value.
-The steady-state response of the sensor, that exists after the stabilization of dynamic effects, gives easy access to the measurand value but this approach is mainly exploited for calibration purposes.
-
-A compensator is an additional dynamic system that acts on the transient response aiming to reduce the sensor transient time.
-The compensation is motivated by the need of inverting the sensor dynamic effects to recreate the input.
-The convolution of the compensator impulse response with the sensor transient response yields the input estimate.
-Therefore, the design of a compensator is based on the sensor model and requires a deconvolution \cite{Eichstadt10}.
-Examples of input estimation using compensation of the sensor transient response include a recursive estimation of the compensator parameters \cite{Shu93}, 
-finite impulse response (FIR) \cite{Elster07, Niedzwiecki16b} filters and 
-infinite impulse response (IIR) filters \cite{Pintelon90, Elster08}.
-The filters in these works estimate in real-time the unknown input value.
-
-An alternative to the compensation approach is to use digital signal processing methods that are independent of the sensor model.
-A data-driven method that estimates the unknown level of step inputs by processing the sensor step response was introduced in \cite{Markovsky15ieee}. 
-This data-driven input estimation method avoids the sensor modeling stage and estimates directly the input.
-This method reduces the estimation time compared to a conventional compensator.
-The step input estimation method performance was demonstrated by simulations and experiments on a digital signal processor (DSP) of low cost \cite{Markovsky15cep}.
-The uncertainty of the step input estimation method has not been assessed before.
-
-To validate the input estimation methods it is necessary to assess the uncertainty associated with their estimates \cite{daSilva12, Ferrero06}.
-There are uncertainty propagation studies for model-based compensators such as the FIR and IIR filters for acceleration measurements where the uncertainty is computed in real time \cite{Elster07, Elster08, Link09}.
-In these works, the uncertainty expression is based on the transfer function or state space representations of the LTI sensor and filter systems.
-Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \cite{Pietrzak14} for mass and in \cite{Ogorevc16} for temperature sensors.
-A deconvolution method is implemented to estimate the input waveform in \cite{Hale09} and the uncertainty is obtained from the input estimate covariance. 
-The impact that the signal processing data-driven dynamic error correction has on the uncertainty is investigated in \cite{Saggin01}. 
-A statistical analysis of the data-driven step input estimation method \cite{Markovsky15cep} was investigated in \cite{Quintana19} and the method uncertainty was obtained with a Monte Carlo simulation study.
-
-This paper provides an uncertainty assessment of the data-driven step input estimation method in a real-life application.
-The measurements were conducted in a weighing system based on a load cell sensor.
-We observed that even when the whiteness assumptions of the measurement noise are not fulfilled, the step input estimation method still is able to provide a good estimation. 
-We found that the mean squared error of the input estimate is near the Cram\'er-Rao lower bound of the EIV problem.
-A confidence interval is provided for the input estimate in terms of the number of samples required to satisfy the accuracy specifications of the user.
-
-The novelty of the paper is threefold.
-First, using the results of \cite{Quintana19}, that describes a statistical analysis of structured errors in variables (EIV) problems, in this paper we describe the statistical properties of a the data-driven step input estimation method in both simulation and a real-life experiments.
-Second, this manuscript also presents the Cram\'er-Rao lower bound for unbiased estimators of the structured and correlated EIV problem that the step input estimation method formulates.
-Using this bound we have the minimum mean-squared error (MSE) for this estimation problem that we use compare with the MSE computed from the predictions obtained after the statistical analysis.
-The third novelty in this manuscript is the model order selection for the step input estimation method where we use the MSE to select the order that provides the smaller MSE with the lowest computational complexity.
+After observing the simulation results, the data-driven affine input estimation method is suitable for real-time applications since it requires low computational resources. 
+The ML method is more appropriate for off-line processing of the sensor transient response, but
+it can estimate also the parameters of a sensor model, and the initial conditions of the sensor.
+The main drawback of the ML method is the need of high computational resources.
 

### 2020-01-27 ChapterPreliminaries/ChapterPreliminaries.tex
Story 182: Intro complete

@@
 
-\section{preliminaries}
+\section{Preliminaries}
 
@@
 where $\mathbf{A} \in \mathbb{R}^{n \times n}$, $\mathbf{B} \in \mathbb{R}^{n \times m}$, $\mathbf{C} \in \mathbb{R}^{p \times n}$, and $\mathbf{D} \in \mathbb{R}^{p \times m}$, are the model matrices, $\nu$ is the measurement noise, $n$ is the system order, $m$ is the number of inputs, and $p$ is the number of outputs.
-Although we may think that most sensors are single-input single output (SISO) like temperature sensors, there are sensors with single input and multiple outputs such as gas sensors \cite{Munther19}, and sensors with multiple inputs and multiple outputs like the three-axis accelerometers \cite{DEmilia16}, the radio-frequency intruder-detection sensors \cite{Ushiki13}, and the radar sensor \cite{Kueppers17}.
+Although we may think that most sensors are single-input single output (SISO) like temperature sensors, there are sensors with single input and multiple outputs such as gas sensors \citep{Munther19}, and sensors with multiple inputs and multiple outputs like the three-axis accelerometers \citep{DEmilia16}, the radio-frequency intruder-detection sensors \citep{Ushiki13}, and the radar sensor \citep{Kueppers17}.
 
@@
 With a step input $\mathbf{u} = \bar{\mathbf{u}} s$, where $s(t) = 1$, for $t \geq 0$, and $s(t) = 0$, for $t < 0$, the discrete-time state space representation of the LTI sensor is equivalent to an augmented autonomous system
-\begin{equation} \begin{aligned} \mathbf{x}_\text{a}(k+1) &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_\text{a}(k) , \quad \mathbf{y}(k) = \underbrace{ \begin{bmatrix} \mathbf{C} & \mathbf{D} \end{bmatrix} }_{\mathbf{C}_\text{a}} \mathbf{x}_\text{a}(k), \quad \text{where} \ \ \mathbf{x}_\text{a}(k) = \begin{bmatrix} \mathbf{x}(k) \\ \mathbf{u}(k) \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}(0) .
-\label{eqn:augmented} \end{aligned} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{x}_\text{a}(k+1) &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_\text{a}(k) , \quad \text{where} \ \ \mathbf{x}_\text{a}(k) = \begin{bmatrix} \mathbf{x}(k) \\ \mathbf{u}(k) \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\
+\mathbf{y}(k) &= \underbrace{ \begin{bmatrix} \mathbf{C} & \mathbf{D} \end{bmatrix} }_{\mathbf{C}_\text{a}} \mathbf{x}_\text{a}(k) . \label{eqn:augmented} \end{aligned} \end{equation}
 The eigenvalues $\lambda$ of the augmented autonomous system are found using
@@
 Thus, we can express 
-\[ \mathbfcal{H}(\mathbf{y}) = \begin{bmatrix} y(1) & y(2) & \cdots & y(n) \\ y(2) & y(3) & \cdots & y(n+1) \\ \iddots & \iddots & \iddots \\ y(n) & y(n+1) & \cdots & y(2n-1) \end{bmatrix} = \begin{bmatrix} \widehat{\mathbf{C}}_\text{a} \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a} \\ \vdots \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a}^{n} \end{bmatrix}  \begin{bmatrix} \mathbf{x}_\text{a}(0) & \mathbf{x}_\text{a}(1) & \cdots & \mathbf{x}_\text{a}(n) \end{bmatrix} = \mathbfcal{O}_\text{a} \mathbf{X}_\text{ini}, \]
+\[ \underbrace{ \begin{bmatrix} y(1) & y(2) & \cdots & y(n) \\ y(2) & y(3) & \cdots & y(n+1) \\ \iddots & \iddots & \iddots \\ y(n) & y(n+1) & \cdots & y(2n-1) \end{bmatrix} }_{ \mathbfcal{H}(\mathbf{y}) }= \underbrace{ \begin{bmatrix} \widehat{\mathbf{C}}_\text{a} \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a} \\ \vdots \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a}^{n} \end{bmatrix} }_{ \mathbfcal{O}_\text{a} }  \underbrace{ \begin{bmatrix} \mathbf{x}_\text{a}(0) & \mathbf{x}_\text{a}(1) & \cdots & \mathbf{x}_\text{a}(n) \end{bmatrix} }_{ \mathbf{X}_\text{ini} } , \]
 where $\mathbf{X}_\text{ini}$ is a matrix with the initial conditions of the $n$ free responses. 
@@
  This EIV problem admits a least-squares (LS) solution and the recursive LS enables the online implementation for real-time step input estimation applications.
- The step input estimation method was described in \cite{Markovsky15cep}.
+ The step input estimation method was described in \citep{Markovsky15cep}.
 
@@
 
+\begin{comment}
 \section{Step input estimation method - Ramp input}
@@
 where $\widebar{\mathbf{K}}$ is exact data information and $\mathbf{E}$ is the corresponding perturbation noise.
-Details of the method formulation are described in \cite{Markovsky15cep}.
+Details of the method formulation are described in \citep{Markovsky15cep}.
 
@@
 \begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} g & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ g & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ g & \Delta \widetilde{y}(T-n) & \Delta \widetilde{y}(T-n+1) & \cdots & \Delta \widetilde{y}(T-1) \end{bmatrix} , \label{eqn:matrix} \end{equation}
-is a Hankel matrix of $(T - n)-\text{block rows}$, constructed from consecutive differences $\Delta \widetilde{y}(t) = \widetilde{y}(t) - \widetilde{y}(t-1)$ of the observed transient response, augmented in the left side with a $(T - n)\text{-vector}$ of elements equal to the known dc-gain $g$, (see \cite{Markovsky15cep}).
+is a Hankel matrix of $(T - n)-\text{block rows}$, constructed from consecutive differences $\Delta \widetilde{y}(t) = \widetilde{y}(t) - \widetilde{y}(t-1)$ of the observed transient response, augmented in the left side with a $(T - n)\text{-vector}$ of elements equal to the known dc-gain $g$, (see \citep{Markovsky15cep}).
 The perturbations $\bm{\epsilon}$ enter in matrix $\widetilde{\mathbf{K}}$ and we can express 
@@
 The recursive LS method allows for a real-time implementation of the step input estimation method.
-Details of the step input estimation method are described in \cite{Markovsky15cep}.
+Details of the step input estimation method are described in \citep{Markovsky15cep}.
 
@@
 
-The estimation of the step input level is obtained as the solution of the minimization problem \cite{Markovsky15ieee, Markovsky15cep} 
+The estimation of the step input level is obtained as the solution of the minimization problem \citep{Markovsky15ieee, Markovsky15cep} 
 \begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert  \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x} \right\Vert^2_2 \label{eqn:min_ls} \end{equation}
@@
 
+\end{comment}
+
 \newpage

### 2020-01-27 ChapterRampInput/ChapterRampInput.tex
Story 182: Intro complete

@@
 
-\section{Introduction  - Ramp input}
+\section{Introduction}
 
+The goal of a measurement is to accurately estimate the input value using the sensor transient response.
+The input might vary during the measurement being the simplest case the input that changes with a constant rate, described with an affine function.
+The inputs that vary at a constant rate activate the sensor gradually. 
+An affine input estimation method is needed to estimate the slope and the interception parameters of the affine function.
+
+This chapter describes two signal processing methods for the estimation of affine input parameters.
+These methods are a data-driven subspace method and a maximum-likelihood (ML) estimation method based on local-optimization.
+The subspace is a recursive method independent of the sensor model, and the ML method iteratively minimizes a cost function using a simulated sensor response in a receding time horizon.
+
+In this chapter, the dynamic weighing is chosen as an implementation example.
+The dynamic weighing of an affine input is an interesting metrology application because the affine input turns a linear time invariant sensor model into a time-varying sensor model.
+The affine input estimation methods provide an insight on the properties of the dynamic measurement of time-varying inputs.
+
+The effectiveness and the uncertainty of the methods are evaluated with simulation studies.
+The uncertainty is assessed, for the subspace method, using a Taylor expansion of the exponentially weighted recursive least squares estimate, and
+for the ML method, using the derivatives of the residual error obtained from the to-be-minimized cost function.
+
+\begin{comment}
 Measurements estimate the unknown value of a physical quantity, namely the measurand.
@@
 The covariance of the optimization method estimate is found using the inverse of the Hessian matrix \cite{Pintelon12Book}.
+
+\end{comment}
 
@@
 \centering
-\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_2.pdf} 
+\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_2.pdf} 
 \caption{ \label{fig:sensor_weight} The sensor transient response $y$ to an affine input excitation $u = at+b$ is processed by the estimation methods to estimate the parameters $a$ and $b$.
@@
 These values were obtained after calibration of the method and fixed during the simulation study.}
-\begin{tabular}{|c c c c c c|} 
+\begin{tabular}{c| c c c c c} 
  \hline
@@
 \centering
-\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_3.pdf} 
+\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_3.pdf} 
 \caption{ \label{fig:rele_dd_40dB_s1} The relative errors of the affine input parameters estimation decrease as the subspace method processes more samples.
@@
 \centering
-\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_4.pdf} 
+\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_4.pdf} 
 \caption{ \label{fig:rele_SNR_dd_10000} The minimum value of the estimation relative errors obtained with the subspace method is less than 2  \% regardless of the SNR between 20 dB and 60 dB. }
@@
 \centering
-\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_5.pdf} 
+\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_5.pdf} 
 \caption{ \label{fig:CRLB_MSE_ab_dd_40dB_MC_10000} When the SNR of the sensor response is 40 dB, the mean squared errors of the slope estimate $\widehat{a}$ and the interception estimate $\widehat{b}$, obtained by the subspace method, are two orders of magnitude above the theoretical minimum variance given by the Cram\'er-Rao lower bound.}
@@
 \centering
-\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_6.pdf} 
+\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_6.pdf} 
 \caption{ \label{fig:CRLB_MSE_SNR_ab_dd_MC_10000} The Cram\'er-Rao lower bounds of the estimates $\mathrm{CRLB}_a$ and $\mathrm{CRLB}_b$ determine the minimum uncertainty that can be achieved and increases with the measurement noise.
@@
 
-\begin{tabular}{|L|L L L L L|} 
+\begin{tabular}{L|L L L L} 
  \hline
- \mathrm{Time \ [s]} & \mathrm{Mass \ [kg]} & 0.1 & 0.3 & 0.5 & 1.0 \\ [0.5ex] 
+ & \mathrm{Mass} & \\ [0.5ex] 
+ \mathrm{Time} & 0.1 \ kg & 0.3 \ kg & 0.5 \ kg & 1.0 \ kg \\ 
  \hline
- 0.05 & \hspace{7.5mm} \mathrm{MSE}_{\hat{a}}: & 3.55\mathrm{x}10^{0} (@40 \ \mathrm{dB}) & 13.5\mathrm{x}10^{0} (@50 \ \mathrm{dB}) & 3.80\mathrm{x}10^{0} (@20 \ \mathrm{dB}) & 8.29\mathrm{x}10^{0}  (@50 \ \mathrm{dB}) \\
- & \hspace{7.5mm} \mathrm{MSE}_{\hat{b}}: & 3.10\mathrm{x}10^{-3} (@40 \ \mathrm{dB})  & 1.11\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) & 1.07\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) & 1.92\mathrm{x}10^{-2} (@60  \ \mathrm{dB})\\
- 0.1 & \hspace{7.5mm} \mathrm{MSE}_{\hat{a}}: & 3.08\mathrm{x}10^{-2} (@30 \ \mathrm{dB}) & 1.16\mathrm{x}10^{0} (@40  \ \mathrm{dB}) & 8.34\mathrm{x}10^{-1} (@60 \ \mathrm{dB}) & 1.27\mathrm{x}10^{0} (@50  \ \mathrm{dB}) \\
- & \hspace{7.5mm} \mathrm{MSE}_{\hat{b}}: & 5.99\mathrm{x}10^{-5} (@30 \ \mathrm{dB}) & 1.17\mathrm{x}10^{-1} (@50  \ \mathrm{dB}) & 2.10\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) & 3.76\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) \\
- 0.5 & \hspace{7.5mm} \mathrm{MSE}_{\hat{a}}: & 3.03\mathrm{x}10^{-2} (@20 \ \mathrm{dB}) & 2.98\mathrm{x}10^{-1} (@20  \ \mathrm{dB}) & 3.25\mathrm{x}10^{0} (@20 \ \mathrm{dB}) & 8.60\mathrm{x}10^{-2} (@20  \ \mathrm{dB}) \\
- & \hspace{7.5mm} \mathrm{MSE}_{\hat{b}}: & 3.10\mathrm{x}10^{-5} (@50 \ \mathrm{dB}) & 9.86\mathrm{x}10^{-4} (@40 \ \mathrm{dB}) & 2.06\mathrm{x}10^{-5} (@50 \ \mathrm{dB}) & 1.74\mathrm{x}10^{-5} (@20 \ \mathrm{dB}) \\ [0.5ex] % [\mathrm{samples}]
+\multicolumn{1}{l|}{0.05 \ s} & \\
+ \multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{a}} \ [g^2]} & 3.5\mathrm{x}10^{0}_{(@40 \ \mathrm{dB})} & 13.5\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})} & 3.8\mathrm{x}10^{0}_{(@20 \ \mathrm{dB})} & 8.3\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})} \\
+\multicolumn{1}{r|}{\hspace{1mm}\mathrm{MSE}_{\hat{b}} \ [g^2]} &  3.1\mathrm{x}10^{-3}_{(@40 \ \mathrm{dB})}  & 1.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} & 1.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} & 1.9\mathrm{x}10^{-2}_{(@60 \ \mathrm{dB})} \\
+\multicolumn{1}{l|}{0.1 \ s} & \\
+ \multicolumn{1}{r|}{\hspace{1mm}\mathrm{MSE}_{\hat{a}} \ [g^2]} & 3.1\mathrm{x}10^{-2}_{(@30 \ \mathrm{dB})} & 1.1\mathrm{x}10^{0}_{(@40 \ \mathrm{dB})} & 8.3\mathrm{x}10^{-1}_{(@60 \ \mathrm{dB})} & 1.3\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})} \\
+\multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{b}} \ [g^2]} & 6.0\mathrm{x}10^{-5}_{(@30 \ \mathrm{dB})} & 1.2\mathrm{x}10^{-1}_{(@50 \ \mathrm{dB})} & 2.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} & 3.7\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} \\
+\multicolumn{1}{l|}{0.5 \ s} & \\
+ \multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{a}} \ [g^2]} & 3.0\mathrm{x}10^{-2}_{(@20 \ \mathrm{dB})} & 3.0\mathrm{x}10^{-1}_{(@20 \ \mathrm{dB})} & 3.2\mathrm{x}10^{0}_{(@20 \ \mathrm{dB})} & 8.6\mathrm{x}10^{-2}_{(@20 \ \mathrm{dB})} \\
+\multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{b}} \ [g^2]} & 3.1\mathrm{x}10^{-5}_{(@50 \ \mathrm{dB})} & 9.8\mathrm{x}10^{-4}_{(@40 \ \mathrm{dB})} & 2.0\mathrm{x}10^{-5}_{(@50 \ \mathrm{dB})} & 1.7\mathrm{x}10^{-5}_{(@20 \ \mathrm{dB})} \\ [0.5ex] % [\mathrm{samples}]
  \hline
@@
 
-\begin{tabular}{|C| C C C C|} 
+\begin{tabular}{C| C C C} 
 \hline  
- &  & \sigma_{\mathrm{s}}: \hspace{4mm} 0\% & \hspace{8.1mm} 5\% & \hspace{8.1mm} 10\% \\ [0.5ex] 
+& \sigma_{\mathrm{s}} \\ [0.5ex] 
+\sigma_{{a},{b}}  & 0\% & 5\% & 10\% \\  
  \hline
- \sigma_{{a},{b}}: \hspace{1.6mm} 0\% & \widehat{a} \ [\mathrm{kg/s}] & 0.990 \pm 9.6 \% \ (@50 \ \mathrm{dB}) &  0.999 \pm 8.2 \% \ (@20 \ \mathrm{dB}) & 0.996 \pm 8.2 \% \ (@50 \ \mathrm{dB}) \\ 
-  & \widehat{b} \ \mathrm{[g]} & \hspace{-1.75mm} 10.0 \pm 10.3 \% \ (@20 \ \mathrm{dB}) & 9.91 \pm 14.2 \% \ (@30 \ \mathrm{dB}) &  9.99 \pm 21.7 \% \ (@60 \ \mathrm{dB}) \\ 
- \hspace{6.65mm} 5\% & \widehat{a} \ [\mathrm{kg/s}] & 0.994 \pm 9.7 \% \ (@50 \ \mathrm{dB}) & 0.993 \pm 11.3 \% \ (@20 \ \mathrm{dB}) & 0.997 \pm 8.0 \% \ (@50 \ \mathrm{dB}) \\  
-  & \widehat{b} \ \mathrm{[g]} & 9.98 \pm 22.4 \% \ (@50 \ \mathrm{dB}) & \hspace{-1.75mm} 10.02 \pm 33.2 \% \ (@30 \ \mathrm{dB}) & 9.94 \pm 16.6 \% \ (@50 \ \mathrm{dB}) \\ 
-\hspace{4.88mm} 10\% & \widehat{a} \ [\mathrm{kg/s}] & 1.00 \pm 20.3 \% \ (@40 \ \mathrm{dB}) & 1.00 \pm 22.0 \% \ (@50 \ \mathrm{dB}) & 0.996 \pm 17.7 \% \ (@40 \ \mathrm{dB}) \\    
- & \widehat{b} \ \mathrm{[g]} & 9.97 \pm 46.8 \% \ (@40 \ \mathrm{dB}) & 9.78 \pm 58.8 \% \ (@50 \ \mathrm{dB}) & 9.91 \pm 41.6 \% \ (@40 \ \mathrm{dB}) \\ [0.5ex] 
+ \multicolumn{1}{l|}{0\%} \\
+ \multicolumn{1}{l|}{\hspace{2mm} \widehat{a} \ [\mathrm{kg/s}]} & 1.0 \pm 9.6 \%_{(@50 \ \mathrm{dB})} &  1.0 \pm 8.2 \%_{(@20 \ \mathrm{dB})} & 1.0 \pm 8.2 \%_{(@50 \ \mathrm{dB})} \\ 
+ \multicolumn{1}{l|}{\hspace{2mm} \widehat{b} \ \mathrm{[g]}} & 10.0 \pm 10.3 \%_{(@20 \ \mathrm{dB})} & 9.9 \pm 14.2 \%_{(@30 \ \mathrm{dB})} &  10.0 \pm 21.7 \%_{(@60 \ \mathrm{dB})} \\ 
+ \multicolumn{1}{l|}{5\%} \\
+ \multicolumn{1}{l|}{\hspace{2mm} \widehat{a} \ [\mathrm{kg/s}]} & 1.0 \pm 9.7 \%_{(@50 \ \mathrm{dB})} & 1.0 \pm 11.3 \%_{(@20 \ \mathrm{dB})} & 1.0 \pm 8.0 \%_{(@50 \ \mathrm{dB})} \\  
+ \multicolumn{1}{l|}{\hspace{2mm} \widehat{b} \ \mathrm{[g]}} & 10.0 \pm 22.4 \%_{(@50 \ \mathrm{dB})} & 10.0 \pm 33.2 \%_{(@30 \ \mathrm{dB})} & 9.9 \pm 16.6 \%_{(@50 \ \mathrm{dB})} \\ 
+\multicolumn{1}{l|}{10\%} \\
+\multicolumn{1}{l|}{\hspace{2mm} \widehat{a} \ [\mathrm{kg/s}]} & 1.0 \pm 20.3 \%_{(@40 \ \mathrm{dB})} & 1.0 \pm 22.0 \%_{(@50 \ \mathrm{dB})} & 1.0 \pm 17.7 \%_{(@40 \ \mathrm{dB})} \\    
+ \multicolumn{1}{l|}{\hspace{2mm} \widehat{b} \ \mathrm{[g]}} & 9.9 \pm 46.8 \%_{(@40 \ \mathrm{dB})} & 9.8 \pm 58.8 \%_{(@50 \ \mathrm{dB})} & 9.9 \pm 41.6 \%_{(@40 \ \mathrm{dB})} \\ [0.5ex] 
 \hline
@@
 \centering
-\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_7.pdf} 
+\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_7.pdf} 
 \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor initial conditions are estimated with the ML method.
@@
 \centering
-\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_8.pdf} 
+\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_8.pdf} 
 \caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian.
@@
 \centering
-\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_9.pdf} 
+\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_9.pdf} 
 \caption{ \label{fig:rele_tv_40dB_s1} The relative errors of the time-varying filter estimation converge slower than with the subspace method. 

### 2020-01-27 ChapterStatisticalAnalysis/ChapterStatisticalAnalysis.tex
Story 182: Intro complete

@@
 \chapter{Statistical Analysis}\label{chap:StatisticalAnalysis}
+
 \begin{quote}
@@
 
-\section{Introduction  - Statistical analysis}
+\section{Introduction}
 
-Errors-in-variables (EIV) are linear estimation problems in which the regression matrix and the regressor are perturbed \cite{VanHuffel91Book}, \cite{Markovsky07overview}.
+The linear estimation problems with perturbation in both the regression matrix and the regressor are errors-in-variables (EIV) problems.
+To evaluate the properties of the solution to an EIV problem it is often required to calculate the statistical moments of the estimate.
+The calculation of the estimate statistical moments demands a larger effort when the regression matrix of a EIV problem is structured, and when is correlated to the regressor.
+
+Structured total least squares and structured total maximum likelihood are two examples of methods that solve structured EIV problems, but that require large computational resources.
+For online applications where a real-time solution of the structured EIV problem is desired, the recursive least-squares (RLS) method is a suboptimal but computationally cheaper estimator.
+
+To obtain the statistical moments of the RLS estimation for a structured and correlated EIV problem, and to assess the estimate uncertainty, that have not been obtained before, a statistical analysis was conducted.
+
+This chapter describes the methodology and the results of the statistical analysis.
+The mathematical expectation of the RLS estimate, approximated by a second-order Taylor series expansion, allows a quantification of the first and second-order moments of the LS estimate.
+The results of the statistical analysis provide an insight of the impact that the structure of the regression matrix and the correlation between the regression matrix and the regressor perturbations have on the uncertainty of the RLS estimate.
+
+\begin{comment}
+Errors-in-variables (EIV) are linear estimation problems in which the regression matrix and the regressor are perturbed \citep{VanHuffel91Book}, \citep{Markovsky07overview}.
 In structured EIV problems, the regression matrix has a given structure that depends on the problem formulation.
-Hankel, Toeplitz, or other application-specific matrices appear in problems of metrology \cite{Markovsky15cep}, system identification \cite{Soderstrom07}, image restoration \cite{Feiz17}, nuclear magnetic resonance spectroscopy \cite{Cai16}, direction-of-arrival estimation \cite{Pan18}, and time-of-arrival estimation \cite{Jia18}.
+Hankel, Toeplitz, or other application-specific matrices appear in problems of metrology \citep{Markovsky15cep}, system identification \citep{Soderstrom07}, image restoration \citep{Feiz17}, nuclear magnetic resonance spectroscopy \citep{Cai16}, direction-of-arrival estimation \citep{Pan18}, and time-of-arrival estimation \citep{Jia18}.
 
@@
 The only observed signal is the sensor output.
-To estimate a step input, the regression matrix, and the regressor are built from the step response observations \cite{Markovsky15cep}, and
+To estimate a step input, the regression matrix, and the regressor are built from the step response observations \citep{Markovsky15cep}, and
 the structure in the regression matrix is block-Hankel.
-This data-driven estimation methodology reduces the estimation time of the classical two-stage approach where a sensor model is first identified, and later the input is estimated using the sensor model \cite{Azam15, Niedzwiecki16a}.
+This data-driven estimation methodology reduces the estimation time of the classical two-stage approach where a sensor model is first identified, and later the input is estimated using the sensor model \citep{Azam15, Niedzwiecki16a}.
 %The problem is EIV because the perturbation of the sensor response gets into the regression matrix.
@@
 Total least-squares (TLS) is the typical estimator for unstructured EIV problems and is consistent when the perturbations have zero mean, and the covariance is a given positive definite matrix.
-When the perturbations are i.i.d. normally distributed, the solution of the TLS is equivalent to that of the maximum likelihood estimator (ML) \cite{Markovsky07overview}. 
-For structured EIV problems, the TLS estimator does not give general results since each specific structure requires a particular treatment \cite{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where finding the global optimum is not guaranteed \cite{Rhode14recursive}.
+When the perturbations are i.i.d. normally distributed, the solution of the TLS is equivalent to that of the maximum likelihood estimator (ML) \citep{Markovsky07overview}. 
+For structured EIV problems, the TLS estimator does not give general results since each specific structure requires a particular treatment \citep{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where finding the global optimum is not guaranteed \citep{Rhode14recursive}.
 Moreover, the computational complexity of TLS and ML inhibits real-time implementation to solve structured EIV problems.
@@
 Some of the reported works that propose LS estimators for structured EIV problems include 
-the design of a fast algorithm for matrices with small displacement rank \cite{Mastronardi07fast},
-the study of the estimator consistency \cite{Palanthandalam10parameter},
-the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models \cite{Kiviet12high} \cite{Kiviet14improved}, and
-a discussion of the causes of bias and inconsistency in homogeneous estimators \cite{Yeredor04homogeneous}.
+the design of a fast algorithm for matrices with small displacement rank \citep{Mastronardi07fast},
+the study of the estimator consistency \citep{Palanthandalam10parameter},
+the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models \citep{Kiviet12high} \citep{Kiviet14improved}, and
+a discussion of the causes of bias and inconsistency in homogeneous estimators \citep{Yeredor04homogeneous}.
 
@@
 The uncertainty of the reported LS estimators for structured EIV problems has not been addressed, and then, remains unknown.
-The estimator uncertainty is expressed using the estimation bias and covariance \cite{Pintelon12Book}.
+The estimator uncertainty is expressed using the estimation bias and covariance \citep{Pintelon12Book}.
 To know the LS estimator uncertainty, we quantified the bias and the covariance of the LS solution of EIV problems in the unstructured and structured cases. 
-This extends the perturbation analysis of the LS estimator of unstructured and uncorrelated problems that was investigated in \cite{Stewart90SPT} and in \cite{Vaccaro94}.
+This extends the perturbation analysis of the LS estimator of unstructured and uncorrelated problems that was investigated in \citep{Stewart90SPT} and in \citep{Vaccaro94}.
 This paper presents a study of the statistics of the LS estimate of unstructured and structured EIV problems. 
  We provide a discussion of the unstructured case as a reference, to get an insight of the impact that the structure of the regression matrix and the correlation between the regression matrix and the regressor perturbations have on the uncertainty of the LS estimate.
-The structured case is motivated by the metrology estimation problem \cite{Markovsky15cep}, where the EIV problem has a Hankel structure, and the perturbations are correlated.
+The structured case is motivated by the metrology estimation problem \citep{Markovsky15cep}, where the EIV problem has a Hankel structure, and the perturbations are correlated.
 The study of the metrology input estimate illustrates a methodology to conduct statistical analysis for any structured EIV problem.
@@
  By following this methodology, the bias and variance of the solutions of EIV problems with other structures is determined, and therefore, the uncertainty of the estimate.
+\end{comment}
 
@@
 The derivation of equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) is described in Appendix 1.
-We use the results described in \cite{Vaccaro94} \S 3 and \cite{Stewart90SPT} \S 2.1 for the expected values of products of unstructured matrices with perturbations.
+We use the results described in \citep{Vaccaro94} \S 3 and \citep{Stewart90SPT} \S 2.1 for the expected values of products of unstructured matrices with perturbations.
 Equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) depend on the unobservable true values $\mathbf{x}$, $\mathbf{K}$, and on the variance of the perturbations.
@@
 
-\newtheorem{thm}{Theorem}
-\newtheorem{lem}[thm]{Lemma}
-\newproof{pf}{Proof}
-
-\small
+% \small    {\small}
 % <how to set font size here to 10 px ? />
 
-\begin{lem}{\small}\label{lem:lemma1}
+\begin{lem} \label{lem:lemma1}
 Let $\mathbf{E} \in {\rm I\!R}^{\left(T-n\right) \times \left(n+1\right)}$ be the partitioned matrix 
@@
 \text{where} \ \mathbf{z}_{i1} = \mathbf{0}, \ \text{and} \ 
-z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{2, j-i} \end{bmatrix} \right), \
-\text{for} \ i = 1, \cdots, T-n, \ \text{and} \ j = 2, \cdots, n+1. \\
+z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{2, j-i} \end{bmatrix} \right), \\
+& \quad \text{for} \ i = 1, \cdots, T-n, \ \text{and} \ j = 2, \cdots, n+1. \\
 & \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \mathbf{E} \right\} = \mathbf{Z}, \ 
 \text{where} \ \mathbf{z}_{1j} = \mathbf{0}, \ \mathbf{z}_{i1} = \mathbf{0}, \ \text{and} \ 
-z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \ \mathbf{D}_{T-n \times T-n}^{2, j-i+1} \right) , \
-\text{for} \ i = 2, \cdots, n+1, \ \text{and} \ j=2, \cdots, n+1 . \\ 
+z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \ \mathbf{D}_{T-n \times T-n}^{2, j-i+1} \right) , \\
+& \quad \text{for} \ i = 2, \cdots, n+1, \ \text{and} \ j=2, \cdots, n+1 . \\ 
 & \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \mathbf{E}^\top \right\} = \mathbf{Z}, \
-\text{where} \ z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} 0 & \mathbf{0}_{n}^\top \\ \mathbf{0}_{n} & \mathbf{D}_{n \times n}^{2, j-i+1} \end{bmatrix} \right), \
-\text{for} \ i = 1,\cdots,T-n, \ \text{and} \ j=1,\cdots,T-n. \\ 
+\text{where} \ z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} 0 & \mathbf{0}_{n}^\top \\ \mathbf{0}_{n} & \mathbf{D}_{n \times n}^{2, j-i+1} \end{bmatrix} \right), \\
+& \quad \text{for} \ i = 1,\cdots,T-n, \ \text{and} \ j=1,\cdots,T-n. \\ 
 & \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \bm{\epsilon} \right\} = \mathbf{z}, \
-\text{where} \ z_i = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{1,n+1-i} \end{bmatrix} \right) , \
-\text{for} \ i = 1,\cdots,T-n . \\
+\text{where} \ z_i = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{1,n+1-i} \end{bmatrix} \right) , \\
+& \quad \text{for} \ i = 1,\cdots,T-n . \\
 & \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \bm{\epsilon} \right\} = \mathbf{z}, \
-\text{where} \ z_1 = 0, \ \text{and} \ z_i = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \ \mathbf{D}_{T-n \times T-n}^{1, n+2-i} \right), \
-\text{for} \ i = 2,\cdots,n+1 . \\
+\text{where} \ z_1 = 0, \ \text{and} \ z_i = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \ \mathbf{D}_{T-n \times T-n}^{1, n+2-i} \right), \\
+& \quad \text{for} \ i = 2,\cdots,n+1 . \\
 & \mathbb{E} \left\{ \mathbf{E} \mathbf{a} \bm{\epsilon}^\top \right\} = \mathbf{Z}, \
-\text{where each column} \ \mathbf{Z}_{j} = - \sigma_{\bm{\epsilon}}^2 \ \mathbf{D}_{T-n \times n+1}^{1,-j} \mathbf{R}_{n+1} \ \mathbf{a} , \
-\text{for} \ j = 1,\cdots,T-n , \ \text{with} \ \mathbf{R}_{n+1} = \begin{bmatrix} & \mathbf{R}_n \\ 0 \end{bmatrix} , \\ & \quad
+\text{where each column} \ \mathbf{Z}_{j} = - \sigma_{\bm{\epsilon}}^2 \ \mathbf{D}_{T-n \times n+1}^{1,-j} \mathbf{R}_{n+1} \ \mathbf{a} , \\
+& \quad \text{for} \ j = 1,\cdots,T-n , \ \text{with} \ \mathbf{R}_{n+1} = \begin{bmatrix} & \mathbf{R}_n \\ 0 \end{bmatrix} , \\ & \quad
 \text{where} \ \mathbf{R}_n \ \text{is a reversal matrix.}
@@
 
-The proof of the lemma is given in Appendix 2.
+The proof of the lemma is given in Appendix \ref{appendix:lemma1}.
 
@@
 \begin{equation} \mathrm{CRLB}(\mathbf{x}) = \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\mathbf{x}}) }{\partial (\widehat{\mathbf{x}})} \right)^\top \mathbf{Fi}^{-1}(\mathbf{x}) \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\mathbf{x}}) }{\partial (\widehat{\mathbf{x}})} \right), \label{eqn:CRLB} \end{equation}
-where $\mathbf{b} (\widehat{\mathbf{x}})$ is the bias of the estimate and $\mathbf{Fi}(\widehat{\mathbf{x}})$ is the Fisher information matrix \cite{Pintelon12Book}.
+where $\mathbf{b} (\widehat{\mathbf{x}})$ is the bias of the estimate and $\mathbf{Fi}(\widehat{\mathbf{x}})$ is the Fisher information matrix \citep{Pintelon12Book}.
 The Fisher information matrix is defined as the expected value of the Hessian of the negative likelihood function 
@@
 
-The minimization problem (\ref{eqn:min_ls}) is equivalent to a structured EIV problem that can be expressed as a linear in the measurements problem \cite{Pintelon12Book}
+The minimization problem (\ref{eqn:min_ls}) is equivalent to a structured EIV problem that can be expressed as a linear in the measurements problem \citep{Pintelon12Book}
 \begin{equation} e (\widehat{\mathbf{x}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\mathbf{x}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\mathbf{x}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 .
@@
 The size of the Fisher information matrix $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
-Moreover, in Chapter 19 of \cite{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ can be obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ after doing inversion by parts, giving
+Moreover, in Chapter 19 of \citep{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ can be obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ after doing inversion by parts, giving
 \begin{equation} \mathbf{Fi}(\mathbf{x}) = \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)^\top \left( \mathbf{M}_1( \mathbf{x} ) \ \mathbf{C}_{\mathbf{z}} \ \mathbf{M}_1^\top( \mathbf{x} ) \right)^{-1} \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)
@@
 \end{equation}
-The standard deviation of the estimate $\widehat{\mathbf{x}}_{\left[1\right]}$ is used to obtain the standard error $\sigma_\mathrm{e}$ of the MC simulation, which decreases with respect to the square root of the number of MC runs $N_{\mathrm{MC}}$ \cite{Hammersley75}:
+The standard deviation of the estimate $\widehat{\mathbf{x}}_{\left[1\right]}$ is used to obtain the standard error $\sigma_\mathrm{e}$ of the MC simulation, which decreases with respect to the square root of the number of MC runs $N_{\mathrm{MC}}$ \citep{Hammersley75}:
 \begin{equation} \sigma_\mathrm{e} = \frac{ \sigma \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) }{\sqrt{N_{\mathrm{MC}}}}, \quad \mathrm{where} \quad \sigma^2 \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) = \frac{1}{N_{\mathrm{MC}}-1} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( \left( \widehat{\mathbf{x}}_{\left[1\right]} \right)_i - \mu \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) \right)^2 } .
@@
 
-Figure \ref{bias_sigma_NMC_unstr_str_n2} shows the empirical bias and the bias predictions, with their corresponding standard errors, for the unstructured (plots in the left) and structured (plots in the right) EIV problems.
+Figure \ref{bias_sigma_NMC_unstr_str_n2} shows the empirical bias and the bias predictions, with their corresponding standard errors, for the unstructured (plots in the top) and structured (plots in the bottom) EIV problems.
 The simulation settings for the structured and correlated EIV problem are described in the following subsection.
@@
  \centering
-%    \begin{minipage}{0.45\textwidth}
-%        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_1l.pdf} 
- %   \end{minipage}
- %   \begin{minipage}{0.45\textwidth}
-  %      \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_1r.pdf} 
-   % \end{minipage}
-  \caption{ \label{bias_sigma_NMC_unstr_str_n2} We observe the results of the LS solutions for the unstructured (left) and the structured (right) EIV problems.
+ \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_1.pdf} 
+  \caption{ \label{bias_sigma_NMC_unstr_str_n2} 
+  We observe the results of the LS solutions for the unstructured (top) and the structured (bottom) EIV problems.
 These results are the empirical bias $b_{\mathrm{e}}$, the predicted bias from exact data $b_{\mathrm{p}}$, the predicted bias from observed data $b_{\widetilde{\mathrm{p}}}$, the empirical standard error $\sigma_{\mathrm{e}}$, and the standard error from the estimations using observed data $\sigma_{\widetilde{\mathrm{p}}}$.
@@
   \centering
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_2l.pdf} 
-    \end{minipage}
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_2r.pdf} 
-    \end{minipage}
-  \caption{ \label{fig:b_bt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (left) between the predicted bias and the empirical bias are proportional to the perturbation noise variance as it is expected, and the relative errors (right) are smaller than 5\% for SNR below 70 dB.
+  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_2.pdf} 
+  \caption{ \label{fig:b_bt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (top) between the predicted bias and the empirical bias are proportional to the perturbation noise variance as it is expected, and the relative errors (bottom) are smaller than 5\% for SNR below 70 dB.
 The bias prediction computed from exact data ${b}_\mathrm{p}$ is very similar to that computed using observed data $b_{\widetilde{\mathrm{p}}}$.
@@
   \centering
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_3l.pdf} 
-    \end{minipage}
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_3r.pdf} 
-    \end{minipage}
-  \caption{ \label{fig:v_vt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (left) between the predicted variances and the empirical variance are proportional to the perturbation noise variance, and the relative errors (right) are smaller than 5\% for SNR higher that 40 dB.
+  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_3.pdf} 
+  \caption{ \label{fig:v_vt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (top) between the predicted variances and the empirical variance are proportional to the perturbation noise variance, and the relative errors (bottom) are smaller than 5\% for SNR higher that 40 dB.
 The variance prediction computed using exact data ${v}_\mathrm{p}$ is very similar to that computed from observed data $v_{\widetilde{\mathrm{p}}}$.
@@
   \centering
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_4l.pdf} 
-    \end{minipage}
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_4r.pdf} 
-    \end{minipage}
-  \caption{ \label{fig:bv_btvt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (left) between the prediction with observed data and the prediction with exact data are proportional to the perturbation noise variance.
-The use of observed data instead of exact data in the prediction formulas is valid when the SNR is above 35 dB since the relative errors (right) are smaller than 5\%.
+  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_4.pdf} 
+  \caption{ \label{fig:bv_btvt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (top) between the prediction with observed data and the prediction with exact data are proportional to the perturbation noise variance.
+The use of observed data instead of exact data in the prediction formulas is valid when the SNR is above 35 dB since the relative errors (bottom) are smaller than 5\%.
 }
@@
 \centering
-\includegraphics[width=0.5\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_5.pdf} 
+\includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_5.pdf} 
 \caption{ \label{fig:y} The structured EIV problem is constructed with 200 samples of the step response to ensure that only transient data is used.
@@
  \begin{equation} {b}_\mathrm{e} = \frac{1}{N_{\mathrm{MC}}} \sum_{i=1}^{N_{\mathrm{MC}}}{ \widehat{u}_i - u } \approx \mu \left( \widehat{u} \right) - u. \end{equation}
-The standard error associated to this empirical bias estimation \cite{Hammersley75} is defined as 
+The standard error associated to this empirical bias estimation \citep{Hammersley75} is defined as 
 \begin{equation} \sigma_\mathrm{e} = \frac{\sigma \left( \widehat{u} \right) }{\sqrt{N_{\mathrm{MC}}}}, \quad \mathrm{where} \quad \sigma^2 \left( \widehat{u} \right) = \frac{1}{N_{\mathrm{MC}}-1} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( { \widehat{u}}_i - \mu \left( \widehat{u} \right) \right)^2 } .
@@
   \centering
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_6l.pdf} 
-    \end{minipage}
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_6r.pdf} 
-    \end{minipage}
-  \caption{ \label{fig:b_bt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (left) between the predicted bias and the empirical bias $b_{\mathrm{e}}$ are proportional to the perturbation noise variance, and the relative errors (right) are smaller than 5\% only for SNR above 40 dB.}
+  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_6.pdf} 
+  \caption{ \label{fig:b_bt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (top) between the predicted bias and the empirical bias $b_{\mathrm{e}}$ are proportional to the perturbation noise variance, and the relative errors (bottom) are smaller than 5\% only for SNR above 40 dB.}
 \end{figure}
@@
   \centering
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_7l.pdf} 
-    \end{minipage}
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_7r.pdf} 
-    \end{minipage}
-  \caption{ \label{fig:v_vt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (left) between the predicted and the empirical variance are proportional to the perturbation variance, and the relative errors (right) between the predicted and the empirical variance are smaller than 5\% for SNR above 45 dB.
+  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_7.pdf} 
+  \caption{ \label{fig:v_vt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (top) between the predicted and the empirical variance are proportional to the perturbation variance, and the relative errors (bottom) between the predicted and the empirical variance are smaller than 5\% for SNR above 45 dB.
 }
@@
   \centering
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_8l.pdf} 
-    \end{minipage}
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_8r.pdf} 
-    \end{minipage}
-  \caption{ \label{fig:bv_btvt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (left) between the predictions computed from observed data and those from exact data are proportional to the perturbation noise variance.
-According to the relative errors (right), the substitution is valid for SNR higher than 45 dB.
+  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_8.pdf} 
+  \caption{ \label{fig:bv_btvt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (top) between the predictions computed from observed data and those from exact data are proportional to the perturbation noise variance.
+According to the relative errors (bottom), the substitution is valid for SNR higher than 45 dB.
 }
@@
   \centering
-\includegraphics[width=0.5\textwidth]{./ChapterStatisticalAnalysis/fig/Fig_9.pdf}
+  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_9.pdf}
   \caption{\label{fig:MSE_CRLB}
@@
 The bias and variance expressions obtained after the statistical analysis depend on each specific structure.
+
+\begin{comment}
 
@@
 
-  The results of the expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$, were described by the authors of this paper in \cite{Quintana19}.
-The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \cite{Vaccaro94, Stewart90SPT}.
+  The results of the expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$, were described by the authors of this paper in \citep{Quintana19}.
+The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \citep{Vaccaro94, Stewart90SPT}.
 It was shown that the bias and variance expressions (\ref{eqn:biasST}) and (\ref{eqn:varST}) are valid predictions of the first and second moments of the LS estimate of a Hankel structured EIV problem.
@@
 
-To find the Cram\'er-Rao lower bound (CRLB) of the structured EIV estimation problem (\ref{eqn:min_ls}), we consider that this structured and correlated estimation problem can be expressed as a linear in the measurements problem \cite{Pintelon12Book}
+To find the Cram\'er-Rao lower bound (CRLB) of the structured EIV estimation problem (\ref{eqn:min_ls}), we consider that this structured and correlated estimation problem can be expressed as a linear in the measurements problem \citep{Pintelon12Book}
 \begin{equation} e (\widehat{\mathbf{x}}, \mathbf{z}) = \mathbf{M}_1( \widehat{\mathbf{x}} ) \ \mathbf{z} = \underbrace{\begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\mathbf{x}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix}}_{\mathbf{M}_1( \widehat{\mathbf{x}} )} \underbrace{\begin{bmatrix} \mathbf{y} \\ \mathrm{vec} ( \mathbf{K} ) \end{bmatrix}}_{\mathbf{z}} = 0 .
@@
 The size of the Fisher information matrix $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
-However, the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ is obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$, by doing inversion by parts \cite{Pintelon12Book} \S 19, and can be expressed as
+However, the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ is obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$, by doing inversion by parts \citep{Pintelon12Book} \S 19, and can be expressed as
 \begin{equation} \mathbf{Fi}(\mathbf{x}) = \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)^\top \left( \mathbf{M}_1( \mathbf{x} ) \mathbf{C}_{\mathbf{z}}  \mathbf{M}_1^\top( \mathbf{x} ) \right)^{-1}  \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right) .
@@
 
-The expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$, were studied in \cite{Quintana19} and their results are described in the following Lemma:
+The expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$, were studied in \citep{Quintana19} and their results are described in Lemma \ref{lem:lemma1}:
 
-\newtheorem{thm}{Theorem}
-\newtheorem{lem}[thm]{Lemma}
-
-\normalsize % \small
+% \normalsize % \small
 %   <how to set font size here to 10 px ? />
 
-\begin{lem}{\small}\label{lem:lemma1}
+\begin{lem}
 Let $\mathbf{E}$ be the matrix defined in (\ref{eqn:matrixE}),
@@
 
-The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \cite{Vaccaro94, Stewart90SPT}.
+The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \citep{Vaccaro94, Stewart90SPT}.
 It was shown that the bias and variance expressions (\ref{eqn:biasST}) and (\ref{eqn:varST}) are good predictions of the first and second moments of the LS estimate of a Hankel structured EIV problem.
@@
 
-In the rest of this section we describe the Cram\'er-Rao lower bound (CRB) of the structured EIV estimation problem (\ref{eqn:min_ls}), that can be expressed as a linear in the measurements problem \cite{Pintelon12Book}
+In the rest of this section we describe the Cram\'er-Rao lower bound (CRB) of the structured EIV estimation problem (\ref{eqn:min_ls}), that can be expressed as a linear in the measurements problem \citep{Pintelon12Book}
 \begin{equation} e (\widehat{\mathbf{x}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\mathbf{x}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\mathbf{x}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 .
@@
 The size of the Fisher information matrix $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
-Moreover, in Chapter 19 of \cite{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ can be obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ after doing inversion by parts, giving
+Moreover, in Chapter 19 of \citep{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ can be obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ after doing inversion by parts, giving
 \begin{equation} \mathbf{Fi}(\mathbf{x}) = \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)^\top \left( \mathbf{M}_1( \mathbf{x} ) \mathbf{C}_{\mathbf{z}}  \mathbf{M}_1^\top( \mathbf{x} ) \right)^{-1}  \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)
@@
 and for an unbiased estimator it is $\mathrm{CRB}_{\mathrm{ub}}(\mathbf{x}) = \mathbf{Fi}^{-1}(\mathbf{x})$
+\end{comment}
 

### 2020-01-27 Gus-thesis.tex
Story 182: Intro complete

@@
 %%
-\documentclass[english,british,BCOR=5mm,DIV=12,fontsize=10pt]{scrbook}
+\documentclass[english, british, BCOR=5mm, DIV=12, fontsize=10pt]{scrbook}
 \usepackage[T1]{fontenc}
 \usepackage[latin9]{inputenc}
-\usepackage[paperwidth=17cm,paperheight=24cm]{geometry}
-\geometry{verbose,tmargin=2.5cm,bmargin=3cm,lmargin=2.5cm,rmargin=2cm}
-\setcounter{secnumdepth}{1}
-\setcounter{tocdepth}{1}
+\usepackage[paperwidth=17cm, paperheight=24cm]{geometry}
+\geometry{verbose, tmargin=2.5cm, bmargin=3cm, lmargin=2.5cm, rmargin=2cm}
+\setcounter{secnumdepth}{2}
+\setcounter{tocdepth}{2}
 \setlength{\parskip}{\medskipamount}
@@
 \definecolor{note_fontcolor}{rgb}{0.800781, 0.800781, 0.800781}
-\usepackage{babel, array, verbatim, float, booktabs, units, amsmath, amssymb, graphicx, tikz}
-\usepackage{mathdots, amsthm, textcomp, times, mathptmx, mathtools, bm, amsbsy}
-\usepackage[unicode=true,pdfusetitle, bookmarks=true, bookmarksnumbered=false, bookmarksopen=false, breaklinks=false,pdfborder={0 0 0}, pdfborderstyle={}, backref=false, colorlinks=false] {hyperref}
+\usepackage{amsmath, amssymb, amsthm, amsbsy, array, babel, booktabs, bm, float, graphicx, mathdots, mathptmx, mathtools, newproof, textcomp, times, units, verbatim}
+\usepackage[unicode=true,pdfusetitle, bookmarks=true, bookmarksnumbered=false, bookmarksopen=false, breaklinks=true,pdfborder={0 0 0}, pdfborderstyle={}, backref=false, colorlinks=false] {hyperref}
 \usepackage[outdir=./]{epstopdf}
+\usepackage{tikz, circuitikz}
 \usetikzlibrary{shapes,arrows,calc,patterns,decorations.pathmorphing,decorations.markings,positioning}
+
+\newtheorem{thm}{Theorem}
+\newtheorem{lem}[thm]{Lemma}
+\newproof{pf}{Proof}
 
@@
 {\textcolor{red}\bgroup}{\egroup}
+
+%\hyphenation{Lon-g-las-t-na-me}
+\usepackage[square]{natbib}
+\bibliographystyle{plainnat} % wmaainf  apalike
 
@@
 
-\include{ChapterRampInput/ChapterRampInput}
-
 \include{ChapterStatisticalAnalysis/ChapterStatisticalAnalysis}
@@
 \include{ChapterExperimentalValidation/ChapterExperimentalValidation}
+
+\include{ChapterRampInput/ChapterRampInput}
 
@@
 
-\bibliographystyle{apalike} % wmaainf
 \bibliography{Gus-thesis}

### 2020-01-27 Summary/Summary.tex
Story 182: Intro complete

@@
 
+\vspace{-1.2cm}
+
+Sensors are dynamical systems and the physical quantity is the input excitation.
+To have fast measurements, the input must be estimated during the sensor transient state.
+Otherwise, users have to wait for the sensor steady state when the response is proportional to the input through the sensor static gain.
+One approach to get fast input estimation is filtering the sensor transient response with another dynamical system aiming to reconstruct the input and to compensate the estimation time.
+Another approach is processing the transient response on a digital signal processor (DSP).
+Compensators are model based and DSPs allow model free methods that can reduce the estimation time.
+
+There exists a model-free method that directly estimates the step input level from the sensor step response.
+This method was formulated as a Hankel structured errors-in-variables (EIV) problem where the regression matrix and the regressor are correlated.
+The structured EIV problem is solved with recursive least-squares to allow real-time implementation.
+The range of application of this step input estimation method is wide because it is independent from sensor models.
+
+The stochastic properties of the step input estimation method are unknown and are not straightforwardly evident.
+To validate the method for metrology applications, we studied the estimate uncertainty by doing a statistical analysis of the step input estimation.
+The estimate uncertainty can be defined in terms of the estimate bias and variance.
+The statistical analysis of the estimation method yields expressions that predict the estimate first and second moments.
+
+We computed the Cram\'er-Rao bound of the structured EIV problem to determine the minimum theoretical variance of the estimate.
+This bound was compared to the estimate mean squared error (MSE) in simulations and in experiments with temperature and weighing sensors.
+It was found that the input estimation is biased but with small variance, and that the estimate MSE is less than one order of magnitude larger than the Cram\'er-Rao bound.
+In the practice, the step input estimation method was found to be robust when the Gaussian and white measurement noise assumption was not satisfied.
+
+\begin{comment}
 \section{Summary - Ramp input}
@@
 It was found that the predicted mean-squared error of the input estimate is close to an approximate Cram\'er-Rao lower bound for biased estimators.
+\end{comment}
+
+%\newpage

### 2020-01-30 ChapterAppendices/ChapterAppendices.tex
Preliminaries ready

@@
 
-\section{Appendix - Derivation of bias and covariance expressions.  \label{appendix:biascovSEIV}}
+\section{Appendix - Derivation of bias and covariance expressions.  \label{appendix:biascovSEIV} }
 
@@
 
-On the other hand, due to the correlation, the expressions that approximate the expected value of the LS estimate of the structured and EIV problem (\ref{eqn:min_ls}) have a different form:
+On the other hand, due to the correlation, the expressions that approximate the expected value of the LS estimate of the structured and correlated EIV problem (\ref{eqn:ddsiemnd}) have a different form:
 \begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  +  \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E}  \right\} \mathbf{x} \\
@@
 
-\section{Appendix - Proof of Lemma \ref{lem:lemma1} \label{appendix:lemma1}}
+\section{Appendix - Proof of Lemma \ref{lem:lemma1} \label{appendix:lemma1} }
 
@@
 
-\section{Appendix - Jacobians in Ramp input ML estimation  \label{appendix:Jacobians}}
+\section{Appendix - Jacobians in affine input estimation by ML method \label{appendix:Jacobians} }
 

### 2020-01-30 ChapterConclusion/ChapterConclusion.tex
Preliminaries ready

@@
 
-\chapter[Conclusions and future work]{Conclusions and \linebreak future work} \label{chap:concl}
+\chapter{Conclusions and \linebreak future work} \label{chap:concl}
 
@@
 
-In this final chapter, we first summarize the contributions made to both parts of the thesis: \textbf{Part \ref{part:Filter}} and \textbf{Part \ref{part:PA}}. Then, we propose
-some possible further improvements to extend this work in future research.
+%In this final chapter, we first summarize the contributions made to both parts of the thesis: \textbf{Part \ref{part:Filter}} and \textbf{Part \ref{part:PA}}.
+Then, I propose some possible further improvements to extend this work in future research.
 
-\newpage
-\section*{Conclusions}
+%\newpage
+%\section*{Conclusions}
 

### 2020-01-30 ChapterExperimentalValidation/ChapterExperimentalValidation.tex
Preliminaries ready

@@
 \begin{quote}
-\emph{The results of this chapter ...}\vfill{}
+\emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I.
+Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Experimental validation of a data-driven step input estimation method for dynamic measurements", IEEE Transactions on Instrumentation and Measurement Journal, 2019, doi: 10.1109/TIM.2019.2951865.
+\nocite{QuintanaTIM} }\vfill{}
 \end{quote}
 
- \section{Introduction}
+ \vfill{}
+
+% \section{Introduction}
 
@@
 The MC simulation performed $N_{MC} = 10^4$ runs of the data-driven step input estimation with different realizations of the measurement noise $\bm{\epsilon}$.
-We are interested in the first element of $\widehat{\mathbf{x}}$, which is the input estimate $\widehat{u}$.
-The measurement noise variance was selected to have a signal-to-noise ratio (SNR) in the interval [30 dB, 80 dB], according to
-\begin{equation} \mathrm{SNR} = 20 \log_{10}{ \dfrac{ \sqrt{ \dfrac{1}{T} \int\limits_{0}^{T}{ y(\tau)^2  \mathrm{d} \tau } } }{ \sigma_{ \epsilon }} } \end{equation}
+The MC simulation was conducted processing $T = 5000$ samples of a simulated transient step response $\widehat{\mathbf{y}}$ generated by a stable linear time-invariant (LTI) system of order $n = 5$, with a sampling frequency of $f_s=4$ kHz.
+This system is a state space model obtained with the System Identification Toolbox using the measured step response of the actual sensor described in the Practical Implementation Section.
+This model represents a weighing sensor, and in the simulations the sensor is excited with a mass of 138.32 g following a step input profile. 
+As can be seen in Figure \ref{fig:uh_sim}, the steady state response of the weighing sensor model is practically reached after 500 samples because from there on the relative error between the transient response and the steady-state response is smaller than 0.2\%.
 
-The MC simulation was conducted processing $T = 5000$ samples of a simulated transient step response $\widehat{\mathbf{y}}$ generated by a stable LTI system of order $n = 5$, with a sampling frequency of $f_s=4$ kHz.
-This system is a state space model obtained with the System Identification Toolbox using the measured step response of the actual sensor described in the Practical Implementation Section.
-The steady state response of the system is practically reached after 500 samples because from there on the relative error between the transient response and the steady-state response is smaller than 0.2\%, see Fig.
-\ref{fig:uh_sim}.
+\begin{comment}
+, see Figure \ref{fig:weighing_sensor}
+\begin{figure}[htb!]
+\centering
 
-A sensor is a dynamic system, therefore, a fast measurement process must necessarily cope with the system’s transient response.
-In that respect we must distinguish between the transient response of the system under test, and the transient phase of the measurement process (i.e.
-before the process has settled on a final measurement outcome).
-Notice that the transient phase of the measurement process is considerably smaller than the settling time of the system under test, and this is a major advantage of the step input estimation method.
+\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
+\tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
+\tikzstyle{damper}=[thick,decoration={markings,  
+  mark connection node=dmp,
+  mark=at position 0.5 with 
+  {
+    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
+    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
+    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
+  }
+}, decorate]
+\tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]
+
+\node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
+\node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$0.138 s(t)$};
+
+\node (ground1) at (M.south) [ground,yshift=-1.5cm,xshift=-0.625cm,anchor=north] {};
+\draw (ground1.north west) -- (ground1.north east);
+\draw [spring] (ground1.north) -- ($(M.south east)!(ground1.north)!(M.south west)$);
+
+\node (groundc) at (M.south) [ground,yshift=-1.5cm,anchor=north] {}; 
+\draw (groundc.north west) -- (groundc.north east);
+
+\node (ground2) at (M.south) [ground,yshift=-1.5cm,xshift=0.625cm,anchor=north] {};
+\draw (ground2.north west) -- (ground2.north east);
+\draw [damper] (ground2.north) -- ($(M.south east)!(ground2.north)!(M.south west)$);
+
+\node[draw=none,fill=none] at (-0.9cm,-1cm) {$k_{\mathrm{s}}$};
+\node[draw=none,fill=none] at (0.15cm,-1cm) {$k_{\mathrm{d}}$};
+\node[draw=none,fill=none] at (2.0cm,1.0cm) {$y$};
+\draw [-latex,thick]  ++(2.2cm,-1cm) -- +(0cm,2.25cm);
+
+\draw [-latex,thick] (M.east) ++(0,0) -- +(1cm,0);
+\draw [line width=0.25mm] (2.2cm,-1cm) -- (2.2cm,1cm);
+\draw [line width=0.25mm] (2.1cm,-1cm) -- (2.3cm,-1cm);
+\draw [line width=0.25mm] (2.1cm,1cm) -- (2.3cm,1cm);
+\draw [line width=0.25mm] (2.1cm,-0.5cm) -- (2.3cm,-0.5cm);
+\draw [line width=0.25mm] (2.1cm,0.5cm) -- (2.3cm,0.5cm);
+\draw [line width=0.25mm] (2.15cm,-0.25cm) -- (2.25cm,-0.25cm);
+\draw [line width=0.25mm] (2.15cm,0.25cm) -- (2.25cm,0.25cm);
+\draw [line width=0.25mm] (2.15cm,-0.75cm) -- (2.25cm,-0.75cm);
+\draw [line width=0.25mm] (2.15cm,0.75cm) -- (2.25cm,0.75cm);
+\draw [line width=0.25mm] (2.1cm,0cm) -- (2.3cm,0cm);
+
+\end{tikzpicture}
+
+\caption{\label{fig:weighing_sensor} A mass-spring-damper model of 5-th order is identified using the MATLAB System Identification toolbox and the step response of a load cell sensor excited with a mass of 138.32 g in step input experiments.
+The identified model is used to perform simulations under different measurement noise levels to validate the data-driven step input estimation method, and the expressions that predict the bias and the variance of the step input level estimate.}
+\end{figure}
+\end{comment}
 
@@
 \end{figure}
+
+A sensor is a dynamic system, therefore, a fast measurement process must necessarily cope with the system transient response.
+In that respect we must distinguish between the transient response of the system under test, and the transient phase of the measurement process (i.e.
+before the process has settled on a final measurement outcome).
+Notice that the transient phase of the measurement process is considerably smaller than the settling time of the system under test, and this is a major advantage of the step input estimation method.
+
+We are interested in the first element of $\widehat{\mathbf{x}}$, which is the input estimate $\widehat{u}$.
+The measurement noise variance was selected to have a signal-to-noise ratio (SNR) in the interval [30 dB, 80 dB], according to
+\begin{equation} \mathrm{SNR} = 20 \log_{10}{ \dfrac{ \sqrt{ \dfrac{1}{T} \int\limits_{0}^{T}{ y(\tau)^2  \mathrm{d} \tau } } }{ \sigma_{ \epsilon }} } \end{equation}
 
@@
 
-Fig. \ref{bias_sigma_sim_unstr_str_n2} shows the empirical bias, the bias predictions and the standard errors of the MC simulation.
+Figure \ref{bias_sigma_sim_unstr_str_n2} shows the empirical bias, the bias predictions and the standard errors of the MC simulation.
 It can be seen that the empirical bias $b_\mathrm{e}$ and the predicted bias $b_{\widetilde{\mathrm{p}}}$ are proportional to the perturbation noise variance while the standard errors $\sigma_\mathrm{e}$ and $\sigma_{\widetilde{\mathrm{p}}}$ are proportional to the perturbation noise standard deviation.
@@
 where $b$ and $v$ are the bias and the variance of the step input estimate, can be applied to the obtained empirical and predicted results and can be compared to the Cram\'er-Rao lower bound (CRB).
-Fig.
-\ref{fig:MSE_CRB} shows that $\mathrm{MSE}_\mathrm{e} = b_\mathrm{e}^2 + v_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}  = b_{\widetilde{\mathrm{p}}}^2 + v_{\widetilde{\mathrm{p}}}$ have the same proportionality with respect to the measurement noise variance as the bound for an unbiased estimator $\mathrm{CRB_{ub}}$.
+Figure \ref{fig:MSE_CRB} shows that $\mathrm{MSE}_\mathrm{e} = b_\mathrm{e}^2 + v_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}  = b_{\widetilde{\mathrm{p}}}^2 + v_{\widetilde{\mathrm{p}}}$ have the same proportionality with respect to the measurement noise variance as the bound for an unbiased estimator $\mathrm{CRB_{ub}}$.
 For SNR above 35 dB, $\mathrm{MSE}_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ are equivalent, and below 35 dB the difference between them is of less than a factor of 10.
@@
 We obtained an approximation of the $\mathrm{CRB}_{\mathrm{b}}$ for our biased estimator using the partial derivative of the bias in expression (\ref{eqn:biasST}).
-Fig.
-\label{fig:MSE_CRB} shows that the bounds for the unbiased and biased estimators are almost equal because the partial derivatives of the bias are negligible w.r.t.
+Figure \label{fig:MSE_CRB} shows that the bounds for the unbiased and biased estimators are almost equal because the partial derivatives of the bias are negligible w.r.t.
 1 in Equation (\ref{eqn:CRB_EIV}).
@@
 In order to get more insight into the step input estimation method we conducted another simulation study. 
-The step input estimation method assumes the order $n$ is given in (\ref{eqn:matrixK}).
+The step input estimation method assumes the order $n$ is given in (\ref{eqn:min_seiv}).
 In this simulation, the step input estimation method processed the step response generated by a $5-th$ order system using different values of $n$ in the interval from 2 to 100.
@@
 For each order $n$ and SNR value, 100 step input estimations are performed from independent noise realizations. 
-Fig. \ref{fig:msee_vs_n_sim} shows the average of the squared biases and the variances, and the MSEs of the input estimate using the first 500 samples.
+Figure \ref{fig:msee_vs_n_sim} shows the average of the squared biases and the variances, and the MSEs of the input estimate using the first 500 samples.
 It is evident that the estimation variance and MSE depend on the SNR.
@@
 This is the asymptotic behavior of the estimation statistical moments with respect to the number of regressors.
-Nevertheless, the simulation results presented in Fig.
-\ref{fig:msee_vs_n_sim} show that the variance first increases for small values of $n$, followed by a decrement and finally after $n \approx 40$ the variances exhibit a slow and steady increment.
+Nevertheless, the simulation results presented in Figure \ref{fig:msee_vs_n_sim} show that the variance first increases for small values of $n$, followed by a decrement and finally after $n \approx 40$ the variances exhibit a slow and steady increment.
 This apparent contradiction does not prove the invalidity of the estimation method since the results presented correspond to a finite sample size and the asymptotic results cannot be applied.
@@
 The optimum value of $n$ is not necessarily equal to the order of the generating system and varies for each SNR.
-According to the plots in Fig. \ref{fig:msee_vs_n_sim}, there are orders that provide local minima of the step input estimation MSEs.
-From the right hand side of Fig. \ref{fig:msee_vs_n_sim}, the orders that give the first two MSE minima were identified and those values are listed in Table \ref{Tbl:orders}.
+According to the plots in Figure \ref{fig:msee_vs_n_sim}, there are orders that provide local minima of the step input estimation MSEs.
+From the right hand side of Figure \ref{fig:msee_vs_n_sim}, the orders that give the first two MSE minima were identified and those values are listed in Table \ref{Tbl:orders}.
 For each SNR, there is a first minimum at a low order and a second minimum at a high order.
@@
 The step input estimation method processed 100 measured sensor step responses, assuming the sensor is of $\mathrm{7-}th$ order.
-Fig. \ref{fig:uh_exp} shows a typical measured transient response $\widetilde{y}$ and an example of the estimated input $\widehat{u}$.
+Figure \ref{fig:uh_exp} shows a typical measured transient response $\widetilde{y}$ and an example of the estimated input $\widehat{u}$.
 
-The empirical bias $b_{\mathrm{e}}$ is the difference between the average of the 100 estimates $\widehat{u}$ and the mass calibration value \linebreak $u = 138.32$ g, at each instant of time, and the standard error $\sigma_\mathrm{e}$ is the standard deviation of the mean estimate of the responses processed, i.e.,
+The empirical bias $b_{\mathrm{e}}$ is the difference between the average of the 100 estimates $\widehat{u}$ and the mass calibration value $u = 138.32$ g, at each instant of time, and the standard error $\sigma_\mathrm{e}$ is the standard deviation of the mean estimate of the responses processed, i.e.,
 \begin{equation} \begin{aligned} \widehat{\mu}_{e} &= \frac{1}{100} \sum_{i=1}^{100}{ \widehat{u}_{i} }, \quad {b}_{\mathrm{e}} = \widehat{\mu}_{e} - u, \quad \mathrm{and} \\  \sigma_{\mathrm{e}} &= \frac{\widehat{\sigma}}{\sqrt{100}}, \quad \mathrm{where} \ \widehat{\sigma}^2 = \frac{1}{99} \sum_{i=1}^{100}{ \left( \widehat{u}_i - \widehat{\mu}_{e} \right)^2 } .
@@
 These expressions require the measurement noise variance $\sigma_\epsilon^2$ to obtain the bias and variance prediction.
-One way to estimate the measurement noise variance is computing the variance of each sensor steady state response, see Fig. \ref{fig:y_ss}.
+One way to estimate the measurement noise variance is computing the variance of each sensor steady state response, see Figure \ref{fig:y_ss}.
 Later in this section we will explore another way to estimate the measurement noise variance.
@@
 
-Fig.
-\ref{fig:b_sigma_exp_str} shows the empirical bias $b_{\mathrm{e}}$ and the standard error $\sigma_{\mathrm{e}}$ that result after processing the first $T=500$ samples of the 100 measured step responses $\widetilde{\mathbf{y}}$.
+Figure \ref{fig:b_sigma_exp_str} shows the empirical bias $b_{\mathrm{e}}$ and the standard error $\sigma_{\mathrm{e}}$ that result after processing the first $T=500$ samples of the 100 measured step responses $\widetilde{\mathbf{y}}$.
 The standard error is smaller than the bias.
@@
 The power spectrum of the signal $\widetilde{\mathbf{y}}$ is given in decibels by $\widetilde{\mathbf{Y}}_\mathrm{dB} = 20 \log_{10}{ |\widetilde{\mathbf{Y}}| } $.
-Figure \ref{fig:YYhR_n7} shows the corresponding power spectra of the sensor response $\mathbf{Y}_\mathrm{dB}$, the simulated response $\widehat{\mathbf{Y}}_\mathrm{dB}$, and the residual $\mathbf{R}_\mathrm{dB}$.
+Figure \ref{fig:YYhR_n7} shows the corresponding power spectra of the sensor response $\widetilde{\mathbf{Y}}_\mathrm{dB}$, the simulated response $\widehat{\mathbf{Y}}_\mathrm{dB}$, and the residual $\mathbf{R}_\mathrm{dB}$.
 There are frequency components near the main resonance peak in the magnitude spectrum of the residual.
@@
 The 5 dB difference provides a conservative bound since the bias and variance computed from 50 dB of SNR are higher than those obtained using the variance estimation from the steady-state response. 
-Fig.
-\ref{fig:b_sigma_exp_str2} shows a comparison of the results obtained with both measurement noise variance estimations after processing the first $T=500$ samples of the step response $\mathbf{y}$.
+Figure \ref{fig:b_sigma_exp_str2} shows a comparison of the results obtained with both measurement noise variance estimations after processing the first $T=500$ samples of the step response $\widetilde{\mathbf{y}}$.
 Using expression (\ref{eqn:biasST}), the bias prediction  $b_{\widetilde{\mathrm{p}}2}$ obtained using an SNR of 50 dB approximates more closely the empirical bias than $b_{\widetilde{\mathrm{p}}1}$ obtained using an SNR of 55dB.
@@
 We investigated another aspect of the step input estimation method performance when processing measured step responses.
-The step input estimation method requires an assumption of the generating system order in the formulation of the estimation problem (\ref{eqn:min_ls}).
+The step input estimation method requires an assumption of the generating system order in the formulation of the estimation problem (\ref{eqn:min_seiv}).
 The estimation method performance is assessed under different assumptions of the values of $n$ in the interval from 2 to 10. 
 For each value of $n$, 100 step input estimations are computed from measured transient responses and the empirical MSEs are compared.
-Fig. \ref{fig:msee_vs_n_exp} shows that, similar to the observations made in the simulation study, the MSEs have two local minima at $n=7$ and $n=48$.
+Figure \ref{fig:msee_vs_n_exp} shows that, similar to the observations made in the simulation study, the MSEs have two local minima at $n=7$ and $n=48$.
 It is recommended to use $n = 7$ in the estimation method to provide a small estimation MSE without the higher computational complexity that $n=48$ implies.

### 2020-01-30 ChapterIntroduction/ChapterIntroduction.tex
Preliminaries ready

@@
 \chapter{Introduction} \label{chap:Introduction}
+
+\vfill{}
 
@@
 These expressions estimate the bias and covariance that the data-driven step input estimation method will exhibit for given sample size and perturbation level.
-I compared the mean squared error of the LS estimate to the minimum variance specified by the Cram\'er-Rao lower bound of the structured EIV problem, to determine the conditions in which the data-driven step input estimation method is appropriate for practical applications.
+I compared the mean squared error of the LS estimate to the minimum variance specified by the Cram\'er-Rao lower bound of the structured EIV problem, to determine the conditions in which the data-driven step input estimation method is appropriate for practical applications \citep{QuintanaCSDA}.
 
@@
 One of the assumptions of the step input estimation method formulation is that the measurement noise is Gaussian white. 
-Nevertheless, the estimation results show that the method is still able to provide useful input estimations.
+Nevertheless, the estimation results show that the method is still able to provide useful input estimations \citep{QuintanaTIM}.
 
-The empirical bias an covariance obtained after repeating the weighing experiment was compared to the bias and covariance estimations obtained in the previous research work \citep{Quintana19}, and to the minimal variance given by the  Cram\'er-Rao lower bound (CRLB) of the structured EIV problem.
+The empirical bias an covariance obtained after repeating the weighing experiment was compared to the bias and covariance estimations obtained in the previous research work \citep{QuintanaCSDA}, and to the minimal variance given by the  Cram\'er-Rao lower bound (CRLB) of the structured EIV problem.
 I found that the mean squared error (MSE) of the step input estimate is near the CRLB, and the distance between the MSE and the CRLB provides a confidence interval for the input estimate with respect to the level of the measurement noise.
@@
 
-The proposed data-driven method for the estimation of the affine input parameters.
-The affine input is parameterized as a straight line model where the slope and the interception are the parameters of interest.
+In the proposed data-driven method for the estimation of the affine input \citep{QuintanaMEAS},
+the affine input is parameterized as a straight line model where the slope and the interception are the parameters of interest.
 The data-driven affine input estimation method formulates a structured errors-invariables (EIV) problem, similar to the one fourmulated for the step input estimation.

### 2020-01-30 ChapterPreliminaries/ChapterPreliminaries.tex
Preliminaries ready

@@
 
-\section{Preliminaries}
+\section{Input estimation is a linear system problem}
 
@@
 
+\subsection{Input estimation with system model}
+
 The discrete-time state-space representation of an LTI system is
@@
 \begin{equation} 
-\underbrace{ \begin{bmatrix} y(0) \\ y(1) \\ y(2) \\ \vdots \\ y(T) \end{bmatrix} }_{\mathbf{y}}}
- = \underbrace{ \begin{bmatrix} \mathbf{C} \\ \mathbf{C} \mathbf{A} \\ \mathbf{C} \mathbf{A}^2 \\ \vdots \\ \mathbf{C} \mathbf{A}^T \end{bmatrix} }_{\mathbfcal{O}}} \mathbf{x}(0) +
- \underbrace{ \begin{bmatrix} \mathbf{D} \\ \mathbf{C} \mathbf{B} & \mathbf{D} \\ \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \\ \vdots & \ddots \\ \mathbf{C} \mathbf{A}^{T-1} \mathbf{B} & \cdots  &  \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \end{bmatrix} }_{\mathbfcal{T}}}
-\underbrace{ \begin{bmatrix} u(0) \\ u(1) \\ u(2) \\ \vdots \\ u(T) \end{bmatrix} }_{\mathbf{u}}} ,
+\underbrace{ \begin{bmatrix} y(0) \\ y(1) \\ y(2) \\ \vdots \\ y(T) \end{bmatrix} }_{\mathbf{y}}
+ = \underbrace{ \begin{bmatrix} \mathbf{C} \\ \mathbf{C} \mathbf{A} \\ \mathbf{C} \mathbf{A}^2 \\ \vdots \\ \mathbf{C} \mathbf{A}^T \end{bmatrix} }_{\mathbfcal{O}} \mathbf{x}(0) +
+ \underbrace{ \begin{bmatrix} \mathbf{D} \\ \mathbf{C} \mathbf{B} & \mathbf{D} \\ \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \\ \vdots & \ddots \\ \mathbf{C} \mathbf{A}^{T-1} \mathbf{B} & \cdots  &  \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \end{bmatrix} }_{\mathbfcal{T}} \underbrace{ \begin{bmatrix} u(0) \\ u(1) \\ u(2) \\ \vdots \\ u(T) \end{bmatrix} }_{\mathbf{u}} ,
  \label{eqn:knownmodel} \end{equation}
 to find the input. 
-This system of equations is constructed from the recursions of (\ref{eqn:dtsslti}), where $\mathbfcal{O}$ is the observability matrix of the system, and $\mathbfcal{T}$ is a Toeplitz matrix of the Markov parameters of the system.
-# formulated as the input estimation problem from the response of a system.
+This system of equations $\mathbf{y} = \mathbfcal{O} \mathbf{x}(0) + \mathbfcal{T} \mathbf{u}$ is constructed from the recursions of (\ref{eqn:dtsslti}), where $\mathbfcal{O}$ is the observability matrix of the system, and $\mathbfcal{T}$ is a Toeplitz matrix of the Markov parameters of the system.
+% formulated as the input estimation problem from the response of a system.
 
@@
 If the initial conditions and the measurement noise are unknown, the typical approach is to feed the output $\mathbf{y}$ to an additional system.
-This additional system is modeled to invert the dynamics of the sensor by doing an operation that is equivalent to deconvolution, aiming to estimate the input $\mathbf{u}$ with the output $\widehat{\mathbf{u}}$ of the additional system.
-This system is called compensator because the transient time of the estimated input $\widehat{\mathbf{u}}$ is smaller than the transient time of $\mathbf{y}$, see Figure \ref{fig:compensator}.
+This additional system is built to invert the dynamics of the sensor by doing an operation that is equivalent to deconvolution, see Figure \ref{fig:compensator}.
+The output $\widehat{\mathbf{u}}$ of the additional system aims to estimate the input $\mathbf{u}$.
+This system is called compensator because the transient time of the input estimation $\widehat{\mathbf{u}}$ is smaller than the transient time of $\mathbf{y}$.
 
@@
 
-If a model of the sensor is not available, the first idea may be to identify a model using the inputs and outputs, and later to estimate the input using the identified model.
-However, the input is unknown and the model must be identified using only the sensor output.
-Modeling the input as a multiple of the unit step function allows an analytical treatment of the input estimation problem.
-This chapter describes step input estimation methods derived from linear systems theory. 
-The description starts with the case in which the model is known, continues with the unknown sensor model case that performs model identification, and ends with a direct estimation of the input without model identification.
-# The suitability of data-driven methods on-line implementation is higher because there is no need for high computational power to identify the system.
+In the next subsections the input is modeled as a multiple of the unit step function, to describe methods to estimate the input step level in two conditions: when the model of the sensor is a-priori given and when the method is not available.
 
-Step input estimation with system model
+\subsubsection{Step input estimation with system model}
 
@@
 
-Step input estimation without system model
+\subsection{Step input estimation without system model}
 
-Reduction to autonomous equivalent model
+If a model of the sensor is not available, it might be needed to identify a model using the inputs and outputs to design later a compensator using the identified model.
+However, the input is unknown and the model should be identified using only the sensor output.
 
-A first method, when the model of the LTI sensor is unknown and we have exact data, identifies a model of the sensor from the step response, and the estimates the input using the identified sensor model.
+\subsubsection{Reduction to autonomous equivalent model}
+
+One method identifies a model of the sensor from the step response, assuming we have exact data, and later estimates the input step level using the identified sensor model.
 The model identification consists in the estimation of the matrices $\widehat{\mathbf{A}}_\text{a}$ and $\widehat{\mathbf{C}}_\text{a}$, and the initial conditions $\widehat{\mathbf{x}}(0)$.
 To do this, consider that a Hankel matrix $\mathbfcal{H}(\mathbf{y}) \in \mathbb{R}^{n \times n}$, constructed from any linearly independent $n$ autonomous responses from the sensor initial conditions, is full column rank.
-Thus, we can express 
+We can express 
 \[ \underbrace{ \begin{bmatrix} y(1) & y(2) & \cdots & y(n) \\ y(2) & y(3) & \cdots & y(n+1) \\ \iddots & \iddots & \iddots \\ y(n) & y(n+1) & \cdots & y(2n-1) \end{bmatrix} }_{ \mathbfcal{H}(\mathbf{y}) }= \underbrace{ \begin{bmatrix} \widehat{\mathbf{C}}_\text{a} \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a} \\ \vdots \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a}^{n} \end{bmatrix} }_{ \mathbfcal{O}_\text{a} }  \underbrace{ \begin{bmatrix} \mathbf{x}_\text{a}(0) & \mathbf{x}_\text{a}(1) & \cdots & \mathbf{x}_\text{a}(n) \end{bmatrix} }_{ \mathbf{X}_\text{ini} } , \]
@@
 
-Identification of the model followed by input estimation
+\subsubsection{Identification of the model followed by input estimation}
 
-A second method is proposed to estimate the step input when the model of the LTI system is unknown but its steady-state gain $\mathbf{G}$ is known.   
-Using the steady-state gain $\mathbf{G}$, we can express $\bar{\mathbf{y}} = \mathbf{G} \bar{\mathbf{u}}$, where $\bar{\mathbf{u}}$ is the input exact value and $\bar{\mathbf{y}}$ is the sensor steady state response.
+There is a second method that estimates the step input when the model of the LTI sensor is unknown but its static gain $\mathbf{G}$ is given.
+Using the static gain $\mathbf{G}$, we can express $\bar{\mathbf{y}} = \mathbf{G} \bar{\mathbf{u}}$, where $\bar{\mathbf{u}}$ is the input exact value and $\bar{\mathbf{y}}$ is the sensor steady state response.
 The total response of the system is the sum of the transient and the steady-state responses.
@@
 \[ \underbrace{ \begin{bmatrix}y(0) \\ y(1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \mathbf{C}_\text{a} \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a} \\ \vdots & \vdots \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a}^T \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{\mathbf{u}} \\ \mathbf{x}(0) \end{bmatrix} .\]
-It is necessary to estimate the observability matrix of the augmented system, followed by the estimation of the input  $\mathbf{u}$, and the initial conditions $\mathbf{x}_{\text{ini}}$, using least squares
+Then, it is necessary to estimate the observability matrix of the augmented system, followed by the estimation of the input  $\mathbf{u}$, and the initial conditions $\mathbf{x}_{\text{ini}}$, using least squares
 \[ \begin{bmatrix} \widehat{\mathbf{u}} \\ \widehat{\mathbf{x}}_{\text{ini}} \end{bmatrix} = \left( \mathbf{K}^\top \mathbf{K} \right)^{-1} \mathbf{K}^\top \mathbf{y}. \]
 
-Data-driven estimation of an unknown step input given system response
+\subsubsection{Data-driven estimation of an unknown step input given a sensor step response}
+%Another alternative to performing model identification is presented, it is a direct estimation of the input without model identification.
+%# The suitability of data-driven methods on-line implementation is higher because there is no need for high computational power to identify the system.
 
-A third method directly estimates the input value from the step response, without an identification of the sensor model.
+A third method directly estimates the input step level from the step response, without the identification of a sensor model.
 Applying the first difference operator, $\Delta = \sigma - 1$, to the system state space representation (\ref{eqn:dtsslti}), we have
@@
 where $\sigma$ is the shift operator, defined as $(\sigma^\tau y) (t) = y(t + \tau)$, and
-$\Delta \mathbf{u}(k) = \mathbf{0}$, for $k \geq 0$, because we have a step input, and
+$\Delta \mathbf{u}(k) = \mathbf{0}$, for $k \geq 0$, and
 $\Delta \mathbf{x}(0) = (\mathbf{A} - \mathbf{I}) \mathbf{x}(0) + \mathbf{B} \bar{\mathbf{u}}$.
@@
 When the response $\Delta \mathbf{y}$ is persistingly exciting of order $L$, i.e., when the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta \mathbf{y})$ of $L+1$ block rows constructed from $\Delta \mathbf{y}$ satisfies
-\[\mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta \mathbf{y} \right) \right) \leq L,  \]
+\begin{equation} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta \mathbf{y} \right) \right) \leq L,  \end{equation}
 the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta \mathbf{y})$ provides a linear map to the free responses of the augmented autonomous system (\ref{eqn:augmented}).  
@@
 that is equivalent to
-\[ \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \mathbf{G} & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ \mathbf{G} & \Delta y(T-n) & \Delta y(T-n+1) & \cdots & \Delta y(T-1) \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{\mathbf{u}} \\ \bm{\ell} \end{bmatrix} .\]
+\begin{equation} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \mathbf{G} & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ \mathbf{G} & \Delta y(T-n) & \Delta y(T-n+1) & \cdots & \Delta y(T-1) \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} \bar{\mathbf{u}} \\ \bm{\ell} \end{bmatrix} }_{\mathbf{x}} , \label{eqn:ddsiemexd} \end{equation}
+where $T$ is the number of samples, and the vector ${\ell}$ is a linear transformation of the system initial conditions \citep{Markovsky15cep}.  
+The solution to this system of equations exist and is unique when we have exact data.
+Nevertheless, in the practice, the observations of the sensor step response are perturbed by noise.
 
-The step response observations ${y}(t) \in \mathbb{R}^{}$ for $t = 1,\ldots,T$, where $T$ is the number of samples, is perturbed by normally distributed measurement noise $\epsilon$ with zero mean and given variance $\sigma_{\epsilon}^2$.
- We have \begin{equation} y = \bar{y} + \epsilon, \quad \mathrm{and} \quad K = \bar{K} + E, \label{eqn:y0plusnoise} \end{equation} 
- where $\bar{y}$ is the exact system response and $E$ is the matrix that whose elements are the corresponding noise perturbations to the exact data elements in $\bar{K}$, and the vector ${\ell}$ is a linear transformation of the system initial conditions.
- The observed response of the system is a step-invariant discretization of the continuous-time response.
+\subsection{Data-driven step input estimation in presence of noise}
 
- The system of equations ${y} = {K} x$ is an errors-in-variables (EIV) minimization problem with block Hankel structure.
- The perturbation noise present in ${K}$ is correlated to the perturbation of the elements in ${y}$.
- This EIV problem admits a least-squares (LS) solution and the recursive LS enables the online implementation for real-time step input estimation applications.
- The step input estimation method was described in \citep{Markovsky15cep}.
+Assuming that the observations of the step response $\mathbf{y}$ are perturbed by measurement noise $\bm{\epsilon}$ of zero mean and given variance $\sigma_{\bm{\epsilon}}^2$, we can express 
+\begin{equation} \widetilde{\mathbf{y}} = \mathbf{y} + \bm{\epsilon} , \label{eqn:y0plusnoise} \end{equation} 
+and
+\begin{equation} \widetilde{\mathbf{K}} = \mathbf{K} + \mathbf{E} . \label{eqn:K0plusnoise} \end{equation}The matrix $\mathbf{E}$ is constructed with the noise data, and is given as
+\begin{equation} \mathbf{E} = \begin{bmatrix} 0 & \Delta \epsilon(1) & \Delta \epsilon(2) & \cdots & \Delta \epsilon(n) \\ 0 & \Delta \epsilon(2) & \Delta \epsilon(3) & \cdots & \Delta \epsilon(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ 0 & \Delta \epsilon(T-n) & \Delta \epsilon(T-n+1) & \cdots & \Delta \epsilon(T-1) \end{bmatrix} .
+\label{eqn:matrixE} \end{equation}
+Therefore, the matrix $\widetilde{\mathbf{K}}$ is expressed as
+\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} g & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ g & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ g & \Delta \widetilde{y}(T-n) & \Delta \widetilde{y}(T-n+1) & \cdots & \Delta \widetilde{y}(T-1) \end{bmatrix} .
+\label{eqn:matrixK} \end{equation}
 
- The uncertainty assessment is compulsory in metrology
+The underlying system of equations 
+\begin{equation} \widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \mathbf{x} \label{eqn:ddsiemnd} \end{equation}
+is another representation of the minimization problem
+\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert  \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x} \right\Vert^2_2 .
+\label{eqn:min_seiv} \end{equation}
+where $\widetilde{\mathbf{y}} = \begin{bmatrix} \widetilde{y}(n+1) & \ldots & \widetilde{y}(T) \end{bmatrix}^\top$, and
+$\widehat{\mathbf{x}} = \begin{bmatrix} \widehat{\mathbf{u}} & \widehat{\bm{\ell}}^\top \end{bmatrix}^\top$. 
+This minimization problem is an errors-in-variables (EIV) problem with Hankel structure, and correlation between the regression matrix $\widetilde{\mathbf{K}}$ and the regressor vector $\widetilde{\mathbf{y}}$.
+
+The data-driven step input estimation method converts the output-error simultaneous model identification and input estimation problem into an errors-in-variables (EIV) input estimation problem.
+The cost of avoiding the parametric sensor modeling is to deal with a more difficult stochastic framework.
+
+For metrology applications, it is desired to have a fast online solution of the problem(\ref{eqn:ddsiemnd}).
+The least-squares (LS) approximate solution of this system of equations offers a simple alternative, in its recursive form, to implement the estimation method in real-time.
+The least squares admit a recursive implementation that avoids the inversion of matrix $\widetilde{\mathbf{K}}$ that increases its size with respect to the sample size $T$.
+Instead, the recursive least-squares (RLS) updates the solution of the system of equations considering the previous value of the estimation.
+
+The LS solution is used even though it might exhibit bias due to the correlation of the perturbation in $\widetilde{\mathbf{K}}$ with the perturbation in $\widetilde{\mathbf{y}}$.
+To conduct the statistical analysis of the step input estimation method, it is more convenient to use the standard least-squares terminology.
+The statistical analysis results obtained from LS treatment are fully compatible with the RLS estimation results.
+The well known bias and covariance results for the LS estimator cannot be invoked because they assume that the additive perturbation only affects the regressor, and that there is no correlation between the regressor and the regression matrix.
+
+\begin{comment}
 
@@
 Persistency of excitation of the input is a necessary identifiability condition in exact system identification.
-
-**** Rik: persistence of excitations stems from identification theory, not system theory
+\end{comment}
 
 \begin{comment}
-\section{Step input estimation method - Ramp input}
-
-This section describes the step input estimation method and presents a formulation of the affine input estimation problem.
-The step input estimation method does not require a sensor model to estimate the unknown level of the step input by processing the sensor step response.
-The affine input estimation problem is formulated as a signal processing estimation problem where the parameters of the affine input can be obtained from the sensor response.
-
 \subsection{Step input estimation method}
@@
 
-\section{Step input estimation method - Statistical analysis}
+\subsection{Step input estimation method - Statistical analysis}
 
@@
 
-\section{Step input estimation method - Experimental validation}
+\subsection{Step input estimation method - Experimental validation}
 
@@
 
+\subsection{Step input estimation method - Ramp input}
+
+This section describes the step input estimation method and presents a formulation of the affine input estimation problem.
+The step input estimation method does not require a sensor model to estimate the unknown level of the step input by processing the sensor step response.
+The affine input estimation problem is formulated as a signal processing estimation problem where the parameters of the affine input can be obtained from the sensor response.
+
+\subsection{Step input estimation method}
+The step input estimation method estimates the unknown value $\widebar{u} \in {\rm I\!R}^{}$, that is the level of a step input $u = \widebar{u} s$, where $s$ is the unit step function.
+The step input $u$ is applied to a stable linear time-invariant sensor of order $n$ and static gain $\gamma \in {\rm I\!R}^{}$.
+The input estimate $\widehat{u}$ is obtained by processing the measured sequence of output observations $\big( {y}(0), \ldots, {y}(T) \big)$, where ${y}(t) \in {\rm I\!R}^{}$ for $t = 1,\ldots,T$, where $T$ is the sample size, and
+\begin{equation} \mathbf{y} = \widebar{\mathbf{y}} + \bm{\epsilon} . \label{eqn:y0plusnoise} \end{equation} 
+The exact sensor response $\widebar{\mathbf{y}}$ is perturbed by additive measurement noise $\bm{\epsilon}$.
+The measurement noise $\epsilon$ is assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.
+
+The measured sensor response is a zero-order hold discretization of the continuous-time response.
+The discrete-time sensor response is considered to be piecewise constant.
+
+The step input can be estimated by solving the minimization problem
+\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert  \mathbf{y} - \mathbf{K} \mathbf{x} \right\Vert^2_2 . \label{eqn:min_ls} \end{equation}
+where $\mathbf{y} = \begin{bmatrix} {y}(n+1) & \ldots & {y}(T) \end{bmatrix}^\top$, 
+$\widehat{\mathbf{x}} = \begin{bmatrix} \widehat{u} & \widehat{\bm{\ell}}^\top \end{bmatrix}^\top$, is a vector whose first element is the input estimation $\widehat{u}$ and the $n\text{-vector} \ \widehat{\bm{\ell}}$ is a linear transformation of the sensor initial conditions.
+The matrix
+\begin{equation} \mathbf{K} = \begin{bmatrix} \gamma & \Delta {y}(1) & \Delta {y}(2) & \cdots & \Delta {y}(n) \\ \gamma & \Delta {y}(2) & \Delta {y}(3) & \cdots & \Delta {y}(n+1) \\ \vdots & \iddots & \iddots & \iddots & \\ \gamma & \Delta {y}(T-n) & \Delta {y}(T-n+1) & \cdots & \Delta {y}(T-1) \label{eqn:matrix} \end{bmatrix} \end{equation}
+is a Hankel matrix of $(T - n)-\text{block rows}$, constructed from the consecutive differences $\Delta {y}(t) = {y}(t) - {y}(t-1)$ of the measured transient response, augmented in the left with a $(T - n)\text{-vector}$ of elements equal to the sensor static gain $\gamma$.
+
+The minimization problem (\ref{eqn:min_ls}) is a structured errors-in-variables (EIV) problem.
+The structure of the EIV problem is due to the presence of the Hankel matrix in $\mathbf{K}$.
+The measurement noise $\bm{\epsilon}$ enters in the regression matrix $\mathbf{K}$ and we can express 
+\begin{equation} \mathbf{K} = \widebar{\mathbf{K}} + \mathbf{E}, \label{eqn:K0plusnoise} \end{equation} 
+where $\widebar{\mathbf{K}}$ is exact data information and $\mathbf{E}$ is the corresponding perturbation noise.
+Details of the method formulation are described in \cite{Markovsky15cep}.
+
+The recursive least squares (RLS) algorithm provides a solution to the structured EIV problem and enables real-time implementations of the estimation method.
+The RLS has an exponential forgetting factor that selects the observations of the transient response to perform the estimation.
+With the exponential  forgetting, the subspace estimation method tracks the evolution of any time-varying applied input.
+The affine input is one case in which the input evolves proportionally to time.
+
+The weighing system admits a state space representation where the states $x_1=y$ and $x_2=\dot{y}$ are the position and the speed of the weighing scale: 
+\[ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{\widebar{a} t + \widebar{b} + m} & \dfrac{-(k_{\mathrm{d}} + \widebar{a})}{\widebar{a} t + \widebar{b} + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 \\ g \end{bmatrix},  \quad y = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .
+\]
+
+\begin{figure}[htb!]
+\centering
+
+\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
+\tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
+\tikzstyle{damper}=[thick,decoration={markings,  
+  mark connection node=dmp,
+  mark=at position 0.5 with 
+  {
+    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
+    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
+    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
+  }
+}, decorate]
+\tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]
+
+\node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
+\node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u(t)$};
+
+\node (ground1) at (M.south) [ground,yshift=-1.5cm,xshift=-0.625cm,anchor=north] {};
+\draw (ground1.north west) -- (ground1.north east);
+\draw [spring] (ground1.north) -- ($(M.south east)!(ground1.north)!(M.south west)$);
+
+\node (groundc) at (M.south) [ground,yshift=-1.5cm,anchor=north] {}; 
+\draw (groundc.north west) -- (groundc.north east);
+
+\node (ground2) at (M.south) [ground,yshift=-1.5cm,xshift=0.625cm,anchor=north] {};
+\draw (ground2.north west) -- (ground2.north east);
+\draw [damper] (ground2.north) -- ($(M.south east)!(ground2.north)!(M.south west)$);
+
+\node[draw=none,fill=none] at (-0.9cm,-1cm) {$k_{\mathrm{s}}$};
+\node[draw=none,fill=none] at (0.15cm,-1cm) {$k_{\mathrm{d}}$};
+\node[draw=none,fill=none] at (2.0cm,1.0cm) {$y$};
+\draw [-latex,thick]  ++(2.2cm,-1cm) -- +(0cm,2.25cm);
+
+\draw [-latex,thick] (M.east) ++(0,0) -- +(1cm,0);
+\draw [line width=0.25mm] (2.2cm,-1cm) -- (2.2cm,1cm);
+\draw [line width=0.25mm] (2.1cm,-1cm) -- (2.3cm,-1cm);
+\draw [line width=0.25mm] (2.1cm,1cm) -- (2.3cm,1cm);
+\draw [line width=0.25mm] (2.1cm,-0.5cm) -- (2.3cm,-0.5cm);
+\draw [line width=0.25mm] (2.1cm,0.5cm) -- (2.3cm,0.5cm);
+\draw [line width=0.25mm] (2.15cm,-0.25cm) -- (2.25cm,-0.25cm);
+\draw [line width=0.25mm] (2.15cm,0.25cm) -- (2.25cm,0.25cm);
+\draw [line width=0.25mm] (2.15cm,-0.75cm) -- (2.25cm,-0.75cm);
+\draw [line width=0.25mm] (2.15cm,0.75cm) -- (2.25cm,0.75cm);
+\draw [line width=0.25mm] (2.1cm,0cm) -- (2.3cm,0cm);
+
+\end{tikzpicture}
+
+\caption{\label{fig:msd_system} A second order mass-spring-damper model represents the dynamic weighing system.
+The dynamics of the system depend on the affine input.
+The weighing system is time-varying when the applied input changes with respect to time.}
+\end{figure}
+
+In this paper we use the dynamic weighing example to illustrate the implementation of the affine input estimation methods.
+
 \end{comment}

### 2020-01-30 ChapterRampInput/ChapterRampInput.tex
Preliminaries ready

@@
 
-\chapter{Ramp input estimation method}\label{chap:RampInput}
+\chapter{Affine input estimation method}\label{chap:AffineInput}
 
 \begin{quote}
-\emph{The results of this chapter ...}\vfill{}
+\emph{The results of this chapter were published in Quintana Carapia, G., and Markovsky, I.,
+"Input parameters estimation from time-varying measurements", Measurement Journal, Vol. 153, 2020,
+ISSN 0263-2241, doi:10.1016/j.measurement.2019.107418. \nocite{QuintanaMEAS} }\vfill{}
 \end{quote}
 
-\section{Introduction}
+\vfill{}
+
+%\section{Introduction}
 
@@
 
-The affine input is modeled as a straight line $u = {a} t + {b}$ with parameters the slope $a$ and the interception $b$.
+The affine input is modeled as a straight line $u(t) = {a} t + {b}$ with parameters the slope $a$ and the interception $b$.
 The affine input estimation problem is formulated as a signal processing problem as follows.
@@
 \paragraph{Problem} 
-Given the sequence of measured output observations $\mathbf{y} = \left( y(1),\ldots,y(T)\right), y(t) \in {\rm I\!R}$, of a stable linear time-invariant system of order $n$, and static gain $\gamma$, generated by an affine input $u = \widebar{a} t + \widebar{b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u} = \widehat{a} t + \widehat{b}$ approximates $u$.
-The measured observations $\mathbf{y} = \widebar{\mathbf{y}} + \bm{\epsilon}$ are exact sensor responses $\widebar{\mathbf{y}}$ perturbed by additive noise  $\bm{\epsilon}$ assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.
+Given the sequence of measured output observations $\widetilde{\mathbf{y}} = \left( \widetilde{y}(1),\ldots, \widetilde{y}(T)\right)$, with $\widetilde{y}(t) \in {\rm I\!R}$, of a stable linear time-invariant system of order $n$, and static gain $\gamma$, generated by an affine input $u(t) = \widebar{a} t + \widebar{b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u}(t) = \widehat{a} t + \widehat{b}$ approximates $u$(t).
+The measured observations $\widetilde{\mathbf{y}} = \widebar{\mathbf{y}} + \bm{\epsilon}$ are exact sensor responses $\widebar{\mathbf{y}}$ perturbed by additive noise  $\bm{\epsilon}$ assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.
 
@@
 
-RLS is a special case of the EWRLS that can solve the minimization problem (\ref{eqn:min_ls}).
+RLS is a special case of the EWRLS that can solve the errors-in-variables (EIV) minimization problem (\ref{eqn:min_seiv}).
 The minimization problem is a structured EIV problem and its EWRLS solution is equivalent to 
-\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert \mathbf{W}^{1/2} \left( \mathbf{y} - \mathbf{K} \mathbf{x} \right) \right\Vert^2_2 .
+\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert \mathbf{W}^{1/2} \left( \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x} \right) \right\Vert^2_2 .
 \label{eqn:min_ewrls} \end{equation}
-where $\mathbf{y}$, $\widehat{\mathbf{x}}$, and $\mathbf{K}$ are defined in (\ref{eqn:min_ls}) and (\ref{eqn:matrix}) and
-$\mathbf{W} \in \mathbb{R}^{(T-n) \times (T-n)}$ is a diagonal matrix of descending powers of the weight $\lambda \in [0, 1)$, \textit{i.e.}, $W = \mathrm{diag}\left(\lambda^{T-n}, \ldots, \lambda^2, \lambda^1 \right)$.
-The weight $\lambda$ is a data selection forgetting factor since it enables to apply different weights to the residuals $\mathbf{y} - \mathbf{K} \mathbf{x}$.
+where $\widetilde{\mathbf{y}}$, $\widehat{\mathbf{x}}$, and $\widetilde{\mathbf{K}}$ are defined in (\ref{eqn:min_seiv}) and (\ref{eqn:K0plusnoise}) and
+$\mathbf{W} \in \mathbb{R}^{(T-n) \times (T-n)}$ is a diagonal matrix of descending powers of the weight $\lambda \in [0, 1)$, \textit{i.e.}, $\mathbf{W} = \mathrm{diag}\left(\lambda^{T-n}, \ldots, \lambda^2, \lambda^1 \right)$.
+The weight $\lambda$ is a data selection forgetting factor since it enables to apply different weights to the residuals $\widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x}$.
 When $\lambda=1$, we have the same solution as RLS.
@@
 
-The profile of an affine input excitation can be reconstructed by solving the structured errors-in-variables problem  (\ref{eqn:min_ewrls}) where $y$ is the corresponding response of a stable dynamic system, $\mathbf{K}$ is constructed from the output observations, and $\lambda<1$.
+The profile of an affine input excitation can be reconstructed by solving the structured errors-in-variables problem  (\ref{eqn:min_ewrls}) where $y$ is the corresponding response of a stable dynamic system, $\widetilde{\mathbf{K}}$ is constructed from the output observations, and $\lambda<1$.
 
-After the affine input $\widehat{u}$ has been estimated, the affine input parameters $a$ and $b$ are estimated by fitting $\widehat{u}$ to the straight line $\widehat{u} = a t + b$ using linear regression, as follows
+After the affine input $\widehat{\mathbf{u}}$ has been estimated, the affine input parameters $a$ and $b$ are estimated by fitting $\widehat{\mathbf{u}}$ to a straight line using linear regression, as follows
 \begin{equation} 
@@
 \end{equation}
-where $\tau$ is a number of samples used as a tuning parameter that counteracts the time delay that exists when a LTI system is excited with an affine input $u$.
+where $\tau$ is a number of samples used as a tuning parameter that counteracts the time delay that exists when a LTI system is excited with an affine input $\mathbf{u}$.
 The subspace method can process online the sensor response to the affine excitation.
-For each new observation $y(t)$, the estimation $\widehat{u}$ is updated followed by the update of the slope $\widehat{a}$ and the interception $\widehat{b}$.
+For each new observation $y(t)$, the estimation $\widehat{\mathbf{u}}$ is updated followed by the update of the slope $\widehat{a}$ and the interception $\widehat{b}$.
 The values of the tuning parameters $\lambda$ and $\tau$ can be obtained in the calibration of the method using the response of the sensor.
@@
 
+\subsubsection{Statistical analysis of the subspace method}
+
+To obtain the first and second moments of the step input estimate $\widehat{\mathbf{u}}$, we need to study the solution
+\begin{equation} \widehat{\mathbf{x}} = ( \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{K}}  )^{-1} \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{y}} , \label{eqn:xhatW} \end{equation}
+of the overdetermined structured errors-in-variables EIV problem (\ref{eqn:min_ewrls}).
+Using a second order Taylor series expansion of the inverse matrix we can approximate the LS solution as
+\begin{equation} \widehat{\mathbf{x}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{C}^{-1} (\mathbf{K} + \mathbf{E})^\top \mathbf{W} (\mathbf{y} + \bm{\epsilon}).
+\label{eqn:xhatexpW} \end{equation}
+where 
+\begin{equation} \mathbf{C} = \mathbf{K}^\top \mathbf{W} \mathbf{K}, \quad \text{and} \quad \mathbf{M} = \mathbf{C}^{-1} ( \mathbf{K}^\top \mathbf{W} \mathbf{E} + \mathbf{E}^\top \mathbf{W} \mathbf{K} + \mathbf{E}^\top \mathbf{W} \mathbf{E} ).
+\end{equation}
+
+The Taylor series approximation of $\widehat{\mathbf{x}}$ enables the calculation of the estimate bias and covariance since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion.
+The bias and the covariance of the estimate $\widehat{\mathbf{x}}$ are obtained from the definitions (\ref{eqn:biasdef}), and (\ref{eqn:covdef}).
+%The bias and the covariance of the estimate $\widehat{\mathbf{x}}$ are obtained from \begin{equation}  \mathbf{b} \left(\widehat{\mathbf{x}} \right) = \mathbf{\mu} - \widebar{\mathbf{x}}, \label{eqn:biasdef} \end{equation}  \begin{equation} \begin{aligned} \mathrm{\mathbf{Cov}} \left( \widehat{\mathbf{x}} \right) & = \mathbb{E} \left\{ \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)  \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)^\top \right\} .
+\end{aligned} \label{eqn:covdef} \end{equation} where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}$ is the expected value of the estimate and $\widebar{\mathbf{x}}$ is the true value.
+Considering the structure of the EIV problem, the bias and the covariance of the estimate approximation (\ref{eqn:xhatexpW}) can be expressed as
+\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{C}^{-1} \left(  \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_1 - \mathbf{B}_2 \right) \mathbf{x} - \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasEW} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{K}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{W} \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{x} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varEW} \end{equation}
+where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \Big\}$, $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger \mathbf{W}$, and $\mathbf{K}^\dagger$ is the pseudo-inverse matrix of $\mathbf{K}$.
+
+The bias and covariance given by expressions (\ref{eqn:biasEW}) and (\ref{eqn:varEW}) depend on the unobservable true values $\mathbf{x}$ and $\mathbf{K}$.
+The measured observations are in the sensor step response $\widetilde{\mathbf{y}}$, and from its observations we construct $\widetilde{\mathbf{K}}$ and compute $\widehat{\mathbf{x}}$.
+The substitution of the measured data in the expressions gives an approximation of the estimate bias and covariance. 
+We have then
+\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{C}}^{-1} \left(  \left( \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\mathbf{x}} - \left( \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasSTW} \end{equation}
+  \begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \mathbf{W} \widetilde{\mathbf{K}}^{\dagger \top} - \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{x} \right) \widetilde{\mathbf{b}}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varSTW} \end{equation}
+  where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \bm{\epsilon}^\top \Big\}$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger \mathbf{W}$.
+
+The results of the expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$ can be found using Lemma \ref{lem:lemma1}.
+The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \citep{Vaccaro94, Stewart90SPT}.
+It was shown that the bias and variance expressions (\ref{eqn:biasSTW}) and (\ref{eqn:varSTW}) are valid predictions of the first and second moments of the LS estimate of a Hankel structured EIV problem.
+The problem formulated by the step input estimation method belongs to this type of structured EIV problems and can use the derived expressions to find the bias and variance of the input estimate $\widehat{u}$.
+The bias of the estimate $\widehat{\mathbf{u}}$ is the fist element of $\widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right)$ and the variance of $\widehat{\mathbf{u}}$ is the first element in the main diagonal of $\widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right)$.
+
 \subsection{Maximum-likelihood method}
 Using a model of the sensor, the ML method estimates the sensor initial conditions and the parameters of the applied affine input.
-The affine input parameters $a$ and $b$ are the slope and the interception of the straight line model $u = at + b$.
+The affine input parameters $a$ and $b$ are the slope and the interception of the straight line model $u(t) = at + b$.
 
 Algorithm \ref{Alg1} lists the steps of the proposed ML method.
-This is an iterative minimization method that uses the Jacobian of the residual error function $\mathbf{r} = \widehat{\mathbf{y}} - \mathbf{y}$ to search the direction that minimizes the difference between the measured sensor response $\mathbf{y}$ and the simulated response $\widehat{\mathbf{y}}$.
+This is an iterative minimization method that uses the Jacobian of the residual error function $\mathbf{r} = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}$ to search the direction that minimizes the difference between the measured sensor response $\widetilde{\mathbf{y}}$ and the simulated response $\widehat{\mathbf{y}}$.
 
@@
 \begin{algorithmic}
-  \REQUIRE{$\mathbf{y}$, and sensor model parameters}{}
+  \REQUIRE{$\widetilde{\mathbf{y}}$, and sensor model parameters}{}
   \STATE {Initialize $\mathbf{\theta} = (a, b, \mathbf{x}_{\text{ini}})$}{}
-  \FOR{each $N$ observations of $\mathbf{y}$}{}
+  \FOR{each $N$ observations of $\widetilde{\mathbf{y}}$}{}
   \STATE {Simulate model response $\widehat{\mathbf{y}}(\mathbf{\theta})$}{}
-  \STATE {Compute error $\mathbf{r}(\mathbf{\theta}) = \mathbf{y} - \widehat{\mathbf{y}}(\mathbf{\theta})$}{}
+  \STATE {Compute error $\mathbf{r}(\mathbf{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\mathbf{\theta})$}{}
   \STATE {Minimize $\mathbf{r}^\top \mathbf{r}$ over $\mathbf{\theta}$}{}
@@
 The optimization variables updates are computed every $N$ new observations.
-The minimization can be done using the Levenberg-Marquardt algorithm \cite{Nocedal06}.
+The minimization can be done using the Levenberg-Marquardt algorithm \citep{Nocedal06}.
 
@@
 The analytic formulation of the Jacobian benefits the estimation method in two ways: it speeds up the minimization and gives direct access to the variance of the estimates.
-The covariance matrix of the ML estimates can be expressed as \cite{Pintelon12Book}
+The covariance matrix of the ML estimates can be expressed as \citep{Pintelon12Book}
 \begin{equation} \mathbf{Cov} \left( \widehat{x} \right) = \left( \left( \dfrac{\partial \mathbf{r} }{ \partial \mathbf{\theta} } \right)^\top \left( \dfrac{\partial \mathbf{r} }{ \partial \mathbf{\theta} } \right) \right)^{-1}.
@@
  \begin{equation} \begin{aligned}
-     & \text{Minimize} \quad \text{over} \ a, b, \mathbf{x}_{\text{ini}} \quad \mathbf{r}^T \mathbf{r} \text{, subject to:} \\ & \ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 & \frac{x_{\text{ini,1}}}{a t + b + m} \\ g & \frac{x_{\text{ini,2}}}{a t + b + m}  \end{bmatrix} \begin{bmatrix} 1 \\ \delta(t)  \end{bmatrix}, \\ & \ \widehat{y} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .
+     & \text{Minimize} \quad \text{over} \ a, b, \mathbf{x}_{\text{ini}} \quad \mathbf{r}^T \mathbf{r} \text{, subject to:} \\ & \ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 & \frac{x_{\text{ini,1}}}{a t + b + m} \\ g & \frac{x_{\text{ini,2}}}{a t + b + m}  \end{bmatrix} \begin{bmatrix} 1 \\ \delta(t)  \end{bmatrix}, \\ & \ \widehat{\mathbf{y}} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .
  \end{aligned} \end{equation}
-where $a$, $b$ are the affine input parameters, $m$, $k_{\mathrm{d}}$, and $k_{\mathrm{s}}$ are the model parameters, the model states $\mathbf{x}$ are the position and the speed of the weighing scale, the difference between the sensor response $\mathbf{y}$ and the simulated response $\widehat{\mathbf{y}}$ is the residual $\mathbf{r} = \widehat{\mathbf{y}} - \mathbf{y}$.
+where $a$, $b$ are the affine input parameters, $m$, $k_{\mathrm{d}}$, and $k_{\mathrm{s}}$ are the model parameters, the model states $\mathbf{x}$ are the position and the speed of the weighing scale, the difference between the sensor response $\widetilde{\mathbf{y}}$ and the simulated response $\widehat{\mathbf{y}}$ is the residual $\mathbf{r} = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}$.
 The model initial conditions $\mathbf{x}_{\text{ini}}$ are considered optimization variables
@@
 \subsection{Time-varying compensation filter}
-The time-varying (TV) filter described in \cite{Pietrzak14} was designed to compensate the measured responses of a conveyor weighing system, considering they are modeled as a saturated ramp.
+The time-varying (TV) filter described in \citep{Pietrzak14} was designed to compensate the measured responses of a conveyor weighing system, considering they are modeled as a saturated ramp.
 The TV filter consists of three low-pass infinite impulse response (IIR) filters in cascade, where the $i-\mathrm{th}$ IIR filter is given by
@@
 for $i = 1,\ldots,3$ and $t=0,\ldots,T$.
-The sensor response is fed to the filter, then $\widehat{y}_0(t) = y(t)$, and the output of the TV filter $\widehat{u}_\mathrm{ltv}(t) = \widehat{y}_3(t)$ is an estimation of the affine input.
+The sensor response is fed to the filter, then $\widehat{y}_0(t) = \widetilde{y}(t)$, and the output of the TV filter $\widehat{u}_\mathrm{ltv}(t) = \widehat{y}_3(t)$ is an estimation of the affine input.
 Since in our case we are processing only the ramp without the saturation, the estimates $\widehat{a}_\mathrm{ltv}$ and $\widehat{b}_\mathrm{ltv}$ of the input parameters are obtained by fitting a straight line to the estimated input $\widehat{u}_\mathrm{ltv}$ using linear regression.
@@
 where $\mu_{\widetilde{a}_\mathrm{ltv}}$, $\mu_{\widetilde{b}_\mathrm{ltv}}$, $\sigma_{\widetilde{a}_\mathrm{ltv}}$ and $\sigma_{\widetilde{b}_\mathrm{ltv}}$ are the mean values and the standard deviations of the estimation errors $\widetilde{a}_\mathrm{ltv} = \widehat{a}_\mathrm{ltv} - \widebar{a}$, and $\widetilde{a}_\mathrm{ltv} = \widehat{b}_\mathrm{ltv} - \widebar{b}$, and where $\widebar{a}$ and $\widebar{b}$ are the true values of the input parameters.
-The values $\mu_{\mathrm{spec}}$ and $\sigma_{\mathrm{spec}}$ are specified in the OIML recommendation R51 \cite{OIML_R51_1} for mass measurements that use a conveyor belt.
+The values $\mu_{\mathrm{spec}}$ and $\sigma_{\mathrm{spec}}$ are specified in the OIML recommendation R51 \citep{OIML_R51_1} for mass measurements that use a conveyor belt.
 
@@
 The parameters of the weighing system are $m = 15$ g, $d = 5.5$ Ns/m and $k = 10250$ N/m.
-The applied affine input $u = 100 t + 10$ represents a mass that changes from $10$ g to $110$ g, at a constant rate, in a time interval of $0.1$ s.
+The applied affine input $u(t) = 100 t + 10$ represents a mass that changes from $10$ g to $110$ g, at a constant rate, in a time interval of $0.1$ s.
 This change of mass represents one example of the weighing input in a conveyor weighing system when an object of 100 g is measured while it is moving at constant speed.
 The sensor response is acquired using sampling time $T_s = 0.1$ ms.
-In Figure \ref{fig:sensor_weight} the input $u$ is represented with the dotted line, the oscillatory curve is the corresponding sensor ramp response $y$, and $\widehat{u}_{\mathrm{}}$ is a typical input estimate obtained with the subspace method.
+In Figure \ref{fig:sensor_weight} the input $u$ is represented with the dotted line, the oscillatory curve is the corresponding sensor ramp response $\widetilde{\mathbf{y}}$, and $\widehat{\mathbf{u}}_{\mathrm{}}$ is a typical input estimate obtained with the subspace method.
 
@@
 \includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_2.pdf} 
-\caption{ \label{fig:sensor_weight} The sensor transient response $y$ to an affine input excitation $u = at+b$ is processed by the estimation methods to estimate the parameters $a$ and $b$.
-In the figure we observe an example of the input estimate $\widehat{u}$ obtained with the subspace method.
-The input parameters are calculated from $\widehat{u}$ using linear regression.}
+\caption{ \label{fig:sensor_weight} The sensor transient response $\widetilde{\mathbf{y}}$ to an affine input excitation $u(t) = at+b$ is processed by the estimation methods to estimate the parameters $a$ and $b$.
+In the figure we observe an example of the input estimate $\widehat{\mathbf{u}}$ obtained with the subspace method.
+The input parameters are calculated from $\widehat{\mathbf{u}}$ using linear regression.}
 \end{figure}
@@
 
-\begin{tabular}{L|L L L L} 
+\begin{tabular}{l|l l l l} 
  \hline
- & \mathrm{Mass} & \\ [0.5ex] 
- \mathrm{Time} & 0.1 \ kg & 0.3 \ kg & 0.5 \ kg & 1.0 \ kg \\ 
+ & Mass & \\ [0.5ex] 
+ Time & 0.1 \ kg & 0.3 \ kg & 0.5 \ kg & 1.0 \ kg \\ 
  \hline
 \multicolumn{1}{l|}{0.05 \ s} & \\
- \multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{a}} \ [g^2]} & 3.5\mathrm{x}10^{0}_{(@40 \ \mathrm{dB})} & 13.5\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})} & 3.8\mathrm{x}10^{0}_{(@20 \ \mathrm{dB})} & 8.3\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})} \\
-\multicolumn{1}{r|}{\hspace{1mm}\mathrm{MSE}_{\hat{b}} \ [g^2]} &  3.1\mathrm{x}10^{-3}_{(@40 \ \mathrm{dB})}  & 1.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} & 1.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} & 1.9\mathrm{x}10^{-2}_{(@60 \ \mathrm{dB})} \\
+ \multicolumn{1}{r|}{\hspace{1mm} $\mathrm{MSE}_{\hat{a}} \ \mathrm{[g^2]}$} & $3.5\mathrm{x}10^{0}_{(@40 \ \mathrm{dB})}$ & $13.5\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})}$ & $3.8\mathrm{x}10^{0}_{(@20 \ \mathrm{dB})}$ & $8.3\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})}$ \\
+\multicolumn{1}{r|}{\hspace{1mm} $\mathrm{MSE}_{\hat{b}} \ \mathrm{[g^2]}$} & $3.1\mathrm{x}10^{-3}_{(@40 \ \mathrm{dB})}$ & $1.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})}$ & $1.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})}$ & $1.9\mathrm{x}10^{-2}_{(@60 \ \mathrm{dB})}$ \\
 \multicolumn{1}{l|}{0.1 \ s} & \\
- \multicolumn{1}{r|}{\hspace{1mm}\mathrm{MSE}_{\hat{a}} \ [g^2]} & 3.1\mathrm{x}10^{-2}_{(@30 \ \mathrm{dB})} & 1.1\mathrm{x}10^{0}_{(@40 \ \mathrm{dB})} & 8.3\mathrm{x}10^{-1}_{(@60 \ \mathrm{dB})} & 1.3\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})} \\
-\multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{b}} \ [g^2]} & 6.0\mathrm{x}10^{-5}_{(@30 \ \mathrm{dB})} & 1.2\mathrm{x}10^{-1}_{(@50 \ \mathrm{dB})} & 2.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} & 3.7\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} \\
+ \multicolumn{1}{r|}{\hspace{1mm} $\mathrm{MSE}_{\hat{a}} \ \mathrm{[g^2]}$} & $3.1\mathrm{x}10^{-2}_{(@30 \ \mathrm{dB})}$ & $1.1\mathrm{x}10^{0}_{(@40 \ \mathrm{dB})}$ & $8.3\mathrm{x}10^{-1}_{(@60 \ \mathrm{dB})}$ & $1.3\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})}$ \\
+\multicolumn{1}{r|}{\hspace{1mm} $\mathrm{MSE}_{\hat{b}} \ \mathrm{[g^2]}$} & $6.0\mathrm{x}10^{-5}_{(@30 \ \mathrm{dB})}$ & $1.2\mathrm{x}10^{-1}_{(@50 \ \mathrm{dB})}$ & $2.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})}$ & $3.7\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})}$ \\
 \multicolumn{1}{l|}{0.5 \ s} & \\
- \multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{a}} \ [g^2]} & 3.0\mathrm{x}10^{-2}_{(@20 \ \mathrm{dB})} & 3.0\mathrm{x}10^{-1}_{(@20 \ \mathrm{dB})} & 3.2\mathrm{x}10^{0}_{(@20 \ \mathrm{dB})} & 8.6\mathrm{x}10^{-2}_{(@20 \ \mathrm{dB})} \\
-\multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{b}} \ [g^2]} & 3.1\mathrm{x}10^{-5}_{(@50 \ \mathrm{dB})} & 9.8\mathrm{x}10^{-4}_{(@40 \ \mathrm{dB})} & 2.0\mathrm{x}10^{-5}_{(@50 \ \mathrm{dB})} & 1.7\mathrm{x}10^{-5}_{(@20 \ \mathrm{dB})} \\ [0.5ex] % [\mathrm{samples}]
+ \multicolumn{1}{r|}{\hspace{1mm} $\mathrm{MSE}_{\hat{a}} \ \mathrm{[g^2]}$} & $3.0\mathrm{x}10^{-2}_{(@20 \ \mathrm{dB})}$ & $3.0\mathrm{x}10^{-1}_{(@20 \ \mathrm{dB})}$ & $3.2\mathrm{x}10^{0}_{(@20 \ \mathrm{dB})}$ & $8.6\mathrm{x}10^{-2}_{(@20 \ \mathrm{dB})}$ \\
+\multicolumn{1}{r|}{\hspace{1mm} $\mathrm{MSE}_{\hat{b}} \ \mathrm{[g^2]}$} & $3.1\mathrm{x}10^{-5}_{(@50 \ \mathrm{dB})}$ & $9.8\mathrm{x}10^{-4}_{(@40 \ \mathrm{dB})}$ & $2.0\mathrm{x}10^{-5}_{(@50 \ \mathrm{dB})}$ & $1.7\mathrm{x}10^{-5}_{(@20 \ \mathrm{dB})}$ \\ [0.5ex] % [\mathrm{samples}]
  \hline
 \end{tabular}
-\
 \label{table:differentmasses}
@@
 
-\begin{tabular}{C| C C C} 
+\begin{tabular}{c| c c c} 
 \hline  
-& \sigma_{\mathrm{s}} \\ [0.5ex] 
-\sigma_{{a},{b}}  & 0\% & 5\% & 10\% \\  
+$\sigma_{{a},{b}}$ & $\sigma_{\mathrm{s}}$: \ 0\% & 5\% & 10\% \\ [0.5ex] 
  \hline
  \multicolumn{1}{l|}{0\%} \\
- \multicolumn{1}{l|}{\hspace{2mm} \widehat{a} \ [\mathrm{kg/s}]} & 1.0 \pm 9.6 \%_{(@50 \ \mathrm{dB})} &  1.0 \pm 8.2 \%_{(@20 \ \mathrm{dB})} & 1.0 \pm 8.2 \%_{(@50 \ \mathrm{dB})} \\ 
- \multicolumn{1}{l|}{\hspace{2mm} \widehat{b} \ \mathrm{[g]}} & 10.0 \pm 10.3 \%_{(@20 \ \mathrm{dB})} & 9.9 \pm 14.2 \%_{(@30 \ \mathrm{dB})} &  10.0 \pm 21.7 \%_{(@60 \ \mathrm{dB})} \\ 
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{a} \ \mathrm{[kg/s]}$} & $1.0 \pm 9.6 \%_{(@50 \ \mathrm{dB})}$ &  $1.0 \pm 8.2 \%_{(@20 \ \mathrm{dB})}$ & $1.0 \pm 8.2 \%_{(@50 \ \mathrm{dB})}$ \\ 
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{b} \ \mathrm{[g]}$} & $10.0 \pm 10.3 \%_{(@20 \ \mathrm{dB})}$ & $9.9 \pm 14.2 \%_{(@30 \ \mathrm{dB})}$ & $10.0 \pm 21.7 \%_{(@60 \ \mathrm{dB})}$ \\ 
  \multicolumn{1}{l|}{5\%} \\
- \multicolumn{1}{l|}{\hspace{2mm} \widehat{a} \ [\mathrm{kg/s}]} & 1.0 \pm 9.7 \%_{(@50 \ \mathrm{dB})} & 1.0 \pm 11.3 \%_{(@20 \ \mathrm{dB})} & 1.0 \pm 8.0 \%_{(@50 \ \mathrm{dB})} \\  
- \multicolumn{1}{l|}{\hspace{2mm} \widehat{b} \ \mathrm{[g]}} & 10.0 \pm 22.4 \%_{(@50 \ \mathrm{dB})} & 10.0 \pm 33.2 \%_{(@30 \ \mathrm{dB})} & 9.9 \pm 16.6 \%_{(@50 \ \mathrm{dB})} \\ 
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{a} \ \mathrm{[kg/s]}$} & $1.0 \pm 9.7 \%_{(@50 \ \mathrm{dB})}$ & $1.0 \pm 11.3 \%_{(@20 \ \mathrm{dB})}$ & $1.0 \pm 8.0 \%_{(@50 \ \mathrm{dB})}$ \\  
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{b} \ \mathrm{[g]}$} & $10.0 \pm 22.4 \%_{(@50 \ \mathrm{dB})}$ & $10.0 \pm 33.2 \%_{(@30 \ \mathrm{dB})}$ & $9.9 \pm 16.6 \%_{(@50 \ \mathrm{dB})}$ \\ 
 \multicolumn{1}{l|}{10\%} \\
-\multicolumn{1}{l|}{\hspace{2mm} \widehat{a} \ [\mathrm{kg/s}]} & 1.0 \pm 20.3 \%_{(@40 \ \mathrm{dB})} & 1.0 \pm 22.0 \%_{(@50 \ \mathrm{dB})} & 1.0 \pm 17.7 \%_{(@40 \ \mathrm{dB})} \\    
- \multicolumn{1}{l|}{\hspace{2mm} \widehat{b} \ \mathrm{[g]}} & 9.9 \pm 46.8 \%_{(@40 \ \mathrm{dB})} & 9.8 \pm 58.8 \%_{(@50 \ \mathrm{dB})} & 9.9 \pm 41.6 \%_{(@40 \ \mathrm{dB})} \\ [0.5ex] 
+\multicolumn{1}{l|}{\hspace{2mm} $\widehat{a} \ \mathrm{[kg/s]}$} & $1.0 \pm 20.3 \%_{(@40 \ \mathrm{dB})}$ & $1.0 \pm 22.0 \%_{(@50 \ \mathrm{dB})}$ & $1.0 \pm 17.7 \%_{(@40 \ \mathrm{dB})}$ \\    
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{b} \ \mathrm{[g]}$} & $9.9 \pm 46.8 \%_{(@40 \ \mathrm{dB})}$ & $9.8 \pm 58.8 \%_{(@50 \ \mathrm{dB})}$ & $9.9 \pm 41.6 \%_{(@40 \ \mathrm{dB})}$ \\ [0.5ex]
 \hline
 \end{tabular}
-\
 \label{table:dd_sensitivity}
@@
 The perturbation uncertainty was simulated by adding normally distributed random noise with standard deviation equal to 0\%, 5\% and 10\% of the corresponding true values of the perturbed parameters.
-A Monte Carlo simulation with $10^3$ runs was conducted for each SNR in the SNR innterval of interest, and in Table \ref{table:ml_sensitivity} are shown the maximum values of the observed estimation uncertainties.
+A Monte Carlo simulation with $10^3$ runs was conducted for each SNR in the SNR interval of interest, and in Table \ref{table:ml_sensitivity} are shown the maximum values of the observed estimation uncertainties.
 The results show that the speed uncertainty $\sigma_{\mathrm{s}}$ has a small impact on the estimation uncertainty.
 On the contrary, the input parameters uncertainties $\sigma_{a,b}$, and the uncertainties of the model parameters $\sigma_{m,d,k}$ cause a large increment in the uncertainty of the estimation.
-
-\begin{comment}
 
@@
 
-\begin{tabular}{|C| C C C C|} 
+\begin{tabular}{c| c c c} 
 \hline  
- &  & \sigma_{\mathrm{s}} \hspace{10mm} 0\% & \hspace{15mm} 5\% & \hspace{14mm} 10\% \\ [0.5ex] 
+$\sigma_{{a},{b}}$ & $\sigma_{\mathrm{s}}$: \ 0\% & 5\% & 10\% \\ [0.5ex] 
 \hline
-\sigma_{{a},{b}} \hspace{5.6mm} 0\% & \widehat{a} \ [\mathrm{kg/s}] & 0.999 \pm 0.7 \% \ (@20 \ \mathrm{dB}) & 1.0 \pm 0.7 \% \ (@20 \ \mathrm{dB}) & 0.999 \pm 0.4 \% \ (@30 \ \mathrm{dB}) \\ 
-& \widehat{b} \ \mathrm{[g]} & \hspace{-1.75mm} 10.0 \pm 1.1 \% \ (@20 \ \mathrm{dB}) & 9.99 \pm 1.1 \% \ (@20 \ \mathrm{dB}) & 10.0 \pm 0.5 \% \ (@30 \ \mathrm{dB}) \\
-& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.6 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) &  0.1 \pm 0.2 \% \ (@30 \ \mathrm{dB}) \\
-& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.1 \pm 122 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 130 \% \ (@40 \ \mathrm{dB}) &  0.09 \pm 38 \% \ (@50 \ \mathrm{dB}) \\
-
-\hspace{11.375mm} 5\% & \widehat{a} \ [\mathrm{kg/s}] & 1.0 \pm 9.0 \% \ (@50 \ \mathrm{dB}) & 0.997 \pm 5.2 \% \ (@40 \ \mathrm{dB}) & 1.0 \pm 5.6 \% \ (@40 \ \mathrm{dB}) \\  
-& \widehat{b} \ \mathrm{[g]} & 9.88 \pm 25.4 \% \ (@50 \ \mathrm{dB}) & 9.97 \pm 5.2 \% \ (@20 \ \mathrm{dB}) & 10.01 \pm 4.7 \% \ (@40 \ \mathrm{dB}) \\ 
-& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 0.57 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 6.24 \% \ (@30 \ \mathrm{dB}) \\
-& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.1 \pm 116 \% \ (@40 \ \mathrm{dB}) & 0.09 \pm 128 \% \ (@40 \ \mathrm{dB}) & 0.11 \pm 115 \% \ (@40 \ \mathrm{dB}) \\
-
-\hspace{9.595mm} 10\% & \widehat{a} \ [\mathrm{kg/s}] & 1.00 \pm 10.4 \% \ (@30 \ \mathrm{dB}) & 0.995 \pm 10.3 \% \ (@20 \ \mathrm{dB}) & 0.996 \pm 10.3 \% \ (@20 \ \mathrm{dB}) \\    
-& \widehat{b} \ \mathrm{[g]} & 9.94 \pm 10.5 \% \ (@20 \ \mathrm{dB}) & 9.96 \pm 10.4 \% \ (@20 \ \mathrm{dB}) & 10.0 \pm 10.5 \% \ (@20 \ \mathrm{dB}) \\
-& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) \\
-& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.11 \pm 108 \% \ (@50 \ \mathrm{dB}) & 0.11 \pm 110 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 137 \% \ (@50 \ \mathrm{dB}) \\
-
-\sigma_{{m},{d},{k}}  \hspace{2mm} 0\% & \widehat{a} \ [\mathrm{kg/s}] & 0.999 \pm 0.7 \% \ (@20 \ \mathrm{dB}) & 1.0 \pm 0.69 \% \ (@20 \ \mathrm{dB}) & 1.0 \pm 0.33 \% \ (@30 \ \mathrm{dB}) \\ 
-& \widehat{b} \ \mathrm{[g]} & \hspace{-1.75mm} 10.0 \pm 1.1 \% \ (@20 \ \mathrm{dB}) & 10.0 \pm 1.13 \% \ (@20 \ \mathrm{dB}) &  10.0 \pm 0.46 \% \ (@30 \ \mathrm{dB}) \\
-& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.6 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 0.19 \% \ (@30 \ \mathrm{dB}) &  0.1 \pm 0.18 \% \ (@30 \ \mathrm{dB}) \\
-& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.1 \pm 122 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 130 \% \ (@40 \ \mathrm{dB}) & 0.09 \pm 38 \% \ (@50 \ \mathrm{dB}) \\
-
-\hspace{11.375mm} 5\% & \widehat{a} \ [\mathrm{kg/s}] & 0.994 \pm 15.3 \% \ (@20g \ \mathrm{dB}) & 1.0 \pm 0.7 \% \ (@20 \ \mathrm{dB}) & 1.01 \pm 21.5 \% \ (@20 \ \mathrm{dB}) \\  
-& \widehat{b} \ \mathrm{[g]} & 10.3 \pm 79.4 \% \ (@20 \ \mathrm{dB}) & 10.0 \pm 1.2 \% \ (@20 \ \mathrm{dB}) & 9.91 \pm 19.0 \% \ (@20 \ \mathrm{dB}) \\ 
-& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.2 \% \ (@30 \ \mathrm{dB}) & 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 0.19 \% \ (@20 \ \mathrm{dB}) \\
-& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.10 \pm 113 \% \ (@40 \ \mathrm{dB}) & 0.1 \pm 119 \% \ (@40 \ \mathrm{dB}) & 0.11 \pm 323 \% \ (@30 \ \mathrm{dB}) \\
-
-\hspace{9.595mm} 10\% & \widehat{a} \ [\mathrm{kg/s}] & 1.01 \pm 22 \% \ (@30 \ \mathrm{dB}) & 1.01 \pm 16.0 \% \ (@30 \ \mathrm{dB}) & 0.987 \pm 28 \% \ (@50 \ \mathrm{dB}) \\    
-& \widehat{b} \ \mathrm{[g]} & 0.1 \pm 26 \% \ (@30 \ \mathrm{dB}) & 9.93 \pm 14.8 \% \ (@30 \ \mathrm{dB}) & 10.4 \pm 93 \% \ (@50 \ \mathrm{dB}) \\ 
-& \widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]} & \hspace{-1.75mm} 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) & 0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) &  0.1 \pm 0.6 \% \ (@20 \ \mathrm{dB}) \\
-& \widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]} & \hspace{-1.75mm} 0.1 \pm 121 \% \ (@40 \ \mathrm{dB}) & 0.09 \pm 136 \% \ (@40 \ \mathrm{dB}) & 0.11 \pm 111 \% \ (@40 \ \mathrm{dB}) \\
-[0.5ex] 
+ \multicolumn{1}{l|}{0\%} \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{a} \ \mathrm{[kg/s]}$} & $1.0 \pm 0.7 \%_{(@20 \ \mathrm{dB})}$ &  $1.0 \pm 0.7 \%_{(@20 \ \mathrm{dB})}$ & $1.0 \pm 0.4 \%_{(@30 \ \mathrm{dB})}$ \\ 
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{b} \ \mathrm{[g]}$} & $10.0 \pm 1.1 \%_{(@20 \ \mathrm{dB})}$ & $10.0 \pm 1.1 \%_{(@20 \ \mathrm{dB})}$ & $10.0 \pm 0.5 \%_{(@30 \ \mathrm{dB})}$ \\ 
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]}$} & $0.1 \pm 0.6 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 0.6 \%_{(@20 \ \mathrm{dB})}$ & $0.1 \pm 0.2 \%_{(@30 \ \mathrm{dB})}$ \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]}$} & $0.1 \pm 122 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 130 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 38 \%_{(@50 \ \mathrm{dB})}$ \\
+ \multicolumn{1}{l|}{5\%} \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{a} \ \mathrm{[kg/s]}$} & $1.0 \pm 9.0 \%_{(@50 \ \mathrm{dB})}$ & $1.0 \pm 5.2 \%_{(@40 \ \mathrm{dB})}$ & $1.0 \pm 5.6 \%_{(@40 \ \mathrm{dB})}$ \\  
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{b} \ \mathrm{[g]}$} & $9.9 \pm 25.4 \%_{(@50 \ \mathrm{dB})}$ & $10.0 \pm 5.2 \%_{(@20 \ \mathrm{dB})}$ & $10.0 \pm 4.7 \%_{(@40 \ \mathrm{dB})}$ \\ 
+\multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]}$} & $0.1 \pm 0.6 \%_{(@20 \ \mathrm{dB})}$ & $0.1 \pm 0.6 \%_{(@20 \ \mathrm{dB})}$ & $0.1 \pm 6.2 \%_{(@30 \ \mathrm{dB})}$ \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]}$} & $40.1 \pm 116 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 128 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 115 \%_{(@40 \ \mathrm{dB})}$ \\
+ \multicolumn{1}{l|}{10\%} \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{a} \ \mathrm{[kg/s]}$} & $1.0 \pm 10.4 \%_{(@30 \ \mathrm{dB})}$ & $1.0 \pm 10.3 \%_{(@20 \ \mathrm{dB})}$ & $1.0 \pm 10.3 \%_{(@20 \ \mathrm{dB})}$ \\ 
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{b} \ \mathrm{[g]}$} & $9.9 \pm 10.5 \%_{(@20 \ \mathrm{dB})}$ & $9.9 \pm 10.4 \%_{(@20 \ \mathrm{dB})}$ & $10.0 \pm 10.5 \%_{(@20 \ \mathrm{dB})}$ \\ 
+\multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]}$} & $0.1 \pm 0.6 \%_{(@20 \ \mathrm{dB})}$ & $0.1 \pm 0.6 \%_{(@20 \ \mathrm{dB})}$ & $0.1 \pm 0.6 \%_{(@20 \ \mathrm{dB})}$ \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]}$} & $0.1 \pm 108 \%_{(@50 \ \mathrm{dB})}$ & $0.1 \pm 110 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 137 \%_{(@50 \ \mathrm{dB})}$ \\
+ \hline
+ $\sigma_{{m},{d},{k}}$ & $\sigma_{\mathrm{s}}$: \ 0\% & 5\% & 10\% \\
+ \hline
+ \multicolumn{1}{l|}{0\%} \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{a} \ \mathrm{[kg/s]}$} & $1.0 \pm 0.7 \%_{(@20 \ \mathrm{dB})}$ &  $1.0 \pm 0.7 \%_{(@20 \ \mathrm{dB})}$ & $1.0 \pm 0.3 \%_{(@30 \ \mathrm{dB})}$ \\ 
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{b} \ \mathrm{[g]}$} & $10.0 \pm 1.1 \%_{(@20 \ \mathrm{dB})}$ & $10.0 \pm 1.1 \%_{(@20 \ \mathrm{dB})}$ & $10.0 \pm 0.4 \%_{(@30 \ \mathrm{dB})}$ \\ 
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]}$} & $0.1 \pm 0.6 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 0.2 \%_{(@30 \ \mathrm{dB})}$ & $0.1 \pm 0.2 \%_{(@30 \ \mathrm{dB})}$ \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]}$} & $0.1 \pm 122 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 130 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 38 \%_{(@50 \ \mathrm{dB})}$ \\
+ \multicolumn{1}{l|}{5\%} \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{a} \ \mathrm{[kg/s]}$} & $1.0 \pm 15.3 \%_{(@20 \ \mathrm{dB})}$ & $1.0 \pm 0.7 \%_{(@20 \ \mathrm{dB})}$ & $1.0 \pm 21.5 \%_{(@20 \ \mathrm{dB})}$ \\  
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{b} \ \mathrm{[g]}$} & $10.3 \pm 79.4 \%_{(@20 \ \mathrm{dB})}$ & $10.0 \pm 1.2 \%_{(@20 \ \mathrm{dB})}$ & $9.9 \pm 19.0 \%_{(@20 \ \mathrm{dB})}$ \\ 
+\multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]}$} & $0.1 \pm 0.2 \%_{(@30 \ \mathrm{dB})}$ & $0.1 \pm 0.6 \%_{(@20 \ \mathrm{dB})}$ & $0.1 \pm 0.2 \%_{(@20 \ \mathrm{dB})}$ \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]}$} & $0.1 \pm 113 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 119 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 323 \%_{(@30 \ \mathrm{dB})}$ \\
+ \multicolumn{1}{l|}{10\%} \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{a} \ \mathrm{[kg/s]}$} & $1.0 \pm 22 \%_{(@30 \ \mathrm{dB})}$ & $1.0 \pm 16 \%_{(@30 \ \mathrm{dB})}$ & $1.0 \pm 28 \%_{(@50 \ \mathrm{dB})}$ \\ 
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{b} \ \mathrm{[g]}$} & $0.1 \pm 26 \%_{(@30 \ \mathrm{dB})}$ & $9.9 \pm 14.8 \%_{(@30 \ \mathrm{dB})}$ & $10.4 \pm 93 \%_{(@50 \ \mathrm{dB})}$ \\ 
+\multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,1}} \ \mathrm{[g]}$} & $0.1 \pm 0.6 \%_{(@20 \ \mathrm{dB})}$ & $0.1 \pm 0.6 \%_{(@20 \ \mathrm{dB})}$ & $0.1 \pm 0.6 \%_{(@20 \ \mathrm{dB})}$ \\
+ \multicolumn{1}{l|}{\hspace{2mm} $\widehat{x}_{\mathrm{ini,2}} \ \mathrm{[g/s]}$} & $0.1 \pm 121 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 136 \%_{(@40 \ \mathrm{dB})}$ & $0.1 \pm 111 \%_{(@40 \ \mathrm{dB})}$ \\  [0.5ex]
 \hline
 \end{tabular}
-\
+
 \label{table:ml_sensitivity}
 \end{table}
-
-\end{comment}
 
@@
 The upper value $f_u$ and the decay rate $\alpha$ were found using optimization (\ref{eqn:tv_optim}). 
-We chose the values $\mu_{\mathrm{spec}}=0.5$ and $\sigma_{\mathrm{spec}}=0.24$ as they are specified in the OIML recommendation \cite{OIML_R51_1} for a mass of 100 g measured in a conveyor belt.
+We chose the values $\mu_{\mathrm{spec}}=0.5$ and $\sigma_{\mathrm{spec}}=0.24$ as they are specified in the OIML recommendation \citep{OIML_R51_1} for a mass of 100 g measured in a conveyor belt.
 The optimized values of the frequency upper value and the decay rate, using a dataset of 100 transient responses, were $f_u = 26.94$ Hz and $\alpha = 5.71$.

### 2020-01-30 ChapterStatisticalAnalysis/ChapterStatisticalAnalysis.tex
Preliminaries ready

@@
 
-\chapter{Statistical Analysis}\label{chap:StatisticalAnalysis}
+\chapter{Statistical Analysis} \label{chap:StatisticalAnalysis}
 
 \begin{quote}
-\emph{The results of this chapter ...}\vfill{}
+\emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I.
+Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Bias and covariance of the least squares estimate in a structured errors-in-variables problem", Computational Statistics & Data Analysis journal, Vol.
+144, 2020, ISSN 0167-9473, doi:10.1016/j.csda.2019.106893.
+\nocite{QuintanaCSDA} }\vfill{}
 \end{quote}
 
-\section{Introduction}
+\vfill{}
+
+%\section{Introduction}
 
@@
 
-\section{Statistical analysis}
+\section{Statistical analysis of the data-driven step input estimation method}
 
-For an overdetermined system of linear equations, the least-squares (LS) solution is given by
+For an overdetermined system of linear equations, constructed by the data-driven step input estimation method (\ref{eqn:min_seiv}), the least-squares (LS) solution is given by
 \begin{equation} \widehat{\mathbf{x}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
 where $\widetilde{\mathbf{K}}^\dagger$ is the pseudo-inverse of $\widetilde{\mathbf{K}}$.
-The objective of the statistical analysis is to obtain the bias and the covariance of the solution $\widehat{\mathbf{x}}$.
+The objective of the statistical analysis is to obtain the bias and the covariance of the solution $\widehat{\mathbf{x}}$ to the structured and correlated EIV problem (\ref{eqn:ddsiemnd}).
 The bias and the covariance are computed using the mathematical expectation operator $\mathbb{E}\{\cdot\}$.
-First we substitute (\ref{eqn:y0plusnoise}) and (\ref{eqn:K0plusnoise}) in (\ref{eqn:xhat}) obtaining
+First we substitute the expressions (\ref{eqn:y0plusnoise}) and (\ref{eqn:K0plusnoise}) in Equation (\ref{eqn:xhat}) to obtain
 \begin{equation*} \widehat{\mathbf{x}} = \left( (\mathbf{K}+\mathbf{E})^\top (\mathbf{K}+\mathbf{E})  \right)^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}), \end{equation*} 
@@
 This subsection describes the statistics of a structured EIV problem with correlation between the perturbations of the regression matrix and the regressor. 
-The structured EIV problem is given by the step input estimation method (\ref{eqn:min_ls}).
+The structured EIV problem is given by the step input estimation method (\ref{eqn:min_seiv}).
 The correlation is a consequence of the construction of the block-Hankel matrix in the regression matrix $\widetilde{\mathbf{K}}$ with the first difference of the elements in the regressor $\widetilde{\mathbf{y}}$.
@@
 
-The minimization problem (\ref{eqn:min_ls}) is equivalent to a structured EIV problem that can be expressed as a linear in the measurements problem \citep{Pintelon12Book}
+The minimization problem (\ref{eqn:min_seiv}) is a structured EIV problem that can be expressed as a linear in the measurements problem \citep{Pintelon12Book}
 \begin{equation} e (\widehat{\mathbf{x}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\mathbf{x}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\mathbf{x}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 .
@@
 
-The Cram\'er-Rao lower bound for an biased estimator of the minimization problem (\ref{eqn:min_ls}) is given by
+The Cram\'er-Rao lower bound for an biased estimator of the minimization problem (\ref{eqn:min_seiv}) is given by
 \begin{equation} \mathrm{CRB}_{\mathrm{b}}(\mathbf{x}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right)^\top \ \mathbf{Fi}^{-1}(\mathbf{x}) \ \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right), \label{eqn:CRB_EIV} \end{equation}
@@
 Two Monte Carlo (MC) simulation studies were conducted to test the obtained bias and variance formulas.
-One MC simulation was devoted to the unstructured EIV problem and the other to the structured EIV problem that corresponds to the step input estimation method (\ref{eqn:min_ls}).
+One MC simulation was devoted to the unstructured EIV problem and the other to the structured EIV problem that corresponds to the step input estimation method (\ref{eqn:min_seiv}).
 The MC simulations performed $N_{MC} = 10^6$ runs of the LS estimator with different realizations of the perturbation $[\begin{matrix} \mathbf{E} & \bm{\epsilon}] \end{matrix}$.
@@
 
-The MC simulation of the structured EIV problem (\ref{eqn:min_ls}) solution was conducted processing $T = 200$ samples of the transient response $\widetilde{\mathbf{y}}$ generated by a linear time invariant system of order $n = 2$, after a step input excitation with $u = 1 \ \mathrm{units}$, where the units represent any physical quantity.
+The MC simulation of the structured EIV problem (\ref{eqn:ddsiemnd}) solution was conducted processing $T = 200$ samples of the transient response $\widetilde{\mathbf{y}}$ generated by a linear time invariant system of order $n = 2$, after a step input excitation with $u = 1 \ \mathrm{units}$, where the units represent any physical quantity.
 The processed step response is shown in Figure \ref{fig:y}.

### 2020-01-30 Gus-thesis.tex
Preliminaries ready

@@
 \usepackage[latin9]{inputenc}
-\usepackage[paperwidth=17cm, paperheight=24cm]{geometry}
+\usepackage[paperwidth=17cm, paperheight=24cm]{geometry} % \usepackage[hmargin=1.5cm, textheight=25cm]{geometry}
 \geometry{verbose, tmargin=2.5cm, bmargin=3cm, lmargin=2.5cm, rmargin=2cm}
@@
 \definecolor{note_fontcolor}{rgb}{0.800781, 0.800781, 0.800781}
-\usepackage{amsmath, amssymb, amsthm, amsbsy, array, babel, booktabs, bm, float, graphicx, mathdots, mathptmx, mathtools, newproof, textcomp, times, units, verbatim}
-\usepackage[unicode=true,pdfusetitle, bookmarks=true, bookmarksnumbered=false, bookmarksopen=false, breaklinks=true,pdfborder={0 0 0}, pdfborderstyle={}, backref=false, colorlinks=false] {hyperref}
+\usepackage{amsfonts, amsmath, amssymb, amsthm, amsbsy, array, babel, booktabs, bm, float, graphicx, lineno, mathdots, mathptmx, mathtools, newproof, pgf, subcaption, textcomp, times, units, verbatim, xcolor}
+\usepackage[unicode=true, pdfusetitle, bookmarks=true, bookmarksnumbered=false, bookmarksopen=false, breaklinks=true, pdfborder={0 0 0}, pdfborderstyle={}, backref=false, colorlinks=false]{hyperref}
 \usepackage[outdir=./]{epstopdf}
 \usepackage{tikz, circuitikz}
-\usetikzlibrary{shapes,arrows,calc,patterns,decorations.pathmorphing,decorations.markings,positioning}
+\usetikzlibrary{shapes, arrows, calc, patterns, decorations.pathmorphing, decorations.markings, positioning}
 
@@
 \newproof{pf}{Proof}
+
+\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}
+
+\usepackage{algorithm, algorithmic}
 

### 2020-02-02 ChapterConclusion/ChapterConclusion.tex
First draft completed!

--- 2020-01-27 ChapterConclusion/ChapterConclusion.tex
@@
 
-In this final chapter, we first summarize the contributions made to both parts of the thesis: \textbf{Part \ref{part:Filter}} and \textbf{Part \ref{part:PA}}. Then, we propose
-some possible further improvements to extend this work in future research.
+%In this final chapter, I first summarize the contributions made in the thesis. Then, I propose some possible further improvements to extend this work in future research.
+
+%\newpage
+%\section*{Conclusions}
+
+In this tesis is described a process to validate a signal processing method for practical applications.
+The signal processing method takes the transient step response of a linear time-invariant sensor and provides an estimation of the level of the originating step input.
+The signal processing method is a subspace data-driven estimation method and constitutes an alternative to the common sensor response processing approaches based in compensation filters.
+Contrary to the compensation filters, this data-driven estimation method is model independent, and improves the estimation time because the model parameters do not have to be estimated neither previously nor simultaneously to the input estimation.
+The improvement in speed estimation mathes the data-driven method suitable for online applications, such as real-time measurements.
+
+In order to convince the metrology community of the benefits and effectiveness of the data-driven method, is was pending to quantify the uncertainty of is resulting estimation.
+The uncertainty of the this method remained unknown because is not straightforward to determine the statistical propertines of its input estimate.
+The data-driven method formulates a minimization problem that is a structured and correlated errors-in-variables (EIV) problem.
+In order to facilitate the practical online applications, the solution to the minimization problem is proposed in terms of the recursive least-squares (LS) estimator.
+The structure and the correlation made necessary to conduct research towards the the first and second moments statistical moments of the LS solution.
+
+The first part of the research reported in this thesis deals with the determination of the mean value and the covariance that the LS estimator has when it solves a structured and correlated EIV problem.
+The Taylor series expansion of the LS solution permits the study of the expected value of the LS estimate.
+The series expansion considers the structure of the regression matrix and the correlation between the regression matrix and the regressor.
+It was found that the element-wise treatment is not a strict requisite for the statistical analysis, since a matrix form treatment is possible.
+As a result, the statistical analysis yielded expressions that approximate the bias and the covariance of the LS estimate, for different amount of samples and perturbation levels.
+These expressions also help to understand the impact that the different matrix structures and the correlation bring into EIV problems and their solutions.
+
+The LS bias and covariance predictions were validated first by Monte Carlo simulations. 
+In the simulations, the empirical statistical moments of the LS estimation were compared to the predictions.
+The simulations were performed with an extensive set of values in the workspace of the sample size of the step response observations, and the perturbation levels.
+The learning obtained from the simulation results are the operating conditions in which the estimation method is effective.
+The operation conditions were the data-driven estimation method was found to be effective include a region of signal-to-noise (SNR) near to 40 dB, were many real-life applications are currently used.
+This is a positive indication for the usability of the method in metrology.
+Another encouraging result is that the mean squared error of the data-driven method estimate is considerably near to the minimal theoretical variance that is defined by the Cram\'er-Rao lower bound of the formulated EIV problem.
+
+Experimental research in a series of experimental measurements completes the validation of the data-driven step input estimation method.
+The temperature and mass measurements are adequate for testing the implementation of the method.
+These two physical magnitudes are demanded applications in scientific and industrial fields.
+The mass measurement demands more effort from the step input estimation method in a real-life application.
+Mass measurement sensors are easily perturbed by vibrations of the environment.
+Furthermore, the mechanical constructions of the weighing sensors increase the order complexity of the sensor model.
+A non-surprising observation in the experimental mass measurements implementation is that the measurement noise collected from the weighing sensor is not Gaussian white noise.
+Moreover, the weighing sensor used, a load cell sensor, has a high order, in theory infinite. 
+Nevertheless, the data-driven step input estimation method performed very well under these conditions. 
+The estimation method showed robustness against non-white noise and provided acceptable results for system orders selected between 5 and 7.
+
+A third aspect of the conducted research included the exploration for the estimation of other input models with signal processing methods. 
+The affine input is one input model that is one complexity level higher than the step input and then it was selected to be considered for further study.
+The recursive LS estimator is a specific case of the exponentially weighted recursive LS (EWRLS), that is able to process the affine input response to estimate the actual input.
+To do this, the EWRLS estimator uses a tuning parameter that selects the data samples.
+During the study of the data-driven method adaptation for the affine input, it was observed that the affine input applied to a weighing sensor turns the LTI sensor into a time-varying (TV) sensor. 
+Thus, the estimation problem became more interesting.
+The adaptive affine input estimation method was simulated under different assumptions and showed robustness when processing time-varying sensor responses.
+To compare the results of this adaptive method, a maximum-likelihood (ML) estimator based on local optimization, and a previously reported digital time-varying filter were implemented.
+The adaptive affine input estimation method outperformed the time-varying filter by presenting lower estimation time, and the ML method by requiring less computational effort.
+
+\section*{Future work}
+
+The use of model independent signal processing methods is a research field that surely will produce interesting results in the near future. 
+With the increasing power of digital signal processors, the design of new methods is an opportunity that cannot be disregarded.
+
+The data-driven method studied in this thesis is one example of an alternative to dynamic measurements under a different paradigm.
+With respect to this method, one conclusion of the analysis presented can be that the data-driven step input estimation method is not statistically efficient because the estimation shows bias.
+A topic for future research is the efficiency increase, perhaps by designing fast and optimal estimators for structured and correlated EIV problems.
+
+On the other hand, for model-based estimation methods, such as the described ML method that has a high computational cost, there is a need also for efficient implementations to enable online optimization in receding-horizon schemes.
+With such efficient methods, the practical implementation of the ML methods can become feasible for real-time measurements.
 
 \newpage
-\section*{Conclusions}
-
-Conclusions - Statistical analysis
-
- We conducted a statistical analysis of a structured errors-in-variables (EIV) estimation problem with correlation to find the first and second moments of its least-squares solution.
-This estimation problem occurs in metrology when we estimate the value of a measurand directly from the sensor transient response.
-The data-driven estimation of the physical quantity is formulated as a structured EIV problem with correlation that uses the observed transient response to construct both the regression matrix and the regressor.
-The real-time implementation of the method uses a recursive least squares algorithm that is simple and has low computational complexity.
-The assessment of the uncertainty is done using the estimate bias and variance.
-
-The conducted statistical analysis produced expressions that predict the estimate bias and variance for given sample size and perturbation level of the observed response.
-The Monte Carlo simulation validated the predictions.
-We compared the results of solving an unstructured and uncorrelated EIV problem with a structured and correlated EIV problem to understand how the structure and the correlation impacts in the estimation.
-We found that the predictions in the structured case are more susceptible to perturbations.
-This is due to the two approximations involved, a second-order Taylor series expansion of the estimate, and the substitution of perturbed data on the prediction expressions.
-The relative error results indicate that the estimate bias, and variance are predicted using the derived expressions, and the observed data.
-The mean squared error of the estimate is close to the results of the maximum likelihood estimate given by the Cram\'er-Rao lower bound.
-
-The bias and variance can be accurately predicted, provided that the Taylor series expansion is valid.
-This constraint has to be taken into account to ensure the effectiveness of the method in practical applications.
-In the example, it was observed that when the SNR lies outside the validity region, the bias and variance estimation was at most three times larger than the empirical values.
-
-The methodology presented in this paper can be applied to estimate the uncertainty of the solutions to other structured EIV problems.
-The bias and variance expressions obtained after the statistical analysis depend on each specific structure.
-
-Conclusions - Experimental validation
-
-In this paper we investigated the statistical properties of a data-driven step input estimation method in a real-life application.
-The step input estimation method is a structured and correlated errors-in-variables problem that is solved with recursive least squares.
-A statistical analysis was conducted using the ordinary least squares condensed notation. 
-This statistical analysis of the input estimate provides expressions that approximate the estimation bias and variance assuming that the measurement noise is Gaussian white noise. 
-The variance approximation is useful to assess the uncertainty of the input estimate.
-In simulation we observed that the mean squared error of the input estimate is close to the theoretical minimum that uses the Cram\'er-Rao lower bound for biased estimators.
-Since the data-driven step input estimation method is not statistically efficient, there is room for improvement. This is a topic for future research.
-In the practical experiments, the measurement noise is not white.
-The noise variance obtained from the sensor steady state response underestimates the measurement noise variance, that was observed 5 dB larger in the power spectrum due to nonlinearities of the sensor.
-Considering this difference in the measurement noise variance, we introduced a conservative bound of the measurement noise variance so that the first and second moments of the input estimate are more accurately predicted.
-Using the variance approximation, we can assess the uncertainty of the input estimate with respect to the number of samples processed by the data-driven step input estimation method.
-The step input estimation method is useful in practical applications where the whiteness assumption of the measurement noise is not fulfilled.
-
-Conclusions - Ramp input
-
-An adaptive subspace method was proposed for the estimation of affine input parameters given the measurement of the caused sensor transient response. 
-The subspace estimation method is a recursive method that allows online implementation.
-This method tracks the input of a system, using exponential forgetting, to process the system response.
-The subspace method is model-free and estimates directly the input parameters without identifying a sensor model.
-Therefore, it can be applied to the measurement of different physical magnitudes.
-In the specific weighing example described in the manuscript, the input is an affine function.
-The method is also applicable when the sensor is time-varying.
-The subspace method is computationally cheap, simple and suitable for implementation on digital signal processor of low computational power.
-
-A maximum-likelihood estimator based on local optimization was designed to obtain a comparative reference for the other methods.
-The maximum-likelihood method estimates the affine input parameters and also model parameters and the sensor's initial conditions.
-This method simulates, in a receding horizon scheme, the response of a sensor model to estimate the input and minimizes the sum of the squares of the residual between the measured and the estimated responses.
-The main drawback of the maximum-likelihood method is its computational cost and efficient implementation of the method is left for future work.
-
-A linear time-invariant weighing system is used as an test example for the estimation methods.
-The weighing system becomes time-varying when an affine input excites the system.
-The estimation methods are compared in a simulation study where the time-varying sensor response is perturbed by measurement noise, that is assumed white of zero mean and known variance.
-The subspace method results are also compared to those of an existing digital time-varying filter.
-The coefficients of the time-varying filter require offline optimization.
-The estimation results obtained with the subspace method converges two times faster and is one order of magnitude smaller than those obtained with the time-varying filter.
-The empirical mean squared errors of the subspace method estimation is two orders of magnitude larger than the theoretical minimum given by the Cram\'er-Rao Lower bound.
-
-Future work of this research is the practical implementation of the subspace method for real-time measurements.
 
 \begin{comment}
-\section{Conclusions - Ramp input}
-
-An adaptive subspace method was proposed for the estimation of affine input parameters given the measurement of the caused sensor transient response. 
-The subspace estimation method is a recursive method that allows online implementation.
-This method tracks the input of a system, using exponential forgetting, to process the system response.
-The subspace method is model-free and estimates directly the input parameters without identifying a sensor model.
-Therefore, it can be applied to the measurement of different physical magnitudes.
-In the specific weighing example described in the manuscript, the input is an affine function.
-The method is also applicable when the sensor is time-varying.
-The subspace method is computationally cheap, simple and suitable for implementation on digital signal processor of low computational power.
-
-A maximum-likelihood estimator based on local optimization was designed to obtain a comparative reference for the other methods.
-The maximum-likelihood method estimates the affine input parameters and also model parameters and the sensor's initial conditions.
-This method simulates, in a receding horizon scheme, the response of a sensor model to estimate the input and minimizes the sum of the squares of the residual between the measured and the estimated responses.
-The main drawback of the maximum-likelihood method is its computational cost and efficient implementation of the method is left for future work.
-
-A linear time-invariant weighing system is used as an test example for the estimation methods.
-The weighing system becomes time-varying when an affine input excites the system.
-The estimation methods are compared in a simulation study where the time-varying sensor response is perturbed by measurement noise, that is assumed white of zero mean and known variance.
-The subspace method results are also compared to those of an existing digital time-varying filter.
-The coefficients of the time-varying filter require offline optimization.
-The estimation results obtained with the subspace method converges two times faster and is one order of magnitude smaller than those obtained with the time-varying filter.
-The empirical mean squared errors of the subspace method estimation is two orders of magnitude larger than the theoretical minimum given by the Cram\'er-Rao Lower bound.
-
-Future work of this research is the practical implementation of the subspace method for real-time measurements.
-
 \section{Conclusions - Statistical analysis}
@@
 
-\section*{Future work}
+\section{Conclusions - Ramp input}
+
+An adaptive subspace method was proposed for the estimation of affine input parameters given the measurement of the caused sensor transient response. 
+The subspace estimation method is a recursive method that allows online implementation.
+This method tracks the input of a system, using exponential forgetting, to process the system response.
+The subspace method is model-free and estimates directly the input parameters without identifying a sensor model.
+Therefore, it can be applied to the measurement of different physical magnitudes.
+In the specific weighing example described in the manuscript, the input is an affine function.
+The method is also applicable when the sensor is time-varying.
+The subspace method is computationally cheap, simple and suitable for implementation on digital signal processor of low computational power.
+
+A maximum-likelihood estimator based on local optimization was designed to obtain a comparative reference for the other methods.
+The maximum-likelihood method estimates the affine input parameters and also model parameters and the sensor's initial conditions.
+This method simulates, in a receding horizon scheme, the response of a sensor model to estimate the input and minimizes the sum of the squares of the residual between the measured and the estimated responses.
+The main drawback of the maximum-likelihood method is its computational cost and efficient implementation of the method is left for future work.
+
+A linear time-invariant weighing system is used as an test example for the estimation methods.
+The weighing system becomes time-varying when an affine input excites the system.
+The estimation methods are compared in a simulation study where the time-varying sensor response is perturbed by measurement noise, that is assumed white of zero mean and known variance.
+The subspace method results are also compared to those of an existing digital time-varying filter.
+The coefficients of the time-varying filter require offline optimization.
+The estimation results obtained with the subspace method converges two times faster and is one order of magnitude smaller than those obtained with the time-varying filter.
+The empirical mean squared errors of the subspace method estimation is two orders of magnitude larger than the theoretical minimum given by the Cram\'er-Rao Lower bound.
+
+Future work of this research is the practical implementation of the subspace method for real-time measurements.
 \end{comment}
-
-text
-
-\newpage

### 2020-03-01 Gus-thesis-R1-answer-review.tex
Second review version

\documentclass[11pt]{article}
\date{\vspace{-10ex}}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\usepackage{amsmath, amsfonts, mathtools, hyperref, tikz, pgf, subcaption, mathdots, color, bm}
\usepackage[a4paper, total={6.5in, 9.0in}]{geometry}
\usepackage{mathptmx, epsfig, amssymb, amsthm, setspace, hanging}

\begin{document}

\title{Author' response to the promotors' reports on the first draft of the thesis document.}

\maketitle

\hspace{-0.6cm}We appreciate the referees useful suggestions.
In this document we quote in \textbf{bold} face the comments from the reports.
Our replies are in normal face.
The passages quoted from the revised manuscript are highlighted in \color{blue}blue\color{black}.

\section*{Reviewer 1}

\begin{enumerate}

\item \textbf{I find the current version of the paper to be a strong improvement over the original form. There are, however, several points where the paper can be improved to make it more attractive and informative for the readers.}

Thank you for your suggestions in the previous report, which helped us to improve the manuscript.

\item \textbf{I am still missing a derivation of (13), (14) as well as (17), (18). As a potential
reader I would prefer to see the derivations in an appendix, instead of just
trusting the authors. A reserve plan would be to issue a technical report,
made publically available on the web, with the full derivations.}

The derivation of the bias and covariance expressions has been added in Appendix 1.

\item \textbf{In page 3, eq (2). Say that this is the LS solution, and preferably give also the underlying system of equations, that is $\tilde{y}=\tilde{K}x$ , for possible reference in the text (for example just before eq (6)). Further, already here it is appropriate to state and explain that a recursive solution is desired, and hence an LS estimate is examined (even if it has some bias). It is a bit too late and too meagre to just state this in passing at p 4, line 61.}

The paragraph after equation 5 has been modified as follows:

\color{blue} 
The underlying system of equations $\widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \mathbf{x}$ in the minimization problem (2)  is an errors-in-variables (EIV) problem with Hankel structure.
For metrology applications, the least-squares (LS) solution of the system of equations offers a simple alternative, in its recursive form, to implement the estimation method in real-time.
The LS solution is examined even when it may have some bias because the perturbation errors in $\widetilde{\mathbf{K}}$ are correlated to the perturbations in $\widetilde{\mathbf{y}}$.
\color{black}

\item \textbf{In page 4, line 61. Help the reader and give a comment why the classical LS results do not apply.}

The sentence has been modified as follows:

\color{blue} 
The classical LS results for the bias and the covariance cannot be invoked because LS assumes that the additive perturbation only affects the regressor, and that there is no correlation between the regressor and the regression matrix.
\color{black}

\item \textbf{In page 4, around eq (9) and (10). The current presentation is not good. It is not enough that the spectral radius is smaller than one for the approximation in (9) to be accurate [think of a case where the radius is 0.99, for example]. I find it much better to start with assuming that the SNR is high (enough).
Then one can conclude that E and M are small, and that the approximation (beginning of a convergent series expansion (9)) makes sense. You may add that the neglected term is $O(\Vert M \Vert^3)$.}

The description of the applied Taylor series expansion is now presented as:

\color{blue}
Applying a second-order Taylor expansion of the inverse matrix
\begin{equation} (\mathbf{I} + \mathbf{M})^{-1} \approx \mathbf{I} - \mathbf{M} + \mathbf{M}^2, \tag{9} \end{equation} 
 that is valid when the SNR is sufficiently high, and therefore $\mathbf{E}$ and $\mathbf{M}$ are small, satisfying the constraint on the spectral radius $\| \mathbf{M} \| < 1$. 
 The neglected term in the Taylor series expansion is of the order $O(\Vert M \Vert^3)$.
We can express the estimate as
\begin{equation} \widehat{\mathbf{x}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}). \tag{10} \end{equation} 
\color{black}

\item \textbf{In page 5, second line. Change 'approximation' to 'approximation (10)'.}

The sentence now reads:

\color{blue} 
Now that the perturbation variables $\bm{\epsilon}$ and $\mathbf{E}$ are no longer inside an inverse matrix, we can compute the mathematical expectation of expressions derived from the Taylor series approximation (10) of $\widehat{\mathbf{x}}$. 
\color{black}

\item  \textbf{In page 5, line 70. Change 'expansion' to 'approximation'.}

The sentence now reads:

\color{blue} 
Moreover, the second-order approximation disregards moments of order four and higher.
\color{black}

\item  \textbf{In page 5, eq (13). Should not the subscript E be in bold here?}

Yes, the equation is now presented as:

\color{blue} 
\begin{equation} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - T \right) \mathbf{Q}^{-1} \mathbf{x} \tag{13} \end{equation}
\color{black}

\item \textbf{In page 5, eq (13), (14). Comment at this stage what the subscript p stands for.}

We have added the following sentence after equations (13) and (14):

\color{blue} 
Moreover, the second-order approximation disregards moments of order four and higher.
\color{black}

\item \textbf{In page 9, line 104. I would find it more natural to write 'Thus' than 'Then' here.}

The sentence now reads:

\color{blue} 
Thus, it is relevant to monitor the evolution of the largest $\mathbf{M}$ eigenvalue to detect the lower limit of the SNR for the predictions to be valid.

\color{black}

\item \textbf{In page 12, line 152. It would be nice to present the response of the system as a
graph.}

The response of the system used to construct the structured EIV problem has been added in Figure 5.

\item \textbf{In page 15, line 191. I would change 'predicted' $\rightarrow$ 'accurately predicted'.}

The sentence now reads:

\color{blue} 
The bias and variance can be accurately predicted, provided that the Taylor series expansion is valid.
\color{black}

\end{enumerate}

\section*{Reviewer 2}

\begin{enumerate}

\item  \textbf{In my opinion, the updated version of the paper has clarified several aspects of the idea, but I would suggest that the equation and description in section IV (Simulation results), could be improved.}

Thank you for your suggestions in the previous report, which helped us to improve the manuscript.

Section 4 has been revised and improved to remove unnecesary redundancy and to clarify the discussion of the results.

\end{enumerate}

\section*{Editor}

\begin{enumerate}

\item \textbf{Follow all the editorial guidelines when you prepare the revised manuscript 
(eg abstract in the third person and no statements like this paper etc.).}

\begin{enumerate}

\item \textbf{Delete the redundant terms "This paper" from the abstract.  E.g.}

\begin{enumerate}

\item \textbf{A structured errors-in-variables (EIV) problem arising in metrology is studied}.

\item \textbf{Expressions to predict the bias and variance of LS estimators applied to unstructured and structured EIV problems are derived}.

These sentences in the abstract have been modified accordingly.

\end{enumerate}

\item \textbf{Use the journal's citation style.}

The format of the References Section has been revised accordingy to follow the journal style.

\item \textbf{Delete the '2010 MSC: 00-01, 99-00'}

This sentence has been removed.

\end{enumerate}

\item  \textbf{Have the paper proof read and corrected.}

The manuscript has been proof read, and the necessary corrections have been done.

\end{enumerate}

\end{document}

### 2020-03-01 ChapterAppendices/ChapterAppendices.tex
Second review version

@@
 
-\chapter{Appendices }\label{chap:Appendices}
+\chapter{Appendices}\label{chap:Appendices}
 % \begin{quote}
@@
 
-\section{Appendix - Derivation of bias and covariance expressions.  \label{appendix:biascovSEIV} }
+\section{Derivation of bias and covariance expressions.  \label{appendix:biascovSEIV} }
 
 The bias and covariance of the least-squares (LS) estimate (\ref{eqn:xhatexp}) are obtained using the mathematical expectation in the definitions (\ref{eqn:biasdef}) and (\ref{eqn:covdef}).
-For an unstructured and uncorrelated EIV problem, the expected value, and the covariance of $\widehat{\mathbf{x}}$ are approximated by
-\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E}  \right\} \mathbf{x}, \quad \text{and}  \\
-\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{x} \mathbf{x}^\top + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \bm{\epsilon} \bm{\epsilon}^\top \mathbf{K} + \mathbf{K}^\top \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} \right\} \mathbf{Q}^{-1} \\
-& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{x} \mathbf{x}^\top \\
-& + \mathbf{x} \mathbf{x}^\top \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{Q}^{-1} - \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}^\top , \label{eqn:EC_Eu} \end{aligned} \end{equation}
+For an unstructured and uncorrelated EIV problem, the expected value, and the covariance of $\widehat{\bm{\theta}}$ are approximated by
+\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\} & \approx \bm{\theta}  + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E}  \right\} \bm{\theta}, \quad \text{and}  \\
+\mathbf{C} \left( \widehat{\bm{\theta}} \right)  & \approx \bm{\theta} \bm{\theta}^\top + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \bm{\epsilon} \bm{\epsilon}^\top \mathbf{K} + \mathbf{K}^\top \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \mathbf{K} \right\} \mathbf{Q}^{-1} \\
+& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \bm{\theta} \bm{\theta}^\top \\
+& + \bm{\theta} \bm{\theta}^\top \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{Q}^{-1} \\
+& \quad - \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\} \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\}^\top , \label{eqn:EC_Eu} \end{aligned} \end{equation} 
 where we have considered the second order Taylor series approximation (\ref{eqn:TseriesExp}), and we have removed the terms of zero expected value, and the terms of order higher than 2.
 After an elementwise evaluation of the corresponding expected values in (\ref{eqn:EC_Eu}), the expressions result in 
-\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  + \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) = \mathbf{x}  +  \sigma_{\mathbf{E}}^2 \mathbf{Q}^{-1} \left( 2\mathbf{I} + 2n\mathbf{I} - T \mathbf{I} \right) \mathbf{x} , \quad \text{and} \\
-\mathbf{C}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \mathbf{x} \mathbf{x}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \ \mathbf{Q}^{-1} \mathbf{x} \mathbf{x}^\top \mathbf{Q}^{-1} , \end{aligned} \end{equation}
+\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\} & \approx \bm{\theta}  + \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) = \bm{\theta}  +  \sigma_{\mathbf{E}}^2 \mathbf{Q}^{-1} \left( 2\mathbf{I} + 2n\mathbf{I} - T \mathbf{I} \right) \bm{\theta} , \quad \text{and} \\
+\mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \bm{\theta} \bm{\theta}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \ \mathbf{Q}^{-1} \bm{\theta} \bm{\theta}^\top \mathbf{Q}^{-1} , \end{aligned} \end{equation}
 from where equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) are obtained.
@@
 On the other hand, due to the correlation, the expressions that approximate the expected value of the LS estimate of the structured and correlated EIV problem (\ref{eqn:ddsiemnd}) have a different form:
-\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & \approx \mathbf{x}  +  \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E}  \right\} \mathbf{x} \\
-& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{E}^\top \bm{\epsilon} - \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} - \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\} , \quad \text{and}  \\
-\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{x} \mathbf{x}^\top + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \bm{\epsilon} \bm{\epsilon}^\top \mathbf{K} + \mathbf{K}^\top \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} - \mathbf{K}^\top \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \mathbf{K} - \mathbf{K}^\top \bm{\epsilon} \mathbf{x}^\top \mathbf{E}^\top \mathbf{K} \right\} \mathbf{Q}^{-1}  \\
-& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{x} \mathbf{x}^\top \\
-& + \mathbf{x} \mathbf{x}^\top \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{Q}^{-1} - \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}^\top \\
-& + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{E}^\top \bm{\epsilon} - \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} - \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\} \mathbf{x}^\top \\
-& + \mathbf{x} \mathbb{E} \left\{ \bm{\epsilon}^\top \mathbf{E} - \bm{\epsilon}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} - \bm{\epsilon}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} \right\} \mathbf{Q}^{-1}.
+\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\} & \approx \bm{\theta}  +  \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E}  \right\} \bm{\theta} \\
+& \quad + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{E}^\top \bm{\epsilon} - \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} - \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\} , \quad \text{and}  \\
+\mathbf{C} \left( \widehat{\bm{\theta}} \right)  & \approx \bm{\theta} \bm{\theta}^\top \\
+& \quad + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \bm{\epsilon} \bm{\epsilon}^\top \mathbf{K} + \mathbf{K}^\top \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \mathbf{K} - \mathbf{K}^\top \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \mathbf{K} - \mathbf{K}^\top \bm{\epsilon} \bm{\theta}^\top \mathbf{E}^\top \mathbf{K} \right\} \mathbf{Q}^{-1}  \\
+& \quad + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \bm{\theta} \bm{\theta}^\top \\
+& \quad + \bm{\theta} \bm{\theta}^\top \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} - \mathbf{E}^\top \mathbf{E} \right\} \mathbf{Q}^{-1} \\
+& \quad - \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\} \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\}^\top \\ 
+& \quad + \mathbf{Q}^{-1} \mathbb{E} \left\{ \mathbf{E}^\top \bm{\epsilon} - \mathbf{K}^\top \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} - \mathbf{E}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\} \bm{\theta}^\top \\
+& \quad + \bm{\theta} \mathbb{E} \left\{ \bm{\epsilon}^\top \mathbf{E} - \bm{\epsilon}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{E}^\top \mathbf{K} - \bm{\epsilon}^\top \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} \right\} \mathbf{Q}^{-1}.
 \label{eqn:EC_E} \end{aligned} \end{equation}
@@
 We have then
-\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\} & = \mathbf{x} + \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \approx \mathbf{x}  +  \mathbf{Q}^{-1} \mathbf{K}^\top \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} \right\}}_{\mathbf{B}_1} - \mathbf{Q}^{-1} \underbrace{\mathbb{E} \left\{ \mathbf{E}^\top \left( \mathbf{I} - \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \right) \mathbf{E} \right\}}_{\mathbf{B}_2} \mathbf{x} \\
-& - \mathbf{Q}^{-1} \mathbf{K}^\top \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\}}_{\mathbf{B}_3} + \mathbf{Q}^{-1} \underbrace{\mathbb{E} \left\{ \mathbf{E}^\top \left( \mathbf{I} - \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \right) \bm{\epsilon} \right\}}_{\mathbf{B}_4} , \quad \text{and} \\
-\mathbf{C} \left( \widehat{\mathbf{x}} \right)  & \approx \mathbf{Q}^{-1} \mathbf{K}^\top \left( \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \bm{\epsilon}^\top \right\}}_{\sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n}}  + \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_1}  - \underbrace{\mathbb{E} \left\{  \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \right\}}_{\mathbf{C}_2} - \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \mathbf{x}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_2^\top} \right) \mathbf{K} \mathbf{Q}^{-1}  - \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\mathbf{x}} \right).
-\end{aligned} \end{equation}
+\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\} & = \bm{\theta} + \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \\
+& \approx \bm{\theta}  +  \mathbf{Q}^{-1} \mathbf{K}^\top \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{E} \right\}}_{\mathbf{B}_1} - \mathbf{Q}^{-1} \underbrace{\mathbb{E} \left\{ \mathbf{E}^\top \left( \mathbf{I} - \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \right) \mathbf{E} \right\}}_{\mathbf{B}_2} \bm{\theta} \\
+& \quad \ \ \ - \mathbf{Q}^{-1} \mathbf{K}^\top \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\}}_{\mathbf{B}_3} + \mathbf{Q}^{-1} \underbrace{\mathbb{E} \left\{ \mathbf{E}^\top \left( \mathbf{I} - \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \right) \bm{\epsilon} \right\}}_{\mathbf{B}_4} , \quad \text{and} \\
+\mathbf{C} \left( \widehat{\bm{\theta}} \right)  & \approx \mathbf{Q}^{-1} \mathbf{K}^\top \left( \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \bm{\epsilon}^\top \right\}}_{\sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n}}  + \underbrace{\mathbb{E} \left\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_1}  - \underbrace{\mathbb{E} \left\{  \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \right\}}_{\mathbf{C}_2} - \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \bm{\theta}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_2^\top} \right) \mathbf{K} \mathbf{Q}^{-1} \\
+& \quad - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right). \end{aligned} \end{equation} 
 from where the expressions (\ref{eqn:biasE}) and (\ref{eqn:varE}) are obtained.
 
-\section{Appendix - Proof of Lemma \ref{lem:lemma1} \label{appendix:lemma1} }
+\section{Proof of Lemma \ref{lem:lemma1} \label{appendix:lemma1} }
 
@@
 
-\section{Appendix - Jacobians in affine input estimation by ML method \label{appendix:Jacobians} }
+\section[Calculation of Jacobian matrices in the affine input ML estimation method]{Calculation of Jacobian matrices \linebreak in the affine input ML estimation method \label{appendix:Jacobians} }
 
 The entries of the Jacobian matrix are the first order partial derivatives of the residual error $\mathbf{r}$ with respect to the optimization variables.
-The state space representation of the weighing model allows to find the analytical expression of the Jacobian.
+The state-space representation of the weighing model allows to find the analytical expression of the Jacobian.
 The partial derivative of the residual error $\mathbf{r}$ with respect to the optimization variable $a$ is 
-\begin{equation} \mathbf{J}_a=\dfrac{\partial \mathbf{r}}{\partial a} = \dfrac{\partial \widehat{y}}{\partial a} = \begin{bmatrix} 1 & 0  \end{bmatrix} \dfrac{\partial \mathbf{x}}{\partial a} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x}_a \end{equation}
-where we use $\mathbf{x}_a = \partial \mathbf{x}/ \partial a$ to simplify the notation. 
+\begin{equation} \mathbf{J}_a=\dfrac{\partial \mathbf{r}}{\partial a} = \dfrac{\partial \widehat{ \mathbf{y} }}{\partial a} = \begin{bmatrix} 1 & 0  \end{bmatrix} \dfrac{\partial \mathbf{x}}{\partial a} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x}_a \end{equation}
+where we use $\mathbf{x}_a = \partial \mathbf{x} / \partial a$ to simplify the notation. 
 Now, from the derivative of the state equation, we have
@@
     & \dot{\mathbf{x}}_a = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_a 
-    + \begin{bmatrix} 0 & 0 \\ \frac{k_{\mathrm{s}} t}{(a t + b + m)^2} & \frac{k_{\mathrm{d}} t - b - m}{(a t + b + m)^2} \end{bmatrix} \mathbf{x} 
-    - \frac{t \ \delta(t)}{(a t + b + m)^2} \mathbf{x}_{\text{ini}}   . 
+    + \begin{bmatrix} 0 & 0 \\ \frac{k_{\mathrm{s}} t}{(a t + b + m)^2} & \frac{k_{\mathrm{d}} t - b - m}{(a t + b + m)^2} \end{bmatrix} \mathbf{x} , \\
+    & \mathrm{with} \quad  \mathbf{x}_a(0) = \begin{bmatrix} 0 \\ 0 \end{bmatrix}. 
 \end{aligned} \end{equation}
-Then, the partial derivative of the error $\mathbf{r}$ with respect to the optimization variable $a$ results in an additional dynamic system.
+Then, the partial derivative of the residual error $\mathbf{r}$ with respect to the optimization variable $a$ results in an additional dynamic system.
 
@@
     & \dot{\mathbf{x}}_b = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_b 
-    + \begin{bmatrix} 0 & 0 \\ \frac{k_{\mathrm{s}}}{(a t + b + m)^2} & \frac{a+k_{\mathrm{d}}}{(a t + b + m)^2} \end{bmatrix} \mathbf{x} 
- - \frac{\delta(t) }{(a t + b + m)^2}  \mathbf{x}_{\text{ini}} ,
- \\ & \mathbf{J}_b = \begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_b . 
+    + \begin{bmatrix} 0 & 0 \\ \frac{k_{\mathrm{s}}}{(a t + b + m)^2} & \frac{a+k_{\mathrm{d}}}{(a t + b + m)^2} \end{bmatrix} \mathbf{x}  ,
+ \\ & \mathbf{J}_b = \begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_b, \quad \mathrm{with} \quad  \mathbf{x}_b(0) = \begin{bmatrix} 0 \\ 0 \end{bmatrix} . 
 \end{aligned} \end{equation}
-The partial derivatives of the error $\mathbf{r}$ with respect to the initial conditions yield the following two Jacobians
+
+The partial derivatives of the residual error $\mathbf{r}$ with respect to the initial conditions yield 
 \begin{equation} \begin{aligned}
-    & \dot{\mathbf{x}}_{\mathbf{x}_{\text{ini,1}}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,1}}} + \begin{bmatrix} \frac{\delta(t)}{a t + b + m} \\ 0 \end{bmatrix} , \\
-    & \mathbf{J}_{\mathbf{x}_{\text{ini,1}}} = \begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,1}}}.
+    & \dot{\mathbf{x}}_{\mathbf{x}_{\text{ini,1}}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,1}}}  , \\
+    & \mathbf{J}_{\mathbf{x}_{\text{ini,1}}} = \begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,1}}}, \quad \mathrm{with} \quad  \mathbf{x}_{\text{ini,1}}(0) = \begin{bmatrix} 1 \\ 0 \end{bmatrix}
 \end{aligned} \end{equation}
@@
 \begin{equation} \begin{aligned}
-    & \dot{\mathbf{x}}_{\mathbf{x}_{\text{ini,2}}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a+k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,2}}} + \begin{bmatrix} 0 \\ \frac{\delta(t)}{a t + b + m} \end{bmatrix} ,  \\
-    & \mathbf{J}_{\mathbf{x}_{\text{ini,2}}}=\begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,2}}}.
+    & \dot{\mathbf{x}}_{\mathbf{x}_{\text{ini,2}}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a+k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,2}}}  ,  \\
+    & \mathbf{J}_{\mathbf{x}_{\text{ini,2}}} = \begin{bmatrix} 1 & 0 \end{bmatrix} \mathbf{x}_{\mathbf{x}_{\text{ini,2}}}, \quad \mathrm{with} \quad  \mathbf{x}_{\text{ini,2}}(0) = \begin{bmatrix} 0 \\ 1 \end{bmatrix} .
 \end{aligned} \end{equation}
+The last two additional dynamic systems are identical, except by their initialization.
 

### 2020-03-01 ChapterConclusion/ChapterConclusion.tex
Second review version

@@
 
-In this tesis is described a process to validate a signal processing method for practical applications.
+In this thesis is described a process to validate a signal processing method for practical applications.
 The signal processing method takes the transient step response of a linear time-invariant sensor and provides an estimation of the level of the originating step input.
@@
 Contrary to the compensation filters, this data-driven estimation method is model independent, and improves the estimation time because the model parameters do not have to be estimated neither previously nor simultaneously to the input estimation.
-The improvement in speed estimation mathes the data-driven method suitable for online applications, such as real-time measurements.
+The improvement in speed estimation makes the data-driven method suitable for online applications, such as real-time measurements.
 
@@
 The data-driven estimation of the physical quantity is formulated as a structured EIV problem with correlation that uses the observed transient response to construct both the regression matrix and the regressor.
-The real-time implementation of the method uses a recursive least squares algorithm that is simple and has low computational complexity.
+The real-time implementation of the method uses a recursive least-squares algorithm that is simple and has low computational complexity.
 The assessment of the uncertainty is done using the estimate bias and variance.
@@
 In this paper we investigated the statistical properties of a data-driven step input estimation method in a real-life application.
-The step input estimation method is a structured and correlated errors-in-variables problem that is solved with recursive least squares.
-A statistical analysis was conducted using the ordinary least squares condensed notation. 
+The step input estimation method is a structured and correlated errors-in-variables problem that is solved with recursive least-squares.
+A statistical analysis was conducted using the ordinary least-squares condensed notation. 
 This statistical analysis of the input estimate provides expressions that approximate the estimation bias and variance assuming that the measurement noise is Gaussian white noise. 

### 2020-03-01 ChapterExperimentalValidation/ChapterExperimentalValidation.tex
Second review version

@@
 The MC simulation was conducted processing $T = 5000$ samples of a simulated transient step response $\widehat{\mathbf{y}}$ generated by a stable linear time-invariant (LTI) system of order $n = 5$, with a sampling frequency of $f_s=4$ kHz.
-This system is a state space model obtained with the System Identification Toolbox using the measured step response of the actual sensor described in the Practical Implementation Section.
+This system is a state-space model obtained with the System Identification Toolbox using the measured step response of the actual sensor described in the Practical Implementation Section.
 This model represents a weighing sensor, and in the simulations the sensor is excited with a mass of 138.32 g following a step input profile. 
@@
 
-We are interested in the first element of $\widehat{\mathbf{x}}$, which is the input estimate $\widehat{u}$.
+We are interested in the first element of $\widehat{\bm{\theta}}$, which is the input estimate $\widehat{u}$.
 The measurement noise variance was selected to have a signal-to-noise ratio (SNR) in the interval [30 dB, 80 dB], according to
@@
 The step input bias and variance predictions from observed data $b_{\widetilde{\mathrm{p}}}$ and $v_{\widetilde{\mathrm{p}}}$, and the associated standard error $\sigma_{\widetilde{\mathrm{p}}}$, are obtained from
-\begin{equation} \begin{aligned} & b_{\widetilde{\mathrm{p}}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathbf{b}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\mathbf{x}} \right) \big|_{\left[1\right]} }, \ v_{\widetilde{\mathrm{p}}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\mathbf{x}} \right) \big|_{\left[1,1\right]} },  \\ & \mathrm{and} \quad \sigma_{\widetilde{\mathrm{p}}} = \sqrt{   \frac{ \sum_{i=1}^{N_{MC}}{ \left( \mathbf{b}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\mathbf{x}} \right) \big|_{\left[1\right]} - b_{\widetilde{\mathrm{p}}} \right)^2 } }{N_{MC}\left( N_{MC}-1 \right)}  } , \end{aligned} \end{equation}
-where $\widetilde{\mathbf{b}}_{\mathrm{p}}^i \left( \widehat{\mathbf{x}} \right) \big|_{\left[1\right]}$ is the first element in the bias vector and $\widetilde{\mathbf{C}}_{\mathrm{p}}^i \left( \widehat{\mathbf{x}} \right) \big|_{\left[1,1\right]}$ is the first element in the covariance matrix obtained in the $i-\mathrm{th}$ approximations.
+\begin{equation} \begin{aligned} & b_{\widetilde{\mathrm{p}}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathbf{b}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1\right]} }, \ v_{\widetilde{\mathrm{p}}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1,1\right]} },  \\ & \mathrm{and} \quad \sigma_{\widetilde{\mathrm{p}}} = \sqrt{   \frac{ \sum_{i=1}^{N_{MC}}{ \left( \mathbf{b}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1\right]} - b_{\widetilde{\mathrm{p}}} \right)^2 } }{N_{MC}\left( N_{MC}-1 \right)}  } , \end{aligned} \end{equation}
+where $\widetilde{\mathbf{b}}_{\mathrm{p}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1\right]}$ is the first element in the bias vector and $\widetilde{\mathbf{C}}_{\mathrm{p}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1,1\right]}$ is the first element in the covariance matrix obtained in the $i-\mathrm{th}$ approximations.
 The predicted bias and variance from exact data are obtained with one evaluation of the expressions
 (\ref{eqn:biasE}) and (\ref{eqn:varE})
-\begin{equation} b_{\mathrm{p}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \big|_{\left[1\right]} }, \ \mathrm{and} \ v_{\mathrm{p}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathrm{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \big|_{\left[1,1\right]} } .
+\begin{equation} b_{\mathrm{p}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \big|_{\left[1\right]} }, \ \mathrm{and} \ v_{\mathrm{p}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathrm{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \big|_{\left[1,1\right]} } .
 \end{equation}
@@
 This apparent contradiction does not prove the invalidity of the estimation method since the results presented correspond to a finite sample size and the asymptotic results cannot be applied.
-The theoretical explanation of the estimation statistics for finite sample sizes is out of the scope of this paper.
+The theoretical explanation of the estimation statistics for finite sample sizes is out of the scope of this document.
 
@@
 
-In this paper we investigated the statistical properties of a data-driven step input estimation method in a real-life application.
-The step input estimation method is a structured and correlated errors-in-variables problem that is solved with recursive least squares.
-A statistical analysis was conducted using the ordinary least squares condensed notation. 
+In this chapter we investigated the statistical properties of a data-driven step input estimation method in a real-life application.
+The step input estimation method is a structured and correlated errors-in-variables problem that is solved with recursive least-squares.
+A statistical analysis was conducted using the ordinary least-squares condensed notation. 
 This statistical analysis of the input estimate provides expressions that approximate the estimation bias and variance assuming that the measurement noise is Gaussian white noise. 

### 2020-03-01 ChapterIntroduction/ChapterIntroduction.tex
Second review version

@@
 % Context:Metrology One of the objectives in metrology is the estimation of the true value of physical quantities.
-A measurement is a dynamical process and a sensor is a dynamical system.
+A measurement is a dynamic process, and a sensor is a dynamic system.
 The physical quantities of interest are the inputs of the system, and the outputs are the electrical signals collected from the sensor.
-The inputs interact with the sensor, and there are energy transferences between them that modify the sensor state.
-The sensor output response depends on the applied inputs, and on the sensor initial conditions.
+The inputs interact with the sensor, and there are energy transfers between them that modify the sensor state.
+The sensor output response depends on the applied inputs and on the sensor initial conditions.
 
 % Context:Tradeoff speed vs accuracy
-There is a trade-off between speed and accuracy in a measurement with a linear time invariant sensor.
-The input excitation drives the sensor into a transient state, that is followed by a steady state.
-During the sensor transient state, the response does not represent directly the input, but 
-in steady state, the sensor response is proportional to the input.
+There is a trade-off between speed and accuracy when measuring with a linear time-invariant sensor.
+The input excitation drives the sensor into a transient state, and when the transient response is below the noise level, we say that the sensor is in steady state.
+During the sensor transient state, the response does not directly represent the input, but 
+in steady state, the sensor response is proportional to the excitation.
 The input can be estimated accurately from the sensor steady state response using the sensor static gain.
-However, waiting for the steady state is not advisable for practical applications that need fast measurements.
-In these practical applications, the input must be estimated during the sensor transient state.
+However, waiting for the steady state is not always possible for practical applications that need fast measurements.
+In these practical applications, the input must be estimated during the transient state.
 
@@
 One approach to get a fast input estimation is filtering the sensor transient response with another dynamical system that inverts the dynamics of the sensor.
-The filter output is an input estimate that compensates the measurement time.
-The transient duration of the compensation filter is smaller than that of the sensor.
-The compensator is designed to deconvolve the sensor response and is based on a sensor model.
+The filter output is an input estimate that compensates for the measurement time.
+The transient duration of the compensation filter should be smaller than that of the sensor.
+The compensator is designed after a sensor model to deconvolve the sensor response.
 
 % Context:Digital signal processing
-Another approach relies in the use of digital signal processors (DSP) to estimate the input value from the sensor transient response.
-DSPs offer an extra versatility level since they allow the implementation of methods that do not necessarily recreate dynamic systems, such as digital filters.
+Another approach relies on the use of digital signal processors (DSP) to estimate the input value from the sensor transient response.
+DSPs offer an extra versatility level since they allow to implement methods that do not necessarily simulate the dynamics of linear systems, such as digital filters.
 A suitable data-driven method in a DSP can provide faster input estimations than with the model-based compensators.
@@
 The regression matrix has a block-Hankel structure.
-The correlation exists because both the regression matrix and the regressor are constructed from the transient response perturbed by measurement noise.
-The measurement noise is assumed to have zero mean and given variance.
+The correlation exists because the transient response, perturbed by measurement noise, constructs both the regression matrix and the regressor.
+The measurement noise is assumed to have zero mean and finite variance.
 The method is implemented in real-time using a recursive least-squares (RLS) solution of the structured EIV problem.
@@
 % The philosophy behind the estimation method is avoiding the explicit identification of a system model from input-output data and estimating directly the input from the transient response.
-The main advantage of the data-driven input estimation method is that it does not identify the sensor model but instead it directly estimates the input. 
-This differs from the classic two-stage methodology where the sensor model is first identified before the input is estimated.
-In this method, the output-error (OE) problem is converted into an EIV problem that is harder to solve, but the RLS solution is easy to compute.
-The range of application of the data-driven input estimation method is wide because it is independent of the sensor model.
+The main advantage of the data-driven input estimation method is that it does not identify the sensor model, but instead, it directly estimates the input. 
+The direct estimation differs from the standard two-stage methodology, that first identifies the sensor model and later estimates the input.
+In this approach, the output-error (OE) problem is converted into an EIV problem that is harder to solve, but the RLS solution is easy to compute.
+The range of application of the data-driven input estimation method is extensive because it is independent of the sensor model.
 The main disadvantage of the data-driven input estimation method is that its stochastic properties are not straightforwardly evident. 
@@
 % DM:Experimental validation of the data-driven step input estimation method
-The estimate uncertainty must be assessed to validate the estimation methods for metrology applications.
+The estimation uncertainty assessment is needed to validate the estimation methods for metrology applications.
 The uncertainty of the data-driven step input estimation method \citep{Markovsky15ieee} is unknown.
-The uncertainty can be defined in terms of the estimate bias and variance \citep{Pintelon12Book}, and these can be obtained by conducting an elementwise statistical analysis.
+The estimation bias and variance define the estimation uncertainty \citep{Pintelon12Book}, and they can be obtained by conducting an elementwise statistical analysis. 
 The validation of the step input estimation method requires also to demonstrate its effectiveness on real-life measurements.
-In real-life experiments, the experimens were conducted with sensors of temperature and mass.
-One challenge of using real-life data is that the noise may not fulfill the whiteness assumptions considered in the estimation problem formulation and in the statistical analysis.
+Temperature and mass sensors are suitable devices to make real-life measurement experiments.
+One challenge of using real-life data is that the noise may not fulfill the whiteness assumption considered in the estimation problem formulation and the statistical analysis.
 The simulation and experimental results permit to compare the performance of the step input estimation method.
@@
 The ideas behind the step input estimation method raise the curiosity towards the design of estimation methods for other input models.
-One of this input models is the ramp.
-The ramp input estimation method is, in addition, motivated by the dynamic weighing that is performed in conveyor systems.
+One of these input models is the ramp.
+The dynamic weighing, performed in conveyor systems, also motivates the ramp input estimation method.
 The dynamic weighing estimates the mass of materials or products during their transportation.
 Ideally, when the conveyor belt transports the materials at a constant speed, i.e., the weighing sensor is excited with a ramp profile. 
-The ramp is an affine input model that consists of two parameters, the slope and the interception.
+The ramp is an affine input model that consists of two parameters: the slope and the interception.
 The slope depends on the applied mass and can be used to estimate it from the transient response.
@@
 % general preview of the thesis contents
-This thesis describes three works iniciated after the data-driven step input estimation method.
+This thesis describes three works initiated after the data-driven step input estimation method.
 The first work is a statistical analysis to elucidate the bias and the covariance of the estimate that this method provides.
-With these statistical moments, the estimate uncertainty is built and we can assess the effectiveness of the method.
+These statistical moments build the estimation uncertainty that appraises the effectiveness of the method.
 
@@
 
-Recent studies in metrology deal with the dynamic process effects that exist in measurements.
-Processing the response of linear time invariant sensors to get fast and accurate input estimations motivates the need to estimate inherent dynamical errors \citep{Hessling06} and to develop methods for dynamic correction in time domain \citep{Hessling08a}.
-The industrial requeriments impose new challenges by increasing the number of sensors distributed in the production lines that provide feedback for quick decision taking to control autonomous activation \citep{Esward09}.
+The scientific literature contains metrology studies that deal with the dynamic process effects on measurements.
+Processing the response of linear time-invariant sensors, to get fast and accurate input estimations, motivates the need to estimate inherent dynamical errors \citep{Hessling06} and to develop methods for dynamic correction in the time domain \citep{Hessling08a}.
+The industrial requirements impose new challenges by increasing the number of sensors distributed in the production lines that provide feedback for quick decision making to control autonomous activation \citep{Esward09}.
 
@@
 a real-time rotational speed estimation using correlation \citep{Wang14},
-a development of an biology-inspired electronic nose \citep{Jing16},
+a development of a biology-inspired electronic nose \citep{Jing16},
 and impedance measurements for material damage estimation using cross-correlation \citep{deCastro19}.
-The data-driven step input estimation method \citep{Markovsky15cep} is added to this list, highlighting the fact that can be implemented for the measurement of different physical magnitudes bacause is model independent.
+The data-driven step input estimation method \citep{Markovsky15cep} is added to this list, highlighting the fact that it can be implemented for the measurement of different physical magnitudes because it is model-independent.
 
 % Uncertainty assessment 
-In metrology, each measurement is an estimation that can be modeled as a random variable.
-Every measurement is inherently perturbed by noise and, therefore, the input estimation is commonly expressed with its two first statistical moments \citep{Ferrero06}.
-In an effort to standardize the uncertainty assessment of the input estimation, a guideline \citep{GUM08}has been prepared and has a overall acceptance in the metrolgy community.
-A review of measurement uncertainty analysis of common use is presented in \citep{daSilva12}.
-The Monte Carlo method has been shown to be an effective uncertainty evaluation tool in \citep{Cox06},
-and supports the usefulness of simulation for quantifying measurement uncertainties by software \citep{Esward16}.
-Another example of the Monte Carlo method is the dynamic measurement uncertainty evaluation of clinical thermometers described in \citep{Ogorevc16}.
+In metrology, a measurement is an estimation, represented as a random variable.
+The measurement noise always exists, and therefore, the input estimation is commonly expressed with its two first statistical moments \citep{Ferrero06}.
+The guideline \citep{GUM08}, accepted by the metrology community, recommends the standardization of the uncertainty assessment.
+The typical measurement uncertainty analysis are reviewed in \citep{daSilva12}.
+The Monte Carlo method is a uncertainty evaluation tool \citep{Cox06},
+that supports the use of simulation techniques for quantifying measurement uncertainties \citep{Esward16}.
+Another example of the Monte Carlo method application is the dynamic measurement uncertainty evaluation of clinical thermometers described in \citep{Ogorevc16}.
 
 Nevertheless, researchers still pinpoint the need to study more the uncertainty assessment methods \citep{Esward09, Hessling10}.
-It is recommended to consider the uncertainties of all the measurement chain components \citep{Diniz17}, and to avoid the direct uncertainty propagation from the calibration towards the to-be-measured quantity.
+Some recommend to consider the uncertainties of all the measurement chain components \citep{Diniz17}, and to avoid the direct uncertainty propagation from the calibration towards the to-be-measured quantity.
 Methods for evaluating the uncertainty associated with the output of compensation filters have been investigated, such as for
@@
 the Kalman filter in \citep{Eichstadt16b}.
-All these works propagate the uncertanity through the filter but it is also necessary to upward the propagation up to the sensor model to include all systematic error contributions \citep{Hessling11}.
+All these works propagate the uncertainty through the filter, but it is also necessary to upward the propagation up to the sensor model to include all systematic error contributions \citep{Hessling11}.
 
@@
 
-This thesis describes research work iniciated after the data-driven step input estimation method.
-This method was proposed for metrology applications but it lacked of an uncertainty assessment, and, therefore, its appropriateness was questionable.
-The first research work conducted is a statistical analysis to elucidate the bias and the covariance of the estimate that the data-driven step input estimation method provides.
-After obtaining these statistical moments, the estimate uncertainty is built, and the effectiveness of the method is evaluated.
-The second research work conducted is a series of experiments that aims to validate the data-driven step input estimation method in real-life applications.
+This thesis describes research work initiated after the data-driven step input estimation method.
+This method aims for metrology applications but it lacked an uncertainty assessment, and, therefore, its appropriateness was questionable.
+The first research work conducted was a statistical analysis to elucidate the bias and the covariance of the estimate that the data-driven step input estimation method provides.
+After obtaining these statistical moments, the estimation uncertainty was assessed, and the effectiveness of the method was evaluated.
+The second research work was a series of experiments conducted to validate the data-driven step input estimation method in real-life applications.
 The experiments were realizations of the step input excitation using temperature and mass sensors.
-The step input estimation method showed robustness when the measurement noise is not Gaussian and white as it was assumed in the theoretical analyses.
-The third research work conducted is the estimation of affine input parameters following the data-driven approach. 
-The data-driven method is proposed for real-time applications, and uses exponential weighing in the recursive least squares solution of a structured errors-in-variables problem, to give preference to recent samples over the older samples.
-The data-driven affine input estimation method methodology and performance is compared those of a maximum likelihood method based.
+The step input estimation method showed robustness when the measurement noise is not Gaussian and white, as it was assumed in the theoretical analyses.
+The third research work conducted was the estimation of affine input parameters following the data-driven approach for real-time applications. 
+The data-driven method uses exponential weighing in the recursive least-squares solution of a structured errors-in-variables problem, to give preference to recent samples over the older samples.
+The methodology and performance of the data-driven affine input estimation method were compared to those of a maximum likelihood method based.
 
@@
 The data-driven step input estimation method formulates a structured errors-in-variables (EIV) problem.
-In linear estimation EIV problems, the regression matrix and the regressor are perturbed by measurement noise \citep{VanHuffel91Book}, \citep{Markovsky07overview}.
+In linear estimation EIV problems, the measurement noise perturbs the regression matrix and the regressor \citep{VanHuffel91Book}, \citep{Markovsky07overview}.
 The regression matrix, in structured EIV problems, has a structure that depends on the problem formulation.
@@
 The data-driven step input estimation method directly estimates the input from the sensor transient response. 
-The perturbations in the EIV problem of interest come from the sensor output, that is the only observed signal \citep{Markovsky15cep}, and is used to construct both the regression matrix and the regressor.
-The structure in this regression matrix is block-Hankel.
-The direct estimation is a data-driven methodology that reduces the input estimation time of the classical two-stage approach where a sensor model is first identified, and later the input is estimated using the sensor model \citep{Azam15, Niedzwiecki16a}.
+The perturbations in the EIV problem of interest come from the sensor output
+This is the only observed signal \citep{Markovsky15cep}, and it constructs both the regression matrix and the regressor.
+The structure in the regression matrix is block-Hankel.
+The direct estimation is a data-driven methodology that reduces the input estimation time of the classical two-stage approach that first identifies a sensor model and later estimates the input using the sensor model \citep{Azam15, Niedzwiecki16a}.
 
-Instead of using a total least-squares (TLS) solution of the structured EIV problem, the recursive least squares solution (RLS) is proposed to get the step input estimate. 
+Instead of using a total least-squares (TLS) solution of the structured EIV problem, the method proposes the recursive least-squares solution (RLS) to get the step input estimate. 
 The TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
-the TLS solution is equivalent to the maximum likelihood (ML) solution when the perturbations of the EIV problems are i.i.d. normally distributed \citep{Markovsky07overview}. 
-However, regarding structured EIV problems, the TLS estimator does not give general results since each structure requires a specific treatment \citep{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where the global optimum is not guaranteed \citep{Rhode14recursive}.
+the TLS solution is equivalent to the maximum likelihood (ML) solution when the disturbances of the EIV problems are i.i.d. normally distributed \citep{Markovsky07overview}. 
+However, regarding structured EIV problems, the TLS estimator does not give general results since each structure requires a specific treatment \citep{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where finding the global optimum is not guaranteed \citep{Rhode14recursive}.
 Moreover, the computational complexity of TLS and ML inhibits their real-time implementation.
@@
 The uncertainty of an estimation method is expressed using the bias and covariance of the estimate \citep{Pintelon12Book}.
-To know the uncertainty of the data-driven step input estimation method, I quantified the bias and the covariance of the LS solution of EIV problems, for unstructured and structured cases. 
-This quantification extends the perturbation analysis of the LS estimator of unstructured and uncorrelated problems that was investigated in \citep{Stewart90SPT} and in \citep{Vaccaro94}.
-The analysis of the unstructured EIV case provides an insight of the impact that the structure of the regression matrix and the correlation between the regression matrix and the regressor perturbations have on the uncertainty of the LS estimate.
+To know the uncertainty of the data-driven step input estimation method, I quantified the bias and covariance of the LS solution of EIV problems, for unstructured and structured cases. 
+The bias and covariance quantification extend the perturbation analysis that investigated in \citep{Stewart90SPT} and in \citep{Vaccaro94}.
+The analysis of the two studied cases provides insight into the impact that the structure and the correlation have on the LS estimation uncertainty.
 The study presented in this thesis illustrates a methodology to conduct statistical analysis for any structured EIV problem.
 
-I derived expresions that quantify the bias and covariance by obtaining the mathematical expectation of the LS estimate approximated by the second-order Taylor series expansion.
+I derived expressions that quantify the bias and covariance by obtaining the mathematical expectation of the LS estimate approximated by the second-order Taylor series expansion.
 Using Monte Carlo simulations, I validated the accuracy of the bias and covariance expressions.
-These expressions estimate the bias and covariance that the data-driven step input estimation method will exhibit for given sample size and perturbation level.
-I compared the mean squared error of the LS estimate to the minimum variance specified by the Cram\'er-Rao lower bound of the structured EIV problem, to determine the conditions in which the data-driven step input estimation method is appropriate for practical applications \citep{QuintanaCSDA}.
+These expressions estimate the bias and covariance that the data-driven step input estimation method will exhibit for a given sample size and perturbation level.
+I compared the mean squared error of the LS estimate to the minimum variance specified by the Cram\'er-Rao lower bound of the structured EIV problem, to determine the conditions under which the data-driven step input estimation method is appropriate for practical applications \citep{QuintanaCSDA}.
 
@@
 infinite impulse response (IIR) filtering \citep{Pintelon90}, \citep{Elster08}.
-The uncertainty propagation for these model-based compensators is based on the transfer function or state space representations of the LTI sensor and filter systems \citep{Link09}, \citep{Hale09}.
+The uncertainty propagation for these model-based compensators is based on the transfer function or state-space representations of the LTI sensor and filter systems \citep{Link09}, \citep{Hale09}.
 Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \citep{Pietrzak14} for mass and in \citep{Ogorevc16} for temperature sensors.
@@
 To validate the data-driven step input estimation method I built a weighing system setup based on a load cell sensor.
-The weighing setup was constructed to ensure repeatability and reproducibility along different experimental realizations.
-Load cell sensors are versatil devices that are found in 
+The weighing setup was constructed to ensure repeatability and reproducibility, along with different experimental realizations.
+Load cell sensors are versatile devices that are found in 
 heart and breathing physiological signal monitoring \citep{Lee16},
@@
 The signal from the load cell sensor in the weighing system is conditioned using operational amplifiers.
-The signal conditioning amplifier has a low-pass filter to prevent aliasing noise.
+The signal conditioning amplifier has a low-pass filter to prevent aliasing from the noise.
 The observed measurement noise has non-white noise properties, mainly due to the characteristics of the load cell sensor.
-One of the assumptions of the step input estimation method formulation is that the measurement noise is Gaussian white. 
+One of the assumptions of the step input estimation method formulation is that the measurement noise is Gaussian and white. 
 Nevertheless, the estimation results show that the method is still able to provide useful input estimations \citep{QuintanaTIM}.
 
-The empirical bias an covariance obtained after repeating the weighing experiment was compared to the bias and covariance estimations obtained in the previous research work \citep{QuintanaCSDA}, and to the minimal variance given by the  Cram\'er-Rao lower bound (CRLB) of the structured EIV problem.
+The empirical bias and covariance obtained after repeating the weighing experiment was compared to the bias and covariance estimations obtained in the previous research work \citep{QuintanaCSDA}, and to the minimal variance given by the  Cram\'er-Rao lower bound (CRLB) of the structured EIV problem.
 I found that the mean squared error (MSE) of the step input estimate is near the CRLB, and the distance between the MSE and the CRLB provides a confidence interval for the input estimate with respect to the level of the measurement noise.
@@
 The data-driven affine input estimation method formulates a structured errors-invariables (EIV) problem, similar to the one fourmulated for the step input estimation.
-An exponential weight is added to the recursive least squares.
+An exponential weight is added to the recursive least-squares.
 This is a forgetting factor that considers that the newer samples are more relevant for the input parameters estimation.

### 2020-03-01 ChapterPreliminaries/ChapterPreliminaries.tex
Second review version

@@
 \vfill{}
-
-\section{Input estimation is a linear system problem}
 
@@
 
-\subsection{Input estimation with system model}
-
+The input estimation is a linear system problem, and can be done with or without a sensor model.
 The discrete-time state-space representation of an LTI system is
@@
 
-In the next subsections the input is modeled as a multiple of the unit step function, to describe methods to estimate the input step level in two conditions: when the model of the sensor is a-priori given and when the method is not available.
+In the next sections the input is modeled as a multiple of the unit step function, to describe methods for estimating the input step level in two conditions: when the model of the sensor is a-priori given and when the method is not available.
 
-\subsubsection{Step input estimation with system model}
+\section{Step input estimation with system model}
 
-With a step input $\mathbf{u} = \bar{\mathbf{u}} s$, where $s(t) = 1$, for $t \geq 0$, and $s(t) = 0$, for $t < 0$, the discrete-time state space representation of the LTI sensor is equivalent to an augmented autonomous system
+With a step input $\mathbf{u} = \bar{\mathbf{u}} s$, where $s(t) = 1$, for $t \geq 0$, and $s(t) = 0$, for $t < 0$, the discrete-time state-space representation of the LTI sensor is equivalent to an augmented autonomous system
 \begin{equation} \begin{aligned} \mathbf{x}_\text{a}(k+1) &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_\text{a}(k) , \quad \text{where} \ \ \mathbf{x}_\text{a}(k) = \begin{bmatrix} \mathbf{x}(k) \\ \mathbf{u}(k) \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\
@@
 
-\subsection{Step input estimation without system model}
+\section{Step input estimation without system model}
 
@@
 
-\subsubsection{Reduction to autonomous equivalent model}
+%\subsubsection{Reduction to autonomous equivalent model}
 
-One method identifies a model of the sensor from the step response, assuming we have exact data, and later estimates the input step level using the identified sensor model.
+A feasible method can be one that identifies a model of the sensor from the step response, assuming we have exact data, and later estimates the input step level using the identified sensor model.
 The model identification consists in the estimation of the matrices $\widehat{\mathbf{A}}_\text{a}$ and $\widehat{\mathbf{C}}_\text{a}$, and the initial conditions $\widehat{\mathbf{x}}(0)$.
@@
 
-\subsubsection{Identification of the model followed by input estimation}
+%\subsubsection{Identification of the model followed by input estimation}
 
-There is a second method that estimates the step input when the model of the LTI sensor is unknown but its static gain $\mathbf{G}$ is given.
+A second method can estimate the step input when the model of the LTI sensor is unknown but its static gain $\mathbf{G}$ is given.
 Using the static gain $\mathbf{G}$, we can express $\bar{\mathbf{y}} = \mathbf{G} \bar{\mathbf{u}}$, where $\bar{\mathbf{u}}$ is the input exact value and $\bar{\mathbf{y}}$ is the sensor steady state response.
@@
 \[ \underbrace{ \begin{bmatrix}y(0) \\ y(1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \mathbf{C}_\text{a} \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a} \\ \vdots & \vdots \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a}^T \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{\mathbf{u}} \\ \mathbf{x}(0) \end{bmatrix} .\]
-Then, it is necessary to estimate the observability matrix of the augmented system, followed by the estimation of the input  $\mathbf{u}$, and the initial conditions $\mathbf{x}_{\text{ini}}$, using least squares
+Then, it is necessary to estimate the observability matrix of the augmented system, followed by the estimation of the input  $\mathbf{u}$, and the initial conditions $\mathbf{x}_{\text{ini}}$, using least-squares
 \[ \begin{bmatrix} \widehat{\mathbf{u}} \\ \widehat{\mathbf{x}}_{\text{ini}} \end{bmatrix} = \left( \mathbf{K}^\top \mathbf{K} \right)^{-1} \mathbf{K}^\top \mathbf{y}. \]
 
-\subsubsection{Data-driven estimation of an unknown step input given a sensor step response}
+%\subsubsection{Data-driven estimation of an unknown step input given a sensor step response}
 %Another alternative to performing model identification is presented, it is a direct estimation of the input without model identification.
@@
 
-A third method directly estimates the input step level from the step response, without the identification of a sensor model.
-Applying the first difference operator, $\Delta = \sigma - 1$, to the system state space representation (\ref{eqn:dtsslti}), we have
+A third method can directly estimate the input step level from the step response, without the identification of a sensor model.
+Applying the first difference operator, $\Delta = \sigma - 1$, to the system state-space representation (\ref{eqn:dtsslti}), we have
 \begin{equation} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta \mathbf{y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
@@
 that is equivalent to
-\begin{equation} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \mathbf{G} & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ \mathbf{G} & \Delta y(T-n) & \Delta y(T-n+1) & \cdots & \Delta y(T-1) \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} \bar{\mathbf{u}} \\ \bm{\ell} \end{bmatrix} }_{\mathbf{x}} , \label{eqn:ddsiemexd} \end{equation}
+\begin{equation} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \mathbf{G} & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ \mathbf{G} & \Delta y(T-n) & \Delta y(T-n+1) & \cdots & \Delta y(T-1) \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} \bar{\mathbf{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
 where $T$ is the number of samples, and the vector ${\ell}$ is a linear transformation of the system initial conditions \citep{Markovsky15cep}.  
@@
 
-\subsection{Data-driven step input estimation in presence of noise}
+%\subsection{Data-driven step input estimation in presence of noise}
 
@@
 Therefore, the matrix $\widetilde{\mathbf{K}}$ is expressed as
-\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} g & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ g & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ g & \Delta \widetilde{y}(T-n) & \Delta \widetilde{y}(T-n+1) & \cdots & \Delta \widetilde{y}(T-1) \end{bmatrix} .
+\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} \mathbf{G} & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ \mathbf{G} & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ \mathbf{G} & \Delta \widetilde{y}(T-n) & \Delta \widetilde{y}(T-n+1) & \cdots & \Delta \widetilde{y}(T-1) \end{bmatrix} .
 \label{eqn:matrixK} \end{equation}
@@
 The underlying system of equations 
-\begin{equation} \widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \mathbf{x} \label{eqn:ddsiemnd} \end{equation}
+\begin{equation} \widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \bm{\theta} \label{eqn:ddsiemnd} \end{equation}
 is another representation of the minimization problem
-\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert  \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x} \right\Vert^2_2 .
+\begin{equation} \widehat{\bm{\theta}} = \underset{\bm{\theta}}{\mathrm{argmin}} \ \left\Vert  \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \bm{\theta} \right\Vert^2_2 .
 \label{eqn:min_seiv} \end{equation}
 where $\widetilde{\mathbf{y}} = \begin{bmatrix} \widetilde{y}(n+1) & \ldots & \widetilde{y}(T) \end{bmatrix}^\top$, and
-$\widehat{\mathbf{x}} = \begin{bmatrix} \widehat{\mathbf{u}} & \widehat{\bm{\ell}}^\top \end{bmatrix}^\top$. 
+$\widehat{\bm{\theta}} = \begin{bmatrix} \widehat{\mathbf{u}} & \widehat{\bm{\ell}}^\top \end{bmatrix}^\top$. 
 This minimization problem is an errors-in-variables (EIV) problem with Hankel structure, and correlation between the regression matrix $\widetilde{\mathbf{K}}$ and the regressor vector $\widetilde{\mathbf{y}}$.
@@
 The least-squares (LS) approximate solution of this system of equations offers a simple alternative, in its recursive form, to implement the estimation method in real-time.
-The least squares admit a recursive implementation that avoids the inversion of matrix $\widetilde{\mathbf{K}}$ that increases its size with respect to the sample size $T$.
+Least-squares admit a recursive implementation that avoids the inversion of matrix $\widetilde{\mathbf{K}}$ that increases its size with respect to the sample size $T$.
 Instead, the recursive least-squares (RLS) updates the solution of the system of equations considering the previous value of the estimation.

### 2020-03-01 ChapterRampInput/ChapterRampInput.tex
Second review version

@@
 The goal of a measurement is to accurately estimate the input value using the sensor transient response.
-The input might vary during the measurement being the simplest case the input that changes with a constant rate, described with an affine function.
+Here we consider the case where the input might vary during the measurement.
+The simplest case is when the input changes with a constant rate, and is described with an affine function.
 The inputs that vary at a constant rate activate the sensor gradually. 
-An affine input estimation method is needed to estimate the slope and the interception parameters of the affine function.
+An affine input estimation method is needed to estimate the slope and the intercept parameters of the affine function.
 
-This chapter describes two signal processing methods for the estimation of affine input parameters.
+This chapter describes two signal processing methods for estimating affine input parameters.
 These methods are a data-driven subspace method and a maximum-likelihood (ML) estimation method based on local-optimization.
@@
 In this chapter, the dynamic weighing is chosen as an implementation example.
-The dynamic weighing of an affine input is an interesting metrology application because the affine input turns a linear time invariant sensor model into a time-varying sensor model.
+The dynamic weighing of an affine input is an interesting metrology application because the affine input turns a linear time-invariant sensor model into a time-varying sensor model.
 The affine input estimation methods provide an insight on the properties of the dynamic measurement of time-varying inputs.
@@
 The effectiveness and the uncertainty of the methods are evaluated with simulation studies.
-The uncertainty is assessed, for the subspace method, using a Taylor expansion of the exponentially weighted recursive least squares estimate, and
+The uncertainty is assessed, for the subspace method, using a Taylor series expansion of the exponentially weighted recursive least-squares estimate, and
 for the ML method, using the derivatives of the residual error obtained from the to-be-minimized cost function.
@@
 
-\subsection{Affine input estimation problem}
+\section{Affine input estimation problem}
 
-The affine input is modeled as a straight line $u(t) = {a} t + {b}$ with parameters the slope $a$ and the interception $b$.
+The affine input is modeled as a straight line $u(t) = {a} t + {b}$ with parameters the slope $a$ and the intercept $b$.
 The affine input estimation problem is formulated as a signal processing problem as follows.
@@
 \paragraph{Problem} 
-Given the sequence of measured output observations $\widetilde{\mathbf{y}} = \left( \widetilde{y}(1),\ldots, \widetilde{y}(T)\right)$, with $\widetilde{y}(t) \in {\rm I\!R}$, of a stable linear time-invariant system of order $n$, and static gain $\gamma$, generated by an affine input $u(t) = \widebar{a} t + \widebar{b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u}(t) = \widehat{a} t + \widehat{b}$ approximates $u$(t).
+Given the sequence of measured output observations $\widetilde{\mathbf{y}} = \left( \widetilde{y}(1),\ldots, \widetilde{y}(T)\right)$, with $\widetilde{y}(t) \in {\rm I\!R}$, of a stable linear time-invariant system of order $n$, and static gain $\mathbf{G}$, generated by an affine input $u(t) = \widebar{a} t + \widebar{b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u}(t) = \widehat{a} t + \widehat{b}$ approximates $u$(t).
 The measured observations $\widetilde{\mathbf{y}} = \widebar{\mathbf{y}} + \bm{\epsilon}$ are exact sensor responses $\widebar{\mathbf{y}}$ perturbed by additive noise  $\bm{\epsilon}$ assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.
@@
 The weighing of objects in a conveyor belt gives the sensor input an ideal straight line profile when the conveyor belt moves at a constant speed.
-The straight line represents the mass coming gradually into the weighing scale sensor in the conveyor belt.
-The mass can be estimated from the slope $a$ of the straight line model. 
+The straight line represents the mass coming gradually into the weighing scale sensor in the conveyor belt, and
+the mass can be estimated from the slope $a$ of the straight line model. 
 The mechanical vibrations of the conveyor belt perturb the input and the sensor response is affected by measurement noise.
-The interest is to estimate the mass of the object using the sensor response observations.
+Estimating the mass of the object using the sensor response observations is the ultimate goal.
 
 Consider the weighing scale modeled as a second order mass-spring-damper system, such as the one shown in the diagram of Figure \ref{fig:msd_system}.
-The application of an affine input turns the linear time-invariant system into a linear time-varying system, whose
-dynamics depends on the input $u(t) = \widebar{a} t + \widebar{b}$, as it is described by the differential equation:
-\begin{equation} \dfrac{d}{dt} \left( \left( \widebar{a} t + \widebar{b} + m \right) \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = \left( \widebar{a} t + \widebar{b} + m \right) g \end{equation}
-where $m$ is the mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.
-
-The weighing system admits a state space representation where the states $x_1=y$ and $x_2=\dot{y}$ are the position and the speed of the weighing scale: 
-\[ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{\widebar{a} t + \widebar{b} + m} & \dfrac{-(k_{\mathrm{d}} + \widebar{a})}{\widebar{a} t + \widebar{b} + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 \\ g \end{bmatrix},  \quad y = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .
-\]
 
@@
 
-In this paper we use the dynamic weighing example to illustrate the implementation of the affine input estimation methods.
+The application of an affine input turns the linear time-invariant system into a linear time-varying system, whose
+dynamics depends on the input $u(t) = \widebar{a} t + \widebar{b}$, as it is described by the differential equation:
+\begin{equation} \dfrac{d}{dt} \left( \left( \widebar{a} t + \widebar{b} + m \right) \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = \left( \widebar{a} t + \widebar{b} + m \right) g \end{equation}
+where $m$ is the mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.
+
+The weighing system admits a state-space representation where the states $x_1=y$ and $x_2=\dot{y}$ are the position and the speed of the weighing scale: 
+\begin{equation} \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{\widebar{a} t + \widebar{b} + m} & \dfrac{-(k_{\mathrm{d}} + \widebar{a})}{\widebar{a} t + \widebar{b} + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 \\ g \end{bmatrix}, \quad y = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .
+\label{eqn:ltvsystem} \end{equation}
+
+In this chapter we use the dynamic weighing example to illustrate the implementation of the affine input estimation methods.
 
@@
 
-In this section we describe the proposed subspace method to solve the affine input estimation problem.
-The method is motivated by the step input estimation method that is formulated as a structured errors-in-variables (EIV) problem and is solved using recursive least squares (RLS).
-The exponentially weighted recursive least squares (EWRLS) is a generalization of RLS that allows the extension of the estimation method to reconstruct the affine input.
+This section describes the subspace method for solving the affine input estimation problem.
+The procedure is motivated by the step input estimation method that is formulated as a structured errors-in-variables (EIV) problem and is solved using recursive least-squares (RLS).
+The exponentially weighted recursive least-squares (EWRLS) is a generalization of RLS that allows for the extension of the estimation method to reconstruct the affine input.
 
-A maximum-likelihood (ML) method that performs simultaneous system identification and input estimation is described and its results are used as reference to compare the subspace method results.
+A maximum-likelihood (ML) method that performs simultaneous system identification and input estimation is described and its results are used as a reference for the subspace method results.
 
@@
 \subsection{Subspace method}
+% \linebreak
+The errors-in-variables (EIV) minimization problem (\ref{eqn:min_seiv}) is extended to estimate the affine input parameters.
+The profile of an affine input excitation can be reconstructed by solving the structured EIV problem  
+\begin{equation} \widehat{\bm{\theta}} = \underset{\bm{\theta}}{\mathrm{argmin}} \ \left\Vert \bm{\Lambda}^{1/2} \left( \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \bm{\theta} \right) \right\Vert^2_2 .
+\label{eqn:min_ewrls} \end{equation}
+where $\widetilde{\mathbf{y}}$ is the observed affine response, $\widetilde{\mathbf{K}}$ is defined in (\ref{eqn:K0plusnoise}), and
+the matrix $\bm{\Lambda} \in \mathbb{R}^{(T-n) \times (T-n)}$ is a diagonal matrix of descending powers of the weight factor $\lambda \in [0, 1)$, \textit{i.e.},  $\bm{\Lambda} = \mathrm{diag}\left(\lambda^{T-n}, \ldots, \lambda^2, \lambda^1 \right)$.
+The weight factor $\lambda$ is a data selection forgetting factor since it enables to apply different weights to the residuals $\widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \bm{\theta}$.
 
-RLS is a special case of the EWRLS that can solve the errors-in-variables (EIV) minimization problem (\ref{eqn:min_seiv}).
-The minimization problem is a structured EIV problem and its EWRLS solution is equivalent to 
-\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert \mathbf{W}^{1/2} \left( \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x} \right) \right\Vert^2_2 .
-\label{eqn:min_ewrls} \end{equation}
-where $\widetilde{\mathbf{y}}$, $\widehat{\mathbf{x}}$, and $\widetilde{\mathbf{K}}$ are defined in (\ref{eqn:min_seiv}) and (\ref{eqn:K0plusnoise}) and
-$\mathbf{W} \in \mathbb{R}^{(T-n) \times (T-n)}$ is a diagonal matrix of descending powers of the weight $\lambda \in [0, 1)$, \textit{i.e.}, $\mathbf{W} = \mathrm{diag}\left(\lambda^{T-n}, \ldots, \lambda^2, \lambda^1 \right)$.
-The weight $\lambda$ is a data selection forgetting factor since it enables to apply different weights to the residuals $\widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x}$.
-When $\lambda=1$, we have the same solution as RLS.
+The EWRLS solution to the problem (\ref{eqn:min_ewrls}) is implemented to facilitate the online implementation of the affine input estimation method.
+When $\lambda=1$, the EWRLS solution is identical to that of the RLS.
 When $\lambda<1$, the older residuals are weighted with lower values than the residuals of recent observations.
-In this way, the solution of the minimization problem depends more on new data.
+In this way, the solution of the minimization problem depends more on newer data, and less on older data.
 
-The profile of an affine input excitation can be reconstructed by solving the structured errors-in-variables problem  (\ref{eqn:min_ewrls}) where $y$ is the corresponding response of a stable dynamic system, $\widetilde{\mathbf{K}}$ is constructed from the output observations, and $\lambda<1$.
+In Figure \ref{fig:sensor_response} we can see an example of the affine input $\mathbf{u}$ that excites a weighing sensor, and the generated response $\widetilde{\mathbf{y}}$.
+After the transient effects have diminished, the sensor response is a ramp of the same slope as the input, but with a different intercept due to an offset $\tau$.
+The offset $\tau$ is explained by the parameter $m$ in the sensor representation given in Figure \ref{fig:msd_system}.
+This is the mass of the sensor itself, that is added to the mass of the object under measurement.	
+From the point of view of the subspace method, the value of $m$ is unknown, and there is no interest in its estimation.
 
-After the affine input $\widehat{\mathbf{u}}$ has been estimated, the affine input parameters $a$ and $b$ are estimated by fitting $\widehat{\mathbf{u}}$ to a straight line using linear regression, as follows
+The EWRLS solution to the problem (\ref{eqn:min_ewrls}) provides an affine input  estimation $\widehat{\mathbf{u}}$. 
+The affine input parameters $a$ and $b$ are estimated, in a second step, by fitting $\widehat{\mathbf{u}}$ to a straight line using linear regression, as follows
 \begin{equation} 
-  \begin{bmatrix} 1 & 1  \\ \vdots & \vdots \\ T-\tau+1 & 1 \end{bmatrix} 
+  \begin{bmatrix} 1 & 1  \\ \vdots & \vdots \\ T & 1 \end{bmatrix} 
   \begin{bmatrix} \widehat{a} \\ \widehat{b}  \end{bmatrix} = 
-  \begin{bmatrix} \widehat{u}(\tau) \\ \vdots \\ \widehat{u}(T) \end{bmatrix} \label{eqn:LS}
+  \begin{bmatrix} \widehat{u}(1)-\tau \\ \vdots \\ \widehat{u}(T)-\tau \end{bmatrix}, \label{eqn:LS}
 \end{equation}
-where $\tau$ is a number of samples used as a tuning parameter that counteracts the time delay that exists when a LTI system is excited with an affine input $\mathbf{u}$.
-The subspace method can process online the sensor response to the affine excitation.
-For each new observation $y(t)$, the estimation $\widehat{\mathbf{u}}$ is updated followed by the update of the slope $\widehat{a}$ and the interception $\widehat{b}$.
+where it is considered the existance of the offset $\tau$.
+
+For each new observation $y(t)$, the estimation $\widehat{\mathbf{u}}$ is updated followed by the RLS update of the slope $\widehat{a}$ and the intercept $\widehat{b}$.
 The values of the tuning parameters $\lambda$ and $\tau$ can be obtained in the calibration of the method using the response of the sensor.
+
+The calibration procedure consists in the measurement of an object with a reference mass.
+During the calibration, $\lambda$ and $\tau$ are tuned by changing their values aiming to reduce the estimation errors of $\widehat{a}$ and $\widehat{b}$.
+This reduction of the estimation errors is observed in the convergence of the affine input parameters towards their true values.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_2_1.pdf} 
+\caption{ \label{fig:sensor_response} When the transient part of the sensor response $\widetilde{\mathbf{y}}$ asymptotically decreases, a ramp with the same slope $a$ of the  affine input $u(t) = at+b$ predominates.
+The intercept of the prevailing ramp is the intercept $b$ added to and offset $\tau$.}
+\end{figure}
 
@@
 To obtain the first and second moments of the step input estimate $\widehat{\mathbf{u}}$, we need to study the solution
-\begin{equation} \widehat{\mathbf{x}} = ( \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{K}}  )^{-1} \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{y}} , \label{eqn:xhatW} \end{equation}
+\begin{equation} \widehat{\bm{\theta}} = ( \widetilde{\mathbf{K}}^\top \bm{\Lambda} \widetilde{\mathbf{K}}  )^{-1} \widetilde{\mathbf{K}}^\top \bm{\Lambda} \widetilde{\mathbf{y}} , \label{eqn:xhatW} \end{equation}
 of the overdetermined structured errors-in-variables EIV problem (\ref{eqn:min_ewrls}).
 Using a second order Taylor series expansion of the inverse matrix we can approximate the LS solution as
-\begin{equation} \widehat{\mathbf{x}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{C}^{-1} (\mathbf{K} + \mathbf{E})^\top \mathbf{W} (\mathbf{y} + \bm{\epsilon}).
+\begin{equation} \widehat{\bm{\theta}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{C}^{-1} (\mathbf{K} + \mathbf{E})^\top \bm{\Lambda} (\mathbf{y} + \bm{\epsilon}).
 \label{eqn:xhatexpW} \end{equation}
 where 
-\begin{equation} \mathbf{C} = \mathbf{K}^\top \mathbf{W} \mathbf{K}, \quad \text{and} \quad \mathbf{M} = \mathbf{C}^{-1} ( \mathbf{K}^\top \mathbf{W} \mathbf{E} + \mathbf{E}^\top \mathbf{W} \mathbf{K} + \mathbf{E}^\top \mathbf{W} \mathbf{E} ).
+\begin{equation} \mathbf{C} = \mathbf{K}^\top \bm{\Lambda} \mathbf{K}, \quad \text{and} \quad \mathbf{M} = \mathbf{C}^{-1} ( \mathbf{K}^\top \bm{\Lambda} \mathbf{E} + \mathbf{E}^\top \bm{\Lambda} \mathbf{K} + \mathbf{E}^\top \bm{\Lambda} \mathbf{E} ).
 \end{equation}
 
-The Taylor series approximation of $\widehat{\mathbf{x}}$ enables the calculation of the estimate bias and covariance since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion.
-The bias and the covariance of the estimate $\widehat{\mathbf{x}}$ are obtained from the definitions (\ref{eqn:biasdef}), and (\ref{eqn:covdef}).
-%The bias and the covariance of the estimate $\widehat{\mathbf{x}}$ are obtained from \begin{equation}  \mathbf{b} \left(\widehat{\mathbf{x}} \right) = \mathbf{\mu} - \widebar{\mathbf{x}}, \label{eqn:biasdef} \end{equation}  \begin{equation} \begin{aligned} \mathrm{\mathbf{Cov}} \left( \widehat{\mathbf{x}} \right) & = \mathbb{E} \left\{ \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)  \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)^\top \right\} .
-\end{aligned} \label{eqn:covdef} \end{equation} where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}$ is the expected value of the estimate and $\widebar{\mathbf{x}}$ is the true value.
+The Taylor series approximation of $\widehat{\bm{\theta}}$ enables the calculation of the estimation bias and covariance since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion.
+The bias and the covariance of the estimate $\widehat{\bm{\theta}}$ are obtained from the definitions (\ref{eqn:biasdef}), and (\ref{eqn:covdef}).
+%The bias and the covariance of the estimate $\widehat{\bm{\theta}}$ are obtained from \begin{equation}  \mathbf{b} \left(\widehat{\bm{\theta}} \right) = \mathbf{\mu} - \widebar{\bm{\theta}}, \label{eqn:biasdef} \end{equation}  \begin{equation} \begin{aligned} \mathrm{\mathbf{Cov}} \left( \widehat{\bm{\theta}} \right) & = \mathbb{E} \left\{ \left( \widehat{\bm{\theta}} - \mathbf{\mu} \right)  \left( \widehat{\bm{\theta}} - \mathbf{\mu} \right)^\top \right\} .
+\end{aligned} \label{eqn:covdef} \end{equation} where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\}$ is the expected value of the estimate and $\widebar{\bm{\theta}}$ is the true value.
 Considering the structure of the EIV problem, the bias and the covariance of the estimate approximation (\ref{eqn:xhatexpW}) can be expressed as
-\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{C}^{-1} \left(  \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_1 - \mathbf{B}_2 \right) \mathbf{x} - \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasEW} \end{equation}
-\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{K}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{W} \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{x} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varEW} \end{equation}
-where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \Big\}$, $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger \mathbf{W}$, and $\mathbf{K}^\dagger$ is the pseudo-inverse matrix of $\mathbf{K}$.
+\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{C}^{-1} \left(  \left( \mathbf{K}^\top \bm{\Lambda} \mathbf{B}_1 - \mathbf{B}_2 \right) \bm{\theta} - \left( \mathbf{K}^\top \bm{\Lambda} \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasEW} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \bm{\Lambda} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \bm{\Lambda} \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right), \end{aligned} \label{eqn:varEW} \end{equation}
+where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\Lambda} \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Lambda} \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\Lambda} \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Lambda} \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \Big\}$, $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger \bm{\Lambda}$, and $\mathbf{K}^\dagger$ is the pseudo-inverse matrix of $\mathbf{K}$.
 
-The bias and covariance given by expressions (\ref{eqn:biasEW}) and (\ref{eqn:varEW}) depend on the unobservable true values $\mathbf{x}$ and $\mathbf{K}$.
-The measured observations are in the sensor step response $\widetilde{\mathbf{y}}$, and from its observations we construct $\widetilde{\mathbf{K}}$ and compute $\widehat{\mathbf{x}}$.
-The substitution of the measured data in the expressions gives an approximation of the estimate bias and covariance. 
+The bias and covariance given by expressions (\ref{eqn:biasEW}) and (\ref{eqn:varEW}) depend on the unobservable true values $\bm{\theta}$ and $\mathbf{K}$.
+The measured observations are in the sensor step response $\widetilde{\mathbf{y}}$, and from its observations we construct $\widetilde{\mathbf{K}}$ and compute $\widehat{\bm{\theta}}$.
+The substitution of the measured data in the expressions gives an approximation of the estimation bias and covariance. 
 We have then
-\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{C}}^{-1} \left(  \left( \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\mathbf{x}} - \left( \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasSTW} \end{equation}
-  \begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \mathbf{W} \widetilde{\mathbf{K}}^{\dagger \top} - \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{x} \right) \widetilde{\mathbf{b}}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varSTW} \end{equation}
-  where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \bm{\epsilon}^\top \Big\}$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger \mathbf{W}$.
+\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{C}}^{-1} \left(  \left( \widetilde{\mathbf{K}}^\top \bm{\Lambda} \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\bm{\theta}} - \left( \widetilde{\mathbf{K}}^\top \bm{\Lambda} \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasSTW} \end{equation}
+  \begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \bm{\Lambda} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \bm{\Lambda} \widetilde{\mathbf{K}}^{\dagger \top} - \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \widetilde{\mathbf{b}}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right), \end{aligned} \label{eqn:varSTW} \end{equation}
+  where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\Lambda} \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Lambda} \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\Lambda} \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Lambda} \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \bm{\epsilon}^\top \Big\}$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger \bm{\Lambda}$.
 
-The results of the expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$ can be found using Lemma \ref{lem:lemma1}.
-The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \citep{Vaccaro94, Stewart90SPT}.
+The results of the expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$ can be found using Lemma \ref{lem:lemma1}, on page \pageref{lem:lemma1}.
+The bias and covariance were obtained to extend the previous analysis conducted on EIV estimation problems without an imposed structure \citep{Vaccaro94, Stewart90SPT}.
 It was shown that the bias and variance expressions (\ref{eqn:biasSTW}) and (\ref{eqn:varSTW}) are valid predictions of the first and second moments of the LS estimate of a Hankel structured EIV problem.
-The problem formulated by the step input estimation method belongs to this type of structured EIV problems and can use the derived expressions to find the bias and variance of the input estimate $\widehat{u}$.
-The bias of the estimate $\widehat{\mathbf{u}}$ is the fist element of $\widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right)$ and the variance of $\widehat{\mathbf{u}}$ is the first element in the main diagonal of $\widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right)$.
+The problem formulated by the step input estimation method belongs to this type of structured EIV problems and we can use the derived expressions to find the bias and variance of the input estimate $\widehat{u}$.
+The bias of the estimate $\widehat{\mathbf{u}}$ is the fist element of $\widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right)$ and the variance of $\widehat{\mathbf{u}}$ is the first element in the main diagonal of $\widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right)$.
 
 \subsection{Maximum-likelihood method}
-Using a model of the sensor, the ML method estimates the sensor initial conditions and the parameters of the applied affine input.
-The affine input parameters $a$ and $b$ are the slope and the interception of the straight line model $u(t) = at + b$.
+Using a model of the sensor, the ML method simultaneously estimates the parameters of the applied affine input $u(t) = at + b$, and some of the sensor model parameters.
+This means that the ML method implementation depends on the problem under study and not always can be estimated all the sensor model parameters.  
+The identifiability conditions of each sensor model parameter is the requirement to fulfill in order to be estimated with the ML method.
+%The affine input parameters $a$ and $b$ are the slope and the intercept of the straight line model $u(t) = at + b$.
 
-Algorithm \ref{Alg1} lists the steps of the proposed ML method.
-This is an iterative minimization method that uses the Jacobian of the residual error function $\mathbf{r} = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}$ to search the direction that minimizes the difference between the measured sensor response $\widetilde{\mathbf{y}}$ and the simulated response $\widehat{\mathbf{y}}$.
+The ML method is an iterative minimization method, described in Algorithm \ref{Alg1}, that minimizes the residual $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$.
+This residual is the difference between the measured sensor response $\widetilde{\mathbf{y}}$ and a simulated response $\widehat{\mathbf{y}}(\bm{\theta})$, that depends on the optimization variables $\bm{\theta}$.
+The ML method uses the Jacobian matrix of the residual $\mathbf{r}$ to search the direction in which the residual decreases.
+The Jacobian matrix $\partial \mathbf{r} / \partial \bm{\theta}$ can be analytically obtained using the sensor mathematical model.
 
@@
 \begin{algorithmic}
-  \REQUIRE{$\widetilde{\mathbf{y}}$, and sensor model parameters}{}
-  \STATE {Initialize $\mathbf{\theta} = (a, b, \mathbf{x}_{\text{ini}})$}{}
+  \REQUIRE{$\widetilde{\mathbf{y}}$, and required sensor model parameters}{}
+  \STATE {Initialize $\bm{\theta} = (a, b, \mathrm{and the to-be-estimated \ model \ parameters})$}{}
   \FOR{each $N$ observations of $\widetilde{\mathbf{y}}$}{}
-  \STATE {Simulate model response $\widehat{\mathbf{y}}(\mathbf{\theta})$}{}
-  \STATE {Compute error $\mathbf{r}(\mathbf{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\mathbf{\theta})$}{}
-  \STATE {Minimize $\mathbf{r}^\top \mathbf{r}$ over $\mathbf{\theta}$}{}
-  \STATE {\hspace{0.5cm}  using analytic Jacobian $\partial \mathbf{r} / \partial \mathbf{\theta}$}{}
-  \STATE {Update $\mathbf{\theta}$}{}
+  \STATE {Simulate model response $\widehat{\mathbf{y}}(\bm{\theta})$}{}
+  \STATE {Compute error $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$}{}
+  \STATE {Minimize $\mathbf{r}^\top \mathbf{r}$ over $\bm{\theta}$}{}
+  \STATE {\hspace{0.5cm}  using analytic Jacobian matrix $\mathbf{J} = \partial \mathbf{r}(\bm{\theta}) / \partial \bm{\theta}$}{}
+  \STATE {Update $\bm{\theta}$}{}
 \ENDFOR
-\ENSURE {Optimized parameters $\widehat{a}, \widehat{b}$, and $\widehat{\mathbf{x}}_{\text{ini}}$}{}
+\ENSURE {Optimized parameters $\widehat{a}, \widehat{b}$, and estimated model parameters.} }{}
 
@@
 \end{algorithm}
-
-To initialize the optimization variables $a$, $b$, and $\mathbf{x}_{\text{ini}}$,
-we use the subspace estimation method using at least the first $2n+2$ samples.
-With the initial affine input parameters we simulate a sensor excitation and, since we are using few samples, we have an approximation of the initial conditions.
-The optimization variables updates are computed every $N$ new observations.
-The minimization can be done using the Levenberg-Marquardt algorithm \citep{Nocedal06}.
 
@@
 The covariance matrix of the ML estimates can be expressed as \citep{Pintelon12Book}
-\begin{equation} \mathbf{Cov} \left( \widehat{x} \right) = \left( \left( \dfrac{\partial \mathbf{r} }{ \partial \mathbf{\theta} } \right)^\top \left( \dfrac{\partial \mathbf{r} }{ \partial \mathbf{\theta} } \right) \right)^{-1}.
+\begin{equation} \mathbf{Cov} \left( \widehat{\bm{\theta}} \right) = \sigma_{\bm{\epsilon}}^2 \left( \left( \dfrac{\partial \mathbf{r} }{ \partial \bm{\theta} } \right)^\top \left( \dfrac{\partial \mathbf{r} }{ \partial \bm{\theta} } \right) \right)^{-1}.
 \label{eqn:covOpt} \end{equation}
 
+The ML estimation method is asymptotically efficient, and the ML estimation variance values are the mininum values that any estimator can attain.
+Therefore, the variance of the ML method estimates should be smaller than the CRLB of the structured EIV minimization problem (\ref{eqn:FIM}).
+
 \subsubsection{ML affine input estimation example}
-We use the previously described dynamic weighing system to illustrate the ML method.
-In this case, the formulation of the problem is: 
+
+The previously described dynamic weighing system is used to illustrate the ML method implementation.
+In this case, the problem can be formulated as: 
  \begin{equation} \begin{aligned}
-     & \text{Minimize} \quad \text{over} \ a, b, \mathbf{x}_{\text{ini}} \quad \mathbf{r}^T \mathbf{r} \text{, subject to:} \\ & \ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 & \frac{x_{\text{ini,1}}}{a t + b + m} \\ g & \frac{x_{\text{ini,2}}}{a t + b + m}  \end{bmatrix} \begin{bmatrix} 1 \\ \delta(t)  \end{bmatrix}, \\ & \ \widehat{\mathbf{y}} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .
- \end{aligned} \end{equation}
-where $a$, $b$ are the affine input parameters, $m$, $k_{\mathrm{d}}$, and $k_{\mathrm{s}}$ are the model parameters, the model states $\mathbf{x}$ are the position and the speed of the weighing scale, the difference between the sensor response $\widetilde{\mathbf{y}}$ and the simulated response $\widehat{\mathbf{y}}$ is the residual $\mathbf{r} = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}$.
-The model initial conditions $\mathbf{x}_{\text{ini}}$ are considered optimization variables
-and appear in the augmented column of the input matrix.
+     & \text{Minimize} \quad \text{over} \ a, b, \mathbf{x}_{\text{ini}} \quad \mathbf{r}^T \mathbf{r} \text{, subject to:} \\ & \ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 \\ g \end{bmatrix}, \\ & \ \widehat{\mathbf{y}} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .
+ \label{eqn:ltv_abxini} \end{aligned} \end{equation}
+where the optimization variables $\bm{\theta} = \left(a, b, \mathbf{x}_{\mathrm{ini}} \right)$ are the affine input parameters $a$, $b$, and the sensor model initial conditions.
 
-The analytic Jacobian for the weighing system example is described in the appendix.
+The model parameters $m$, $k_{\mathrm{d}}$, and $k_{\mathrm{s}}$ cannot be estimated simultaneously due to identifiability issues.
+In the state equation of the time-varying sensor model, the parameters $k_{\mathrm{d}}$ and $a$ are entangled and cannot be explicitly separated.
+The same occurs for the parameters $m$ and $b$, and for $k_{\mathrm{s}}$ with respect to $a$ and $b$.
+
+To define the initial value of the optimization variables $a$, $b$,
+the subspace estimation method can be used with, at least, the first $2n+2$ transient response samples, where $n=2$ is the order of the sensor model.
+With the initial affine input parameters, a sensor response can be simulated, and, since we are using few samples, the first samples of the simulated response also are an approximation of the initial conditions $\mathbf{x}_{\mathrm{ini}}$.
+The optimization variables are updated every $N$ new observations.
+The minimization algorithm can be, for example, the Levenberg-Marquardt algorithm \citep{Nocedal06}.
+
+The analytic Jacobian matrix $\mathbf{J}$ of the residual $\mathbf{r}(\bm{\theta})$ can be obtained from the first derivative of the state-space representation (\ref{eqn:ltv_abxini}), with respect to the optimization variables $\bm{\theta}$.
+The residual $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$ is the error difference between the observed sensor response $\widetilde{\mathbf{y}}$ and the simulated sensor response $\widehat{\mathbf{y}}(\bm{\theta})$ to the affine input.
+More details of the Jacobian matrix calculation, for the sensor model under study, are given in the appendix.
 
 \subsection{Time-varying compensation filter}
+
 The time-varying (TV) filter described in \citep{Pietrzak14} was designed to compensate the measured responses of a conveyor weighing system, considering they are modeled as a saturated ramp.
@@
 
-The time-varying coefficients $k_1(t)$ and $k_2(t)$ are computed from
+The time-varying coefficients $k_1(t)$ and $k_2(t)$ are computed from \citep{Pietrzak14}
 \begin{equation} k_1(t) = \frac{f_c(t) - \frac{k_3}{\pi T_s}}{f_c(t) + \frac{k_3}{\pi T_s}}, \quad k_2(t) = \frac{1 + k_1(t)}{2} , \quad k_3 = \sqrt{ \sqrt[3]{2}  - 1} \end{equation}
 where $T_s$ is the sampling time and $f_c(t)$ is a heuristic "cutoff" frequency
-\begin{equation} f_c(t) = f_u + \left( f_l - f_u \right) \beta^{\left( t-1 \right) /  \alpha \left( T-1 \right) } \end{equation}
+\begin{equation} f_c(t) = f_u + \left( f_l - f_u \right) \beta^{\frac{t-1}{\alpha \left( T-1 \right)} } \end{equation}
 that changes between the lower $f_l$ and upper $f_u$ limits, where the coefficient $\beta$ is lower than one, and $\alpha$ is the decay rate. 
@@
 A second order weighing system was excited with an affine input to get the transient response.
-The parameters of the weighing system are $m = 15$ g, $d = 5.5$ Ns/m and $k = 10250$ N/m.
+The parameters of the weighing system are $m$ = 15 g, $k_{\mathrm{d}$ = 5.5 Ns/m and $kg_{\mathrm{s}$ = 10250 N/m.
 The applied affine input $u(t) = 100 t + 10$ represents a mass that changes from $10$ g to $110$ g, at a constant rate, in a time interval of $0.1$ s.
 This change of mass represents one example of the weighing input in a conveyor weighing system when an object of 100 g is measured while it is moving at constant speed.
-The sensor response is acquired using sampling time $T_s = 0.1$ ms.
-In Figure \ref{fig:sensor_weight} the input $u$ is represented with the dotted line, the oscillatory curve is the corresponding sensor ramp response $\widetilde{\mathbf{y}}$, and $\widehat{\mathbf{u}}_{\mathrm{}}$ is a typical input estimate obtained with the subspace method.
+A total of 1000 samples of the sensor response are acquired with sampling time $T_s = 0.1$ ms.
+In Figure \ref{fig:sensor_weight} the input $\mathbf{u}$ is represented with the black dotted line, the red oscillatory curve is the corresponding sensor ramp response $\widetilde{\mathbf{y}}$, and the blue curve $\widehat{\mathbf{u}}_{\mathrm{}}$ is a typic al input estimate obtained with the subspace method.
 
@@
 In the figure we observe an example of the input estimate $\widehat{\mathbf{u}}$ obtained with the subspace method.
-The input parameters are calculated from $\widehat{\mathbf{u}}$ using linear regression.}
+The input parameters are calculated from $\widehat{\mathbf{u}} - \tau$ using linear regression.}
 \end{figure}
@@
 In each simulation, the sensor response was perturbed with an independent realization of additive normally distributed measurement noise.
-The added perturbation noise has signal-to-noise ratios (SNR) in the interval [20 dB, 60 dB], values that are realistic in practical applications.
-The SNR is defined as the ratio of signal power to the noise power, that is equivalent to the root-mean-square value of the true signal to the variance of the perturbation noise, and in dB is given as
+The added perturbation noise has signal-to-noise ratio (SNR) in the interval [20 dB, 60 dB].
+These SNR values are realistic in practical applications.
+The SNR is defined as the ratio of signal power to the noise power, that is equivalent to the root-mean-square (RMS) value of the true signal to the standard deviation of the perturbation noise, and in dB is given as
 
-\begin{equation} \mathrm{SNR} = 20 \log_{10}{\dfrac{\dfrac{1}{T}\int\limits_0^T{\widebar{y}(t)^2 d t}}{\sigma_{\bm{\epsilon}}}} \end{equation}
+\begin{equation} \mathrm{SNR} = 20 \log_{10}{\dfrac{ \sqrt{ \dfrac{1}{T}\int\limits_0^T{\widebar{y}(t)^2 d t} } } { \sigma_{\bm{\epsilon}}}} \end{equation}
 
@@
 The subspace method processed online the sensor transient response.
-The first estimation was obtained with $2n+1$ samples.
+The first estimation was obtained with $2n+1$ samples, where $n=2$ is the assumed order of the sensor.
 The method updated recursively the value of the estimated parameters for each new collected sample, using the forgetting factor $\lambda$ listed in Table \ref{table:lambdas}.
@@
 \centering
-\caption{The values of the forgetting factor $\lambda$ and the samples shift $\tau$ that configure the subspace method for the different values of SNR.
-These values were obtained after calibration of the method and fixed during the simulation study.}
+\caption{Here are listed the selected values of the forgetting factor $\lambda$ and the offset $\tau$ that configure the subspace method for the different values of SNR.
+These values were obtained after calibration of the method and were fixed during the simulation study.}
 \begin{tabular}{c| c c c c c} 
@@
  $\lambda$ & 0.939 & 0.940 & 0.955 & 0.959 & 0.959 \\ % forgetting factor
- $\tau$ \  & 15 & 15 & 17 & 14 & 20 \\ [0.5ex] % [\mathrm{samples}]
+ $\tau$ [g] \  & 15 & 15 & 17 & 14 & 20 \\ [0.5ex] % [\mathrm{samples}]
  \hline
@@
 \includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_3.pdf} 
-\caption{ \label{fig:rele_dd_40dB_s1} The relative errors of the affine input parameters estimation decrease as the subspace method processes more samples.
+\caption{ \label{fig:rele_dd_40dB_s1} The relative errors of the affine input parameters estimates decrease as the subspace method processes more samples.
 The relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ are smaller than 5 \% after 400 and 500 samples, respectively ($T_s = 0.1$ ms).
@@
 The Cram\'er-Rao lower bound (CRLB) of the errors-in-variables problem formulated by the subspace method was numerically computed for different sample size using Equation (\ref{eqn:FIM}).
-The CRLB is the minimum variance that the estimates $\widehat{a}$ and $\widehat{b}$ can have.
+The CRLB is the minimum variance that the estimates $\widehat{a}$ and $\widehat{b}$ can have from the solution of the structured EIV minimization problem. 
 The average of $10^4$ runs with independent noise realizations allows to find the empirical mean squared error (MSE) of the estimates, defined as
@@
 \includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_5.pdf} 
-\caption{ \label{fig:CRLB_MSE_ab_dd_40dB_MC_10000} When the SNR of the sensor response is 40 dB, the mean squared errors of the slope estimate $\widehat{a}$ and the interception estimate $\widehat{b}$, obtained by the subspace method, are two orders of magnitude above the theoretical minimum variance given by the Cram\'er-Rao lower bound.}
+\caption{ \label{fig:CRLB_MSE_ab_dd_40dB_MC_10000} When the SNR of the sensor response is 40 dB, the mean squared errors of the slope estimate $\widehat{a}$ and the intercept estimate $\widehat{b}$, obtained by the subspace method, are two orders of magnitude above the theoretical minimum variance given by the Cram\'er-Rao lower bound.}
 \end{figure}
@@
 According to these results, the input parameters uncertainties $\sigma_{a,b}$ affect more the uncertainty of the estimation than the speed uncertainty $\sigma_{\mathrm{s}}$. 
-The parameter that is more affected by the input parameters uncertainty is the interception $\widehat{b}$, since the uncertainty of the slope $\widehat{a}$ is smaller.
+The parameter that is more affected by the input parameters uncertainty is the intercept $\widehat{b}$, since the uncertainty of the slope $\widehat{a}$ is smaller.
 
@@
 The speed $\sigma_{\mathrm{s}}$, and the input parameters $\sigma_{\mathrm{s}}$ uncertainties are 0\%, 5\%, and 10\% of their true values.
-The table shows the maximum values of the estimates uncertainty.
+The table shows the maximum values of the estimation uncertainty.
 The speed uncertainty causes a smaller spread of the estimates than the input parameters uncertainty.}
@@
 \subsection{Results of the maximum-likelihood method}
-The maximum-likelihood (ML) method processed off-line the sensor transient response.
-The ML method used the first 50 samples to initialize the optimization variables and updated the variables every $N = 1$ sample.
+
+The maximum-likelihood (ML) method processed off-line the sensor transient response, perturbed with measurement noise of 40 dB SNR.
+The ML method used the first 50 samples to initialize the optimization variables and updated the variables every $N = 5$ samples.
 In Figure \ref{fig:rele_dd_40dB_s1} are shown the relative errors of the estimates $\widehat{a}$, $\widehat{b}$, $\widehat{x}_{\mathrm{ini,1}}$, and $\widehat{x}_{\mathrm{ini,2}}$.
@@
 
-Using the analytic Jacobian and Equation (\ref{eqn:covOpt}), we computed the covariance of the estimate.
-In Figure \ref{fig:cov_lo_40dB_s1} are shown the variances of the optimization variables taken from the diagonal of the covariance matrix. 
-We can see that the variances of $\widehat{a}$ and $\widehat{b}$ decrease faster than the variances of $\widehat{x}_{\mathrm{ini,1}}$ and $\widehat{x}_{\mathrm{ini,2}}$ as more samples are processed.
+The estimation covariance was computed using the analytic Jacobian and Equation (\ref{eqn:covOpt}).
+Figure \ref{fig:cov_lo_40dB_s1} shows the variances of the optimization variables, that are on the diagonal of the covariance matrix $\mathbm{J}$. 
+We can see that the estimation variance decrease as more samples are processed.
+Moreover, the estimation variances of $\widehat{a}$ and $\widehat{b}$ obtained with the ML method are lower than 
+the corresponding estimation MSE errors obtained from a Monte Carlo simulation of the subspace method (see Figure \ref{fig:CRLB_MSE_SNR_ab_dd_MC_10000}).
 
@@
 Figure \ref{fig:rele_tv_40dB_s1} shows the relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ computed with the TV filter after processing the sensor transient response.
-The relative error of the slope estimate is below 5 \% after 300 samples but the relative error of the interception estimate is near 10 \%.
+The relative error of the slope estimate is below 5 \% after 300 samples but the relative error of the intercept estimate is near 10 \%.
 The convergence rate of the estimate $\widehat{a}$ was similar to that of the subspace method.
@@
 
-The subspace method obtains an estimation of the affine input parameters with a recursive least squares solution of a structured errors-in-variables problem.
+The subspace method obtains an estimation of the affine input parameters with a recursive least-squares solution of a structured errors-in-variables problem.
 Updating the parameter estimates without matrix inversion simplifies the method implementation on digital signal processors of low cost.
-The price we pay by computing the least squares solution of an errors-in-variables problem is an increase in the bias of the estimates.
-Nevertheless, the empirical mean squared errors of the estimates are at most two orders of magnitude larger than the Cram\'er-Rao lower bound, meaning that the estimates uncertainty is low, even when the SNR is lower than 40 dB.
+The price we pay by computing the least-squares solution of an errors-in-variables problem is an increase in the bias of the estimates.
+Nevertheless, the empirical mean squared errors of the estimates are at most two orders of magnitude larger than the Cram\'er-Rao lower bound, meaning that the estimation uncertainty is low, even when the SNR is lower than 40 dB.
 
@@
 
-The subspace method can estimate the affine input parameters from the sensor response using few parameters, the sensor order $n$, the sensor static gain $\gamma$, and the RLS forgetting factor $\lambda$.
+The subspace method can estimate the affine input parameters from the sensor response using few parameters, the sensor order $n$, the sensor static gain $\mathbf{G}$, and the RLS forgetting factor $\lambda$.
 The subspace method does not necessarily requires optimization of $\lambda$ using a dataset of measured sensor responses.
@@
 The results of the sensitivity analysis show how the uncertainty of the subspace method estimates is affected when the ramp input is subject to perturbation.
-The impact on the uncertainty on the slope $\widehat{a}$ and the interception $\widehat{b}$ parameters is different.
+The impact on the uncertainty on the slope $\widehat{a}$ and the intercept $\widehat{b}$ parameters is different.
 The ramp speed uncertainty $\sigma_{\mathrm{s}}$ is added to the uncertainty of the parameter $\widehat{b}$, but does not contribute to the uncertainty of the parameter $\widehat{a}$.
@@
 
-The results of the sensitivity analysis of the ML method show that the uncertainty of the ramp input speed $\sigma_{\mathrm{s}}$ does not have an impact on the estimates uncertainty.
+The results of the sensitivity analysis of the ML method show that the uncertainty of the ramp input speed $\sigma_{\mathrm{s}}$ does not have an impact on the estimation uncertainty.
 Similar to the subspace case, the uncertainty of the ramp input parameters $\sigma_{a,b}$ is additive to the uncertainty of the estimated parameters $\widehat{a}$ and $\widehat{b}$, but does not contribute to the uncertainty of the first element of the initial conditions.
-On the other hand, the estimates uncertainty is affected by the perturbation on the model parameters $\sigma_{m,d,k}$.
+On the other hand, the estimation uncertainty is affected by the perturbation on the model parameters $\sigma_{m,d,k}$.
 It is observed that the uncertainty of the estimated parameters $\widehat{a}$ and $\widehat{b}$ increases two and three times the uncertainty of the model parameters.
@@
 
-An adaptive subspace method was proposed for the estimation of affine input parameters given the measurement of the caused sensor transient response. 
+An adaptive subspace method was proposed for estimating affine input parameters given the measurement of the caused sensor transient response. 
 The subspace estimation method is a recursive method that allows online implementation.
@@
 
-A linear time-invariant weighing system is used as an test example for the estimation methods.
+A linear time-invariant weighing system is used as a test example for the estimation methods.
 The weighing system becomes time-varying when an affine input excites the system.
-The estimation methods are compared in a simulation study where the time-varying sensor response is perturbed by measurement noise, that is assumed white of zero mean and known variance.
+The estimation methods are compared in a simulation study where the time-varying sensor response is perturbed by measurement noise, that is assumed to be white, of zero mean, and with known finite variance.
 The subspace method results are also compared to those of an existing digital time-varying filter.

### 2020-03-01 ChapterStatisticalAnalysis/ChapterStatisticalAnalysis.tex
Second review version

@@
 \emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I.
-Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Bias and covariance of the least squares estimate in a structured errors-in-variables problem", Computational Statistics & Data Analysis journal, Vol.
+Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Bias and covariance of the least-squares estimate in a structured errors-in-variables problem", Computational Statistics & Data Analysis journal, Vol.
 144, 2020, ISSN 0167-9473, doi:10.1016/j.csda.2019.106893.
@@
 
-Structured total least squares and structured total maximum likelihood are two examples of methods that solve structured EIV problems, but that require large computational resources.
+Structured total least-squares and structured total maximum likelihood are two examples of methods that solve structured EIV problems, but that require large computational resources.
 For online applications where a real-time solution of the structured EIV problem is desired, the recursive least-squares (RLS) method is a suboptimal but computationally cheaper estimator.
 
-To obtain the statistical moments of the RLS estimation for a structured and correlated EIV problem, and to assess the estimate uncertainty, that have not been obtained before, a statistical analysis was conducted.
+To obtain the statistical moments of the RLS estimation for a structured and correlated EIV problem, and to assess the estimation uncertainty, that have not been obtained before, a statistical analysis was conducted.
 
@@
 Via Monte Carlo simulations, we validated the accuracy of the bias, and the covariance approximations.
-The derived approximations predict the LS estimate bias, and the covariance, for given sample size and perturbation level.
+The derived approximations predict the LS estimation bias, and the covariance, for given sample size and perturbation level.
 The predicted variance gives the uncertainty of the LS estimate.
@@
 For an overdetermined system of linear equations, constructed by the data-driven step input estimation method (\ref{eqn:min_seiv}), the least-squares (LS) solution is given by
-\begin{equation} \widehat{\mathbf{x}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
+\begin{equation} \widehat{\bm{\theta}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
 where $\widetilde{\mathbf{K}}^\dagger$ is the pseudo-inverse of $\widetilde{\mathbf{K}}$.
-The objective of the statistical analysis is to obtain the bias and the covariance of the solution $\widehat{\mathbf{x}}$ to the structured and correlated EIV problem (\ref{eqn:ddsiemnd}).
+The objective of the statistical analysis is to obtain the bias and the covariance of the solution $\widehat{\bm{\theta}}$ to the structured and correlated EIV problem (\ref{eqn:ddsiemnd}).
 The bias and the covariance are computed using the mathematical expectation operator $\mathbb{E}\{\cdot\}$.
 First we substitute the expressions (\ref{eqn:y0plusnoise}) and (\ref{eqn:K0plusnoise}) in Equation (\ref{eqn:xhat}) to obtain
-\begin{equation*} \widehat{\mathbf{x}} = \left( (\mathbf{K}+\mathbf{E})^\top (\mathbf{K}+\mathbf{E})  \right)^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}), \end{equation*} 
+\begin{equation*} \widehat{\bm{\theta}} = \left( (\mathbf{K}+\mathbf{E})^\top (\mathbf{K}+\mathbf{E})  \right)^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}), \end{equation*} 
 which is equivalent to
-\begin{equation} \widehat{\mathbf{x}} = \left( \mathbf{I} + \mathbf{M} \right)^{-1} \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}), \label{eqn:1steq} \end{equation} 
+\begin{equation} \widehat{\bm{\theta}} = \left( \mathbf{I} + \mathbf{M} \right)^{-1} \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}), \label{eqn:1steq} \end{equation} 
 where 
@@
 We can express the estimate as
-\begin{equation} \widehat{\mathbf{x}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}).
+\begin{equation} \widehat{\bm{\theta}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}).
 \label{eqn:xhatexp} \end{equation}
 
-Now that the perturbation variables $\bm{\epsilon}$ and $\mathbf{E}$ are no longer inside an inverse matrix, we can compute the mathematical expectation of expressions derived from the Taylor series approximation (\ref{eqn:xhatexp}) of $\widehat{\mathbf{x}}$.
-The bias of the estimate $\widehat{\mathbf{x}}$ is obtained from
-\begin{equation} \mathbf{b} \left(\widehat{\mathbf{x}} \right) = \bm{\mu}\left(\widehat{\mathbf{x}} \right) - \mathbf{x}, \label{eqn:biasdef} \end{equation}
-where $\bm{\mu} \left(\widehat{\mathbf{x}} \right) = \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}$, and $\mathbf{x} = \mathbf{K}^\dagger \mathbf{y} = \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{y}$ is the true value.
-The covariance of the estimate $\widehat{\mathbf{x}}$ is obtained from
-\begin{equation} \begin{aligned} \mathbf{C} \left( \widehat{\mathbf{x}} \right) & = \mathbb{E} \left\{ \left( \widehat{\mathbf{x}} - \bm{\mu} \right) \left( \widehat{\mathbf{x}} - \bm{\mu} \right)^\top \right\} .
+Now that the perturbation variables $\bm{\epsilon}$ and $\mathbf{E}$ are no longer inside an inverse matrix, we can compute the mathematical expectation of expressions derived from the Taylor series approximation (\ref{eqn:xhatexp}) of $\widehat{\bm{\theta}}$.
+The bias of the estimate $\widehat{\bm{\theta}}$ is obtained from
+\begin{equation} \mathbf{b} \left(\widehat{\bm{\theta}} \right) = \bm{\mu}\left(\widehat{\bm{\theta}} \right) - \bm{\theta}, \label{eqn:biasdef} \end{equation}
+where $\bm{\mu} \left(\widehat{\bm{\theta}} \right) = \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\}$, and $\bm{\theta} = \mathbf{K}^\dagger \mathbf{y} = \mathbf{Q}^{-1} \mathbf{K}^\top \mathbf{y}$ is the true value.
+The covariance of the estimate $\widehat{\bm{\theta}}$ is obtained from
+\begin{equation} \begin{aligned} \mathbf{C} \left( \widehat{\bm{\theta}} \right) & = \mathbb{E} \left\{ \left( \widehat{\bm{\theta}} - \bm{\mu} \right) \left( \widehat{\bm{\theta}} - \bm{\mu} \right)^\top \right\} .
 \end{aligned} \label{eqn:covdef} \end{equation}
@@
 
-First, the statistics of the least squares estimator (\ref{eqn:xhat}) of an unstructured EIV problem is discussed, 
+First, the statistics of the least-squares estimator (\ref{eqn:xhat}) of an unstructured EIV problem is discussed, 
 provided that the perturbations of the regression matrix and the regressor are i.i.d.
@@
 the analytic approximation of the bias (\ref{eqn:biasdef}) is given by 
-\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - T \right) \mathbf{Q}^{-1} \mathbf{x} , \end{aligned} \label{eqn:biasEu} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - T \right) \mathbf{Q}^{-1} \bm{\theta} , \end{aligned} \label{eqn:biasEu} \end{equation}
 and the covariance (\ref{eqn:covdef}) is approximated by 
-\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \mathbf{x} \mathbf{x}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \ \mathbf{Q}^{-1} \mathbf{x} \mathbf{x}^\top \mathbf{Q}^{-1} , \end{aligned} \label{eqn:varEu} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \bm{\theta} \bm{\theta}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \ \mathbf{Q}^{-1} \bm{\theta} \bm{\theta}^\top \mathbf{Q}^{-1} , \end{aligned} \label{eqn:varEu} \end{equation}
 where the subscript $\mathrm{p}$ stands for prediction of the bias and covariance.
@@
 We use the results described in \citep{Vaccaro94} \S 3 and \citep{Stewart90SPT} \S 2.1 for the expected values of products of unstructured matrices with perturbations.
-Equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) depend on the unobservable true values $\mathbf{x}$, $\mathbf{K}$, and on the variance of the perturbations.
-The observed variables are $\widetilde{\mathbf{y}}$, $\widetilde{\mathbf{K}}$, and from them we compute $\widehat{\mathbf{x}}$.
+Equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) depend on the unobservable true values $\bm{\theta}$, $\mathbf{K}$, and on the variance of the perturbations.
+The observed variables are $\widetilde{\mathbf{y}}$, $\widetilde{\mathbf{K}}$, and from them we compute $\widehat{\bm{\theta}}$.
 It is proposed to directly substitute the observed variables in the analytic expressions.
-The substitution gives an approximation of the estimate bias and covariance using the observed data.
+The substitution gives an approximation of the estimation bias and covariance using the observed data.
 We have then
-\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - T \right) \widetilde{\mathbf{Q}}^{-1} \widehat{\mathbf{x}} , \end{aligned} \label{eqn:biasEuST} \end{equation}
-\begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \sigma_\epsilon^2 \ \widetilde{\mathbf{Q}}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \right) \widetilde{\mathbf{Q}}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \widetilde{\mathbf{Q}}^{-1} \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \widetilde{\mathbf{Q}}^{-1} .
+\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - T \right) \widetilde{\mathbf{Q}}^{-1} \widehat{\bm{\theta}} , \end{aligned} \label{eqn:biasEuST} \end{equation}
+\begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_\epsilon^2 \ \widetilde{\mathbf{Q}}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \right) \widetilde{\mathbf{Q}}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \widetilde{\mathbf{Q}}^{-1} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \widetilde{\mathbf{Q}}^{-1} .
 \end{aligned} \label{eqn:varEuST} \end{equation}
-In order to have a prediction of the estimate bias and covariance, the variances $\sigma_{\mathbf{E}}^2$ and $\sigma_{\bm{\epsilon}}^2$ and the observed variables $\widetilde{\mathbf{y}}$, $\widetilde{\mathbf{K}}$, and $\widehat{\mathbf{x}}$ are needed.
+In order to have a prediction of the estimation bias and covariance, the variances $\sigma_{\mathbf{E}}^2$ and $\sigma_{\bm{\epsilon}}^2$ and the observed variables $\widetilde{\mathbf{y}}$, $\widetilde{\mathbf{K}}$, and $\widehat{\bm{\theta}}$ are needed.
 
@@
 After removing the terms with negligible expected value, and considering the structure of matrix $\mathbf{K}$, the estimation bias (\ref{eqn:biasdef}) is approximated by 
-\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{Q}^{-1} \left( \left( \mathbf{K}^\top \mathbf{B}_1 - \mathbf{B}_2 \right) x - \left( \mathbf{K}^\top \mathbf{B}_3 - \mathbf{B}_4 \right) \right) , \end{aligned} \label{eqn:biasE} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{Q}^{-1} \left( \left( \mathbf{K}^\top \mathbf{B}_1 - \mathbf{B}_2 \right) x - \left( \mathbf{K}^\top \mathbf{B}_3 - \mathbf{B}_4 \right) \right) , \end{aligned} \label{eqn:biasE} \end{equation}
 whereas, the estimation covariance (\ref{eqn:covdef}) is approximated by 
-\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{K}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\mathbf{x}} \right) , \end{aligned} \label{eqn:varE} \end{equation}
-where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \Big\}$, $\mathbf{P} = \mathbf{K} \mathbf{K}^\dagger$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{P}$.
+\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right) , \end{aligned} \label{eqn:varE} \end{equation}
+where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \Big\}$, $\mathbf{P} = \mathbf{K} \mathbf{K}^\dagger$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{P}$.
 The derivation of equations (\ref{eqn:biasE}) and (\ref{eqn:varE}) is described in Appendix 1.
@@
 
-The formulas for the bias and covariance (\ref{eqn:biasE}) and (\ref{eqn:varE}) depend on the perturbation variance and on the unobservable variables $\mathbf{x}$ and $\mathbf{K}$.
-The substitution of the observed variables in the expression gives an approximation of the estimate bias and covariance based on the observed system response.
+The formulas for the bias and covariance (\ref{eqn:biasE}) and (\ref{eqn:varE}) depend on the perturbation variance and on the unobservable variables $\bm{\theta}$ and $\mathbf{K}$.
+The substitution of the observed variables in the expression gives an approximation of the estimation bias and covariance based on the observed system response.
 We have then
-\begin{equation} \begin{aligned} \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{Q}}^{-1} \left( \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\mathbf{x}} - \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasST} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{Q}}^{-1} \left( \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\bm{\theta}} - \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasST} \end{equation}
 and
-\begin{equation} \begin{aligned} \mathbf{C}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \widetilde{\mathbf{K}}^{\dagger \top} - \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right) \mathbf{b}_{\widetilde{\mathrm{p}}}^\top \left( \widehat{\mathbf{x}} \right) , \end{aligned} \label{eqn:varST} \end{equation}
-where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \bm{\epsilon}^\top \Big\}$, $\widetilde{\mathbf{C}} = \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}}$, $\widetilde{\mathbf{P}} = \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{P}}$.
+\begin{equation} \begin{aligned} \mathbf{C}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \widetilde{\mathbf{K}}^{\dagger \top} - \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\widetilde{\mathrm{p}}}^\top \left( \widehat{\bm{\theta}} \right) , \end{aligned} \label{eqn:varST} \end{equation}
+where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \bm{\epsilon}^\top \Big\}$, $\widetilde{\mathbf{C}} = \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}}$, $\widetilde{\mathbf{P}} = \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{P}}$.
 
@@
 The Cram\'er-Rao lower bound (CRLB) provides the lower limit on the variance of the estimate
-\begin{equation} \mathrm{CRLB}(\mathbf{x}) = \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\mathbf{x}}) }{\partial (\widehat{\mathbf{x}})} \right)^\top \mathbf{Fi}^{-1}(\mathbf{x}) \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\mathbf{x}}) }{\partial (\widehat{\mathbf{x}})} \right), \label{eqn:CRLB} \end{equation}
-where $\mathbf{b} (\widehat{\mathbf{x}})$ is the bias of the estimate and $\mathbf{Fi}(\widehat{\mathbf{x}})$ is the Fisher information matrix \citep{Pintelon12Book}.
+\begin{equation} \mathrm{CRLB}(\bm{\theta}) = \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\bm{\theta}}) }{\partial (\widehat{\bm{\theta}})} \right)^\top \mathbf{Fi}^{-1}(\bm{\theta}) \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\bm{\theta}}) }{\partial (\widehat{\bm{\theta}})} \right), \label{eqn:CRLB} \end{equation}
+where $\mathbf{b} (\widehat{\bm{\theta}})$ is the bias of the estimate and $\mathbf{Fi}(\widehat{\bm{\theta}})$ is the Fisher information matrix \citep{Pintelon12Book}.
 The Fisher information matrix is defined as the expected value of the Hessian of the negative likelihood function 
-\begin{equation} \mathbf{Fi}(x) = - \mathbb{E} \left\{ \frac{\partial ^2 l (\widehat{\mathbf{x}}) }{\partial \widehat{\mathbf{x}}^2 } \right\}, \end{equation}
-where the partial derivatives are evaluated in $\widehat{\mathbf{x}} = \mathbf{x}$.
+\begin{equation} \mathbf{Fi}(x) = - \mathbb{E} \left\{ \frac{\partial ^2 l (\widehat{\bm{\theta}}) }{\partial \widehat{\bm{\theta}}^2 } \right\}, \end{equation}
+where the partial derivatives are evaluated in $\widehat{\bm{\theta}} = \bm{\theta}$.
 
 The minimization problem (\ref{eqn:min_seiv}) is a structured EIV problem that can be expressed as a linear in the measurements problem \citep{Pintelon12Book}
-\begin{equation} e (\widehat{\mathbf{x}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\mathbf{x}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\mathbf{x}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 .
+\begin{equation} e (\widehat{\bm{\theta}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\bm{\theta}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\bm{\theta}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 .
 \label{eqn:M1vecyK} \end{equation}
 where $\widetilde{\mathbf{z}} = \mathbf{z} + \epsilon_{\mathbf{z}}$.
-The CRB requires that the true model $\mathbf{M}_1( \mathbf{x} ) \ \mathbf{z} = 0$ exists.
+The CRB requires that the true model $\mathbf{M}_1( \bm{\theta} ) \ \mathbf{z} = 0$ exists.
 Under the assumption of the measurement perturbation $\epsilon_{\mathbf{z}}$ being normally distributed with covariance matrix $\mathbf{C}_{\mathbf{z}}$, the loglikelihood function of the structured EIV problem is
-\begin{equation} \ln{ l(\widetilde{\mathbf{z}}, \widehat{\mathbf{z}}, \widehat{\mathbf{x}}) } = - \frac{1}{2} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right)^\top \mathbf{C}_{\mathbf{z}}^{-1} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right) + \mathrm{constant}, \end{equation}
-where $\widehat{\mathbf{z}}$ are parameters of the measurements $\widetilde{\mathbf{z}}$ that have to be estimated and satisfy $\mathbf{M}_1( \widehat{\mathbf{x}} ) \ \widehat{\mathbf{z}} = 0$.
-The size of the Fisher information matrix $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
-Moreover, in Chapter 19 of \citep{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ can be obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ after doing inversion by parts, giving
-\begin{equation} \mathbf{Fi}(\mathbf{x}) = \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)^\top \left( \mathbf{M}_1( \mathbf{x} ) \ \mathbf{C}_{\mathbf{z}} \ \mathbf{M}_1^\top( \mathbf{x} ) \right)^{-1} \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)
+\begin{equation} \ln{ l(\widetilde{\mathbf{z}}, \widehat{\mathbf{z}}, \widehat{\bm{\theta}}) } = - \frac{1}{2} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right)^\top \mathbf{C}_{\mathbf{z}}^{-1} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right) + \mathrm{constant}, \end{equation}
+where $\widehat{\mathbf{z}}$ are parameters of the measurements $\widetilde{\mathbf{z}}$ that have to be estimated and satisfy $\mathbf{M}_1( \widehat{\bm{\theta}} ) \ \widehat{\mathbf{z}} = 0$.
+The size of the Fisher information matrix $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
+Moreover, in Chapter 19 of \citep{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\bm{\theta})$ can be obtained from $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ after doing inversion by parts, giving
+\begin{equation} \mathbf{Fi}(\bm{\theta}) = \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)^\top \left( \mathbf{M}_1( \bm{\theta} ) \ \mathbf{C}_{\mathbf{z}} \ \mathbf{M}_1^\top( \bm{\theta} ) \right)^{-1} \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)
  , \label{eqn:FIM} \end{equation} 
-where the partial derivatives are evaluated at the true values $\mathbf{x}$, and the covariance matrix of the measurements is
+where the partial derivatives are evaluated at the true values $\bm{\theta}$, and the covariance matrix of the measurements is
 \begin{equation} \mathbf{C}_{\mathbf{z}} = \sigma_{\epsilon}^2 \begin{bmatrix} \mathbf{I}_{T-n} & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{1,n} & \mathbf{D}_{T-n \times T-n}^{1,n-1} & \cdots & \mathbf{D}_{T-n \times T-n}^{1,1} \\ \mathbf{0}_{T-n} & \mathbf{0}_{T-n} & \mathbf{0}_{T-n} & \mathbf{0}_{T-n} & \cdots & \mathbf{0}_{T-n} \\ \left( \mathbf{D}_{T-n \times T-n}^{1,n} \right)^\top & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{2,1} & \mathbf{D}_{T-n \times T-n}^{2,0} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,2-n} \\ \left( \mathbf{D}_{T-n \times T-n}^{1,n-1} \right)^\top & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{2,2} & \mathbf{D}_{T-n \times T-n}^{2,1} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,3-n} \\ \vdots & \vdots & \vdots & \vdots & & \vdots \\ \left( \mathbf{D}_{T-n \times T-n}^{1,1} \right)^\top & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{2,n} & \mathbf{D}_{T-n \times T-n}^{2,n-1} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,1} \end{bmatrix} .
@@
 The Cram\'er-Rao lower bound for an biased estimator of the minimization problem (\ref{eqn:min_seiv}) is given by
-\begin{equation} \mathrm{CRB}_{\mathrm{b}}(\mathbf{x}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right)^\top \ \mathbf{Fi}^{-1}(\mathbf{x}) \ \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right), \label{eqn:CRB_EIV} \end{equation}
-and for an unbiased estimator it is $\mathrm{CRB}_{\mathrm{ub}}(\mathbf{x}) = \mathbf{Fi}^{-1}(\mathbf{x})$.
+\begin{equation} \mathrm{CRB}_{\mathrm{b}}(\bm{\theta}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right)^\top \ \mathbf{Fi}^{-1}(\bm{\theta}) \ \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right), \label{eqn:CRB_EIV} \end{equation}
+and for an unbiased estimator it is $\mathrm{CRB}_{\mathrm{ub}}(\bm{\theta}) = \mathbf{Fi}^{-1}(\bm{\theta})$.
 
@@
 For high enough SNR, the constraint $\| \mathbf{M} \| < 1$ is valid and the Taylor series expansion  (\ref{eqn:TseriesExp}) holds. 
-In this case, the derived expressions predict the LS estimate bias and variance.
+In this case, the derived expressions predict the LS estimation bias and variance.
 Thus, it is relevant to monitor the evolution of the largest $\mathbf{M}$ eigenvalue to detect the lower limit of the SNR  that allows the validity of the predictions.
 
-The MC simulations provide empirical values of the bias and variance, through the sample mean and sample variance of the LS estimate $\widehat{\mathbf{x}}$.
-The results presented in this Section focus the interest in the first element of $\widehat{\mathbf{x}}$, because in the structured EIV problem, the step input estimate is $\widehat{u} = \widehat{\mathbf{x}}_{\left[1\right]}$.
+The MC simulations provide empirical values of the bias and variance, through the sample mean and sample variance of the LS estimate $\widehat{\bm{\theta}}$.
+The results presented in this Section focus the interest in the first element of $\widehat{\bm{\theta}}$, because in the structured EIV problem, the step input estimate is $\widehat{u} = \widehat{\bm{\theta}}_{\left[1\right]}$.
 
@@
 These settings are similar to those of the structured case, aiming to have comparable situations.
-%In the following results we focus our interest in the first element of the estimate $\widehat{\mathbf{x}}$.
+%In the following results we focus our interest in the first element of the estimate $\widehat{\bm{\theta}}$.
 
-The difference between the sample mean of the estimate $\widehat{\mathbf{x}}_{\left[1\right]}$ and its true value $\mathbf{x}_{\left[1\right]}$ is the empirical bias $b_\mathrm{e}$.
-\begin{equation} {b}_\mathrm{e} = \frac{1}{N_{\mathrm{MC}}} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( \widehat{\mathbf{x}}_{\left[1\right]} \right)_i -  \mathbf{x}_{\left[1\right]} } \approx \mu \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) - \mathbf{x}_{\left[1\right]} .
+The difference between the sample mean of the estimate $\widehat{\bm{\theta}}_{\left[1\right]}$ and its true value $\bm{\theta}_{\left[1\right]}$ is the empirical bias $b_\mathrm{e}$.
+\begin{equation} {b}_\mathrm{e} = \frac{1}{N_{\mathrm{MC}}} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( \widehat{\bm{\theta}}_{\left[1\right]} \right)_i -  \bm{\theta}_{\left[1\right]} } \approx \mu \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) - \bm{\theta}_{\left[1\right]} .
 \end{equation}
-The standard deviation of the estimate $\widehat{\mathbf{x}}_{\left[1\right]}$ is used to obtain the standard error $\sigma_\mathrm{e}$ of the MC simulation, which decreases with respect to the square root of the number of MC runs $N_{\mathrm{MC}}$ \citep{Hammersley75}:
-\begin{equation} \sigma_\mathrm{e} = \frac{ \sigma \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) }{\sqrt{N_{\mathrm{MC}}}}, \quad \mathrm{where} \quad \sigma^2 \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) = \frac{1}{N_{\mathrm{MC}}-1} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( \left( \widehat{\mathbf{x}}_{\left[1\right]} \right)_i - \mu \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) \right)^2 } .
+The standard deviation of the estimate $\widehat{\bm{\theta}}_{\left[1\right]}$ is used to obtain the standard error $\sigma_\mathrm{e}$ of the MC simulation, which decreases with respect to the square root of the number of MC runs $N_{\mathrm{MC}}$ \citep{Hammersley75}:
+\begin{equation} \sigma_\mathrm{e} = \frac{ \sigma \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) }{\sqrt{N_{\mathrm{MC}}}}, \quad \mathrm{where} \quad \sigma^2 \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) = \frac{1}{N_{\mathrm{MC}}-1} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( \left( \widehat{\bm{\theta}}_{\left[1\right]} \right)_i - \mu \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) \right)^2 } .
 \label{eqn:stderr} \end{equation}
@@
 The standard error of the bias prediction $b_{\widetilde{\mathrm{p}}}$ is given by
-\begin{equation} \sigma_{\widetilde{\mathrm{p}}} = \sqrt{ \frac{1}{N_{\mathrm{MC}} \left( N_{\mathrm{MC}}-1 \right)} \sum_{i=1}^{N_{\mathrm{MC}}} { \left( \left( \mathbf{b}_{\widetilde{\mathrm{p}} \left[1\right]} \right)_i - b_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}}_{\left[1\right]} \right) \right)^2 } } .
+\begin{equation} \sigma_{\widetilde{\mathrm{p}}} = \sqrt{ \frac{1}{N_{\mathrm{MC}} \left( N_{\mathrm{MC}}-1 \right)} \sum_{i=1}^{N_{\mathrm{MC}}} { \left( \left( \mathbf{b}_{\widetilde{\mathrm{p}} \left[1\right]} \right)_i - b_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) \right)^2 } } .
 \end{equation}
@@
 Processing 200 samples ensures that the step input estimation is computed from transient data only.
-%In the following results we focus our interest in the first element of the estimate $\widehat{\mathbf{x}}_{\left[1\right]}$, where the step input level estimate $\widehat{u}$ is located.
+%In the following results we focus our interest in the first element of the estimate $\widehat{\bm{\theta}}_{\left[1\right]}$, where the step input level estimate $\widehat{u}$ is located.
 
@@
 The data-driven estimation of the physical quantity is formulated as a structured EIV problem with correlation that uses the observed transient response to construct both the regression matrix and the regressor.
-The real-time implementation of the method uses a recursive least squares algorithm that is simple and has low computational complexity.
-The assessment of the uncertainty is done using the estimate bias and variance.
+The real-time implementation of the method uses a recursive least-squares algorithm that is simple and has low computational complexity.
+The assessment of the uncertainty is done using the estimation bias and variance.
 
-The conducted statistical analysis produced expressions that predict the estimate bias and variance for given sample size and perturbation level of the observed response.
+The conducted statistical analysis produced expressions that predict the estimation bias and variance for given sample size and perturbation level of the observed response.
 The Monte Carlo simulation validated the predictions.
@@
 This is due to the two approximations involved, a second-order Taylor series expansion of the estimate, and the substitution of perturbed data on the prediction expressions.
-The relative error results indicate that the estimate bias, and variance are predicted using the derived expressions, and the observed data.
+The relative error results indicate that the estimation bias, and variance are predicted using the derived expressions, and the observed data.
 The mean squared error of the estimate is close to the results of the maximum likelihood estimate given by the Cram\'er-Rao lower bound.
@@
 
-The methodology presented in this paper can be applied to estimate the uncertainty of the solutions to other structured EIV problems.
+The methodology presented in this chapter can be applied to estimate the uncertainty of the solutions to other structured EIV problems.
 The bias and variance expressions obtained after the statistical analysis depend on each specific structure.
@@
 To obtain the first and second moments of the step input estimate $\widehat{u}$, we need to study the solution
-\begin{equation} \widehat{\mathbf{x}} = ( \mathbf{K}^\top \mathbf{W} \mathbf{K}  )^{-1} \mathbf{K}^\top \mathbf{W} \mathbf{y} , \label{eqn:xhat} \end{equation}
+\begin{equation} \widehat{\bm{\theta}} = ( \mathbf{K}^\top \mathbf{W} \mathbf{K}  )^{-1} \mathbf{K}^\top \mathbf{W} \mathbf{y} , \label{eqn:xhat} \end{equation}
 of the overdetermined structured errors-in-variables EIV problem (\ref{eqn:min_ewrls}).
 Using a second order Taylor series expansion of the inverse matrix we can approximate the LS solution as
-\begin{equation} \widehat{\mathbf{x}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{C}^{-1} (\widebar{\mathbf{K}} + \mathbf{E})^\top \mathbf{W} (\widebar{\mathbf{y}} + \bm{\epsilon}).
+\begin{equation} \widehat{\bm{\theta}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{C}^{-1} (\widebar{\mathbf{K}} + \mathbf{E})^\top \mathbf{W} (\widebar{\mathbf{y}} + \bm{\epsilon}).
 \label{eqn:xhatexp} \end{equation}
@@
 
-The Taylor series approximation of $\widehat{\mathbf{x}}$ enables the calculation of the estimate bias and covariance since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion.
-The bias and the covariance of the estimate $\widehat{\mathbf{x}}$ are obtained from
-\begin{equation}  \mathbf{b} \left(\widehat{\mathbf{x}} \right) = \mathbf{\mu} - \widebar{\mathbf{x}}, \label{eqn:biasdef} \end{equation}
-\begin{equation} \begin{aligned} \mathrm{\mathbf{Cov}} \left( \widehat{\mathbf{x}} \right) & = \mathbb{E} \left\{ \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)  \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)^\top \right\} .
+The Taylor series approximation of $\widehat{\bm{\theta}}$ enables the calculation of the estimation bias and covariance since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion.
+The bias and the covariance of the estimate $\widehat{\bm{\theta}}$ are obtained from
+\begin{equation}  \mathbf{b} \left(\widehat{\bm{\theta}} \right) = \mathbf{\mu} - \widebar{\bm{\theta}}, \label{eqn:biasdef} \end{equation}
+\begin{equation} \begin{aligned} \mathrm{\mathbf{Cov}} \left( \widehat{\bm{\theta}} \right) & = \mathbb{E} \left\{ \left( \widehat{\bm{\theta}} - \mathbf{\mu} \right)  \left( \widehat{\bm{\theta}} - \mathbf{\mu} \right)^\top \right\} .
 \end{aligned} \label{eqn:covdef} \end{equation}
-where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}$ is the expected value of the estimate and $\widebar{\mathbf{x}}$ is the true value. 
+where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\}$ is the expected value of the estimate and $\widebar{\bm{\theta}}$ is the true value. 
 Considering the structure of the EIV problem, the bias and the covariance of the estimate approximation (\ref{eqn:xhatexp}) can be expressed as
-\begin{equation} \begin{aligned} \widebar{\mathbf{b}}_{\mathrm{p}} \left( \widehat{x} \right) & \approx \widebar{\mathbf{C}}^{-1} \left(  \left( \widebar{\mathbf{K}}^\top \mathbf{W} \widebar{\mathbf{B}}_1 - \widebar{\mathbf{B}}_2 \right) \mathbf{x} - \left( \widebar{\mathbf{K}}^\top \mathbf{W} \widebar{\mathbf{B}}_3 - \widebar{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasE} \end{equation}
-\begin{equation} \begin{aligned} \widebar{\mathrm{\mathbf{Cov}}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \widebar{\mathbf{K}}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \widebar{\mathbf{C}}_1 - \widebar{\mathbf{C}}_2 - \widebar{\mathbf{C}}_2^\top \right) \mathbf{W} \widebar{\mathbf{K}}^{\dagger \top}  - \widebar{\mathbf{b}}_{\mathrm{p}} \left( \widehat{x} \right) \widebar{\mathbf{b}}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varE} \end{equation}
-where $\widebar{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widebar{\mathbf{K}}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\widebar{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widebar{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widebar{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widebar{\mathbf{K}}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\widebar{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widebar{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widebar{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \Big\}$, \linebreak $\widebar{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \Big\}$, $\widebar{\mathbf{P}}_\perp = \mathbf{I} - \widebar{\mathbf{K}} \widebar{\mathbf{K}}^\dagger \mathbf{W}$, and $\mathbf{K}^\dagger$ is the pseudo-inverse matrix of $\mathbf{K}$.
+\begin{equation} \begin{aligned} \widebar{\mathbf{b}}_{\mathrm{p}} \left( \widehat{x} \right) & \approx \widebar{\mathbf{C}}^{-1} \left(  \left( \widebar{\mathbf{K}}^\top \mathbf{W} \widebar{\mathbf{B}}_1 - \widebar{\mathbf{B}}_2 \right) \bm{\theta} - \left( \widebar{\mathbf{K}}^\top \mathbf{W} \widebar{\mathbf{B}}_3 - \widebar{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasE} \end{equation}
+\begin{equation} \begin{aligned} \widebar{\mathrm{\mathbf{Cov}}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \widebar{\mathbf{K}}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \widebar{\mathbf{C}}_1 - \widebar{\mathbf{C}}_2 - \widebar{\mathbf{C}}_2^\top \right) \mathbf{W} \widebar{\mathbf{K}}^{\dagger \top}  - \widebar{\mathbf{b}}_{\mathrm{p}} \left( \widehat{x} \right) \widebar{\mathbf{b}}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varE} \end{equation}
+where $\widebar{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widebar{\mathbf{K}}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\widebar{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widebar{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widebar{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widebar{\mathbf{K}}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\widebar{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widebar{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widebar{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \Big\}$, \linebreak $\widebar{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \Big\}$, $\widebar{\mathbf{P}}_\perp = \mathbf{I} - \widebar{\mathbf{K}} \widebar{\mathbf{K}}^\dagger \mathbf{W}$, and $\mathbf{K}^\dagger$ is the pseudo-inverse matrix of $\mathbf{K}$.
 
-The bias and covariance given by expressions (\ref{eqn:biasE}) and (\ref{eqn:varE}) depend on the unobservable true values $\widebar{\mathbf{x}}$ and $\widebar{\mathbf{K}}$.
-The measured observations are in the sensor step response $\mathbf{y}$, and from its observations we construct $\mathbf{K}$ and compute $\widehat{\mathbf{x}}$.
-The substitution of the measured data in the expressions gives an approximation of the estimate bias and covariance. 
+The bias and covariance given by expressions (\ref{eqn:biasE}) and (\ref{eqn:varE}) depend on the unobservable true values $\widebar{\bm{\theta}}$ and $\widebar{\mathbf{K}}$.
+The measured observations are in the sensor step response $\mathbf{y}$, and from its observations we construct $\mathbf{K}$ and compute $\widehat{\bm{\theta}}$.
+The substitution of the measured data in the expressions gives an approximation of the estimation bias and covariance. 
 We have then
-\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{C}^{-1} \left(  \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_1 - \mathbf{B}_2 \right) \widehat{\mathbf{x}} - \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasST} \end{equation}
-  \begin{equation} \begin{aligned} {\mathrm{\mathbf{Cov}}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{K}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{W} \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{x} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varST} \end{equation}
-  where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \bm{\epsilon}^\top \Big\}$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger \mathbf{W}$.
+\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{C}^{-1} \left(  \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_1 - \mathbf{B}_2 \right) \widehat{\bm{\theta}} - \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasST} \end{equation}
+  \begin{equation} \begin{aligned} {\mathrm{\mathbf{Cov}}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{W} \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{x} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varST} \end{equation}
+  where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \bm{\epsilon}^\top \Big\}$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger \mathbf{W}$.
 
@@
 The problem formulated by the step input estimation method belongs to this type of structured EIV problems and can use the derived expressions to find the bias and variance of the input estimate $\widehat{u}$.
-The bias of the estimate $\widehat{u}$ is the fist element of $\mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right)$ and the variance of $\widehat{u}$ is the first element in the main diagonal of $\mathrm{\mathbf{Cov}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right)$.
+The bias of the estimate $\widehat{u}$ is the fist element of $\mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right)$ and the variance of $\widehat{u}$ is the first element in the main diagonal of $\mathrm{\mathbf{Cov}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right)$.
 
@@
 To find the Cram\'er-Rao lower bound (CRLB) of the structured EIV estimation problem (\ref{eqn:min_ls}), we consider that this structured and correlated estimation problem can be expressed as a linear in the measurements problem \citep{Pintelon12Book}
-\begin{equation} e (\widehat{\mathbf{x}}, \mathbf{z}) = \mathbf{M}_1( \widehat{\mathbf{x}} ) \ \mathbf{z} = \underbrace{\begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\mathbf{x}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix}}_{\mathbf{M}_1( \widehat{\mathbf{x}} )} \underbrace{\begin{bmatrix} \mathbf{y} \\ \mathrm{vec} ( \mathbf{K} ) \end{bmatrix}}_{\mathbf{z}} = 0 .
+\begin{equation} e (\widehat{\bm{\theta}}, \mathbf{z}) = \mathbf{M}_1( \widehat{\bm{\theta}} ) \ \mathbf{z} = \underbrace{\begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\bm{\theta}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix}}_{\mathbf{M}_1( \widehat{\bm{\theta}} )} \underbrace{\begin{bmatrix} \mathbf{y} \\ \mathrm{vec} ( \mathbf{K} ) \end{bmatrix}}_{\mathbf{z}} = 0 .
 \label{eqn:M1vecyK} \end{equation}
 where $\mathbf{z} = \widebar{\mathbf{z}} + \bm{\epsilon}_{\mathbf{z}}$.
-To have the CRLB, it is necessary the existence of the true model $\mathbf{M}_1( \mathbf{x} ) \ \mathbf{z} = 0$.
+To have the CRLB, it is necessary the existence of the true model $\mathbf{M}_1( \bm{\theta} ) \ \mathbf{z} = 0$.
 The measurement perturbation $\bm{\epsilon}_{\mathbf{z}}$ is assumed to be normally distributed with covariance matrix $\mathbf{C}_{\mathbf{z}}$, and then, the loglikelihood function of the structured and correlated EIV problem is
-\begin{equation} \ln{ l(\mathbf{z}, \widehat{\mathbf{z}}, \widehat{\mathbf{x}}) } = - \frac{1}{2} \left( \mathbf{z} - \widehat{\mathbf{z}} \right)^\top \mathbf{C}_{\mathbf{z}}^{-1} \left( \mathbf{z} - \widehat{\mathbf{z}} \right) + \mathrm{constant}, \end{equation}
-where the elements of $\widehat{\mathbf{z}}$ are the estimated parameters of the measurements $\mathbf{z}$,  that satisfy $\mathbf{M}_1(\widehat{\mathbf{x}}) \ \widehat{\mathbf{z}} = 0$.
-The size of the Fisher information matrix $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
-However, the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ is obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$, by doing inversion by parts \citep{Pintelon12Book} \S 19, and can be expressed as
-\begin{equation} \mathbf{Fi}(\mathbf{x}) = \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)^\top \left( \mathbf{M}_1( \mathbf{x} ) \mathbf{C}_{\mathbf{z}}  \mathbf{M}_1^\top( \mathbf{x} ) \right)^{-1}  \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right) .
+\begin{equation} \ln{ l(\mathbf{z}, \widehat{\mathbf{z}}, \widehat{\bm{\theta}}) } = - \frac{1}{2} \left( \mathbf{z} - \widehat{\mathbf{z}} \right)^\top \mathbf{C}_{\mathbf{z}}^{-1} \left( \mathbf{z} - \widehat{\mathbf{z}} \right) + \mathrm{constant}, \end{equation}
+where the elements of $\widehat{\mathbf{z}}$ are the estimated parameters of the measurements $\mathbf{z}$,  that satisfy $\mathbf{M}_1(\widehat{\bm{\theta}}) \ \widehat{\mathbf{z}} = 0$.
+The size of the Fisher information matrix $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
+However, the Fisher information matrix $\mathbf{Fi}(\bm{\theta})$ is obtained from $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$, by doing inversion by parts \citep{Pintelon12Book} \S 19, and can be expressed as
+\begin{equation} \mathbf{Fi}(\bm{\theta}) = \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)^\top \left( \mathbf{M}_1( \bm{\theta} ) \mathbf{C}_{\mathbf{z}}  \mathbf{M}_1^\top( \bm{\theta} ) \right)^{-1}  \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right) .
  \label{eqn:FIM}   \end{equation} 
-The partial derivatives are evaluated at the true values $\mathbf{x}$. 
+The partial derivatives are evaluated at the true values $\bm{\theta}$. 
 Thus, the covariance matrix of the measurements is
@@
 The Cram\'er-Rao lower bound for a biased estimator of the minimization problem (\ref{eqn:min_ls}) is 
-\begin{equation} \mathrm{CRLB}_{\mathrm{b}}(\mathbf{x}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right)^\top \mathbf{Fi}^{-1}(\mathbf{x}) \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right), \label{eqn:CRB_EIV} \end{equation}
-whereas for an unbiased estimator, it is $\mathrm{CRLB}_{\mathrm{ub}}(\mathbf{x}) = \mathbf{Fi}^{-1}(\mathbf{x})$.
+\begin{equation} \mathrm{CRLB}_{\mathrm{b}}(\bm{\theta}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right)^\top \mathbf{Fi}^{-1}(\bm{\theta}) \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right), \label{eqn:CRB_EIV} \end{equation}
+whereas for an unbiased estimator, it is $\mathrm{CRLB}_{\mathrm{ub}}(\bm{\theta}) = \mathbf{Fi}^{-1}(\bm{\theta})$.
 
@@
 To obtain the first and second moments of the step input estimate $\widehat{u}$, we need to study the least-squares (LS) solution 
-\begin{equation} \widehat{\mathbf{x}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}}  )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
+\begin{equation} \widehat{\bm{\theta}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}}  )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
 of the overdetermined structured errors-in-variables EIV problem (\ref{eqn:min_ls}), 
@@
 Using a second order Taylor series expansion of the inverse matrix we can approximate the LS solution as
-\begin{equation} \widehat{\mathbf{x}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}).
+\begin{equation} \widehat{\bm{\theta}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}).
 \label{eqn:xhatexp} \end{equation}
@@
 
-The Taylor series approximation of $\widehat{\mathbf{x}}$ enables the calculation of the bias and covariance of $\widehat{\mathbf{x}}$ since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion.
-The bias and the covariance of the estimate $\widehat{\mathbf{x}}$ are obtained from
-\begin{equation}  \mathbf{b} \left(\widehat{\mathbf{x}} \right) = \mathbf{\mu} - \mathbf{x}, \label{eqn:biasdef} \end{equation}
-\begin{equation} \begin{aligned} \mathrm{\mathbf{C}} \left( \widehat{\mathbf{\mathbf{x}}} \right) & = \mathbb{E} \left\{ \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)  \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)^\top \right\} .
+The Taylor series approximation of $\widehat{\bm{\theta}}$ enables the calculation of the bias and covariance of $\widehat{\bm{\theta}}$ since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion.
+The bias and the covariance of the estimate $\widehat{\bm{\theta}}$ are obtained from
+\begin{equation}  \mathbf{b} \left(\widehat{\bm{\theta}} \right) = \mathbf{\mu} - \bm{\theta}, \label{eqn:biasdef} \end{equation}
+\begin{equation} \begin{aligned} \mathrm{\mathbf{C}} \left( \widehat{\mathbf{\bm{\theta}}} \right) & = \mathbb{E} \left\{ \left( \widehat{\bm{\theta}} - \mathbf{\mu} \right)  \left( \widehat{\bm{\theta}} - \mathbf{\mu} \right)^\top \right\} .
 \end{aligned} \label{eqn:covdef} \end{equation}
-where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}$, and $\mathbf{x} = \mathbf{K}^\dagger \mathbf{y}$ is the true value. 
+where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\}$, and $\bm{\theta} = \mathbf{K}^\dagger \mathbf{y}$ is the true value. 
 Considering the structure of the EIV problem, the bias and the covariance of the approximation (\ref{eqn:xhatexp}) can be expressed as
-\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{Q}^{-1} \left(  \left( \mathbf{K}^\top \mathbf{B}_1 - \mathbf{B}_2 \right) \mathbf{x} - \left( \mathbf{K}^\top \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasE} \end{equation}
-\begin{equation} \begin{aligned} \mathrm{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{K}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{K}^{\dagger \top}, \end{aligned} \label{eqn:varE} \end{equation}
-where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \Big\}$, \linebreak $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \Big\}$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger$.
+\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{Q}^{-1} \left(  \left( \mathbf{K}^\top \mathbf{B}_1 - \mathbf{B}_2 \right) \bm{\theta} - \left( \mathbf{K}^\top \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasE} \end{equation}
+\begin{equation} \begin{aligned} \mathrm{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{K}^{\dagger \top}, \end{aligned} \label{eqn:varE} \end{equation}
+where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \Big\}$, \linebreak $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \Big\}$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger$.
 
@@
 
-The bias and covariance given by expressions (\ref{eqn:biasE}) and (\ref{eqn:varE}) depend on the unobservable true values $\mathbf{x}$, $\mathbf{K}$.
-The measured variable is the sensor step response $\widetilde{\mathbf{y}}$, and from its observations we construct $\widetilde{\mathbf{K}}$ and compute $\widehat{\mathbf{x}}$.
+The bias and covariance given by expressions (\ref{eqn:biasE}) and (\ref{eqn:varE}) depend on the unobservable true values $\bm{\theta}$, $\mathbf{K}$.
+The measured variable is the sensor step response $\widetilde{\mathbf{y}}$, and from its observations we construct $\widetilde{\mathbf{K}}$ and compute $\widehat{\bm{\theta}}$.
 The substitution of the measured data in the expressions gives an approximation of the bias and covariance estimation.
 We have then
-\begin{equation} \begin{aligned} \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{Q}}^{-1} \left(  \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\mathbf{x}} - \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasST} \end{equation}
-\begin{equation} \begin{aligned} \mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \widetilde{\mathbf{K}}^{\dagger \top}, \end{aligned} \label{eqn:varST} \end{equation}
-where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{\mathbf{C}}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \bm{\epsilon}^\top \Big\}$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger$.
+\begin{equation} \begin{aligned} \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{Q}}^{-1} \left(  \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\bm{\theta}} - \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasST} \end{equation}
+\begin{equation} \begin{aligned} \mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \widetilde{\mathbf{K}}^{\dagger \top}, \end{aligned} \label{eqn:varST} \end{equation}
+where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{\mathbf{C}}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \bm{\epsilon}^\top \Big\}$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger$.
 
@@
 
-The assessment of the uncertainty of the estimate $\widehat{\mathbf{x}}$ is obtained from the covariance matrix $\mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right)$.
-In this way, the uncertainty of the step input estimate $\widehat{u}$ is described by the variance in the first element on the diagonal of $\mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}} \left( \widehat{\mathbf{x}} \right)$.
+The assessment of the uncertainty of the estimate $\widehat{\bm{\theta}}$ is obtained from the covariance matrix $\mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right)$.
+In this way, the uncertainty of the step input estimate $\widehat{u}$ is described by the variance in the first element on the diagonal of $\mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right)$.
 
 In the rest of this section we describe the Cram\'er-Rao lower bound (CRB) of the structured EIV estimation problem (\ref{eqn:min_ls}), that can be expressed as a linear in the measurements problem \citep{Pintelon12Book}
-\begin{equation} e (\widehat{\mathbf{x}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\mathbf{x}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\mathbf{x}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 .
+\begin{equation} e (\widehat{\bm{\theta}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\bm{\theta}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\bm{\theta}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 .
 \label{eqn:M1vecyK} \end{equation}
 where $\widetilde{\mathbf{z}} = \mathbf{z} + \epsilon_{\mathbf{z}}$.
-The CRB requires that the true model $\mathbf{M}_1( \mathbf{x} ) \ \mathbf{z} = 0$ exists.
+The CRB requires that the true model $\mathbf{M}_1( \bm{\theta} ) \ \mathbf{z} = 0$ exists.
 Under the assumption of the measurement perturbation $\epsilon_{\mathbf{z}}$ being normally distributed with covariance matrix $\mathbf{C}_{\mathbf{z}}$, the loglikelihood function of the structured EIV problem is
-\begin{equation} \ln{ l(\widetilde{\mathbf{z}}, \widehat{\mathbf{z}}, \widehat{\mathbf{x}}) } = - \frac{1}{2} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right)^\top \mathbf{C}_{\mathbf{z}}^{-1} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right) + \mathrm{constant}, \end{equation}
-where $\widehat{\mathbf{z}}$ are parameters of the measurements $\widetilde{\mathbf{z}}$ that have to be estimated and satisfy $\mathbf{M}_1( \widehat{\mathbf{x}} ) \ \widehat{\mathbf{z}} = 0$.
-The size of the Fisher information matrix $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
-Moreover, in Chapter 19 of \citep{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\mathbf{x})$ can be obtained from $\mathbf{Fi}(\mathbf{x}, \mathbf{z})$ after doing inversion by parts, giving
-\begin{equation} \mathbf{Fi}(\mathbf{x}) = \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)^\top \left( \mathbf{M}_1( \mathbf{x} ) \mathbf{C}_{\mathbf{z}}  \mathbf{M}_1^\top( \mathbf{x} ) \right)^{-1}  \left( \frac{\partial e (\widehat{\mathbf{x}}, \mathbf{z}) }{\partial \mathbf{x} } \right)
+\begin{equation} \ln{ l(\widetilde{\mathbf{z}}, \widehat{\mathbf{z}}, \widehat{\bm{\theta}}) } = - \frac{1}{2} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right)^\top \mathbf{C}_{\mathbf{z}}^{-1} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right) + \mathrm{constant}, \end{equation}
+where $\widehat{\mathbf{z}}$ are parameters of the measurements $\widetilde{\mathbf{z}}$ that have to be estimated and satisfy $\mathbf{M}_1( \widehat{\bm{\theta}} ) \ \widehat{\mathbf{z}} = 0$.
+The size of the Fisher information matrix $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
+Moreover, in Chapter 19 of \citep{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\bm{\theta})$ can be obtained from $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ after doing inversion by parts, giving
+\begin{equation} \mathbf{Fi}(\bm{\theta}) = \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)^\top \left( \mathbf{M}_1( \bm{\theta} ) \mathbf{C}_{\mathbf{z}}  \mathbf{M}_1^\top( \bm{\theta} ) \right)^{-1}  \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)
  \label{eqn:FIM}   \end{equation} 
-where the partial derivatives are evaluated at the true values $\mathbf{x}$, and the covariance matrix of the measurements is
+where the partial derivatives are evaluated at the true values $\bm{\theta}$, and the covariance matrix of the measurements is
 
@@
 The Cram\'er-Rao lower bound for an biased estimator of the minimization problem (\ref{eqn:min_ls}) is given by
-\begin{equation} \mathrm{CRB}_{\mathrm{b}}(\mathbf{x}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right)^\top \mathbf{Fi}^{-1}(\mathbf{x}) \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\mathbf{x}} \right) }{\partial \mathbf{x} } \right), \label{eqn:CRB_EIV} \end{equation}
-and for an unbiased estimator it is $\mathrm{CRB}_{\mathrm{ub}}(\mathbf{x}) = \mathbf{Fi}^{-1}(\mathbf{x})$
+\begin{equation} \mathrm{CRB}_{\mathrm{b}}(\bm{\theta}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right)^\top \mathbf{Fi}^{-1}(\bm{\theta}) \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right), \label{eqn:CRB_EIV} \end{equation}
+and for an unbiased estimator it is $\mathrm{CRB}_{\mathrm{ub}}(\bm{\theta}) = \mathbf{Fi}^{-1}(\bm{\theta})$
 \end{comment}

### 2020-03-01 Gus-thesis-R1-review-report.tex
Second review version

\documentclass[11pt]{article}
\date{\vspace{-10ex}}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Added lines by Gustavo Quintana to facilitate internal review process
\usepackage{amsmath, amsfonts, mathtools, hyperref, tikz, pgf, subcaption, mathdots}
\usepackage[a4paper, total={6.5in, 9in}]{geometry}
\usepackage{setspace}
\usetikzlibrary{shapes, arrows, calc, patterns, decorations.pathmorphing, decorations.markings, positioning, external}
\tikzexternalize
\tikzset{external/system call={pdflatex \tikzexternalcheckshellescape -halt-on-error -interaction=batchmode -jobname "\image" "\texsource" && pdftops -eps "\image.pdf"}}
\tikzexternalize[shell escape=-enable-write18]
\usepackage{algorithm, algorithmic}
%\usepackage[noend]{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
%\renewcommand{\algorithmicreturn}{\textbf{Initialize:}}

\makeatletter
\newcommand*\rel@kern[1]{\kern#1\dimexpr\macc@kerna}
\newcommand*\widebar[1]{%
  \begingroup
  \def\mathaccent##1##2{%
    \rel@kern{0.8}%
    \overline{\rel@kern{-0.8}\macc@nucleus\rel@kern{0.2}}%
    \rel@kern{-0.2}%
  }%
  \macc@depth\@ne
  \let\math@bgroup\@empty \let\math@egroup\macc@set@skewchar
  \mathsurround\z@ \frozen@everymath{\mathgroup\macc@group\relax}%
  \macc@set@skewchar\relax
  \let\mathaccentV\macc@nested@a
  \macc@nested@a\relax111{#1}%
  \endgroup
}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Promotors' reports on the first draft of the thesis document.}

\maketitle

\section*{Ivan Markovsky}

High level comments on the thesis draft:

\begin{enumerate}

\item Don't number the Summary along with the other chapters.

\item EWRLS and TV are not standard acronyms. Don't use them.

\item Reorganize the content so that there are only two levels in the
table of contents (chapter and section). For example in the introduction
either move 2.2.1. Statistical analysis, 2.2.2. Experimental validation,
2.2.3. Affine input estimation to separate sections or don't show them
in the table of contents.
\item In Chapter 3. Preliminaries there is only one section. This is a sign
of bad organization. The section heading should be the chapter heading,
or else add other sections.
\item  The introduction begins with two pages of text that as far as I see
are not a summary of the chapter. This text should form a section on its
own.
\item  Apply in practice what you have learned in the course of Jean-luc
Doumont. In particular, start every chapter with a short summary of what
is in the chapter. Don't do this formally because it is a rule but try
to convey the highlights of the chapter and the links among the sections
in just 1/2 page.

\end{enumerate}

\section*{Rik Pintelon}

\begin{enumerate}

\item

\item

\item

\item

\item

\item

\item

\item

\item

\item

\item

\end{enumerate}

\end{document}

### 2020-03-01 Summary/Summary.tex
Second review version

--- 2020-01-12 Summary/Summary.tex
@@
 
+\vspace{-1.2cm}
+
+Sensors are dynamical systems, and their input excitations are the physical quantities. 
+The input must be estimated during the sensor transient state to expedite the measurements. 
+Otherwise, users have to wait for the sensor steady state when the response is proportional to the input through the sensor static gain. 
+One approach to get fast input estimation is filtering the sensor transient response with another dynamical system aiming to compensate the estimation time by reconstructing the input. 
+Another approach is processing the transient response on a digital signal processor (DSP). 
+Contrary to model-based compensators, DSPs allow model-free methods that can reduce the measurement time.
+
+There exists a model-free method that directly estimates the step input level from the sensor step response. 
+The formulation of this method is a Hankel-structure errors-in-variables (EIV) problem, where the regression matrix and the regressor are correlated.
+The structured EIV problem is solved with recursive least-squares to allow for real-time implementation.
+The range of application of this step input estimation method is broad because of its sensor model independence.
+
+The stochastic properties of the step input estimation method are unknown and are not straightforwardly evident.
+To validate the method for metrology applications, we studied the estimation uncertainty by doing a statistical analysis of the step input estimation.
+The estimation bias and variance builds the estimation uncertainty.
+The statistical analysis of the estimation method yields expressions that predict the estimate first and second moments.
+
+We computed the Cram\'er-Rao bound of the structured EIV problem to determine the minimum theoretical variance of its estimate.
+Both in simulations and experiments, with temperature and weighing sensors, the sample mean squared error (MSE) is compared to the Cram\'er-Rao bound.
+The simulation results show that the data-driven step input estimation is biased but with small variance, and that the estimation MSE is less than one order of magnitude larger than the Cram\'er-Rao bound.
+In practical experiments, the step input estimation method is robust when the measurement noise is not Gaussian and white.
+
+\begin{comment}
 \section{Summary - Ramp input}

### 2020-03-20 ChapterConclusion/ChapterConclusion.tex
Third review

--- 2020-01-30 ChapterConclusion/ChapterConclusion.tex
@@
 
-\chapter{Conclusions and \linebreak future work} \label{chap:concl}
+\chapter[Conclusions and future work]{Conclusions and \linebreak future work} \label{chap:concl}
 
-\vfill{}
+%\vfill{}
 
-%In this final chapter, we first summarize the contributions made to both parts of the thesis: \textbf{Part \ref{part:Filter}} and \textbf{Part \ref{part:PA}}.
-Then, I propose some possible further improvements to extend this work in future research.
+%In this final chapter, I first summarize the contributions made in the thesis. Then, I propose some possible further improvements to extend this work in future research.
 
@@
 
-Conclusions - Statistical analysis
+This thesis describes the validation of a signal processing method for metrology applications.
+The signal processing method takes the transient step response of a linear time-invariant sensor and provides an estimation of the level of the originating step input.
+The signal processing method is a subspace data-driven estimation method and constitutes an alternative to typical sensor response processing approaches based on compensation filters.
+Contrary to the compensation filters, this data-driven estimation method is model-independent and reduces the estimation time. 
+The reason for the time reduction is the bypassing of the model parameters estimation, which is not performed previously nor simultaneously to the input estimation.
+The improvement in speed estimation makes the data-driven method suitable for real-time measurements.
+
+Quantifying the estimation uncertainty was necessary to convince the metrology community of the data-driven method advantages.
+The uncertainty assessment is not straightforward.
+The data-driven method formulates a minimization problem that is a block-Hankel structured and correlated errors-in-variables (EIV) problem.
+The recursive least-squares (LS) solution to the minimization problem enables the online implementation of the method.
+The structure and correlation of the EIV problem required research to find the first and second statistical moments of the LS solution.
+
+The first part of the research reported in this thesis deals with the determination of the mean value and the covariance that the LS estimator has when it solves a structured and correlated EIV problem.
+The Taylor series expansion of the LS solution permits the study of the expected value of the LS estimate.
+The series expansion takes into account the structure of the regression matrix and the correlation between the regression matrix and the regressor.
+With the Taylor series expansion, the element-wise treatment is not a strict requirement for the statistical analysis since it is possible to use matrix form in the computations.
+As a result, the statistical analysis yielded expressions that approximate the bias and the covariance of the LS estimate for given sample size and perturbation level.
+These expressions also help to understand the impact that the different matrix structures and the correlation bring into EIV problems and their solutions.
+
+The LS bias and covariance predictions were validated first by Monte Carlo simulations. 
+In the simulations are obtained the empirical statistical moments of the LS estimation, and their predictions calculated using the derived expressions.
+The simulations used an extensive set of values in the workspace of the step response sample size and perturbation levels.
+The learning obtained from the simulation results is the operating conditions in which the estimation method is effective.
+The operation conditions were the data-driven estimation method is effective include a region of signal-to-noise (SNR) near to 40 dB, which is typical in many real-life applications.
+This effectiveness near an SNR of 40 dB is a positive indication for the usability of the method in metrology.
+Another encouraging result is that the mean squared error of the data-driven method estimate is considerably near to the minimal theoretical variance that is defined by the Cram\'er-Rao lower bound of the formulated EIV problem.
+
+The temperature and mass measurements are adequate for testing the implementation of the method.
+These two physical magnitudes are demanded applications in scientific and industrial fields, and there are available sensors of low cost in the market.
+The mass measurement demands more effort from the step input estimation method in a real-life application.
+Mass measurement sensors are affected by environment vibrations.
+Moreover, the mechanical constructions of the weighing sensors increase the order complexity of the sensor model.
+A load cell sensor, used as the weighing sensor, has a high order, that in theory, is infinite. 
+A non-surprising observation in the implementation of experimental measurements is that the measurement noise collected from the weighing sensor is not Gaussian white noise.
+Nevertheless, the implementation of the data-driven step input estimation method showed good performance results under these conditions. 
+The estimation method showed robustness against non-white noise and provided acceptable results for system orders selected between 5 and 7.
+
+A third aspect of the conducted research included the exploration of estimating inputs based on different models with signal processing methods. 
+A first choice is the affine input model, which is one complexity level higher than the step input.
+The exploration demonstrated that the adaptation of the data-driven method with the exponentially weighted recursive LS (EWRLS) is useful to estimate the affine input.
+More general than the recursive LS estimator, the EWRLS uses a tuning parameter to select the data samples relevant for the solution computation.
+
+The affine input estimation problem became more interesting because the application of the affine input to a weighing sensor turns the LTI system into a time-varying (TV) system. 
+The adaptive data-driven affine input estimation method was simulated under different assumptions and showed robustness when processing time-varying sensor responses.
+Two other estimation methods are used as a reference to compare the results of the adaptive method.
+These methods are a maximum-likelihood (ML) estimator based on local optimization and a previously reported digital time-varying filter.
+The adaptive affine input estimation method outperformed the time-varying filter by presenting lower estimation time, and the ML method by requiring less computational effort.
+
+\section*{Future work}
+
+The use of model independent signal processing methods is a research field that surely will produce interesting results in the near future. 
+With the increasing power of digital signal processors, the design of new methods is an opportunity that cannot be disregarded.
+
+The data-driven method studied in this thesis is one example of an alternative to dynamic measurements under a different paradigm.
+With respect to this method, one conclusion of the analysis presented can be that the data-driven step input estimation method is not statistically efficient because the estimation shows bias.
+A topic for future research is the efficiency increase, perhaps by designing fast and optimal estimators for structured and correlated EIV problems.
+
+On the other hand, for model-based estimation methods, such as the described ML method that has a high computational cost, there is a need also for efficient implementations to enable online optimization in receding-horizon schemes.
+With such efficient methods, the practical implementation of the ML methods can become feasible for real-time measurements.
+
+\newpage
+
+\begin{comment}
+\section{Conclusions - Statistical analysis}
 
@@
 
-Conclusions - Experimental validation
+\section{Conclusions - Experimental validation}
 
 In this paper we investigated the statistical properties of a data-driven step input estimation method in a real-life application.
-The step input estimation method is a structured and correlated errors-in-variables problem that is solved with recursive least squares.
-A statistical analysis was conducted using the ordinary least squares condensed notation. 
+The step input estimation method is a structured and correlated errors-in-variables problem that is solved with recursive least-squares.
+A statistical analysis was conducted using the ordinary least-squares condensed notation. 
 This statistical analysis of the input estimate provides expressions that approximate the estimation bias and variance assuming that the measurement noise is Gaussian white noise. 
@@
 
-Conclusions - Ramp input
-
-An adaptive subspace method was proposed for the estimation of affine input parameters given the measurement of the caused sensor transient response. 
-The subspace estimation method is a recursive method that allows online implementation.
-This method tracks the input of a system, using exponential forgetting, to process the system response.
-The subspace method is model-free and estimates directly the input parameters without identifying a sensor model.
-Therefore, it can be applied to the measurement of different physical magnitudes.
-In the specific weighing example described in the manuscript, the input is an affine function.
-The method is also applicable when the sensor is time-varying.
-The subspace method is computationally cheap, simple and suitable for implementation on digital signal processor of low computational power.
-
-A maximum-likelihood estimator based on local optimization was designed to obtain a comparative reference for the other methods.
-The maximum-likelihood method estimates the affine input parameters and also model parameters and the sensor's initial conditions.
-This method simulates, in a receding horizon scheme, the response of a sensor model to estimate the input and minimizes the sum of the squares of the residual between the measured and the estimated responses.
-The main drawback of the maximum-likelihood method is its computational cost and efficient implementation of the method is left for future work.
-
-A linear time-invariant weighing system is used as an test example for the estimation methods.
-The weighing system becomes time-varying when an affine input excites the system.
-The estimation methods are compared in a simulation study where the time-varying sensor response is perturbed by measurement noise, that is assumed white of zero mean and known variance.
-The subspace method results are also compared to those of an existing digital time-varying filter.
-The coefficients of the time-varying filter require offline optimization.
-The estimation results obtained with the subspace method converges two times faster and is one order of magnitude smaller than those obtained with the time-varying filter.
-The empirical mean squared errors of the subspace method estimation is two orders of magnitude larger than the theoretical minimum given by the Cram\'er-Rao Lower bound.
-
-Future work of this research is the practical implementation of the subspace method for real-time measurements.
-
-\begin{comment}
 \section{Conclusions - Ramp input}
@@
 Future work of this research is the practical implementation of the subspace method for real-time measurements.
-
-\section{Conclusions - Statistical analysis}
-
- We conducted a statistical analysis of a structured errors-in-variables (EIV) estimation problem with correlation to find the first and second moments of its least-squares solution.
-This estimation problem occurs in metrology when we estimate the value of a measurand directly from the sensor transient response.
-The data-driven estimation of the physical quantity is formulated as a structured EIV problem with correlation that uses the observed transient response to construct both the regression matrix and the regressor.
-The real-time implementation of the method uses a recursive least squares algorithm that is simple and has low computational complexity.
-The assessment of the uncertainty is done using the estimate bias and variance.
-
-The conducted statistical analysis produced expressions that predict the estimate bias and variance for given sample size and perturbation level of the observed response.
-The Monte Carlo simulation validated the predictions.
-We compared the results of solving an unstructured and uncorrelated EIV problem with a structured and correlated EIV problem to understand how the structure and the correlation impacts in the estimation.
-We found that the predictions in the structured case are more susceptible to perturbations.
-This is due to the two approximations involved, a second-order Taylor series expansion of the estimate, and the substitution of perturbed data on the prediction expressions.
-The relative error results indicate that the estimate bias, and variance are predicted using the derived expressions, and the observed data.
-The mean squared error of the estimate is close to the results of the maximum likelihood estimate given by the Cram\'er-Rao lower bound.
-
-The bias and variance can be accurately predicted, provided that the Taylor series expansion is valid.
-This constraint has to be taken into account to ensure the effectiveness of the method in practical applications.
-In the example, it was observed that when the SNR lies outside the validity region, the bias and variance estimation was at most three times larger than the empirical values.
-
-The methodology presented in this paper can be applied to estimate the uncertainty of the solutions to other structured EIV problems.
-The bias and variance expressions obtained after the statistical analysis depend on each specific structure.
-
-\section{Conclusions - Experimental validation}
-
-In this paper we investigated the statistical properties of a data-driven step input estimation method in a real-life application.
-The step input estimation method is a structured and correlated errors-in-variables problem that is solved with recursive least squares.
-A statistical analysis was conducted using the ordinary least squares condensed notation. 
-This statistical analysis of the input estimate provides expressions that approximate the estimation bias and variance assuming that the measurement noise is Gaussian white noise. 
-The variance approximation is useful to assess the uncertainty of the input estimate.
-In simulation we observed that the mean squared error of the input estimate is close to the theoretical minimum that uses the Cram\'er-Rao lower bound for biased estimators.
-Since the data-driven step input estimation method is not statistically efficient, there is room for improvement. This is a topic for future research.
-In the practical experiments, the measurement noise is not white.
-The noise variance obtained from the sensor steady state response underestimates the measurement noise variance, that was observed 5 dB larger in the power spectrum due to nonlinearities of the sensor.
-Considering this difference in the measurement noise variance, we introduced a conservative bound of the measurement noise variance so that the first and second moments of the input estimate are more accurately predicted.
-Using the variance approximation, we can assess the uncertainty of the input estimate with respect to the number of samples processed by the data-driven step input estimation method.
-The step input estimation method is useful in practical applications where the whiteness assumption of the measurement noise is not fulfilled.
-
-\section*{Future work}
 \end{comment}
-
-text
-
-\newpage

### 2020-03-20 ChapterExperimentalValidation/ChapterExperimentalValidation.tex
Third review

@@
 \begin{quote}
+\vspace{-0.75cm}
 \emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I.
@@
 
- \vfill{}
+%\vfill{}
 
@@
 
-The influence of the input is observed in the sensor output, and thus it makes sense to estimate the input using the sensor response.
+The input excitation influences the sensor output, and thus it makes sense to estimate the input using the sensor response.
 When the sensor reaches its steady-state, the input is a straightforwardly estimated from the sensor output making use of the sensor static gain.
-To obtain a fast estimation, the sensor transient response must be considered instead of the steady state response.
+The sensor transient response must be considered, instead of the steady state response, to obtain a fast input estimation.
 
-A common approach is to filter the sensor transient response with another dynamic system to reduce the sensor transient time by inverting the sensor dynamics.
-To design such a compensator system it is necessary a sensor model.
+A common approach is to filter the sensor transient response with another dynamic system that inverts the sensor dynamics, aiming to compensate the sensor transient time.
+It is necessary to have a sensor model to design a compensator system. 
 A different approach uses digital signal processing methods that are independent of the sensor model.
 One of these methods is the data-driven method that estimates the unknown level of step inputs by processing the sensor step response. 
-To assess the uncertainty of the data-driven step input estimation method, a experimental validation of this method was conducted in a real-life metrology application. 
-This chapter describes the measurement experiments with a weighing sensor system and discuss the results observed.
+To assess the uncertainty of the data-driven step input estimation method, it processed step responses in a real-life metrology application. 
+This chapter describes the measurement experiments with a weighing sensor system and discusses the results observed.
 The results are put in perspective to simulations and consider the measurement noise that is additive to the response of the sensor.
 
-The experimental results validate the data-driven step input estimation method, and indicate in which intervals of SNR and sample size the effectiveness of the method is guaranteed.
+The experimental results validate the data-driven step input estimation method and indicate in which intervals of SNR and sample size the method is reliable.
 

### 2020-03-20 ChapterIntroduction/ChapterIntroduction.tex
Third review

@@
 
-\vfill{}
+%\vfill{}
 
@@
 One approach to get a fast input estimation is filtering the sensor transient response with another dynamical system that inverts the dynamics of the sensor.
-The filter output is an input estimate that compensates for the measurement time.
+The filter output is an input estimate that compensates for the time span of the sensor transient state.
 The transient duration of the compensation filter should be smaller than that of the sensor.
@@
 % The philosophy behind the estimation method is avoiding the explicit identification of a system model from input-output data and estimating directly the input from the transient response.
-The main advantage of the data-driven input estimation method is that it does not identify the sensor model, but instead, it directly estimates the input. 
+The main advantage of the data-driven step input estimation method is that it does not identify the sensor model, but instead, it directly estimates the input. 
 The direct estimation differs from the standard two-stage methodology, that first identifies the sensor model and later estimates the input.
@@
 One challenge of using real-life data is that the noise may not fulfill the whiteness assumption considered in the estimation problem formulation and the statistical analysis.
-The simulation and experimental results permit to compare the performance of the step input estimation method.
+The simulation and experimental results permit to observe and compare the performance of the step input estimation method.
 
@@
 Ideally, when the conveyor belt transports the materials at a constant speed, i.e., the weighing sensor is excited with a ramp profile. 
-The ramp is an affine input model that consists of two parameters: the slope and the interception.
+The ramp is an affine input model that consists of two parameters: the slope and the intercept.
 The slope depends on the applied mass and can be used to estimate it from the transient response.
@@
 % general preview of the thesis contents
-This thesis describes three works initiated after the data-driven step input estimation method.
-The first work is a statistical analysis to elucidate the bias and the covariance of the estimate that this method provides.
-These statistical moments build the estimation uncertainty that appraises the effectiveness of the method.
+%This thesis describes three works initiated after the data-driven step input estimation method.
+%The first work is a statistical analysis to elucidate the bias and the covariance of the estimate that this method provides.
+%These statistical moments build the estimation uncertainty that appraises the effectiveness of the method.
 
@@
 The first research work conducted was a statistical analysis to elucidate the bias and the covariance of the estimate that the data-driven step input estimation method provides.
-After obtaining these statistical moments, the estimation uncertainty was assessed, and the effectiveness of the method was evaluated.
+After obtaining these statistical moments, the estimation uncertainty was assessed, and thus, the effectiveness of the method was appraised.
 The second research work was a series of experiments conducted to validate the data-driven step input estimation method in real-life applications.
@@
 In the proposed data-driven method for the estimation of the affine input \citep{QuintanaMEAS},
-the affine input is parameterized as a straight line model where the slope and the interception are the parameters of interest.
+the affine input is parameterized as a straight line model where the slope and the intercept are the parameters of interest.
 The data-driven affine input estimation method formulates a structured errors-invariables (EIV) problem, similar to the one fourmulated for the step input estimation.

### 2020-03-20 ChapterRampInput/ChapterRampInput.tex
Third review

@@
 \begin{quote}
+\vspace{-0.5cm}
 \emph{The results of this chapter were published in Quintana Carapia, G., and Markovsky, I.,
@@
 
-The goal of a measurement is to accurately estimate the input value using the sensor transient response.
-Here we consider the case where the input might vary during the measurement.
-The simplest case is when the input changes with a constant rate, and is described with an affine function.
+The goal of any measurement is accurately estimating the input using the response of the sensor in the transient state.
+Here we consider the case where the input varies during the measurement.
+The simplest case is when the input changes with a constant rate, such as being described with an affine function.
 The inputs that vary at a constant rate activate the sensor gradually. 
@@
 This chapter describes two signal processing methods for estimating affine input parameters.
-These methods are a data-driven subspace method and a maximum-likelihood (ML) estimation method based on local-optimization.
+These methods are a data-driven subspace method and a maximum-likelihood (ML) estimation method based on local optimization.
 The subspace is a recursive method independent of the sensor model, and the ML method iteratively minimizes a cost function using a simulated sensor response in a receding time horizon.
 
-In this chapter, the dynamic weighing is chosen as an implementation example.
-The dynamic weighing of an affine input is an interesting metrology application because the affine input turns a linear time-invariant sensor model into a time-varying sensor model.
-The affine input estimation methods provide an insight on the properties of the dynamic measurement of time-varying inputs.
+In this chapter, the dynamic weighing is an implementation example.
+The dynamic weighing of an affine input is a demanding metrology application because the affine excitation turns the linear time-invariant weighing sensor into a time-varying system.
+Then, the affine input estimation methods provide insight into the properties of the dynamic measurement of time-varying inputs.
 
-The effectiveness and the uncertainty of the methods are evaluated with simulation studies.
-The uncertainty is assessed, for the subspace method, using a Taylor series expansion of the exponentially weighted recursive least-squares estimate, and
+The effectiveness and uncertainty of the proposed methods are evaluated with simulation studies.
+The uncertainty is assessed, for the adaptive data-driven method, using a Taylor series expansion of the EWRLS estimate, and
 for the ML method, using the derivatives of the residual error obtained from the to-be-minimized cost function.

### 2020-03-20 ChapterStatisticalAnalysis/ChapterStatisticalAnalysis.tex
Third review

@@
 \begin{quote}
+\vspace{-0.5cm}
 \emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I.
@@
 The linear estimation problems with perturbation in both the regression matrix and the regressor are errors-in-variables (EIV) problems.
-To evaluate the properties of the solution to an EIV problem it is often required to calculate the statistical moments of the estimate.
-The calculation of the estimate statistical moments demands a larger effort when the regression matrix of a EIV problem is structured, and when is correlated to the regressor.
+It is required to calculate the statistical moments of the estimate to evaluate the properties of the solution of an EIV problem.
+The calculation of the estimate statistical moments demands a more considerable effort when the regression matrix of an EIV problem is structured, and when the regression matrix and the regressor are correlated.
 
 Structured total least-squares and structured total maximum likelihood are two examples of methods that solve structured EIV problems, but that require large computational resources.
-For online applications where a real-time solution of the structured EIV problem is desired, the recursive least-squares (RLS) method is a suboptimal but computationally cheaper estimator.
+For online applications, where a real-time solution of the structured EIV problem is imperative, the recursive least-squares (RLS) method provides a suboptimal but computationally cheaper online solution.
 
-To obtain the statistical moments of the RLS estimation for a structured and correlated EIV problem, and to assess the estimation uncertainty, that have not been obtained before, a statistical analysis was conducted.
-
+Statistical analysis is indispensable to obtain the statistical moments of the RLS solution for a structured and correlated EIV problem, and to assess the estimation uncertainty.
 This chapter describes the methodology and the results of the statistical analysis.
 The mathematical expectation of the RLS estimate, approximated by a second-order Taylor series expansion, allows a quantification of the first and second-order moments of the LS estimate.
-The results of the statistical analysis provide an insight of the impact that the structure of the regression matrix and the correlation between the regression matrix and the regressor perturbations have on the uncertainty of the RLS estimate.
+The statistical analysis results provide insight into the impact that the structure of the regression matrix and its correlation with the regressor has on the uncertainty of the RLS estimate.
 

### 2020-03-20 Extras/Summary.tex
Third review

--- 2020-03-01 Summary/Summary.tex
@@
 
-\vspace{-1.2cm}
+\vspace{-1.5cm}
 
-Sensors are dynamical systems, and their input excitations are the physical quantities. 
-The input must be estimated during the sensor transient state to expedite the measurements. 
-Otherwise, users have to wait for the sensor steady state when the response is proportional to the input through the sensor static gain. 
-One approach to get fast input estimation is filtering the sensor transient response with another dynamical system aiming to compensate the estimation time by reconstructing the input. 
+Sensors are dynamical systems, and the physical quantities are their input excitations. 
+The input magnitude must be estimated during the sensor transient state to expedite the measurements. 
+Otherwise, users have to wait for the sensor steady state when the response is proportional to the input. 
+In this case, the proportionality constant is the sensor static gain. 
+One approach to get fast input estimation is filtering the sensor transient response with another dynamical system aiming to compensate the estimation time by regenerating the input. 
 Another approach is processing the transient response on a digital signal processor (DSP). 
@@
 
-There exists a model-free method that directly estimates the step input level from the sensor step response. 
+There exists a data-driven method that directly estimates the step input level from the sensor step response. 
 The formulation of this method is a Hankel-structure errors-in-variables (EIV) problem, where the regression matrix and the regressor are correlated.
 The structured EIV problem is solved with recursive least-squares to allow for real-time implementation.
-The range of application of this step input estimation method is broad because of its sensor model independence.
+The range of application of this step input estimation method is broad because of its independence from the sensor model .
 
@@
 To validate the method for metrology applications, we studied the estimation uncertainty by doing a statistical analysis of the step input estimation.
-The estimation bias and variance builds the estimation uncertainty.
-The statistical analysis of the estimation method yields expressions that predict the estimate first and second moments.
+The estimation bias and variance build the estimation uncertainty.
+The statistical analysis of the estimation method yields expressions that predict the estimation first and second moments and enable the uncertainty assessment.
 

### 2020-03-20 Gus-thesis.tex
First draft completed!
Second review version
Third review

@@
 %% It has been adapted slightly by Gustavo Quintana.
-%% Compile with: pdflatex --shell-escape -synctex=1 -interaction=nonstopmode Gus-thesis.tex 
+%% Compile with: 
+%% cd C:\GusDocs\org\thesis\Gus-thesis
+%% pdflatex --shell-escape -synctex=1 -interaction=nonstopmode Gus-thesis.tex 
 %%
@@
 	\textbf{\small PROMOTERS}		& &	{\small prof. dr. ir. Ivan Markovsky}				\\
-				& &		{\footnotesize{}Vrije Universiteit Brussel, Belgium}\\[0.3em]
+				& &		{\footnotesize{}ELEC Department, Vrije Universiteit Brussel, Belgium}\\[0.3em]
 				& &	\small prof. dr. ir. Rik Pintelon				\\
-				& &		{\footnotesize{}ELEC Department, Vrije Universiteit Brussel, Belgium.}	\\[0.3em]
+				& &		{\footnotesize{}ELEC Department, Vrije Universiteit Brussel, Belgium}	\\[0.3em]
 				\\
 	\textbf{\small MEMBERS OF THE JURY}	&	&\small prof. dr. ir. Patrick Guillaume (President)		\\
-				&	&{\footnotesize{}MECH Department, Vrije Universiteit Brussel, Belgium.}	\\[0.3em]
+				&	&{\footnotesize{}MECH Department, Vrije Universiteit Brussel, Belgium}	\\[0.3em]
 				% &								\\
 				&	&\small prof. dr. ir. Roger Vounckx (Vice-President)				\\
-				&	&{\footnotesize{}ETRO Department, Vrije Universiteit Brussel, Belgium.}	\\[0.3em]
+				&	&{\footnotesize{}ETRO Department, Vrije Universiteit Brussel, Belgium}	\\[0.3em]
 				% &								\\
 				&	&\small dr. ir. Philippe Dreesen (Secretary)					\\
-				&	&{\footnotesize{}ELEC Department, Vrije Universiteit Brussel, Belgium.}				\\[0.3em]
+				&	&{\footnotesize{}ELEC Department, Vrije Universiteit Brussel, Belgium}				\\[0.3em]
 				% &								\\
 				&	&\small dr. ir. Nikolaos Deligiannis \\
-				&	&{\footnotesize{}ETRO Department, Vrije Universiteit Brussel, Belgium.} 			\\[0.3em]
+				&	&{\footnotesize{}ETRO Department, Vrije Universiteit Brussel, Belgium} 			\\[0.3em]
 				% &								\\
 				&	&\small prof. dr. ir. Lyudmila Mihaylova 		\\
-				&	&{\footnotesize{}ACSE Department, University of Sheffield, United Kingdom.} \\
+				&	&{\footnotesize{}ACSE Department, University of Sheffield, United Kingdom} \\
 								% &								\\
 				&	&\small prof. dr. ir. Stephane Chretien  		\\
-				&	&{\footnotesize{}ERIC Laboratory, University Lyon 2, France.} \\
+				&	&{\footnotesize{}ERIC Laboratory, University Lyon 2, France} \\
 								% &								\\
 				&	&\small prof. dr. ir. Guillaume  Merc\`ere 		\\
-				&	&{\footnotesize{}LIAS Laboratory, Universit\'e de Poitiers, France.} \\
+				&	&{\footnotesize{}LIAS Laboratory, Universit\'e de Poitiers, France} \\
 \end{tabular}
@@
 
-{\footnotesize{}Cover design by ...}\\
-{\footnotesize{}Background image by ...}{\footnotesize\par}
-{\footnotesize{}~}{\footnotesize\par}
+%{\footnotesize{}Cover design by ...}\\
+%{\footnotesize{}Background image by ...}{\footnotesize\par}
+%{\footnotesize{}~}{\footnotesize\par}
 
-{\footnotesize{}$\copyright$ February 2019 Gustavo Quintana Carapia.}\\
+{\footnotesize{}$\copyright$ February 2020, Gustavo Quintana Carapia.}\\
 {\footnotesize{}~}{\footnotesize\par}
@@
 
-\include{Extras/Acknowledgements}
+%\include{Extras/Acknowledgements}
 
@@
 
+\include{Extras/Summary}
+
 \pagenumbering{arabic}
-
-\include{Summary/Summary}
 
@@
 
+\vspace{-0.5cm}
+
 \addcontentsline{toc}{chapter}{Bibliography}

### 2020-04-07 ChapterConclusion/ChapterConclusion.tex
Final version before internal defense

@@
-%!TEX root = ..\Thesis.tex
+%!TEX root = ..\Gus-thesis.tex
 \glsresetall
@@
 The data-driven method formulates a minimization problem that is a block-Hankel structured and correlated errors-in-variables (EIV) problem.
-The recursive least-squares (LS) solution to the minimization problem enables the online implementation of the method.
+The recursive least-squares (RLS) solution to the minimization problem enables the online implementation of the method.
 The structure and correlation of the EIV problem required research to find the first and second statistical moments of the LS solution.
@@
 A first choice is the affine input model, which is one complexity level higher than the step input.
-The exploration demonstrated that the adaptation of the data-driven method with the exponentially weighted recursive LS (EWRLS) is useful to estimate the affine input.
-More general than the recursive LS estimator, the EWRLS uses a tuning parameter to select the data samples relevant for the solution computation.
+The exploration demonstrated that the adaptation of the data-driven method with the exponentially weighted RLS is useful to estimate the affine input.
+More general than the RLS estimator, the exponentially weighted RLS uses a tuning parameter to select the data samples relevant for the solution computation.
 
@@
 With such efficient methods, the practical implementation of the ML methods can become feasible for real-time measurements.
-
-\newpage
 
@@
 \end{comment}
+
+%\newpage

### 2020-04-07 ChapterIntroduction/ChapterIntroduction.tex
Final version before internal defense

@@
-%!TEX root = ..\Thesis.tex
+%!TEX root = ..\Gus-thesis.tex
+\glsresetall
+
 \chapter{Introduction} \label{chap:Introduction}
@@
 There is a trade-off between speed and accuracy when measuring with a linear time-invariant sensor.
-The input excitation drives the sensor into a transient state, and when the transient response is below the noise level, we say that the sensor is in steady state.
-During the sensor transient state, the response does not directly represent the input, but 
+The input excitation drives the sensor into a transient regime, and when the transient response is below the noise level, we say that the sensor is in steady state.
+During the sensor transient regime, the response does not directly represent the input, but 
 in steady state, the sensor response is proportional to the excitation.
@@
 However, waiting for the steady state is not always possible for practical applications that need fast measurements.
-In these practical applications, the input must be estimated during the transient state.
+In these practical applications, the input must be estimated during the transient regime.
 
@@
 One approach to get a fast input estimation is filtering the sensor transient response with another dynamical system that inverts the dynamics of the sensor.
-The filter output is an input estimate that compensates for the time span of the sensor transient state.
+The filter output is an input estimate that compensates for the time span of the sensor transient regime.
 The transient duration of the compensation filter should be smaller than that of the sensor.
@@
 The validation of the step input estimation method requires also to demonstrate its effectiveness on real-life measurements.
-Temperature and mass sensors are suitable devices to make real-life measurement experiments.
-One challenge of using real-life data is that the noise may not fulfill the whiteness assumption considered in the estimation problem formulation and the statistical analysis.
-The simulation and experimental results permit to observe and compare the performance of the step input estimation method.
+Temperature and mass sensors are suitable for real-life measurement experiments.
+One challenge of using real-life sensor outputs is that the noise may not fulfill the whiteness assumed in the estimation problem formulation and in the statistical analysis.
+The experimental and simulation results permit to evaluate the step input estimation method performance.
 
 % DM:Estimation of affine input parameters
-The ideas behind the step input estimation method raise the curiosity towards the design of estimation methods for other input models.
-One of these input models is the ramp.
-The dynamic weighing, performed in conveyor systems, also motivates the ramp input estimation method.
+The step input estimation method concepts raise the curiosity towards the design of estimation methods for other input models.
+The affine input model, for example, is found in conveyor dynamic weighing systems.
 The dynamic weighing estimates the mass of materials or products during their transportation.
-Ideally, when the conveyor belt transports the materials at a constant speed, i.e., the weighing sensor is excited with a ramp profile. 
-The ramp is an affine input model that consists of two parameters: the slope and the intercept.
-The slope depends on the applied mass and can be used to estimate it from the transient response.
-An adaptation of the step input estimation method can estimate the parameters of affine inputs.
+When the conveyor belt moves at a constant speed, the weighing sensor is excited with an affine profile, 
+described by the slope and the intercept parameters.
+The affine input parameters can be estimated by adapting the step input estimation method.
 
@@
 
-The scientific literature contains metrology studies that deal with the dynamic process effects on measurements.
-Processing the response of linear time-invariant sensors, to get fast and accurate input estimations, motivates the need to estimate inherent dynamical errors \citep{Hessling06} and to develop methods for dynamic correction in the time domain \citep{Hessling08a}.
-The industrial requirements impose new challenges by increasing the number of sensors distributed in the production lines that provide feedback for quick decision making to control autonomous activation \citep{Esward09}.
+The scientific literature contains metrology studies that deal with the dynamic process effects of measurements.
+This subsection presents a list of relevant works published in relation with the thesis topic. 
+These works study theoretical aspects like the estimation of inherent dynamical errors \citep{Hessling06}, the dynamic correction in time domain \citep{Hessling08a}; and experimental aspects like the increment of the amount of sensors in production lines for quick decision making in autonomous control \citep{Esward09}.
 
-To implement the input estimation methods in the practice, the compensation filters based on deconvolution \citep{Eichstadt10} are the preferred methodology to develop 
+To implement input estimation methods in the practice, the compensation filters based on deconvolution \citep{Eichstadt10} are the preferred methodology to develop 
 analog filters \citep{Jafaripanah05}, 
@@
 and a multiple choice set of filters \citep{Huang16}. 
-All these methods have in common the use of the measurement system model to build the compensator, for instance, after the model parameters.
+All these methods have in common the use of the measurement system model parameters to build the compensator.
 
@@
 that supports the use of simulation techniques for quantifying measurement uncertainties \citep{Esward16}.
-Another example of the Monte Carlo method application is the dynamic measurement uncertainty evaluation of clinical thermometers described in \citep{Ogorevc16}.
+Another example of the Monte Carlo method application is a dynamic measurement uncertainty evaluation of clinical thermometers \citep{Ogorevc16}.
 
@@
 Some recommend to consider the uncertainties of all the measurement chain components \citep{Diniz17}, and to avoid the direct uncertainty propagation from the calibration towards the to-be-measured quantity.
-Methods for evaluating the uncertainty associated with the output of compensation filters have been investigated, such as for
+Methods for evaluating the uncertainty associated with the output of compensation filters have been investigated for
 a discrete-time infinite-response filter in \citep{Link09},
@@
 
-This thesis describes research work initiated after the data-driven step input estimation method.
+This thesis describes research work initiated after the data-driven step input estimation method \citep{Markovsky15cep}.
 This method aims for metrology applications but it lacked an uncertainty assessment, and, therefore, its appropriateness was questionable.
@@
 The data-driven method uses exponential weighing in the recursive least-squares solution of a structured errors-in-variables problem, to give preference to recent samples over the older samples.
-The methodology and performance of the data-driven affine input estimation method were compared to those of a maximum likelihood method based.
+The methodology and performance of the data-driven affine input estimation method were compared to those of a maximum-likelihood method.
 
@@
 The data-driven step input estimation method directly estimates the input from the sensor transient response. 
-The perturbations in the EIV problem of interest come from the sensor output
+The perturbations in the EIV problem of interest come from the sensor output.
 This is the only observed signal \citep{Markovsky15cep}, and it constructs both the regression matrix and the regressor.
 The structure in the regression matrix is block-Hankel.
-The direct estimation is a data-driven methodology that reduces the input estimation time of the classical two-stage approach that first identifies a sensor model and later estimates the input using the sensor model \citep{Azam15, Niedzwiecki16a}.
+The data-driven direct estimation methodology  estimates the input faster than the classical two-stage approach, which first identifies a sensor model and later estimates the input using the sensor model \citep{Azam15, Niedzwiecki16a}.
 
-Instead of using a total least-squares (TLS) solution of the structured EIV problem, the method proposes the recursive least-squares solution (RLS) to get the step input estimate. 
+Instead of using a total least-squares (TLS) solution of the structured EIV problem, the step input estimation method uses the recursive least-squares solution (RLS). 
 The TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
@@
 a discussion of the causes of bias and inconsistency in homogeneous estimators \citep{Yeredor04homogeneous}.
-The literature does not address the uncertainty of the reported LS estimators for structured EIV problems.
+The literature does not address the uncertainty of LS estimators for structured EIV problems.
 
@@
 The bias and covariance quantification extend the perturbation analysis that investigated in \citep{Stewart90SPT} and in \citep{Vaccaro94}.
-The analysis of the two studied cases provides insight into the impact that the structure and the correlation have on the LS estimation uncertainty.
+The conducted analysis provides insight into the impact that the structure and the correlation have on the LS estimation uncertainty.
 The study presented in this thesis illustrates a methodology to conduct statistical analysis for any structured EIV problem.
@@
 
-I worked on a method to estimate the parameters of inputs that vary at a constant rate, influenced by the data-driven signal processing method that estimates the step level value using subspace techniques \citep{Markovsky15cep}, \citep{Markovsky15ieee}.
+I developed a method to estimate the parameters of inputs that vary at a constant rate, influenced by the data-driven signal processing method that estimates the step level value using subspace techniques \citep{Markovsky15cep}, \citep{Markovsky15ieee}.
 An input varies at a constant rate in applications where the magnitude of interest activates the sensor gradually. 
@@
 
-In the proposed data-driven method for the estimation of the affine input \citep{QuintanaMEAS},
+In the proposed data-driven method that estimates the affine input \citep{QuintanaMEAS},
 the affine input is parameterized as a straight line model where the slope and the intercept are the parameters of interest.
-The data-driven affine input estimation method formulates a structured errors-invariables (EIV) problem, similar to the one fourmulated for the step input estimation.
+The data-driven affine input estimation method formulates a structured errors-invariables (EIV) problem, similar to the one formulated for the step input estimation.
 An exponential weight is added to the recursive least-squares.
@@
 
-\newpage
+%\newpage

### 2020-04-07 ChapterPreliminaries/ChapterPreliminaries.tex
Third review
Final version before internal defense

@@
 A sensor is considered to be a causal linear time-invariant (LTI) system.
-The unknown quantity is the input $\mathbf{u}$ of the sensor, and the consequences of this excitation are a change in the sensor state $\mathbf{x}$, from the initial conditions $\mathbf{x}_{\text{ini}}$, and a transient response in the sensor output $\mathbf{y}$, see Figure \ref{fig:sysLTI}.
+The unknown quantity is the input ${u}$ of the sensor, and the consequences of this excitation are a change in the sensor state $\mathbf{x}$, from the initial conditions $\mathbf{x}_{\text{ini}}$, and a transient response in the sensor output ${y}$, see Figure \ref{fig:sysLTI}.
 A measurement estimates the input value using the sensor response.
@@
  \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini}}$};
- \node[draw=none,fill=none] [right=of NB1,xshift=-0.5cm,yshift=0.25cm] {$\mathbf{y}$};
+ \node[draw=none,fill=none] [right=of NB1,xshift=-0.5cm,yshift=0.25cm] {${y}$};
  \draw [-latex,thick] (NB1.west) +(-1.0,0) -- +(0.0cm,0);
 
- \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {$\mathbf{u}$};
+ \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {${u}$};
 
@@
  \caption{Block diagram of an LTI sensor.
-The input $\mathbf{u}$ excites the sensor and generates the sensor response $\mathbf{y}$.
+The input ${u}$ excites the sensor and generates the sensor response ${y}$.
 To estimate the input value it is necessary to process the response.} \label{fig:sysLTI}
@@
 The input estimation is a linear system problem, and can be done with or without a sensor model.
-The discrete-time state-space representation of an LTI system is
-\begin{equation} \begin{aligned} \mathbf{x}(k+1) &= \mathbf{A} \mathbf{x}(k) + \mathbf{B} \mathbf{u}(k), \quad \text{with} \quad \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\ 
-\mathbf{y}(k) &= \mathbf{C} \mathbf{x}(k) + \mathbf{D} \mathbf{u}(k) + \mathbf{\nu}(k),  \label{eqn:dtsslti} \end{aligned} \end{equation}
-where $\mathbf{A} \in \mathbb{R}^{n \times n}$, $\mathbf{B} \in \mathbb{R}^{n \times m}$, $\mathbf{C} \in \mathbb{R}^{p \times n}$, and $\mathbf{D} \in \mathbb{R}^{p \times m}$, are the model matrices, $\nu$ is the measurement noise, $n$ is the system order, $m$ is the number of inputs, and $p$ is the number of outputs.
-Although we may think that most sensors are single-input single output (SISO) like temperature sensors, there are sensors with single input and multiple outputs such as gas sensors \citep{Munther19}, and sensors with multiple inputs and multiple outputs like the three-axis accelerometers \citep{DEmilia16}, the radio-frequency intruder-detection sensors \citep{Ushiki13}, and the radar sensor \citep{Kueppers17}.
+The discrete-time state-space representation of a single-input single output (SISO) LTI system is
+\begin{equation} \begin{aligned} \mathbf{x}(k+1) &= \mathbf{A} \mathbf{x}(k) + \mathbf{B} {u}(k), \quad \text{with} \quad \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\ 
+{y}(k) &= \mathbf{C} \mathbf{x}(k) + {D} {u}(k) + {\epsilon}(k),  \label{eqn:dtsslti} \end{aligned} \end{equation}
+where $\mathbf{A} \in \mathbb{R}^{n \times n}$, $\mathbf{B} \in \mathbb{R}^{n}$, $\mathbf{C} \in \mathbb{R}^{1 \times n}$, and ${D} \in \mathbb{R}$, are the model matrices, $\epsilon$ is the measurement noise, and $n$ is the system order.
+% $m$ is the number of inputs, and $p$ is the number of outputs.
 
-The discrete-time state-space representation suggests that when the model, and the initial conditions are known, and in absence of measurement noise, it is sufficient to solve the system of equations 
+The methodology presented in this work is explained with SISO LTI systems, but can also be generalized to systems with multiple inputs and outputs.
+An example of SISO LTI sensor is the temperature sensor, but there are sensors that have single input and multiple outputs, like the gas sensors described in \citep{Munther19}, and sensors that have multiple inputs and multiple outputs, like the three-axis accelerometers \citep{DEmilia16}, the radio-frequency intruder-detection sensors \citep{Ushiki13}, and the radar sensors \citep{Kueppers17}.
+
+The discrete-time state-space representation suggests that, when the model and the initial conditions are known, and in absence of measurement noise, it is sufficient to solve the system of equations 
 \begin{equation} 
@@
 The input estimation problem is more complex when one or more of these assumptions are not fulfilled.
-If the initial conditions and the measurement noise are unknown, the typical approach is to feed the output $\mathbf{y}$ to an additional system.
+If the initial conditions and the measurement noise are unknown, the typical approach is to feed the output ${y}$ to an additional system.
 This additional system is built to invert the dynamics of the sensor by doing an operation that is equivalent to deconvolution, see Figure \ref{fig:compensator}.
-The output $\widehat{\mathbf{u}}$ of the additional system aims to estimate the input $\mathbf{u}$.
-This system is called compensator because the transient time of the input estimation $\widehat{\mathbf{u}}$ is smaller than the transient time of $\mathbf{y}$.
+The output $\widehat{{u}}$ of the additional system aims to estimate the input ${u}$.
+This system is called compensator because the transient time of the input estimation $\widehat{{u}}$ is smaller than the transient time of ${y}$.
 
@@
  \draw [-latex,thick] (NB2.east) ++(0,0) -- +(1.0cm,0);
- \node[draw=none,fill=none] [right=of NB2,xshift=-0.5cm,yshift=0.25cm] {$\widehat{\mathbf{u}}$};
+ \node[draw=none,fill=none] [right=of NB2,xshift=-0.5cm,yshift=0.25cm] {$\widehat{{u}}$};
 
  \end{tikzpicture}
- \caption{Input estimation using a compensator that processes the sensor response $\mathbf{y}$.
-The compensator is an additional system that reverts the dynamics of the sensor.} \label{fig:compensator}
+ \caption{Input estimation using a compensator that processes the sensor response ${y}$. The compensator is an additional system that reverts the dynamics of the sensor.} \label{fig:compensator}
  \end{figure}
@@
 
-With a step input $\mathbf{u} = \bar{\mathbf{u}} s$, where $s(t) = 1$, for $t \geq 0$, and $s(t) = 0$, for $t < 0$, the discrete-time state-space representation of the LTI sensor is equivalent to an augmented autonomous system
-\begin{equation} \begin{aligned} \mathbf{x}_\text{a}(k+1) &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_\text{a}(k) , \quad \text{where} \ \ \mathbf{x}_\text{a}(k) = \begin{bmatrix} \mathbf{x}(k) \\ \mathbf{u}(k) \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\
-\mathbf{y}(k) &= \underbrace{ \begin{bmatrix} \mathbf{C} & \mathbf{D} \end{bmatrix} }_{\mathbf{C}_\text{a}} \mathbf{x}_\text{a}(k) . \label{eqn:augmented} \end{aligned} \end{equation}
+With a step input ${u} = \bar{{u}} s$, where $s(k) = 1$, for $k \geq 0$, and $s(k) = 0$, for $k < 0$, the discrete-time state-space representation of the LTI sensor is equivalent to an augmented autonomous system
+\begin{equation} \begin{aligned} \mathbf{x}_\text{a}(k+1) &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_\text{a}(k) , \quad \text{where} \ \ \mathbf{x}_\text{a}(k) = \begin{bmatrix} \mathbf{x}(k) \\ {u}(k) \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\
+{y}(k) &= \underbrace{ \begin{bmatrix} \mathbf{C} & \mathbf{D} \end{bmatrix} }_{\mathbf{C}_\text{a}} \mathbf{x}_\text{a}(k) . \label{eqn:augmented} \end{aligned} \end{equation}
 The eigenvalues $\lambda$ of the augmented autonomous system are found using
@@
 
-Since the input $\mathbf{u}(t+1) = \mathbf{u}(t)$, for $t \geq 0$, is an augmented state of the autonomous system, and the system is known, a state estimator is sufficient to estimate the input.
+Since the input ${u}(k+1) = {u}(k)$, for $k \geq 0$, is an augmented state of the autonomous system, and the system is known, a state estimator is sufficient to estimate the input.
 In these conditions, the Kalman filter estimates recursively the input value.
@@
 The model identification consists in the estimation of the matrices $\widehat{\mathbf{A}}_\text{a}$ and $\widehat{\mathbf{C}}_\text{a}$, and the initial conditions $\widehat{\mathbf{x}}(0)$.
-To do this, consider that a Hankel matrix $\mathbfcal{H}(\mathbf{y}) \in \mathbb{R}^{n \times n}$, constructed from any linearly independent $n$ autonomous responses from the sensor initial conditions, is full column rank.
+To do this, consider that a Hankel matrix $\mathbfcal{H}({y}) \in \mathbb{R}^{n \times n}$, constructed from any linearly independent $n$ autonomous responses from the sensor initial conditions, is full column rank.
 We can express 
-\[ \underbrace{ \begin{bmatrix} y(1) & y(2) & \cdots & y(n) \\ y(2) & y(3) & \cdots & y(n+1) \\ \iddots & \iddots & \iddots \\ y(n) & y(n+1) & \cdots & y(2n-1) \end{bmatrix} }_{ \mathbfcal{H}(\mathbf{y}) }= \underbrace{ \begin{bmatrix} \widehat{\mathbf{C}}_\text{a} \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a} \\ \vdots \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a}^{n} \end{bmatrix} }_{ \mathbfcal{O}_\text{a} }  \underbrace{ \begin{bmatrix} \mathbf{x}_\text{a}(0) & \mathbf{x}_\text{a}(1) & \cdots & \mathbf{x}_\text{a}(n) \end{bmatrix} }_{ \mathbf{X}_\text{ini} } , \]
+\[ \underbrace{ \begin{bmatrix} y(1) & y(2) & \cdots & y(n) \\ y(2) & y(3) & \cdots & y(n+1) \\ \iddots & \iddots & \iddots \\ y(n) & y(n+1) & \cdots & y(2n-1) \end{bmatrix} }_{ \mathbfcal{H}({y}) }= \underbrace{ \begin{bmatrix} \widehat{\mathbf{C}}_\text{a} \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a} \\ \vdots \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a}^{n} \end{bmatrix} }_{ \mathbfcal{O}_\text{a} }  \underbrace{ \begin{bmatrix} \mathbf{x}_\text{a}(0) & \mathbf{x}_\text{a}(1) & \cdots & \mathbf{x}_\text{a}(n) \end{bmatrix} }_{ \mathbf{X}_\text{ini} } , \]
 where $\mathbf{X}_\text{ini}$ is a matrix with the initial conditions of the $n$ free responses. 
-A singular value decomposition of $\mathbfcal{H}(\mathbf{y})$ 
-\[ \mathbf{U} \bm{\Sigma} \mathbf{V} = \mathbfcal{H}(\mathbf{y}), \]
+A singular value decomposition of $\mathbfcal{H}({y})$ 
+\[ \mathbf{U} \bm{\Sigma} \mathbf{V} = \mathbfcal{H}({y}), \]
 permits the estimation of the observability matrix $\mathbfcal{O}_\text{a}$ and the initial conditions $\mathbf{X}_\text{ini}$, for example, by choosing
@@
 
-A second method can estimate the step input when the model of the LTI sensor is unknown but its static gain $\mathbf{G}$ is given.
-Using the static gain $\mathbf{G}$, we can express $\bar{\mathbf{y}} = \mathbf{G} \bar{\mathbf{u}}$, where $\bar{\mathbf{u}}$ is the input exact value and $\bar{\mathbf{y}}$ is the sensor steady state response.
+A second method can estimate the step input when the model of the LTI sensor is unknown but its static gain $G$ is given.
+Using the static gain $G$, we can express $\bar{{y}} = G \bar{{u}}$, where $\bar{{u}}$ is the input exact value and $\bar{{y}}$ is the sensor steady state response.
 The total response of the system is the sum of the transient and the steady-state responses.
 Thus, considering the augmented autonomous model, we can write
-\[ \mathbf{y} = \mathbf{G} \ \bar{\mathbf{u}} \ + \ \mathbfcal{O}_\text{a} \ \mathbf{x}(0) , \]
+\[ {y} = G \ \bar{{u}} \ + \ \mathbfcal{O}_\text{a} \ \mathbf{x}(0) , \]
 that in matrix form is
-\[ \underbrace{ \begin{bmatrix}y(0) \\ y(1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \mathbf{C}_\text{a} \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a} \\ \vdots & \vdots \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a}^T \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{\mathbf{u}} \\ \mathbf{x}(0) \end{bmatrix} .\]
-Then, it is necessary to estimate the observability matrix of the augmented system, followed by the estimation of the input  $\mathbf{u}$, and the initial conditions $\mathbf{x}_{\text{ini}}$, using least-squares
-\[ \begin{bmatrix} \widehat{\mathbf{u}} \\ \widehat{\mathbf{x}}_{\text{ini}} \end{bmatrix} = \left( \mathbf{K}^\top \mathbf{K} \right)^{-1} \mathbf{K}^\top \mathbf{y}. \]
+\[ \underbrace{ \begin{bmatrix}y(0) \\ y(1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \mathbf{C}_\text{a} \\ G & \mathbf{C}_\text{a} \mathbf{A}_\text{a} \\ \vdots & \vdots \\ G & \mathbf{C}_\text{a} \mathbf{A}_\text{a}^T \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{{u}} \\ \mathbf{x}(0) \end{bmatrix} .\]
+Then, it is necessary to estimate the observability matrix of the augmented system, followed by the estimation of the input  ${u}$, and the initial conditions $\mathbf{x}_{\text{ini}}$, using least-squares
+\[ \begin{bmatrix} \widehat{{u}} \\ \widehat{\mathbf{x}}_{\text{ini}} \end{bmatrix} = \left( \mathbf{K}^\top \mathbf{K} \right)^{-1} \mathbf{K}^\top \mathbf{y}. \]
 
@@
 Applying the first difference operator, $\Delta = \sigma - 1$, to the system state-space representation (\ref{eqn:dtsslti}), we have
-\begin{equation} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta \mathbf{y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
-where $\sigma$ is the shift operator, defined as $(\sigma^\tau y) (t) = y(t + \tau)$, and
-$\Delta \mathbf{u}(k) = \mathbf{0}$, for $k \geq 0$, and
-$\Delta \mathbf{x}(0) = (\mathbf{A} - \mathbf{I}) \mathbf{x}(0) + \mathbf{B} \bar{\mathbf{u}}$.
+\begin{equation} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta {y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
+where $\sigma$ is the shift operator, defined as $(\sigma^\tau y) (k) = y(k + \tau)$, and
+$\Delta {u}(k) = \mathbf{0}$, for $k \geq 0$, and
+$\Delta \mathbf{x}(0) = (\mathbf{A} - \mathbf{I}) \mathbf{x}(0) + \mathbf{B} \bar{{u}}$.
 The resulting system (\ref{eqn:ssalti}) is autonomous.
-When the response $\Delta \mathbf{y}$ is persistingly exciting of order $L$, i.e., when the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta \mathbf{y})$ of $L+1$ block rows constructed from $\Delta \mathbf{y}$ satisfies
-\begin{equation} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta \mathbf{y} \right) \right) \leq L,  \end{equation}
-the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta \mathbf{y})$ provides a linear map to the free responses of the augmented autonomous system (\ref{eqn:augmented}).  
+When the response $\Delta {y}$ is persistingly exciting of order $L$, i.e., when the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ of $L+1$ block rows constructed from $\Delta {y}$ satisfies
+\begin{equation} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) \leq L,  \end{equation}
+the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ provides a linear map to the free responses of the augmented autonomous system (\ref{eqn:augmented}).  
 In this case, the total response of the system is given by 
-\[ \mathbf{y} = \mathbf{G} \ \bar{\mathbf{u}} + \mathbfcal{H}(\Delta \mathbf{y}) \ \bm\ell} \]
+\[ \mathbf{y} = G \ \bar{{u}} + \mathbfcal{H}(\Delta {y}) \ \bm\ell} \]
 that is equivalent to
-\begin{equation} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \mathbf{G} & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ \mathbf{G} & \Delta y(T-n) & \Delta y(T-n+1) & \cdots & \Delta y(T-1) \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} \bar{\mathbf{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
-where $T$ is the number of samples, and the vector ${\ell}$ is a linear transformation of the system initial conditions \citep{Markovsky15cep}.  
+\begin{equation} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ G & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ G & \Delta y(T-n) & \Delta y(T-n+1) & \cdots & \Delta y(T-1) \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} \bar{{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
+where $T$ is the number of samples, and the vector $\bm{\ell}$ is a linear transformation of the system initial conditions \citep{Markovsky15cep}.  
 The solution to this system of equations exist and is unique when we have exact data.
@@
 
-Assuming that the observations of the step response $\mathbf{y}$ are perturbed by measurement noise $\bm{\epsilon}$ of zero mean and given variance $\sigma_{\bm{\epsilon}}^2$, we can express 
+Assuming that the observations of the vector $\mathbf{y}$ are perturbed by measurement noise $\bm{\epsilon}$ of zero mean and given variance $\sigma_{\bm{\epsilon}}^2$, we can express 
 \begin{equation} \widetilde{\mathbf{y}} = \mathbf{y} + \bm{\epsilon} , \label{eqn:y0plusnoise} \end{equation} 
@@
 Therefore, the matrix $\widetilde{\mathbf{K}}$ is expressed as
-\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} \mathbf{G} & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ \mathbf{G} & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ \mathbf{G} & \Delta \widetilde{y}(T-n) & \Delta \widetilde{y}(T-n+1) & \cdots & \Delta \widetilde{y}(T-1) \end{bmatrix} .
+\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(T-n) & \Delta \widetilde{y}(T-n+1) & \cdots & \Delta \widetilde{y}(T-1) \end{bmatrix} .
 \label{eqn:matrixK} \end{equation}
@@
 where $\widetilde{\mathbf{y}} = \begin{bmatrix} \widetilde{y}(n+1) & \ldots & \widetilde{y}(T) \end{bmatrix}^\top$, and
-$\widehat{\bm{\theta}} = \begin{bmatrix} \widehat{\mathbf{u}} & \widehat{\bm{\ell}}^\top \end{bmatrix}^\top$. 
+$\widehat{\bm{\theta}} = \begin{bmatrix} \widehat{{u}} & \widehat{\bm{\ell}}^\top \end{bmatrix}^\top$. 
 This minimization problem is an errors-in-variables (EIV) problem with Hankel structure, and correlation between the regression matrix $\widetilde{\mathbf{K}}$ and the regressor vector $\widetilde{\mathbf{y}}$.

### 2020-04-07 ChapterPreliminaries/ChapterPreliminariesBK.tex
Final version before internal defense

--- 2020-03-01 ChapterPreliminaries/ChapterPreliminaries.tex
@@
 The discrete-time state-space representation of an LTI system is
-\begin{equation} \begin{aligned} \mathbf{x}(k+1) &= \mathbf{A} \mathbf{x}(k) + \mathbf{B} \mathbf{u}(k), \quad \text{with} \quad \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\ 
-\mathbf{y}(k) &= \mathbf{C} \mathbf{x}(k) + \mathbf{D} \mathbf{u}(k) + \mathbf{\nu}(k),  \label{eqn:dtsslti} \end{aligned} \end{equation}
-where $\mathbf{A} \in \mathbb{R}^{n \times n}$, $\mathbf{B} \in \mathbb{R}^{n \times m}$, $\mathbf{C} \in \mathbb{R}^{p \times n}$, and $\mathbf{D} \in \mathbb{R}^{p \times m}$, are the model matrices, $\nu$ is the measurement noise, $n$ is the system order, $m$ is the number of inputs, and $p$ is the number of outputs.
+\begin{equation} \begin{aligned} \mathbf{x}_{k+1} &= \mathbf{A} \mathbf{x}_{k} + \mathbf{B} \mathbf{u}_{k}, \quad \text{with} \quad \mathbf{x}_{\text{ini}} = \mathbf{x}_{0} \\ 
+\mathbf{y}_{k} &= \mathbf{C} \mathbf{x}_{k} + \mathbf{D} \mathbf{u}_{k} + \bm{\nu}_{k},  \label{eqn:dtsslti} \end{aligned} \end{equation}
+where $\mathbf{A} \in \mathbb{R}^{n \times n}$, $\mathbf{B} \in \mathbb{R}^{n \times m}$, $\mathbf{C} \in \mathbb{R}^{p \times n}$, and $\mathbf{D} \in \mathbb{R}^{p \times m}$, are the model matrices, $\bm{\nu}$ is the measurement noise, $n$ is the system order, $m$ is the number of inputs, $p$ is the number of outputs, and $k$ is the time instant.
 Although we may think that most sensors are single-input single output (SISO) like temperature sensors, there are sensors with single input and multiple outputs such as gas sensors \citep{Munther19}, and sensors with multiple inputs and multiple outputs like the three-axis accelerometers \citep{DEmilia16}, the radio-frequency intruder-detection sensors \citep{Ushiki13}, and the radar sensor \citep{Kueppers17}.
@@
 \begin{equation} 
-\underbrace{ \begin{bmatrix} y(0) \\ y(1) \\ y(2) \\ \vdots \\ y(T) \end{bmatrix} }_{\mathbf{y}}
- = \underbrace{ \begin{bmatrix} \mathbf{C} \\ \mathbf{C} \mathbf{A} \\ \mathbf{C} \mathbf{A}^2 \\ \vdots \\ \mathbf{C} \mathbf{A}^T \end{bmatrix} }_{\mathbfcal{O}} \mathbf{x}(0) +
- \underbrace{ \begin{bmatrix} \mathbf{D} \\ \mathbf{C} \mathbf{B} & \mathbf{D} \\ \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \\ \vdots & \ddots \\ \mathbf{C} \mathbf{A}^{T-1} \mathbf{B} & \cdots  &  \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \end{bmatrix} }_{\mathbfcal{T}} \underbrace{ \begin{bmatrix} u(0) \\ u(1) \\ u(2) \\ \vdots \\ u(T) \end{bmatrix} }_{\mathbf{u}} ,
+\underbrace{ \begin{bmatrix} \mathbf{y}_{0} \\ \mathbf{y}_{1} \\ \mathbf{y}_{2} \\ \vdots \\ \mathbf{y}_{T} \end{bmatrix} }_{\mathbf{y}}
+ = \underbrace{ \begin{bmatrix} \mathbf{C} \\ \mathbf{C} \mathbf{A} \\ \mathbf{C} \mathbf{A}^2 \\ \vdots \\ \mathbf{C} \mathbf{A}^T \end{bmatrix} }_{\mathbfcal{O}} \mathbf{x}_{0} +
+ \underbrace{ \begin{bmatrix} \mathbf{D} \\ \mathbf{C} \mathbf{B} & \mathbf{D} \\ \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \\ \vdots & \ddots \\ \mathbf{C} \mathbf{A}^{T-1} \mathbf{B} & \cdots  &  \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \end{bmatrix} }_{\mathbfcal{T}} \underbrace{ \begin{bmatrix} \mathbf{u}_{0} \\ \mathbf{u}_{1} \\ \mathbf{u}_{2} \\ \vdots \\ \mathbf{u}_{T} \end{bmatrix} }_{\mathbf{u}} ,
  \label{eqn:knownmodel} \end{equation}
 to find the input. 
-This system of equations $\mathbf{y} = \mathbfcal{O} \mathbf{x}(0) + \mathbfcal{T} \mathbf{u}$ is constructed from the recursions of (\ref{eqn:dtsslti}), where $\mathbfcal{O}$ is the observability matrix of the system, and $\mathbfcal{T}$ is a Toeplitz matrix of the Markov parameters of the system.
+This system of equations $\mathbf{y} = \mathbfcal{O} \mathbf{x}_{0} + \mathbfcal{T} \mathbf{u}$ is constructed from the recursions of (\ref{eqn:dtsslti}), where $\mathbfcal{O}$ is the observability matrix of the system, and $\mathbfcal{T}$ is a Toeplitz matrix of the Markov parameters of the system.
 % formulated as the input estimation problem from the response of a system.
@@
  \draw [-latex,thick] (NB2.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
- \node[draw=none,fill=none] [above=of NB2,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini,c}}$};
+ \node[draw=none,fill=none] [above=of NB2,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini,C}}$};
  \draw [-latex,thick] (NB2.east) ++(0,0) -- +(1.0cm,0);
@@
 
-With a step input $\mathbf{u} = \bar{\mathbf{u}} s$, where $s(t) = 1$, for $t \geq 0$, and $s(t) = 0$, for $t < 0$, the discrete-time state-space representation of the LTI sensor is equivalent to an augmented autonomous system
-\begin{equation} \begin{aligned} \mathbf{x}_\text{a}(k+1) &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_\text{a}(k) , \quad \text{where} \ \ \mathbf{x}_\text{a}(k) = \begin{bmatrix} \mathbf{x}(k) \\ \mathbf{u}(k) \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\
-\mathbf{y}(k) &= \underbrace{ \begin{bmatrix} \mathbf{C} & \mathbf{D} \end{bmatrix} }_{\mathbf{C}_\text{a}} \mathbf{x}_\text{a}(k) . \label{eqn:augmented} \end{aligned} \end{equation}
+With a step input $\mathbf{u} = \bar{\mathbf{u}} s$, where $s(k) = 1$, for $k \geq 0$, and $s(k) = 0$, for $k < 0$, the discrete-time state-space representation of the LTI sensor is equivalent to an augmented autonomous system
+\begin{equation} \begin{aligned} \mathbf{x}_{k+1,\text{a}} &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_{k,\text{a}} , \quad \text{where} \ \ \mathbf{x}_{k,\text{a}} = \begin{bmatrix} \mathbf{x}_{k} \\ \mathbf{u}_{k} \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}_{0} \\
+\mathbf{y}_{k} &= \underbrace{ \begin{bmatrix} \mathbf{C} & \mathbf{D} \end{bmatrix} }_{\mathbf{C}_\text{a}} \mathbf{x}_{k,\text{a}} . \label{eqn:augmented} \end{aligned} \end{equation}
 The eigenvalues $\lambda$ of the augmented autonomous system are found using
@@
 
-Since the input $\mathbf{u}(t+1) = \mathbf{u}(t)$, for $t \geq 0$, is an augmented state of the autonomous system, and the system is known, a state estimator is sufficient to estimate the input.
+Since the input $\mathbf{u}_{k+1} = \mathbf{u}_{k}$, for $k \geq 0$, is an augmented state of the autonomous system, and the system is known, a state estimator is sufficient to estimate the input.
 In these conditions, the Kalman filter estimates recursively the input value.
@@
 A feasible method can be one that identifies a model of the sensor from the step response, assuming we have exact data, and later estimates the input step level using the identified sensor model.
-The model identification consists in the estimation of the matrices $\widehat{\mathbf{A}}_\text{a}$ and $\widehat{\mathbf{C}}_\text{a}$, and the initial conditions $\widehat{\mathbf{x}}(0)$.
+The model identification consists in the estimation of the matrices $\widehat{\mathbf{A}}_\text{a}$ and $\widehat{\mathbf{C}}_\text{a}$, and the initial conditions $\widehat{\mathbf{x}}_{0}$.
 To do this, consider that a Hankel matrix $\mathbfcal{H}(\mathbf{y}) \in \mathbb{R}^{n \times n}$, constructed from any linearly independent $n$ autonomous responses from the sensor initial conditions, is full column rank.
 We can express 
-\[ \underbrace{ \begin{bmatrix} y(1) & y(2) & \cdots & y(n) \\ y(2) & y(3) & \cdots & y(n+1) \\ \iddots & \iddots & \iddots \\ y(n) & y(n+1) & \cdots & y(2n-1) \end{bmatrix} }_{ \mathbfcal{H}(\mathbf{y}) }= \underbrace{ \begin{bmatrix} \widehat{\mathbf{C}}_\text{a} \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a} \\ \vdots \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a}^{n} \end{bmatrix} }_{ \mathbfcal{O}_\text{a} }  \underbrace{ \begin{bmatrix} \mathbf{x}_\text{a}(0) & \mathbf{x}_\text{a}(1) & \cdots & \mathbf{x}_\text{a}(n) \end{bmatrix} }_{ \mathbf{X}_\text{ini} } , \]
+\[ \underbrace{ \begin{bmatrix} \mathbf{y}_{1} & \mathbf{y}_{2} & \cdots & \mathbf{y}_{n} \\ \mathbf{y}_{2} & \mathbf{y}_{3} & \cdots & \mathbf{y}_{n+1} \\ \iddots & \iddots & \iddots \\ \mathbf{y}_{n} & \mathbf{y}_{n+1} & \cdots & \mathbf{y}_{2n-1} \end{bmatrix} }_{ \mathbfcal{H}(\mathbf{y}) }= \underbrace{ \begin{bmatrix} \widehat{\mathbf{C}}_\text{a} \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a} \\ \vdots \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a}^{n} \end{bmatrix} }_{ \mathbfcal{O}_\text{a} }  \underbrace{ \begin{bmatrix} \mathbf{x}_{0,\text{a}} & \mathbf{x}_{1,\text{a}} & \cdots & \mathbf{x}_{n,\text{a}} \end{bmatrix} }_{ \mathbf{X}_\text{ini} } , \]
 where $\mathbf{X}_\text{ini}$ is a matrix with the initial conditions of the $n$ free responses. 
@@
 Thus, considering the augmented autonomous model, we can write
-\[ \mathbf{y} = \mathbf{G} \ \bar{\mathbf{u}} \ + \ \mathbfcal{O}_\text{a} \ \mathbf{x}(0) , \]
+\[ \mathbf{y} = \mathbf{G} \ \bar{\mathbf{u}} \ + \ \mathbfcal{O}_\text{a} \ \mathbf{x}_{0} , \]
 that in matrix form is
-\[ \underbrace{ \begin{bmatrix}y(0) \\ y(1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \mathbf{C}_\text{a} \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a} \\ \vdots & \vdots \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a}^T \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{\mathbf{u}} \\ \mathbf{x}(0) \end{bmatrix} .\]
+\[ \underbrace{ \begin{bmatrix}\mathbf{y}_{0} \\ \mathbf{y}_{1} \\ \vdots \\ \mathbf{y}_{T} \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \mathbf{C}_\text{a} \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a} \\ \vdots & \vdots \\ \mathbf{G} & \mathbf{C}_\text{a} \mathbf{A}_\text{a}^T \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{\mathbf{u}} \\ \mathbf{x}_{0} \end{bmatrix} .\]
 Then, it is necessary to estimate the observability matrix of the augmented system, followed by the estimation of the input  $\mathbf{u}$, and the initial conditions $\mathbf{x}_{\text{ini}}$, using least-squares
@@
 Applying the first difference operator, $\Delta = \sigma - 1$, to the system state-space representation (\ref{eqn:dtsslti}), we have
-\begin{equation} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta \mathbf{y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
-where $\sigma$ is the shift operator, defined as $(\sigma^\tau y) (t) = y(t + \tau)$, and
-$\Delta \mathbf{u}(k) = \mathbf{0}$, for $k \geq 0$, and
-$\Delta \mathbf{x}(0) = (\mathbf{A} - \mathbf{I}) \mathbf{x}(0) + \mathbf{B} \bar{\mathbf{u}}$.
+\begin{equation} \begin{aligned} \Delta \mathbf{x}_{k+1} = \mathbf{A} \Delta \mathbf{x}_{k}, \quad \Delta \mathbf{y}_{k} = \mathbf{C} \Delta \mathbf{x}_{k}, \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}_{0} , \label{eqn:ssalti} \end{aligned} \end{equation}
+where $\sigma$ is the shift operator, defined as $(\sigma^\tau \mathbf{y})_{k} = \mathbf{y}_{t + \tau}$, and
+$\Delta \mathbf{u}_{k} = \mathbf{0}$, for $k \geq 0$, and
+$\Delta \mathbf{x}_{0} = (\mathbf{A} - \mathbf{I}) \mathbf{x}_{0} + \mathbf{B} \bar{\mathbf{u}}$.
 The resulting system (\ref{eqn:ssalti}) is autonomous.
@@
 that is equivalent to
-\begin{equation} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \mathbf{G} & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ \mathbf{G} & \Delta y(T-n) & \Delta y(T-n+1) & \cdots & \Delta y(T-1) \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} \bar{\mathbf{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
+\begin{equation} \underbrace{ \begin{bmatrix} \mathbf{y}_{n+1} \\ \vdots \\ \mathbf{y}_{T} \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{G} & \Delta \mathbf{y}_{1} & \Delta \mathbf{y}_{2} & \cdots & \Delta \mathbf{y}_{n} \\ \mathbf{G} & \Delta \mathbf{y}_{2} & \Delta \mathbf{y}_{3} & \cdots & \Delta \mathbf{y}_{n+1} \\ \vdots & \iddots & \iddots & \iddots \\ \mathbf{G} & \Delta \mathbf{y}_{T-n} & \Delta \mathbf{y}_{T-n+1} & \cdots & \Delta \mathbf{y}_{T-1} \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} \bar{\mathbf{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
 where $T$ is the number of samples, and the vector ${\ell}$ is a linear transformation of the system initial conditions \citep{Markovsky15cep}.  
@@
 \begin{equation} \widetilde{\mathbf{K}} = \mathbf{K} + \mathbf{E} . \label{eqn:K0plusnoise} \end{equation}The matrix $\mathbf{E}$ is constructed with the noise data, and is given as
-\begin{equation} \mathbf{E} = \begin{bmatrix} 0 & \Delta \epsilon(1) & \Delta \epsilon(2) & \cdots & \Delta \epsilon(n) \\ 0 & \Delta \epsilon(2) & \Delta \epsilon(3) & \cdots & \Delta \epsilon(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ 0 & \Delta \epsilon(T-n) & \Delta \epsilon(T-n+1) & \cdots & \Delta \epsilon(T-1) \end{bmatrix} .
+\begin{equation} \mathbf{E} = \begin{bmatrix} 0 & \Delta \bm{\epsilon}_{1} & \Delta \bm{\epsilon}_{2} & \cdots & \Delta \bm{\epsilon}_{n} \\ 0 & \Delta \bm{\epsilon}_{2} & \Delta \bm{\epsilon}_{3} & \cdots & \Delta \bm{\epsilon}_{n+1} \\ \vdots & \vdots & \vdots & & \vdots \\ 0 & \Delta \bm{\epsilon}_{T-n} & \Delta \bm{\epsilon}_{T-n+1} & \cdots & \Delta \bm{\epsilon}_{T-1} \end{bmatrix} .
 \label{eqn:matrixE} \end{equation}
 Therefore, the matrix $\widetilde{\mathbf{K}}$ is expressed as
-\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} \mathbf{G} & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ \mathbf{G} & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ \mathbf{G} & \Delta \widetilde{y}(T-n) & \Delta \widetilde{y}(T-n+1) & \cdots & \Delta \widetilde{y}(T-1) \end{bmatrix} .
+\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} \mathbf{G} & \Delta \widetilde{\mathbf{y}}_{1} & \Delta \widetilde{\mathbf{y}}_{2} & \cdots & \Delta \widetilde{\mathbf{y}}_{n} \\ \mathbf{G} & \Delta \widetilde{\mathbf{y}}_{2} & \Delta \widetilde{\mathbf{y}}_{3} & \cdots & \Delta \widetilde{\mathbf{y}}_{n+1} \\ \vdots & \vdots & \vdots & & \vdots \\ \mathbf{G} & \Delta \widetilde{\mathbf{y}}_{T-n} & \Delta \widetilde{\mathbf{y}}_{T-n+1} & \cdots & \Delta \widetilde{\mathbf{y}}_{T-1} \end{bmatrix} .
 \label{eqn:matrixK} \end{equation}
@@
 \label{eqn:min_seiv} \end{equation}
-where $\widetilde{\mathbf{y}} = \begin{bmatrix} \widetilde{y}(n+1) & \ldots & \widetilde{y}(T) \end{bmatrix}^\top$, and
-$\widehat{\bm{\theta}} = \begin{bmatrix} \widehat{\mathbf{u}} & \widehat{\bm{\ell}}^\top \end{bmatrix}^\top$. 
+where $\widetilde{\mathbf{y}} = \begin{bmatrix} \widetilde{\mathbf{y}}_{n+1}^\top & \ldots & \widetilde{\mathbf{y}}_{T}^\top \end{bmatrix}^\top$, and
+$\widehat{\bm{\theta}} = \begin{bmatrix} \widehat{\mathbf{u}}^\top & \widehat{\bm{\ell}}^\top \end{bmatrix}^\top$. 
 This minimization problem is an errors-in-variables (EIV) problem with Hankel structure, and correlation between the regression matrix $\widetilde{\mathbf{K}}$ and the regressor vector $\widetilde{\mathbf{y}}$.

### 2020-04-07 ChapterRampInput/ChapterRampInput.tex
Final version before internal defense

@@
+%!TEX root = ..\Gus-thesis.tex
 \glsresetall
@@
 
-The goal of any measurement is accurately estimating the input using the response of the sensor in the transient state.
+The goal of any measurement is accurately estimating the input using the response of the sensor in the transient regime.
 Here we consider the case where the input varies during the measurement.
@@
 The effectiveness and uncertainty of the proposed methods are evaluated with simulation studies.
-The uncertainty is assessed, for the adaptive data-driven method, using a Taylor series expansion of the EWRLS estimate, and
+The uncertainty is assessed, for the adaptive data-driven method, using a Taylor series expansion of the exponentially weighted recursive least-squares estimate, and
 for the ML method, using the derivatives of the residual error obtained from the to-be-minimized cost function.
@@
 The procedure is motivated by the step input estimation method that is formulated as a structured errors-in-variables (EIV) problem and is solved using recursive least-squares (RLS).
-The exponentially weighted recursive least-squares (EWRLS) is a generalization of RLS that allows for the extension of the estimation method to reconstruct the affine input.
+The exponentially weighted recursive least-squares is a generalization of RLS that allows for the extension of the estimation method to reconstruct the affine input.
 
@@
 
-The EWRLS solution to the problem (\ref{eqn:min_ewrls}) is implemented to facilitate the online implementation of the affine input estimation method.
-When $\lambda=1$, the EWRLS solution is identical to that of the RLS.
+The exponentially weighted RLS solution to the problem (\ref{eqn:min_ewrls}) is implemented to facilitate the online implementation of the affine input estimation method.
+When $\lambda=1$, the exponentially weighted RLS solution is identical to that of the RLS.
 When $\lambda<1$, the older residuals are weighted with lower values than the residuals of recent observations.
@@
 
-The EWRLS solution to the problem (\ref{eqn:min_ewrls}) provides an affine input  estimation $\widehat{\mathbf{u}}$. 
+The exponentially weighted RLS solution to the problem (\ref{eqn:min_ewrls}) provides an affine input  estimation $\widehat{\mathbf{u}}$. 
 The affine input parameters $a$ and $b$ are estimated, in a second step, by fitting $\widehat{\mathbf{u}}$ to a straight line using linear regression, as follows

### 2020-04-07 ChapterStatisticalAnalysis/ChapterStatisticalAnalysis.tex
Final version before internal defense

@@
-%!TEX root = ..\Thesis.tex
-
+%!TEX root = ..\Gus-thesis.tex
 \glsresetall
@@
 \begin{quote}
-\vspace{-0.5cm}
+%\vspace{-0.5cm}
 \emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I.
@@
 After removing the terms with negligible expected value, and considering the structure of matrix $\mathbf{K}$, the estimation bias (\ref{eqn:biasdef}) is approximated by 
-\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{Q}^{-1} \left( \left( \mathbf{K}^\top \mathbf{B}_1 - \mathbf{B}_2 \right) x - \left( \mathbf{K}^\top \mathbf{B}_3 - \mathbf{B}_4 \right) \right) , \end{aligned} \label{eqn:biasE} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{Q}^{-1} \left( \left( \mathbf{K}^\top \mathbf{B}_1 - \mathbf{B}_2 \right) \mathbf{x} - \left( \mathbf{K}^\top \mathbf{B}_3 - \mathbf{B}_4 \right) \right) , \end{aligned} \label{eqn:biasE} \end{equation}
 whereas, the estimation covariance (\ref{eqn:covdef}) is approximated by 
@@
 \begin{comment}
-
 \section{STATISTICAL ANALYSIS - Ramp input}

### 2020-04-07 Extras/Summary.tex
Final version before internal defense

@@
 \chapter{Summary} \label{chap:Summary}
-
 \vspace{-1.5cm}
-
-Sensors are dynamical systems, and the physical quantities are their input excitations. 
-The input magnitude must be estimated during the sensor transient state to expedite the measurements. 
-Otherwise, users have to wait for the sensor steady state when the response is proportional to the input. 
-In this case, the proportionality constant is the sensor static gain. 
-One approach to get fast input estimation is filtering the sensor transient response with another dynamical system aiming to compensate the estimation time by regenerating the input. 
-Another approach is processing the transient response on a digital signal processor (DSP). 
-Contrary to model-based compensators, DSPs allow model-free methods that can reduce the measurement time.
+Measurements are dynamical processes where the input is estimated using the sensor response. 
+The users cannot wait for the sensor to reach a steady regime when the sensor response closely approximates the sensor static gain times the input. 
+To obtain a fast input estimation, one approach is filtering the sensor transient response with another dynamical system, aiming to compensate the estimation time.
+The compensator is commonly built after a sensor model to perform deconvolution.  
+Another approach is estimating the input with data-driven methods that are independent from the sensor model.
+Current digital signal processors (DSP) allow the implementation of both model-based and data-driven estimation methods.
 
@@
 The formulation of this method is a Hankel-structure errors-in-variables (EIV) problem, where the regression matrix and the regressor are correlated.
-The structured EIV problem is solved with recursive least-squares to allow for real-time implementation.
-The range of application of this step input estimation method is broad because of its independence from the sensor model .
+The structured EIV problem is solved with recursive least-squares for real-time implementation.
+The independence from the sensor model of the estimation method broadens its range of application.
 
-The stochastic properties of the step input estimation method are unknown and are not straightforwardly evident.
-To validate the method for metrology applications, we studied the estimation uncertainty by doing a statistical analysis of the step input estimation.
-The estimation bias and variance build the estimation uncertainty.
-The statistical analysis of the estimation method yields expressions that predict the estimation first and second moments and enable the uncertainty assessment.
+The uncertainty of the step input estimation method is unknown and is not straightforwardly evident.
+To validate the method for metrology applications, we studied the stochastic properties of the step input estimation.
+A statistical analysis yields the estimation bias and variance, and thus, the estimation uncertainty.
 
-We computed the Cram\'er-Rao bound of the structured EIV problem to determine the minimum theoretical variance of its estimate.
-Both in simulations and experiments, with temperature and weighing sensors, the sample mean squared error (MSE) is compared to the Cram\'er-Rao bound.
-The simulation results show that the data-driven step input estimation is biased but with small variance, and that the estimation MSE is less than one order of magnitude larger than the Cram\'er-Rao bound.
-In practical experiments, the step input estimation method is robust when the measurement noise is not Gaussian and white.
+Both with simulations and experiments, the sample mean squared error (MSE) of the step input estimation method is compared to its analytical Cram\'er-Rao bound.
+The Cram\'er-Rao bound determines the minimum theoretical variance of the estimation.
+It was found that the data-driven step input estimation is biased but has small variance, and the estimation MSE is less than one order of magnitude larger than the Cram\'er-Rao bound.
+Moreover, in practical experiments, the step input estimation method is robust when the measurement noise is not Gaussian and white.
 

### 2020-05-21 ChapterPreliminaries/ChapterPreliminaries.tex
Final final version before internal defense

@@
 \begin{equation} \begin{aligned} \mathbf{x}_\text{a}(k+1) &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_\text{a}(k) , \quad \text{where} \ \ \mathbf{x}_\text{a}(k) = \begin{bmatrix} \mathbf{x}(k) \\ {u}(k) \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\
-{y}(k) &= \underbrace{ \begin{bmatrix} \mathbf{C} & \mathbf{D} \end{bmatrix} }_{\mathbf{C}_\text{a}} \mathbf{x}_\text{a}(k) . \label{eqn:augmented} \end{aligned} \end{equation}
+{y}(k) &= \underbrace{ \begin{bmatrix} \mathbf{C} & D \end{bmatrix} }_{\mathbf{C}_\text{a}} \mathbf{x}_\text{a}(k) . \label{eqn:augmented} \end{aligned} \end{equation}
 The eigenvalues $\lambda$ of the augmented autonomous system are found using
@@
 The resulting system (\ref{eqn:ssalti}) is autonomous.
-When the response $\Delta {y}$ is persistingly exciting of order $L$, i.e., when the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ of $L+1$ block rows constructed from $\Delta {y}$ satisfies
+When the response $\Delta {y}$ is persistently exciting of order $L$, i.e., when the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ of $L+1$ block rows constructed from $\Delta {y}$ satisfies
 \begin{equation} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) \leq L,  \end{equation}
@@
 In this case, the total response of the system is given by 
-\[ \mathbf{y} = G \ \bar{{u}} + \mathbfcal{H}(\Delta {y}) \ \bm\ell} \]
+\begin{equation*} \mathbf{y} = G \bar{u} + \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell} \end{equation*}
 that is equivalent to

### 2020-05-21 internal-defense/internal-defense BK.tex
Final final version before internal defense

% Compile with: 
% cd C:\GusDocs\org\thesis\Gus-thesis\internal-defense\
% pdflatex --shell-escape -synctex=1 -interaction=nonstopmode internal-defense.tex

\documentclass[presentation]{beamer}
%\usepackage[space]{grffile}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[normalem]{ulem}
\usepackage{algorithm, algorithmic, amsfonts, amsmath, amssymb, amsthm, amsbsy, array, babel, booktabs, bm, capt-of, circuitikz, color, float, graphicx, grffile, hyperref, lineno, longtable, mathdots, mathptmx, mathtools, minted, pgf, rotating, subfigure, textcomp, tikz, times, units, verbatim, wrapfig, xcolor} % 
\usepackage[outdir=./]{epstopdf}
\usetikzlibrary{shapes, arrows, calc, patterns, decorations.pathmorphing, decorations.markings, positioning}

%newproof, subcaption, 
%\newtheorem{thm}{Theorem}
%\newtheorem{lem}[thm]{Lemma}
%\newproof{pf}{Proof}

\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}

\input{C:/GusDocs/org/texinputs/mystyle-beamer-org}
\titlegraphic{\includegraphics[width=5cm]{C:/GusDocs/org/texinputs/vublogo/RGB/VUB_logo.pdf}}

% \useinnertheme{circles}
%\setbeamertemplate{itemize items}[circle]
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}{\hfill \usebeamertemplate***{navigation symbols} \hspace*{0.3cm} \vspace*{0.3cm} \leavevmode \hbox{\begin{beamercolorbox}[wd=0.9\paperwidth,ht=2.25ex,dp=1ex,right]{date in head/foot} \insertframenumber{} \hspace*{1ex} \end{beamercolorbox}} \vskip0pt}
\setbeamerfont{footline}{size=\fontsize{8}{9}}
\addtocounter{framenumber}{0}
\thispagestyle{fancy}
\usetheme{}
\usecolortheme{}

\author{Gustavo Quintana Carapia}
\date{}
\title{Statistical analysis and experimental validation \linebreak of data-driven dynamic measurement methods}

\newcommand\Wider[2][3em]{%
\makebox[\linewidth][c]{%
  \begin{minipage}{\dimexpr\textwidth+#1\relax}
  \raggedright#2
  \end{minipage}%
  }%
}

\begin{document}

\begin{frame}[noframenumbering,plain]
\Wider{\maketitle}
\end{frame}

%\linebreak

\begin{frame}[label={slide:context}]{A measurement is a dynamic process}
\begin{itemize}
\item The sensor interacts with its environment.
\item The error is inevitably present in the sensor response.
\item The aim is to estimate the input from the sensor response.
\end{itemize}
\begin{figure}[htb!]
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}] 
 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {Sensor};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini}}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=1.0cm,yshift=0.25cm] {${y}$};
 \draw [-latex,thick] (NB1.west) +(-1.0,0) -- +(0.0cm,0);
 \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {${u}$}; 
 \draw (2,0.0) circle (2.5mm);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.00mm,yshift=0.0mm]{+};
 \draw [-latex,thick] (NB1.west) +(3.0,0) -- +(4.0cm,0); 
  \draw [-latex,thick] (NB1.west) ++(2.75cm,1.0) -- +(0.0cm,-0.75);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.250cm,yshift=0.75cm] {$\epsilon$};
\end{tikzpicture}
\end{figure}
%\begin{columns}
%\begin{column}{0.35\columnwidth}
%\begin{block}{}
%\end{block}
%\end{column}
%\begin{column}{0.65\columnwidth}
%\begin{block}{}
%\begin{center}
%\includegraphics[width=.9\linewidth]{./ytilde.eps}
%\end{center}
%\end{block}
%\end{column}
%\end{columns}
\end{frame}

%\begin{frame}[label={slides:have}]{Methods for step input estimation.}
%\begin{columns}
%\begin{column}{0.5\columnwidth}
%\begin{block}{Known model}
%Kalman filter.
%State space recursions.
%\end{block}
%\end{column}
%\begin{column}{0.5\columnwidth}
%\begin{block}{Unknown model}
%Data-driven.
%Local optimization.    
%\end{block}
%\end{column}
%\end{columns}
%\end{frame}

\begin{frame}[label={slide:have}]{A compensator is typically used \linebreak to estimate the input from the sensor response}
\begin{itemize}
\item The additional convolution inverts the sensor dynamics.
\end{itemize}
\begin{figure}[htb!]
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}] 
 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {Sensor};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini}}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=1.0cm,yshift=0.25cm] {${y}$};
 \draw [-latex,thick] (NB1.west) +(-1.0,0) -- +(0.0cm,0);
 \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {${u}$};
 \draw (2,0.0) circle (2.5mm);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.00mm,yshift=0.0mm]{+};
 \draw [-latex,thick] (NB1.west) +(3.0,0) -- +(4.0cm,0); 
  \draw [-latex,thick] (NB1.west) ++(2.75cm,1.0) -- +(0.0cm,-0.75);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.250cm,yshift=0.75cm] {$\epsilon$};
\node (NB2) [minimum width=1.5cm, minimum height=0.75cm, xshift=4.5cm, yshift=0.0cm] {Compensator}; 
 \draw [-latex,thick] (NB2.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB2,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini,c}}$};
 \draw [-latex,thick] (NB2.east) ++(0,0) -- +(1.0cm,0);
 \node[draw=none,fill=none] [right=of NB2,xshift=-0.5cm,yshift=0.25cm] {$\widehat{{u}}$};
 \end{tikzpicture}
 \end{figure}
\end{frame}

\begin{frame}[label={slide:have}]{Digital signal processors can help \linebreak to reduce the measurement time even more}
Appropriate DSP programming  can 
\begin{itemize}
\item emulate dynamical systems, or 
\item implement data-driven methods. 
\end{itemize}
\begin{figure}[htb!]
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}] 
 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {Sensor};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini}}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=1.0cm,yshift=0.25cm] {${y}$};
 \draw [-latex,thick] (NB1.west) +(-1.0,0) -- +(0.0cm,0);
 \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {${u}$};
 \draw (2,0.0) circle (2.5mm);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.00mm,yshift=0.0mm]{+};
 \draw [-latex,thick] (NB1.west) +(3.0,0) -- +(4.0cm,0); 
  \draw [-latex,thick] (NB1.west) ++(2.75cm,1.0) -- +(0.0cm,-0.75);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.250cm,yshift=0.75cm] {$\epsilon$};
\node (NB2) [minimum width=1.5cm, minimum height=0.75cm, xshift=4.0cm, yshift=0.0cm] {DSP}; 
 \draw [-latex,thick] (NB2.east) ++(0,0) -- +(1.0cm,0);
 \node[draw=none,fill=none] [right=of NB2,xshift=-0.5cm,yshift=0.25cm] {$\widehat{{u}}$};
 \end{tikzpicture}
 \end{figure}
\end{frame}

\begin{frame}[label={slide:have}]{In this work we focus on a data-driven \linebreak step input estimation method}
\begin{itemize}
\item The method is formulated as a structured and correlated errors-in-variables problem. 
\item The online solution is computed using recursive least squares. 
\end{itemize}
\end{frame}

\begin{frame}[label={slide:need}]{The uncertainty of the data-driven step input estimation method was unknown}
\begin{itemize}
\item The validation of the method was necessary to foster \linebreak the method utilization in metrology applications. 
\end{itemize}
\end{frame}

\begin{frame}[label={slide:objective}]{A statistical analysis was conducted to find the data-driven step input estimation method uncertainty}
\begin{itemize}
\item This analysis unveils the step input estimation mean and variance, and thus, the method uncertainty.  
\item The estimation mean and variance knowledge permits to evaluate the method and its effectiveness. 
\end{itemize}
\end{frame}

\begin{frame}[label={slide:plan}]{The conducted research work covered}
\begin{enumerate}
\item A statistical analysis of structured and correlated errors-in-variables problems. 
\item An experimental validation based on the uncertainty of \linebreak the step input estimation method. 
\item The proposal of two affine input estimation methods. 
\end{enumerate}
\end{frame}

\begin{frame}[label={slide:preliminaries1}]{To formulate the input estimation methods, we consider a measurement is a linear system problem}
\begin{figure}[htb!]
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}] 
 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {Sensor};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini}}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=1.0cm,yshift=0.25cm] {${y}$};
 \draw [-latex,thick] (NB1.west) +(-1.0,0) -- +(0.0cm,0);
 \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {${u}$}; 
 \draw (2,0.0) circle (2.5mm);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.00mm,yshift=0.0mm]{+};
 \draw [-latex,thick] (NB1.west) +(3.0,0) -- +(4.0cm,0); 
  \draw [-latex,thick] (NB1.west) ++(2.75cm,1.0) -- +(0.0cm,-0.75);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.250cm,yshift=0.75cm] {$\epsilon$};
\end{tikzpicture}
\end{figure}
\begin{itemize}
\item  With LTI sensor, the measurement dynamics \linebreak can be described using state-space representation: 
\end{itemize}
\begin{equation*} \begin{aligned} \mathbf{x}(k+1) &= \mathbf{A} \mathbf{x}(k) + \mathbf{B} {u}(k), \quad \text{with} \quad \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\ 
{y}(k) &= \mathbf{C} \mathbf{x}(k) + {D} {u}(k) + {\epsilon}(k) \end{aligned} \end{equation*}
\end{frame}

\begin{frame}[label={slide:preliminaries2}]{If the sensor model and initial conditions are known, and we have the exact data sensor response}
\begin{itemize}
\item  then, the input can be estimated from: 
\end{itemize}
\begin{equation*} 
\underbrace{ \begin{bmatrix} y(0) \\ y(1) \\ y(2) \\ \vdots \\ y(T) \end{bmatrix} }_{\mathbf{y}}
 = \underbrace{ \begin{bmatrix} \mathbf{C} \\ \mathbf{C} \mathbf{A} \\ \mathbf{C} \mathbf{A}^2 \\ \vdots \\ \mathbf{C} \mathbf{A}^T \end{bmatrix} }_{\mathbfcal{O}} \mathbf{x}(0) +
 \underbrace{ \begin{bmatrix} \mathbf{D} \\ \mathbf{C} \mathbf{B} & \mathbf{D} \\ \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \\ \vdots & \ddots \\ \mathbf{C} \mathbf{A}^{T-1} \mathbf{B} & \cdots  &  \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \end{bmatrix} }_{\mathbfcal{T}} \underbrace{ \begin{bmatrix} u(0) \\ u(1) \\ u(2) \\ \vdots \\ u(T) \end{bmatrix} }_{\mathbf{u}} \end{equation*}
\end{frame}

\begin{frame}[label={slide:preliminaries3}]{A step input ${u} = \bar{{u}} s$ permits a representation \linebreak with an augmented state space }
\begin{itemize}
\item The step input adds a pole at (1,0) to the original system: 
\end{itemize}
\begin{equation*} \begin{aligned} \mathbf{x}_\text{a}(k+1) &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_\text{a}(k) , \quad \text{where} \ \ \mathbf{x}_\text{a}(k) = \begin{bmatrix} \mathbf{x}(k) \\ {u}(k) \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\
{y}(k) &= \underbrace{ \begin{bmatrix} \mathbf{C} & D \end{bmatrix} }_{\mathbf{C}_\text{a}} \mathbf{x}_\text{a}(k) \end{aligned} \end{equation*}
\end{frame}

\begin{frame}[label={slide:preliminaries4}]{Without sensor model, one alternative to estimate the step input  starts by formulating}
\small
\begin{equation*} \begin{aligned} & \underbrace{ \begin{bmatrix} y(1) & y(2) & \cdots & y(n) \\ y(2) & y(3) & \cdots & y(n+1) \\ \iddots & \iddots & \iddots \\ y(n) & y(n+1) & \cdots & y(2n-1) \end{bmatrix} }_{ \mathbfcal{H}({y}) } = \underbrace{ \begin{bmatrix} \widehat{\mathbf{C}}_\text{a} \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a} \\ \vdots \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a}^{n} \end{bmatrix} }_{ \mathbfcal{O}_\text{a} }  \underbrace{ \begin{bmatrix} \mathbf{x}_\text{a}(0) & \mathbf{x}_\text{a}(1) & \cdots & \mathbf{x}_\text{a}(n) \end{bmatrix} }_{ \mathbf{X}_\text{ini} } \end{aligned} \end{equation*}
\normalsize
 A singular value decomposition of $\mathbfcal{H}({y})$ permits the estimation of the observability matrix $\mathbfcal{O}_\text{a}$ and the initial conditions $\mathbf{X}_\text{ini}$ \nolinebreak
\begin{equation*}  \widehat{\mathbfcal{O}}_\text{a} = \mathbf{U} \sqrt{\bm{\Sigma}} \quad \mathrm{and} \quad \widehat{\mathbf{X}}_\text{ini} = \sqrt{\bm{\Sigma}} \mathbf{V}n, \quad \mathrm{from} \quad \mathbf{U} \bm{\Sigma} \mathbf{V} = \mathbfcal{H}({y}) \end{equation*}
\end{frame}

\begin{frame}[label={slide:preliminaries5}]{Once we have estimated the observability matrix $\mathbfcal{O}_\text{a}$ and the initial conditions $\mathbf{X}_\text{ini}$, we can write}
\begin{equation*}  {y} = G \ \bar{{u}} \ + \ \widehat{\mathbfcal{O}}_\text{a} \ \mathbf{x}_\text{a}(0) , \end{equation*}
\begin{itemize}
\item that is equivalent to: 
%\begin{equation*}  \underbrace{ \begin{bmatrix}y(0) \\ y(1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \mathbf{C}_\text{a} \\ G & \mathbf{C}_\text{a} \mathbf{A}_\text{a} \\ \vdots & \vdots \\ G & \mathbf{C}_\text{a} \mathbf{A}_\text{a}^T \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{{u}} \\ \mathbf{x}_\text{a}(0) \end{bmatrix} \end{equation*}
\begin{equation*}  \underbrace{ \begin{bmatrix}y(0) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G \otimes \mathbf{1}_{T+1} & \widehat{\mathbfcal{O}}_\text{a} \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{{u}} \\ \mathbf{x}_\text{a}(0) \end{bmatrix} \end{equation*}
\item and admits a least-squares solution
\begin{equation*}  \begin{bmatrix} \widehat{{u}} \\ \widehat{\mathbf{x}}_{\text{ini}} \end{bmatrix} = \left( \mathbf{K}^\top \mathbf{K} \right)^{-1} \mathbf{K}^\top \mathbf{y} \end{equation*}
\end{itemize}
\end{frame}

\begin{frame}[label={slide:preliminaries6}]{Another alternative to estimate the step input \linebreak starts by differentiating the state-space representation} 
\begin{equation*} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta {y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) \end{aligned} \end{equation*}
\begin{columns}
\begin{column}{0.35\columnwidth}
where: \linebreak $\Delta = \sigma - 1$, and \linebreak $(\sigma^\tau y) (k) = y(k + \tau)$.
\end{column}
\begin{column}{0.65\columnwidth}
If $\Delta {y}$ is persistently exciting of order $L$, 
\begin{equation*} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) \leq L  \end{equation*} 
\end{column}
\end{columns}
\linebreak 
then, we can write the total response of the system as  
%\begin{equation*} \mathbf{y} = G \bar{u} + \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell} \end{equation*}
%that is equivalent to
%\begin{equation*} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ G & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ G & \Delta y(T-n) & \Delta y(T-n+1) & \cdots & \Delta y(T-1) \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} \bar{{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} \end{equation*}
\begin{equation*} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G \otimes \mathbf{1}_{T+1} & \mathbfcal{H}\left(\Delta {y}\right) \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} \bar{{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} \end{equation*}
\end{frame}

\begin{frame}[label={slide:preliminaries7}]{The data-driven step input estimation method is a structured and correlated errors-in-variables problem}
considering the addition of the measurement noise
\begin{equation*} \begin{aligned} \widetilde{\mathbf{y}} &= \mathbf{y} + \bm{\epsilon} , \\
\widetilde{\mathbf{K}} &= \mathbf{K} + \mathbf{E} , \end{aligned} \end{equation*}
we have
\begin{equation*} \underbrace{ \begin{bmatrix} \widetilde{y}(n+1) \\ \vdots \\ \widetilde{y}(T) \end{bmatrix}}_{\widetilde{\mathbf{y}}} = \underbrace{ \begin{bmatrix} G \otimes \mathbf{1}_{T+1} & \mathbfcal{H}\left(\Delta {\widetilde{y}}\right) \end{bmatrix}}_{\widetilde{\mathbf{K}}} \underbrace{ \begin{bmatrix} \bar{{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} \end{equation*}
%and  $\mathbf{E}$ is constructed with the noise data
%\begin{equation*} \mathbf{E} = \begin{bmatrix} 0 & \Delta \epsilon(1) & \Delta \epsilon(2) & \cdots & \Delta \epsilon(n) \\ 0 & \Delta \epsilon(2) & \Delta \epsilon(3) & \cdots & \Delta \epsilon(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ 0 & \Delta \epsilon(T-n) & \Delta \epsilon(T-n+1) & \cdots & \Delta \epsilon(T-1) \end{bmatrix} \end{equation*}
\end{frame}

\begin{frame}[label={slide:statistical1}]{What are the statistical moments of the least-squares solution of a structured and correlated errors-in-variables problem?}
\begin{equation*} \widehat{\bm{\theta}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} \end{equation*}
\end{frame}

\begin{frame}[label={slide:statistical1_2}]{Since the noise is assumed additive}
\begin{equation*} \begin{aligned} \widehat{\bm{\theta}} &= \left( (\mathbf{K}+\mathbf{E})^\top (\mathbf{K}+\mathbf{E})  \right)^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}) \\
&= \left( \mathbf{I} + \mathbf{M} \right)^{-1} \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}) \end{aligned}\end{equation*} 
\linebreak 
where
\begin{equation*} \begin{aligned}  \mathbf{Q} &= \mathbf{K}^\top \mathbf{K}, \quad \text{and} \\
\quad \mathbf{M} &= \mathbf{Q}^{-1} ( \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{E} ). \end{aligned} \end{equation*} 
\end{frame}

\begin{frame}[label={slide:statistical2}]{A second order Taylor series aproximation of the inverse matrix permits to study the LS solution}
Using
\begin{equation*} (\mathbf{I} + \mathbf{M})^{-1} \approx \mathbf{I} - \mathbf{M} + \mathbf{M}^2, \end{equation*} 
the LS solution is approximated by
\begin{equation*} \widehat{\bm{\theta}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}) \end{equation*} 
\linebreak
To find the LS solution bias and covariance, we use
\begin{equation*} \begin{aligned} \bm{\mu} \left(\widehat{\bm{\theta}} \right) &= \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\}, \quad \mathbf{b} \left(\widehat{\bm{\theta}} \right) = \bm{\mu}\left(\widehat{\bm{\theta}} \right) - \bm{\theta} \\
\mathbf{C} \left( \widehat{\bm{\theta}} \right) &= \mathbb{E} \left\{ \left( \widehat{\bm{\theta}} - \bm{\mu} \right) \left( \widehat{\bm{\theta}} - \bm{\mu} \right)^\top \right\} . \end{aligned} \end{equation*} 
\end{frame}

\begin{frame}[label={slide:statistical3}]{For an unstructured and uncorrelated EIV problem, the bias and covariance of the LS solution are}
\begin{equation*} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - T \right) \mathbf{Q}^{-1} \bm{\theta} \\
\\
\mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \bm{\theta} \bm{\theta}^\top \right) \mathbf{Q}^{-1} \\
\quad & - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \ \mathbf{Q}^{-1} \bm{\theta} \bm{\theta}^\top \mathbf{Q}^{-1}  \end{aligned} \end{equation*}
\end{frame}

\begin{frame}[label={slide:statistical4}]{It is proposed to substitute the observed variables in the derived expressions.}
The observed variables are $\widetilde{\mathbf{y}}$, $\widetilde{\mathbf{K}}$, and from them we compute $\widehat{\bm{\theta}}$.
The substitution gives an approximation of the estimation bias and covariance using the observed data.
\begin{equation*} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - T \right) \widetilde{\mathbf{Q}}^{-1} \widehat{\bm{\theta}} \\
\\
\widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_\epsilon^2 \ \widetilde{\mathbf{Q}}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \right) \widetilde{\mathbf{Q}}^{-1} \\
&\quad - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \widetilde{\mathbf{Q}}^{-1} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \widetilde{\mathbf{Q}}^{-1} \end{aligned} \end{equation*}
\end{frame}

\begin{frame}[label={slide:statistical5}]{For the structured and correlated EIV problem in the data-driven step input estimation method}
the bias and covariance of the LS solution are
\begin{equation*} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{Q}^{-1} \left( \left( \mathbf{K}^\top \mathbf{B}_1 - \mathbf{B}_2 \right) \mathbf{x} - \left( \mathbf{K}^\top \mathbf{B}_3 - \mathbf{B}_4 \right) \right) \\ \\
\mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right)  \end{aligned} \end{equation*}
where the matrices $\mathbf{B}_1$ to $\mathbf{B}_1$, and $\mathbf{C}_1$ to $\mathbf{C}_2$ can be obtained considering the structure and correlation of the problem.] 
\end{frame}

\begin{frame}[label={slide:statistical6}]{The substitution of the observed variables in the expressions also approximates the estimation bias and covariance}
\begin{equation*} \begin{aligned} \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{Q}}^{-1} \left( \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\bm{\theta}} - \left( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right) \\ \\
\mathbf{C}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \widetilde{\mathbf{K}}^{\dagger \top} - \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\widetilde{\mathrm{p}}}^\top \left( \widehat{\bm{\theta}} \right) \end{aligned} \end{equation*}
\end{frame}

\begin{frame}[label={slide:statistical7}]{We can find the Cram\'er-Rao lower bound of the structured and correlated errors-in-variables problem}
It is the lower limit on the estimation variance.
\begin{equation*} \text{CRLB}(\bm{\theta}) = \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\bm{\theta}}) }{\partial (\widehat{\bm{\theta}})} \right)^\top \mathbf{Fi}^{-1}(\bm{\theta}) \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\bm{\theta}}) }{\partial (\widehat{\bm{\theta}})} \right) \end{equation*}
Where the Fisher information matrix is  
\begin{equation*} \mathbf{Fi}(x) = - \mathbb{E} \left\{ \frac{\partial ^2 l (\widehat{\bm{\theta}}) }{\partial \widehat{\bm{\theta}}^2 } \right\} \end{equation*}
\end{frame}

\begin{frame}[label={slide:statistical7_2}]{The structured EIV problem can be expressed as a linear in the measurements problem}

\begin{equation*} e (\widehat{\bm{\theta}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\bm{\theta}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\bm{\theta}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 .  \end{equation*}
where $\widetilde{\mathbf{z}} = \mathbf{z} + \epsilon_{\mathbf{z}}$, and $\mathbf{C}_{\mathbf{z}}$ is the covariance matrix of $\epsilon_{\mathbf{z}}$.

Then, the loglikelihood function is
\begin{equation*} \ln{ l(\widetilde{\mathbf{z}}, \widehat{\mathbf{z}}, \widehat{\bm{\theta}}) } = - \frac{1}{2} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right)^\top \mathbf{C}_{\mathbf{z}}^{-1} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right) + \mathrm{constant}, \end{equation*}
and the Fisher information matrix is
\begin{equation*} \mathbf{Fi}(\bm{\theta}) = \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)^\top \left( \mathbf{M}_1( \bm{\theta} ) \ \mathbf{C}_{\mathbf{z}} \ \mathbf{M}_1^\top( \bm{\theta} ) \right)^{-1} \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right) 
 \end{equation*}
\end{frame}

\begin{frame}[label={slide:statistical8}]{Simulation results for unstructured (top) and \linebreak structured EIV problems (bottom)}
\begin{itemize}
\item The absolute value of the bias and the standard error are proportional to the noise variance and standard deviation.
\item The bias predictions coincide with the empirical bias, and 
\item the standard errors are smaller than the estimation bias.
\end{itemize}
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Stat_Fig_1.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:statistical9}]{The relative errors of the bias and variance estimation show the accuracy of the derived expressions}
\begin{itemize}
\item The empirical results are closely approximated 
%for unstructured (top) and structured EIV problems (bottom)
\end{itemize}
\begin{figure}[htb!]
\centering
\includegraphics[width=1\columnwidth]{./fig/Stat_Fig_2.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:statistical10}]{The MSE of the step input estimation is larger than its CRLB at most by one order of magnitude}
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Stat_Fig_3.pdf} 
\includegraphics[width=0.55\columnwidth]{./fig/Stat_Fig_4.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:statistical11}]{Observations of the statistical analysis}
\begin{itemize}
\item The predictions in the structured case are more susceptible to perturbations.
\item The bias and variance predictions accuracy depend on the Taylor series validity.
\item This methodology can be applied to assess the uncertainty to other structured EIV problems solutions.
\item The bias and variance expressions obtained depend on each specific structure.
\end{itemize}
\end{frame}

\begin{frame}[label={slide:experimental-validation1}]{A description of the experimental validation of the step input estimation method}
The diagram represents the setup used to measure mass. 
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Exp_Fig_5.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:experimental-validation2}]{The step input estimation by processing a simulated (left) and an measured (right) sensor response}
A model of the sensor was obtained using the system identification toolbox.
\begin{figure}[htb!]
\centering
\includegraphics[width=1\columnwidth]{./fig/Exp_Fig_1.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:experimental-validation3}]{The bias and variance estimation estimation results allow to evaluate the effectiveness with respect to SNR}
In simulation, with an order $n=5$ the SNR region of validity is largar then 40 dB. 
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Exp_Fig_2.pdf} 
\includegraphics[width=0.55\columnwidth]{./fig/Exp_Fig_3.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:experimental-validation4}]{The uncertainty of the step input estimation does not decrease for larger order $n$}
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Exp_Fig_4.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:experimental-validation5}]{The empirical and the estimation of the bias and variance of the step input for different sample size}
These results are the average from 100 experiments. 
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Exp_Fig_8.pdf} 
\includegraphics[width=0.55\columnwidth]{./fig/Exp_Fig_11.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:experimental-validation6}]{The measurement noise contains frequency components characteristic of mechanical systems}
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Exp_Fig_7.pdf} 
\includegraphics[width=0.55\columnwidth]{./fig/Exp_Fig_9.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:experimental-validation7}]{The uncertainty of the step input estimation does not depend on the order $n$}
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Exp_Fig_12.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:experimental-validation8}]{Observations of the experimental validation of the step input estimation method}
\begin{itemize}
\item In simulation, the input estimation MSE is close to the CRLB theoretical minimum for biased estimators.
\item The step input estimation method is useful in practical applications where the whiteness assumption of the measurement noise is not fulfilled. 
\item The noise variance obtained from the sensor steady state response underestimates the measurement noise variance.
\item Using the predictions, the uncertainty assessment is provided for given sample size and noise variance.
\end{itemize}

\end{frame}

\begin{frame}[label={slide:affine-input-estimation1}]{When an affine input $u=at+b$ is applied to weighing system, it becomes time varying}
\begin{figure}[htb!]
\centering
\includegraphics[width=0.3\columnwidth]{./fig/Aff_Fig_1.pdf} 
\end{figure}
\begin{equation*} \dfrac{d}{dt} \left( \left( \widebar{a} t + \widebar{b} + m \right) \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = \left( \widebar{a} t + \widebar{b} + m \right) g \end{equation*}
\end{frame}

\begin{frame}[label={slide:affine-input-estimation2}]{This is a typical sensor response to an affine input}
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Aff_Fig_2.pdf} 
\includegraphics[width=0.55\columnwidth]{./fig/Aff_Fig_3.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:affine-input-estimation3}]{Data-driven affine input estimation method results with respect to sample size}
In simulation, the input parameters converge towards the true value.

The MSEs come also near the CRLBs, by one order of magnitude. 
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Aff_Fig_4.pdf} 
\includegraphics[width=0.55\columnwidth]{./fig/Aff_Fig_6.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:affine-input-estimation4}]{Data-driven affine input estimation method results with respect to SNR}
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Aff_Fig_5.pdf} 
\includegraphics[width=0.55\columnwidth]{./fig/Aff_Fig_7.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:affine-input-estimation5}]{Maximum-likelihood affine input estimation method results with respect to sample size}
The parameters converge after three iterations, but the required computational power is large.
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Aff_Fig_8.pdf} 
\includegraphics[width=0.55\columnwidth]{./fig/Aff_Fig_9.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:affine-input-estimation6}]{Time varying affine input estimation method results}
A conventional time varying filter, designed for weighing with affine inputs, has a poor performance in the same conditions. 
\begin{figure}[htb!]
\centering
\includegraphics[width=0.55\columnwidth]{./fig/Aff_Fig_10.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slide:affine-input-estimation7}]{Observations of the affine input estimation methods}
\begin{itemize}
\item The subspace method estimates directly the affine input parameters, even when the sensor is time-varying.
\item The subspace method is computationally cheap, simple and suitable for implementation on digital signal processor of low computational power. 
\item The maximum-likelihood method can estimate also model parameters or initial conditions, but needs a lot of computational resources.
\item The proposed methods outperform a conventional time-varying filter.
\end{itemize}
\end{frame}

\begin{frame}[label={slide:conclusions}]{Global conclusions of the conducted research 1}
\begin{itemize}
\item  The data-driven step input estimation method is valid for metrology applications.
\item We have means to assess the estimation uncertainty by analysing the LS solution of a structured and correlated EIV problem.
\item The derived expressions predict the input estimation bias and covariance for given sample size and perturbation level.
\end{itemize}
\end{frame}

\begin{frame}[label={slide:conclusions}]{Global conclusions of the conducted research 2}
\begin{itemize}
\item The input estimation MSE is considerably near to the minimal theoretical variance of the structured EIV problem.
\item The implementation of the data-driven step input estimation method showed robustness under non Gaussian white noise from spurious mechanical vibrations. 
\item The adaptive data-driven affine input estimation method is also robust when processing time-varying sensor responses.
\end{itemize}
\end{frame}

\begin{frame}[label={slide:conclusions}]{Future work}
\begin{itemize}
\item Data-driven signal processing methods will continue to produce interesting results.
\item Statistically efficient input estimation methods can be studied to reduce or eliminate bias.
\item Efficient online optimization methods are required to simplify the  implementation of measurement methods.   
\end{itemize}
\end{frame}

\end{document}

### 2020-05-21 internal-defense/internal-defense-feedback.tex
Final final version before internal defense

% Compile with: 
% cd C:\GusDocs\org\thesis\Gus-thesis\internal-defense\
% pdflatex --shell-escape -synctex=1 -interaction=nonstopmode internal-defense-feedback.tex

\documentclass[12pt,english]{article}
\title{Feedback given during the \linebreak internal-defence presentation rehearsal}
\author{Gustavo Quintana Carapia}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[normalem]{ulem}
\usepackage{algorithm, algorithmic, amsfonts, amsmath, amssymb, amsthm, amsbsy, array, babel, booktabs, bm, capt-of, circuitikz, color, float, graphicx, grffile, hyperref, lineno, longtable, mathdots, mathptmx, mathtools, minted, pgf, rotating, subfigure, textcomp, tikz, times, units, verbatim, wrapfig, xcolor} % 
\usepackage[outdir=./]{epstopdf}
\usetikzlibrary{shapes, arrows, calc, patterns, decorations.pathmorphing, decorations.markings, positioning}

\usepackage[a4paper,bindingoffset=0.2in,%
            left=1in,right=1in,top=1in,bottom=1in,%
            footskip=.25in]{geometry}
\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}

\date{\today}
% Hint: \title{what ever}, \author{who care} and \date{when ever} could stand 
% before or after the \begin{document} command 
% BUT the \maketitle command MUST come AFTER the \begin{document} command! 
\begin{document}

\maketitle

\section{General comments}

\paragraph{Figures}
Reduce information, there is no time to describe everything. Consider the key results. Describe all the shown elements.

\paragraph{Introduction slides}
Reduce the time spent in them.

\paragraph{Presentation}
Rehearse several times and find out what I am doing wrong to correct myself. 
The repetition of the presentation should help me to foresee next slide/topic of the speech and to make it sound more natural. Reduce the artificial voices like 'ah', 'mmm'. Take care on the words articulation. Pay attention to these details.

The pros and cons should be very clear.

Attention getter is not good.

It would be nice to show the Lego implementation  results.

\section{Slides}

\paragraph{1}
$\epsilon$ is not mentioned.
Emphasize why it is important to have bias and variance estimation
Uncertainty bounds are necessary in metrology.
Giving a value without bias/variance bounds is unacceptable in metrology.

\paragraph{3}
Mention the pros and cons of avoiding model approach. Compensator still exhibits transient.

\paragraph{6}
Why do you jump immediately to discrete system? How do you get the discrete time sensor model? For which class of inputs is this method valid?

%\newgeometry{left=0.8in,right=0.8in}

If the input $\mathbf{u}$ belongs to the class of piecewise constant signals
\begin{equation*} \mathbf{u} (t) = \mathbf{u} (kT), \quad \text{where} \quad kT \leq t < (k+1)T \end{equation*}
where $T$ is the sampling time, then, the state-space representation of a linear time-invariant system in continuous time
\begin{equation*} \dot{ \mathbf{x} } = \mathbf{F} \mathbf{x} + \mathbf{G} \mathbf{u}, \quad \mathbf{y} = \mathbf{C} \mathbf{x} + \mathbf{D} \mathbf{u}, \end{equation*} 
corresponds to the discrete-time representation
\begin{equation*}  \mathbf{x} ((k+1)T) = \mathbf{A} \mathbf{x}(kT) + \mathbf{B} \mathbf{u}(kT), \quad \mathbf{y}(kT) = \mathbf{C} \mathbf{x}(kT) + \mathbf{D} \mathbf{u}(kT) , \end{equation*} 
where the matrices $\mathbf{A}$ and $\mathbf{B}$ are given by 
\begin{equation*}  \mathbf{A}  = e^{\mathbf{F}T}, \quad \mathbf{B} = \int_0^T {e^{\mathbf{A} \tau} \mathbf{G} d \tau}. \end{equation*} 
Therefore, when the input $\mathbf{u}$ belongs to the class of piecewise constant signals, the discrete-time state-space equation is an exact representation of the continuous-time system.

%\newgeometry{left=1in,right=1in}

\paragraph{7-8}
Why is D not involved in the equation?

\textit{In the illustration examples, $D$ is a scalar}

\paragraph{10}
Emphasize that the noise makes its way into the matrix K by removing the parametric model. 
This is the price we pay from converting the OE probem into an EIV problem.

\paragraph{14}
Mention that it is a second order Taylor series expansion.

\paragraph{15, 21 and 23}
If you dont mention a variable in the plots, don't show it. Redo figures if necessary.

\paragraph{24}
Mention that the estimation MSE is smallest for orders $n$ larger than the true order. Why is this happening?

\textit{The estimation MSE is smallest for orders $n$ larger than the true order because increasing the order increases
\begin{itemize}
	\item the number of columns in the regression matrix,
	\item the complexity of the problem, and  
	\item the risk for the problem to become ill-posed.
\end{itemize}
The models with order higher than the true order describes all the dynamics of the sensor plus extra dynamics. 
Thus, my hypothesis is that increasing the model order above the true order would increase the variability and reduce the impact of the bias due to disturbing noise.
The results support this hypothesis. }

\paragraph{28}
What is the high level contribution of the ML method?

\paragraph{30, Conclusions}
Focus on the inportant aspects of the work.
\begin{itemize}
	\item the data driven has a price to pay: the noise makes its way into the regression matrix,
	\item the bias and variance are needed to describe the uncertainty of the method,   
	\item the MSE is smaller when the data-driven method assumes an order higher than the true model. Even if the LTI behaves well, the bias due to disturbing noise is captured by increasing the model order beyond the true value.
\end{itemize}

\paragraph{30, Questions}
\begin{itemize}
	\item What is the class of inputs for which the method would work?
	\item What is the most important figure of merit? -> MSE
	\item nothing is said about the confidence bounds for the user,
	\item an important topic has not been mentioned: the statistical analysis has been validated in simulation but there is no sufficient mathematical support to check the error introduced by the substitutions. However, the current state-of-the-art in statistics does not allow the determination of this results for finite sample size. The best we can do, as it has been of common acceptance in the community, is to trust in the results of empirical studies such as the Monte Carlo simulation. There is a rule of thumb: the asymptotic results can be used for finite sample size provided that the number of samples is ten times larger than the number of to-be-estimated parameters.
	\item add to the presentation the two latter 
	\item could you estimate the model parameters given the step input and the step response? No, the step input does not provide enough persistency of excitation to support the identification from input-output.
	\item read the Torsten S\"oderstr\"om paper \& book to have it in perspective and say how the presented work relates to his work  
	\item say how the data-driven approach can be generalized to other identification problems: my idea is to emphasize that the methodology can be followed to find the MSE for solutions to other problems.

\end{itemize}

\end{document}

### 2020-05-21 internal-defense/internal-defense.tex
Final final version before internal defense

% Compile with: 
% cd C:\GusDocs\org\thesis\Gus-thesis\internal-defense\
% pdflatex --shell-escape -synctex=1 -interaction=nonstopmode internal-defense.tex

\documentclass[presentation]{beamer}
%\usepackage[space]{grffile}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[normalem]{ulem}
\usepackage{algorithm, algorithmic, amsfonts, amsmath, amssymb, amsthm, amsbsy, array, babel, booktabs, bm, capt-of, circuitikz, color, float, graphicx, grffile, hyperref, lineno, longtable, mathdots, mathptmx, mathtools, minted, pgf, rotating, subfigure, textcomp, tikz, times, units, verbatim, wrapfig, xcolor} % 
\usepackage[outdir=./]{epstopdf}
\usetikzlibrary{shapes, arrows, calc, patterns, decorations.pathmorphing, decorations.markings, positioning}

%newproof, subcaption, 
%\newtheorem{thm}{Theorem}
%\newtheorem{lem}[thm]{Lemma}
%\newproof{pf}{Proof}

\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}

\input{C:/GusDocs/org/texinputs/mystyle-beamer-org}
\titlegraphic{\includegraphics[width=5cm]{C:/GusDocs/org/texinputs/vublogo/RGB/VUB_logo.pdf}}

\renewcommand{\indent}{\hspace*{5em}}
% \useinnertheme{circles}
\setbeamertemplate{itemize items}[circle]
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}{\hfill \usebeamertemplate***{navigation symbols} \hspace*{0.3cm} \vspace*{0.3cm} \leavevmode \hbox{\begin{beamercolorbox}[wd=0.9\paperwidth,ht=2.25ex,dp=1ex,right]{date in head/foot} \insertframenumber{} \hspace*{1ex} \end{beamercolorbox}} \vskip0pt}
\setbeamerfont{footline}{size=\fontsize{8}{9}}
\addtocounter{framenumber}{0}
\thispagestyle{fancy}
\usetheme{}
\usecolortheme{}

\author{Gustavo Quintana Carapia}
\date{}
\title{Statistical analysis and experimental validation \linebreak of data-driven dynamic measurement methods}

\newcommand\Wider[2][3em]{%
\makebox[\linewidth][c]{%
  \begin{minipage}{\dimexpr\textwidth+#1\relax}
  \raggedright#2
  \end{minipage}%
  }%
}

\begin{document}

\begin{frame}[noframenumbering,plain]
\Wider{\maketitle}
\end{frame}

%\linebreak

\begin{frame}[label={slide:context}]{A measurement is a dynamic process that \linebreak estimates the input from the sensor response}
\begin{itemize}
\item The sensor is a dynamic system.
\item The sensor response is perturbed with errors.
\end{itemize}
\begin{figure}[htb!]
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}] 
 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {Sensor};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini}}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=-0.5cm,yshift=0.25cm] {${y}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=1.0cm,yshift=0.25cm] {$\widetilde{y}$};
 \draw [-latex,thick] (NB1.west) +(-1.0,0) -- +(0.0cm,0);
 \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {${u}$}; 
 \draw (2,0.0) circle (2.5mm);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.00mm,yshift=0.0mm]{+};
 \draw [-latex,thick] (NB1.west) +(3.0,0) -- +(4.0cm,0); 
  \draw [-latex,thick] (NB1.west) ++(2.75cm,1.0) -- +(0.0cm,-0.75);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.250cm,yshift=0.75cm] {$\epsilon$};
\end{tikzpicture}
\end{figure}
\begin{itemize}
	\color{blue}
	\item The aim is to assess the uncertainty of \linebreak the data-driven input estimation method.
\end{itemize}
\end{frame}

\begin{frame}[label={slides:have}]{How is the input estimated from the sensor response?}
\begin{itemize}
	\color{blue}
	\item With sensor model
	\begin{itemize}
		\item Kalman filter.
		\item Compensator system.
	\end{itemize}
\end{itemize}
\begin{figure}[htb!]
 \centering
 \begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}] 
 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {Sensor};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini}}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=-0.5cm,yshift=0.25cm] {${y}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=1.0cm,yshift=0.25cm] {$\widetilde{y}$};
 \draw [-latex,thick] (NB1.west) +(-1.0,0) -- +(0.0cm,0);
 \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {${u}$};
 \draw (2,0.0) circle (2.5mm);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.00mm,yshift=0.0mm]{+};
 \draw [-latex,thick] (NB1.west) +(3.0,0) -- +(4.0cm,0); 
  \draw [-latex,thick] (NB1.west) ++(2.75cm,1.0) -- +(0.0cm,-0.75);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.250cm,yshift=0.75cm] {$\epsilon$};
 \node (NB2) [minimum width=1.5cm, minimum height=0.75cm, xshift=4.5cm, yshift=0.0cm] {Compensator}; 
 \draw [-latex,thick] (NB2.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB2,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini,c}}$};
 \draw [-latex,thick] (NB2.east) ++(0,0) -- +(1.0cm,0);
 \node[draw=none,fill=none] [right=of NB2,xshift=-0.5cm,yshift=0.25cm] {$\widehat{{u}}$};
 \end{tikzpicture}
\end{figure}
\begin{itemize}
	\color{blue}
	\item Without sensor model
	\begin{itemize}
	\item System identification and input estimation.
	\item Optimization.
	\item Data-driven methods.
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={slide:have}]{Digital Signal Processors \linebreak offer an alternative to classical approaches}
A DSP, with appropriate algorithms, can 
\begin{itemize}
	\item emulate dynamical systems, or 
	\item implement data-driven methods. 
\end{itemize}
\begin{figure}[htb!]
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}] 
 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {Sensor};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini}}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=-0.5cm,yshift=0.25cm] {${y}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=1.0cm,yshift=0.25cm] {$\widetilde{y}$};
 \draw [-latex,thick] (NB1.west) +(-1.0,0) -- +(0.0cm,0);
 \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {${u}$};
 \draw (2,0.0) circle (2.5mm);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.00mm,yshift=0.0mm]{+};
 \draw [-latex,thick] (NB1.west) +(3.0,0) -- +(4.0cm,0); 
  \draw [-latex,thick] (NB1.west) ++(2.75cm,1.0) -- +(0.0cm,-0.75);
 \node[draw=none,fill=none] [right=of NB1,xshift=0.250cm,yshift=0.75cm] {$\epsilon$};
\node (NB2) [minimum width=1.5cm, minimum height=0.75cm, xshift=4.0cm, yshift=0.0cm] {DSP}; 
 \draw [-latex,thick] (NB2.east) ++(0,0) -- +(1.0cm,0);
 \node[draw=none,fill=none] [right=of NB2,xshift=-0.5cm,yshift=0.25cm] {$\widehat{{u}}$};
 \end{tikzpicture}
 \end{figure}
 \begin{itemize}
 	\color{blue}
	\item The statistical properties of data-driven methods \linebreak are not straightforward  evident. 
\end{itemize}
\end{frame}

\begin{frame}[label={slides:preview}]{Statistical analysis and experimental validation \linebreak of data-driven dynamic measurement methods}
\begin{itemize}
	\item Formulation of a data-driven step input estimation method.
	\vspace{0.5cm}
	\color{gray}
	\item Thesis contribution.
	\begin{itemize}
		\color{gray}
 		\item Statistical analysis, structured EIV problems.
		\item Experimental validation.
		\item Affine input estimation methods.
 	\end{itemize}
\end{itemize}
\color{black}
\end{frame}

\begin{frame}[label={slide:preliminaries1}]{The input estimation methods are \linebreak formulated as output-error problems}
\begin{figure}[htb!]
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}] 
 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {$\left( \mathbf{A}, \mathbf{B}, \mathbf{C}, D \right)$};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{ini}}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=-0.5cm,yshift=0.25cm] {${y}$};
 \draw [-latex,thick] (NB1.west) +(-1.0,0) -- +(0.0cm,0);
 \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {${u}$}; 
% \draw (2.3,0.0) circle (2.5mm);
% \node[draw=none,fill=none] [right=of NB1,xshift=1.0cm,yshift=0.25cm] {${y}$};
% \node[draw=none,fill=none] [right=of NB1,xshift=0.00mm,yshift=0.0mm]{+};
% \draw [-latex,thick] (NB1.west) +(3.6,0) -- +(4.6cm,0); 
% \draw [-latex,thick] (NB1.west) ++(3.3cm,1.0) -- +(0.0cm,-0.75);
% \node[draw=none,fill=none] [right=of NB1,xshift=0.250cm,yshift=0.75cm] {$\epsilon$};
\end{tikzpicture}
\end{figure}
\begin{itemize}
	\item With model and exact data, we can write
	\vspace{2mm}
    %\item  With LTI sensor, the measurement dynamics \linebreak can be described using state-space representation: 
\end{itemize}
    %\begin{equation*} \begin{aligned} \mathbf{x}(k+1) &= \mathbf{A} \mathbf{x}(k) + \mathbf{B} {u}(k), \quad \text{with} \quad \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\ {y}(k) &= \mathbf{C} \mathbf{x}(k) + {D} {u}(k) + {\epsilon}(k) \end{aligned} \end{equation*}
    \begin{equation*} \underbrace{ \begin{bmatrix} y(0) \\ y(1) \\ y(2) \\ \vdots \\ y(N) \end{bmatrix} }_{\mathbf{y}}  = \underbrace{ \begin{bmatrix} \mathbf{C} \\ \mathbf{C} \mathbf{A} \\ \mathbf{C} \mathbf{A}^2 \\ \vdots \\ \mathbf{C} \mathbf{A}^N \end{bmatrix} }_{\mathbfcal{O}} \mathbf{x}(0) + \underbrace{ \begin{bmatrix} \mathbf{D} \\ \mathbf{C} \mathbf{B} & \mathbf{D} \\ \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \\ \vdots & \ddots \\ \mathbf{C} \mathbf{A}^{N-1} \mathbf{B} & \cdots  &  \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \end{bmatrix} }_{\mathbfcal{T}} \underbrace{ \begin{bmatrix} u(0) \\ u(1) \\ u(2) \\ \vdots \\ u(N) \end{bmatrix} }_{\mathbf{u}} \end{equation*}
\end{frame}

\begin{frame}[label={slide:preliminaries3}]{With a step input, the sensor admits an \linebreak augmented autonomous state space representation}
\begin{figure}
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}] 
 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {$ \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}}, \underbrace{ \begin{bmatrix} \mathbf{C} & D \end{bmatrix} }_{\mathbf{C}_\text{a}} $};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-0.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\mathbf{x}_{\text{a,ini}}$};
 \node[draw=none,fill=none] [right=of NB1,xshift=-0.5cm,yshift=0.25cm] {${y}$};
% \draw [-latex,thick] (NB1.west) +(-1.5,0) -- +(0.0cm,0);
% \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {$u=\bar{u}s$}; 
% \draw (2.7,0.0) circle (2.5mm);
% \node[draw=none,fill=none] [right=of NB1,xshift=1.0cm,yshift=0.25cm] {${y}$};
% \node[draw=none,fill=none] [right=of NB1,xshift=0.00mm,yshift=0.0mm]{+};
% \draw [-latex,thick] (NB1.west) +(4.4,0) -- +(5.4cm,0); 
% \draw [-latex,thick] (NB1.west) ++(4.2cm,1.0) -- +(0.0cm,-0.75);
% \node[draw=none,fill=none] [right=of NB1,xshift=0.250cm,yshift=0.75cm] {$\epsilon$};
\end{tikzpicture}
\end{figure}
\begin{itemize}
	\item Without sensor model, we can estimate first $\mathbfcal{O}_\text{a}$ from
	\vspace{2mm}
   %\item The step input adds a pole at (1,0) to the original system: 
\end{itemize}
  %\begin{equation*} \begin{aligned} \mathbf{x}_\text{a}(k+1) &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_\text{a}(k) , \quad \text{where} \ \ \mathbf{x}_\text{a}(k) = \begin{bmatrix} \mathbf{x}(k) \\ {u}(k) \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\ {y}(k) &= \underbrace{ \begin{bmatrix} \mathbf{C} & D \end{bmatrix} }_{\mathbf{C}_\text{a}} \mathbf{x}_\text{a}(k) \end{aligned} \end{equation*}
  \small
\begin{equation*} \begin{aligned} & \underbrace{ \begin{bmatrix} y(1) & y(2) & \cdots & y(n) \\ y(2) & y(3) & \cdots & y(n+1) \\ \iddots & \iddots & \iddots \\ y(n) & y(n+1) & \cdots & y(2n-1) \end{bmatrix} }_{ \mathbfcal{H}({y}) } = \underbrace{ \begin{bmatrix} {\mathbf{C}}_\text{a} \\ {\mathbf{C}}_\text{a} {\mathbf{A}}_\text{a} \\ \vdots \\ {\mathbf{C}}_\text{a} {\mathbf{A}}_\text{a}^{n} \end{bmatrix} }_{ {\mathbfcal{O}}_\text{a} }  \underbrace{ \begin{bmatrix} \mathbf{x}_\text{a}(1) & \mathbf{x}_\text{a}(2) & \cdots & \mathbf{x}_\text{a}(n) \end{bmatrix} }_{ \mathbf{X}_\text{ini} } \end{aligned} \end{equation*}
\normalsize
\end{frame}

\begin{frame}[label={slide:preliminaries5}]{After estimating $\widehat{\mathbfcal{O}}_\text{a}$, we can write \linebreak the total response of the system}
\begin{equation*}  {y} = G \ u \ + \ \widehat{\mathbfcal{O}}_\text{a} \ \mathbf{x}_\text{a}(0) , \end{equation*}
\begin{itemize}
\item that is equivalent to
%\begin{equation*}  \underbrace{ \begin{bmatrix}y(0) \\ y(1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \mathbf{C}_\text{a} \\ G & \mathbf{C}_\text{a} \mathbf{A}_\text{a} \\ \vdots & \vdots \\ G & \mathbf{C}_\text{a} \mathbf{A}_\text{a}^T \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{{u}} \\ \mathbf{x}_\text{a}(0) \end{bmatrix} \end{equation*}
\begin{equation*}  \underbrace{ \begin{bmatrix} y(0) \\ \vdots \\ y(N) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{1}_{N+1} \otimes G & \widehat{\mathbfcal{O}}_\text{a} \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} u \\ \mathbf{x}_\text{a}(0) \end{bmatrix} \end{equation*}
% \item and admits a least-squares solution
% \begin{equation*}  \begin{bmatrix} \widehat{{u}} \\ \widehat{\mathbf{x}}_{\text{ini}} \end{bmatrix} = \left( \mathbf{K}^\top \mathbf{K} \right)^{-1} \mathbf{K}^\top \mathbf{y} \end{equation*}
\item where $G$ is the sensor static gain.
\end{itemize}
\end{frame}

\begin{frame}[label={slide:preliminaries6}]{To estimate directly the step input,  obtain \linebreak the first difference of the state-space representation} 
\begin{figure}
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}] 
 \node (NB1) [minimum width=1.5cm,minimum height=0.75cm,xshift=0.0cm, yshift=0.0cm] {\begin{equation*} \begin{aligned} \Delta \mathbf{x}(k+1) &= \mathbf{A} \Delta \mathbf{x}(k) \\ \Delta {y}(k) &= \mathbf{C} \Delta \mathbf{x}(k) \end{aligned} \end{equation*}};
 \draw [-latex,thick] (NB1.east) ++(0,0) -- +(1.0cm,0);
 \draw [-latex,thick] (NB1.north) ++(-1.5cm,0.5) -- +(0.0cm,-0.5);
 \node[draw=none,fill=none] [above=of NB1,xshift=0.0cm,yshift=-1.0cm] {$\Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0)$};
 \node[draw=none,fill=none] [right=of NB1,xshift=-0.5cm,yshift=0.25cm] {$\Delta y$};
% \draw [-latex,thick] (NB1.west) +(-1.5,0) -- +(0.0cm,0);
% \node[draw=none,fill=none] [left=of NB1, xshift=0.750cm,yshift=0.25cm] {$u=\bar{u}s$}; 
% \draw (2.7,0.0) circle (2.5mm);
% \node[draw=none,fill=none] [right=of NB1,xshift=1.0cm,yshift=0.25cm] {${y}$};
% \node[draw=none,fill=none] [right=of NB1,xshift=0.00mm,yshift=0.0mm]{+};
% \draw [-latex,thick] (NB1.west) +(4.4,0) -- +(5.4cm,0); 
% \draw [-latex,thick] (NB1.west) ++(4.2cm,1.0) -- +(0.0cm,-0.75);
% \node[draw=none,fill=none] [right=of NB1,xshift=0.250cm,yshift=0.75cm] {$\epsilon$};
\end{tikzpicture}
\end{figure}
%\begin{equation*} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta {y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) \end{aligned} \end{equation*}
where: $\Delta \mathbf{x}(k) = \mathbf{x}(k+1) - \mathbf{x}(k)$ 
\vspace{4mm} 
% \begin{columns}
%  \begin{column}{0.375\columnwidth}
%   where: \linebreak $\Delta \mathbf{x}(k) = \mathbf{x}(k+1) - \mathbf{x}(k)$ 
%   %\newline
%  \end{column}
%   \begin{column}{0.625\columnwidth}
%    If $\Delta {y}$ is persistently exciting of order $L$, \vspace{-4mm} 
%    \[ \text{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) \leq L \] 
%   \end{column}
% \end{columns}
\begin{itemize}
\item If $\Delta {y}$ is persistently exciting enough, the total response is  
\end{itemize}
%\begin{equation*} \mathbf{y} = G \bar{u} + \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell} \end{equation*}
%that is equivalent to
%\begin{equation*} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ G & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ G & \Delta y(T-n) & \Delta y(T-n+1) & \cdots & \Delta y(T-1) \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} \bar{{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} \end{equation*}
\begin{equation*} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(N) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} \mathbf{1}_{N-n} \otimes G & \mathbfcal{H}\left(\Delta {y}\right) \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} u \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} \end{equation*}
\end{frame}

\begin{frame}[label={slide:preliminaries7}]{The data-driven step input estimation method converts \linebreak  the output-error into an errors-in-variables problem}
\begin{equation*} \underbrace{ \begin{bmatrix} \widetilde{y}(n+1) \\ \vdots \\ \widetilde{y}(N) \end{bmatrix}}_{\widetilde{\mathbf{y}}} = \underbrace{ \begin{bmatrix} \mathbf{1}_{N-n} \otimes G & \mathbfcal{H}\left(\Delta {\widetilde{y}}\right) \end{bmatrix}}_{\widetilde{\mathbf{K}}} \underbrace{ \begin{bmatrix} u \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} \end{equation*}
%and  $\mathbf{E}$ is constructed with the noise data
%\begin{equation*} \mathbf{E} = \begin{bmatrix} 0 & \Delta \epsilon(1) & \Delta \epsilon(2) & \cdots & \Delta \epsilon(n) \\ 0 & \Delta \epsilon(2) & \Delta \epsilon(3) & \cdots & \Delta \epsilon(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ 0 & \Delta \epsilon(T-n) & \Delta \epsilon(T-n+1) & \cdots & \Delta \epsilon(T-1) \end{bmatrix} \end{equation*}
\begin{itemize}
	\item additive measurement noise $\widetilde{\mathbf{y}} = \mathbf{y} + \bm{\epsilon},$ and $\widetilde{\mathbf{K}} = \mathbf{K} + \mathbf{E}$, 
	\item structured and correlated, 
	\item no evident statistical properties,
	\item hard to get unbiased solution, 
	\item admits a least-squares solution.
\end{itemize}
\end{frame}

\begin{frame}[label={slide:need}]{The data-driven step input estimation \linebreak uncertainty assessment was pending}
This requires to find 
\begin{itemize}
	\item the structured errors-in-variables problem 
	\begin{itemize}
		\item stochastic properties,
	\end{itemize}
	\item the data-driven step input estimation 
	\begin{itemize}
		\item statistical moments, 
		\item mean-squared-error,  
	\end{itemize}
\end{itemize}
\vspace{5mm}
\begin{itemize}
	\color{blue}
	\item The uncertainty assessment
	\begin{itemize}
		\color{blue}
		\item provides confidence bounds w.r.t sample size,
		\item fosters the method utilization in metrology.
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={slides:preview}]{Statistical analysis and experimental validation \linebreak of data-driven dynamic measurement methods}
\begin{itemize}
	\color{gray}
	\item Formulation of a data-driven step input estimation method.
	\vspace{0.5cm}
	\color{black}
	\item Thesis contribution.
	\begin{itemize}
 		\item Statistical analysis, structured EIV problems.
		\color{gray}
		\item Experimental validation.
		\item Affine input estimation methods.
 	\end{itemize}
\end{itemize}
\color{black}
\end{frame}

\begin{frame}[label={slide:statistical1}]{To study the least-squares solution of \linebreak an errors-in-variables problem}
\begin{equation*} \widehat{\bm{\theta}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \end{equation*}
%\vspace{5mm}
where the noise is assumed additive
%\vspace{-2.5mm} 
\begin{equation*} \widehat{\bm{\theta}} = \left( (\mathbf{K}+\mathbf{E})^\top (\mathbf{K}+\mathbf{E})  \right)^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}) , \end{equation*} 
\vspace{5mm} 
\begin{itemize}
	\color{blue}
	\item I calculated 
	\begin{itemize}
		\color{blue}
		\item the second order Taylor series expansion of $\widehat{\bm{\theta}}$. 
		\item  expressions that predict the bias and covariance,
		\begin{itemize}
			\color{blue}
			\item for unstructured and structured EIV problems.  
		\end{itemize}
	\end{itemize}
\end{itemize}

%\vspace{-2.5mm} 
%\begin{equation*} \begin{aligned} \widehat{\bm{\theta}} &= \big( \mathbf{I} + \underbrace{\mathbf{Q}^{-1} ( \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{E} )}_{\mathbf{M}} \big)^{-1} \big( \underbrace{\mathbf{K}^\top \mathbf{K}}_{\mathbf{Q}} \big)^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}) \end{aligned}\end{equation*} 
\end{frame}

\begin{frame}[label={slide:statistical8}]{Monte Carlo simulation shows that the expressions for structured EIV  are more susceptible to perturbations}
% \begin{itemize}
% \item The absolute value of the bias and the standard error are proportional to the noise variance and standard deviation.
% \item The bias predictions coincide with the empirical bias, and 
% \item the standard errors are smaller than the estimation bias.
% \end{itemize}
\begin{columns}
\begin{column}{0.35\columnwidth}
\begin{itemize}
	\item Unstructured EIV \linebreak
	\hspace*{-5mm} - uncorrelated, \linebreak
	\hspace*{-5mm} - $\begin{bmatrix} \mathbf{K} &\mathbf{y}\end{bmatrix}$ randomly generated. \linebreak
	\newline
	\item Structured EIV \linebreak
	\hspace*{-5mm} - step input estimation from $2^{\text{nd}}$ order model response. 
\end{itemize}
% \includegraphics[width=1\columnwidth]{./fig/Stat_Fig_4.pdf} 
\end{column}
\begin{column}{0.65\columnwidth}
 \begin{figure}
  \centering
  \includegraphics[width=1\columnwidth]{./fig/Stat_Fig_1u.pdf} 
 \end{figure}
 \begin{figure}
  \centering
  \includegraphics[width=1\columnwidth]{./fig/Stat_Fig_1s.pdf} 
 \end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[label={slide:statistical10}]{For a $2^{\text{nd}}$ order system, the step input estimation MSE is larger than the CRLB by one order of magnitude}
\begin{figure}
 \centering
 \includegraphics[width=0.65\columnwidth]{./fig/Stat_Fig_3.pdf} 
\end{figure}
\end{frame}

\begin{frame}[label={slides:preview}]{Statistical analysis and experimental validation \linebreak of data-driven dynamic measurement methods}
\begin{itemize}
	\color{gray}
	\item Formulation of a data-driven step input estimation method.
	\vspace{0.5cm}
	\color{black}
	\item Thesis contribution.
	\begin{itemize}
		\color{gray}
		\item Bias and variance analysis, structured EIV problems.
		\color{black}
		\item Experimental validation.
		\color{gray}
		\item Affine input estimation methods.
 	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={slide:experimental-validation1}]{I designed and built a weighing system \linebreak to collect experimental data from step inputs}
\ctikzset{bipoles/resistor/height=0.075}
\ctikzset{bipoles/resistor/width=0.25}
\begin{figure}[!htb]
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
\tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
\tikzstyle{damper}=[thick,decoration={markings,  
  mark connection node=dmp,
  mark=at position 0.5 with 
  {
    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
  }
}, decorate]
\tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]

% load cell
\node (M1) [minimum width=1.6cm,minimum height=0.4cm, xshift=-3cm, yshift=-1cm] {};

\node (ground1) at (M1.east) [ground,yshift=0.0cm,rotate=90,xshift=0.0cm,anchor=north] {};
\draw (ground1.north east) -- (ground1.north west);

\node at (M1.north) [fill=none, xshift=-0.75cm, yshift=0.750cm] (2.0cm, 1.50cm) {\small Mass};
\draw (M1.north) [-latex,thick]  ++(-0.75cm,0.5cm) -- +(0cm,-0.5cm);
\draw (-3.3750cm,-0.935cm) arc(30:330:1.25mm);
\draw (-2.625cm,-1.05cm) arc(210:510:1.25mm);
\draw (M1.west) [thick]  ++(0.4cm, 0.0750cm) -- +(0.80cm,0.0cm);
\draw (M1.west) [thick]  ++(0.4cm,-0.0750cm) -- +(0.80cm,0.0cm);

\node at (M1.south) [align=center, draw=none,fill=none, xshift=0.0cm, yshift=-0.5cm] (2.0cm, 1.50cm) {\small load cell \\ \small Tedea 1004};

% amplifier
\draw (-1.6,-0.0) -- (-1.6,-2.0) -- (1.2,-1) -- cycle;
    \draw (-1.5,-1) to[R] (-1.1,-0.6) to[R] (-0.7,-1);
    \draw (-1.5,-1) to[R] (-1.1,-1.4) to[R] (-0.7,-1);
\draw [-latex,thick]  ++(1.2,-1) -- +(0.8,0);

\node (filter) [minimum width=0.50cm,minimum height=0.50cm, xshift=-0.0cm, yshift=-1cm] {};
\draw [thick]  ++(-0.275,-0.9) -- +(0.275,0.0);
\draw [thick]  ++(0,-0.9) -- +(0.25,-0.250);
\node at (M1.east) [draw=none, fill=none, xshift=4.0cm, yshift=0.25cm] (2.0cm, 1.50cm) {$\widetilde{y}$};
% wire
\draw plot [smooth] coordinates {(-2.3,-0.8) (-2.2,-0.5) (-1.7,-0.5) (-1.65,-0.9) (-1.6,-1)};

\node at (M1.south) [align=center, draw=none,fill=none, xshift=3.75cm, yshift=-0.750cm] (2.0cm, 1.50cm) {\small amplifier and filter};

\node (ADC) [align=center, minimum width=1.2cm,minimum height=0.9cm, xshift=2.6cm, yshift=-1cm] {\small ADC \\ \small 16 bits};
\end{tikzpicture}
\end{figure}
\begin{itemize}
	\item A $5^{\text{th}}$ order sensor model was obtained \linebreak with the sysId toolbox.
\end{itemize}
\end{frame}

\begin{frame}[label={slide:experimental-validation2}]{These are typical step input estimation results \linebreak produced by the data-driven method}
\begin{figure}[htb!]
\centering
\hspace*{-5mm} \includegraphics[width=1.1\columnwidth]{./fig/Exp_Fig_1.pdf} 
\end{figure}
\begin{columns}
\begin{column}{0.5\columnwidth}
Processing a $5^{\text{th}}$ order \linebreak model response.
\end{column}
\begin{column}{0.5\columnwidth}
Processing a measured \linebreak sensor response.
\end{column}
\end{columns}
\end{frame}

\begin{frame}[label={slide:experimental-validation3}]{Monte Carlo simulation shows that, for SNR > 40 dB \linebreak the bias and variance are well predicted}
\begin{figure}
 \centering
 \includegraphics[width=0.65\columnwidth]{./fig/Exp_Fig_2.pdf} 
\end{figure}
\begin{figure}
 \centering
 \includegraphics[width=0.65\columnwidth]{./fig/Exp_Fig_3.pdf} 
\end{figure}
\begin{itemize}
	\color{blue}
	\item The data-driven step input estimation MSE is at most \linebreak one order of magnitude larger than the EIV problem CRLB.
\end{itemize}
\end{frame}

\begin{frame}[label={slide:experimental-validation6}]{There are periodic components \linebreak in the measurement noise}
\begin{figure}
\centering
\includegraphics[width=0.65\columnwidth]{./fig/Exp_Fig_7.pdf} 
\end{figure}
\begin{figure}
\centering 
\includegraphics[width=0.65\columnwidth]{./fig/Exp_Fig_9.pdf} 
\end{figure}
\begin{itemize}
	\color{blue}
	\item Considering the residuals, \linebreak the SNR was adjusted from 55 dB to 50 dB. 
\end{itemize}
\end{frame}

\begin{frame}[label={slide:experimental-validation5}]{After processing 100 measured responses, \linebreak the bias and variance decrease w.r.t. sample size} 
% \begin{figure}
%  \centering
%  \includegraphics[width=0.6\columnwidth]{./fig/Exp_Fig_8.pdf} 
% \end{figure}
\begin{figure}
 \centering
 \includegraphics[width=0.65\columnwidth]{./fig/Exp_Fig_11.pdf} 
\end{figure}
% \begin{itemize}
% 	\color{blue}
% 	\item The step input estimation method is useful \linebreak when the measurement noise is not Gaussian white. 
% \end{itemize}
\end{frame}

\begin{frame}[label={slide:experimental-validation4}]{The estimation MSE is smallest \linebreak for orders larger than the true order}
\begin{figure}
  \centering
  \includegraphics[width=0.65\columnwidth]{./fig/Exp_Fig_4.pdf} 
\end{figure}
\begin{itemize}
$10^4$ simulated responses from $5^{\text{th}}$ order model.
\end{itemize}
\end{frame}

% \begin{frame}[label={slide:experimental-validation8}]{Observations of the experimental validation of the step input estimation method}
% \begin{itemize}
% \item In simulation, the input estimation MSE is close to the CRLB theoretical minimum for biased estimators.
% \item The step input estimation method is useful in practical applications where the whiteness assumption of the measurement noise is not fulfilled. 
% \item The noise variance obtained from the sensor steady state response underestimates the measurement noise variance.
% \item Using the predictions, the uncertainty assessment is provided for given sample size and noise variance.
% \end{itemize}
% \end{frame}

\begin{frame}[label={slides:preview}]{Statistical analysis and experimental validation \linebreak of data-driven dynamic measurement methods}
\begin{itemize}
	\color{gray}
	\item Formulation of a data-driven step input estimation method.
	\vspace{0.5cm}
	\item Thesis contribution.
	\begin{itemize}
 		\color{gray}
 		\item Statistical analysis, structured EIV problems.
		\item Experimental validation.
		\color{black}
		\item Affine input estimation methods.
 	\end{itemize}
\end{itemize}
\color{black}
\end{frame}

\begin{frame}[label={slide:affine-input-estimation1}]{Can we estimate affine inputs \linebreak with a data-driven method?}
\begin{figure}
\centering
\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
\tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
\tikzstyle{damper}=[thick,decoration={markings,  
  mark connection node=dmp,
  mark=at position 0.5 with 
  {
    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
  }
}, decorate]
\tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]

\node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
\node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u=at+b$};

\node (ground1) at (M.south) [ground,yshift=-1.5cm,xshift=-0.625cm,anchor=north] {};
\draw (ground1.north west) -- (ground1.north east);
\draw [spring] (ground1.north) -- ($(M.south east)!(ground1.north)!(M.south west)$);

\node (groundc) at (M.south) [ground,yshift=-1.5cm,anchor=north] {}; 
\draw (groundc.north west) -- (groundc.north east);

\node (ground2) at (M.south) [ground,yshift=-1.5cm,xshift=0.625cm,anchor=north] {};
\draw (ground2.north west) -- (ground2.north east);
\draw [damper] (ground2.north) -- ($(M.south east)!(ground2.north)!(M.south west)$);

\node[draw=none,fill=none] at (-0.9cm,-1cm) {$k_{\mathrm{s}}$};
\node[draw=none,fill=none] at (0.15cm,-1cm) {$k_{\mathrm{d}}$};
\node[draw=none,fill=none] at (2.0cm,1.0cm) {$y$};
\draw [-latex,thick]  ++(2.2cm,-1cm) -- +(0cm,2.25cm);

\draw [-latex,thick] (M.east) ++(0,0) -- +(1cm,0);
\draw [line width=0.25mm] (2.2cm,-1cm) -- (2.2cm,1cm);
\draw [line width=0.25mm] (2.1cm,-1cm) -- (2.3cm,-1cm);
\draw [line width=0.25mm] (2.1cm,1cm) -- (2.3cm,1cm);
\draw [line width=0.25mm] (2.1cm,-0.5cm) -- (2.3cm,-0.5cm);
\draw [line width=0.25mm] (2.1cm,0.5cm) -- (2.3cm,0.5cm);
\draw [line width=0.25mm] (2.15cm,-0.25cm) -- (2.25cm,-0.25cm);
\draw [line width=0.25mm] (2.15cm,0.25cm) -- (2.25cm,0.25cm);
\draw [line width=0.25mm] (2.15cm,-0.75cm) -- (2.25cm,-0.75cm);
\draw [line width=0.25mm] (2.15cm,0.75cm) -- (2.25cm,0.75cm);
\draw [line width=0.25mm] (2.1cm,0cm) -- (2.3cm,0cm);
\end{tikzpicture}
\end{figure}
\begin{itemize}
	\item An affine input turns this weighing system into time-varying.
\end{itemize}
%\begin{equation*} \dfrac{d}{dt} \left( \left( \widebar{a} t + \widebar{b} + m \right) \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = \left( \widebar{a} t + \widebar{b} + m \right) g \end{equation*}
\begin{figure}[htb!]
\centering
\includegraphics[width=0.65\columnwidth]{./fig/Aff_Fig_3.pdf} 
\end{figure}
\end{frame}

%\setbeamerfont{frametitle}{size=\small}
\begin{frame}[label={slide:affine-input-estimation3}]{A data-driven affine input estimation method is adapted using exponentially weighted recursive LS}
\begin{figure}
\centering
\includegraphics[width=0.65\columnwidth]{./fig/Aff_Fig_4.pdf} 
\end{figure}
\begin{figure}
\centering
\hspace*{-3.5mm} \includegraphics[width=0.67\columnwidth]{./fig/Aff_Fig_6.pdf} 
\end{figure}
\begin{itemize}
	\color{blue}
	\item Inherits statistical properties from \linebreak the data-driven step input estimation method. 
\end{itemize}
\end{frame}

\begin{frame}[label={slide:affine-input-estimation5}]{A maximum-likelihood affine input estimation method estimates simultaneously also the initial conditions}
\begin{figure}
\centering
\includegraphics[width=0.6\columnwidth]{./fig/Aff_Fig_8.pdf} 
\end{figure}
\begin{figure}
\centering
\hspace*{0.75mm} \includegraphics[width=0.59\columnwidth]{./fig/Aff_Fig_9.pdf} 
\end{figure}
\begin{itemize}
	\color{blue}
	\item The ML method provides the best possible estimation, \linebreak but requires large computational power. 
\end{itemize}
\end{frame}

\begin{frame}[label={slide:affine-input-estimation6}]{The proposed methods outperform \linebreak a conventional time-varying filter}
\begin{itemize}
	\color{blue}
	\item Processing the same sensor response. 
	\begin{itemize}
		\item Data-driven method:
		\begin{figure}
  			\centering
  			\includegraphics[width=0.65\columnwidth]{./fig/Aff_Fig_4.pdf} 
		\end{figure}
		\item Time-varying filter:
		\begin{figure}
  			\centering
  			\includegraphics[width=0.65\columnwidth]{./fig/Aff_Fig_10.pdf} 
		\end{figure}
	\end{itemize}
\end{itemize}
\end{frame}

% \begin{frame}[label={slide:affine-input-estimation7}]{Observations of the affine input estimation methods}
% \begin{itemize}
% \item The subspace method estimates directly the affine input parameters, even when the sensor is time-varying.
% \item The subspace method is computationally cheap, simple and suitable for implementation on digital signal processor of low computational power. 
% \item The maximum-likelihood method can estimate also model parameters or initial conditions, but needs a lot of computational resources.
% \item The proposed methods outperform a conventional time-varying filter.
% \end{itemize}
% \end{frame}

\begin{frame}[label={slide:conclusions}]{In summary}
\begin{itemize}
	\item In the data-driven step input estimation method
	\vspace{3mm}
	\begin{itemize}
		\item the noise enters in the regression matrix,
		\vspace{2mm}
		\item the obtained bias and variance expressions can be used to describe the input uncertainty,
		\vspace{2mm}
		\item users can set the sample size to reach a required MSE,
		\vspace{2mm}
		\item the MSE is smallest for large model order.
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={slide:conclusions}]{Conclusions}
\begin{itemize}
	\item  The data-driven input estimation methods
	\vspace{3mm}
	\begin{itemize}
		\item are valid for metrology applications,
		\vspace{2mm}
		\item are completed with the uncertainty assessment,
		\vspace{2mm}
		\item are robust under non Gaussian white noise,
		\vspace{2mm}
		\item can process responses from time-varying sensors.  
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={slide:conclusions}]{Future work}
\begin{itemize}
	\item Design and implementation of
	\vspace{3mm}
	\begin{itemize}
		\item data-driven methods to estimate other input models,
		\vspace{2mm}
		\item efficient solution methods for structured EIV problems,
		\vspace{2mm}
		\item efficient online optimization methods to reduce its computational burden.
	\end{itemize}   
\end{itemize}
\end{frame}

\begin{frame}
\Wider{\maketitle}
\end{frame}

\end{document}

### 2020-08-21 ChapterAppendices/ChapterAppendices.tex
Final version before internal defense
Second draft before public defense

@@
 
-\section{Derivation of bias and covariance expressions.  \label{appendix:biascovSEIV} }
+\section{Derivation of bias and covariance expressions  \label{appendix:biascovSEIV} }
 
@@
 After an elementwise evaluation of the corresponding expected values in (\ref{eqn:EC_Eu}), the expressions result in 
-\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\} & \approx \bm{\theta}  + \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) = \bm{\theta}  +  \sigma_{\mathbf{E}}^2 \mathbf{Q}^{-1} \left( 2\mathbf{I} + 2n\mathbf{I} - T \mathbf{I} \right) \bm{\theta} , \quad \text{and} \\
-\mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \bm{\theta} \bm{\theta}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \ \mathbf{Q}^{-1} \bm{\theta} \bm{\theta}^\top \mathbf{Q}^{-1} , \end{aligned} \end{equation}
+\begin{equation} \begin{aligned} \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\} & \approx \bm{\theta}  + \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) = \bm{\theta}  +  \sigma_{\mathbf{E}}^2 \mathbf{Q}^{-1} \left( 2\mathbf{I} + 2n\mathbf{I} - \color{blue}N\color{black} \mathbf{I} \right) \bm{\theta} , \quad \text{and} \\
+\mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \bm{\theta} \bm{\theta}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - \color{blue}N\color{black} \right)^2 \ \mathbf{Q}^{-1} \bm{\theta} \bm{\theta}^\top \mathbf{Q}^{-1} , \end{aligned} \end{equation}
 from where equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) are obtained.
@@
 & \quad \ \ \ - \mathbf{Q}^{-1} \mathbf{K}^\top \underbrace{\mathbb{E} \left\{ \mathbf{E} \mathbf{Q}^{-1} \mathbf{K}^\top \bm{\epsilon} \right\}}_{\mathbf{B}_3} + \mathbf{Q}^{-1} \underbrace{\mathbb{E} \left\{ \mathbf{E}^\top \left( \mathbf{I} - \mathbf{K} \mathbf{Q}^{-1} \mathbf{K}^\top \right) \bm{\epsilon} \right\}}_{\mathbf{B}_4} , \quad \text{and} \\
-\mathbf{C} \left( \widehat{\bm{\theta}} \right)  & \approx \mathbf{Q}^{-1} \mathbf{K}^\top \left( \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \bm{\epsilon}^\top \right\}}_{\sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n}}  + \underbrace{\mathbb{E} \left\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_1}  - \underbrace{\mathbb{E} \left\{  \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \right\}}_{\mathbf{C}_2} - \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \bm{\theta}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_2^\top} \right) \mathbf{K} \mathbf{Q}^{-1} \\
+\mathbf{C} \left( \widehat{\bm{\theta}} \right)  & \approx \mathbf{Q}^{-1} \mathbf{K}^\top \left( \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \bm{\epsilon}^\top \right\}}_{\sigma_{\bm{\epsilon}}^2 \mathbf{I}_{\color{blue}N\color{black}-n}}  + \underbrace{\mathbb{E} \left\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_1}  - \underbrace{\mathbb{E} \left\{  \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \right\}}_{\mathbf{C}_2} - \underbrace{\mathbb{E} \left\{ \bm{\epsilon} \bm{\theta}^\top \mathbf{E}^\top \right\}}_{\mathbf{C}_2^\top} \right) \mathbf{K} \mathbf{Q}^{-1} \\
 & \quad - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right). \end{aligned} \end{equation} 
@@
 \begin{equation} z_{ij} = \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \mathbf{E} \right\}_{ij} = \mathbb{E} \left\{ \mathbf{e}_i \mathbf{A} \mathbf{E}_j \right\} = \mathrm{tr} \left( \mathbf{A} \ \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} \right) , \end{equation}
-where $\mathbf{e}_i$, and $\mathbf{E}_j$ are the $i\text{-th}$ row, and the $j\text{-th}$ column of $\mathbf{E}$, for $i = 1, \ldots, T-n$, and $j = 2, \ldots, n+1$.
+where $\mathbf{e}_i$, and $\mathbf{E}_j$ are the $i\text{-th}$ row, and the $j\text{-th}$ column of $\mathbf{E}$, for $i = 1, \ldots, \color{blue}N\color{black}-n$, and $j = 2, \ldots, n+1$.
 The matrix $\mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\}$ is the product of $\sigma_{\bm{\epsilon}}^2$ times a matrix whose elements are 0 in the first column, 2 in the $(j-i-1) \text{-th}$ diagonal, and -1 in the $(j-i-2) \text{-th}$, and $(j-i) \text{-th}$ diagonals, with zeros elsewhere.
 By using the definition of the second differential operator, we express
-\begin{equation} \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} = \sigma_{\bm{\epsilon}}^2 \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{2, j-i} \end{bmatrix}. \end{equation}
+\begin{equation} \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} = \sigma_{\bm{\epsilon}}^2 \begin{bmatrix} \mathbf{0}_{\color{blue}N\color{black}-n} & \mathbf{D}_{\color{blue}N\color{black}-n \times n}^{2, j-i} \end{bmatrix}.
+\end{equation}
 
@@
 Therefore we have
-\begin{equation} \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} = \sigma_{\bm{\epsilon}}^2 \mathbf{D}_{T-n \times T-n}^{2,j-i+1}. \end{equation}
+\begin{equation} \mathbb{E} \left\{ \mathbf{E}_j \mathbf{e}_i \right\} = \sigma_{\bm{\epsilon}}^2 \mathbf{D}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n}^{2,j-i+1}.
+\end{equation}
 
@@
 The vector $\bm{\epsilon}$ in the expected values $\mathbb{E} \left\{ \mathbf{E} \mathbf{A} \epsilon \right\}$, $\mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \epsilon \right\}$, and $\mathbb{E} \left\{ \mathbf{E} \mathbf{A} \bm{\epsilon}^\top \right\}$ is
-\begin{equation} \bm{\epsilon} = \begin{bmatrix} \epsilon(n+1) & \epsilon(n+2) & \cdots & \epsilon(T) \end{bmatrix}^\top, \end{equation}
+\begin{equation} \bm{\epsilon} = \begin{bmatrix} \epsilon(n+1) & \epsilon(n+2) & \cdots & \epsilon(\color{blue}N\color{black}) \end{bmatrix}^\top, \end{equation}
 as it is imposed by the input estimation method formulation.

### 2020-08-21 ChapterExperimentalValidation/ChapterExperimentalValidation.tex
Final version before internal defense
Second draft before public defense

@@
 \emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I.
-Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Experimental validation of a data-driven step input estimation method for dynamic measurements", IEEE Transactions on Instrumentation and Measurement Journal, 2019, doi: 10.1109/TIM.2019.2951865.
+Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Experimental validation of a data-driven step input estimation method for dynamic measurements", IEEE Transactions on Instrumentation and Measurement Journal, \color{blue} vol.
+69, no.
+7, pp.
+4843-4851, July 2020, \color{black} doi: 10.1109/TIM.2019.2951865.
 \nocite{QuintanaTIM} }\vfill{}
@@
 
-A measurement is a dynamic process. 
-The sensor is a dynamic system that experiments changes after an input excitation.
-The sensor output is first a transient state response and followed by a steady state response.
-
-The input excitation influences the sensor output, and thus it makes sense to estimate the input using the sensor response.
-When the sensor reaches its steady-state, the input is a straightforwardly estimated from the sensor output making use of the sensor static gain.
-The sensor transient response must be considered, instead of the steady state response, to obtain a fast input estimation.
+A measurement is a dynamic process \color{blue} and a sensor is a dynamic system that is excited by the unknown and to-be-measured input.
+The input application drives the sensor into a transient regime response.
+The input excitation influences the sensor output, and the input has to be estimated using the sensor transient response.
+If the sensor reached its steady-state, the input would be straightforwardly estimated from the sensor output making use of the sensor static gain.
+As is discussed in \citet{Dienstfrey14}, the lower is the sensor bandwidth, the higher is the need for methods based on processing the sensor transient response to obtain fast input estimation.
 
 A common approach is to filter the sensor transient response with another dynamic system that inverts the sensor dynamics, aiming to compensate the sensor transient time.
-It is necessary to have a sensor model to design a compensator system. 
-A different approach uses digital signal processing methods that are independent of the sensor model.
-One of these methods is the data-driven method that estimates the unknown level of step inputs by processing the sensor step response. 
-To assess the uncertainty of the data-driven step input estimation method, it processed step responses in a real-life metrology application. 
-This chapter describes the measurement experiments with a weighing sensor system and discusses the results observed.
-The results are put in perspective to simulations and consider the measurement noise that is additive to the response of the sensor.
+Examples of this compensation approach are the FIR filter described in \citet{Elster07}, the time-varying continuous-time filter conceived in \citet{Piskorowski08}, and the IIR filters presented in \citet{Link09} and \citet{Eichstadt10}.
+The main drawback of these approaches is that the compensator design is based on a sensor model.
+The sensor model requirement implies that the uncertainty on the parameters contributes to the uncertainty budget, as it is explained in \citet{Hessling11}.
+The uncertainty propagation through the compensation systems has been studied in \citet{Elster07, Link07} for the estimated model parameters, in \citet{Eichstadt16b} for a Kalman filter estimator, and in \citet{DEmilia16} considering the models of the different elements in the measurement system\color{black}.
 
-The experimental results validate the data-driven step input estimation method and indicate in which intervals of SNR and sample size the method is reliable.
+An approach different to compensation uses digital signal processing methods that are independent of the sensor model.
+\color{blue} This is especially relevant when the measurement system is inaccessible as the temperature sensor on board of a spatial probe described in \citet{Saggin01}, or when the formulation of the estimation method bypasses the model identification and estimates directly the input as in the method introduced in \citet{Markovsky15cep} that estimates the unknown level of step inputs.
+This data-driven method step input estimation method did not have a metric to describe its uncertainty, but in the previous chapter of this work a methodology is described to obtain the bias and the variance of the step input estimation.
+The methodology yields expressions for the bias and variance, and the expressions are validated via Monte Carlo simulation as is suggested in the work of \citet{Cox06}. 
+The estimation uncertainty is then described with the estimation first two statistical moments.
+This chapter describes the experiments performed to assess the uncertainty of the data-driven step input estimation method under real life data. 
+The estimation method processed step responses from a  weighing sensor experimental setup. 
+The experimental results are put in perspective to the simulations and validate the data-driven step input estimation method, showing which intervals of SNR and sample size ensure the reliability of the method.
+
+The weighing system is constructed using a load cell as the sensing element.
+The load cell is a popular and versatile device that has been used in weighing systems in the works of \citet{Piskorowski08}, \citet{Boschetti13}, \citet{Kesilmis16}, and \citet{Guo18}.
+Other magnitudes can also be measured with a load cell, examples of these are found in the works of \citet{Rossander15} that measure forces in the blades of a wind turbine, \citet{Hernandez06} that improve the quality of confort of bus seats measuring the forces they are subjet to, \citet{Alaziz17} and \citet{Zahradka18} that monitor the quality of sleep by sensing the body motion on the bed, \citet{Ballo16} that measure accelerations on a dummy in frontal impact tests.
+
+\color{black}
 
@@
 The convolution of the compensator impulse response with the sensor transient response yields the input estimate.
-Therefore, the design of a compensator is based on the sensor model and requires a deconvolution \citep{Eichstadt10}.
-Examples of input estimation using compensation of the sensor transient response include a recursive estimation of the compensator parameters \citep{Shu93}, 
-finite impulse response (FIR) \citep{Elster07, Niedzwiecki16b} filters and 
-infinite impulse response (IIR) filters \citep{Pintelon90, Elster08}.
+Therefore, the design of a compensator is based on the sensor model and requires a deconvolution \citet{Eichstadt10}.
+Examples of input estimation using compensation of the sensor transient response include a recursive estimation of the compensator parameters \citet{Shu93}, 
+finite impulse response (FIR) \citet{Elster07, Niedzwiecki16b} filters and 
+infinite impulse response (IIR) filters \citet{Pintelon90, Elster08}.
 The filters in these works estimate in real-time the unknown input value.
@@
 An alternative to the compensation approach is to use digital signal processing methods that are independent of the sensor model.
-A data-driven method that estimates the unknown level of step inputs by processing the sensor step response was introduced in \citep{Markovsky15ieee}. 
+A data-driven method that estimates the unknown level of step inputs by processing the sensor step response was introduced in \citet{Markovsky15ieee}. 
 This data-driven input estimation method avoids the sensor modeling stage and estimates directly the input.
 This method reduces the estimation time compared to a conventional compensator.
-The step input estimation method performance was demonstrated by simulations and experiments on a digital signal processor (DSP) of low cost \citep{Markovsky15cep}.
+The step input estimation method performance was demonstrated by simulations and experiments on a digital signal processor (DSP) of low cost \citet{Markovsky15cep}.
 The uncertainty of the step input estimation method has not been assessed before.
 
-To validate the input estimation methods it is necessary to assess the uncertainty associated with their estimates \citep{daSilva12, Ferrero06}.
-There are uncertainty propagation studies for model-based compensators such as the FIR and IIR filters for acceleration measurements where the uncertainty is computed in real time \citep{Elster07, Elster08, Link09}.
+To validate the input estimation methods it is necessary to assess the uncertainty associated with their estimates \citet{daSilva12, Ferrero06}.
+There are uncertainty propagation studies for model-based compensators such as the FIR and IIR filters for acceleration measurements where the uncertainty is computed in real time \citet{Elster07, Elster08, Link09}.
 In these works, the uncertainty expression is based on the transfer function or state space representations of the LTI sensor and filter systems.
-Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \citep{Pietrzak14} for mass and in \citep{Ogorevc16} for temperature sensors.
-A deconvolution method is implemented to estimate the input waveform in \citep{Hale09} and the uncertainty is obtained from the input estimate covariance. 
-The impact that the signal processing data-driven dynamic error correction has on the uncertainty is investigated in \citep{Saggin01}. 
-A statistical analysis of the data-driven step input estimation method \citep{Markovsky15cep} was investigated in \citep{Quintana19} and the method uncertainty was obtained with a Monte Carlo simulation study.
+Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \citet{Pietrzak14} for mass and in \citet{Ogorevc16} for temperature sensors.
+A deconvolution method is implemented to estimate the input waveform in \citet{Hale09} and the uncertainty is obtained from the input estimate covariance. 
+The impact that the signal processing data-driven dynamic error correction has on the uncertainty is investigated in \citet{Saggin01}. 
+A statistical analysis of the data-driven step input estimation method \citet{Markovsky15cep} was investigated in \citet{Quintana19} and the method uncertainty was obtained with a Monte Carlo simulation study.
 
@@
 The novelty of the paper is threefold.
-First, using the results of \citep{Quintana19}, that describes a statistical analysis of structured errors in variables (EIV) problems, in this paper we describe the statistical properties of a the data-driven step input estimation method in both simulation and a real-life experiments.
+First, using the results of \citet{Quintana19}, that describes a statistical analysis of structured errors in variables (EIV) problems, in this paper we describe the statistical properties of a the data-driven step input estimation method in both simulation and a real-life experiments.
 Second, this manuscript also presents the Cram\'er-Rao lower bound for unbiased estimators of the structured and correlated EIV problem that the step input estimation method formulates.
@@
 The MC simulation performed $N_{MC} = 10^4$ runs of the data-driven step input estimation with different realizations of the measurement noise $\bm{\epsilon}$.
-The MC simulation was conducted processing $T = 5000$ samples of a simulated transient step response $\widehat{\mathbf{y}}$ generated by a stable linear time-invariant (LTI) system of order $n = 5$, with a sampling frequency of $f_s=4$ kHz.
+The MC simulation was conducted processing $\color{blue}N\color{black} = 5000$ samples of a simulated transient step response $\widehat{\mathbf{y}}$ generated by a stable linear time-invariant (LTI) system of order $n = 5$, with a sampling frequency of $f_s=4$ kHz.
 This system is a state-space model obtained with the System Identification Toolbox using the measured step response of the actual sensor described in the Practical Implementation Section.
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_1.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_1.pdf} 
 \caption{ \label{fig:uh_sim} 
@@
 We are interested in the first element of $\widehat{\bm{\theta}}$, which is the input estimate $\widehat{u}$.
-The measurement noise variance was selected to have a signal-to-noise ratio (SNR) in the interval [30 dB, 80 dB], according to
-\begin{equation} \mathrm{SNR} = 20 \log_{10}{ \dfrac{ \sqrt{ \dfrac{1}{T} \int\limits_{0}^{T}{ y(\tau)^2  \mathrm{d} \tau } } }{ \sigma_{ \epsilon }} } \end{equation}
+The measurement noise variance was selected to have a signal-to-noise ratio (SNR) in the interval [30 dB, 80 dB], according to \color{blue}(\ref{eqn:SNR})\color{black}.
 
@@
 In each of the $N_{MC}$ runs, we compute the predictions of the step input estimation bias and variance from measured data using Equations (\ref{eqn:biasST}) and (\ref{eqn:varST}). 
-The step input bias and variance predictions from observed data $b_{\widetilde{\mathrm{p}}}$ and $v_{\widetilde{\mathrm{p}}}$, and the associated standard error $\sigma_{\widetilde{\mathrm{p}}}$, are obtained from
-\begin{equation} \begin{aligned} & b_{\widetilde{\mathrm{p}}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathbf{b}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1\right]} }, \ v_{\widetilde{\mathrm{p}}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1,1\right]} },  \\ & \mathrm{and} \quad \sigma_{\widetilde{\mathrm{p}}} = \sqrt{   \frac{ \sum_{i=1}^{N_{MC}}{ \left( \mathbf{b}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1\right]} - b_{\widetilde{\mathrm{p}}} \right)^2 } }{N_{MC}\left( N_{MC}-1 \right)}  } , \end{aligned} \end{equation}
+The step input bias and variance predictions from observed data $b_{\widetilde{\mathrm{p}}}$ and $v_{\widetilde{\mathrm{p}}}$, and the associated standard error $\sigma_{\widetilde{\mathrm{p}}\mathrm{,err}}$, are obtained from
+\begin{equation} \begin{aligned} & b_{\widetilde{\mathrm{p}}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathbf{b}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1\right]} }, \ v_{\widetilde{\mathrm{p}}} = \frac{1}{N_{MC}} \sum_{i=1}^{N_{MC}}{ \mathrm{\mathbf{C}}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1,1\right]} },  \\ & \mathrm{and} \quad \sigma_{\widetilde{\mathrm{p}}\mathrm{,err}} = \sqrt{   \frac{ \sum_{i=1}^{N_{MC}}{ \left( \mathbf{b}_{\widetilde{\mathrm{p}}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1\right]} - b_{\widetilde{\mathrm{p}}} \right)^2 } }{N_{MC}\left( N_{MC}-1 \right)}  } , \end{aligned} \end{equation}
 where $\widetilde{\mathbf{b}}_{\mathrm{p}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1\right]}$ is the first element in the bias vector and $\widetilde{\mathbf{C}}_{\mathrm{p}}^i \left( \widehat{\bm{\theta}} \right) \big|_{\left[1,1\right]}$ is the first element in the covariance matrix obtained in the $i-\mathrm{th}$ approximations.
@@
 Figure \ref{bias_sigma_sim_unstr_str_n2} shows the empirical bias, the bias predictions and the standard errors of the MC simulation.
-It can be seen that the empirical bias $b_\mathrm{e}$ and the predicted bias $b_{\widetilde{\mathrm{p}}}$ are proportional to the perturbation noise variance while the standard errors $\sigma_\mathrm{e}$ and $\sigma_{\widetilde{\mathrm{p}}}$ are proportional to the perturbation noise standard deviation.
+It can be seen that the empirical bias $b_\mathrm{e}$ and the predicted bias $b_{\widetilde{\mathrm{p}}}$ are proportional to the perturbation noise variance while the standard errors $\sigma_\mathrm{e,err}$ and $\sigma_{\widetilde{\mathrm{p}}\mathrm{,err}}$ are proportional to the perturbation noise standard deviation.
 For SNR below 40 dB there is a difference of a small order of magnitude between the empirical bias $b_\mathrm{e}$ and the bias prediction $b_{\widetilde{\mathrm{p}}}$.
 
-The standard errors of the MC simulation $\sigma_\mathrm{e}$ and $\sigma_{\widetilde{\mathrm{p}}}$ are smaller than  $b_\mathrm{e}$ and $b_{\widetilde{\mathrm{p}}}$.
+The standard errors of the MC simulation $\sigma_\mathrm{e,err}$ and $\sigma_{\widetilde{\mathrm{p}}\mathrm{,err}}$ are smaller than  $b_\mathrm{e}$ and $b_{\widetilde{\mathrm{p}}}$.
 The estimates are spread near the sample mean and the uncertainty is smaller than the bias. 
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_2.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_2.pdf} 
 \caption{ \label{bias_sigma_sim_unstr_str_n2} 
-The results of the Monte Carlo simulation of the step input estimation method are the empirical bias $b_{\mathrm{e}}$, the predicted bias using exact data $b_{\mathrm{p}}$, the predicted bias using measured data $b_{\widetilde{\mathrm{p}}}$, the empirical standard error $\sigma_{\mathrm{e}}$, and the predicted standard error $\sigma_{\widetilde{\mathrm{p}}}$.
+The results of the Monte Carlo simulation of the step input estimation method are the empirical bias $b_{\mathrm{e}}$, the predicted bias using exact data $b_{\mathrm{p}}$, the predicted bias using measured data $b_{\widetilde{\mathrm{p}}}$, the empirical standard error $\sigma_{\mathrm{e,err}}$, and the predicted standard error $\sigma_{\widetilde{\mathrm{p}}\mathrm{,err}}$.
 The estimation biases are proportional to the perturbation variance and the estimation standard errors are proportional to the perturbation standard deviation.
@@
 \begin{equation} \mathrm{MSE} = b^2 + v,   \end{equation}
-where $b$ and $v$ are the bias and the variance of the step input estimate, can be applied to the obtained empirical and predicted results and can be compared to the Cram\'er-Rao lower bound (CRB).
-Figure \ref{fig:MSE_CRB} shows that $\mathrm{MSE}_\mathrm{e} = b_\mathrm{e}^2 + v_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}  = b_{\widetilde{\mathrm{p}}}^2 + v_{\widetilde{\mathrm{p}}}$ have the same proportionality with respect to the measurement noise variance as the bound for an unbiased estimator $\mathrm{CRB_{ub}}$.
+where $b$ and $v$ are the bias and the variance of the step input estimate, can be applied to the obtained empirical and predicted results and can be compared to the Cram\'er-Rao lower bound (CRLB).
+Figure \ref{fig:MSE_CRLB} shows that $\mathrm{MSE}_\mathrm{e} = b_\mathrm{e}^2 + v_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}  = b_{\widetilde{\mathrm{p}}}^2 + v_{\widetilde{\mathrm{p}}}$ have the same proportionality with respect to the measurement noise variance as the bound for an unbiased estimator $\mathrm{CRLB_{ub}}$.
 For SNR above 35 dB, $\mathrm{MSE}_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ are equivalent, and below 35 dB the difference between them is of less than a factor of 10.
 
-We obtained an approximation of the $\mathrm{CRB}_{\mathrm{b}}$ for our biased estimator using the partial derivative of the bias in expression (\ref{eqn:biasST}).
-Figure \label{fig:MSE_CRB} shows that the bounds for the unbiased and biased estimators are almost equal because the partial derivatives of the bias are negligible w.r.t.
-1 in Equation (\ref{eqn:CRB_EIV}).
-By adding the square of the predicted bias to the biased estimator bound $\mathrm{CRB_{b}}$ we obtain an approximation of the minimum MSE that the biased estimator can achieve.
-This minimum MSE is close to the CRBs for large SNR but the square of the bias causes an increase of the MSE around 35 dB.
-The differences between the CRBs and $\mathrm{MSE}_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ are of one order of magnitude for large SNR and become small for SNR lower than 40 dB.
+We obtained an approximation of the $\mathrm{CRLB}_{\mathrm{b}}$ for our biased estimator using the partial derivative of the bias in expression (\ref{eqn:biasST}).
+Figure \label{fig:MSE_CRLB} shows that the bounds for the unbiased and biased estimators are almost equal because the partial derivatives of the bias are negligible w.r.t.
+1 in Equation (\ref{eqn:CRLB_EIV}).
+By adding the square of the predicted bias to the biased estimator bound $\mathrm{CRLB_{b}}$ we obtain an approximation of the minimum MSE that the biased estimator can achieve.
+This minimum MSE is close to the CRLBs for large SNR but the square of the bias causes an increase of the MSE around 35 dB.
+The differences between the CRLBs and $\mathrm{MSE}_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ are of one order of magnitude for large SNR and become small for SNR lower than 40 dB.
 This difference is the cost of solving a structured EIV problem with a simple LS method.
-%Therefore, for large measurement noise variance, the step input estimation method produces results that are comparable to maximum-likelihood estimation, since $\mathrm{MSE}_{\widetilde{\mathrm{p}}} \approx \mathrm{CRB_{b}} + b_{\widetilde{\mathrm{p}}}^2$.
+%Therefore, for large measurement noise variance, the step input estimation method produces results that are comparable to maximum-likelihood estimation, since $\mathrm{MSE}_{\widetilde{\mathrm{p}}} \approx \mathrm{CRLB_{b}} + b_{\widetilde{\mathrm{p}}}^2$.
 
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_3.pdf}
-\caption{\label{fig:MSE_CRB} The observation instant is fixed at 500 samples.
+\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_3.pdf}
+\caption{\label{fig:MSE_CRLB} The observation instant is fixed at 500 samples.
 The bias and the MSEs decrease for large SNR. 
 The empirical $\mathrm{MSE_e}$ and the predicted $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ of the step input estimation are one order of magnitude larger than the Cram\'er-Rao lower bound.
-Adding the CRB for a biased estimator with the predicted bias squared we have the minimum MSE, that grows in the interval [25 dB, 45 dB].
-Below 35 dB the difference between $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ and the CRB is less than a factor of 10.}
+Adding the CRLB for a biased estimator with the predicted bias squared we have the minimum MSE, that grows in the interval [25 dB, 45 dB].
+Below 35 dB the difference between $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ and the CRLB is less than a factor of 10.}
 \end{figure}
 
-There are two features of the system step response that make the CRB small.
+There are two features of the system step response that make the CRLB small.
 One is the measurement noise variance and the second is sample size.
@@
 \centering
-\includegraphics[width=\columnwidth]{./ChapterExperimentalValidation/fig/Fig_4.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_4.pdf} 
 \caption{ \label{fig:msee_vs_n_sim} 
@@
 An experimental setup was constructed to test the step input estimation method.
-The implementation is a weighing system that uses a load cell Tedea Huntleigh 1004 \citep{tedea1004}.
+The implementation is a weighing system that uses a load cell Tedea Huntleigh 1004, whose specifications are found in  \citet{tedea1004}.
 The maximum rating of the load cell is 600 g.
@@
 The step responses were collected and stored as datasets for further analysis.
+\color{blue}The datasets, and the code used to generate the results in this section \href{https://drive.google.com/drive/folders/1EfQb4nccg6x4kGcZropKRug3WkeNGX-m?usp=sharing}{\color{blue}are publicly available} in the sense of reproducible research research in the following url: \href{https://drive.google.com/drive/folders/1EfQb4nccg6x4kGcZropKRug3WkeNGX-m?usp=sharing}{https://drive.google.com/drive/folders/1EfQb4nccg6x4kGcZropKRug3WkeNGX-m?usp=sharing}\color{black}.
 The number of samples collected for each step response is $N=20000$.
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_6.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_6.pdf} 
 \caption{ \label{fig:uh_exp} 
@@
 
-The empirical bias $b_{\mathrm{e}}$ is the difference between the average of the 100 estimates $\widehat{u}$ and the mass calibration value $u = 138.32$ g, at each instant of time, and the standard error $\sigma_\mathrm{e}$ is the standard deviation of the mean estimate of the responses processed, i.e.,
-\begin{equation} \begin{aligned} \widehat{\mu}_{e} &= \frac{1}{100} \sum_{i=1}^{100}{ \widehat{u}_{i} }, \quad {b}_{\mathrm{e}} = \widehat{\mu}_{e} - u, \quad \mathrm{and} \\  \sigma_{\mathrm{e}} &= \frac{\widehat{\sigma}}{\sqrt{100}}, \quad \mathrm{where} \ \widehat{\sigma}^2 = \frac{1}{99} \sum_{i=1}^{100}{ \left( \widehat{u}_i - \widehat{\mu}_{e} \right)^2 } .
+The empirical bias $b_{\mathrm{e}}$ is the difference between the average of the 100 estimates $\widehat{u}$ and the mass calibration value $u = 138.32$ g, at each instant of time, and the standard error $\sigma_\mathrm{e,err}$ is the standard deviation of the mean estimate of the responses processed, i.e.,
+\begin{equation} \begin{aligned} \widehat{\mu}_{\mathrm{e}} &= \frac{1}{100} \sum_{i=1}^{100}{ \widehat{u}_{i} }, \quad {b}_{\mathrm{e}} = \widehat{\mu}_{\mathrm{e}} - u, \quad \mathrm{and} \\  \sigma_{\mathrm{e,err}} &= \frac{\widehat{\sigma_\mathrm{e}}}{\sqrt{100}}, \quad \mathrm{where} \ \widehat{\sigma_\mathrm{e}}^2 = \frac{1}{99} \sum_{i=1}^{100}{ \left( \widehat{u}_i - \widehat{\mu}_{\mathrm{e}} \right)^2 } .
 \end{aligned} \end{equation}
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_7.pdf}
+\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_7.pdf}
 \caption{ \label{fig:y_ss} 
@@
 
-Figure \ref{fig:b_sigma_exp_str} shows the empirical bias $b_{\mathrm{e}}$ and the standard error $\sigma_{\mathrm{e}}$ that result after processing the first $T=500$ samples of the 100 measured step responses $\widetilde{\mathbf{y}}$.
+Figure \ref{fig:b_sigma_exp_str} shows the empirical bias $b_{\mathrm{e}}$ and the standard error $\sigma_{\mathrm{e,err}}$ that result after processing the first $\color{blue}N\color{black}=500$ samples of the 100 measured step responses $\widetilde{\mathbf{y}}$.
 The standard error is smaller than the bias.
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_8.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_8.pdf} 
 \caption{ \label{fig:b_sigma_exp_str} 
-The results of estimating the step input level after processing 100 measured step responses are the empirical bias $b_{\mathrm{e}}$ and the empirical standard error $\sigma_{\mathrm{e}}$. 
+The results of estimating the step input level after processing 100 measured step responses are the empirical bias $b_{\mathrm{e}}$ and the empirical standard error $\sigma_{\mathrm{e,err}}$. 
 The estimation bias and the estimation standard error decrease as more samples are processed. 
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_9.pdf}
+\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_9.pdf}
 \caption{ \label{fig:YYhR_n7} 
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_10.pdf}
+\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_10.pdf}
 \caption{ \label{fig:meas_SNR} 
@@
 The 5 dB difference provides a conservative bound since the bias and variance computed from 50 dB of SNR are higher than those obtained using the variance estimation from the steady-state response. 
-Figure \ref{fig:b_sigma_exp_str2} shows a comparison of the results obtained with both measurement noise variance estimations after processing the first $T=500$ samples of the step response $\widetilde{\mathbf{y}}$.
+Figure \ref{fig:b_sigma_exp_str2} shows a comparison of the results obtained with both measurement noise variance estimations after processing the first $\color{blue}N\color{black}=500$ samples of the step response $\widetilde{\mathbf{y}}$.
 Using expression (\ref{eqn:biasST}), the bias prediction  $b_{\widetilde{\mathrm{p}}2}$ obtained using an SNR of 50 dB approximates more closely the empirical bias than $b_{\widetilde{\mathrm{p}}1}$ obtained using an SNR of 55dB.
-In accordance, the standard error of the bias predictions $\sigma_{\widetilde{\mathrm{p}}2}$ is larger than $\sigma_{\widetilde{\mathrm{p}}1}$ and is a conservative measure of the input estimation uncertainty.
+In accordance, the standard error of the bias predictions $\sigma_{\widetilde{\mathrm{p}}\mathrm{,err2}}$ is larger than $\sigma_{\widetilde{\mathrm{p}}\mathrm{,err1}}$ and is a conservative measure of the input estimation uncertainty.
 In conclusion, using the noise variance estimated from the residual prevents underestimating the step input estimation uncertainty.
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_11.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_11.pdf} 
 \caption{ \label{fig:b_sigma_exp_str2} 
 Comparative view of the bias prediction using two different noise variance estimations. 
-Estimating the variance from the step response residual gives a bias prediction $b_{\widetilde{\mathrm{p}}2}$ and a standard error $\sigma_{\widetilde{\mathrm{p}}2}$ that are slightly higher than using the noise variance estimated from the steady state response $\sigma_{\widetilde{\mathrm{p}}1}$ and $\sigma_{\widetilde{\mathrm{p}}1}$.
+Estimating the variance from the step response residual gives a bias prediction $b_{\widetilde{\mathrm{p}}2}$ and a standard error $\sigma_{\widetilde{\mathrm{p}}\mathrm{,err2}}$ that are slightly higher than $b_{\widetilde{\mathrm{p}}1}$ and $\sigma_{\widetilde{\mathrm{p}}\mathrm{,err1}}$ which correspond to computations that use the noise variance estimated from the steady state response.
 The bias prediction $b_{\widetilde{\mathrm{p}}2}$ approximates better the empirical bias. 
-The standard error $\sigma_{\widetilde{\mathrm{p}}2}$ provides a conservative value of the input estimation uncertainty. }
+The standard error $\sigma_{\widetilde{\mathrm{p}}\mathrm{,err2}}$ provides a conservative value of the input estimation uncertainty. }
 \end{figure}
@@
 The step input estimation method requires an assumption of the generating system order in the formulation of the estimation problem (\ref{eqn:min_seiv}).
-The estimation method performance is assessed under different assumptions of the values of $n$ in the interval from 2 to 10. 
+The estimation method performance is assessed under different assumptions of the values of $n$ in the interval from 2 to 100. 
 For each value of $n$, 100 step input estimations are computed from measured transient responses and the empirical MSEs are compared.
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{./ChapterExperimentalValidation/fig/Fig_12.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_12.pdf} 
 \caption{ \label{fig:msee_vs_n_exp} 

### 2020-08-21 ChapterIntroduction/ChapterIntroduction.tex
Second draft before public defense

@@
 The inputs interact with the sensor, and there are energy transfers between them that modify the sensor state.
-The sensor output response depends on the applied inputs and on the sensor initial conditions.
+The sensor output response depends on the applied inputs and on the sensor's initial conditions.
+
+\color{blue}
+
+Take for example the case of a thermometer.
+To measure temperature, the thermometer is put in contact with the substance or object of interest.
+According to the Newton's law of cooling, the temperature $y$ of the thermometer changes proportionally to its difference with the temperature $u$ of the matter under study:
+\begin{equation}  \dfrac{dy}{dt} = k_{\mathrm{T}} \left( u - y \right) \label{eqn:NewtonCooling} \end{equation}
+where the proportionality constant $k_{\mathrm{T}}$ is a parameter of the thermometer's first order model.
+The input $u$ is unknown and its estimation is the objective of the measurement.
+It is customary to take the reading from the thermometer once the temperature stops changing.
+
+Another popular example of a measurement is found in weighing. 
+In most of modern scales, the object of interest is placed on top of a weighing sensor.
+A simple model of the weighing sensor is a second order mass-spring-damper model shown in Figure \ref{fig:simple_msd_system}.
+In this model, the unknown mass $u(t)$, that can be time-dependent, causes a displacement $y$ of the scale and the dynamics of the sensor are described by the differential equation:
+\begin{equation} \dfrac{d}{dt} \left( \left(u(t)+m\right) \ \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = \left(u(t)+m\right)  g \end{equation}
+where $m$ is the own mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.
+
+For illustration purposes, Figures \ref{fig:thermometer} and \ref{fig:scale} show the responses of the thermometer and the weighing sensors, respectively, after the application of a constant input $u$, which can be represented by a step input.
+
+\begin{figure}[htb!]
+\centering
+
+\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
+\tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
+\tikzstyle{damper}=[thick,decoration={markings,  
+  mark connection node=dmp,
+  mark=at position 0.5 with 
+  {
+    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
+    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
+    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
+  }
+}, decorate]
+\tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]
+
+\node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
+\node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u(t)$};
+
+\node (ground1) at (M.south) [ground,yshift=-1.5cm,xshift=-0.625cm,anchor=north] {};
+\draw (ground1.north west) -- (ground1.north east);
+\draw [spring] (ground1.north) -- ($(M.south east)!(ground1.north)!(M.south west)$);
+
+\node (groundc) at (M.south) [ground,yshift=-1.5cm,anchor=north] {}; 
+\draw (groundc.north west) -- (groundc.north east);
+
+\node (ground2) at (M.south) [ground,yshift=-1.5cm,xshift=0.625cm,anchor=north] {};
+\draw (ground2.north west) -- (ground2.north east);
+\draw [damper] (ground2.north) -- ($(M.south east)!(ground2.north)!(M.south west)$);
+
+\node[draw=none,fill=none] at (-0.9cm,-1cm) {$k_{\mathrm{s}}$};
+\node[draw=none,fill=none] at (0.15cm,-1cm) {$k_{\mathrm{d}}$};
+\node[draw=none,fill=none] at (2.0cm,1.0cm) {$y$};
+\draw [-latex,thick]  ++(2.2cm,-1cm) -- +(0cm,2.25cm);
+
+\draw [-latex,thick] (M.east) ++(0,0) -- +(1cm,0);
+\draw [line width=0.25mm] (2.2cm,-1cm) -- (2.2cm,1cm);
+\draw [line width=0.25mm] (2.1cm,-1cm) -- (2.3cm,-1cm);
+\draw [line width=0.25mm] (2.1cm,1cm) -- (2.3cm,1cm);
+\draw [line width=0.25mm] (2.1cm,-0.5cm) -- (2.3cm,-0.5cm);
+\draw [line width=0.25mm] (2.1cm,0.5cm) -- (2.3cm,0.5cm);
+\draw [line width=0.25mm] (2.15cm,-0.25cm) -- (2.25cm,-0.25cm);
+\draw [line width=0.25mm] (2.15cm,0.25cm) -- (2.25cm,0.25cm);
+\draw [line width=0.25mm] (2.15cm,-0.75cm) -- (2.25cm,-0.75cm);
+\draw [line width=0.25mm] (2.15cm,0.75cm) -- (2.25cm,0.75cm);
+\draw [line width=0.25mm] (2.1cm,0cm) -- (2.3cm,0cm);
+
+\end{tikzpicture}
+
+\caption{\label{fig:simple_msd_system} \color{blue} A second order mass-spring-damper model represents the weighing sensor.
+The application of the mass $u$ causes a change in the position $y$ of the scale.
+\color{black} }
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_2.pdf} 
+\caption{ \label{fig:thermometer} 
+\color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_{\mathrm{T}}=0.5 s^{-1}$, caused by a step input of 67 $^{\circ} \mathrm{C}$  when the thermometer was initially at 23 $^{\circ} \mathrm{C}$.
+In this example, we should wait 10 s to have an estimation of the input with a relative error lower than 0.5 \%.
+\color{black} }
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_3.pdf} 
+\caption{ \label{fig:scale} 
+\color{blue} This simulation illustrates the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium.
+The sensor parameters are $k_{\mathrm{d}} = 5 \ \mathrm{N \ s/m}$, $m = 5 \ \mathrm{g}$, and $k_{\mathrm{s}} = 10250 \ \mathrm{N/m}$.
+In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement with a relative error smaller than 0.72 \%.
+\color{black}  }
+\end{figure}
+
+In \citet{Eichstadt12} an illustration on a force measurement sensor can be found. 
+The sensor is modeled as a cascade of two damped mass-spring systems. 
+The transfer function of the sensor is $H(s) = H_1(s) \ H_2(s)$, where each of the two transfer functions are described by
+\begin{equation} H_i(s) = \dfrac{\left( 2 \pi f_{0,i}\right)^2}{s^2  + 2 \delta_i \left( 2 \pi f_{0,i}\right) s  +  \left( 2 \pi f_{0,i}\right)^2} 
+\end{equation}  
+where $\delta_1 = 0.1$, $\delta_2 = 0.2$, $f_{0,1} = 1.9$ Hz, and $f_{0,2} = 7$ Hz are the system parameters.
+The response of this sensor to a unit step input, shown in Figure \ref{fig:force}, takes at least 5 s to in which the transient response is predominant.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_4.pdf} 
+\caption{ \label{fig:force} 
+\color{blue} The simulation of a force sensor based on a double mass-spring system in cascade also shows that the stabilization after a unit step input is not instantaneous.
+In the plot we infer that the user has to wait 5 s to estimate the force from the sensor response that is 0.25 \% far from its value in equilibrium.
+\color{black}  }
+\end{figure}
+
+The handbook of sensors written by \citet{Fraden16Book} describes a pressure sensor that detects sudden changes in the pressure inside a room.
+These changes can be caused by the opening of a door or by the movement of a person.
+The principle of operation of detecting minute pressure gradients is the flowmeter that compares the pressure at both outlets of a tube.   
+An experimental identification of such a flowmeter is described in \citet{Wiklund02}, where a second order model is proposed to explain the dynamics with the transfer function
+\begin{equation} H(s) = \dfrac{ e^{-\tau_{delay}s}}{\left( 1 +  \tau_1 s \right) \left( 1 +  \tau_2 s \right)} 
+\end{equation}  
+where $\tau_{delay} = 0.07$ s is the dead time, and $\tau_{1} = 0.01$ s and $\tau_{2} = 1.6$ s are two time constants.
+The step response simulation of this dynamic model results in the sensor step response shown in Figure \ref{fig:pressure}.  
+We have a response to a unit step function that is mostly and increasing exponential that requires 7 s approximately to reach a value that is 98.7 \% ot its maximum value.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_5.pdf} 
+\caption{ \label{fig:pressure} 
+\color{blue} The response of a flowmeter to a sudden change in its pressure gradient acrros the extremes of a tube is an overdamped transient response.
+Ideally we would expect the the transient response follows closely the shape of the unit step input applied, but instead an exponential response is observed, that last for more than 7 s.
+\color{black}  }
+\end{figure}
+
+\color{black}
 
@@
 % Context:Data-driven step input estimation method
-An example of a data-driven method devised for a DSP is the direct step input level estimation from the sensor step response \citep{Markovsky15cep}.
+An example of a data-driven method devised for a DSP is the direct step input level estimation from the sensor step response introduced in \citet{Markovsky15cep}.
 This method formulates a Hankel structured errors-in-variables (EIV) problem with correlation.
@@
 The main advantage of the data-driven step input estimation method is that it does not identify the sensor model, but instead, it directly estimates the input. 
-The direct estimation differs from the standard two-stage methodology, that first identifies the sensor model and later estimates the input.
+The direct estimation differs from the standard two-stage methodology that first identifies the sensor model and later estimates the input.
 In this approach, the output-error (OE) problem is converted into an EIV problem that is harder to solve, but the RLS solution is easy to compute.
@@
 The estimation uncertainty assessment is needed to validate the estimation methods for metrology applications.
-The uncertainty of the data-driven step input estimation method \citep{Markovsky15ieee} is unknown.
-The estimation bias and variance define the estimation uncertainty \citep{Pintelon12Book}, and they can be obtained by conducting an elementwise statistical analysis. 
+The uncertainty of the data-driven step input estimation method described in \citet{Markovsky15ieee} is unknown.
+The estimation bias and variance define the estimation uncertainty, as it is explained in \citet{Pintelon12Book}, and they can be obtained by conducting an elementwise statistical analysis. 
 The validation of the step input estimation method requires also to demonstrate its effectiveness on real-life measurements.
@@
 This subsection presents a list of relevant works published in relation with the thesis topic. 
-These works study theoretical aspects like the estimation of inherent dynamical errors \citep{Hessling06}, the dynamic correction in time domain \citep{Hessling08a}; and experimental aspects like the increment of the amount of sensors in production lines for quick decision making in autonomous control \citep{Esward09}.
+These works study theoretical aspects like the estimation of inherent dynamical errors presented in \citet{Hessling06}, the dynamic correction in time domain proposed in \citet{Hessling08a}; and experimental aspects like the increment of the amount of sensors in production lines for quick decision making in autonomous control described in \citet{Esward09}.
 
-To implement input estimation methods in the practice, the compensation filters based on deconvolution \citep{Eichstadt10} are the preferred methodology to develop 
-analog filters \citep{Jafaripanah05}, 
-adaptive digital filters \citep{Shu93}, 
-lattice adaptive filters \citep{Hernandez06}, 
-compensators for simultaneous responses of different sensors \citep{Boschetti13}, 
-regularized deconvolution compensators \citep{Dienstfrey14}, 
-and a multiple choice set of filters \citep{Huang16}. 
+To implement input estimation methods in the practice, the compensation filters based on deconvolution explained by \citet{Eichstadt10} are the preferred methodology to develop, for example, 
+analog filters in \citet{Jafaripanah05}, 
+adaptive digital filters in \citet{Shu93}, 
+lattice adaptive filters in \citet{Hernandez06}, 
+compensators for simultaneous responses of different sensors in \citet{Boschetti13}, 
+regularized deconvolution compensators in \citet{Dienstfrey14}, 
+and a multiple choice set of filters in \citet{Huang16}. 
 All these methods have in common the use of the measurement system model parameters to build the compensator.
@@
 A list of measurement methods based on digital signal processors (DSP) include 
-a study of the data-driven dynamic error correction and its impact on the measured temperature \citep{Saggin01},
-a modulation quality measurement of microwave access systems \citep{Angrisani10}, 
-a real-time rotational speed estimation using correlation \citep{Wang14},
-a development of a biology-inspired electronic nose \citep{Jing16},
-and impedance measurements for material damage estimation using cross-correlation \citep{deCastro19}.
-The data-driven step input estimation method \citep{Markovsky15cep} is added to this list, highlighting the fact that it can be implemented for the measurement of different physical magnitudes because it is model-independent.
+a data-driven dynamic error correction and its impact on the measured temperature studied in \citet{Saggin01},
+a modulation quality measurement of microwave access systems presented in \citet{Angrisani10}, 
+a real-time rotational speed estimation method using correlation introduced in \citet{Wang14},
+a biology-inspired electronic nose developed in \citet{Jing16},
+and impedance measurements for material damage estimation method using cross-correlation introduced in \citet{deCastro19}.
+The data-driven step input estimation method proposed by \citet{Markovsky15cep} is added to this list, highlighting the fact that it can be implemented for the measurement of different physical magnitudes because it is model-independent.
 
@@
 In metrology, a measurement is an estimation, represented as a random variable.
-The measurement noise always exists, and therefore, the input estimation is commonly expressed with its two first statistical moments \citep{Ferrero06}.
-The guideline \citep{GUM08}, accepted by the metrology community, recommends the standardization of the uncertainty assessment.
-The typical measurement uncertainty analysis are reviewed in \citep{daSilva12}.
-The Monte Carlo method is a uncertainty evaluation tool \citep{Cox06},
-that supports the use of simulation techniques for quantifying measurement uncertainties \citep{Esward16}.
-Another example of the Monte Carlo method application is a dynamic measurement uncertainty evaluation of clinical thermometers \citep{Ogorevc16}.
+The measurement noise always exists, and therefore, the input estimation is commonly expressed with its two first statistical moments as it is described in \citet{Ferrero06}.
+The guidelines listed in \citet{GUM08}, and accepted by the metrology community, recommend the standardization of the uncertainty assessment.
+The typical measurement uncertainty analysis \color{blue} is \color{black} reviewed in \citet{daSilva12}.
+The Monte Carlo method is an uncertainty evaluation tool according to \citet{Cox06},
+that supports the use of simulation techniques for quantifying measurement uncertainties as suggested by \citet{Esward16}.
+One example of the Monte Carlo method application is a dynamic measurement uncertainty evaluation of clinical thermometers described in \citet{Ogorevc16}.
 
-Nevertheless, researchers still pinpoint the need to study more the uncertainty assessment methods \citep{Esward09, Hessling10}.
-Some recommend to consider the uncertainties of all the measurement chain components \citep{Diniz17}, and to avoid the direct uncertainty propagation from the calibration towards the to-be-measured quantity.
+Nevertheless, researchers like \citet{Esward09}, and \citet{Hessling10} still pinpoint the need to study more the uncertainty assessment methods.
+\citet{Diniz17} recommends to consider the uncertainties of all the measurement chain components, and to avoid the direct uncertainty propagation from the calibration towards the to-be-measured quantity.
 Methods for evaluating the uncertainty associated with the output of compensation filters have been investigated for
-a discrete-time infinite-response filter in \citep{Link09},
-a discrete-time finite-response filter in \citep{Elster07, Elster08}, and
-the Kalman filter in \citep{Eichstadt16b}.
-All these works propagate the uncertainty through the filter, but it is also necessary to upward the propagation up to the sensor model to include all systematic error contributions \citep{Hessling11}.
+a discrete-time infinite-response filter in \citet{Link09},
+a discrete-time finite-response filter in \citet{Elster07, Elster08}, and
+the Kalman filter in \citet{Eichstadt16b}.
+All these works propagate the uncertainty through the filter, but it is also necessary to upward the propagation up to the sensor model to include all systematic error contributions, as it is explained by \citet{Hessling11}.
 
@@
 
-This thesis describes research work initiated after the data-driven step input estimation method \citep{Markovsky15cep}.
+This thesis describes research work initiated after the data-driven step input estimation method proposed by \citet{Markovsky15cep}.
 This method aims for metrology applications but it lacked an uncertainty assessment, and, therefore, its appropriateness was questionable.
@@
 The data-driven step input estimation method formulates a structured errors-in-variables (EIV) problem.
-In linear estimation EIV problems, the measurement noise perturbs the regression matrix and the regressor \citep{VanHuffel91Book}, \citep{Markovsky07overview}.
+In linear estimation EIV problems, the measurement noise perturbs the regression matrix and the regressor, as it is detailed in \citet{VanHuffel91Book} and \citet{Markovsky07overview}.
 The regression matrix, in structured EIV problems, has a structure that depends on the problem formulation.
-The Hankel and Toeplitz matrices appear in problems of metrology \citep{Markovsky15cep}, system identification \citep{Soderstrom07}, image restoration \citep{Feiz17}, nuclear magnetic resonance spectroscopy \citep{Cai16}, direction-of-arrival estimation \citep{Pan18}, and time-of-arrival estimation \citep{Jia18}.
+Hankel and Toeplitz matrices appear in several application problems such as those described by \citet{Markovsky15cep} in metrology, by \citet{Soderstrom07} in system identification, by \citet{Feiz17} in image restoration, by \citet{Cai16} in nuclear magnetic resonance spectroscopy, by \citet{Pan18} in direction-of-arrival estimation, and by \citet{Jia18} in time-of-arrival estimation.
 
-The data-driven step input estimation method directly estimates the input from the sensor transient response. 
-The perturbations in the EIV problem of interest come from the sensor output.
-This is the only observed signal \citep{Markovsky15cep}, and it constructs both the regression matrix and the regressor.
+The data-driven step input estimation method introduced by \citet{Markovsky15cep} directly estimates the input from the sensor transient response. 
+The perturbations in the formulated EIV problem come from the sensor output, since it is the only observed signal, and from this signal both the regression matrix and the regressor are built.
 The structure in the regression matrix is block-Hankel.
-The data-driven direct estimation methodology  estimates the input faster than the classical two-stage approach, which first identifies a sensor model and later estimates the input using the sensor model \citep{Azam15, Niedzwiecki16a}.
+The data-driven direct estimation methodology  estimates the input faster than the classical two-stage approach described in \citet{Azam15} and \citet{Niedzwiecki16a}, which first identifies a sensor model and later estimates the input using the sensor model.
 
 Instead of using a total least-squares (TLS) solution of the structured EIV problem, the step input estimation method uses the recursive least-squares solution (RLS). 
-The TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
-the TLS solution is equivalent to the maximum likelihood (ML) solution when the disturbances of the EIV problems are i.i.d. normally distributed \citep{Markovsky07overview}. 
-However, regarding structured EIV problems, the TLS estimator does not give general results since each structure requires a specific treatment \citep{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where finding the global optimum is not guaranteed \citep{Rhode14recursive}.
-Moreover, the computational complexity of TLS and ML inhibits their real-time implementation.
-The least-squares (LS) estimator is a suboptimal but simple solution to structured EIV problems that admits a recursive form (RLS) suitable for real-time implementation.
+In the review conducted by \citet{Markovsky07overview} it is shown that the TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
+the TLS solution is equivalent to the maximum likelihood (ML) solution when the disturbances of the EIV problems are i.i.d. normally distributed . 
+However, the studies of structured EIV problems presented in \citet{VanHuffel07TLSeditorial} demonstrate that the TLS estimator does not give general results since each structure requires a specific treatment.
+Moreover, \citet{Rhode14recursive} found that the ML estimator leads to non-convex optimization when applied to structured EIV problems, and the global optimum is not guaranteed.
+Also, the computational complexity of TLS and ML inhibits their real-time implementation.
+Thus, the least-squares (LS) estimator is a suboptimal but simple solution to structured EIV problems that admits a recursive form (RLS) suitable for real-time implementation.
 
@@
 In the literature we find 
-the design of a fast algorithm for matrices with small displacement rank \citep{Mastronardi07fast}, 
-the study of the LS estimator consistency \citep{Palanthandalam10parameter},
-the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models \citep{Kiviet12high}, \citep{Kiviet14improved}, and
-a discussion of the causes of bias and inconsistency in homogeneous estimators \citep{Yeredor04homogeneous}.
+the design of a fast algorithm for matrices with small displacement rank in \citet{Mastronardi07fast}, 
+the study of the LS estimator consistency in \citet{Palanthandalam10parameter},
+the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models in \citet{Kiviet12high}, \citet{Kiviet14improved}, and
+a discussion of the causes of bias and inconsistency in homogeneous estimators in \citet{Yeredor04homogeneous}.
 The literature does not address the uncertainty of LS estimators for structured EIV problems.
 
-The uncertainty of an estimation method is expressed using the bias and covariance of the estimate \citep{Pintelon12Book}.
-To know the uncertainty of the data-driven step input estimation method, I quantified the bias and covariance of the LS solution of EIV problems, for unstructured and structured cases. 
-The bias and covariance quantification extend the perturbation analysis that investigated in \citep{Stewart90SPT} and in \citep{Vaccaro94}.
+The uncertainty of an estimation method is expressed in \citet{Pintelon12Book} using the bias and covariance of the estimate.
+To know the uncertainty of the data-driven step input estimation method, we quantified the bias and covariance of the LS solution of EIV problems, for unstructured and structured cases. 
+The bias and covariance quantification extend the perturbation analysis that was investigated in \citet{Stewart90SPT} and in \citet{Vaccaro94}.
 The conducted analysis provides insight into the impact that the structure and the correlation have on the LS estimation uncertainty.
@@
 
-I derived expressions that quantify the bias and covariance by obtaining the mathematical expectation of the LS estimate approximated by the second-order Taylor series expansion.
-Using Monte Carlo simulations, I validated the accuracy of the bias and covariance expressions.
+We derived expressions that quantify the bias and covariance by obtaining the mathematical expectation of the LS estimate approximated by the second-order Taylor series expansion.
+Using Monte Carlo simulations, we validated the accuracy of the bias and covariance expressions.
 These expressions estimate the bias and covariance that the data-driven step input estimation method will exhibit for a given sample size and perturbation level.
-I compared the mean squared error of the LS estimate to the minimum variance specified by the Cram\'er-Rao lower bound of the structured EIV problem, to determine the conditions under which the data-driven step input estimation method is appropriate for practical applications \citep{QuintanaCSDA}.
+We compared in \citet{QuintanaCSDA} the mean squared error of the LS estimate to the minimum variance specified by the Cram\'er-Rao lower bound of the structured EIV problem, to determine the conditions under which the data-driven step input estimation method is appropriate for practical applications.
 
@@
 The validation of the data-driven step input estimation method in a practical application was necessary to demonstrate the usefulness of the method.
-The method performance was illustrated by simulations and temperature experiments on a digital signal processor (DSP) of low cost \citep{Markovsky15cep}.
+The method performance was studied in \citet{Markovsky15cep} using simulations and temperature experiments on a digital signal processor (DSP) of low cost.
 The method estimates the unknown level of step inputs by processing the sensor step response, and 
@@
 Other methods for input estimation mainly compensate the sensor transient response, for example, by 
-recursive estimation of the compensator parameters \citep{Shu93}, 
-finite impulse response (FIR) filtering \citep{Elster07}, \citep{Niedzwiecki16b} filters and 
-infinite impulse response (IIR) filtering \citep{Pintelon90}, \citep{Elster08}.
-The uncertainty propagation for these model-based compensators is based on the transfer function or state-space representations of the LTI sensor and filter systems \citep{Link09}, \citep{Hale09}.
-Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \citep{Pietrzak14} for mass and in \citep{Ogorevc16} for temperature sensors.
+recursive estimation of the compensator parameters in \citet{Shu93}, 
+finite impulse response (FIR) filtering in \citet{Elster07}, \citet{Niedzwiecki16b} and 
+infinite impulse response (IIR) filtering in \citet{Pintelon90}, \citet{Elster08}.
+The uncertainty propagation for these model-based compensators is based on the transfer function or state-space representations of the LTI sensor and filter systems in \citet{Link09} and \citet{Hale09}.
+Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \citet{Pietrzak14} for mass and in \citet{Ogorevc16} for temperature sensors.
 
-To validate the data-driven step input estimation method I built a weighing system setup based on a load cell sensor.
+To validate the data-driven step input estimation method we built a weighing system setup based on a load cell sensor.
 The weighing setup was constructed to ensure repeatability and reproducibility, along with different experimental realizations.
-Load cell sensors are versatile devices that are found in 
-heart and breathing physiological signal monitoring \citep{Lee16},
-clinical analysis of sleep quality \citep{Zahradka18},
-automobile safety studies \citep{Ballo16},
-wind turbine design \citep{Rossander15}, 
-civil engineering \citep{Olmi16}, and 
-sport bicycle design \citep{Casas16}, to name a few.
+Load cell sensors are versatile devices that are applied for
+heart and breathing physiological signal monitoring in \citet{Lee16},
+clinical analysis of sleep quality in \citet{Zahradka18},
+automobile safety studies in \citet{Ballo16},
+wind turbine design in \citet{Rossander15}, 
+civil engineering structure studies in \citet{Olmi16}, and 
+sport bicycle design in \citet{Casas16}, to name a few.
 
@@
 Weighing has been basic for the development of scientific and trade activites. 
-The load cell is now a standard transducer for weight determination and also for the improvement of measurement techniques, such as the geometric approach to processing of load cell responses \citep{Kesilmis16}, the design of new conveyor machinery \citep{Yamani18}, and electronic truck scales \citep{Guo18}.
+The load cell is now a standard transducer for weight determination and also for the improvement of measurement techniques, such as the geometric approach to processing of load cell responses \citet{Kesilmis16}, the design of new conveyor machinery \citet{Yamani18}, and electronic truck scales \citet{Guo18}.
 
-In safety studies, a six axis load cell is devised to quantify accelerations and impact forces exerted on a dummy \citep{Ballo16}.
-In alternative energy developments the load cells are useful to measure the forces on the arms of a vertical axis wind turbine \citep{Rossander15}.
-An academic study of the load that a structure withstands is conducted with strain gage load cells that confirms the numerical results and facilitates the design of complex shaped structures \citep{Olmi16}.
-In sports, the performance of new instrumented crank mechanisms is fostered by the utilization of load cells in the characterization, analysis and validation design stages \citep{Casas16}.
+In safety studies, a six axis load cell is devised to quantify accelerations and impact forces exerted on a dummy \citet{Ballo16}.
+In alternative energy developments the load cells are useful to measure the forces on the arms of a vertical axis wind turbine \citet{Rossander15}.
+An academic study of the load that a structure withstands is conducted with strain gage load cells that confirms the numerical results and facilitates the design of complex shaped structures \citet{Olmi16}.
+In sports, the performance of new instrumented crank mechanisms is fostered by the utilization of load cells in the characterization, analysis and validation design stages \citet{Casas16}.
 
-The versatility of the load cells permits the physiological signal monitoring of the the heart and breathing rates \citep{Lee16}, clinical analysis of sleep quality \citep{Zahradka18} and the classification of the movement intensity of people while they are sleeping \citep{Alaziz17}.
+The versatility of the load cells permits the physiological signal monitoring of the the heart and breathing rates \citet{Lee16}, clinical analysis of sleep quality \citet{Zahradka18} and the classification of the movement intensity of people while they are sleeping \citet{Alaziz17}.
 All of these experimental studies have load cells installed on bed setups.
@@
 An extension of the methodology proposed for the data-driven step input estimation was formulated to estimate the parameters of an affine input that changes at a constant rate,  by signal processing of the sensor transient response.
-This type of ramp inputs is observed in the measurement of mass during the transportation of products on conveyor systems, ranging from few grams \citep{Burmen09} to almost hundreds of kilograms \citep{Tasaki07}.
+This type of ramp inputs is observed in the measurement of mass during the transportation of products on conveyor systems, ranging from few grams \citet{Burmen09} to almost hundreds of kilograms \citet{Tasaki07}.
 The data-driven affine input estimation method is proposed as an alternative to existing compensation filters, such as 
-the time-variant low-pass filters introduced in \citep{Piskorowski08, Pietrzak14}, and
-the combination of filters in cascade proposed in \citep{Niedzwiecki16a}.
+the time-variant low-pass filters introduced in \citet{Piskorowski08, Pietrzak14}, and
+the combination of filters in cascade proposed in \citet{Niedzwiecki16a}.
 \end{comment}
@@
 One of the assumptions of the step input estimation method formulation is that the measurement noise is Gaussian and white. 
-Nevertheless, the estimation results show that the method is still able to provide useful input estimations \citep{QuintanaTIM}.
+Nevertheless, the estimation results discussed in \citet{QuintanaTIM} show that the method is still able to provide useful input estimations.
 
-The empirical bias and covariance obtained after repeating the weighing experiment was compared to the bias and covariance estimations obtained in the previous research work \citep{QuintanaCSDA}, and to the minimal variance given by the  Cram\'er-Rao lower bound (CRLB) of the structured EIV problem.
-I found that the mean squared error (MSE) of the step input estimate is near the CRLB, and the distance between the MSE and the CRLB provides a confidence interval for the input estimate with respect to the level of the measurement noise.
+The empirical bias and covariance obtained after repeating the weighing experiment was compared to the bias and covariance estimations obtained in our previous research work (\citet{QuintanaCSDA}), and to the minimal variance given by the  Cram\'er-Rao lower bound (CRLB) of the structured EIV problem.
+We found that the mean squared error (MSE) of the step input estimate is near the CRLB, and the distance between the MSE and the CRLB provides a confidence interval for the input estimate with respect to the level of the measurement noise.
 
@@
 A dynamic measurement is present when the fluctuations of the measurand impact on the input estimation, such as when a low-bandwidth sensor is excited with a fast changing input.
-The detection of input characteristics is of interest in several scientific and industrial applications for temperature \citep{Saggin01}, pressure \citep{Matthews14}, acceleration \citep{Link07}, force \citep{Vlajic16}, \citep{Hessling08a}, and mass \citep{Shu93}, \citep{Boschetti13}.
+The detection of input characteristics is of interest in several scientific and industrial applications, such as those described for measuring temperature in \citet{Saggin01}, pressure in \citet{Matthews14}, acceleration in \citet{Link07}, force in \citet{Vlajic16}, \citet{Hessling08a}, and mass in \citet{Shu93}, \citet{Boschetti13}.
 
-I developed a method to estimate the parameters of inputs that vary at a constant rate, influenced by the data-driven signal processing method that estimates the step level value using subspace techniques \citep{Markovsky15cep}, \citep{Markovsky15ieee}.
+We developed a method to estimate the parameters of inputs that vary at a constant rate, influenced by the data-driven signal processing method that estimates the step level value using subspace techniques introduced in \citet{Markovsky15cep}, \citet{Markovsky15ieee}.
 An input varies at a constant rate in applications where the magnitude of interest activates the sensor gradually. 
 An example of this affine activation is the measurement of mass while the to-be-weighted object is transported by a conveyor belt, and the profile of the input is a saturated ramp.
-Current solutions to the weighing in motion are low pass filters that estimate the mass using the saturated part of the input \citep{Tasaki07, Pietrzak14}.
+Current solutions described in \citet{Tasaki07}, \citet{Pietrzak14} use low pass filters to estimate the mass in motion based on the saturated part of the ramp input.
 The signal processing affine input estimation methods are motivated by the need to obtain the mass of the object from the ramp before it reaches saturation.
 
-In the proposed data-driven method that estimates the affine input \citep{QuintanaMEAS},
-the affine input is parameterized as a straight line model where the slope and the intercept are the parameters of interest.
+We propose in \citet{QuintanaMEAS} a data-driven method that estimates the affine input,
+which is parameterized as a straight line model where the slope and the intercept are the parameters of interest.
 The data-driven affine input estimation method formulates a structured errors-invariables (EIV) problem, similar to the one formulated for the step input estimation.
@@
 The ML method simulates the response of a sensor model to an affine input, and minimizes a cost function that is the sum of the squared differences between the actual and the simulated sensor responses.
-The ML method resembles the model predictive control approach in the sense that a cost function is minimized iteratively to optimize the parameters of a sensor model using the observed sensor response in a receding time horizon \citep{Mayne14}.
+The ML method resembles the model predictive control approach, explained in \citet{Mayne14}, in the sense that a cost function is minimized iteratively to optimize the parameters of a sensor model using the observed sensor response in a receding time horizon.
 The difference is that the ML method aims to estimate the unknown value of the affine input parameters instead of identifying a model and controlling the dynamic system.

### 2020-08-21 ChapterPreliminaries/ChapterPreliminaries.tex
Second draft before public defense

@@
 
+\color{blue}
+Under certain conditions on the input, a discrete-time representation of the continuous-time dynamics $H(s)$ is exact without any approximation.
+It is demonstrated in \citet{Ljung87Book} and in \citet{MiddletonGoodwinBook} that when the input is piecewise constant, then the step-invariant transformation
+\begin{equation} \mathbf{H}(z) = \left( 1- z^{-1} \right) \mathcal{Z} \left\{ \mathcal{L}^{-1} \left\{ \dfrac{\mathbf{H}(s)}{s} \right\} \right\}   \end{equation}
+is a zero-order hold sampling of the continuous time step response that leads to an exact representation of the system dynamics in discrete time with the transfer function $\mathbf{H}(z)$.
+It worths stating this fact here since the method studied in this thesis estimates the true value of a step input by processing samples of a sensor step response. 
+\color{black}
+
 The methodology presented in this work is explained with SISO LTI systems, but can also be generalized to systems with multiple inputs and outputs.
-An example of SISO LTI sensor is the temperature sensor, but there are sensors that have single input and multiple outputs, like the gas sensors described in \citep{Munther19}, and sensors that have multiple inputs and multiple outputs, like the three-axis accelerometers \citep{DEmilia16}, the radio-frequency intruder-detection sensors \citep{Ushiki13}, and the radar sensors \citep{Kueppers17}.
+An example of SISO LTI sensor is the temperature sensor, but there are sensors that have single input and multiple outputs, like the gas sensors described in \citet{Munther19}, and sensors that have multiple inputs and multiple outputs, like three-axis accelerometers \citet{DEmilia16}, radio-frequency intruder-detection sensors \citet{Ushiki13}, and radar sensors \citet{Kueppers17}.
 
@@
 \begin{equation} 
-\underbrace{ \begin{bmatrix} y(0) \\ y(1) \\ y(2) \\ \vdots \\ y(T) \end{bmatrix} }_{\mathbf{y}}
- = \underbrace{ \begin{bmatrix} \mathbf{C} \\ \mathbf{C} \mathbf{A} \\ \mathbf{C} \mathbf{A}^2 \\ \vdots \\ \mathbf{C} \mathbf{A}^T \end{bmatrix} }_{\mathbfcal{O}} \mathbf{x}(0) +
- \underbrace{ \begin{bmatrix} \mathbf{D} \\ \mathbf{C} \mathbf{B} & \mathbf{D} \\ \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \\ \vdots & \ddots \\ \mathbf{C} \mathbf{A}^{T-1} \mathbf{B} & \cdots  &  \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \mathbf{D} \end{bmatrix} }_{\mathbfcal{T}} \underbrace{ \begin{bmatrix} u(0) \\ u(1) \\ u(2) \\ \vdots \\ u(T) \end{bmatrix} }_{\mathbf{u}} ,
+\underbrace{ \begin{bmatrix} y(0) \\ y(1) \\ y(2) \\ \vdots \\ y(\color{blue}N\color{black}) \end{bmatrix} }_{\mathbf{y}}
+ = \underbrace{ \begin{bmatrix} \mathbf{C} \\ \mathbf{C} \mathbf{A} \\ \mathbf{C} \mathbf{A}^2 \\ \vdots \\ \mathbf{C} \mathbf{A}^{\color{blue}N\color{black}} \end{bmatrix} }_{\mathbfcal{O}} \mathbf{x}(0) +
+ \underbrace{ \begin{bmatrix} \color{blue}D\color{black} \\ \mathbf{C} \mathbf{B} & \color{blue}D\color{black} \\ \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \color{blue}D\color{black} \\ \vdots & \ddots \\ \mathbf{C} \mathbf{A}^{\color{blue}N\color{black}-1} \mathbf{B} & \cdots  &  \mathbf{C} \mathbf{A} \mathbf{B} & \mathbf{C} \mathbf{B} & \color{blue}D\color{black} \end{bmatrix} }_{\mathbfcal{T}} \underbrace{ \begin{bmatrix} u(0) \\ u(1) \\ u(2) \\ \vdots \\ u(\color{blue}N\color{black}) \end{bmatrix} }_{\mathbf{u}} ,
  \label{eqn:knownmodel} \end{equation}
@@
  \end{tikzpicture}
- \caption{Input estimation using a compensator that processes the sensor response ${y}$. The compensator is an additional system that reverts the dynamics of the sensor.} \label{fig:compensator}
+ \caption{Input estimation using a compensator that processes the sensor response ${y}$.
+The compensator is an additional system that \color{blue} inverts \color{black} the dynamics of the sensor.} \label{fig:compensator}
  \end{figure}
@@
 
-With a step input ${u} = \bar{{u}} s$, where $s(k) = 1$, for $k \geq 0$, and $s(k) = 0$, for $k < 0$, the discrete-time state-space representation of the LTI sensor is equivalent to an augmented autonomous system
+With a step input ${u}$, where $u(k) = u$, for $k \geq 0$, and $u(k) = 0$, for $k < 0$, the discrete-time state-space representation of the LTI sensor is equivalent to an augmented autonomous system  
 \begin{equation} \begin{aligned} \mathbf{x}_\text{a}(k+1) &= \underbrace{ \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} }_{\mathbf{A}_\text{a}} \mathbf{x}_\text{a}(k) , \quad \text{where} \ \ \mathbf{x}_\text{a}(k) = \begin{bmatrix} \mathbf{x}(k) \\ {u}(k) \end{bmatrix}, \ \ \mathbf{x}_{\text{ini}} = \mathbf{x}(0) \\
@@
 \[ \left| \lambda \mathbf{I} - \mathbf{A}_\text{a} \right| = \left| \lambda \mathbf{I} - \begin{bmatrix} \mathbf{A} & \mathbf{B} \\ 0 & 1 \end{bmatrix} \right| = \left| \lambda \mathbf{I} - \mathbf{A} \right| \left( \lambda - 1 \right) = 0.\]
-Therefore, the eigenvalues, poles, of the augmented autonomous system (\ref{eqn:augmented}) are the eigenvalues, poles, of the LTI system, with the additional eigenvalue $\lambda = 1$, pole at $(1,0)$.
+Therefore, the eigenvalues (poles) of the augmented autonomous system (\ref{eqn:augmented}) are the eigenvalues, poles, of the LTI system, with the additional eigenvalue $\lambda = 1$, pole at $(1,0)$.
 
@@
 The model identification consists in the estimation of the matrices $\widehat{\mathbf{A}}_\text{a}$ and $\widehat{\mathbf{C}}_\text{a}$, and the initial conditions $\widehat{\mathbf{x}}(0)$.
-To do this, consider that a Hankel matrix $\mathbfcal{H}({y}) \in \mathbb{R}^{n \times n}$, constructed from any linearly independent $n$ autonomous responses from the sensor initial conditions, is full column rank.
-We can express 
-\[ \underbrace{ \begin{bmatrix} y(1) & y(2) & \cdots & y(n) \\ y(2) & y(3) & \cdots & y(n+1) \\ \iddots & \iddots & \iddots \\ y(n) & y(n+1) & \cdots & y(2n-1) \end{bmatrix} }_{ \mathbfcal{H}({y}) }= \underbrace{ \begin{bmatrix} \widehat{\mathbf{C}}_\text{a} \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a} \\ \vdots \\ \widehat{\mathbf{C}}_\text{a} \widehat{\mathbf{A}}_\text{a}^{n} \end{bmatrix} }_{ \mathbfcal{O}_\text{a} }  \underbrace{ \begin{bmatrix} \mathbf{x}_\text{a}(0) & \mathbf{x}_\text{a}(1) & \cdots & \mathbf{x}_\text{a}(n) \end{bmatrix} }_{ \mathbf{X}_\text{ini} } , \]
+To do this, consider that \color{blue}the \color{black} Hankel matrix $\mathbfcal{H}({y}) \in \mathbb{R}^{n \times n}$, constructed from any linearly independent $n$ autonomous responses from the sensor's initial conditions, is full column rank.
+\color{blue}Then, \color{black} we can express \color{blue}the following equality\color{black}  
+\begin{equation} \underbrace{ \begin{bmatrix} y(\color{blue}0\color{black}) & y(\color{blue}1\color{black}) & \cdots & y(\color{blue}n-1\color{black}) \\ y(\color{blue}1\color{black}) & y(\color{blue}2\color{black}) & \cdots & y(\color{blue}n\color{black}) \\ \iddots & \iddots & \iddots \\ y(\color{blue}n-1\color{black}) & y(\color{blue}n\color{black}) & \cdots & y(\color{blue}2n-2\color{black}) \end{bmatrix} }_{ \mathbfcal{H}({y}) }= \underbrace{ \begin{bmatrix} {\mathbf{C}}_\text{a} \\ {\mathbf{C}}_\text{a} {\mathbf{A}}_\text{a} \\ \vdots \\ {\mathbf{C}}_\text{a} {\mathbf{A}}_\text{a}^{n-1} \end{bmatrix} }_{ \mathbfcal{O}_\text{a} }  \underbrace{ \begin{bmatrix} \mathbf{x}_\text{a}(0) & \mathbf{x}_\text{a}(1) & \cdots & \mathbf{x}_\text{a}(n-1) \end{bmatrix} }_{ \mathbf{X}_\text{ini} } , \end{equation}
 where $\mathbf{X}_\text{ini}$ is a matrix with the initial conditions of the $n$ free responses. 
 A singular value decomposition of $\mathbfcal{H}({y})$ 
-\[ \mathbf{U} \bm{\Sigma} \mathbf{V} = \mathbfcal{H}({y}), \]
+\begin{equation} \mathbf{U} \bm{\Sigma} \mathbf{V} = \mathbfcal{H}({y}), \end{equation}
 permits the estimation of the observability matrix $\mathbfcal{O}_\text{a}$ and the initial conditions $\mathbf{X}_\text{ini}$, for example, by choosing
-\[ \widehat{\mathbfcal{O}}_\text{a} = \mathbf{U} \sqrt{\bm{\Sigma}}, \quad \text{and} \quad \widehat{\mathbf{X}}_\text{ini} = \sqrt{\bm{\Sigma}} \mathbf{V} . \]
+\begin{equation} \widehat{\mathbfcal{O}}_\text{a} = \mathbf{U} \sqrt{\bm{\Sigma}}, \quad \text{and} \quad \widehat{\mathbf{X}}_\text{ini} = \sqrt{\bm{\Sigma}} \mathbf{V} . \end{equation}
 The matrices $\widehat{\mathbf{A}}_\text{a}$ and $\widehat{\mathbf{C}}_\text{a}$ can be estimated, from $\widehat{\mathbfcal{O}}_\text{a}$, by solving a system of equations.
@@
 A second method can estimate the step input when the model of the LTI sensor is unknown but its static gain $G$ is given.
-Using the static gain $G$, we can express $\bar{{y}} = G \bar{{u}}$, where $\bar{{u}}$ is the input exact value and $\bar{{y}}$ is the sensor steady state response.
+Using the static gain $G$, we can express ${{y}} = G {{u}}$, where ${{u}}$ is the input exact value and ${{y}}$ is the \color{blue} exact \color{black} sensor steady state response.
 The total response of the system is the sum of the transient and the steady-state responses.
 Thus, considering the augmented autonomous model, we can write
-\[ {y} = G \ \bar{{u}} \ + \ \mathbfcal{O}_\text{a} \ \mathbf{x}(0) , \]
-that in matrix form is
-\[ \underbrace{ \begin{bmatrix}y(0) \\ y(1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \mathbf{C}_\text{a} \\ G & \mathbf{C}_\text{a} \mathbf{A}_\text{a} \\ \vdots & \vdots \\ G & \mathbf{C}_\text{a} \mathbf{A}_\text{a}^T \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} \bar{{u}} \\ \mathbf{x}(0) \end{bmatrix} .\]
-Then, it is necessary to estimate the observability matrix of the augmented system, followed by the estimation of the input  ${u}$, and the initial conditions $\mathbf{x}_{\text{ini}}$, using least-squares
-\[ \begin{bmatrix} \widehat{{u}} \\ \widehat{\mathbf{x}}_{\text{ini}} \end{bmatrix} = \left( \mathbf{K}^\top \mathbf{K} \right)^{-1} \mathbf{K}^\top \mathbf{y}. \]
+\begin{equation} {y} = G \ {{u}} \ + \ \mathbfcal{O}_\text{a} \ \mathbf{x}_{\color{blue}\text{a}\color{black}}(0) , \end{equation}
+\color{blue} which \color{black} in matrix form is
+\begin{equation} \underbrace{ \begin{bmatrix}y(0) \\ y(1) \\ \vdots \\ y(\color{blue}N\color{black}) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \mathbf{C}_\text{a} \\ G & \mathbf{C}_\text{a} \mathbf{A}_\text{a} \\ \vdots & \vdots \\ G & \mathbf{C}_\text{a} \mathbf{A}_\text{a}^{\color{blue}N\color{black}} \end{bmatrix}}_{\mathbf{K}} \begin{bmatrix} {{u}} \\ \mathbf{x}_{\color{blue}\text{a}\color{black}}(0) \end{bmatrix} .\end{equation}
+Then, it is necessary to estimate the observability matrix of the augmented system, followed by the estimation of the input  ${u}$, and the initial conditions, using least-squares
+\begin{equation} \begin{bmatrix} \widehat{{u}} \\ \widehat{\mathbf{x}}_{\color{blue}\text{a}(0)\color{black}} \end{bmatrix} = \left( \mathbf{K}^\top \mathbf{K} \right)^{-1} \mathbf{K}^\top \mathbf{y}.
+\end{equation}
 
@@
 
-A third method can directly estimate the input step level from the step response, without the identification of a sensor model.
-Applying the first difference operator, $\Delta = \sigma - 1$, to the system state-space representation (\ref{eqn:dtsslti}), we have
+A third method can directly estimate the input step level from the step response, without identifying a sensor model.
+To derive this method, we use the first difference operator $\Delta$, defined as $\Delta y(k) = y(k + 1) - y(k)$.
+Applying the first difference operator to the system state-space representation (\ref{eqn:dtsslti}) results in the autonomous system
 \begin{equation} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta {y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
-where $\sigma$ is the shift operator, defined as $(\sigma^\tau y) (k) = y(k + \tau)$, and
-$\Delta {u}(k) = \mathbf{0}$, for $k \geq 0$, and
-$\Delta \mathbf{x}(0) = (\mathbf{A} - \mathbf{I}) \mathbf{x}(0) + \mathbf{B} \bar{{u}}$.
-The resulting system (\ref{eqn:ssalti}) is autonomous.
-When the response $\Delta {y}$ is persistently exciting of order $L$, i.e., when the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ of $L+1$ block rows constructed from $\Delta {y}$ satisfies
-\begin{equation} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) \leq L,  \end{equation}
-the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ provides a linear map to the free responses of the augmented autonomous system (\ref{eqn:augmented}).  
-In this case, the total response of the system is given by 
-\begin{equation*} \mathbf{y} = G \bar{u} + \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell} \end{equation*}
+where $\Delta {u}(k) = \mathbf{0}$, for $k \geq 0$, and
+$\Delta \mathbf{x}(0) = (\mathbf{A} - \mathbf{I}) \mathbf{x}(0) + \mathbf{B} {{u}}$.
+
+If the response $\Delta {y}$ \color{blue} of this autonomous system \color{black}  is persistently exciting of order $L$, then the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ of $L+1$ block rows constructed from $\Delta {y}$ satisfies
+\begin{equation} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) = \color{blue} \mathrm{rank} \left( \begin{bmatrix} \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \iddots & \iddots & \iddots \\ \Delta y(L+1) & \Delta y(L+2) & \cdots & \Delta y(L+n) \end{bmatrix} \right) \color{black} \leq L.
+\end{equation}
+\color{blue}If we assume that the autonomous response $\Delta {y}$ is persistently exciting with a suficiently high order, then the image of the columns of the Hankel matrix $\mathbfcal{H}_{N-n}(\Delta {y})$ is a linear space that contains all the natural responses of the augmented autonomous system (\ref{eqn:augmented})
+\begin{equation} \mathbf{y}_{\mathrm{natural}} = \mathbfcal{H}_{N-n}\left(\Delta {y}\right) \bm{\ell}, \end{equation}
+for any vector $\bm{\ell} \in \mathbb{R}^{n}$.
+Since the autonomous system was derived from the LTI system by augmenting a state and considering that the step input is part of the inital conditions,
+then the total response of the LTI system is the sum of the forced response, due to the applied input ${u}$ and the natural response spanned from the Hankel matrix
+\begin{equation} \mathbf{y} = \mathbf{y}_{\mathrm{forced}}+ \mathbf{y}_{\mathrm{natural}} =G {u} + \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell} ,  \end{equation} \color{black}
 that is equivalent to
-\begin{equation} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(T) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ G & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ G & \Delta y(T-n) & \Delta y(T-n+1) & \cdots & \Delta y(T-1) \end{bmatrix}}_{\mathbf{K}} \underbrace{ \begin{bmatrix} \bar{{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
-where $T$ is the number of samples, and the vector $\bm{\ell}$ is a linear transformation of the system initial conditions \citep{Markovsky15cep}.  
+\begin{equation} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(\color{blue}N\color{black}) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ G & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ G & \Delta y(\color{blue}N\color{black}-n) & \Delta y(\color{blue}N\color{black}-n+1) & \cdots & \Delta y(\color{blue}N\color{black}-1) \end{bmatrix} }_{\mathbf{K}} \underbrace{ \begin{bmatrix} {{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
+where $\color{blue}N\color{black}$ is the number of samples.
+\color{blue} From this point of view, the vector $\bm{\ell}$ is a linear transformation of the system's initial conditions\color{black}.  
 The solution to this system of equations exist and is unique when we have exact data.
-Nevertheless, in the practice, the observations of the sensor step response are perturbed by noise.
+Nevertheless, in practice, the observations of the sensor step response are perturbed by noise.
 
@@
 and
-\begin{equation} \widetilde{\mathbf{K}} = \mathbf{K} + \mathbf{E} . \label{eqn:K0plusnoise} \end{equation}The matrix $\mathbf{E}$ is constructed with the noise data, and is given as
-\begin{equation} \mathbf{E} = \begin{bmatrix} 0 & \Delta \epsilon(1) & \Delta \epsilon(2) & \cdots & \Delta \epsilon(n) \\ 0 & \Delta \epsilon(2) & \Delta \epsilon(3) & \cdots & \Delta \epsilon(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ 0 & \Delta \epsilon(T-n) & \Delta \epsilon(T-n+1) & \cdots & \Delta \epsilon(T-1) \end{bmatrix} .
+\begin{equation} \widetilde{\mathbf{K}} = \mathbf{K} + \mathbf{E} . \label{eqn:K0plusnoise} \end{equation}
+The matrix $\mathbf{E}$ is constructed with the noise data, and is given as
+\begin{equation} \mathbf{E} = \begin{bmatrix} 0 & \Delta \epsilon(1) & \Delta \epsilon(2) & \cdots & \Delta \epsilon(n) \\ 0 & \Delta \epsilon(2) & \Delta \epsilon(3) & \cdots & \Delta \epsilon(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ 0 & \Delta \epsilon(\color{blue}N\color{black}-n) & \Delta \epsilon(\color{blue}N\color{black}-n+1) & \cdots & \Delta \epsilon(\color{blue}N\color{black}-1) \end{bmatrix} .
 \label{eqn:matrixE} \end{equation}
 Therefore, the matrix $\widetilde{\mathbf{K}}$ is expressed as
-\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(T-n) & \Delta \widetilde{y}(T-n+1) & \cdots & \Delta \widetilde{y}(T-1) \end{bmatrix} .
+\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(\color{blue}N\color{black}-n) & \Delta \widetilde{y}(\color{blue}N\color{black}-n+1) & \cdots & \Delta \widetilde{y}(\color{blue}N\color{black}-1) \end{bmatrix} .
 \label{eqn:matrixK} \end{equation}
@@
 \label{eqn:min_seiv} \end{equation}
-where $\widetilde{\mathbf{y}} = \begin{bmatrix} \widetilde{y}(n+1) & \ldots & \widetilde{y}(T) \end{bmatrix}^\top$, and
+where $\widetilde{\mathbf{y}} = \begin{bmatrix} \widetilde{y}(n+1) & \ldots & \widetilde{y}(\color{blue}N\color{black}) \end{bmatrix}^\top$, and
 $\widehat{\bm{\theta}} = \begin{bmatrix} \widehat{{u}} & \widehat{\bm{\ell}}^\top \end{bmatrix}^\top$. 
@@
 
-For metrology applications, it is desired to have a fast online solution of the problem(\ref{eqn:ddsiemnd}).
-The least-squares (LS) approximate solution of this system of equations offers a simple alternative, in its recursive form, to implement the estimation method in real-time.
-Least-squares admit a recursive implementation that avoids the inversion of matrix $\widetilde{\mathbf{K}}$ that increases its size with respect to the sample size $T$.
-Instead, the recursive least-squares (RLS) updates the solution of the system of equations considering the previous value of the estimation.
+The problem (\ref{eqn:ddsiemnd}) admits a least-squares (LS) solution \color{blue} with normal equations given by
+\begin{equation} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}}.  \label{eqn:neq_seiv} \end{equation} 
+The closed form of the LS solution can be expressed as
+\begin{equation} \widehat{\bm{\theta}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
+where $\widetilde{\mathbf{K}}^\dagger$ is the pseudo-inverse matrix of $\widetilde{\mathbf{K}}$.
+As the number of samples $N$ increases, the number of rows of the matrix $\widetilde{\mathbf{K}}$ grows, making the pseudo-inverse computation inefficient since it requires larger flops, memory and time.
+
+For metrology applications, it is desired to have a fast solution that can be obtained in real-time. 
+The recursive algorithm (RLS) is a least-squares alternative to implement the step input estimation method in real-time.
+The RLS recursively updates the solution of the system of equations, for each newly acquired sensor response sample, considering the previous value of the estimation, instead of performing a matrix inversion.
+An expression of the RLS equations is given in \citet{Kailath00book} as
+\begin{equation} \begin{aligned} \widehat{\bm{\theta}}(k) &= \widehat{\bm{\theta}}(k-1) + \bm{\kappa}_{k} \left( \widetilde{y}(k) - \widetilde{\mathbf{k}}_{k} \widehat{\bm{\theta}}(k-1) \right) , \\  \bm{\kappa}_{k} &= \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top / \left( 1 + \widetilde{\mathbf{k}}_{k} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top  \right) \\ \bm{\Psi}(k) &= \left( \mathbf{I} - \bm{\kappa}_{k} \widetilde{\mathbf{k}}_{k} \right) \bm{\Psi}(k-1), \label{eqn:RLS} \end{aligned} \end{equation}
+for $k = 2n+1, 2n+2, \ldots$, where $\widetilde{\mathbf{k}}_{k}$ represents the row of $\widetilde{\mathbf{K}}$ that corresponds to the $k\mathrm{-th}$ sample, $\bm{\kappa}_{k}$ is a gain scalar, and $\Psi(k)$ is a covariance matrix.
+The estimates and the covariance matrix are initialized using the first $n+1$ samples, i.e.,  $\widehat{\bm{\theta}} (2n+1) = \left( \widetilde{\mathbf{K}}_{n+1}^\top \widetilde{\mathbf{K}}_{n+1} \right)^{-1} \widetilde{\mathbf{K}}_{n+1}^\top \widetilde{\mathbf{y}}_{n+1}$ and $\Psi(2n+1) = \left( \widetilde{\mathbf{K}}_{n+1}^\top \widetilde{\mathbf{K}}_{n+1} \right)^{-1}$, where $\widetilde{\mathbf{K}}_{n+1}$ is the matrix $\widetilde{\mathbf{K}}$ with the first $n+1$ rows, and $\widetilde{\mathbf{y}}_{n+1}$ is the vector $\widetilde{\mathbf{y}}$ with the first $n+1$ elements.
+Since the RLS is a recursive implementation of the least-squares computation, the RLS solution is equivalent to the LS solution.
+Moreover, the statistical analysis of the LS solution is valid also for the RLS solution, and is preferred since it is simpler to express the statistics using the closed form of the LS solution.
+
+The computational complexity of the RLS algorithm solution to the system of equations (\ref{eqn:ddsiemexd}) is $O \left( \left( n+1 \right)^2 \right)$.
+The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
+From there on, the RLS algorithm updates the solution with linear complexity.
+Therefore, even though a large number of samples $N$ makes the matrix in the system of equations (\ref{eqn:ddsiemexd}) increase in the number of rows, the estimation of the solution does not increase in complexity.
+The data-driven step input estimation method is then scalable for any sensor of order $n$, and can be executed in devices with limited computational resources, provided that the computation of the  $(n+1) \times (n+1)$ inverse matrix is feasible.
+\color{black}
 
@@
 where $\widebar{\mathbf{K}}$ is exact data information and $\mathbf{E}$ is the corresponding perturbation noise.
-Details of the method formulation are described in \citep{Markovsky15cep}.
+Details of the method formulation are described in \citet{Markovsky15cep}.
 
@@
 \begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} g & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ g & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ g & \Delta \widetilde{y}(T-n) & \Delta \widetilde{y}(T-n+1) & \cdots & \Delta \widetilde{y}(T-1) \end{bmatrix} , \label{eqn:matrix} \end{equation}
-is a Hankel matrix of $(T - n)-\text{block rows}$, constructed from consecutive differences $\Delta \widetilde{y}(t) = \widetilde{y}(t) - \widetilde{y}(t-1)$ of the observed transient response, augmented in the left side with a $(T - n)\text{-vector}$ of elements equal to the known dc-gain $g$, (see \citep{Markovsky15cep}).
+is a Hankel matrix of $(T - n)-\text{block rows}$, constructed from consecutive differences $\Delta \widetilde{y}(t) = \widetilde{y}(t) - \widetilde{y}(t-1)$ of the observed transient response, augmented in the left side with a $(T - n)\text{-vector}$ of elements equal to the known dc-gain $g$, (see \citet{Markovsky15cep}).
 The perturbations $\bm{\epsilon}$ enter in matrix $\widetilde{\mathbf{K}}$ and we can express 
@@
 The recursive LS method allows for a real-time implementation of the step input estimation method.
-Details of the step input estimation method are described in \citep{Markovsky15cep}.
+Details of the step input estimation method are described in \citet{Markovsky15cep}.
 
@@
 
-The estimation of the step input level is obtained as the solution of the minimization problem \citep{Markovsky15ieee, Markovsky15cep} 
+The estimation of the step input level is obtained as the solution of the minimization problem \citet{Markovsky15ieee, Markovsky15cep} 
 \begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert  \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x} \right\Vert^2_2 \label{eqn:min_ls} \end{equation}

### 2020-08-21 ChapterRampInput/ChapterRampInput.tex
Final final version before internal defense
Second draft before public defense

@@
 
-The goal of any measurement is accurately estimating the input using the response of the sensor in the transient regime.
-Here we consider the case where the input varies during the measurement.
-The simplest case is when the input changes with a constant rate, such as being described with an affine function.
-The inputs that vary at a constant rate activate the sensor gradually. 
-An affine input estimation method is needed to estimate the slope and the intercept parameters of the affine function.
+\color{blue} 
+In this thesis document we have discussed that a measurement estimates the unknown value of a physical quantity, also called measurand.
+The physical quantity of interest is applied as an input signal to a dynamic system, the sensor, and its state changes from the initial conditions into a transient state.
+The input value is estimated using the sensor response in transient state.
+A dynamic measurement occurs when the input fluctuations impacts on the input value estimation.
+A typical example of a dynamic measurement problem is a low-bandwidth sensor excited with a fast changing input as is explained in \citet{Dienstfrey14}.
+In this case, the sensor transient response is slower than the variation of the input, making it difficult to track the input evolution.
+The variation of some input characteristics, like the minimum or maximum or the effects of the environment, are considered in applications such as measurements of temperature in \citet{Saggin01}, pressure in \citet{Matthews14}, acceleration in \citet{Link07}, force in \citet{Vlajic16, Hessling08a}, and mass in \cite{Shu93, Boschetti13}.
 
-This chapter describes two signal processing methods for estimating affine input parameters.
-These methods are a data-driven subspace method and a maximum-likelihood (ML) estimation method based on local optimization.
-The subspace is a recursive method independent of the sensor model, and the ML method iteratively minimizes a cost function using a simulated sensor response in a receding time horizon.
+The data-driven step input estimation method, introduced by \citet{Markovsky15cep} and described in Chapter \ref{chap:Preliminaries},
+estimates the unknown step level of an input directly from the sensor step response.
+Since most inputs have variations, contrary to the step response that is constant after its application, sets the need for methods for estimating more input models.
+One of such input models are affine inputs that vary at a constant rate, and they are found in applications where the measurand activates the sensor gradually. 
+An example of gradual activation is the mass measurement of objects in motion, that exists in along conveyor belts.
 
-In this chapter, the dynamic weighing is an implementation example.
-The dynamic weighing of an affine input is a demanding metrology application because the affine excitation turns the linear time-invariant weighing sensor into a time-varying system.
-Then, the affine input estimation methods provide insight into the properties of the dynamic measurement of time-varying inputs.
+Current solutions to the weighing in motion are low pass filters that estimate the mass using a saturated ramp as it is proposed by \citet{Tasaki07}, and by \citet{Niedzwiecki16b}, and also time varying filters are developed in \citet{Piskorowski08} and in \citet{Pietrzak14}.
+Nevertheless, the signal processing affine input estimation methods are motivated by the need to obtain the mass of the object from the ramp before it reaches saturation.
+The ramp is parameterized as a straight line model where the slope and the interception are the parameters of interest of an affine input model.
+This chapter describes two methods for the estimation of the affine input parameters.
 
-The effectiveness and uncertainty of the proposed methods are evaluated with simulation studies.
-The uncertainty is assessed, for the adaptive data-driven method, using a Taylor series expansion of the exponentially weighted recursive least-squares estimate, and
-for the ML method, using the derivatives of the residual error obtained from the to-be-minimized cost function.
+One method is an extension of the data-driven step input estimation method, and then it is a recursive algorithm of low computational cost that can be implemented in real-time.
+The second method is a maximum-likelihood (ML) estimation method based on local-optimization that is proposed for off-line processing of the sensor transient response.
+The ML method is similar to the model predictive control approach, as discussed in \cite{Mayne14}, because a cost function is minimized iteratively to optimize the parameters of a sensor model using the observed sensor response in a receding time horizon.
+The difference is that the ML method aims to estimate the unknown value of the affine input parameters instead of identifying a model and controlling the dynamic system.
+Nevertheless, the ML method can estimate the parameters of the affine input, the parameters of a sensor model, and the initial conditions of the sensor.
+
+The uncertainty of the subspace method is assessed in \cite{QuintanaCSDA} using a Taylor expansion of the estimate and Monte Carlo random sampling approach.
+The Monte Carlo approach requires a large set of generated random samples, and according to \citet{Cox06}, for simple systems it is the recommended method.
+The uncertainty of the ML method is assessed using the derivatives of the residual error that constructs the to-be-minimized cost function.
+The covariance of the optimization method estimate is found in \cite{Pintelon12Book} using the inverse of the Hessian matrix.
+\color{black}
 
@@
 \paragraph{Problem} 
-Given the sequence of measured output observations $\widetilde{\mathbf{y}} = \left( \widetilde{y}(1),\ldots, \widetilde{y}(T)\right)$, with $\widetilde{y}(t) \in {\rm I\!R}$, of a stable linear time-invariant system of order $n$, and static gain $\mathbf{G}$, generated by an affine input $u(t) = \widebar{a} t + \widebar{b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u}(t) = \widehat{a} t + \widehat{b}$ approximates $u$(t).
-The measured observations $\widetilde{\mathbf{y}} = \widebar{\mathbf{y}} + \bm{\epsilon}$ are exact sensor responses $\widebar{\mathbf{y}}$ perturbed by additive noise  $\bm{\epsilon}$ assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.
+Given the sequence of measured output observations $\widetilde{\mathbf{y}} = \left( \widetilde{y}(1),\ldots, \widetilde{y}(\color{blue}N\color{black})\right)$, with $\widetilde{y}(t) \in \mathbb{R}$, of a stable linear time-invariant system of order $n$, and static gain ${G}$, generated by an affine input $u(t) = {a} t + {b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u}(t) = \widehat{a} t + \widehat{b}$ approximates $u$(t).
+The measured observations $\widetilde{\mathbf{y}} = {\mathbf{y}} + \bm{\epsilon}$ are exact sensor responses ${\mathbf{y}}$ perturbed by additive noise  $\bm{\epsilon}$ assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.
 
@@
 \node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
-\node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u(t)$};
+\node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u=at+b$};
 
@@
 The application of an affine input turns the linear time-invariant system into a linear time-varying system, whose
-dynamics depends on the input $u(t) = \widebar{a} t + \widebar{b}$, as it is described by the differential equation:
-\begin{equation} \dfrac{d}{dt} \left( \left( \widebar{a} t + \widebar{b} + m \right) \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = \left( \widebar{a} t + \widebar{b} + m \right) g \end{equation}
+dynamics depends on the input $u(t) = {a} t + {b}$, as it is described by the differential equation:
+\begin{equation} \dfrac{d}{dt} \left( \left( {a} t + {b} + m \right) \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = \left( {a} t + {b} + m \right) g \end{equation}
 where $m$ is the mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.
@@
 The weighing system admits a state-space representation where the states $x_1=y$ and $x_2=\dot{y}$ are the position and the speed of the weighing scale: 
-\begin{equation} \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{\widebar{a} t + \widebar{b} + m} & \dfrac{-(k_{\mathrm{d}} + \widebar{a})}{\widebar{a} t + \widebar{b} + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 \\ g \end{bmatrix}, \quad y = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .
+\begin{equation} \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{{a} t + {b} + m} & \dfrac{-(k_{{d}} + {a})}{{a} t + {b} + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 \\ g \end{bmatrix}, \quad y = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .
 \label{eqn:ltvsystem} \end{equation}
@@
 The profile of an affine input excitation can be reconstructed by solving the structured EIV problem  
-\begin{equation} \widehat{\bm{\theta}} = \underset{\bm{\theta}}{\mathrm{argmin}} \ \left\Vert \bm{\Lambda}^{1/2} \left( \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \bm{\theta} \right) \right\Vert^2_2 .
+\begin{equation} \widehat{\bm{\theta}} = \underset{\bm{\theta}}{\mathrm{argmin}} \ \left\Vert \bm{\Omega}^{1/2} \left( \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \bm{\theta} \right) \right\Vert^2_2 .
 \label{eqn:min_ewrls} \end{equation}
 where $\widetilde{\mathbf{y}}$ is the observed affine response, $\widetilde{\mathbf{K}}$ is defined in (\ref{eqn:K0plusnoise}), and
-the matrix $\bm{\Lambda} \in \mathbb{R}^{(T-n) \times (T-n)}$ is a diagonal matrix of descending powers of the weight factor $\lambda \in [0, 1)$, \textit{i.e.},  $\bm{\Lambda} = \mathrm{diag}\left(\lambda^{T-n}, \ldots, \lambda^2, \lambda^1 \right)$.
-The weight factor $\lambda$ is a data selection forgetting factor since it enables to apply different weights to the residuals $\widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \bm{\theta}$.
+the matrix $\bm{\Omega} \in \mathbb{R}^{(\color{blue}N\color{black}-n) \times (\color{blue}N\color{black}-n)}$ is a diagonal matrix of descending powers of the weight factor $\omega \in [0, 1)$, \textit{i.e.},  $\bm{\Omega} = \mathrm{diag}\left(\omega^{\color{blue}N\color{black}-n}, \ldots, \omega^2, \omega^1 \right)$.
+The weight factor $\omega$ is a data selection forgetting factor since it enables to apply different weights to the residuals $\widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \bm{\theta}$.
 
 The exponentially weighted RLS solution to the problem (\ref{eqn:min_ewrls}) is implemented to facilitate the online implementation of the affine input estimation method.
-When $\lambda=1$, the exponentially weighted RLS solution is identical to that of the RLS.
-When $\lambda<1$, the older residuals are weighted with lower values than the residuals of recent observations.
+\color{blue}
+The exponentially weighted recursive least-squares (RLS) can solve recursively the estimation problem (\ref{eqn:min_seiv}).
+In \citet{Kailath00book} the exponentially weighted RLS algorithm is described as follows
+
+\begin{equation} \begin{aligned} \widehat{\bm{\theta}}(k) &= \widehat{\bm{\theta}}(k-1) + \bm{\kappa}_{k} \left( \widetilde{y}(k) - \widetilde{\mathbf{k}}_{k} \widehat{\bm{\theta}}(k-1) \right) , \\  \bm{\kappa}_{k} &= \omega^{-1} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top / \left( 1 + \omega^{-1} \widetilde{\mathbf{k}}_{k} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top  \right) \\ \bm{\Psi}(k) &= \omega^{-1} \left( \mathbf{I} - \bm{\kappa}_{k} \widetilde{\mathbf{k}}_{k} \right) \bm{\Psi}(k-1), \label{eqn:RLS} \end{aligned} \end{equation}
+for $k = 2n+1, 2n+2, \ldots$, where $\widetilde{\mathbf{k}}_{k}$ represents the row of $\widetilde{\mathbf{K}}$ that corresponds to the $k\mathrm{-th}$ sample, $\bm{\kappa}_{k}$ is a gain scalar, and $\Psi(k)$ is a covariance matrix.
+The estimates and the covariance matrix are initialized using the first $n+1$ samples, i.e.,  
+\begin{equation} \begin{aligned} \widehat{\bm{\theta}} (2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top  \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega}  \widetilde{\mathbf{y}}_{n+1}, \quad \mathrm{and} \\ \Psi(2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} , \end{aligned} \end{equation}
+where $\widetilde{\mathbf{K}}_{n+1}$ is the matrix $\widetilde{\mathbf{K}}$ with the first $n+1$ rows, and $\widetilde{\mathbf{y}}_{n+1}$ is the vector $\widetilde{\mathbf{y}}$ with the first $n+1$ elements\color{black}.
+When $\omega=1$, the exponentially weighted RLS solution is identical to that of the RLS.
+When $\omega<1$, the older residuals are weighted with lower values than the residuals of recent observations.
 In this way, the solution of the minimization problem depends more on newer data, and less on older data.
@@
 In Figure \ref{fig:sensor_response} we can see an example of the affine input $\mathbf{u}$ that excites a weighing sensor, and the generated response $\widetilde{\mathbf{y}}$.
-After the transient effects have diminished, the sensor response is a ramp of the same slope as the input, but with a different intercept due to an offset $\tau$.
-The offset $\tau$ is explained by the parameter $m$ in the sensor representation given in Figure \ref{fig:msd_system}.
+After the transient effects have diminished, the sensor response is a ramp of the same slope as the input, but with a different intercept due to an offset $\omicron$.
+The offset $\omicron$ is explained by the parameter $m$ in the sensor representation given in Figure \ref{fig:msd_system}.
 This is the mass of the sensor itself, that is added to the mass of the object under measurement.	
@@
 \begin{equation} 
-  \begin{bmatrix} 1 & 1  \\ \vdots & \vdots \\ T & 1 \end{bmatrix} 
+  \begin{bmatrix} 1 & 1  \\ \vdots & \vdots \\ \color{blue}N\color{black} & 1 \end{bmatrix} 
   \begin{bmatrix} \widehat{a} \\ \widehat{b}  \end{bmatrix} = 
-  \begin{bmatrix} \widehat{u}(1)-\tau \\ \vdots \\ \widehat{u}(T)-\tau \end{bmatrix}, \label{eqn:LS}
+  \begin{bmatrix} \widehat{u}(1)-\omicron \\ \vdots \\ \widehat{u}(\color{blue}N\color{black})-\omicron \end{bmatrix}, \label{eqn:LS}
 \end{equation}
-where it is considered the existance of the offset $\tau$.
+where it is considered the existance of the offset $\omicron$.
 
 For each new observation $y(t)$, the estimation $\widehat{\mathbf{u}}$ is updated followed by the RLS update of the slope $\widehat{a}$ and the intercept $\widehat{b}$.
-The values of the tuning parameters $\lambda$ and $\tau$ can be obtained in the calibration of the method using the response of the sensor.
+The values of the tuning parameters $\omega$ and $\omicron$ can be obtained in the calibration of the method using the response of the sensor.
 
 The calibration procedure consists in the measurement of an object with a reference mass.
-During the calibration, $\lambda$ and $\tau$ are tuned by changing their values aiming to reduce the estimation errors of $\widehat{a}$ and $\widehat{b}$.
+During the calibration, $\omega$ and $\omicron$ are tuned by changing their values aiming to reduce the estimation errors of $\widehat{a}$ and $\widehat{b}$.
 This reduction of the estimation errors is observed in the convergence of the affine input parameters towards their true values.
@@
 \centering
-\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_2_1.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterRampInput/fig/Fig_2_1.pdf} 
 \caption{ \label{fig:sensor_response} When the transient part of the sensor response $\widetilde{\mathbf{y}}$ asymptotically decreases, a ramp with the same slope $a$ of the  affine input $u(t) = at+b$ predominates.
-The intercept of the prevailing ramp is the intercept $b$ added to and offset $\tau$.}
+The intercept of the prevailing ramp is the intercept $b$ added to and offset $\omicron$.}
 \end{figure}
@@
 To obtain the first and second moments of the step input estimate $\widehat{\mathbf{u}}$, we need to study the solution
-\begin{equation} \widehat{\bm{\theta}} = ( \widetilde{\mathbf{K}}^\top \bm{\Lambda} \widetilde{\mathbf{K}}  )^{-1} \widetilde{\mathbf{K}}^\top \bm{\Lambda} \widetilde{\mathbf{y}} , \label{eqn:xhatW} \end{equation}
+\begin{equation} \widehat{\bm{\theta}} = ( \widetilde{\mathbf{K}}^\top \bm{\Omega} \widetilde{\mathbf{K}}  )^{-1} \widetilde{\mathbf{K}}^\top \bm{\Omega} \widetilde{\mathbf{y}} , \label{eqn:xhatW} \end{equation}
 of the overdetermined structured errors-in-variables EIV problem (\ref{eqn:min_ewrls}).
 Using a second order Taylor series expansion of the inverse matrix we can approximate the LS solution as
-\begin{equation} \widehat{\bm{\theta}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{C}^{-1} (\mathbf{K} + \mathbf{E})^\top \bm{\Lambda} (\mathbf{y} + \bm{\epsilon}).
+\begin{equation} \widehat{\bm{\theta}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{C}^{-1} (\mathbf{K} + \mathbf{E})^\top \bm{\Omega} (\mathbf{y} + \bm{\epsilon}).
 \label{eqn:xhatexpW} \end{equation}
 where 
-\begin{equation} \mathbf{C} = \mathbf{K}^\top \bm{\Lambda} \mathbf{K}, \quad \text{and} \quad \mathbf{M} = \mathbf{C}^{-1} ( \mathbf{K}^\top \bm{\Lambda} \mathbf{E} + \mathbf{E}^\top \bm{\Lambda} \mathbf{K} + \mathbf{E}^\top \bm{\Lambda} \mathbf{E} ).
+\begin{equation} \mathbf{C} = \mathbf{K}^\top \bm{\Omega} \mathbf{K}, \quad \text{and} \quad \mathbf{M} = \mathbf{C}^{-1} ( \mathbf{K}^\top \bm{\Omega} \mathbf{E} + \mathbf{E}^\top \bm{\Omega} \mathbf{K} + \mathbf{E}^\top \bm{\Omega} \mathbf{E} ).
 \end{equation}
@@
 Considering the structure of the EIV problem, the bias and the covariance of the estimate approximation (\ref{eqn:xhatexpW}) can be expressed as
-\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{C}^{-1} \left(  \left( \mathbf{K}^\top \bm{\Lambda} \mathbf{B}_1 - \mathbf{B}_2 \right) \bm{\theta} - \left( \mathbf{K}^\top \bm{\Lambda} \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasEW} \end{equation}
-\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \bm{\Lambda} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \bm{\Lambda} \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right), \end{aligned} \label{eqn:varEW} \end{equation}
-where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\Lambda} \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Lambda} \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\Lambda} \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Lambda} \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \Big\}$, $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger \bm{\Lambda}$, and $\mathbf{K}^\dagger$ is the pseudo-inverse matrix of $\mathbf{K}$.
+\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{C}^{-1} \left(  \left( \mathbf{K}^\top \bm{\Omega} \mathbf{B}_1 - \mathbf{B}_2 \right) \bm{\theta} - \left( \mathbf{K}^\top \bm{\Omega} \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasEW} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \bm{\Omega} \left( \sigma_{\epsilon}^2 \mathbf{I}_{\color{blue}N\color{black}-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \bm{\Omega} \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right), \end{aligned} \label{eqn:varEW} \end{equation}
+where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\Omega} \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Omega} \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\Omega} \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Omega} \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \Big\}$, $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger \bm{\Omega}$, and $\mathbf{K}^\dagger$ is the pseudo-inverse matrix of $\mathbf{K}$.
 
@@
 We have then
-\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{C}}^{-1} \left(  \left( \widetilde{\mathbf{K}}^\top \bm{\Lambda} \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\bm{\theta}} - \left( \widetilde{\mathbf{K}}^\top \bm{\Lambda} \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasSTW} \end{equation}
-  \begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \bm{\Lambda} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \bm{\Lambda} \widetilde{\mathbf{K}}^{\dagger \top} - \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \widetilde{\mathbf{b}}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right), \end{aligned} \label{eqn:varSTW} \end{equation}
-  where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\Lambda} \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Lambda} \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\Lambda} \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Lambda} \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \bm{\epsilon}^\top \Big\}$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger \bm{\Lambda}$.
+\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{C}}^{-1} \left(  \left( \widetilde{\mathbf{K}}^\top \bm{\Omega} \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\bm{\theta}} - \left( \widetilde{\mathbf{K}}^\top \bm{\Omega} \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasSTW} \end{equation}
+  \begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \bm{\Omega} \left( \sigma_{\epsilon}^2 \mathbf{I}_{\color{blue}N\color{black}-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \bm{\Omega} \widetilde{\mathbf{K}}^{\dagger \top} - \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \widetilde{\mathbf{b}}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right), \end{aligned} \label{eqn:varSTW} \end{equation}
+  where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\Omega} \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Omega} \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\Omega} \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Omega} \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \bm{\epsilon}^\top \Big\}$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger \bm{\Omega}$.
 
 The results of the expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$ can be found using Lemma \ref{lem:lemma1}, on page \pageref{lem:lemma1}.
-The bias and covariance were obtained to extend the previous analysis conducted on EIV estimation problems without an imposed structure \citep{Vaccaro94, Stewart90SPT}.
+The bias and covariance were obtained to extend the previous analysis conducted on EIV estimation problems without an imposed structure in \citet{Vaccaro94} and \citet{Stewart90SPT}.
 It was shown that the bias and variance expressions (\ref{eqn:biasSTW}) and (\ref{eqn:varSTW}) are valid predictions of the first and second moments of the LS estimate of a Hankel structured EIV problem.
@@
 \subsection{Maximum-likelihood method}
-Using a model of the sensor, the ML method simultaneously estimates the parameters of the applied affine input $u(t) = at + b$, and some of the sensor model parameters.
-This means that the ML method implementation depends on the problem under study and not always can be estimated all the sensor model parameters.  
-The identifiability conditions of each sensor model parameter is the requirement to fulfill in order to be estimated with the ML method.
+
+% This means that the ML method implementation depends on the problem under study and not always can be estimated all the sensor model parameters.  
+% The identifiability conditions of each sensor model parameter is the requirement to fulfill in order to be estimated with the ML method.
 %The affine input parameters $a$ and $b$ are the slope and the intercept of the straight line model $u(t) = at + b$.
 
-The ML method is an iterative minimization method, described in Algorithm \ref{Alg1}, that minimizes the residual $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$.
-This residual is the difference between the measured sensor response $\widetilde{\mathbf{y}}$ and a simulated response $\widehat{\mathbf{y}}(\bm{\theta})$, that depends on the optimization variables $\bm{\theta}$.
-The ML method uses the Jacobian matrix of the residual $\mathbf{r}$ to search the direction in which the residual decreases.
-The Jacobian matrix $\partial \mathbf{r} / \partial \bm{\theta}$ can be analytically obtained using the sensor mathematical model.
+Using a model of the sensor, the maximum-likelihood ML method simultaneously estimates the parameters of the applied affine input $u(t) = at + b$, and some of the sensor model parameters, \color{blue} using the observed sensor response $\widetilde{\mathbf{y}}$\color{black}.
+\color{blue} This method fits iteratively the response of the model $\widehat{\mathbf{y}}(\bm{\theta})$ to the sensor response by searching for the parameters $\bm{\theta}$ that reduce the error difference between both responses.
+%the residual $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$ 
+Algorithm \ref{Alg1} describes the formulation of an objective function $f = \mathbf{r}^\top \mathbf{r}$, where $\mathbf{r}$ is the residual $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$, that is minimized through the iterations.
+The objective function is the sum of the squares of the differences between the samples of both responses.
+Since the measurement noise is assumed Gaussian distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$, where the samples of this perturbation are independent and identically distributed, the minimization of the objective function $f$ maximizes the likelihood of fitting the actual sensor response with the model response.
 
-\begin{algorithm}
+The Jacobian matrix of the residual $\mathbf{J}\bm{\theta} = \partial \mathbf{r}(\bm{\theta}) / \partial \bm{\theta}$ can be analytically derived using the sensor model.
+With the Jacobian matrix, the ML method searches the direction in which the residual decreases towards a local minimum.
+Unfortunately, this ML method cannot guarantee the global minimum since it depends strongly on the initialization of the optimization parameters, as it is discussed in \citet{Nocedal06}.
+Therefore, it is proposed to initialize the to-be-optimized parameters $a$ and $b$ using the results of the subspace method obtained after a convenient number of samples so that the affine input parameters are set close to the optimal values.
+The initialization of the remaining model parameters under study can be done using the sensor model representation.
+An illustrative example of the ML will be provided in a posterior subsection.  
+\color{black}
+
+\begin{algorithm} \tag{6.1}
 \caption{ML Affine input estimation.}\label{Alg1}
@@
   \REQUIRE{$\widetilde{\mathbf{y}}$, and required sensor model parameters}{}
-  \STATE {Initialize $\bm{\theta} = (a, b, \mathrm{and the to-be-estimated \ model \ parameters})$}{}
-  \FOR{each $N$ observations of $\widetilde{\mathbf{y}}$}{}
+  \STATE {Initialize $\bm{\theta} = (a, b, \text{and \ the \ to-be-estimated \ model \ parameters})$}{}
+  \FOR{each $N_{\mathrm{s}}$ observations of $\widetilde{\mathbf{y}}$}{}
   \STATE {Simulate model response $\widehat{\mathbf{y}}(\bm{\theta})$}{}
-  \STATE {Compute error $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$}{}
-  \STATE {Minimize $\mathbf{r}^\top \mathbf{r}$ over $\bm{\theta}$}{}
-  \STATE {\hspace{0.5cm}  using analytic Jacobian matrix $\mathbf{J} = \partial \mathbf{r}(\bm{\theta}) / \partial \bm{\theta}$}{}
+  \STATE {Obtain residual $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$}{}
+  \STATE {Minimize $f = \mathbf{r}^\top \mathbf{r}$ over $\bm{\theta}$}{}
+  \STATE {\hspace{0.5cm}  using analytic Jacobian matrix $\mathbf{J}\bm{\theta} = \partial \mathbf{r}(\bm{\theta}) / \partial \bm{\theta}$}{}
   \STATE {Update $\bm{\theta}$}{}
 \ENDFOR
-\ENSURE {Optimized parameters $\widehat{a}, \widehat{b}$, and estimated model parameters.} }{}
+\ENSURE {Optimized parameters $\widehat{a}, \ \widehat{b}$, \ and estimated model parameters.} {}
 
@@
 The analytic formulation of the Jacobian benefits the estimation method in two ways: it speeds up the minimization and gives direct access to the variance of the estimates.
-The covariance matrix of the ML estimates can be expressed as \citep{Pintelon12Book}
-\begin{equation} \mathbf{Cov} \left( \widehat{\bm{\theta}} \right) = \sigma_{\bm{\epsilon}}^2 \left( \left( \dfrac{\partial \mathbf{r} }{ \partial \bm{\theta} } \right)^\top \left( \dfrac{\partial \mathbf{r} }{ \partial \bm{\theta} } \right) \right)^{-1}.
+The covariance matrix of the ML estimates can be expressed as in \citet{Pintelon12Book}
+\begin{equation} \mathbf{C} \left( \widehat{\bm{\theta}} \right) = \sigma_{\bm{\epsilon}}^2 \left( \left( \dfrac{\partial \mathbf{r} }{ \partial \bm{\theta} } \right)^\top \left( \dfrac{\partial \mathbf{r} }{ \partial \bm{\theta} } \right) \right)^{-1}.
 \label{eqn:covOpt} \end{equation}
@@
  \begin{equation} \begin{aligned}
-     & \text{Minimize} \quad \text{over} \ a, b, \mathbf{x}_{\text{ini}} \quad \mathbf{r}^T \mathbf{r} \text{, subject to:} \\ & \ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 \\ g \end{bmatrix}, \\ & \ \widehat{\mathbf{y}} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .
+     & \text{Minimize} \quad \text{over} \ a, b, \mathbf{x}_{\text{ini}} \quad \mathbf{r}^\top \mathbf{r} \text{, subject to:} \\ & \ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 \\ g \end{bmatrix}, \\ & \ \widehat{\mathbf{y}} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .
  \label{eqn:ltv_abxini} \end{aligned} \end{equation}
-where the optimization variables $\bm{\theta} = \left(a, b, \mathbf{x}_{\mathrm{ini}} \right)$ are the affine input parameters $a$, $b$, and the sensor model initial conditions.
+where the optimization variables $\bm{\theta} = \left(a, b, \mathbf{x}_{\mathrm{ini}} \right)$ are the affine input parameters $a$, $b$, and the sensor model's initial conditions.
 
@@
 With the initial affine input parameters, a sensor response can be simulated, and, since we are using few samples, the first samples of the simulated response also are an approximation of the initial conditions $\mathbf{x}_{\mathrm{ini}}$.
-The optimization variables are updated every $N$ new observations.
-The minimization algorithm can be, for example, the Levenberg-Marquardt algorithm \citep{Nocedal06}.
+The optimization variables are updated every $N_{\mathrm{s}}$ new observations.
+The minimization algorithm can be, for example, the Levenberg-Marquardt algorithm described in \citet{Nocedal06}.
 
@@
 
-The time-varying (TV) filter described in \citep{Pietrzak14} was designed to compensate the measured responses of a conveyor weighing system, considering they are modeled as a saturated ramp.
+The time-varying (TV) filter described in \citet{Pietrzak14} was designed to compensate the measured responses of a conveyor weighing system, considering they are modeled as a saturated ramp.
 The TV filter consists of three low-pass infinite impulse response (IIR) filters in cascade, where the $i-\mathrm{th}$ IIR filter is given by
 \begin{equation} \widehat{y}_i(t) + k_1(t) \widehat{y}_i(t-1) = k_2(t) \left( \widehat{y}_{i-1}(t) + \widehat{y}_{i-1}(t-1) \right) \end{equation}
-for $i = 1,\ldots,3$ and $t=0,\ldots,T$.
+for $i = 1,\ldots,3$ and $t=0, \ldots, \color{blue}N\color{black}$.
 The sensor response is fed to the filter, then $\widehat{y}_0(t) = \widetilde{y}(t)$, and the output of the TV filter $\widehat{u}_\mathrm{ltv}(t) = \widehat{y}_3(t)$ is an estimation of the affine input.
@@
 
-The time-varying coefficients $k_1(t)$ and $k_2(t)$ are computed from \citep{Pietrzak14}
+The time-varying coefficients $k_1(t)$ and $k_2(t)$ are computed in \citet{Pietrzak14} from 
 \begin{equation} k_1(t) = \frac{f_c(t) - \frac{k_3}{\pi T_s}}{f_c(t) + \frac{k_3}{\pi T_s}}, \quad k_2(t) = \frac{1 + k_1(t)}{2} , \quad k_3 = \sqrt{ \sqrt[3]{2}  - 1} \end{equation}
 where $T_s$ is the sampling time and $f_c(t)$ is a heuristic "cutoff" frequency
-\begin{equation} f_c(t) = f_u + \left( f_l - f_u \right) \beta^{\frac{t-1}{\alpha \left( T-1 \right)} } \end{equation}
+\begin{equation} f_c(t) = f_u + \left( f_l - f_u \right) \beta^{\frac{t-1}{\alpha \left( \color{blue}N\color{black}-1 \right)} } \end{equation}
 that changes between the lower $f_l$ and upper $f_u$ limits, where the coefficient $\beta$ is lower than one, and $\alpha$ is the decay rate. 
 The lower frequency value $f_l$ and the coefficient $\beta$ are fixed and the variables $f_u$ and $\alpha$ are optimized off-line by solving the minimization problem 
-\begin{equation} \mathrm{minimize} \ \quad \ \mathrm{over} \ f_u, \alpha \ \quad \ \mathrm{max} \left( \dfrac{\mu_{\widetilde{a}_{\mathrm{ltv}}}}{\mu_{\mathrm{spec}}}, \dfrac{\mu_{\widetilde{b}_{\mathrm{ltv}}}}{\mu_{\mathrm{spec}}}, \dfrac{\sigma_{\widetilde{a}_{\mathrm{ltv}}}}{\sigma_{\mathrm{spec}}}, \dfrac{\sigma_{\widetilde{b}_{\mathrm{ltv}}}}{\sigma_{\mathrm{spec}}} \right) + \mathrm{max} \left( \dfrac{\eta_{i}}{T} \right) \label{eqn:tv_optim} \end{equation}
-where $\mu_{\widetilde{a}_\mathrm{ltv}}$, $\mu_{\widetilde{b}_\mathrm{ltv}}$, $\sigma_{\widetilde{a}_\mathrm{ltv}}$ and $\sigma_{\widetilde{b}_\mathrm{ltv}}$ are the mean values and the standard deviations of the estimation errors $\widetilde{a}_\mathrm{ltv} = \widehat{a}_\mathrm{ltv} - \widebar{a}$, and $\widetilde{a}_\mathrm{ltv} = \widehat{b}_\mathrm{ltv} - \widebar{b}$, and where $\widebar{a}$ and $\widebar{b}$ are the true values of the input parameters.
-The values $\mu_{\mathrm{spec}}$ and $\sigma_{\mathrm{spec}}$ are specified in the OIML recommendation R51 \citep{OIML_R51_1} for mass measurements that use a conveyor belt.
+\begin{equation} \mathrm{minimize} \ \quad \ \mathrm{over} \ f_u, \alpha \ \quad \ \mathrm{max} \left( \dfrac{\mu_{\widetilde{a}_{\mathrm{ltv}}}}{\mu_{\mathrm{spec}}}, \dfrac{\mu_{\widetilde{b}_{\mathrm{ltv}}}}{\mu_{\mathrm{spec}}}, \dfrac{\sigma_{\widetilde{a}_{\mathrm{ltv}}}}{\sigma_{\mathrm{spec}}}, \dfrac{\sigma_{\widetilde{b}_{\mathrm{ltv}}}}{\sigma_{\mathrm{spec}}} \right) + \mathrm{max} \left( \dfrac{\eta_{i}}{\color{blue}N\color{black}} \right) \label{eqn:tv_optim} \end{equation}
+where $\mu_{\widetilde{a}_\mathrm{ltv}}$, $\mu_{\widetilde{b}_\mathrm{ltv}}$, $\sigma_{\widetilde{a}_\mathrm{ltv}}$ and $\sigma_{\widetilde{b}_\mathrm{ltv}}$ are the mean values and the standard deviations of the estimation errors $\widetilde{a}_\mathrm{ltv} = \widehat{a}_\mathrm{ltv} - {a}$, and $\widetilde{a}_\mathrm{ltv} = \widehat{b}_\mathrm{ltv} - {b}$, and where ${a}$ and ${b}$ are the true values of the input parameters.
+The values $\mu_{\mathrm{spec}}$ and $\sigma_{\mathrm{spec}}$ are specified in the OIML recommendation R51 defined in \citet{OIML_R51_1} for mass measurements that use a conveyor belt.
 
@@
 A second order weighing system was excited with an affine input to get the transient response.
-The parameters of the weighing system are $m$ = 15 g, $k_{\mathrm{d}$ = 5.5 Ns/m and $kg_{\mathrm{s}$ = 10250 N/m.
+The parameters of the weighing system are $m$ = 15 g, $k_{\mathrm{d}}$ = 5.5 Ns/m and $kg_{\mathrm{s}}$ = 10250 N/m.
 The applied affine input $u(t) = 100 t + 10$ represents a mass that changes from $10$ g to $110$ g, at a constant rate, in a time interval of $0.1$ s.
@@
 \centering
-\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_2.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterRampInput/fig/Fig_2.pdf} 
 \caption{ \label{fig:sensor_weight} The sensor transient response $\widetilde{\mathbf{y}}$ to an affine input excitation $u(t) = at+b$ is processed by the estimation methods to estimate the parameters $a$ and $b$.
 In the figure we observe an example of the input estimate $\widehat{\mathbf{u}}$ obtained with the subspace method.
-The input parameters are calculated from $\widehat{\mathbf{u}} - \tau$ using linear regression.}
+The input parameters are calculated from $\widehat{\mathbf{u}} - \omicron$ using linear regression.}
 \end{figure}
@@
 These SNR values are realistic in practical applications.
-The SNR is defined as the ratio of signal power to the noise power, that is equivalent to the root-mean-square (RMS) value of the true signal to the standard deviation of the perturbation noise, and in dB is given as
-
-\begin{equation} \mathrm{SNR} = 20 \log_{10}{\dfrac{ \sqrt{ \dfrac{1}{T}\int\limits_0^T{\widebar{y}(t)^2 d t} } } { \sigma_{\bm{\epsilon}}}} \end{equation}
+The SNR is defined as the ratio of signal power to the noise power, that is equivalent to the root-mean-square (RMS) value of the true signal to the standard deviation of the perturbation noise, and in dB is given by \color{blue}(\ref{eqn:SNR})\color{black}.
 
@@
 The first estimation was obtained with $2n+1$ samples, where $n=2$ is the assumed order of the sensor.
-The method updated recursively the value of the estimated parameters for each new collected sample, using the forgetting factor $\lambda$ listed in Table \ref{table:lambdas}.
+The method updated recursively the value of the estimated parameters for each new collected sample, using the forgetting factor $\omega$ listed in Table \ref{table:lambdas}.
 In Figure \ref{fig:rele_dd_40dB_s1} we observe the relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ obtained when SNR = 40 dB. 
@@
 \centering
-\caption{Here are listed the selected values of the forgetting factor $\lambda$ and the offset $\tau$ that configure the subspace method for the different values of SNR.
+\caption{Here are listed the selected values of the forgetting factor $\omega$ and the offset $\omicron$ that configure the subspace method for the different values of SNR.
 These values were obtained after calibration of the method and were fixed during the simulation study.}
@@
  \hline
- $\lambda$ & 0.939 & 0.940 & 0.955 & 0.959 & 0.959 \\ % forgetting factor
- $\tau$ [g] \  & 15 & 15 & 17 & 14 & 20 \\ [0.5ex] % [\mathrm{samples}]
+ $\omega$ & 0.939 & 0.940 & 0.955 & 0.959 & 0.959 \\ % forgetting factor
+ $\omicron$ [g] \  & 15 & 15 & 17 & 14 & 20 \\ [0.5ex] % [\mathrm{samples}]
  \hline
@@
 \centering
-\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_3.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterRampInput/fig/Fig_3.pdf} 
 \caption{ \label{fig:rele_dd_40dB_s1} The relative errors of the affine input parameters estimates decrease as the subspace method processes more samples.
@@
 \centering
-\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_4.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterRampInput/fig/Fig_4.pdf} 
 \caption{ \label{fig:rele_SNR_dd_10000} The minimum value of the estimation relative errors obtained with the subspace method is less than 2  \% regardless of the SNR between 20 dB and 60 dB. }
@@
 \centering
-\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_5.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterRampInput/fig/Fig_5.pdf} 
 \caption{ \label{fig:CRLB_MSE_ab_dd_40dB_MC_10000} When the SNR of the sensor response is 40 dB, the mean squared errors of the slope estimate $\widehat{a}$ and the intercept estimate $\widehat{b}$, obtained by the subspace method, are two orders of magnitude above the theoretical minimum variance given by the Cram\'er-Rao lower bound.}
@@
 \centering
-\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_6.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterRampInput/fig/Fig_6.pdf} 
 \caption{ \label{fig:CRLB_MSE_SNR_ab_dd_MC_10000} The Cram\'er-Rao lower bounds of the estimates $\mathrm{CRLB}_a$ and $\mathrm{CRLB}_b$ determine the minimum uncertainty that can be achieved and increases with the measurement noise.
@@
 \subsection{Results of the maximum-likelihood method}
-
-The maximum-likelihood (ML) method processed off-line the sensor transient response, perturbed with measurement noise of 40 dB SNR.
-The ML method used the first 50 samples to initialize the optimization variables and updated the variables every $N = 5$ samples.
-In Figure \ref{fig:rele_dd_40dB_s1} are shown the relative errors of the estimates $\widehat{a}$, $\widehat{b}$, $\widehat{x}_{\mathrm{ini,1}}$, and $\widehat{x}_{\mathrm{ini,2}}$.
+\color{blue} 
+The maximum-likelihood (ML) method estimates the affine input parameters from the sensor model transient response.
+A total of 100 runs of the method were obtained, using different realizations of the measurement noise.
+The measurement noise variance was set to have an SNR of 40 dB.
+The ML method used the first 50 samples to initialize the optimization variables and updated the variables every $N_{\mathrm{s}} = 5$ samples.
+Figure \ref{fig:rele_lo_40dB_s10} shows the average of the observed relative errors in the estimation of the parameters $\widehat{a}$, $\widehat{b}$, $\widehat{x}_{\mathrm{ini,1}}$, and $\widehat{x}_{\mathrm{ini,2}}$\color{black}.
 The convergence of the ML estimates gives relative errors below 5 \% after three iterations.
@@
 \centering
-\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_7.pdf} 
-\caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor initial conditions are estimated with the ML method.
-After three iterations the relative errors of the estimates are smaller than 5 \%.
-}
+\includegraphics[width=0.69\columnwidth]{./ChapterRampInput/fig/Fig_7.pdf} 
+\caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method.
+After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and every of the estimates.
+The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%.
+\color{black} }
 \end{figure}
 
-The estimation covariance was computed using the analytic Jacobian and Equation (\ref{eqn:covOpt}).
-Figure \ref{fig:cov_lo_40dB_s1} shows the variances of the optimization variables, that are on the diagonal of the covariance matrix $\mathbm{J}$. 
-We can see that the estimation variance decrease as more samples are processed.
+The estimation covariance was computed using the analytic Jacobian and the Equation (\ref{eqn:covOpt}).
+Figure \ref{fig:cov_lo_40dB_s1} shows \color{blue} the average \color{black} of the variances that were observed on the diagonal of the covariance matrix $\mathbf{J}$. 
+We can see that the estimation variances decrease as more samples are processed.
 Moreover, the estimation variances of $\widehat{a}$ and $\widehat{b}$ obtained with the ML method are lower than 
@@
 \centering
-\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_8.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterRampInput/fig/Fig_8.pdf} 
 \caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian.
-The variances of the affine input parameters estimates decrease faster than the variances of the initial conditions estimates.
-}
+\color{blue} The estimates $\widehat{x}_{\mathrm{ini,1}}$ and $\widehat{a}$ have the smaller and larger variances, respectively.
+\color{black}  }
 \end{figure}
 
-The ML method is computationally more expensive than the subspace method because the ML method simulates the response of a sensor model to optimize the input parameters and the sensor initial conditions.
+The ML method is computationally more expensive than the subspace method because the ML method simulates the response of a sensor model to optimize the input parameters and the sensor's initial conditions.
 
@@
 The upper value $f_u$ and the decay rate $\alpha$ were found using optimization (\ref{eqn:tv_optim}). 
-We chose the values $\mu_{\mathrm{spec}}=0.5$ and $\sigma_{\mathrm{spec}}=0.24$ as they are specified in the OIML recommendation \citep{OIML_R51_1} for a mass of 100 g measured in a conveyor belt.
+We chose the values $\mu_{\mathrm{spec}}=0.5$ and $\sigma_{\mathrm{spec}}=0.24$ as they are specified in the OIML recommendation defined in \citet{OIML_R51_1} for a mass of 100 g measured in a conveyor belt.
 The optimized values of the frequency upper value and the decay rate, using a dataset of 100 transient responses, were $f_u = 26.94$ Hz and $\alpha = 5.71$.
@@
 \centering
-\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_9.pdf} 
+\includegraphics[width=0.69\columnwidth]{./ChapterRampInput/fig/Fig_9.pdf} 
 \caption{ \label{fig:rele_tv_40dB_s1} The relative errors of the time-varying filter estimation converge slower than with the subspace method. 
@@
 
-The subspace method can estimate the affine input parameters from the sensor response using few parameters, the sensor order $n$, the sensor static gain $\mathbf{G}$, and the RLS forgetting factor $\lambda$.
-The subspace method does not necessarily requires optimization of $\lambda$ using a dataset of measured sensor responses.
-It is required to tune $\lambda$ online during the calibration of the system and later $\lambda$ remains fixed during the measurements.
+The subspace method can estimate the affine input parameters from the sensor response using few parameters, the sensor order $n$, the sensor static gain ${G}$, and the RLS forgetting factor $\omega$.
+The subspace method does not necessarily requires optimization of $\omega$ using a dataset of measured sensor responses.
+It is required to tune $\omega$ online during the calibration of the system and later $\omega$ remains fixed during the measurements.
 

### 2020-08-21 ChapterStatisticalAnalysis/ChapterStatisticalAnalysis.tex
Second draft before public defense

@@
 \emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I.
-Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Bias and covariance of the least-squares estimate in a structured errors-in-variables problem", Computational Statistics & Data Analysis journal, Vol.
+Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Bias and covariance of the least-squares estimate in a structured errors-in-variables problem", Computational Statistics \& Data Analysis journal, Vol.
 144, 2020, ISSN 0167-9473, doi:10.1016/j.csda.2019.106893.
@@
 The linear estimation problems with perturbation in both the regression matrix and the regressor are errors-in-variables (EIV) problems.
-It is required to calculate the statistical moments of the estimate to evaluate the properties of the solution of an EIV problem.
-The calculation of the estimate statistical moments demands a more considerable effort when the regression matrix of an EIV problem is structured, and when the regression matrix and the regressor are correlated.
+\color{blue} The formulation of the data-driven step input estimation problem is an EIV problem.
+This formulation gives a Hankel structure to the perturbed regression matrix, which is correlated with the regressor vector.
+The solution to the structured EIV problem is proposed by means of a recursive least squares algorithm to benefit from its simplicity for online implementations in devices with small computational resources.
+The properties of this least-squares (LS) approximation cannot be inferred from classical results\color{black}.
+\color{blue} The statistical analysis aims \color{black} to calculate the statistical moments of the \color{blue} least-squares \color{black} estimate to evaluate these properties.
+The statistical analysis of the LS estimate demands \color{blue} an additional effort \color{black} when the regression matrix is structured, and when it is correlated with the regressor, in a EIV problem.
 
-Structured total least-squares and structured total maximum likelihood are two examples of methods that solve structured EIV problems, but that require large computational resources.
-For online applications, where a real-time solution of the structured EIV problem is imperative, the recursive least-squares (RLS) method provides a suboptimal but computationally cheaper online solution.
+\color{blue} 
+The total least-squares methods reviewed in \citet{Markovsky07overview} are not appropriate to solve structured EIV problems.
+To overcome the challenges imposed by the structure, methods such as the structured total least-squares described in \citet{Lemmerling02} and the structured total maximum likelihood presented in \citet{Beck10} are proposed.
+Nevertheless, the structured total least-squares requires optimization and even the recursive alternatives such as the one developed by \citet{Rhode14recursive} are not suitable for online applications in devices with reduced computational resources.
+Moreover, current versions of the structured total maximum likelihood method consider uncorrelated cases only. 
+The instrumental-variables methods described in \citet{Soderstrom18} are other alternatives to implement recursive algorithms that solve EIV problems.
+The generalized instrumental-variables estimator is one example of them, that is mainly used in the identification of linear systems from observed input and output data.
+However, it is not straightforward to use the generalized instrumental-variables estimator to implement the direct estimation of the input from the observed output bypassing the model identification, like the data-driven input estimation method does.
+Thus, the recursive LS method is proposed in \citet{Markovsky15cep} to get a real-time approximation to the structured EIV problem solution.
+Even though this estimation is suboptimal, we benefit from its simplicity and suitability for implementation in devices with limited computational resources.
 
-Statistical analysis is indispensable to obtain the statistical moments of the RLS solution for a structured and correlated EIV problem, and to assess the estimation uncertainty.
-This chapter describes the methodology and the results of the statistical analysis.
-The mathematical expectation of the RLS estimate, approximated by a second-order Taylor series expansion, allows a quantification of the first and second-order moments of the LS estimate.
+As it is described in the works of \citet{Ferrero06} and \citet{Hessling13b}, and in the review of \citet{daSilva12}, the measurement results are meaningful only when the input estimation is provided together with an assessment of its uncertainty.
+For an unbiased estimator, the uncertainty is assessed from the estimation variance.
+On the other hand, for a biased estimator the mean-squared error is the main metric used to describe the uncertainty.
+Therefore, it is necessary to know the first two statistical moments of the estimation.
+
+This chapter describes a methodology to study the statistical moments of the LS solution for the structured EIV problem introduced in \citet{Markovsky15cep, Markovsky15ieee} for the data-driven step input estimation method.
+The described statistical analysis obtains the statistical moments of the LS solution, and they allow to assess the estimation uncertainty in terms of the mean-squared error.
 The statistical analysis results provide insight into the impact that the structure of the regression matrix and its correlation with the regressor has on the uncertainty of the RLS estimate.
 
+\color{black}
+
 \begin{comment}
-Errors-in-variables (EIV) are linear estimation problems in which the regression matrix and the regressor are perturbed \citep{VanHuffel91Book}, \citep{Markovsky07overview}.
+Errors-in-variables (EIV) are linear estimation problems in which the regression matrix and the regressor are perturbed \citet{VanHuffel91Book}, \citet{Markovsky07overview}.
 In structured EIV problems, the regression matrix has a given structure that depends on the problem formulation.
-Hankel, Toeplitz, or other application-specific matrices appear in problems of metrology \citep{Markovsky15cep}, system identification \citep{Soderstrom07}, image restoration \citep{Feiz17}, nuclear magnetic resonance spectroscopy \citep{Cai16}, direction-of-arrival estimation \citep{Pan18}, and time-of-arrival estimation \citep{Jia18}.
+Hankel, Toeplitz, or other application-specific matrices appear in problems of metrology \citet{Markovsky15cep}, system identification \citet{Soderstrom07}, image restoration \citet{Feiz17}, nuclear magnetic resonance spectroscopy \citet{Cai16}, direction-of-arrival estimation \citet{Pan18}, and time-of-arrival estimation \citet{Jia18}.
 
@@
 The only observed signal is the sensor output.
-To estimate a step input, the regression matrix, and the regressor are built from the step response observations \citep{Markovsky15cep}, and
+To estimate a step input, the regression matrix, and the regressor are built from the step response observations \citet{Markovsky15cep}, and
 the structure in the regression matrix is block-Hankel.
-This data-driven estimation methodology reduces the estimation time of the classical two-stage approach where a sensor model is first identified, and later the input is estimated using the sensor model \citep{Azam15, Niedzwiecki16a}.
+This data-driven estimation methodology reduces the estimation time of the classical two-stage approach where a sensor model is first identified, and later the input is estimated using the sensor model \citet{Azam15, Niedzwiecki16a}.
 %The problem is EIV because the perturbation of the sensor response gets into the regression matrix.
@@
 Total least-squares (TLS) is the typical estimator for unstructured EIV problems and is consistent when the perturbations have zero mean, and the covariance is a given positive definite matrix.
-When the perturbations are i.i.d. normally distributed, the solution of the TLS is equivalent to that of the maximum likelihood estimator (ML) \citep{Markovsky07overview}. 
-For structured EIV problems, the TLS estimator does not give general results since each specific structure requires a particular treatment \citep{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where finding the global optimum is not guaranteed \citep{Rhode14recursive}.
+When the perturbations are i.i.d. normally distributed, the solution of the TLS is equivalent to that of the maximum likelihood estimator (ML) \citet{Markovsky07overview}. 
+For structured EIV problems, the TLS estimator does not give general results since each specific structure requires a particular treatment \citet{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where finding the global optimum is not guaranteed \citet{Rhode14recursive}.
 Moreover, the computational complexity of TLS and ML inhibits real-time implementation to solve structured EIV problems.
@@
 Some of the reported works that propose LS estimators for structured EIV problems include 
-the design of a fast algorithm for matrices with small displacement rank \citep{Mastronardi07fast},
-the study of the estimator consistency \citep{Palanthandalam10parameter},
-the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models \citep{Kiviet12high} \citep{Kiviet14improved}, and
-a discussion of the causes of bias and inconsistency in homogeneous estimators \citep{Yeredor04homogeneous}.
+the design of a fast algorithm for matrices with small displacement rank \citet{Mastronardi07fast},
+the study of the estimator consistency \citet{Palanthandalam10parameter},
+the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models \citet{Kiviet12high} \citet{Kiviet14improved}, and
+a discussion of the causes of bias and inconsistency in homogeneous estimators \citet{Yeredor04homogeneous}.
 
@@
 The uncertainty of the reported LS estimators for structured EIV problems has not been addressed, and then, remains unknown.
-The estimator uncertainty is expressed using the estimation bias and covariance \citep{Pintelon12Book}.
+The estimator uncertainty is expressed using the estimation bias and covariance \citet{Pintelon12Book}.
 To know the LS estimator uncertainty, we quantified the bias and the covariance of the LS solution of EIV problems in the unstructured and structured cases. 
-This extends the perturbation analysis of the LS estimator of unstructured and uncorrelated problems that was investigated in \citep{Stewart90SPT} and in \citep{Vaccaro94}.
+This extends the perturbation analysis of the LS estimator of unstructured and uncorrelated problems that was investigated in \citet{Stewart90SPT} and in \citet{Vaccaro94}.
 This paper presents a study of the statistics of the LS estimate of unstructured and structured EIV problems. 
  We provide a discussion of the unstructured case as a reference, to get an insight of the impact that the structure of the regression matrix and the correlation between the regression matrix and the regressor perturbations have on the uncertainty of the LS estimate.
-The structured case is motivated by the metrology estimation problem \citep{Markovsky15cep}, where the EIV problem has a Hankel structure, and the perturbations are correlated.
+The structured case is motivated by the metrology estimation problem \citet{Markovsky15cep}, where the EIV problem has a Hankel structure, and the perturbations are correlated.
 The study of the metrology input estimate illustrates a methodology to conduct statistical analysis for any structured EIV problem.
@@
 For an overdetermined system of linear equations, constructed by the data-driven step input estimation method (\ref{eqn:min_seiv}), the least-squares (LS) solution is given by
-\begin{equation} \widehat{\bm{\theta}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
-where $\widetilde{\mathbf{K}}^\dagger$ is the pseudo-inverse of $\widetilde{\mathbf{K}}$.
+\ref{eqn:xhat}.
 The objective of the statistical analysis is to obtain the bias and the covariance of the solution $\widehat{\bm{\theta}}$ to the structured and correlated EIV problem (\ref{eqn:ddsiemnd}).
@@
 \begin{equation*} \widehat{\bm{\theta}} = \left( (\mathbf{K}+\mathbf{E})^\top (\mathbf{K}+\mathbf{E})  \right)^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}), \end{equation*} 
-which is equivalent to
+which is equivalent to \color{blue}
+\begin{equation*} \widehat{\bm{\theta}} = \left( \left( \mathbf{K}^\top \mathbf{K} \right)  \left( \mathbf{I} + \left( \mathbf{K}^\top \mathbf{K} \right)^{-1} \left( \mathbf{K}^\top \mathbf{E} + \mathbf{E}^\top \mathbf{K} + \mathbf{E}^\top \mathbf{E} \right) \right)  \right)^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}), \end{equation*}
+and \color{black}
 \begin{equation} \widehat{\bm{\theta}} = \left( \mathbf{I} + \mathbf{M} \right)^{-1} \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}), \label{eqn:1steq} \end{equation} 
@@
 \begin{equation} (\mathbf{I} + \mathbf{M})^{-1} \approx \mathbf{I} - \mathbf{M} + \mathbf{M}^2, \label{eqn:TseriesExp} \end{equation} 
-that is valid when the SNR is sufficiently high, and therefore $\mathbf{E}$ and $\mathbf{M}$ are small, satisfying the constraint on the spectral radius $\| \mathbf{M} \| < 1$. 
+that is valid when the SNR is sufficiently high, and therefore $\mathbf{E}$ and $\mathbf{M}$ are small, satisfying the constraint on the spectral radius $\color{blue} \rho(\mathbf{M}) = \color{black} \| \mathbf{M} \| < 1$.
 The neglected term in the Taylor series expansion is of the order $O(\Vert M \Vert^3)$.
-We can express the estimate as
+\color{blue} An a priori bound on the error introduced by the second order Taylor series expansion can be expressed in terms of the spectral radius $\rho(\mathbf{M})$.
+Considering that the Taylor series expansion is an infinite sum we have that
+\begin{equation} \begin{aligned} (\mathbf{I} + \mathbf{M})^{-1} &= \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \\
+&=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \right) \\
+&=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \end{aligned} \end{equation} 
+Therefore we can obtain a bound for the second error expansion error if we take a matrix norm as follows
+\begin{equation} \begin{aligned} \| (\mathbf{I} + \mathbf{M})^{-1} - ( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 ) \| &=  \| \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \| \\
+& \leq \| \mathbf{M}^3 \| \| (\mathbf{I} + \mathbf{M})^{-1} \| \leq \dfrac{\| \mathbf{M} \|^3}{ 1 -  \| \mathbf{M} \|} = \dfrac{\rho(\mathbf{M})^3}{1 - \rho(\mathbf{M})}  \end{aligned} \end{equation}
+
+Using the second order Taylor expansion of the inverse matrix\color{black}, we can express the estimate as
 \begin{equation} \widehat{\bm{\theta}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{Q}^{-1} (\mathbf{K}+\mathbf{E})^\top (\mathbf{y}+\bm{\epsilon}).
@@
 The covariance of the estimate $\widehat{\bm{\theta}}$ is obtained from
-\begin{equation} \begin{aligned} \mathbf{C} \left( \widehat{\bm{\theta}} \right) & = \mathbb{E} \left\{ \left( \widehat{\bm{\theta}} - \bm{\mu} \right) \left( \widehat{\bm{\theta}} - \bm{\mu} \right)^\top \right\} .
+\begin{equation} \begin{aligned} \mathbf{C} \left( \widehat{\bm{\theta}} \right) & = \mathbb{E} \left\{ \left( \widehat{\bm{\theta}} - \bm{\mu} \color{blue} \left(\widehat{\bm{\theta}} \right) \color{black} \right) \left( \widehat{\bm{\theta}} - \bm{\mu} \color{blue} \left(\widehat{\bm{\theta}} \right) \color{black} \right)^\top \right\} .
 \end{aligned} \label{eqn:covdef} \end{equation}
 
-The terms derived from (\ref{eqn:xhatexp}) that do not contribute to the bias and to the covariance are filtered out by the mathematical expectation operator considering the following general rules that are valid regardless of the existence of structure in the regression problem:
+The terms derived from (\ref{eqn:xhatexp}) that do not contribute to the bias and to the covariance are filtered out by the mathematical expectation operator, described in \citet{Papoulis02Book}, considering the following general rules that are valid regardless of the existence of structure in the regression problem:
 \begin{itemize}
@@
 Moreover, the second-order approximation disregards moments of order four and higher.
+
+\color{blue}The normality assumption of the perturbation noise is necessary to provide more prior information into the method formulation.
+The second order Taylor series expansion (\ref{eqn:xhatexp}) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
+Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
+If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance\color{black}.
 
@@
 normally distributed with zero mean, and variances $\sigma_{\mathbf{E}}^2$ and $\sigma_{\bm{\epsilon}}^2$, respectively.
-Therefore, the terms in the Taylor series expansion (\ref{eqn:xhatexp}) that contain products of $\mathbf{E}$ and $\bm{\epsilon}$ have zero expected value.
+Therefore, the terms in the Taylor series expansion that contain products of $\mathbf{E}$ and $\bm{\epsilon}$ have zero expected value.
 After removing the terms without contribution to the bias, and to the covariance, with the mathematical expectation operator,
 the analytic approximation of the bias (\ref{eqn:biasdef}) is given by 
-\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - T \right) \mathbf{Q}^{-1} \bm{\theta} , \end{aligned} \label{eqn:biasEu} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - \color{blue}N\color{black} \right) \mathbf{Q}^{-1} \bm{\theta} , \end{aligned} \label{eqn:biasEu} \end{equation}
 and the covariance (\ref{eqn:covdef}) is approximated by 
-\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \bm{\theta} \bm{\theta}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \ \mathbf{Q}^{-1} \bm{\theta} \bm{\theta}^\top \mathbf{Q}^{-1} , \end{aligned} \label{eqn:varEu} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\bm{\epsilon}}^2 \ \mathbf{Q}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \bm{\theta} \bm{\theta}^\top \right) \mathbf{Q}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - \color{blue}N\color{black} \right)^2 \ \mathbf{Q}^{-1} \bm{\theta} \bm{\theta}^\top \mathbf{Q}^{-1} , \end{aligned} \label{eqn:varEu} \end{equation}
 where the subscript $\mathrm{p}$ stands for prediction of the bias and covariance.
 The derivation of equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) is described in Appendix 1.
-We use the results described in \citep{Vaccaro94} \S 3 and \citep{Stewart90SPT} \S 2.1 for the expected values of products of unstructured matrices with perturbations.
+We use the results described in \citet{Vaccaro94} \S 3 and \citet{Stewart90SPT} \S 2.1 for the expected values of products of unstructured matrices with perturbations.
 Equations (\ref{eqn:biasEu}) and (\ref{eqn:varEu}) depend on the unobservable true values $\bm{\theta}$, $\mathbf{K}$, and on the variance of the perturbations.
@@
 We have then
-\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - T \right) \widetilde{\mathbf{Q}}^{-1} \widehat{\bm{\theta}} , \end{aligned} \label{eqn:biasEuST} \end{equation}
-\begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_\epsilon^2 \ \widetilde{\mathbf{Q}}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \right) \widetilde{\mathbf{Q}}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - T \right)^2 \widetilde{\mathbf{Q}}^{-1} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \widetilde{\mathbf{Q}}^{-1} .
+\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_{\mathbf{E}}^2 \left( 2 + 2n - \color{blue}N\color{black} \right) \widetilde{\mathbf{Q}}^{-1} \widehat{\bm{\theta}} , \end{aligned} \label{eqn:biasEuST} \end{equation}
+\begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \sigma_\epsilon^2 \ \widetilde{\mathbf{Q}}^{-1} + \sigma_{\mathbf{E}}^2 \ \mathrm{trace} \left( \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \right) \widetilde{\mathbf{Q}}^{-1} - \sigma_{\mathbf{E}}^4 \left( 2 + 2n - \color{blue}N\color{black} \right)^2 \widetilde{\mathbf{Q}}^{-1} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \widetilde{\mathbf{Q}}^{-1} .
 \end{aligned} \label{eqn:varEuST} \end{equation}
@@
 whereas, the estimation covariance (\ref{eqn:covdef}) is approximated by 
-\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right) , \end{aligned} \label{eqn:varE} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{\color{blue}N\color{black}-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right) , \end{aligned} \label{eqn:varE} \end{equation}
 where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \Big\}$, $\mathbf{P} = \mathbf{K} \mathbf{K}^\dagger$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{P}$.
@@
 \begin{lem} \label{lem:lemma1}
-Let $\mathbf{E} \in {\rm I\!R}^{\left(T-n\right) \times \left(n+1\right)}$ be the partitioned matrix 
-\[ \mathbf{E} = \begin{bmatrix} \mathbf{0}_{T-n \times 1} & \bm{\mathcal{H}}(\bm{\epsilon}) \mathbf{D}_{n+1 \times n}^{1,0} \end{bmatrix}, \]
- where $\bm{\mathcal{H}}(\bm{\epsilon}) \in {\rm I\!R}^{\left(T-n\right) \times \left(n+1\right)}$ is the block-Hankel matrix of $T-n$ rows and $n$ columns
-\[ \bm{\mathcal{H}}(\bm{\epsilon}) = \begin{bmatrix} \epsilon(0) & \epsilon(1) & \cdots & \epsilon(n) \\ \epsilon(1) & \epsilon(2) & \cdots & \epsilon(n+1) \\ \vdots & \vdots & & \vdots \\ \epsilon(T-n-1) & \epsilon(T-n) & \cdots & \epsilon(T-1) \end{bmatrix} , \]
+Let $\mathbf{E} \in \mathbb{R}^{\left(\color{blue}N\color{black}-n\right) \times \left(n+1\right)}$ be the partitioned matrix 
+\[ \mathbf{E} = \begin{bmatrix} \mathbf{0}_{\color{blue}N\color{black}-n \times 1} & \bm{\mathcal{H}}(\bm{\epsilon}) \mathbf{D}_{n+1 \times n}^{1,0} \end{bmatrix}, \]
+ where $\bm{\mathcal{H}}(\bm{\epsilon}) \in {\rm I\!R}^{\left(\color{blue}N\color{black}-n\right) \times \left(n+1\right)}$ is the block-Hankel matrix of $\color{blue}N\color{black}-n$ rows and $n$ columns
+\[ \bm{\mathcal{H}}(\bm{\epsilon}) = \begin{bmatrix} \epsilon(0) & \epsilon(1) & \cdots & \epsilon(n) \\ \epsilon(1) & \epsilon(2) & \cdots & \epsilon(n+1) \\ \vdots & \vdots & & \vdots \\ \epsilon(\color{blue}N\color{black}-n-1) & \epsilon(\color{blue}N\color{black}-n) & \cdots & \epsilon(\color{blue}N\color{black}-1) \end{bmatrix} , \]
 constructed from samples of the i.i.d. normally distributed random variable $\bm{\epsilon} \sim \mathcal{N}(0, \sigma_{\bm{\epsilon}}^2)$.
-Let $\mathbf{D}_{r \times c}^{1,k}$ and $\mathbf{D}_{r \times c}^{2,k}$ be the first and second-order finite differences matricial operators of dimensions $r \times c$ starting from the subdiagonal $k$, for example,
+Let $\mathbf{D}_{r \times c}^{1,d}$ and $\mathbf{D}_{r \times c}^{2,d}$ be the first and second-order finite differences matricial operators of dimensions $r \times c$ starting from the subdiagonal $d$, for example,
 \[ \mathbf{D}_{4 \times 3}^{1,-1} = \begin{bmatrix} 0 & 0 & 0 \\ -1 & 0 & 0 \\ 1 & -1 & 0 \\ 0 & 1 & -1 \end{bmatrix}, \ \mathbf{D}_{4 \times 3}^{2,0} = \begin{bmatrix}-1 & 0 & 0 \\ 2 & -1 & 0 \\ - 1 & 2 & -1 \\ 0 & -1 & 2 \end{bmatrix} .
@@
 \text{where} \ \mathbf{z}_{i1} = \mathbf{0}, \ \text{and} \ 
-z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{2, j-i} \end{bmatrix} \right), \\
-& \quad \text{for} \ i = 1, \cdots, T-n, \ \text{and} \ j = 2, \cdots, n+1. \\
+z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} \mathbf{0}_{\color{blue}N\color{black}-n} & \mathbf{D}_{\color{blue}N\color{black}-n \times n}^{2, j-i} \end{bmatrix} \right), \\
+& \quad \text{for} \ i = 1, \cdots, \color{blue}N\color{black}-n, \ \text{and} \ j = 2, \cdots, n+1. \\
 & \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \mathbf{E} \right\} = \mathbf{Z}, \ 
 \text{where} \ \mathbf{z}_{1j} = \mathbf{0}, \ \mathbf{z}_{i1} = \mathbf{0}, \ \text{and} \ 
-z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \ \mathbf{D}_{T-n \times T-n}^{2, j-i+1} \right) , \\
+z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \ \mathbf{D}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n}^{2, j-i+1} \right) , \\
 & \quad \text{for} \ i = 2, \cdots, n+1, \ \text{and} \ j=2, \cdots, n+1 . \\ 
@@
 \text{where} \ z_{ij} = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} 0 & \mathbf{0}_{n}^\top \\ \mathbf{0}_{n} & \mathbf{D}_{n \times n}^{2, j-i+1} \end{bmatrix} \right), \\
-& \quad \text{for} \ i = 1,\cdots,T-n, \ \text{and} \ j=1,\cdots,T-n. \\ 
+& \quad \text{for} \ i = 1,\cdots,\color{blue}N\color{black}-n, \ \text{and} \ j=1,\cdots,\color{blue}N\color{black}-n. \\ 
 & \mathbb{E} \left\{ \mathbf{E} \mathbf{A} \bm{\epsilon} \right\} = \mathbf{z}, \
-\text{where} \ z_i = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times n}^{1,n+1-i} \end{bmatrix} \right) , \\
-& \quad \text{for} \ i = 1,\cdots,T-n . \\
+\text{where} \ z_i = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \begin{bmatrix} \mathbf{0}_{\color{blue}N\color{black}-n} & \mathbf{D}_{\color{blue}N\color{black}-n \times n}^{1,n+1-i} \end{bmatrix} \right) , \\
+& \quad \text{for} \ i = 1,\cdots,\color{blue}N\color{black}-n . \\
 & \mathbb{E} \left\{ \mathbf{E}^\top \mathbf{A} \bm{\epsilon} \right\} = \mathbf{z}, \
-\text{where} \ z_1 = 0, \ \text{and} \ z_i = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \ \mathbf{D}_{T-n \times T-n}^{1, n+2-i} \right), \\
+\text{where} \ z_1 = 0, \ \text{and} \ z_i = \sigma_{\bm{\epsilon}}^2 \ \mathrm{Tr} \left( \mathbf{A} \ \mathbf{D}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n}^{1, n+2-i} \right), \\
 & \quad \text{for} \ i = 2,\cdots,n+1 . \\
 & \mathbb{E} \left\{ \mathbf{E} \mathbf{a} \bm{\epsilon}^\top \right\} = \mathbf{Z}, \
-\text{where each column} \ \mathbf{Z}_{j} = - \sigma_{\bm{\epsilon}}^2 \ \mathbf{D}_{T-n \times n+1}^{1,-j} \mathbf{R}_{n+1} \ \mathbf{a} , \\
-& \quad \text{for} \ j = 1,\cdots,T-n , \ \text{with} \ \mathbf{R}_{n+1} = \begin{bmatrix} & \mathbf{R}_n \\ 0 \end{bmatrix} , \\ & \quad
+\text{where each column} \ \mathbf{Z}_{j} = - \sigma_{\bm{\epsilon}}^2 \ \mathbf{D}_{\color{blue}N\color{black}-n \times n+1}^{1,-j} \mathbf{R}_{n+1} \ \mathbf{a} , \\
+& \quad \text{for} \ j = 1,\cdots,\color{blue}N\color{black}-n , \ \text{with} \ \mathbf{R}_{n+1} = \begin{bmatrix} & \mathbf{R}_n \\ 0 \end{bmatrix} , \\ & \quad
 \text{where} \ \mathbf{R}_n \ \text{is a reversal matrix.}
@@
 These operations mainly compute the trace of a product of the corresponding deterministic matrix $\mathbf{A}$, and a matrix constructed from the finite differences matricial operator, which can be of first order $\mathbf{D}^{1}$ or of second order $\mathbf{D}^{2}$.
-The total number of operations in the computation of the bias and the covariance is $O\left( T^3 + n^2 \right)$, but this order can be reduced since the $\mathbf{D}$ matrices are sparse.
+The total number of operations in the computation of the bias and the covariance is $O\left( \color{blue}N\color{black}^3 + n^2 \right)$, but this order can be reduced since the $\mathbf{D}$ matrices are sparse.
 
@@
 and
-\begin{equation} \begin{aligned} \mathbf{C}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \widetilde{\mathbf{K}}^{\dagger \top} - \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\widetilde{\mathrm{p}}}^\top \left( \widehat{\bm{\theta}} \right) , \end{aligned} \label{eqn:varST} \end{equation}
+\begin{equation} \begin{aligned} \mathbf{C}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{\color{blue}N\color{black}-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \widetilde{\mathbf{K}}^{\dagger \top} - \mathbf{b}_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\widetilde{\mathrm{p}}}^\top \left( \widehat{\bm{\theta}} \right) , \end{aligned} \label{eqn:varST} \end{equation}
 where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \bm{\epsilon}^\top \Big\}$, $\widetilde{\mathbf{C}} = \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}}$, $\widetilde{\mathbf{P}} = \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{P}}$.
@@
 
-The Cram\'er-Rao lower bound (CRLB) provides the lower limit on the variance of the estimate
-\begin{equation} \mathrm{CRLB}(\bm{\theta}) = \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\bm{\theta}}) }{\partial (\widehat{\bm{\theta}})} \right)^\top \mathbf{Fi}^{-1}(\bm{\theta}) \left( \mathbf{I} + \frac{\partial \mathbf{b} (\widehat{\bm{\theta}}) }{\partial (\widehat{\bm{\theta}})} \right), \label{eqn:CRLB} \end{equation}
-where $\mathbf{b} (\widehat{\bm{\theta}})$ is the bias of the estimate and $\mathbf{Fi}(\widehat{\bm{\theta}})$ is the Fisher information matrix \citep{Pintelon12Book}.
+The Cram\'er-Rao lower bound (CRLB) provides the lower limit on the estimation variance.
+For a biased estimator of the minimization problem (\ref{eqn:min_seiv}), the CRLB is given by
+\begin{equation} \mathrm{CRLB}_{\mathrm{b}}(\bm{\theta}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right)^\top \ \mathbf{Fi}^{-1}(\bm{\theta}) \ \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right), \label{eqn:CRLB_EIV} \end{equation}
+and for an unbiased estimator it is $\mathrm{CRLB}_{\mathrm{ub}}(\bm{\theta}) = \mathbf{Fi}^{-1}(\bm{\theta})$,
+where $\mathbf{b} (\widehat{\bm{\theta}})$ is the estimation bias and $\mathbf{Fi}(\widehat{\bm{\theta}})$ is the Fisher information matrix described in \citet{Pintelon12Book}.
+
 The Fisher information matrix is defined as the expected value of the Hessian of the negative likelihood function 
-\begin{equation} \mathbf{Fi}(x) = - \mathbb{E} \left\{ \frac{\partial ^2 l (\widehat{\bm{\theta}}) }{\partial \widehat{\bm{\theta}}^2 } \right\}, \end{equation}
+\begin{equation} \mathbf{Fi}(\color{blue}\theta\color{black}) = - \mathbb{E} \left\{ \frac{\partial ^2 l (\widehat{\bm{\theta}}) }{\partial \widehat{\bm{\theta}}^2 } \right\}, \end{equation}
 where the partial derivatives are evaluated in $\widehat{\bm{\theta}} = \bm{\theta}$.
 
-The minimization problem (\ref{eqn:min_seiv}) is a structured EIV problem that can be expressed as a linear in the measurements problem \citep{Pintelon12Book}
-\begin{equation} e (\widehat{\bm{\theta}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\bm{\theta}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\bm{\theta}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 .
-\label{eqn:M1vecyK} \end{equation}
-where $\widetilde{\mathbf{z}} = \mathbf{z} + \epsilon_{\mathbf{z}}$.
-The CRB requires that the true model $\mathbf{M}_1( \bm{\theta} ) \ \mathbf{z} = 0$ exists.
-Under the assumption of the measurement perturbation $\epsilon_{\mathbf{z}}$ being normally distributed with covariance matrix $\mathbf{C}_{\mathbf{z}}$, the loglikelihood function of the structured EIV problem is
+The minimization problem (\ref{eqn:min_seiv}) can be expressed as a linear in the measurements problem, as it is described in \citet{Pintelon12Book},
+\begin{equation} e (\widehat{\bm{\theta}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\bm{\theta}} ) \ \widetilde{\mathbf{z}} = \underbrace{ \begin{bmatrix} \mathbf{I}_{\color{blue}N\color{black}-n} & - \widehat{\bm{\theta}}^\top \otimes \mathbf{I}_{\color{blue}N\color{black}-n} \end{bmatrix} }_{\mathbf{M}_1( \widehat{\bm{\theta}} )} \underbrace{ \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} }_{\widetilde{\mathbf{z}}} = 0 , \label{eqn:M1vecyK} \end{equation}
+where $\widetilde{\mathbf{z}} = \mathbf{z} + \bm{\epsilon}_{\mathbf{z}}$ is a vector constructed by stacking the perturbation noise of the regressor and the regression matrix, and the true model must satisfy the equality $\mathbf{M}_1( \bm{\theta} ) \ \mathbf{z} = 0$.
+Since the true measurements $\mathbf{z}$ are unknown, they have to be estimated and are parameterized as $\widehat{\mathbf{z}}$.
+Under the assumption of the measurement perturbation $\bm{\epsilon}_{\mathbf{z}}$ being normally distributed with zero mean and covariance matrix 
+\begin{equation} \mathbf{C}_{\mathbf{z}} = \sigma_{\bm{\epsilon}}^2 
+\begin{bmatrix} \mathbf{I}_{\color{blue}N\color{black}-n} & \mathbf{0}_{\color{blue}N\color{black}-n} & \mathbf{D}^{1,n}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} & \mathbf{D}^{1,n-1}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} & \cdots & \mathbf{D}^{1,1}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} \\ \mathbf{0}_{\color{blue}N\color{black}-n} & \mathbf{0}_{\color{blue}N\color{black}-n} & \mathbf{0}_{\color{blue}N\color{black}-n} & \mathbf{0}_{\color{blue}N\color{black}-n} & \cdots & \mathbf{0}_{\color{blue}N\color{black}-n} \\ \left( \mathbf{D}^{1,n}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} \right)^\top & \mathbf{0}_{\color{blue}N\color{black}-n} & \mathbf{D}^{2,1}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} & \mathbf{D}^{2,0}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} & \cdots & \mathbf{D}^{2,2-n}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} \\ \left( \mathbf{D}^{1,n-1}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} \right)^\top & \mathbf{0}_{\color{blue}N\color{black}-n} & \mathbf{D}^{2,2}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} & \mathbf{D}^{2,1}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} & \cdots & \mathbf{D}^{2,3-n}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} \\ \vdots & \vdots & \vdots & \vdots & & \vdots \\ \left( \mathbf{D}^{1,1}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} \right)^\top & \mathbf{0}_{\color{blue}N\color{black}-n} & \mathbf{D}^{2,n}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} & \mathbf{D}^{2,n-1}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} & \cdots & \mathbf{D}^{2,1}_{\color{blue}N\color{black}-n \times \color{blue}N\color{black}-n} \end{bmatrix} \label{eqn:Cz} \end{equation}
+the loglikelihood function of the structured EIV problem is
 \begin{equation} \ln{ l(\widetilde{\mathbf{z}}, \widehat{\mathbf{z}}, \widehat{\bm{\theta}}) } = - \frac{1}{2} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right)^\top \mathbf{C}_{\mathbf{z}}^{-1} \left( \widetilde{\mathbf{z}} - \widehat{\mathbf{z}} \right) + \mathrm{constant}, \end{equation}
-where $\widehat{\mathbf{z}}$ are parameters of the measurements $\widetilde{\mathbf{z}}$ that have to be estimated and satisfy $\mathbf{M}_1( \widehat{\bm{\theta}} ) \ \widehat{\mathbf{z}} = 0$.
-The size of the Fisher information matrix $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
-Moreover, in Chapter 19 of \citep{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\bm{\theta})$ can be obtained from $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ after doing inversion by parts, giving
-\begin{equation} \mathbf{Fi}(\bm{\theta}) = \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)^\top \left( \mathbf{M}_1( \bm{\theta} ) \ \mathbf{C}_{\mathbf{z}} \ \mathbf{M}_1^\top( \bm{\theta} ) \right)^{-1} \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)
- , \label{eqn:FIM} \end{equation} 
-where the partial derivatives are evaluated at the true values $\bm{\theta}$, and the covariance matrix of the measurements is
-\begin{equation} \mathbf{C}_{\mathbf{z}} = \sigma_{\epsilon}^2 \begin{bmatrix} \mathbf{I}_{T-n} & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{1,n} & \mathbf{D}_{T-n \times T-n}^{1,n-1} & \cdots & \mathbf{D}_{T-n \times T-n}^{1,1} \\ \mathbf{0}_{T-n} & \mathbf{0}_{T-n} & \mathbf{0}_{T-n} & \mathbf{0}_{T-n} & \cdots & \mathbf{0}_{T-n} \\ \left( \mathbf{D}_{T-n \times T-n}^{1,n} \right)^\top & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{2,1} & \mathbf{D}_{T-n \times T-n}^{2,0} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,2-n} \\ \left( \mathbf{D}_{T-n \times T-n}^{1,n-1} \right)^\top & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{2,2} & \mathbf{D}_{T-n \times T-n}^{2,1} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,3-n} \\ \vdots & \vdots & \vdots & \vdots & & \vdots \\ \left( \mathbf{D}_{T-n \times T-n}^{1,1} \right)^\top & \mathbf{0}_{T-n} & \mathbf{D}_{T-n \times T-n}^{2,n} & \mathbf{D}_{T-n \times T-n}^{2,n-1} & \cdots & \mathbf{D}_{T-n \times T-n}^{2,1} \end{bmatrix} .
- \label{eqn:Cz} \end{equation}
 
-The Cram\'er-Rao lower bound for an biased estimator of the minimization problem (\ref{eqn:min_seiv}) is given by
-\begin{equation} \mathrm{CRB}_{\mathrm{b}}(\bm{\theta}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right)^\top \ \mathbf{Fi}^{-1}(\bm{\theta}) \ \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right), \label{eqn:CRB_EIV} \end{equation}
-and for an unbiased estimator it is $\mathrm{CRB}_{\mathrm{ub}}(\bm{\theta}) = \mathbf{Fi}^{-1}(\bm{\theta})$.
+The size of the Fisher information matrix $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ depends on the number of to-be-estimated parameters $\widehat{\mathbf{z}}$, that grows with the sample size.
+In Chapter 19 of \citet{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\bm{\theta})$ that depends only on $\bm{\theta}$ can be obtained after doing inversion by parts, resulting in
+\begin{equation} \mathbf{Fi}(\bm{\theta}) = \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)^\top \left( \mathbf{M}_1( \bm{\theta} ) \ \mathbf{C}_{\mathbf{z}} \ \mathbf{M}_1^\top( \bm{\theta} ) \right)^{-1} \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right) .
+\label{eqn:FIM} \end{equation}
+
+\color{blue}
+\subsection{Instrumental variables formulation of the data-driven step input estimation method}
+
+The errors-in-variables problem formulation (\ref{eqn:min_seiv}) of the data-driven step input estimation method can be converted into an instrumental-variables problem.
+A guide to do this conversion is found in \citet{Soderstrom18} and in \citet{Pintelon12Book}. 
+The instrumental variable $\widetilde{\mathbf{Z}}$ has to be selected so that the normal equations (\ref{eqn:neq_seiv}) are expresed instead as
+\begin{equation} \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{y}} . \label{eqn:neq_siv} \end{equation}
+The selection of the instrumental variable has to be done in such a way that is correlated with the regression matrix $\widetilde{\mathbf{K}}$, and uncorrelated with the perturbations $\bm{\epsilon}$.
+Considering the block-Hankel matrix $\widetilde{\mathbf{K}}$ (\ref{eqn:matrixK}) with $N=2n+1$,
+\begin{equation} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(n+1) & \Delta \widetilde{y}(n+2) & \cdots & \Delta \widetilde{y}(2n) \end{bmatrix} , \label{eqn:matrixK_r} \end{equation}
+then, the instrumental variable $\widetilde{\mathbf{Z}}$ should not contain the elements that construct $\widetilde{\mathbf{K}}$, i.e., the elements of the vector $\left( \widetilde{y}(1), \ldots, \widetilde{y}(2n+1) \right)^\top$.
+A possibility to generalize this ideas is expressing the instrumental variable as
+\begin{equation} \widetilde{\mathbf{Z}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-4n) & \Delta \widetilde{y}(t-4n+1) & \cdots & \Delta \widetilde{y}(t-3n-1) \\ G & \Delta \widetilde{y}(t-4n+1) & \Delta \widetilde{y}(t-4n+2) & \cdots & \Delta \widetilde{y}(t-3n) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-3n) & \Delta \widetilde{y}(t-3n+1) & \cdots & \Delta \widetilde{y}(t-2n-1) \end{bmatrix} , \label{eqn:matrixZ_t} \end{equation}
+for a regression matrix
+\begin{equation} \widetilde{\mathbf{K}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-2n+1) & \Delta \widetilde{y}(t-2n+2) & \cdots & \Delta \widetilde{y}(t-n) \\ G & \Delta \widetilde{y}(t-2n+2) & \Delta \widetilde{y}(t-2n+3) & \cdots & \Delta \widetilde{y}(t-n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-n+1) & \Delta \widetilde{y}(t-n+2) & \cdots & \Delta \widetilde{y}(t) \end{bmatrix} , \label{eqn:matrixK_t} \end{equation}
+for $t = 4n+1, \ldots, N$.
+
+The difference with the EIV formulation is that this implementation of the IV method requires to wait more, until $4n+1$ samples of the sensor response are read, instead of $2n+1$ samples. 
+For sensors of small dynamic order this twice-the-time waiting does not represent a big change, but as the sensor order increases, the need for waiting can turn the method into impractical since a lot of time should pass before starting to estimate the input.
+
+\subsubsection{Recursive generalized instrumental variables method}
+
+There exists a generalized instrumental variables method (GIVE) that was proposed to identify dynamic models from input/output data. 
+GIVE is a bias-eliminating method that uses a generalized instrumental vector constructed from delayed values of the input and output signals \cite{Soderstrom18}.
+Since the aim of this thesis is to analyze a method that directly estimates the input from the observed sensor output, without identifying explicitly a dynamic model of the sensor, the GIVE method is out of the scope of this work.
+Future research can include studying the formulation of a data-driven input estimation method in the spirit of the GIVE method. 
+Nevertheless, a brief description of the recursive GIVE method is given in the following paragraphs.
+
+\subsubsection{Generalized instrumental variables method}
+For an errors-in-variables problem expressed as
+\begin{equation} \widetilde{y}(t) = \widetilde{\bm{\phi}}(t) \bm{\theta}   \end{equation}
+where $\widetilde{y}(t)$ is the output perturbed by measurement noise, $\bm{\theta}$ is the vector of to-be-estimated parameters, and $\widetilde{\bm{\phi}}(t)$ is the regressor vector constructed from delayed samples of the measured input $\widetilde{u}(t)$ and output $\widetilde{y}(t)$.
+The least-squares estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is the solution of the normal equation
+\begin{equation} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} \widehat{\bm{\theta}}_{\mathrm{LS}} = \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} , \label{eqn:norm} \end{equation}
+where $\widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}}$ is the sample covariance matrix of the regressor vector, and $\widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}}$ is the sample cross-covariance matrix of the regressor vector and the output.
+When the number of samples becomes infinitely large, these covariances matrices can be expressed as a sum of the contributions from exact data and measurement noise contributions
+\begin{equation} \begin{aligned} \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} &= \mathbb{E} \{ \widetilde{\bm{\phi}} \widetilde{\bm{\phi}}^\top \} = \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \\ \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}}  &=  \mathbb{E} \{ \widetilde{\bm{\phi}} \widetilde{y}^\top \} = \mathbf{r}_{\bm{\phi} y} + \mathbf{r}_{\bm{\eta} \bm{\epsilon}} , \end{aligned} \end{equation}
+where we are assuming that $\widetilde{y} =  y + \bm{\epsilon}$, $\widetilde{\bm{\phi}} = \bm{\phi} + \bm{\eta}$, and $\mathbf{r}_{\bm{\phi} y} = \mathbf{R}_{{\bm{\phi}}} \bm{\theta}$.
+The bias of the least-square solution can be expressed as 
+\begin{equation} \mathbf{R}_{\widetilde{\bm{\phi}}} \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) = \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{R}_{\widetilde{\bm{\phi}}} \bm{\theta} = \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} - \mathbf{R}_{\bm{\eta}} \bm{\theta} , \end{equation}
+
+If we assume that the measurement noise is white for both the input and output, 
+\begin{equation}  \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} = \mathbb{E}\{ \bm{\eta} \bm{\epsilon}^\top \} = \mathbf{0} \quad \mathrm{and} \quad \mathbf{R}_{\bm{\eta}} = \begin{bmatrix} \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{n_y} & \mathbf{0} \\ \mathbf{0} & \sigma_u^2 \mathbf{I}_{n_u} \end{bmatrix}, \end{equation}
+ where $\sigma_{\bm{\epsilon}}^2$ and $\sigma_u^2$ are the variances of the output and the input, respectively. The normal equation (\ref{eqn:norm}) implies that 
+\begin{equation} \left( \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \right)  \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) \neq \mathbf{0}, \end{equation}
+and the estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is not consistent.
+Instead, a bias-compensating least squares estimation can be obtained with  
+\begin{equation} \widehat{\bm{\theta}}_{\mathrm{BCLS}} = \left( \mathbf{R}_{\widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\eta}} \right)^{-1}  \left(  \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{r}_{\bm{\eta} \bm{\epsilon}} \right) .
+\end{equation}
+
+To formulate the generalized instrumental variable estimator, a generalized instrumental vector $\mathbf{z}(t)$ is built from delayed samples of the input and the output 
+\begin{equation*} \mathbf{z}(t) = \left( y(t), y(t-1), \ldots, y(t-n_y), u(t), u(t-1), \ldots, u(t-n_u) \right)^{\top} . \end{equation*}
+The dimension of the generalized instrumental vector should be $n_{\mathbf{z}} \geq \mathrm{dim} ( \vartheta ) = n_u + n_y + 2$, because the total parameter vector $\vartheta = \left( \theta^\top, \rho^\top \right)^\top$ considers also the variances of the measurement noise $\rho = \left( \sigma_{\bm{\epsilon}}^2, \sigma_u^2 \right)^\top$.
+Since the generalized instrumental vector $\mathbf{z}(t)$ and the equation error $\widetilde{y}(t) - \widetilde{\bm{\phi}}(t) \bm{\theta}$ must be correlated, we can express the system of equations as
+\begin{equation} \left( \mathbf{R}_{\widetilde{\mathbf{z}} \widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \right) \bm{\theta} = \left(  \mathbf{r}_{\widetilde{\mathbf{z}} \widetilde{y}} - \mathbf{r}_{\bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) \right) , \label{eqn:GIVE} \end{equation}
+where $\widetilde{\mathbf{z}} = \mathbf{z} + \bm{\zeta}$.
+
+\subsubsection{Recursive generalized instrumental variables method}
+If we define
+\begin{equation} \begin{aligned} \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right)  &=  \mathbf{r}_{\bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \bm{\theta}  =  \mathbf{J} \left( \bm{\theta} \right) \bm{\rho}, \\
+\mathbf{R} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau=1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{\bm{\phi}}^\top \left( \tau \right) , \\
+\mathbf{q} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau=1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{y} \left( \tau \right) \\
+\mathbf{b} \left( t \right)  &=  \mathbf{q} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right),
+ \end{aligned} \end{equation}
+then the equation (\ref{eqn:GIVE}) is equivalent to
+\begin{equation} \mathbf{R} \left( t \right) \bm{\theta} = \mathbf{b} \left( t \right) . \end{equation}
+This representation allows a recursive form
+\begin{equation} \begin{aligned} 
+\mathbf{R} \left( t+1 \right)  &= \mathbf{R} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{\bm{\phi}}^\top \left( t+1 \right) - \mathbf{R} \left( t \right) \right), \\
+\mathbf{b} \left( t+1 \right)  &= \mathbf{b} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{b} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right) \right), \\
+\widehat{\bm{\theta}} \left( t+1 \right)  &=  \left( \mathbf{R}^\top \left( t+1 \right) \mathbf{R} \left( t+1 \right) \right)^{-1} \mathbf{R}^\top \left( t+1 \right) \mathbf{b} \left( t+1 \right) \\
+\mathbf{q} \left( t+1 \right)  &= \mathbf{q} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{q} \left( t \right) \right) \\
+\widehat{\bm{\rho}} \left( t+1 \right)  &=  \mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right) \left( \mathbf{q} \left( t \right) - \mathbf{R} \left( t \right) \widehat{\bm{\theta}} \left( t \right) \right) ,
+ \end{aligned} \end{equation}
+where $\mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right)$ is the pseudo-inverse of $\mathbf{J} \left( \bm{\theta} \right)$.
+
+The complexity of this recursive generalized instrumental variables estimator is ${O}(n^3)$, which is larger than the ${O}(n^2)$ of the recursive least squares.
+Another drawback of the recursive generalized IV estimator is that it computes a pseudo-inverse matrix in each iteration step.
+If we would need to formulate the data-driven step input estimation problem in the generalized IV framework, it would require to study the formulation in detail because currently the IV framework is based on the assumption that both measurements of the input and output are available, but for the step input estimation problem only the output is available.
+Therefore, it is not straightforward to implement the recursive generalized IV estimator to estimate the step input, and this can lead to future research.
+\color{black}
 
@@
 Two Monte Carlo (MC) simulation studies were conducted to test the obtained bias and variance formulas.
-One MC simulation was devoted to the unstructured EIV problem and the other to the structured EIV problem that corresponds to the step input estimation method (\ref{eqn:min_seiv}).
-The MC simulations performed $N_{MC} = 10^6$ runs of the LS estimator with different realizations of the perturbation $[\begin{matrix} \mathbf{E} & \bm{\epsilon}] \end{matrix}$.
-In the structured EIV problem the perturbation $\mathbf{E}$ is correlated with $\bm{\epsilon}$.
-The perturbations variance was selected to have a signal-to-noise ratio (SNR) in the interval [30 dB, 80 dB], according to 
-\begin{equation} \mathrm{SNR} = 20 \log_{10}{ \dfrac{ \sqrt{ \dfrac{1}{T} \sum\limits_{t=1}^{T}{ y(t)^2 } } }{ \sigma_{ \bm{\epsilon}}} } . \end{equation} 
-For high enough SNR, the constraint $\| \mathbf{M} \| < 1$ is valid and the Taylor series expansion  (\ref{eqn:TseriesExp}) holds. 
-In this case, the derived expressions predict the LS estimation bias and variance.
+\color{blue} There are two sets of formulas.
+One set was obtained for the unstructured and uncorrelated EIV problem, and, the other set, for the structured and correlated EIV problem. 
+The problem solved by the data-driven step input estimation method (\ref{eqn:min_seiv}) belongs to the second set.
+The objective of the MC simulations is to compare the results of the bias and variance formulas in these two cases. 
+The comparison is an evaluation of the effect that the structure and correlation have in the problem solution.
+To make a fair comparison, the same exact data and the perturbation variance is used in both types of EIV problems.
+
+The exact data is the system of linear equations (\ref{eqn:ddsiemexd}) built by the data-driven step input estimation method.
+To build the system of equations, the processed signal ${y}$ is the exact response of a sensor to the input $0.1s(t)$, where $s(t)$ is the unit step response.  
+The sensor is modeled as a linear time-invariant mass-spring-damper system of order $n = 2$, that is given by 
+\begin{equation} 
+\mathbf{x}(k+1) = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{m+M} & \dfrac{-k_{\mathrm{d}}}{m+M} \end{bmatrix} \mathbf{x}(k) + \begin{bmatrix} 0\\ -g\end{bmatrix} u(k),  
+\quad y(k) = \begin{bmatrix} -1 & 0 \end{bmatrix} \mathbf{x}(k), \label{eqn:msdst}
+\end{equation}
+where $m=0.015$ kg, $k_{\mathrm{s}}=500$ kg/m, $k_{\mathrm{d}}=0.5$ kg s/m, $M=0.1$ kg, and $g=9.81 \ \mathrm{m/s^2}$.   
+The exact sensor response is shown in Figure \ref{fig:y}.
+
+In the simulations, the sample size considered is $N=200$, since there the transient regime of the sensor model is still evident. 
+In the case of the structured EIV problem, this sample size satisfies the requirements of the data-driven step input estimation method.
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_1.pdf} 
+\caption{ \label{fig:y} \color{blue} The exact response of the model is used in the simulations of the two EIV problem types.
+For the structured EIV problem, it is important to consider the transient response.
+Thus, we use the first 200 samples.
+The dashed lines show the values of ${y}$ that are 2\% above and below the steady-regime output value\color{black}.
+}
+\end{figure}
+
+\color{black}
+
+%In the structured EIV problem the perturbation $\mathbf{E}$ is correlated with $\bm{\epsilon}$.
+The MC simulations performed $N_{MC} = 10^6$ runs of the LS solver with different realizations of the additive perturbations $[\begin{matrix} \mathbf{E} & \bm{\epsilon}] \end{matrix}$.
+The perturbation variance \color{blue} $\sigma_{\bm{\epsilon}}^2$ \color{black} was selected to have a signal-to-noise ratio (SNR) in the interval [30 dB, 80 dB], according to 
+\begin{equation} \mathrm{SNR} = 20 \log_{10}{ \dfrac{ \sqrt{ \dfrac{1}{\color{blue}N\color{black}} \sum\limits_{t=1}^{\color{blue}N\color{black}}{ y(t)^2 } } }{ \sigma_{ \bm{\epsilon}}} } .
+\label{eqn:SNR} \end{equation}
+
+The constraint $\| \mathbf{M} \| < 1$ and the Taylor series expansion  (\ref{eqn:TseriesExp}) hold for SNR sufficiently high, and 
+in this conditions, the derived expressions \color{blue} predict \color{black} the LS estimation bias and variance.
 Thus, it is relevant to monitor the evolution of the largest $\mathbf{M}$ eigenvalue to detect the lower limit of the SNR  that allows the validity of the predictions.
@@
 The MC simulation of the unstructured EIV problem solution was conducted using the following settings.
-The exact data elements in the matrix $\mathbf{K} \in {\rm I\!R}^{\left(T-n\right) \times \left(n+1\right)}$  are normally distributed with zero mean and variance one, for $T = 200$ and $n = 2$.
-The to-be-estimated exact data is the vector $x = \begin{bmatrix} 1 & 2 & 3 \end{bmatrix}^\top$.
-The perturbations $\mathbf{E}$ and $\bm{\epsilon}$ are normally distributed with zero mean and variances subject to $\sigma_{\mathbf{E}}^2 = 2 \sigma_{\bm{\epsilon}}^2$.
-These settings are similar to those of the structured case, aiming to have comparable situations.
+The perturbations $\mathbf{E}$ and $\bm{\epsilon}$ are normally distributed with zero mean and variances $\sigma_{\mathbf{E}}^2 = 2 \sigma_{\bm{\epsilon}}^2$.
+\color{blue} 
+This relationship between the variances is also present in the structured EIV problem, and then we have similar assumptions.
+However, for the uncorrelated and unstructured EIV case, the perturbations $\mathbf{E}$ and $\bm{\epsilon}$ are independent.\color{black}
+
 %In the following results we focus our interest in the first element of the estimate $\widehat{\bm{\theta}}$.
@@
 \end{equation}
-The standard deviation of the estimate $\widehat{\bm{\theta}}_{\left[1\right]}$ is used to obtain the standard error $\sigma_\mathrm{e}$ of the MC simulation, which decreases with respect to the square root of the number of MC runs $N_{\mathrm{MC}}$ \citep{Hammersley75}:
-\begin{equation} \sigma_\mathrm{e} = \frac{ \sigma \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) }{\sqrt{N_{\mathrm{MC}}}}, \quad \mathrm{where} \quad \sigma^2 \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) = \frac{1}{N_{\mathrm{MC}}-1} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( \left( \widehat{\bm{\theta}}_{\left[1\right]} \right)_i - \mu \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) \right)^2 } .
+The standard deviation of the estimate $\widehat{\bm{\theta}}_{\left[1\right]}$ is used to obtain the standard error $\sigma_\mathrm{e,err}$ of the MC simulation, which decreases with respect to the square root of the number of MC runs $N_{\mathrm{MC}}$ as it is explained in \citet{Hammersley75}:
+\begin{equation} \sigma_\mathrm{e,err} = \frac{ \sigma_\mathrm{e} \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) }{\sqrt{N_{\mathrm{MC}}}}, \quad \mathrm{where} \quad \sigma_\mathrm{e}^2 \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) = \frac{1}{N_{\mathrm{MC}}-1} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( \left( \widehat{\bm{\theta}}_{\left[1\right]} \right)_i - \mu \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) \right)^2 } .
 \label{eqn:stderr} \end{equation}
@@
 The standard error of the bias prediction $b_{\widetilde{\mathrm{p}}}$ is given by
-\begin{equation} \sigma_{\widetilde{\mathrm{p}}} = \sqrt{ \frac{1}{N_{\mathrm{MC}} \left( N_{\mathrm{MC}}-1 \right)} \sum_{i=1}^{N_{\mathrm{MC}}} { \left( \left( \mathbf{b}_{\widetilde{\mathrm{p}} \left[1\right]} \right)_i - b_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) \right)^2 } } .
+\begin{equation} \sigma_{\widetilde{\mathrm{p}}, \mathrm{err}} = \sqrt{ \frac{1}{N_{\mathrm{MC}} \left( N_{\mathrm{MC}}-1 \right)} \sum_{i=1}^{N_{\mathrm{MC}}} { \left( \left( \mathbf{b}_{\widetilde{\mathrm{p}} \left[1\right]} \right)_i - b_{\widetilde{\mathrm{p}}} \left( \widehat{\bm{\theta}}_{\left[1\right]} \right) \right)^2 } } .
 \end{equation}
@@
 
-Figure \ref{bias_sigma_NMC_unstr_str_n2} shows the empirical bias and the bias predictions, with their corresponding standard errors, for the unstructured (plots in the top) and structured (plots in the bottom) EIV problems.
+Figure \ref{bias_sigma_NMC_unstr_str_n2} shows the empirical bias and the bias predictions, with their corresponding standard errors, \color{blue} for the unstructured EIV problem\color{black}.
 The simulation settings for the structured and correlated EIV problem are described in the following subsection.
-In the figure, it can be observed that the empirical bias $b_\mathrm{e}$ is proportional to the perturbation noise variance while the standard error $\sigma_\mathrm{e}$ is proportional to the perturbation noise standard deviation.
-The bias predictions $b_\mathrm{p}$ and $b_{\widetilde{\mathrm{p}}}$ are accurate since they coincide with the empirical bias $b_\mathrm{e}$ in all the SNR interval for unstructured EIV problems, 
-In both unstructured and structured EIV cases, the standard errors of the MC estimates $\sigma_{\mathrm{e}}$ and $\sigma_{\widetilde{\mathrm{p}}}$ are smaller than the bias estimates $b_{\mathrm{e}}$ and $b_{\widetilde{\mathrm{p}}}$.
+In the figure, it can be observed that the empirical bias $b_\mathrm{e}$ is proportional to the perturbation noise variance while the standard error $\sigma_\mathrm{e, err}$ is proportional to the perturbation noise standard deviation.
+The bias predictions $b_\mathrm{p}$ and $b_{\widetilde{\mathrm{p}}}$ are accurate since they coincide with the empirical bias $b_\mathrm{e}$ in all the SNR interval for unstructured EIV problems. 
+In both unstructured and structured EIV cases, the standard errors of the MC estimates $\sigma_{\mathrm{e, err}}$ and $\sigma_{\widetilde{\mathrm{p}}, \mathrm{err}}$ are smaller than the bias estimates $b_{\mathrm{e}}$ and $b_{\widetilde{\mathrm{p}}}$.
 The bias estimates are spread near their sample means. 
@@
  \centering
- \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_1.pdf} 
+ \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_2.pdf} 
   \caption{ \label{bias_sigma_NMC_unstr_str_n2} 
-  We observe the results of the LS solutions for the unstructured (top) and the structured (bottom) EIV problems.
-These results are the empirical bias $b_{\mathrm{e}}$, the predicted bias from exact data $b_{\mathrm{p}}$, the predicted bias from observed data $b_{\widetilde{\mathrm{p}}}$, the empirical standard error $\sigma_{\mathrm{e}}$, and the standard error from the estimations using observed data $\sigma_{\widetilde{\mathrm{p}}}$.
+  We observe the results of the LS solutions \color{blue} for the unstructured EIV problem\color{black}.
+These results are the empirical bias $b_{\mathrm{e}}$, the predicted bias from exact data $b_{\mathrm{p}}$, the predicted bias from observed data $b_{\widetilde{\mathrm{p}}}$, the empirical standard error $\sigma_{\mathrm{e, err}}$, and the standard error from the estimations using observed data $\sigma_{\widetilde{\mathrm{p}}, \mathrm{err}}$.
 The estimation biases are proportional to the perturbation variance and the estimation standard errors are proportional to the perturbation standard deviation.
@@
   \centering
-  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_2.pdf} 
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_3.pdf} 
   \caption{ \label{fig:b_bt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (top) between the predicted bias and the empirical bias are proportional to the perturbation noise variance as it is expected, and the relative errors (bottom) are smaller than 5\% for SNR below 70 dB.
@@
   \centering
-  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_3.pdf} 
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_4.pdf} 
   \caption{ \label{fig:v_vt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (top) between the predicted variances and the empirical variance are proportional to the perturbation noise variance, and the relative errors (bottom) are smaller than 5\% for SNR higher that 40 dB.
@@
   \centering
-  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_4.pdf} 
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_5.pdf} 
   \caption{ \label{fig:bv_btvt_abse_rele_unstr_e7} The MC simulation shows that when we solve an unstructured EIV problem with LS, the absolute errors (top) between the prediction with observed data and the prediction with exact data are proportional to the perturbation noise variance.
@@
 
-The MC simulation of the structured EIV problem (\ref{eqn:ddsiemnd}) solution was conducted processing $T = 200$ samples of the transient response $\widetilde{\mathbf{y}}$ generated by a linear time invariant system of order $n = 2$, after a step input excitation with $u = 1 \ \mathrm{units}$, where the units represent any physical quantity.
-The processed step response is shown in Figure \ref{fig:y}.
-The steady state response of the system is reached after 400 samples because from there on the relative error between the instantaneous values of the transient response and the steady-state response value is smaller than 2\%.
-Processing 200 samples ensures that the step input estimation is computed from transient data only.
-%In the following results we focus our interest in the first element of the estimate $\widehat{\bm{\theta}}_{\left[1\right]}$, where the step input level estimate $\widehat{u}$ is located.
+\color{blue}The MC simulation of the structured EIV problem (\ref{eqn:ddsiemnd}) solution was conducted using the step response $\widetilde{{y}}$ of the sensor model (\ref{eqn:msdst}).
+We use the same exact data response as for the previous subsection, but in this case the perturbations of the regression matrix $\widetilde{\mathbf{K}}$ have Hankel structure and are correlated with the perturbations of the regressor vector $\widetilde{\mathbf{y}}$.
+The perturbations $\bm{\epsilon}$ are normally distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$, and are also used to build the perturbations $\mathbf{E}$.
+Due to the Hankel structure in Equation (\ref{eqn:matrixE}), the variances of the perturbations $\mathbf{E}$ and $\bm{\epsilon}$ are related by $\sigma_{\mathbf{E}}^2 \approx 2 \sigma_{\bm{\epsilon}}^2$.
+
+The spectral radius of the matrix $\mathbf{K}$ in Equation (\ref{eqn:TseriesExp}) is observed for the SNR values in the interval of interest.
+As can be observed in Figure \ref{fig:spectral_radius}, for SNR values larger than 35 dB, the perturbation levels make the spectral radius $\| \mathbf{M} \| < 1$, and, therefore, ensure the approximation given by the Taylor series expansion of the inverse matrix.
+Below an SNR of 35 dB, the spectral radius of $\mathbf{K}$  grows above 1. 
+In the rest of this subsection we will evaluate the effect of the error introduced by the approximation.
 
@@
  \centering
-\includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_5.pdf} 
-\caption{ \label{fig:y} The structured EIV problem is constructed with 200 samples of the step response to ensure that only transient data is used.
-The relative errors between the values of $\mathbf{y}$ and the steady state value are smaller than 2\% after 400 samples, as it is indicated with dashed lines.
+ \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_6.pdf} 
+  \caption{ \label{fig:spectral_radius} 
+  \color{blue} The spectral radius $\| \mathbf{M} \|$ is evaluated for different perturbation levels.
+The approximation of the inverse matrix with the Taylor series expansion gives small error when $\| \mathbf{M} \| < 1$.
+The SNR values that increase the spectral radius above one are those below 35 dB\color{black}.
 }
@@
 
+We are interested in the mean-squared error of the of the step input estimation, obtained empirically with the Monte Carlo simulation\color{black}.
 The empirical bias is the sample mean of the $N_{\mathrm{MC}}$ estimates minus the true value
  \begin{equation} {b}_\mathrm{e} = \frac{1}{N_{\mathrm{MC}}} \sum_{i=1}^{N_{\mathrm{MC}}}{ \widehat{u}_i - u } \approx \mu \left( \widehat{u} \right) - u. \end{equation}
-The standard error associated to this empirical bias estimation \citep{Hammersley75} is defined as 
-\begin{equation} \sigma_\mathrm{e} = \frac{\sigma \left( \widehat{u} \right) }{\sqrt{N_{\mathrm{MC}}}}, \quad \mathrm{where} \quad \sigma^2 \left( \widehat{u} \right) = \frac{1}{N_{\mathrm{MC}}-1} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( { \widehat{u}}_i - \mu \left( \widehat{u} \right) \right)^2 } .
+The standard error associated to this empirical bias estimation is defined in \citet{Hammersley75} as 
+\begin{equation} \sigma_\mathrm{e, err} = \frac{\sigma_\mathrm{e} \left( \widehat{u} \right) }{\sqrt{N_{\mathrm{MC}}}}, \quad \mathrm{where} \quad \sigma_\mathrm{e}^2 \left( \widehat{u} \right) = \frac{1}{N_{\mathrm{MC}}-1} \sum_{i=1}^{N_{\mathrm{MC}}}{ \left( { \widehat{u}}_i - \mu \left( \widehat{u} \right) \right)^2 } .
 \end{equation}
 
-The plots on the right side of Figure \ref{bias_sigma_NMC_unstr_str_n2} show the empirical bias, the bias predictions (\ref{eqn:biasE}) and (\ref{eqn:biasST}), and their corresponding standard errors, for the structured EIV problem.
+The plots \color{blue} in \color{black} Figure \ref{bias_sigma_NMC_str_str_n2} show the empirical bias, the bias predictions (\ref{eqn:biasE}) and (\ref{eqn:biasST}), and their corresponding standard errors, for the structured EIV problem.
 The empirical bias $b_\mathrm{e}$ is proportional to the perturbation noise variance, and
@@
 At an SNR of 30 dB the perturbation noise affects the bias prediction from observed data and it is three times smaller than the empirical bias.
+
+\begin{figure}[htb!]
+ \centering
+ \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_7.pdf} 
+  \caption{ \label{bias_sigma_NMC_str_str_n2} 
+  We observe the results of the LS solutions \color{blue} for the structured EIV problem\color{black}.
+These results are the empirical bias $b_{\mathrm{e}}$, the predicted bias from exact data $b_{\mathrm{p}}$, the predicted bias from observed data $b_{\widetilde{\mathrm{p}}}$, the standard error $\sigma_{\mathrm{e,err}}$ of the empirical bias, and the standard error $\sigma_{\widetilde{\mathrm{p}} \mathrm{,err}}$ of the the estimations bias computed using observed data.
+The estimation biases are proportional to the perturbation variance and the estimation standard errors are proportional to the perturbation standard deviation.
+Since the standard errors are smaller than the biases, the MC simulation is meaningful.
+}
+\end{figure}
 
@@
   \centering
-  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_6.pdf} 
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_8.pdf} 
   \caption{ \label{fig:b_bt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (top) between the predicted bias and the empirical bias $b_{\mathrm{e}}$ are proportional to the perturbation noise variance, and the relative errors (bottom) are smaller than 5\% only for SNR above 40 dB.}
@@
   \centering
-  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_7.pdf} 
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_9.pdf} 
   \caption{ \label{fig:v_vt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (top) between the predicted and the empirical variance are proportional to the perturbation variance, and the relative errors (bottom) between the predicted and the empirical variance are smaller than 5\% for SNR above 45 dB.
@@
   \centering
-  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_8.pdf} 
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_10.pdf} 
   \caption{ \label{fig:bv_btvt_abse_rele_str_e7} The MC simulation shows that when we solve a structured EIV problem with LS, the absolute errors (top) between the predictions computed from observed data and those from exact data are proportional to the perturbation noise variance.
@@
   \centering
-  \includegraphics[width=1\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_9.pdf}
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_11.pdf}
   \caption{\label{fig:MSE_CRLB}
@@
 \end{figure}
+
+\color{blue}
+\subsection{Simulation results for the structured IV solution}
+
+The structured IV problem, with instrumental variable (\ref{eqn:matrixZ_t}), and normal equations (\ref{eqn:neq_siv}), admits also a recursive least-squares solution. 
+A Monte Carlo simulation was conducted to find the empirical bias and variance of the solution to the IV problem.
+This simulation is conducted using the same conditions as for the structured EIV problem, in the previous subsection, and aims to compare the EIV and the IV problem solutions.   
+In Figure \ref{fig:bvMSE_IV} we can observe, for different SNR levels, the empirical bias (top), variance(center), and mean-squared error (bottom) of the solutions to the EIV and the IV problems.
+The empirical bias the IV solution is two orders of magnitude smaller than that of the EIV solution. 
+Nevertheless, the empirical variance of the two solutions are very close together, making the mean-squared error almost identical.
+
+\begin{figure}[!htpb]
+  \centering
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_12.pdf}
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_13.pdf}
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_14.pdf}
+  \caption{\label{fig:bvMSE_IV} \color{blue} In a simulation, the solutions of the structured EIV and IV problems are empirically compared.
+The bias from solving the IV problem is noticeable smaller, by two orders of magnitude, than the one from solving the EIV problem.
+This argument is valid for SNRs higher that 35 dB.
+The obtained variances, on the other hand, are very similar in the same SNR interval.
+As a consequence, the mean-squared errors almost coincide.}
+\end{figure}
+
+\color{black}
 
@@
 
-  The results of the expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$, were described by the authors of this paper in \citep{Quintana19}.
-The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \citep{Vaccaro94, Stewart90SPT}.
+  The results of the expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$, were described by the authors of this paper in \citet{Quintana19}.
+The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \citet{Vaccaro94, Stewart90SPT}.
 It was shown that the bias and variance expressions (\ref{eqn:biasST}) and (\ref{eqn:varST}) are valid predictions of the first and second moments of the LS estimate of a Hankel structured EIV problem.
@@
 
-To find the Cram\'er-Rao lower bound (CRLB) of the structured EIV estimation problem (\ref{eqn:min_ls}), we consider that this structured and correlated estimation problem can be expressed as a linear in the measurements problem \citep{Pintelon12Book}
+To find the Cram\'er-Rao lower bound (CRLB) of the structured EIV estimation problem (\ref{eqn:min_ls}), we consider that this structured and correlated estimation problem can be expressed as a linear in the measurements problem \citet{Pintelon12Book}
 \begin{equation} e (\widehat{\bm{\theta}}, \mathbf{z}) = \mathbf{M}_1( \widehat{\bm{\theta}} ) \ \mathbf{z} = \underbrace{\begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\bm{\theta}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix}}_{\mathbf{M}_1( \widehat{\bm{\theta}} )} \underbrace{\begin{bmatrix} \mathbf{y} \\ \mathrm{vec} ( \mathbf{K} ) \end{bmatrix}}_{\mathbf{z}} = 0 .
@@
 The size of the Fisher information matrix $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
-However, the Fisher information matrix $\mathbf{Fi}(\bm{\theta})$ is obtained from $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$, by doing inversion by parts \citep{Pintelon12Book} \S 19, and can be expressed as
+However, the Fisher information matrix $\mathbf{Fi}(\bm{\theta})$ is obtained from $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$, by doing inversion by parts \citet{Pintelon12Book} \S 19, and can be expressed as
 \begin{equation} \mathbf{Fi}(\bm{\theta}) = \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)^\top \left( \mathbf{M}_1( \bm{\theta} ) \mathbf{C}_{\mathbf{z}}  \mathbf{M}_1^\top( \bm{\theta} ) \right)^{-1}  \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right) .
@@
 The Cram\'er-Rao lower bound for a biased estimator of the minimization problem (\ref{eqn:min_ls}) is 
-\begin{equation} \mathrm{CRLB}_{\mathrm{b}}(\bm{\theta}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right)^\top \mathbf{Fi}^{-1}(\bm{\theta}) \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right), \label{eqn:CRB_EIV} \end{equation}
+\begin{equation} \mathrm{CRLB}_{\mathrm{b}}(\bm{\theta}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right)^\top \mathbf{Fi}^{-1}(\bm{\theta}) \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right), \label{eqn:CRLB_EIV} \end{equation}
 whereas for an unbiased estimator, it is $\mathrm{CRLB}_{\mathrm{ub}}(\bm{\theta}) = \mathbf{Fi}^{-1}(\bm{\theta})$.
@@
 
-The expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$, were studied in \citep{Quintana19} and their results are described in Lemma \ref{lem:lemma1}:
+The expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$, were studied in \citet{Quintana19} and their results are described in Lemma \ref{lem:lemma1}:
 
@@
 
-The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \citep{Vaccaro94, Stewart90SPT}.
+The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \citet{Vaccaro94, Stewart90SPT}.
 It was shown that the bias and variance expressions (\ref{eqn:biasST}) and (\ref{eqn:varST}) are good predictions of the first and second moments of the LS estimate of a Hankel structured EIV problem.
@@
 
-In the rest of this section we describe the Cram\'er-Rao lower bound (CRB) of the structured EIV estimation problem (\ref{eqn:min_ls}), that can be expressed as a linear in the measurements problem \citep{Pintelon12Book}
+In the rest of this section we describe the Cram\'er-Rao lower bound (CRLB) of the structured EIV estimation problem (\ref{eqn:min_ls}), that can be expressed as a linear in the measurements problem \citet{Pintelon12Book}
 \begin{equation} e (\widehat{\bm{\theta}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\bm{\theta}} ) \ \widetilde{\mathbf{z}} = \begin{bmatrix} \mathbf{I}_{T-n} & - \widehat{\bm{\theta}}^T \otimes \mathbf{I}_{T-n} \end{bmatrix} \begin{bmatrix} \widetilde{\mathbf{y}} \\ \mathrm{vec} ( \widetilde{\mathbf{K}} ) \end{bmatrix} = 0 .
@@
 where $\widetilde{\mathbf{z}} = \mathbf{z} + \epsilon_{\mathbf{z}}$.
-The CRB requires that the true model $\mathbf{M}_1( \bm{\theta} ) \ \mathbf{z} = 0$ exists.
+The CRLB requires that the true model $\mathbf{M}_1( \bm{\theta} ) \ \mathbf{z} = 0$ exists.
 Under the assumption of the measurement perturbation $\epsilon_{\mathbf{z}}$ being normally distributed with covariance matrix $\mathbf{C}_{\mathbf{z}}$, the loglikelihood function of the structured EIV problem is
@@
 The size of the Fisher information matrix $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ depends on the number of unknowns in $\widehat{\mathbf{z}}$ and grows with the sample size.
-Moreover, in Chapter 19 of \citep{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\bm{\theta})$ can be obtained from $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ after doing inversion by parts, giving
+Moreover, in Chapter 19 of \citet{Pintelon12Book} it is shown that the Fisher information matrix $\mathbf{Fi}(\bm{\theta})$ can be obtained from $\mathbf{Fi}(\bm{\theta}, \mathbf{z})$ after doing inversion by parts, giving
 \begin{equation} \mathbf{Fi}(\bm{\theta}) = \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)^\top \left( \mathbf{M}_1( \bm{\theta} ) \mathbf{C}_{\mathbf{z}}  \mathbf{M}_1^\top( \bm{\theta} ) \right)^{-1}  \left( \frac{\partial e (\widehat{\bm{\theta}}, \mathbf{z}) }{\partial \bm{\theta} } \right)
@@
 The Cram\'er-Rao lower bound for an biased estimator of the minimization problem (\ref{eqn:min_ls}) is given by
-\begin{equation} \mathrm{CRB}_{\mathrm{b}}(\bm{\theta}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right)^\top \mathbf{Fi}^{-1}(\bm{\theta}) \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right), \label{eqn:CRB_EIV} \end{equation}
-and for an unbiased estimator it is $\mathrm{CRB}_{\mathrm{ub}}(\bm{\theta}) = \mathbf{Fi}^{-1}(\bm{\theta})$
+\begin{equation} \mathrm{CRLB}_{\mathrm{b}}(\bm{\theta}) = \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right)^\top \mathbf{Fi}^{-1}(\bm{\theta}) \left( \mathbf{I}_{n+1} + \frac{\partial \mathbf{b} \left( \widehat{\bm{\theta}} \right) }{\partial \bm{\theta} } \right), \label{eqn:CRLB_EIV} \end{equation}
+and for an unbiased estimator it is $\mathrm{CRLB}_{\mathrm{ub}}(\bm{\theta}) = \mathbf{Fi}^{-1}(\bm{\theta})$
 \end{comment}

### 2020-08-21 Extras/ListofPublications.tex
Story 182: Intro complete
Second review version
Final version before internal defense
Second draft before public defense

@@
 
-\subsubsection*{Journal publication2}
-
-\noindent G.
-Quintana-Carapia, I.
-Markovsky, R.
-Pintelon, P.
-Z.
-Csurcsia and D.
-Verbeke, "Experimental validation of a data-driven step input estimation method for dynamic measurements," \textit{IEEE Transactions on Instrumentation and Measurement}, Vol.
-x, No.
-x, pp.
-xx-xx, 2019, doi: 10.1109/TIM.2019.2951865
-
-\vspace{0.2cm}
+\subsubsection*{Journal publications}
 
@@
 
-\vspace{0.2cm}
+%\vspace{0.1618cm}
+
+\noindent G.
+Quintana-Carapia, I.
+Markovsky, R.
+Pintelon, P.
+Z.
+Csurcsia and D.
+Verbeke, "Experimental validation of a data-driven step input estimation method for dynamic measurements," \textit{IEEE Transactions on Instrumentation and Measurement}, \color{blue} vol.
+69, no.
+7, pp.
+4843-4851, July 2020, \color{black} doi: 10.1109/TIM.2019.2951865.
+
+%\vspace{0.1618cm}
+
 \noindent G.
@@
 
-\vspace{0.2cm}
+\vspace{0.1618cm}
 \subsubsection*{Conference publications}
@@
 
-\vspace{0.2cm}
+%\vspace{0.1618cm}
 
@@
 In: \textit{35th Benelux Meeting on Systems and Control}", Soesterberg, The Netherlands, 2016.
+
+%\newpage

### 2020-08-21 Extras/Nomenclature.tex
Second draft before public defense

% I made my list of abbreviations manually because the glosseries didn't work
\chapter*{Nomenclature}

\color{blue}
A list of the symbols used in this thesis is found in the following list.
\begin{longtable}[c]{>{\raggedleft}p{0.2\columnwidth}>{\raggedright}p{0.7\columnwidth}}
${a}, {b}$  &  true parameters an affine input model \tabularnewline
$\widehat{a}, \widehat{b}$  &  estimated parameters an affine input model \tabularnewline
$\mathbf{A}, \mathbf{B}, \mathbf{C}, D$  &   matrices of an LTI dynamic system state space representation  \tabularnewline
$\widehat{\mathbf{A}}, \widehat{\mathbf{B}}, \widehat{\mathbf{C}}, \widehat{D}$  &   estimated matrices of an LTI dynamic system state space representation  \tabularnewline
$\widehat{\mathbf{A}}_\text{a}, \widehat{\mathbf{C}}_\text{a}$  &  estimated matrices of an autonomous LTI dynamic system state space representation \tabularnewline
$\mathbf{A}_\text{a}, \mathbf{C}_\text{a}$  &  matrices of an autonomous LTI dynamic system state space representation \tabularnewline
$\mathbf{b} \left(\widehat{\bm{\theta}} \right)$  &  bias of the estimate $\bm{\theta}$  \tabularnewline
$\mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right)$  &  predicted bias of the estimate $\bm{\theta}$ obtained from exact data \tabularnewline
$\widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right)$  &  predicted bias of the estimate $\bm{\theta}$ obtained from measured data \tabularnewline
$\mathbf{b}_{\widetilde{\mathrm{p}} \left[1\right]}$   &  first element of the estimation predicted bias obtained from measured data \tabularnewline
$b_\mathrm{e}$  &  empirical bias of the step input estimate  \tabularnewline
$b_{\mathrm{p}}$  &  bias prediction computed using exact data   \tabularnewline
$b_{\widetilde{\mathrm{p}}}$  &  sample mean of the step input estimation bias, computed  from observed data predictions   \tabularnewline
$\mathbf{B}_1, \mathbf{B}_2, \mathbf{B}_3, \mathbf{B}_4$  &  matrices of the bias prediction formulas computed from exact data  \tabularnewline
$\widetilde{\mathbf{B}}_1, \widetilde{\mathbf{B}}_2, \widetilde{\mathbf{B}}_3, \widetilde{\mathbf{B}}_4$  &  matrices of the bias prediction formulas computed from measured data  \tabularnewline
$\mathbf{C} \left( \widehat{\bm{\theta}} \right)$  &  covariance matrix of the estimate $\bm{\theta}$  \tabularnewline
$\mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right)$  &  predicted covariance of the estimate $\bm{\theta}$ obtained from exact data \tabularnewline
$\widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right)$  &  predicted covariance of the estimate $\bm{\theta}$ obtained from measured data \tabularnewline
$\mathbf{C}_{\widetilde{\mathrm{p}} \left[1,1\right]}$  &  first element of the estimation predicted covariance obtained from measured data \tabularnewline
$\mathbf{C}_1, \mathbf{C}_2$  &  matrices of the covariance prediction formulas computed from exact data  \tabularnewline
$\widetilde{\mathbf{C}}_1, \widetilde{\mathbf{C}}_2$  &  matrices of the covariance prediction formulas computed from measured data  \tabularnewline
$\mathbf{C}_{\mathbf{z}}$  &  covariance matrix of a linear in the measurements problem $e (\widehat{\bm{\theta}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\bm{\theta}} ) \ \widetilde{\mathbf{z}}$ \tabularnewline
$\mathrm{CRLB}(\bm{\theta})$  &  Cram\'er-Rao lower bound of the estimate $\bm{\theta}$  \tabularnewline
$\mathbf{D}_{r \times c}^{1,d}$ and $\mathbf{D}_{r \times c}^{2,d}$  &  first and second-order finite differences matricial operators of dimensions $r \times c$ starting from the subdiagonal $d$  \tabularnewline
$\Delta$  &  signals first difference operator, defined as $\Delta y(k) = y(k + 1) - y(k)$  \tabularnewline
$\Delta \mathbf{x}(k)$ &   autonomous system state vector at time $k$ \tabularnewline
$\Delta y(k)$ &   sample of an autonomous system response at time $k$ \tabularnewline
$\mathbf{E}$  &  measurement noise that is added to the regression matrix in a system of linear equations  \tabularnewline
$\bm{\epsilon}$  &   measurement noise that is added to the system response  \tabularnewline
$\bm{\epsilon}(k)$  &   samples of the measurement noise at time $k$  \tabularnewline
$\bm{\epsilon}_{\mathbf{z}}$  &  perturbation noise of a linear in the measurements problem $e (\widehat{\bm{\theta}}, \widetilde{\mathbf{z}}) = \mathbf{M}_1( \widehat{\bm{\theta}} ) \ \widetilde{\mathbf{z}}$ \tabularnewline
$\mathbb{E} \left\{ \cdot \right\}$  &  mathematical expectancy operator  \tabularnewline
$\mathbf{Fi}(\theta)$  &  Fisher information matrix  \tabularnewline
$\mathbf{G}$  &   static gain matrix of an LTI dynamic system  \tabularnewline
$\mathbf{H}(s)$, $\mathbf{H}(z)$  & continuous and discrete time transfer functions of a dynamic system, respectively \tabularnewline
$\mathbfcal{H}({y})$ and $\bm{\mathcal{H}}(\bm{\epsilon})$  &  $n \times n$ Hankel matrices constructed from the system response $y$ and the measurement noise $\bm{\epsilon}$, respectively  \tabularnewline
$\mathbfcal{H}_{L+1}(\Delta {y})$  &  Hankel matrix of $L+1$ block rows constructed from $\Delta {y}$  \tabularnewline
$\mathbf{I}_n$  & $n \times n$ identity matrix   \tabularnewline
$\mathbf{J} = \partial \mathbf{r} / \partial \bm{\theta}$  &  Jacobian matrix of the residual $\mathbf{r}$   \tabularnewline
$k$  &  $k\mathrm{-th}$ instant of time  \tabularnewline
$\mathbf{K}$  &  regression matrix in a system of linear equations constructed from exact data  \tabularnewline
$\widetilde{\mathbf{K}}$  &  regression matrix in a system of linear equations constructed from measured data  \tabularnewline
$\widetilde{\mathbf{k}}_{k}$  &  row of matrix $\widetilde{\mathbf{K}}$ that corresponds to the $k\mathrm{-th}$ sample of the measured response  \tabularnewline
$\mathbf{K}^\dagger$  &  pseudo-inverse of the matrix $\mathbf{K}$  \tabularnewline
$k_{\mathrm{T}}$  &   temperature coefficienct of a first-order thermometer model  \tabularnewline
$\bm{\kappa}$  &  scalar gain in the recursive least-squares algorithm  \tabularnewline
$l$  &  likelihood function of an estimation problem  \tabularnewline
$\bm{\ell}$  &   linear transformation of the initial conditions in the data-driven step input estimation method  \tabularnewline
$\lambda$  &  eigenvalues of a dynamic system  \tabularnewline
$m, k_{\mathrm{s}}, k_{\mathrm{d}}$  &  mass, stiffness and damping coefficients of a second-order weighing system model  \tabularnewline
$\bm{\mu}\left(\widehat{\bm{\theta}} \right)$  &  expected value of the estimate $\bm{\theta}$  \tabularnewline
$\mathrm{MSE}_{\mathrm{e}}$  &  mean-squared error computed using the empirical results in the Monte Carlo simulation  \tabularnewline
$\mathrm{MSE}_{\mathrm{p}}$  &  mean-squared error computed using the predictions from exact data  \tabularnewline
$\mathrm{MSE}_{\widetilde{\mathrm{p}}}$  &  mean-squared error computed using the predictions from observed data  \tabularnewline
$n$  &  order of a dynamic system  \tabularnewline
$N$  &  maximum number of available system response samples   \tabularnewline
$N_{\mathrm{MC}}$  &  number of Monte Carlo runs   \tabularnewline
$\mathbfcal{O}$  &  observability matrix of an LTI dynamic system  \tabularnewline
$\mathbfcal{O}_\text{a}$  &  observability matrix of an autonomous LTI dynamic system  \tabularnewline
$\omega$   &  weighing factor of the recursive least-squares algorithm  \tabularnewline
$\bm{\Omega}$   &  diagonal matrix of descending powers of the weight factor $\omega$  \tabularnewline
$\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger$  &  orthogonal matrix computed from exact data  \tabularnewline
$\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger$  &  orthogonal matrix computed from measured data  \tabularnewline
$\bm{\Psi}$  &  covariance matrix of the recursive least-squares algorithm algorithm  \tabularnewline
$\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$  &  residual difference between the measured response $\widetilde{\mathbf{y}}$ and the simulated response $\widehat{\mathbf{y}}(\bm{\theta})$  \tabularnewline
$\mathbf{R}_n$  &  $n \times n$ reversal matrix   \tabularnewline
$\rho(\mathbf{M})$  &   spectral radius of matrix $\mathbf{M}$ \tabularnewline
$s(t)$ &  unit step input \tabularnewline
$\sigma_\mathrm{e}^2$  &  empirical variance of the step input estimate  \tabularnewline
$\sigma_\mathrm{e,err}$  &  standard error associated to the the estimation empirical bias in the Monte Carlo simulation  \tabularnewline
$\sigma_{\mathbf{E}}^2$  &    variance of the measurement noise matrix $\mathbf{E}$, assuming its columns are stacked in a vector  \tabularnewline
$\sigma_{\widetilde{\mathrm{p}}, \mathrm{err}}$  &  standard error of the of the empirical estimate Monte Carlo simulation  \tabularnewline
$\sigma_{\bm{\epsilon}}^2$  &  variance of the measurement noise $\bm{\epsilon}$ \tabularnewline
$\mathbfcal{T}$  &  Markov parameters matrix of an LTI dynamic system  \tabularnewline
$\bm{\theta}$  &   true value of the to-be-estimated parameters in a linear regression problem  \tabularnewline
$\widehat{\bm{\theta}}$  &   estimated parameters in a linear regression problem  \tabularnewline
$u(t)$ &   input signal applied to a dynamic systems \tabularnewline
$\widehat{u}$ &   estimated input signal \tabularnewline
$u(k)$ &  sample of the input signal at time $k$  \tabularnewline
$\mathbf{u}$  &  vector constructed from samples of the input signal  \tabularnewline
$\mathbf{U}, \bm{\Sigma}, \mathbf{V}$  &    singular value decomposition matrices  \tabularnewline
$v_{{\mathrm{p}}}$  &  variance prediction computed using exact data   \tabularnewline
$v_{\widetilde{\mathrm{p}}}$  &  sample mean of the step input estimation variance, computed  from observed data predictions   \tabularnewline
$\mathbf{x}(k)$ &   system state vector at time $k$ \tabularnewline
$\mathbf{x}_{\text{ini}}$  &  initial conditions of a dynamic system  \tabularnewline
$\mathbf{x}_{\text{ini,c}}$  &  initial conditions of a compensator dynamic system  \tabularnewline
$\mathbf{X}_{\text{ini}}$  &  matrix of the initial conditions of a dynamic system at different instant of time  \tabularnewline
$\mathbf{x}_\text{a}(0)$  &  initial conditions of an autonomous dynamic system  \tabularnewline
$\mathbf{x}_\text{a}(k)$  &   autonomous system state vector at time $k$ \tabularnewline
$y(t)$  &  exact response of a dynamic system\tabularnewline
$\widehat{y}$  &   simulated exact response signal \tabularnewline
$\widetilde{\mathbf{y}} = \mathbf{y} + \bm{\epsilon}$  &  observed system response afected by measurement noise  \tabularnewline
$y(k)$  &  sample of the exact response signal at time $k$\tabularnewline
$\mathbf{y}$  &  vector constructed from samples of the system response signal   \tabularnewline
$\widetilde{\mathbf{z}} = \mathbf{z} + \epsilon_{\mathbf{z}}$   &  vector constructed by stacking the perturbation noise of the regressor and the regression matrix of the EIV problem \tabularnewline
$\widetilde{\mathbf{Z}}$  &  instrumental variable constructed from measured data  \tabularnewline
% \tabularnewline
\end{longtable}
\color{black}

### 2020-08-21 internal defense feedback/Answers-jury-members-report - BK2.tex
Second draft before public defense

\documentclass[11pt]{article}
\date{\vspace{-10ex}}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Added lines by Gustavo Quintana to facilitate internal review process
\usepackage{amsmath, amsfonts, mathtools, hyperref, tikz, pgf, subcaption, mathdots}
\usepackage[a4paper, total={6.5in, 9in}]{geometry}
\usepackage{setspace}
\usetikzlibrary{shapes, arrows, calc, patterns, decorations.pathmorphing, decorations.markings, positioning, external}
\tikzexternalize
\tikzset{external/system call={pdflatex \tikzexternalcheckshellescape -halt-on-error -interaction=batchmode -jobname "\image" "\texsource" && pdftops -eps "\image.pdf"}}
\tikzexternalize[shell escape=-enable-write18]
\usepackage{algorithm, algorithmic, bm}
%\usepackage[noend]{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
%\renewcommand{\algorithmicreturn}{\textbf{Initialize:}}

\makeatletter
\newcommand*\rel@kern[1]{\kern#1\dimexpr\macc@kerna}
\newcommand*\widebar[1]{%
  \begingroup
  \def\mathaccent##1##2{%
    \rel@kern{0.8}%
    \overline{\rel@kern{-0.8}\macc@nucleus\rel@kern{0.2}}%
    \rel@kern{-0.2}%
  }%
  \macc@depth\@ne
  \let\math@bgroup\@empty \let\math@egroup\macc@set@skewchar
  \mathsurround\z@ \frozen@everymath{\mathgroup\macc@group\relax}%
  \macc@set@skewchar\relax
  \let\mathaccentV\macc@nested@a
  \macc@nested@a\relax111{#1}%
  \endgroup
}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\newtheorem{innercustomgeneric}{\customgenericname}
\providecommand{\customgenericname}{}
\newcommand{\newcustomtheorem}[2]{%
  \newenvironment{#1}[1]
  {%
   \renewcommand\customgenericname{#2}%
   \renewcommand\theinnercustomgeneric{##1}%
   \innercustomgeneric
  }
  {\endinnercustomgeneric}
}
\newcustomtheorem{customthm}{Theorem}
\newcustomtheorem{customlemma}{Lemma}

\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}

\begin{document}

\title{Jury members' comments on the private PhD defense of \linebreak Gustavo Quintana-Carapia}

\maketitle

Thank you to the jury members for reviewing the thesis manuscript and for giving suggestions for improvement. 
The manuscript changes due to the comments are highlighted in \color{blue} blue fontcolor\color{black}.

\section*{General comments}

\begin{itemize}
	\item The manuscript text should better position the PhD to related work. Improve the literature review and discuss (and cite) the state of the art.

	{\bfseries The introduction of the chapters now includes the state of the art to position the PhD work with respect to the literature.}

	\item  Add more clearly the motivation for the PhD research work by including motivational examples to the introduction.

	{\bfseries Motivational examples have been added to the Introduction. }

	\item  Add a discussion with comparison of existing EIV methods. Give, with the positive and negative aspects in mind, a better justification for why you use a (R)LS method, and not, e.g., (R)IV method.

	{\bfseries In Section 4.1.4, a discussion that compares IV methods with EIV methods has been added.}

	\item  Discuss more clearly the aspects of scalability of the methods and their limitations in terms of model order, size of data sets (e.g., related to the inversion of LS matrices).
	\item  The equations on pages 12 and 13 seemingly contain errors and should be corrected.

	{\bfseries The equations on pages 12 and 13 have been corrected.}

	\item  The figure shown on page 66 was made with a single experiment; it is desirable to see the (averaged) effect on several experiments.

	{\bfseries A series of 100 runs of the maximum-likelihod estimation method has been simulated and the results have been added to the manuscript in Section 6.3.2.}

\end{itemize}

\section*{Lyudmila Mihaylova}

\begin{itemize}
	\item You develop results for uncertainty quantification; what type of uncertainties  can your approach deal with and which levels of noise can you consider? Elaborate on SNR levels; I’m also very interested in scalability, so how big the matrices can be.
	\item  Concerning the scalability to data: eqn 3.6: how scalable is the approach in all aspects (order, data, linked to inversion for LS matrix); can you deal with hundreds, thousands of data?

	{\bfseries The following paragraph has been added in Section 3.2 to discuss the scalability of the data-driven step input estimation method.}

	\color{blue}
    The computational complexity of the RLS algorithm solution to the system of equations (3.14) is $O \left( \left( n+1 \right)^2 \right)$.
    The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
    From there on, the RLS algorithm updates the solution with linear complexity.
    Therefore, even though a large number of samples $N$ makes the matrix in the system of equations (3.14) increase in the number of rows, the estimation of the solution does not increase in complexity. 
    The data-driven step input estimation method is then scalable for any sensor of order $n$, and can be executed in devices with limited computational resources, provided that the computation of the  $n+1 \times n+1$ inverse matrix is feasible.
    \color{black}

	\item  What recommendation up to scale/order can method work with. What are the limitations? It would be good to have more precise discussion about these aspects in dissertation: what do you mean by large scalability: 100, $10^4$, ...? A more detailed discussion is desirable.
	\item  On slide 25 (noisy graphs) it would be good to see a statistical analysis (more experiments than one).

    {\bfseries 100 runs of the maximum-likelihod estimation method have been simulated and their results have been added to the manuscript in Section 6.3.2.}

	\color{blue} 
    The maximum-likelihood (ML) method estimated the affine input parameters from the sensor model transient response.
    A total of 100 runs of the method were obtained, using different realizations of measurement noise.
    The measurement noise variance was set to have an SNR of 40 dB.
    The ML method used the first 50 samples to initialize the optimization variables and updated the variables every $N_{\mathrm{s}} = 5$ samples.
    Figure \ref{fig:rele_lo_40dB_s10} shows the average of the observed relative errors in the estimation of the parameters $\widehat{a}$, $\widehat{b}$, $\widehat{x}_{\mathrm{ini,1}}$, and $\widehat{x}_{\mathrm{ini,2}}$.

    \renewcommand{\thefigure}{6.8}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_7.pdf} 
    \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method. After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and everyone of the estimates. The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%. \color{black} }
    \end{figure}

    Figure \ref{fig:cov_lo_40dB_s1} shows \color{blue} the average \color{black} of the variances that were observed on the diagonal of the covariance matrix $\mathbf{J}$. 
    We can see that the estimation variances decrease as more samples are processed.
    Moreover, the estimation variances of $\widehat{a}$ and $\widehat{b}$ obtained with the ML method are lower than the corresponding estimation MSE errors obtained from a Monte Carlo simulation of the subspace method.

    \renewcommand{\thefigure}{6.9}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_8.pdf} 
    \caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian. \color{blue} The estimates $\widehat{x}_{\mathrm{ini,1}}$ and $\widehat{a}$ have the smaller and larger variances, respectively. \color{black}  }
    \end{figure}

    \color{black}

\end{itemize}

\section*{Stephane Chretien}

\begin{itemize}
	\item The Taylor expansion is computed of order 2, what is the accuracy of expansion with respect to system, noise level,... Can you come up with a priori bound on error that will allow you to assess the quality of the expansion? Can you predict where the solution is going to be before computing Taylor expansion?

	{\bfseries An a priori bound on the error introduced by the Taylor series expansion can be obtained in terms of the spectral radius of the matrix $\mathbf{M}$. In Section 4.1 the following description has been added: }

    \color{blue} An a priori bound on the error introduced by the second order Taylor series expansion can be expressed in terms of the spectral radius $\rho(\mathbf{M})$.
    Considering that the Taylor series expansion is an infinite sum we have that
    \begin{equation} \tag{4.4} \begin{aligned} (\mathbf{I} + \mathbf{M})^{-1} &= \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \\
    &=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \right) \\
    &=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \end{aligned} \end{equation} 
    Therefore we can obtain a bound for the second error expansion error if we take a matrix norm as follows
    \begin{equation} \tag{4.5} \begin{aligned} \| (\mathbf{I} + \mathbf{M})^{-1} - ( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 ) \| &=  \| \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \| \\
    & \leq \| \mathbf{M}^3 \| \| (\mathbf{I} + \mathbf{M})^{-1} \| \leq \dfrac{\| \mathbf{M} \|^3}{ 1 -  \| \mathbf{M} \|} = \dfrac{\rho(\mathbf{M})^3}{1 - \rho(\mathbf{M})}  \end{aligned} \end{equation} 
	\color{black}

	\item  Concerning other sorts of prior estimations, e.g., given by explicit function theorems. I wonder if you could get something and apply this bound in order to control the bound of the 2nd order expansion. There exist some good results that could help you devise something a priori which can control the error you are making. Neuberger's results (quite old now) but may be handy for this kind of problem. This can be investigated further.

	\color{blue}
	Neuberger's paper \cite{Neuberger07} introduces five theorems that are consequence of the continuous Newton method.
	The Newton method aims to find a zero of the function $F: R^n \rightarrow R^n$ by iterating
	\begin{equation} z_{k+1} = z_k - \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
	for $k=0,1,2,\ldots$, and for a given initial $z_0$, assuming that $\left(F' (y) \right)^{-1}$ exists for some $y \in R^n$.
	A domain of attraction is the set of all initial values $z_0$ that lead to a root of $F$ after the convergence $z_0, z_1, z_2,\ldots$.
	The damped Newton method  
	\begin{equation} z_{k+1} = z_k - \delta_k \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
	prevents convergence issues, such as chaotic domains of attraction, with $\delta_1, \delta_2, \ldots \in (0, 1)$. 
	The continuous Newton method is a sequence of damped Newton methods. To implement the continuous Newton method, select $T>0$ and perfom $m$ runs of the damped Newton method with $\delta_k = T/m$, for $k=1,2,\ldots, m$.
	The result $z_{k+1}$ is the zero estimation $x_m$, and if the sequence of estimates $x_1, x_2, \ldots$ converges, then it converges to the result of the continuous Newton method.

	The objective of the continuous Newton method is to find a function $z: [0, \infty) \rightarrow R^n$, so that
    \begin{equation} z(0) = x \in R^n, \quad z' (t) = - \left(F' (z(t)) \right)^{-1} F(z(t)), \quad t \geq 0, \label{eqn:contNewton}\end{equation}
    subject to the existence of the limit $u=\mathrm{lim}_{t \rightarrow \infty}{z(t)}$, that satisfies $F(u)=0$.
    This implies 
    \begin{equation}  F' (z(t)) z' (t) = - F(z(t)) , \end{equation}
    that is equivalent to
    \begin{equation}  \left(F \circ z\right)' (t) = - F(z(t)) , \end{equation}
    from where we get
    \begin{equation}  F(z(t)) = e^{-t} F(z(0)) . \end{equation}
    Thus, the residual $F(z(t))$ only changes in magnitude and not in direction.
    This property is essential for the results of the theorems in \cite{Neuberger07}. 
    In particular, Theorem 5 gives the conditions for the continuous Newton method to ensure finding a root of the function.

    \begin{customthm}{5} \label{thm:five}
    Suppose that each of $H$, $J$, and $K$ is a Banach space, with $H$ compactly embedded in $J$, that $r>0$, and that $G:B_r(0) \rightarrow K$ is continuous as a function on $J$. Suppose also that $g$ belongs to $K$ and that for each $y$ in $b_r(0)$ there is an $h$ in $B_r(0)$, where $b_r(0)$ and $B_r(0)$ are open and closed balls in $H$ of radius $r$ centered at 0, such that
    \begin{equation} \lim\limits_{t \to 0+} -{\dfrac{1}{t} \left(G(y+th) - G(y)\right) = g}. \end{equation}
    Then, there exists $u$ in $B_r(0)$ such that $G(u)=g$.
    \end{customthm}
    %$F:B_r(0) \rightarrow K$

    The theorem implies that $u$ is a zero of the function  defined as $F(y) = G(y) - g$.  
    Since $G$ is a continuous function and if $\left(G' (y) \right)^{-1}$ exists for each $y$ in $B_r(0)$, a metric for the validity of the Theorem (\ref{thm:five}) hypothesis is the inequality $\left\Vert  \left( G' (y) \right)^{-1} g \right\Vert \leq r$.

    To link the data-driven step-input estimation problem with this theorem, we can take the system of linear equations $\widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \bm{\theta}$, then define $G(y) = \widetilde{\mathbf{K}} \bm{\theta}$, and $g = \widetilde{\mathbf{y}}$, so that we have $F(\bm{\theta}) = G(\bm{\theta}) - g$. The derivative of the function $G(\bm{\theta})$ is given by $G'(\bm{\theta}) = \widetilde{\mathbf{K}} \mathbf{I}$. Using the pseudo-inverse of $\widetilde{\mathbf{K}}$, the metric $\left\Vert  \widetilde{\mathbf{K}}^{\dagger} \widetilde{\mathbf{y}} \right\Vert \leq r$ indicates that the estimate $\hat{\bm{\theta}}$ is found inside a ball of radius $r$.
    Thus,  
	\color{black}

	\item  About trying to find a good transformation from LS to EIV model. I am wondering if you can think of other more accurate starting points instead of LS. Closer to the problem you want to study, but still preserving computational lightness? The closer you start from objective, better the Taylor expansion would be.. Something to investigate as well...
	\item  About robustness: is it an issue in practical problems with heavy tailed noise, outliers,... Perhaps the use of median of means plugged into your approach might produce interesting alternatives for outliers or heavy tailed data. But this is an intricate thing to deal with. Might be interesting to ponder this in future.
	\item  On deep learning: could this be an alternative?
\end{itemize}

\section*{Guillaume Mercere}

\begin{itemize}
	\item In the first or second sentence: “users cannot wait for a steady state regime” . Do you have specific examples of real industrial systems for which it is necessary to have continuous read-out before steady-state is reached? - Further on the industrial part and selling sensors with data-driven parts. Can you explain to experts that your solution is better? Sensors will be used for ages; why not spend one day to do system identification and get a reliable model to make a compensator, instead of using your approach?

	{\bfseries The fast weighing of parcels in the post office is a typical example of the real life systems where the time is limited to perform the measurements. Another one that we see nowadays is the temperature measurement of people in airports or train stations to prevent the spread of corona virus. The data driven method can be implemented in portable devices with limited computational resources. In industrial plants were a time slot can be devoted to identify the measurement system and implement a compensation method, the solution might be useful only for a certain amount of time. It is assumed that the model approximation deviates due to changes in the system elements, leading to inaccuracies in the inpiut estimation. There is a need for adaptive methods to avoid stopping the production lines to repeat the system identification procedure. The data-driven method is such an adaptive method. }

	\item  Chapter 2, Section 2.2.1. Can you give me some classic solutions/algorithms instead of LS in general? (Say I'm not familiar with EIV or TLS?) Usually with LS problems: multiply with $K^\top$ and compute pseudoinverse. Do you know another solution where they don't use K on the left-hand-side but something else? To get rid of noise by using a matrix on the left-hand-side, is related to EIV. There are some recursive solutions for this. It could be interesting to compare with such IV solutions.
	\item  In Chapter 3: at beginning you say to use linear systems theory to model a sensor. Can you discuss this assumption: is it true in practice? Can you guarantee it behaves like an LTI system?
	\item  Equation 3.1: an LTI system, described with SS representation. You only put output noise on this representation. People sometimes also add process noise. Why? How to modify your approach if there is process noise? Is it still white in this case?
	\item  On page 12 standard approach top of page equation: there are some mistakes in the equations! there is no epsilon (noise) in the equation. But then you do an SVD of H and estimate from there. How does it work if you have noise? Have you tried implementing it and see what you get?

	{\bfseries The step input estimation based on singular value decomposition works only with exact data. When  noise perturbes the sensor response the method cannot guarantee the step input estimation. The singular value decomposition of a Hankel matrix constructed from a perturbed sensor response provides only an approximation of the singular values and the singular vectors of this Hankel matrix. Then, the step input estimation obtained with this method from perturbed sensor response is only a rough estimation.     }

	\item  On page 12: I’m not sure the equations at the end of the page are correct, e.g., Y=G ubar ... not sure there is x; same for rest; There seem to be many mistakes here and next pages. Be careful with the equations!

	{\bfseries The equations have been reviewed and the typos were corrected.}

	\item  On page 13, you assume persistency of excitation of order $L$. How do you check this? In your case the input is constant; how to guarantee that the assumption is met?

	{\bfseries The persistency of excitation of a signal is verified by looking at the rank of the Hankel matrix constructed from the consecutive samples of this signal. When the Hankel matrix of $L+1$ block rows has a rank lower or equal than L, then we say that the signal is persistently exciting of order $L$. A step input has a very low order of persistency $L=1$, but the step respose of a system has a larger order of persistency. In fact, we assume that if the persistency of excitation of the step response is sufficiently high, then the columns of the Hankel matrix constructed from the step response span a linear space with all the responses of the equivalent augmented autonomous system where the step input is considered an augmented initial condition.}

	\item  In Chapter 4, p 17, Section 4.1.1: You assume that perturbations are IID and must be normally distributed. Why using an assumption of normally distributed? I would think that the kind of distribution is important for CRLB, but I’m not sure you need it for eqns 4.8, 4.9, 4.10,... IID is important, normality perhaps not.

	{\bfseries The following paragraph has been added in Section 4.1 to discuss on the need for the normality assumption.  }

	\color{blue}
	The normality assumption is necessary to provide more prior information into the method formulation.
    The second order Taylor series expansion (4.4) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
    Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance, which can be disregarded since $3 \sigma_{\bm{\epsilon}}^4 \ll \sigma_{\bm{\epsilon}}^2$\color{black}.

	\item   In equation 4.11 you have $\sigma_eps^2 + \sigma_e^2$; how to get in practice $\sigma_e$? Do you have to do this each time? Recursively?

	{\bfseries The measurementy noise variance is prior information and in Chapter 5 we describe two ways of estimating $\sigma_{\bm{\epsilon}}$ In summary, the measurement noise can be extracted fom the sensor ressponse once the transient effects are smaller than the noise floor, or from the sensor response power spectrum. The online estimation of the measurement noise, simultaneously with the online step input estimation is a topic for future research.   }

	\item   Switch to 4.2, simulation results: first part dedicated to K, randomly generated; not sure if still possible, but in all your results it would be nice to have somewhere explicitly K = ... (even if randomly generated). You use only one realization also for matrices A, B, C, D in the text, so it can be added explicitly.
	\item   in Fig 4.1 the bottom figure also discusses the structured case. Why do you talk about structured results in a section about the unstructured case?

	{\bfseries Section 4.2 has been improved to describe the result next to its discussion.  }

	\item   In 4.1 uses a constant K, but in 4.2.2 you use A, B, C, D, without any links to K. You compare the two curves generated from different data sets… It is strange to compare them. I don't see the link between both.

	{\bfseries Section 4.2 has been reestructured and the simulation for the unstructured case were redone to establish a link between the two cases. The exact data is the same in the two cases and the perturbation noise follows the corresponding rules of each case, with the same variance noise levels. In this way the comparison between the two cases makes more sense.  }

	\item   To show comparison between structured and unstructured case, use unstructured results for structured case; then use unstructured results for structured case and cover all the possibilities. Now you compare curves related to two different simulations.

	{\bfseries \color{red} Interesting question, I want to try it this evening \color{black}}

	\item   I was surprised to read in Chapter 5 something very close to what I read in chapter 4. In both cases generated A, B, C. In 5.1 again oscillating 5th order system, similar conclusions, ... What is the added value of 5.1 wrt 4.2.2?

	{\bfseries The derivation described in Chapter 4 was conducted with a second order mass-spring-damper weighing system example on mind, and the results observed were submitted in the corresponding paper. Afterwards, when we moved to the practical implementation, we faced the increase of the model order after setting up the weighing experiment. The lowest order we can assume is fifth order. So we decided to repeat the simulations and these new results serve two purposes: a transition from simulation to experimental results, and also to get an insight on the impact of the bias and variance predictions after increasing the model order.    }

	\item   On the impact of selected order on results. You were talking about a related bias/variance trade off. So, do you know in the literature some solutions to avoid overfitting? Can you cite something? Can you use something in your case to test for a good order of system?
	\item   Do you know of cross-validation: split data into two parts or several parts; some for estimation, some for validation? Have you tried this? Can this be used here to select good orders? Better results? Influence of system order on bias/variance; if I only use one dataset it is complicated and usually you can increase complexity to mimic a data set, but to avoid overfitting (too high order) use a test/validation data set. with this i should see what is the best value... Have you tried this?
	\item   In Section 6.2.2. Why do you call it a max likelihood approach? To me it is not one. Why do you call it like this? Where is the likelihood? Depending on which likelihood (eg. Gaussian, ...) it leads to LS. Don't' call it ML if you dont give distributions on ... In this case it is better to call it a nonlinear minimization procedure. Not clear why you call it ML.

	{\bfseries The method description in Section 6.2.2 has been modified to explain why the method is a maximum likelihood method.  }

	\item   Concerning the ML approach in simulations, you say that it takes 30 seconds to complete. What is done exactly in these 30 seconds; how many samples, realizations,... I am a bit surprised you require 30 sec for 2 parameters.
\end{itemize}

\section*{Nikos Deligiannis}

\begin{itemize}
	\item When solving algorithms. Do you opt for closed-form solutions, or do you use optimization based algorithms? How to solve LS problems? Do you solve them using pseudo-inverse or do you do gradient descent?
	\item  Today there is an overwhelming amount of data sensed and communicated; How would you apply your method in a scenario where dimensionality increases significantly? This relates to the questions of previous jury members; what is the implication of using numerical versus closed form solutions?
	\item  What if the $K^\top K$ matrix is not invertible? What if you have fewer samples than unknowns; what can you do? Solutions that assume structure like in compressed sensing, regularization,... can you use them, what would be the impact of that?
	\item  You mention that you improve complexity with your approach, but I missed a complexity analysis. Did you do this? Anything in processing time, flops,...?
	\item  You mentioned deep learning could be applied; could you show how? An issue is that neural nets are trained in high SNR data, but applied in engineering with lower SNR. This clogs performance; any ideas? Is this a limitation?
	\item  You mention that you consider Kalman filters; have you seen approaches where they write Deep Learning in form of Kalman Filter; Could this be applicable?
	\item  Recommendation: if you revise the thesis: maybe change "I propose" to "we propose".

	{\bfseries The formulation “I propose” was replaced by “we propose”. }

\end{itemize}

\section*{Philippe Dreesen}

\begin{itemize}
	\item Your main assumption is that a sensor outputs a steady state value that is proportional to the to-be-measured quantity. This means that you assume that the sensor is a linear system. What if this assumption is not met, and the sensor has a saturation for high values? Can you still use your methods?
	\item  Sensors are physical devices and deliver a continuous-time output. The models that you use are in discrete-time. Are there any risks that you miss information by not sampling fast enough?

	{\bfseries In Chapter 3 has been added a discussion on the step invariance transformation. This defines the conditions of the input for describing exactly the system dynamics in discrete time after sampling the continuous system response. If the input is stepwise continuous, then the zero order hold sampling of the step response leads to an exact discrete time representation of the system. The method under study in this thesis was developed for step input, then we can rely on the discrete time representation. There exist corresponding invariant transformations for piecewise linear inputs and for piecewise polynomial inputs that lead to exact representations from first and second order sampling of the system response, respectively.  }

	\item  Related to earlier questions about compressed sensing. There are some recent results on the connections between sparsity and low-rank Hankel matrix completion and approximation. Say that you have an issue with missing data, e.g., from sensor outages. Would it be possible to plug in a procedure that does structured matrix completion and approximation into your methods?
\end{itemize}

\section*{Roger Vounckx}

\begin{itemize}
	\item Two practical questions. On page 35-49 you consider a weight of 138.x grams. You sample relatively slowly (4 kHz) and after 500 samples we can consider that we have correct estimation. If I have a container that fills up and weight is constantly changing. Can we tackle this?
	\item  Consider an optical fiber link with a detector capable of 100 million measurements per second. Can I speed up the link by using your method? You have to do calculations: how high can you go in speed before running into troubles with DSP?
\end{itemize}

\section*{Rik Pintelon}

\begin{itemize}
	\item Returning to a question by Guillaume Mercere: on page 17; do you need normality? You made 2nd order taylor series expansion. Which moment do you need to estimate? Look at the equations page 16. There are 3rd, 4th order moments,... you need to know something more about the noise; this is one of the reasons to make Gaussian assumption: 3rd order moment is 0 (symmetry); 4th order for Gaussian is easy to have an explicit relationship.

    {\bfseries Section 4.1 has been appended with a paragraph to discuss on the need for the normality assumption.  }

	\item  Returning to a question by Philippe Dreesen on missing data. If you can deal with missing data, you can deal with saturation in sensors: you just remove the saturations. What is the rank of the Hankel matrix? Is it full rank or not? The question is if matrix completion can help? It works in low-rank approximations; not if the matrix is of full rank.
\end{itemize}

\section*{Ivan Markovsky}

\begin{itemize}
	\item About scalability of algorithms: You said in the very beginning that it can be very fast compared with system identification. This is the point: can you think of a good answer for these questions on scalability? Nikos Deligiannis mentioned computational analysis. You can directly say for (R)LS what is the computational complexity. Order three if full matrix inversion, now per iteration only linear. So you can scale to very large data sizes.

	{\bfseries The last paragraphs of Section 3.2 have been included to discuss the scalability of the data-driven step input estimation method.  }

\end{itemize}

\section*{Detailed comments provided by jury members}

\section*{Guillaume Mercere}

\begin{itemize}
	\item It is important for me to know why these new developments are generated. That is the reason why I suggest adding sections or paragraphs dedicated to
    \begin{itemize}
        \item the reasons why, in the industry, we need such a new technique. As I said during the defense, do you have a list of sensors for which waiting for the steady regime for getting the measurement is not possible? Why do you need a fast algorithm? What are the reasons why industry people need fast algorithms? In the end, it seems that, for you, it is more important to have a fast biased result than an unbiased estimate. Why? What do you mean by fast?
        \item  the literature on EIV estimation problems. As shortly mentioned in Chapter 2, there are solutions in the literature (by the way, you never mention the recent book written by Torsten Soderstrom in your document which is strange for me). But you do not describe them (even shortly) and, more importantly, you do not criticize them. Not in general but, at least, explain why you do not use them, what are their advantages and drawbacks and, finally, w.r.t. this list of positive and negative points, why you choose a (R)LS solution. E.g., why not using a recursive instrumental variable solution instead?

        {\bfseries A description of IV methods has been added in Section 4.1.4}

        \color{blue}
    \subsection*{4.1.4. Instrumental variables formulation of the data-driven step input estimation method}

    The errors-in-variables problem formulation (3.18) of the data-driven step input estimation method can be converted into an instrumental-variables problem.
    A guide to do this conversion is found in \cite{Soderstrom18} and in \cite{Pintelon12Book}. 
    The instrumental variable $\widetilde{\mathbf{Z}}$ has to be selected so that the normal equations (3.19) are expresed instead as
    \begin{equation} \tag{4.22} \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{y}} . \label{eqn:neq_siv} \end{equation}
    The selection of the instrumental variable has to be done in such a way that is correlated the regression matrix $\widetilde{\mathbf{K}}$, and uncorrelated with the perturbations $\bm{\epsilon}$.
    Assuming that the block-Hankel matrix $\widetilde{\mathbf{K}}$ is given by
    \begin{equation} \tag{4.23} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(n+1) & \Delta \widetilde{y}(n+2) & \cdots & \Delta \widetilde{y}(2n) \end{bmatrix} , \label{eqn:matrixK_r} \end{equation}
    then, the instrumental variable $\widetilde{\mathbf{Z}}$ should not contain the elements that construct $\widetilde{\mathbf{K}}$, i.e., the elements of the vector $\left( \widetilde{y}(1), \ldots, \widetilde{y}(2n+1) \right)^\top$.
    A possibility to generalize this idea is expressing the instrumental variable as
    \begin{equation} \tag{4.24} \widetilde{\mathbf{Z}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-4n) & \Delta \widetilde{y}(t-4n+1) & \cdots & \Delta \widetilde{y}(t-3n-1) \\ G & \Delta \widetilde{y}(t-4n+1) & \Delta \widetilde{y}(t-4n+2) & \cdots & \Delta \widetilde{y}(t-3n) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-3n) & \Delta \widetilde{y}(t-3n+1) & \cdots & \Delta \widetilde{y}(t-2n-1) \end{bmatrix} , \label{eqn:matrixZ_t} \end{equation}
    for a regression matrix
    \begin{equation} \tag{4.25} \widetilde{\mathbf{K}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-2n+1) & \Delta \widetilde{y}(t-2n+2) & \cdots & \Delta \widetilde{y}(t-n) \\ G & \Delta \widetilde{y}(t-2n+2) & \Delta \widetilde{y}(t-2n+3) & \cdots & \Delta \widetilde{y}(t-n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-n+1) & \Delta \widetilde{y}(t-n+2) & \cdots & \Delta \widetilde{y}(t) \end{bmatrix} , \label{eqn:matrixK_t} \end{equation}
    for $t = 4n+1, \ldots, N$.

    The difference with the EIV formulation is that this implementation of the IV method requires to wait more, until $4n+1$ samples of the sensor response are read, instead of $2n+1$ samples. 
    For sensors of small order this twice-the-time waiting does not represent a big change, but as the sensor order increases, the need for waiting can turn the method into impractical since a lot of time should pass before starting to estimate the input.

    There exists a generalized instrumental variables method (GIVE) that was proposed to identify dynamic models from input/output data. 
    GIVE is a bias-eliminating method that uses a generalized instrumental vector constructed from delayed values of the input and output signals \cite{Soderstrom18}.
    Since the aim of this thesis is to analize a method that directly estimates the input from the observed sensor output, without identifying explicitly a dynamic model of the sensor, the GIVE method is out of the scope of this work. 
    A future research can include studying the formulation of a data-driven input estimation method in the spirit of the GIVE method. 
    Nevertheless, a brief description of the GIVE method and its recursive version is given in the following paragraphs.

    \subsubsection*{4.1.4.1 Generalized instrumental variables method}
    For an errors-in-variables problem expressed as
    \begin{equation} \tag{4.26} \widetilde{y}(t) = \widetilde{\bm{\phi}}(t) \bm{\theta}   \end{equation}
    where $\widetilde{y}(t)$ is the output perturbed by measurement noise, $\bm{\theta}$ is the vector of to-be-estimated parameters, and $\widetilde{\bm{\phi}}(t)$ is the regressor vector constructed from delayed samples of the measured input $\widetilde{u}(t)$ and output $\widetilde{y}(t)$.
    The least-squares estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is the solution of the normal equation
    \begin{equation} \tag{4.27} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} \widehat{\bm{\theta}}_{\mathrm{LS}} = \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} , \label{eqn:norm} \end{equation}
    where $\widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}}$ is the covariance matrix of the regressor vector, and $\widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}}$ is the cross-covariance matrix of the regressor vector and the output.
    When the number of samples becomes infinitely large, these covariances matrices can be expressed as a sum of the contributions from exact data and measurement noise contributions
    \begin{equation} \tag{4.28} \begin{aligned}  \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} &= \mathbb{E} \{ \widetilde{\bm{\phi}} \widetilde{\bm{\phi}}^\top \} = \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \\ \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} &= \mathbb{E}\{ \widetilde{\bm{\phi}} \widetilde{y}^\top \} = {\mathbf{r}}_{\bm{\phi} y} + {\mathbf{r}}_{\bm{\eta} \bm{\epsilon}} , \end{aligned} \end{equation}
    where we are assuming that $\widetilde{y} =  y +\bm{\epsilon}$, $\widetilde{\bm{\phi}} = \bm{\phi} + \bm{\eta}$, and $\mathbf{r}_{\bm{\phi} y} = \mathbf{R}_{{\bm{\phi}}} \bm{\theta}$.
    The bias of the least-square solution can be expressed as 
    \begin{equation} \tag{4.29} \mathbf{R}_{\widetilde{\bm{\phi}}} \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) = \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{R}_{\widetilde{\bm{\phi}}} \bm{\theta} = \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} - \mathbf{R}_{\bm{\eta}} \bm{\theta} , \end{equation}

    If we assume that the measurement noise is white for both the input and output, 
    \begin{equation} \tag{4.30} \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} = \mathbb{E}\{ \bm{\eta} \bm{\epsilon}^\top \} = \mathbf{0} \quad \mathrm{and} \quad \mathbf{R}_{\bm{\eta}} = \begin{bmatrix} \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{n_y} & \mathbf{0} \\ \mathbf{0} & \sigma_u^2 \mathbf{I}_{n_u} \end{bmatrix}, \end{equation} 
    where $\sigma_{\bm{\epsilon}}^2$ and $\sigma_u^2$ are the variances of the output and the input, respectively. The normal equation (\ref{eqn:norm}) implies that 
    \begin{equation} \tag{4.31} \left( \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \right)  \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) \neq \mathbf{0}, \end{equation}
    and the estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is not consistent.
    Instead, a bias-compensating least squares estimation can be obtained with  
    \begin{equation} \tag{4.32} \widehat{\bm{\theta}}_{\mathrm{BCLS}} = \left( \mathbf{R}_{\widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\eta}} \right)^{-1}  \left(  \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{r}_{\bm{\eta} \bm{\epsilon}} \right) . \end{equation}

    To formulate the generalized instrumental variable estimator, a generalized instrumental vector $\mathbf{z}(t)$ is built from delayed samples of the input and the output. 
    The dimension of the generalized instrumental vector should be $n_{\mathbf{z}} \geq \mathrm{dim} ( \vartheta ) = n_u + n_y + 2$, since the total parameter vector $\vartheta = \left( \theta^\top, \rho^\top \right)^\top$ considers also the variances of the measurement noise $\rho = \left( \sigma_{\bm{\epsilon}}^2, \sigma_u^2 \right)^\top$. 
    Since the generalized instrumental vector $\mathbf{z}(t)$ and the equation error $\widetilde{y}(t) - \widetilde{\bm{\phi}}(t) \bm{\theta}$ must be correlated, we can express the sistem of equation as 
    \begin{equation} \tag{4.33}\left( \mathbf{R}_{\widetilde{\mathbf{z}} \widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \right) \bm{\theta} = \left(  \mathbf{r}_{\widetilde{\mathbf{z}} \widetilde{y}} - \mathbf{r}_{\bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) \right) , \label{eqn:GIVE} \end{equation}
    where $\widetilde{\mathbf{z}} = \mathbf{z} + \bm{\zeta}$.

    \subsubsection*{4.1.4.2 Recursive generalized instrumental variables method}
    If we define
    \begin{equation} \tag{4.34} \begin{aligned} \mathbf{h} \left( \bm{\rho}, \bm{\theta} \right)  &=  \mathbf{r}_{ \bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \bm{\theta}  =  \mathbf{J} \left( \bm{\theta} \right) \bm{\rho}, \\
    \mathbf{R} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{\bm{\phi}}^\top \left( \tau \right) , \\     
    \mathbf{q} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{y} \left( \tau \right) \\     
    \mathbf{b} \left( t \right)  &=  \mathbf{q} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right)
    , \end{aligned} \end{equation}
    then the equation (\ref{eqn:GIVE}) is equivalent to
    \begin{equation} \tag{4.35} \mathbf{R} \left( t \right) \bm{\theta} = \mathbf{b} \left( t \right) . \end{equation}
    This representation allows a recursive form
    \begin{equation} \tag{4.36} \begin{aligned} 
    \mathbf{R} \left( t+1 \right)  &= \mathbf{R} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{\bm{\phi}}^\top \left( t+1 \right) - \mathbf{R} \left( t \right) \right), \\
    \mathbf{b} \left( t+1 \right)  &= \mathbf{b} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{b} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right) \right), \\
    \widehat{\bm{\theta}} \left( t+1 \right)  &=  \left( \mathbf{R}^\top \left( t+1 \right) \mathbf{R} \left( t+1 \right) \right)^{-1} \mathbf{R}^\top \left( t+1 \right) \mathbf{b} \left( t+1 \right) \\
    \mathbf{q} \left( t+1 \right)  &= \mathbf{q} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{q} \left( t \right) \right) \\
    \widehat{\bm{\rho}} \left( t+1 \right)  &=  \mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right) \left( \mathbf{q} \left( t \right) - \mathbf{R} \left( t \right) \widehat{\bm{\theta}} \left( t \right) \right) , \end{aligned} \end{equation}
    where $\mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right)$ is the pseudo-inverse of $\mathbf{J} \left( \bm{\theta} \right)$.

    The complexity of this recursive generalized instrumental variables estimator is ${O}(n^3)$, that is larger than the ${O}(n^2)$ of the recursive least squares.
    Another drawback of the recursive generalized IV estimator is that it computes a pseudo-inverse matrix in each iteration step.
    If we would need to formulate the data-driven step input estimation problem in the generalized IV framework, it would require to study the formulation in detail because currently the IV framework is based on the assumption that both measurements of the input and output are available, but for the step input estimation problem only the output is available.
    Therefore, it is not straightforward to implement the recursive generalized IV estimator to estimate the step input, and this can lead to future research.
    \color{black}

    \end{itemize}
    \item In the document, you always talk about recursive implementations of least squares solution but, in the end, you never give access to any recursive algorithm description (as you do for what you call the ML method). Why? Do it as well for the recursive weighted least squares solution.

    {\bfseries The following recursive least-squares algorithm description has been added to Section 3.2:}

    \color{blue} 
    The problem (3.17) admits a least-squares (LS) solution with normal equations given by
    \begin{equation} \tag{3.19} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}}.   \end{equation} % \label{eqn:neq_seiv} 
    The closed form of the LS solution can be expressed as
    \begin{equation} \tag{3.20} \widehat{\bm{\theta}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
    where $\widetilde{\mathbf{K}}^\dagger$ is the pseudo-inverse matrix of $\widetilde{\mathbf{K}}$.
    As the number of samples $N$ increases, the number of rows of the matrix $\widetilde{\mathbf{K}}$ grows, making the pseudo-inverse computation inefficient since it requires larger flops, memory and time.

    For metrology applications, it is desired to have a fast solution that can be obtained in real-time. 
    The recursive algorithm (RLS) is a least-squares alternative to implement the step input estimation method in real-time.
    The RLS recursively updates the solution of the system of equations, for each newly acquired sensor response sample, considering the previous value of the estimation, instead of performing a matrix inversion.
    An expression of the RLS equations is given in \cite{Kailath00book} as
    \begin{equation} \tag{3.21} \begin{aligned} \widehat{\bm{\theta}}(t) &= \widehat{\bm{\theta}}(t-1) + \bm{\kappa}_{t} \left( \widetilde{y}(t) - \widetilde{k}_{t} \widehat{\bm{\theta}}(t-1) \right) , \\  \bm{\kappa}_{t} &= \bm{\Psi}(t-1) \widetilde{k}_{t}^\top / \left( 1 + \widetilde{k}_{t} \bm{\Psi}(t-1) \widetilde{k}_{t}^\top  \right) \\ \bm{\Psi}(t) &= \left( \mathbf{I} - \bm{\kappa}_{t} \widetilde{k}_{t} \right) \bm{\Psi}(t-1), \label{eqn:RLS} \end{aligned} \end{equation}
    where $\widetilde{k}_{t}$ represents the row of $\widetilde{K}$ that corresponds to the $t\mathrm{-th}$ sample, $\bm{\kappa}_{t}$ is a gain scalar, $\Psi(t)$ is a covariance matrix, and the estimates and the covariance matrix are initialized respectively with $\widehat{\bm{\theta}}(1)= \widehat{\bm{\theta}}(0)$ and $\Psi(-1) = \Psi(0)$.
    Since the RLS is a recursive implementation of the least-squares computation, the RLS solution is equivalent to the LS solution.
    Moreover, the statistical analysis of the LS solution is valid also for the RLS solution, and is preferred since it is simpler to express the statistics using the closed form of the LS solution.    
    \color{black}

    {\bfseries The following description for the exponentially recursive least-squares algorithm has been added to Section 6.2.1:}

    \color{blue}
    The exponentially weighted recursive least-squares (RLS) can solve recursively the estimation problem (3.20).
    In \cite{Kailath00book} is described the exponentially weighted RLS algorithm, which is as follows

    \begin{equation} \tag{6.4} \begin{aligned}  \widehat{\bm{\theta}}(k) &= \widehat{\bm{\theta}}(k-1) + \bm{\kappa}_{k} \left( \widetilde{y}(k) - \widetilde{\mathbf{k}}_{k} \widehat{\bm{\theta}}(k-1) \right) , \\  \bm{\kappa}_{k} &= \omega^{-1} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top / \left( 1 + \omega^{-1} \widetilde{\mathbf{k}}_{k} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top  \right) \\ \bm{\Psi}(k) &= \omega^{-1} \left( \mathbf{I} - \bm{\kappa}_{k} \widetilde{\mathbf{k}}_{k} \right) \bm{\Psi}(k-1), \label{eqn:RLS} \end{aligned} \end{equation}
    for $k = 2n+1, 2n+2, \ldots$, where $\widetilde{\mathbf{k}}_{k}$ represents the row of $\widetilde{\mathbf{K}}$ that corresponds to the $k\mathrm{-th}$ sample, $\bm{\kappa}_{k}$ is a gain scalar, and $\Psi(k)$ is a covariance matrix.
    The estimates and the covariance matrix are initialized using the first $n+1$ samples, i.e.,  
    \begin{equation} \tag{6.5} \begin{aligned} \widehat{\bm{\theta}} (2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top  \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega}  \widetilde{\mathbf{y}}_{n+1}, \quad \mathrm{and} \\ \Psi(2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} , \end{aligned} \end{equation} 
    where $\widetilde{\mathbf{K}}_{n+1}$ is the matrix $\widetilde{\mathbf{K}}$ with the first $n+1$ rows, and $\widetilde{\mathbf{y}}_{n+1}$ is the vector $\widetilde{\mathbf{y}}$ with the first $n+1$ elements.
    When $\omega=1$, the exponentially weighted RLS solution is identical to that of the RLS.
    When $\omega<1$, the older residuals are weighted with lower values than the residuals of recent observations.
    In this way, the solution of the minimization problem depends more on newer data, and less on older data.

    \color{black}

    \item Double check your equations page 12 and 13. By the way, can you explain where Eq. (3.6) comes from?

    {\bfseries The equations in page 12 and 13 have been reviewed and typos corrected. The derivation of Eq. (3.6) has been improved with more explaination: }

    \color{blue}
    A third method can directly estimate the input step level from the step response, without identifying a sensor model.
    To derive this method, we use the first difference operator, $\Delta = \sigma - 1$, where $\sigma$ is the shift operator, defined as $(\sigma^\tau y) (k) = y(k + \tau)$.
    Applying the first difference operator to the system state-space representation (3.1) results in the autonomous system
    \begin{equation} \tag{3.10} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta {y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
    where $\Delta {u}(k) = \mathbf{0}$, for $k \geq 0$, and
    $\Delta \mathbf{x}(0) = (\mathbf{A} - \mathbf{I}) \mathbf{x}(0) + \mathbf{B} {{u}}$.

    If the response $\Delta {y}$  of this autonomous system is persistently exciting of order $L$, then the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ of $L+1$ block rows constructed from $\Delta {y}$ satisfies
    \begin{equation} \tag{3.11} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) = \mathrm{rank} \left( \begin{bmatrix} \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \iddots & \iddots & \iddots \\ \Delta y(L+1) & \Delta y(L+2) & \cdots & \Delta y(L+n) \end{bmatrix} \right) \leq L,  \end{equation}
    and the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ is a linear map to the natural responses of the augmented autonomous system (3.3) 
    \begin{equation} \tag{3.12} \mathbf{y}_{\mathrm{natural}} = \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell}, \end{equation}
    from any vector $\bm{\ell} \in \mathbb{R}^{n}$.
    The total response of the system is the sum of the forced response, due to the applied input ${u}$ and the natural response
    \begin{equation} \tag{3.13} \mathbf{y} = \mathbf{y}_{\mathrm{forced}}+ \mathbf{y}_{\mathrm{natural}} =G {u} + \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell} ,  \end{equation} 
    that is equivalent to
    \begin{equation} \tag{3.14} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(N) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ G & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ G & \Delta y(N-n) & \Delta y(N-n+1) & \cdots & \Delta y(N-1) \end{bmatrix} }_{\mathbf{K}} \underbrace{ \begin{bmatrix} {{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
    where $N$ is the number of samples.
    From this point of view, the vector $\bm{\ell}$ is a linear transformation of the system's initial conditions.  
    \color{black}

    \item  In Section 4.1.1 and 4.1.2, discuss the normality assumption. Please restructure Section 4.1.3 which is not clear for me.

    {\bfseries The discussion of the normality assumption has been added to Section 4.1, since it is relevant for the two cases considered. Section 4.1.3 has been reestructured to improve its clarity.}

    \item  Why do you always study the statistical properties of your least squares solution with high SNR only? What do you get for SNR = 10 dB for instance? 
    \item Give the values of the matrices you use for simulating data sets so that the reader can test what you have carried out.

    {\bfseries The first paragraphs of Section 4.2, and Fig. 4.1, were added to describe the conditions in which the simulations were conducted from the structured and unstructured EIV problems.}

    \color{blue}
    Two Monte Carlo (MC) simulation studies were conducted to test the obtained bias and variance formulas.
    There are two sets of formulas.
    One set was obtained for the unstructured and uncorrelated EIV problem, and, the other set, for the structured and correlated EIV problem. 
    The problem solved by the data-driven step input estimation method (3.20) belongs to the second set.
    The objective of the MC simulations is to compare the results of the bias and variance formulas in these two cases. 
    The comparison is an evaluation of the effect that the structure and correlation have in the problem solution.
    To make a fair comparison, the same exact data and the perturbation variance is used in both types of EIV problems.

    The exact data is the system of linear equations \ref{eqn:ddsiemexd} built by the data-driven step input estimation method.
    To build the system of equations, the processed signal ${y}$ is the exact response of a sensor to the input $0.1s$, where $s$ is the unit step response.  
    The sensor is modeled as a linear time-invariant mass-spring-damper system of order $n = 2$, that is given by 
    \begin{equation} \tag{4.37}
    \mathbf{x}(k+1) = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{m+M} & \dfrac{-k_{\mathrm{d}}}{m+M} \end{bmatrix} \mathbf{x}(k) + \begin{bmatrix} 0\\ -g\end{bmatrix} u(k),  
    \quad y(k) = \begin{bmatrix} -1 & 0 \end{bmatrix} \mathbf{x}(k), \label{eqn:msdst}
    \end{equation}
    where $m=0.015$ kg, $k_{\mathrm{s}}=500$ kg/m, $k_{\mathrm{d}}=0.5$ kg s/m, $M=0.1$ kg, and $g=9.81 \ \mathrm{m/s^2}$.   
    The exact sensor response is shown in Figure 4.1.

    In the simulations, the sample size considered is $N=200$, since there the transient regime of the sensor model is still evident. 
    In the case of the structured EIV problem, this sample size satisfies the requirements of the data-driven step input estimation method. 
    \color{black}

    \item Please remove Fig. 4.1 b from Section 4.2.1 because the top and bottom curves are obtained with different systems and simulation condition. Or explain why you can compare both. It could be interesting to test each time (i.e., in Section 4.2.1 and 4.2.2) the bias and covariance formulas obtained by considering the structured and unstructured EIV problems.

    {\bfseries Fig. 4.1 b has been relocated in its corresponding section. Now it is Fig. 4.7.}

    \item  When you generate different simulation examples, explain why the new ones are interesting or, more importantly, what they bring w.r.t. the former simulations results. This is the case for instance when I compare Section 4.2.2 and Section 5.1. 
    \item Except if you can convince me, do not use the name "maximum likelihood" when you do not introduce a likelihood you want to maximize. In Chapter 6, you introduce a nonlinear least squares algorithm, nothing else (see the attached document, Numerical Optimization, Chapter 10).

    {\bfseries The method description in Section 6.2.2 has been modified to explain why the method is a maximum likelihood method. The modified paragraphs are as follows: }

    \color{blue} 
    Using a model of the sensor, the maximum-likelihood ML method simultaneously estimates the parameters of the applied affine input $u(t) = at + b$, and some of the sensor model parameters, using the observed sensor response $\widetilde{\mathbf{y}}$.
    This method fits iteratively the response of the model $\widehat{\mathbf{y}}(\bm{\theta})$ to the sensor response by searching for the parameters $\bm{\theta}$ that reduce the error difference between both responses.
    Algorithm 6.1 describes the formulation of an objective function $f = \mathbf{r}^\top \mathbf{r}$, where $\mathbf{r}$ is the residual $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$, that is minimized through the iterations.
    The objective function is the sum of the squares of the differences between the samples of both responses.
    Since the measurement noise is assumed Gaussian distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$, where the samples of this perturbation are independent and identically distributed, the minimization of the objective function $f$ maximizes the likelihood of fitting the actual sensor response with the model response.

    The Jacobian matrix of the residual $\mathbf{J}\bm{\theta} = \partial \mathbf{r}(\bm{\theta}) / \partial \bm{\theta}$ can be analytically derived using the sensor model.
    With the Jacobian matrix, the ML method searches the direction in which the residual decreases towards a local minimum.
    Unfortunately, this ML method cannot guarantee the global minimum since it depends strongly on the initialization of the optimization parameters, as it is discussed in \cite{Nocedal06}.
    Therefore, it is proposed to initialize the to-be-optimized parameters $a$ and $b$ using the results of the subspace method obtained after a convenient number of samples so that the affine input parameters are set close to the optimal values.
    The initialization of the remaining model parameters under study can be done using the sensor model representation.
    An illustrative example of the ML will be provided in a posterior subsection.  
    \color{black}
\end{itemize}

\section*{Nikos Deligiannis}

\begin{itemize}
    \item Comments related to the structure of the thesis:
    \begin{itemize}
	\item In the introduction chapter it is best to describe a set of applications and settings that motivate your research and proposed solutions. This can make your thesis more accessible. Furthermore, you can add a subsection that describes in a high-level manner the remainder and structure of the thesis (describing briefly the contributions of each chapter). You can also mention your contributions there but in a less technical way since the problem statement only comes in the Chapter preliminaries.
	\item In each main chapter (Chapter 4, Chapter 5, Chapter 6) you can add an introduction subsection, a state-of-the-art subsection and then elaborate your contributions and methods.

    {\bfseries The introduction of the chapters now includes the state of the art to position the PhD work with respect to the literature.}

	\item I believe that this structure will improve the understanding of the thesis and light up your contributions.

	{\bfseries Indeed, this structure allows to visualize the thesis contributions.}

	\end{itemize}
	\item Further comments:
	\begin{itemize}
	\item Please discuss the complexity and scalability issues related to your approach in relation to the state of the art.

    {\bfseries The following paragraph has been added in Section 3.2 to discuss the scalability of the data-driven step input estimation method.}

	\color{blue}
    The computational complexity of the RLS algorithm solution to the system of equations (3.14) is $O \left( \left( n+1 \right)^2 \right)$.
    The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
    From there on, the RLS algorithm updates the solution with linear complexity.
    Therefore, even though a large number of samples $N$ makes the matrix in the system of equations (3.14) increase in the number of rows, the estimation of the solution does not increase in complexity. 
    The data-driven step input estimation method is then scalable for any sensor of order $n$, and can be executed in devices with limited computational resources, provided that the computation of the  $n+1 \times n+1$ inverse matrix is feasible.
    \color{black}

	\item Position your work with regard to the state of the art as well as regarding solutions based on low-rank matrix completion and compressed sensing.
	\item I recommend to use the formulation “we propose” than “I propose”

	{\bfseries The formulations similar to “I propose” were replaced by equivalents such as “we propose”. }

	\item You might wish to add a list of notation in the beginning of the thesis.

	{\bfseries A nomenclature list is provided before the summary of the document. }

	\end{itemize}
\end{itemize}

\section*{Philippe Dreesen}

\begin{itemize}
    \item General remarks:
    \begin{itemize}
    \item  It would be good to include some layman’s examples that motivate the work of the thesis already in the introduction. In the current text, there is a nice example on page 52, almost at the end of the thesis. These kinds of examples should appear in the introduction to motivate and position the work.

    {\bfseries The following paragraphs have been added to the Introduction: }

    \color{blue}

    Take for example the case of a thermometer.
    To measure temperature, the thermometer is put in contact with the substance or object of interest.
    Accordingly to the Newton's law of cooling, the temperature $y$ of the thermometer changes proportionally to the its difference with respect to the temperature $u$ of the matter under study:
    \begin{equation} \tag{2.1}  \dfrac{dy}{dt} = k_T \left( u - y \right) \label{eqn:NewtonCooling} \end{equation}
    where the proportionality constant $k_T$ is a parameter of the thermometer first order model.
    The input $u$ is unknown and its estimation is the objective of the measurement.
    It is customary to take the reading from the thermometer once the temperature stops changing.

    Another popular example of a measurement is found in weighing. 
    In most of modern scales, the object of interest if placed on top of a weighing sensor.
    A simple model of the weighing sensor is the second order mass-spring-damper model shown in Figure \ref{fig:simple_msd_system}.
    In this model, the unknown mass $u$ causes a displacement $y$ of the scale and the dynamics of the sensor are described by the differential equation:
    \begin{equation} \tag{2.2} \dfrac{d}{dt} \left( u(t) \ \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = u(t) \ g \end{equation}
    where $m$ is the own mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.

    For illustration purposes, Figures \ref{fig:thermometer} and \ref{fig:scale} show the responses of the thermometer and the weighing sensors, respectively, after the application of a constant input $u$, which can be represented by a step input.

    \renewcommand{\thefigure}{2.1}
    \begin{figure}[htb!]
    \centering

    \begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
    \tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
    \tikzstyle{damper}=[thick,decoration={markings,  
    mark connection node=dmp,
    mark=at position 0.5 with 
    {
    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
    }
    }, decorate]
    \tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]

    \node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
    \node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u(t)$};

    \node (ground1) at (M.south) [ground,yshift=-1.5cm,xshift=-0.625cm,anchor=north] {};
    \draw (ground1.north west) -- (ground1.north east);
    \draw [spring] (ground1.north) -- ($(M.south east)!(ground1.north)!(M.south west)$);

    \node (groundc) at (M.south) [ground,yshift=-1.5cm,anchor=north] {}; 
    \draw (groundc.north west) -- (groundc.north east);

    \node (ground2) at (M.south) [ground,yshift=-1.5cm,xshift=0.625cm,anchor=north] {};
    \draw (ground2.north west) -- (ground2.north east);
    \draw [damper] (ground2.north) -- ($(M.south east)!(ground2.north)!(M.south west)$);

    \node[draw=none,fill=none] at (-0.9cm,-1cm) {$k_{\mathrm{s}}$};
    \node[draw=none,fill=none] at (0.15cm,-1cm) {$k_{\mathrm{d}}$};
    \node[draw=none,fill=none] at (2.0cm,1.0cm) {$y$};
    \draw [-latex,thick]  ++(2.2cm,-1cm) -- +(0cm,2.25cm);

    \draw [-latex,thick] (M.east) ++(0,0) -- +(1cm,0);
    \draw [line width=0.25mm] (2.2cm,-1cm) -- (2.2cm,1cm);
    \draw [line width=0.25mm] (2.1cm,-1cm) -- (2.3cm,-1cm);
    \draw [line width=0.25mm] (2.1cm,1cm) -- (2.3cm,1cm);
    \draw [line width=0.25mm] (2.1cm,-0.5cm) -- (2.3cm,-0.5cm);
    \draw [line width=0.25mm] (2.1cm,0.5cm) -- (2.3cm,0.5cm);
    \draw [line width=0.25mm] (2.15cm,-0.25cm) -- (2.25cm,-0.25cm);
    \draw [line width=0.25mm] (2.15cm,0.25cm) -- (2.25cm,0.25cm);
    \draw [line width=0.25mm] (2.15cm,-0.75cm) -- (2.25cm,-0.75cm);
    \draw [line width=0.25mm] (2.15cm,0.75cm) -- (2.25cm,0.75cm);
    \draw [line width=0.25mm] (2.1cm,0cm) -- (2.3cm,0cm);

    \end{tikzpicture}

    \caption{\label{fig:simple_msd_system}  \color{blue} A second order mass-spring-damper model represents the weighing sensor. The application of the mass $u$ causes a change in the position $y$ of the scale. \color{black} } 
    \end{figure}

    \renewcommand{\thefigure}{2.2}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_2.pdf} 
    \caption{ \label{fig:thermometer} 
    \color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_T=0.5 s^{-1}$, caused by applying a step input of 67 $\circ \mathrm{C}$  when the thermometer was initially at 23 $\circ \mathrm{C}$. In this example, we should wait 10 s to have an accurate estimation of the input. \color{black} }
    \end{figure}

    \renewcommand{\thefigure}{2.3}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_3.pdf} 
    \caption{ \label{fig:scale} 
    \color{blue} This simulation illustrate the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium. The sensor parameters are $k_{\mathrm{d}} = 5 \mathrm{N s/m}$, and $k_{\mathrm{s}} = 10000 \mathrm{N/m}$. In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement. \color{black}  }
    \end{figure}

    \color{black}

    \end{itemize}
    \item Detailed remarks:

    {\bfseries The detailed remarks in this list have been attended accordingly: }

    \begin{itemize}
    \item Citations that are used in-line should have brackets only around the year, not around the authors’ names. For instance on page 3, line -2 is written “reviewed in [Hack and ten Caten, 2012]”. This should be displayed as “reviewed in Hack and ten Caten [2012]. In LaTeX this can typically be done by making the proper distinction between the commands \ citet and \ citep. The former is used for references in text, the latter for parenthetical references.
    \item  On page 1 of the PDF (front of thesis), I believe also N. Deligiannis should be mentioned with the title “prof.”. 
    \item Page 1 of thesis: typo on line 6: “and on the sensor initial conditions” => “sensor’s initial conditions”
    \item Page 3, line -2: “uncertainty analysis are reviewed” => “is reviewed”
    \item Page 3, line -1: remove comma before “that”
    \item Page 3, line -2: fix the in-text reference (see above)
    \item Page 4, lines 11-13: fix the in-text references (see above)
    \item Page 5, line 5: Remove “The” at the beginning of the sentence (The Hankel…)
    \item Page 6, line 10: fix the in-text references (see above)
    \item Page 7, lines 10-11: the sentence on heart and breathing physiological monitoring sounds a bit awkward. Please reformulate.
    \item Page 9, line -1: fix the in-line reference (see above)
    \item Page 10, lines 1-3: in the enumeration of the types of sensors, remove the article ‘the’ before the sensor name. So: “like three-axis sensors [ref], radio-frequency intruder sensors [ref], and radar sensors [ref]. 
    \item Page 10, equation 3.10: The symbol D was not bold in the equations above. Be consistent!
    \item Page 10, caption of figure: replace “reverts” by “inverts”
    \item Page 11, line 9, put “poles” between parentheses instead of between commas.
    \item Page 12, line -6: “that in matrix form is” => “which in matrix form is”
    \item Page 13, line -9: “in the practice” => “in practice”
    \item Page 14, line -14: missing space between “problem” and “(3.11)”
    \item Page 16, halfway page: the sentence “Applying a second-order…” is not a proper sentence (perhaps fix this by removing “that”).
    \item Page 16 introduces seemingly some new notation, e.g., b(.), mu(.) and C(.). While their meaning is clear from the context, it is desirable to give proper definitions when introducing new mathematical tools.
    \item Page 17, line -1: fix in-text references (see above)
    \item Page 20 (and beyond): The abbreviation CRLB is introduced, but sometimes CRB is used instead.
    \item Page 21, line -10: fix in-text reference (See above)
    \item Page 35, second sentence “The sensor is a dynamic…” is not a proper sentence.
    \item Page 50: Try to reorganize this section into two or more paragraphs.
    \item Page 59, line 2: fix in-text reference (See above)
    \item Page 74, line 11: “The learning obtained from…” this is an awkward sentence; please reformulate
    \item Page 81, all three journal publication references do not have (or have wrong?) page numbers.
    \end{itemize}
\end{itemize}

\bibliography{../Gus-thesis/Gus-thesis.bib}

\end{document}

### 2020-08-21 internal defense feedback/Answers-jury-members-report BK.tex
Second draft before public defense

--- 2020-08-21 internal defense feedback/Answers-jury-members-report - BK2.tex
@@
 
-	{\bfseries In Section 4.1.4, a discussion that compares IV methods with EIV methods has been added.}
+	{\bfseries In Section 4.1.4, a discussion that compares IV methods with EIV methods has beeen added.}
 
@@
 	{\bfseries The measurementy noise variance is prior information and in Chapter 5 we describe two ways of estimating $\sigma_{\bm{\epsilon}}$ In summary, the measurement noise can be extracted fom the sensor ressponse once the transient effects are smaller than the noise floor, or from the sensor response power spectrum.
-The online estimation of the measurement noise, simultaneously with the online step input estimation is a topic for future research.
+The online estimation of the measurement noise, simultaneously with the online step input estimation is a topic for future reasearch.
 }

### 2020-08-21 internal defense feedback/Answers-jury-members-report.tex
Second draft before public defense

--- 2020-08-21 internal defense feedback/Answers-jury-members-report - BK2.tex
@@
     \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method.
-After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and everyone of the estimates.
+After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and every of the estimates.
 The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%.
@@
     Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
-    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance, which can be disregarded since $3 \sigma_{\bm{\epsilon}}^4 \ll \sigma_{\bm{\epsilon}}^2$\color{black}.
+    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance\color{black}.
 
@@
 Can we tackle this?
+
+	{\bfseries Yes, that is exactly the idea considered by the affine input estimation method.
+Since the data-driven estimation method can track the changes of the input, the parameters of an input model that describes the constant variation can be recovered if we add the forgetting factor to the recursive solution algorithm.
+}
+
 	\item  Consider an optical fiber link with a detector capable of 100 million measurements per second.
@@
 You have to do calculations: how high can you go in speed before running into troubles with DSP?
+
+	{\bfseries The optical fiber link transmits digital pulses.
+Since the period would be 1 ns, the step input method implementation would require to sample in much shorter time.
+Assuming that the duty cyce is 50 \%, and that the number of samples required for the step detection are 200, then the sampling time should be 1.25 fs or lower.
+This could be a constraint for the commutation circuits and we would be reaching a feasibility limit.
+However, theoretically the method would perform detecting the pulses level changes.
+}
+
 \end{itemize}
@@
     \begin{equation} \tag{4.22} \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{y}} . \label{eqn:neq_siv} \end{equation}
-    The selection of the instrumental variable has to be done in such a way that is correlated the regression matrix $\widetilde{\mathbf{K}}$, and uncorrelated with the perturbations $\bm{\epsilon}$.
+    The selection of the instrumental variable has to be done in such a way that is correlated with the regression matrix $\widetilde{\mathbf{K}}$, and uncorrelated with the perturbations $\bm{\epsilon}$.
     Assuming that the block-Hankel matrix $\widetilde{\mathbf{K}}$ is given by
@@
     \begin{equation} \tag{4.34} \begin{aligned} \mathbf{h} \left( \bm{\rho}, \bm{\theta} \right)  &=  \mathbf{r}_{ \bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \bm{\theta}  =  \mathbf{J} \left( \bm{\theta} \right) \bm{\rho}, \\
-    \mathbf{R} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{\bm{\phi}}^\top \left( \tau \right) , \\     
-    \mathbf{q} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{y} \left( \tau \right) \\     
+    \mathbf{R} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau=1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{\bm{\phi}}^\top \left( \tau \right) , \\     
+    \mathbf{q} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau=1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{y} \left( \tau \right) \\     
     \mathbf{b} \left( t \right)  &=  \mathbf{q} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right)
@@
     \item  Why do you always study the statistical properties of your least squares solution with high SNR only? What do you get for SNR = 10 dB for instance?
+
+    {\bfseries The second Taylor series expansion of the inverse matrix $(\mathbf{I} + \mathbf{M})^{-1}$ is valid when the spectral radius of $\mathbf{M}$ is lower than 1.
+The spectral radius depends on the perturbation noise variance.
+As the noise variance increases, for fixed exact data, the SNR decreases and the spectral radius grows, as well as the error introduced by the Taylor series expansion.
+We observe that for large SNR the Taylor series expansion allows us to use the derived expressions that calculate the bias and the variance.
+When the relative error of these preditions increase, below 40 dB for example, the spectral radius is getting closer and eventually going over 1.
+With SNR = 10 dB the Taylor series expansion of the inverse matrix introduces an error unacceptably large.
+}
+
     \item Give the values of the matrices you use for simulating data sets so that the reader can test what you have carried out.
@@
     \color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_T=0.5 s^{-1}$, caused by applying a step input of 67 $\circ \mathrm{C}$  when the thermometer was initially at 23 $\circ \mathrm{C}$.
-In this example, we should wait 10 s to have an accurate estimation of the input.
+In this example, we should wait 10 s to have an estimation of the input with a relative error lower than 0.5 \%.
 \color{black} }
@@
     \caption{ \label{fig:scale} 
-    \color{blue} This simulation illustrate the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium.
+    \color{blue} This simulation illustrates the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium.
 The sensor parameters are $k_{\mathrm{d}} = 5 \mathrm{N s/m}$, and $k_{\mathrm{s}} = 10000 \mathrm{N/m}$.
-In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement.
+In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement with a relative error smaller than 0.72 \%.
 \color{black}  }

### 2020-08-21 internal defense feedback/Answers-jury-members-report_pending.tex
Second draft before public defense

--- 2020-08-21 internal defense feedback/Answers-jury-members-report - BK2.tex
@@
 
-Thank you to the jury members for reviewing the thesis manuscript and for giving suggestions for improvement. 
-The manuscript changes due to the comments are highlighted in \color{blue} blue fontcolor\color{black}.
-
 \section*{General comments}
@@
 \begin{itemize}
-	\item The manuscript text should better position the PhD to related work. Improve the literature review and discuss (and cite) the state of the art.
-
-	{\bfseries The introduction of the chapters now includes the state of the art to position the PhD work with respect to the literature.}
-
-	\item  Add more clearly the motivation for the PhD research work by including motivational examples to the introduction.
-
-	{\bfseries Motivational examples have been added to the Introduction. }
-
-	\item  Add a discussion with comparison of existing EIV methods.
-Give, with the positive and negative aspects in mind, a better justification for why you use a (R)LS method, and not, e.g., (R)IV method.
-
-	{\bfseries In Section 4.1.4, a discussion that compares IV methods with EIV methods has been added.}
 
 	\item  Discuss more clearly the aspects of scalability of the methods and their limitations in terms of model order, size of data sets (e.g., related to the inversion of LS matrices).
-	\item  The equations on pages 12 and 13 seemingly contain errors and should be corrected.
-
-	{\bfseries The equations on pages 12 and 13 have been corrected.}
-
-	\item  The figure shown on page 66 was made with a single experiment; it is desirable to see the (averaged) effect on several experiments.
-
-	{\bfseries A series of 100 runs of the maximum-likelihod estimation method has been simulated and the results have been added to the manuscript in Section 6.3.2.}
 
@@
 Elaborate on SNR levels; I’m also very interested in scalability, so how big the matrices can be.
-	\item  Concerning the scalability to data: eqn 3.6: how scalable is the approach in all aspects (order, data, linked to inversion for LS matrix); can you deal with hundreds, thousands of data?
-
-	{\bfseries The following paragraph has been added in Section 3.2 to discuss the scalability of the data-driven step input estimation method.}
-
-	\color{blue}
-    The computational complexity of the RLS algorithm solution to the system of equations (3.14) is $O \left( \left( n+1 \right)^2 \right)$.
-    The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
-    From there on, the RLS algorithm updates the solution with linear complexity.
-    Therefore, even though a large number of samples $N$ makes the matrix in the system of equations (3.14) increase in the number of rows, the estimation of the solution does not increase in complexity.
-    The data-driven step input estimation method is then scalable for any sensor of order $n$, and can be executed in devices with limited computational resources, provided that the computation of the  $n+1 \times n+1$ inverse matrix is feasible.
-    \color{black}
 
@@
 A more detailed discussion is desirable.
-	\item  On slide 25 (noisy graphs) it would be good to see a statistical analysis (more experiments than one).
-
-    {\bfseries 100 runs of the maximum-likelihod estimation method have been simulated and their results have been added to the manuscript in Section 6.3.2.}
-
-	\color{blue} 
-    The maximum-likelihood (ML) method estimated the affine input parameters from the sensor model transient response.
-    A total of 100 runs of the method were obtained, using different realizations of measurement noise.
-    The measurement noise variance was set to have an SNR of 40 dB.
-    The ML method used the first 50 samples to initialize the optimization variables and updated the variables every $N_{\mathrm{s}} = 5$ samples.
-    Figure \ref{fig:rele_lo_40dB_s10} shows the average of the observed relative errors in the estimation of the parameters $\widehat{a}$, $\widehat{b}$, $\widehat{x}_{\mathrm{ini,1}}$, and $\widehat{x}_{\mathrm{ini,2}}$.
-
-    \renewcommand{\thefigure}{6.8}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_7.pdf} 
-    \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method.
-After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and everyone of the estimates.
-The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%.
-\color{black} }
-    \end{figure}
-
-    Figure \ref{fig:cov_lo_40dB_s1} shows \color{blue} the average \color{black} of the variances that were observed on the diagonal of the covariance matrix $\mathbf{J}$. 
-    We can see that the estimation variances decrease as more samples are processed.
-    Moreover, the estimation variances of $\widehat{a}$ and $\widehat{b}$ obtained with the ML method are lower than the corresponding estimation MSE errors obtained from a Monte Carlo simulation of the subspace method.
-
-    \renewcommand{\thefigure}{6.9}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_8.pdf} 
-    \caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian.
-\color{blue} The estimates $\widehat{x}_{\mathrm{ini,1}}$ and $\widehat{a}$ have the smaller and larger variances, respectively.
-\color{black}  }
-    \end{figure}
-
-    \color{black}
 
@@
 \begin{itemize}
-	\item The Taylor expansion is computed of order 2, what is the accuracy of expansion with respect to system, noise level,...
-Can you come up with a priori bound on error that will allow you to assess the quality of the expansion?
-Can you predict where the solution is going to be before computing Taylor expansion?
-
-	{\bfseries An a priori bound on the error introduced by the Taylor series expansion can be obtained in terms of the spectral radius of the matrix $\mathbf{M}$.
-In Section 4.1 the following description has been added: }
-
-    \color{blue} An a priori bound on the error introduced by the second order Taylor series expansion can be expressed in terms of the spectral radius $\rho(\mathbf{M})$.
-    Considering that the Taylor series expansion is an infinite sum we have that
-    \begin{equation} \tag{4.4} \begin{aligned} (\mathbf{I} + \mathbf{M})^{-1} &= \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \\
-    &=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \right) \\
-    &=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \end{aligned} \end{equation} 
-    Therefore we can obtain a bound for the second error expansion error if we take a matrix norm as follows
-    \begin{equation} \tag{4.5} \begin{aligned} \| (\mathbf{I} + \mathbf{M})^{-1} - ( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 ) \| &=  \| \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \| \\
-    & \leq \| \mathbf{M}^3 \| \| (\mathbf{I} + \mathbf{M})^{-1} \| \leq \dfrac{\| \mathbf{M} \|^3}{ 1 -  \| \mathbf{M} \|} = \dfrac{\rho(\mathbf{M})^3}{1 - \rho(\mathbf{M})}  \end{aligned} \end{equation}
-	\color{black}
-
-	\item  Concerning other sorts of prior estimations, e.g., given by explicit function theorems.
-I wonder if you could get something and apply this bound in order to control the bound of the 2nd order expansion.
-There exist some good results that could help you devise something a priori which can control the error you are making.
-Neuberger's results (quite old now) but may be handy for this kind of problem.
-This can be investigated further.
-
-	\color{blue}
-	Neuberger's paper \cite{Neuberger07} introduces five theorems that are consequence of the continuous Newton method.
-	The Newton method aims to find a zero of the function $F: R^n \rightarrow R^n$ by iterating
-	\begin{equation} z_{k+1} = z_k - \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
-	for $k=0,1,2,\ldots$, and for a given initial $z_0$, assuming that $\left(F' (y) \right)^{-1}$ exists for some $y \in R^n$.
-	A domain of attraction is the set of all initial values $z_0$ that lead to a root of $F$ after the convergence $z_0, z_1, z_2,\ldots$.
-	The damped Newton method  
-	\begin{equation} z_{k+1} = z_k - \delta_k \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
-	prevents convergence issues, such as chaotic domains of attraction, with $\delta_1, \delta_2, \ldots \in (0, 1)$. 
-	The continuous Newton method is a sequence of damped Newton methods.
-To implement the continuous Newton method, select $T>0$ and perfom $m$ runs of the damped Newton method with $\delta_k = T/m$, for $k=1,2,\ldots, m$.
-	The result $z_{k+1}$ is the zero estimation $x_m$, and if the sequence of estimates $x_1, x_2, \ldots$ converges, then it converges to the result of the continuous Newton method.
-
-	The objective of the continuous Newton method is to find a function $z: [0, \infty) \rightarrow R^n$, so that
-    \begin{equation} z(0) = x \in R^n, \quad z' (t) = - \left(F' (z(t)) \right)^{-1} F(z(t)), \quad t \geq 0, \label{eqn:contNewton}\end{equation}
-    subject to the existence of the limit $u=\mathrm{lim}_{t \rightarrow \infty}{z(t)}$, that satisfies $F(u)=0$.
-    This implies 
-    \begin{equation}  F' (z(t)) z' (t) = - F(z(t)) , \end{equation}
-    that is equivalent to
-    \begin{equation}  \left(F \circ z\right)' (t) = - F(z(t)) , \end{equation}
-    from where we get
-    \begin{equation}  F(z(t)) = e^{-t} F(z(0)) . \end{equation}
-    Thus, the residual $F(z(t))$ only changes in magnitude and not in direction.
-    This property is essential for the results of the theorems in \cite{Neuberger07}. 
-    In particular, Theorem 5 gives the conditions for the continuous Newton method to ensure finding a root of the function.
-
-    \begin{customthm}{5} \label{thm:five}
-    Suppose that each of $H$, $J$, and $K$ is a Banach space, with $H$ compactly embedded in $J$, that $r>0$, and that $G:B_r(0) \rightarrow K$ is continuous as a function on $J$.
-Suppose also that $g$ belongs to $K$ and that for each $y$ in $b_r(0)$ there is an $h$ in $B_r(0)$, where $b_r(0)$ and $B_r(0)$ are open and closed balls in $H$ of radius $r$ centered at 0, such that
-    \begin{equation} \lim\limits_{t \to 0+} -{\dfrac{1}{t} \left(G(y+th) - G(y)\right) = g}. \end{equation}
-    Then, there exists $u$ in $B_r(0)$ such that $G(u)=g$.
-    \end{customthm}
-    %$F:B_r(0) \rightarrow K$
-
-    The theorem implies that $u$ is a zero of the function  defined as $F(y) = G(y) - g$.  
-    Since $G$ is a continuous function and if $\left(G' (y) \right)^{-1}$ exists for each $y$ in $B_r(0)$, a metric for the validity of the Theorem (\ref{thm:five}) hypothesis is the inequality $\left\Vert  \left( G' (y) \right)^{-1} g \right\Vert \leq r$.
-
-    To link the data-driven step-input estimation problem with this theorem, we can take the system of linear equations $\widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \bm{\theta}$, then define $G(y) = \widetilde{\mathbf{K}} \bm{\theta}$, and $g = \widetilde{\mathbf{y}}$, so that we have $F(\bm{\theta}) = G(\bm{\theta}) - g$.
-The derivative of the function $G(\bm{\theta})$ is given by $G'(\bm{\theta}) = \widetilde{\mathbf{K}} \mathbf{I}$.
-Using the pseudo-inverse of $\widetilde{\mathbf{K}}$, the metric $\left\Vert  \widetilde{\mathbf{K}}^{\dagger} \widetilde{\mathbf{y}} \right\Vert \leq r$ indicates that the estimate $\hat{\bm{\theta}}$ is found inside a ball of radius $r$.
-    Thus,  
-	\color{black}
 
@@
 \begin{itemize}
-	\item In the first or second sentence: “users cannot wait for a steady state regime” .
-Do you have specific examples of real industrial systems for which it is necessary to have continuous read-out before steady-state is reached?
-- Further on the industrial part and selling sensors with data-driven parts.
-Can you explain to experts that your solution is better?
-Sensors will be used for ages; why not spend one day to do system identification and get a reliable model to make a compensator, instead of using your approach?
-
-	{\bfseries The fast weighing of parcels in the post office is a typical example of the real life systems where the time is limited to perform the measurements.
-Another one that we see nowadays is the temperature measurement of people in airports or train stations to prevent the spread of corona virus.
-The data driven method can be implemented in portable devices with limited computational resources.
-In industrial plants were a time slot can be devoted to identify the measurement system and implement a compensation method, the solution might be useful only for a certain amount of time.
-It is assumed that the model approximation deviates due to changes in the system elements, leading to inaccuracies in the inpiut estimation.
-There is a need for adaptive methods to avoid stopping the production lines to repeat the system identification procedure.
-The data-driven method is such an adaptive method.
-}
 
@@
 Is it still white in this case?
-	\item  On page 12 standard approach top of page equation: there are some mistakes in the equations!
-there is no epsilon (noise) in the equation.
-But then you do an SVD of H and estimate from there.
-How does it work if you have noise?
-Have you tried implementing it and see what you get?
-
-	{\bfseries The step input estimation based on singular value decomposition works only with exact data.
-When  noise perturbes the sensor response the method cannot guarantee the step input estimation.
-The singular value decomposition of a Hankel matrix constructed from a perturbed sensor response provides only an approximation of the singular values and the singular vectors of this Hankel matrix.
-Then, the step input estimation obtained with this method from perturbed sensor response is only a rough estimation.
-}
-
-	\item  On page 12: I’m not sure the equations at the end of the page are correct, e.g., Y=G ubar ...
-not sure there is x; same for rest; There seem to be many mistakes here and next pages.
-Be careful with the equations!
-
-	{\bfseries The equations have been reviewed and the typos were corrected.}
-
-	\item  On page 13, you assume persistency of excitation of order $L$. How do you check this? In your case the input is constant; how to guarantee that the assumption is met?
-
-	{\bfseries The persistency of excitation of a signal is verified by looking at the rank of the Hankel matrix constructed from the consecutive samples of this signal.
-When the Hankel matrix of $L+1$ block rows has a rank lower or equal than L, then we say that the signal is persistently exciting of order $L$.
-A step input has a very low order of persistency $L=1$, but the step respose of a system has a larger order of persistency.
-In fact, we assume that if the persistency of excitation of the step response is sufficiently high, then the columns of the Hankel matrix constructed from the step response span a linear space with all the responses of the equivalent augmented autonomous system where the step input is considered an augmented initial condition.}
-
-	\item  In Chapter 4, p 17, Section 4.1.1: You assume that perturbations are IID and must be normally distributed.
-Why using an assumption of normally distributed?
-I would think that the kind of distribution is important for CRLB, but I’m not sure you need it for eqns 4.8, 4.9, 4.10,...
-IID is important, normality perhaps not.
-
-	{\bfseries The following paragraph has been added in Section 4.1 to discuss on the need for the normality assumption.  }
-
-	\color{blue}
-	The normality assumption is necessary to provide more prior information into the method formulation.
-    The second order Taylor series expansion (4.4) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
-    Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
-    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance, which can be disregarded since $3 \sigma_{\bm{\epsilon}}^4 \ll \sigma_{\bm{\epsilon}}^2$\color{black}.
-
-	\item   In equation 4.11 you have $\sigma_eps^2 + \sigma_e^2$; how to get in practice $\sigma_e$? Do you have to do this each time? Recursively?
-
-	{\bfseries The measurementy noise variance is prior information and in Chapter 5 we describe two ways of estimating $\sigma_{\bm{\epsilon}}$ In summary, the measurement noise can be extracted fom the sensor ressponse once the transient effects are smaller than the noise floor, or from the sensor response power spectrum.
-The online estimation of the measurement noise, simultaneously with the online step input estimation is a topic for future research.
-}
 
@@
 You use only one realization also for matrices A, B, C, D in the text, so it can be added explicitly.
-	\item   in Fig 4.1 the bottom figure also discusses the structured case. Why do you talk about structured results in a section about the unstructured case?
-
-	{\bfseries Section 4.2 has been improved to describe the result next to its discussion.  }
-
-	\item   In 4.1 uses a constant K, but in 4.2.2 you use A, B, C, D, without any links to K.
-You compare the two curves generated from different data sets… It is strange to compare them.
-I don't see the link between both.
-
-	{\bfseries Section 4.2 has been reestructured and the simulation for the unstructured case were redone to establish a link between the two cases.
-The exact data is the same in the two cases and the perturbation noise follows the corresponding rules of each case, with the same variance noise levels.
-In this way the comparison between the two cases makes more sense.
-}
 
@@
 	{\bfseries \color{red} Interesting question, I want to try it this evening \color{black}}
-
-	\item   I was surprised to read in Chapter 5 something very close to what I read in chapter 4.
-In both cases generated A, B, C.
-In 5.1 again oscillating 5th order system, similar conclusions, ...
-What is the added value of 5.1 wrt 4.2.2?
-
-	{\bfseries The derivation described in Chapter 4 was conducted with a second order mass-spring-damper weighing system example on mind, and the results observed were submitted in the corresponding paper.
-Afterwards, when we moved to the practical implementation, we faced the increase of the model order after setting up the weighing experiment.
-The lowest order we can assume is fifth order.
-So we decided to repeat the simulations and these new results serve two purposes: a transition from simulation to experimental results, and also to get an insight on the impact of the bias and variance predictions after increasing the model order.
-}
 
@@
 Have you tried this?
-	\item   In Section 6.2.2.
-Why do you call it a max likelihood approach?
-To me it is not one.
-Why do you call it like this?
-Where is the likelihood?
-Depending on which likelihood (eg.
-Gaussian, ...) it leads to LS.
-Don't' call it ML if you dont give distributions on ...
-In this case it is better to call it a nonlinear minimization procedure.
-Not clear why you call it ML.
-
-	{\bfseries The method description in Section 6.2.2 has been modified to explain why the method is a maximum likelihood method.  }
 
@@
 	\item  You mention that you consider Kalman filters; have you seen approaches where they write Deep Learning in form of Kalman Filter; Could this be applicable?
-	\item  Recommendation: if you revise the thesis: maybe change "I propose" to "we propose".
-
-	{\bfseries The formulation “I propose” was replaced by “we propose”. }
 
@@
 Can you still use your methods?
-	\item  Sensors are physical devices and deliver a continuous-time output. The models that you use are in discrete-time. Are there any risks that you miss information by not sampling fast enough?
-
-	{\bfseries In Chapter 3 has been added a discussion on the step invariance transformation.
-This defines the conditions of the input for describing exactly the system dynamics in discrete time after sampling the continuous system response.
-If the input is stepwise continuous, then the zero order hold sampling of the step response leads to an exact discrete time representation of the system.
-The method under study in this thesis was developed for step input, then we can rely on the discrete time representation.
-There exist corresponding invariant transformations for piecewise linear inputs and for piecewise polynomial inputs that lead to exact representations from first and second order sampling of the system response, respectively.
-}
 
@@
 
-\section*{Roger Vounckx}
-
-\begin{itemize}
-	\item Two practical questions.
-On page 35-49 you consider a weight of 138.x grams.
-You sample relatively slowly (4 kHz) and after 500 samples we can consider that we have correct estimation.
-If I have a container that fills up and weight is constantly changing.
-Can we tackle this?
-	\item  Consider an optical fiber link with a detector capable of 100 million measurements per second.
-Can I speed up the link by using your method?
-You have to do calculations: how high can you go in speed before running into troubles with DSP?
-\end{itemize}
-
 \section*{Rik Pintelon}
@@
 \begin{itemize}
-	\item Returning to a question by Guillaume Mercere: on page 17; do you need normality?
-You made 2nd order taylor series expansion.
-Which moment do you need to estimate?
-Look at the equations page 16.
-There are 3rd, 4th order moments,...
-you need to know something more about the noise; this is one of the reasons to make Gaussian assumption: 3rd order moment is 0 (symmetry); 4th order for Gaussian is easy to have an explicit relationship.
-
-    {\bfseries Section 4.1 has been appended with a paragraph to discuss on the need for the normality assumption.  }
 
@@
 It works in low-rank approximations; not if the matrix is of full rank.
-\end{itemize}
-
-\section*{Ivan Markovsky}
-
-\begin{itemize}
-	\item About scalability of algorithms: You said in the very beginning that it can be very fast compared with system identification.
-This is the point: can you think of a good answer for these questions on scalability?
-Nikos Deligiannis mentioned computational analysis.
-You can directly say for (R)LS what is the computational complexity.
-Order three if full matrix inversion, now per iteration only linear.
-So you can scale to very large data sizes.
-
-	{\bfseries The last paragraphs of Section 3.2 have been included to discuss the scalability of the data-driven step input estimation method.  }
-
 \end{itemize}
@@
 What do you mean by fast?
-        \item  the literature on EIV estimation problems.
-As shortly mentioned in Chapter 2, there are solutions in the literature (by the way, you never mention the recent book written by Torsten Soderstrom in your document which is strange for me).
-But you do not describe them (even shortly) and, more importantly, you do not criticize them.
-Not in general but, at least, explain why you do not use them, what are their advantages and drawbacks and, finally, w.r.t.
-this list of positive and negative points, why you choose a (R)LS solution.
-E.g., why not using a recursive instrumental variable solution instead?
-
-        {\bfseries A description of IV methods has been added in Section 4.1.4}
-
-        \color{blue}
-    \subsection*{4.1.4. Instrumental variables formulation of the data-driven step input estimation method}
-
-    The errors-in-variables problem formulation (3.18) of the data-driven step input estimation method can be converted into an instrumental-variables problem.
-    A guide to do this conversion is found in \cite{Soderstrom18} and in \cite{Pintelon12Book}. 
-    The instrumental variable $\widetilde{\mathbf{Z}}$ has to be selected so that the normal equations (3.19) are expresed instead as
-    \begin{equation} \tag{4.22} \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{y}} . \label{eqn:neq_siv} \end{equation}
-    The selection of the instrumental variable has to be done in such a way that is correlated the regression matrix $\widetilde{\mathbf{K}}$, and uncorrelated with the perturbations $\bm{\epsilon}$.
-    Assuming that the block-Hankel matrix $\widetilde{\mathbf{K}}$ is given by
-    \begin{equation} \tag{4.23} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(n+1) & \Delta \widetilde{y}(n+2) & \cdots & \Delta \widetilde{y}(2n) \end{bmatrix} , \label{eqn:matrixK_r} \end{equation}
-    then, the instrumental variable $\widetilde{\mathbf{Z}}$ should not contain the elements that construct $\widetilde{\mathbf{K}}$, i.e., the elements of the vector $\left( \widetilde{y}(1), \ldots, \widetilde{y}(2n+1) \right)^\top$.
-    A possibility to generalize this idea is expressing the instrumental variable as
-    \begin{equation} \tag{4.24} \widetilde{\mathbf{Z}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-4n) & \Delta \widetilde{y}(t-4n+1) & \cdots & \Delta \widetilde{y}(t-3n-1) \\ G & \Delta \widetilde{y}(t-4n+1) & \Delta \widetilde{y}(t-4n+2) & \cdots & \Delta \widetilde{y}(t-3n) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-3n) & \Delta \widetilde{y}(t-3n+1) & \cdots & \Delta \widetilde{y}(t-2n-1) \end{bmatrix} , \label{eqn:matrixZ_t} \end{equation}
-    for a regression matrix
-    \begin{equation} \tag{4.25} \widetilde{\mathbf{K}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-2n+1) & \Delta \widetilde{y}(t-2n+2) & \cdots & \Delta \widetilde{y}(t-n) \\ G & \Delta \widetilde{y}(t-2n+2) & \Delta \widetilde{y}(t-2n+3) & \cdots & \Delta \widetilde{y}(t-n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-n+1) & \Delta \widetilde{y}(t-n+2) & \cdots & \Delta \widetilde{y}(t) \end{bmatrix} , \label{eqn:matrixK_t} \end{equation}
-    for $t = 4n+1, \ldots, N$.
-
-    The difference with the EIV formulation is that this implementation of the IV method requires to wait more, until $4n+1$ samples of the sensor response are read, instead of $2n+1$ samples. 
-    For sensors of small order this twice-the-time waiting does not represent a big change, but as the sensor order increases, the need for waiting can turn the method into impractical since a lot of time should pass before starting to estimate the input.
-
-    There exists a generalized instrumental variables method (GIVE) that was proposed to identify dynamic models from input/output data. 
-    GIVE is a bias-eliminating method that uses a generalized instrumental vector constructed from delayed values of the input and output signals \cite{Soderstrom18}.
-    Since the aim of this thesis is to analize a method that directly estimates the input from the observed sensor output, without identifying explicitly a dynamic model of the sensor, the GIVE method is out of the scope of this work.
-    A future research can include studying the formulation of a data-driven input estimation method in the spirit of the GIVE method. 
-    Nevertheless, a brief description of the GIVE method and its recursive version is given in the following paragraphs.
-
-    \subsubsection*{4.1.4.1 Generalized instrumental variables method}
-    For an errors-in-variables problem expressed as
-    \begin{equation} \tag{4.26} \widetilde{y}(t) = \widetilde{\bm{\phi}}(t) \bm{\theta}   \end{equation}
-    where $\widetilde{y}(t)$ is the output perturbed by measurement noise, $\bm{\theta}$ is the vector of to-be-estimated parameters, and $\widetilde{\bm{\phi}}(t)$ is the regressor vector constructed from delayed samples of the measured input $\widetilde{u}(t)$ and output $\widetilde{y}(t)$.
-    The least-squares estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is the solution of the normal equation
-    \begin{equation} \tag{4.27} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} \widehat{\bm{\theta}}_{\mathrm{LS}} = \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} , \label{eqn:norm} \end{equation}
-    where $\widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}}$ is the covariance matrix of the regressor vector, and $\widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}}$ is the cross-covariance matrix of the regressor vector and the output.
-    When the number of samples becomes infinitely large, these covariances matrices can be expressed as a sum of the contributions from exact data and measurement noise contributions
-    \begin{equation} \tag{4.28} \begin{aligned}  \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} &= \mathbb{E} \{ \widetilde{\bm{\phi}} \widetilde{\bm{\phi}}^\top \} = \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \\ \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} &= \mathbb{E}\{ \widetilde{\bm{\phi}} \widetilde{y}^\top \} = {\mathbf{r}}_{\bm{\phi} y} + {\mathbf{r}}_{\bm{\eta} \bm{\epsilon}} , \end{aligned} \end{equation}
-    where we are assuming that $\widetilde{y} =  y +\bm{\epsilon}$, $\widetilde{\bm{\phi}} = \bm{\phi} + \bm{\eta}$, and $\mathbf{r}_{\bm{\phi} y} = \mathbf{R}_{{\bm{\phi}}} \bm{\theta}$.
-    The bias of the least-square solution can be expressed as 
-    \begin{equation} \tag{4.29} \mathbf{R}_{\widetilde{\bm{\phi}}} \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) = \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{R}_{\widetilde{\bm{\phi}}} \bm{\theta} = \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} - \mathbf{R}_{\bm{\eta}} \bm{\theta} , \end{equation}
-
-    If we assume that the measurement noise is white for both the input and output, 
-    \begin{equation} \tag{4.30} \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} = \mathbb{E}\{ \bm{\eta} \bm{\epsilon}^\top \} = \mathbf{0} \quad \mathrm{and} \quad \mathbf{R}_{\bm{\eta}} = \begin{bmatrix} \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{n_y} & \mathbf{0} \\ \mathbf{0} & \sigma_u^2 \mathbf{I}_{n_u} \end{bmatrix}, \end{equation}
-    where $\sigma_{\bm{\epsilon}}^2$ and $\sigma_u^2$ are the variances of the output and the input, respectively. The normal equation (\ref{eqn:norm}) implies that 
-    \begin{equation} \tag{4.31} \left( \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \right)  \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) \neq \mathbf{0}, \end{equation}
-    and the estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is not consistent.
-    Instead, a bias-compensating least squares estimation can be obtained with  
-    \begin{equation} \tag{4.32} \widehat{\bm{\theta}}_{\mathrm{BCLS}} = \left( \mathbf{R}_{\widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\eta}} \right)^{-1}  \left(  \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{r}_{\bm{\eta} \bm{\epsilon}} \right) .
-\end{equation}
-
-    To formulate the generalized instrumental variable estimator, a generalized instrumental vector $\mathbf{z}(t)$ is built from delayed samples of the input and the output. 
-    The dimension of the generalized instrumental vector should be $n_{\mathbf{z}} \geq \mathrm{dim} ( \vartheta ) = n_u + n_y + 2$, since the total parameter vector $\vartheta = \left( \theta^\top, \rho^\top \right)^\top$ considers also the variances of the measurement noise $\rho = \left( \sigma_{\bm{\epsilon}}^2, \sigma_u^2 \right)^\top$.
-    Since the generalized instrumental vector $\mathbf{z}(t)$ and the equation error $\widetilde{y}(t) - \widetilde{\bm{\phi}}(t) \bm{\theta}$ must be correlated, we can express the sistem of equation as
-    \begin{equation} \tag{4.33}\left( \mathbf{R}_{\widetilde{\mathbf{z}} \widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \right) \bm{\theta} = \left(  \mathbf{r}_{\widetilde{\mathbf{z}} \widetilde{y}} - \mathbf{r}_{\bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) \right) , \label{eqn:GIVE} \end{equation}
-    where $\widetilde{\mathbf{z}} = \mathbf{z} + \bm{\zeta}$.
-
-    \subsubsection*{4.1.4.2 Recursive generalized instrumental variables method}
-    If we define
-    \begin{equation} \tag{4.34} \begin{aligned} \mathbf{h} \left( \bm{\rho}, \bm{\theta} \right)  &=  \mathbf{r}_{ \bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \bm{\theta}  =  \mathbf{J} \left( \bm{\theta} \right) \bm{\rho}, \\
-    \mathbf{R} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{\bm{\phi}}^\top \left( \tau \right) , \\     
-    \mathbf{q} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{y} \left( \tau \right) \\     
-    \mathbf{b} \left( t \right)  &=  \mathbf{q} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right)
-    , \end{aligned} \end{equation}
-    then the equation (\ref{eqn:GIVE}) is equivalent to
-    \begin{equation} \tag{4.35} \mathbf{R} \left( t \right) \bm{\theta} = \mathbf{b} \left( t \right) . \end{equation}
-    This representation allows a recursive form
-    \begin{equation} \tag{4.36} \begin{aligned} 
-    \mathbf{R} \left( t+1 \right)  &= \mathbf{R} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{\bm{\phi}}^\top \left( t+1 \right) - \mathbf{R} \left( t \right) \right), \\
-    \mathbf{b} \left( t+1 \right)  &= \mathbf{b} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{b} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right) \right), \\
-    \widehat{\bm{\theta}} \left( t+1 \right)  &=  \left( \mathbf{R}^\top \left( t+1 \right) \mathbf{R} \left( t+1 \right) \right)^{-1} \mathbf{R}^\top \left( t+1 \right) \mathbf{b} \left( t+1 \right) \\
-    \mathbf{q} \left( t+1 \right)  &= \mathbf{q} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{q} \left( t \right) \right) \\
-    \widehat{\bm{\rho}} \left( t+1 \right)  &=  \mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right) \left( \mathbf{q} \left( t \right) - \mathbf{R} \left( t \right) \widehat{\bm{\theta}} \left( t \right) \right) , \end{aligned} \end{equation}
-    where $\mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right)$ is the pseudo-inverse of $\mathbf{J} \left( \bm{\theta} \right)$.
-
-    The complexity of this recursive generalized instrumental variables estimator is ${O}(n^3)$, that is larger than the ${O}(n^2)$ of the recursive least squares.
-    Another drawback of the recursive generalized IV estimator is that it computes a pseudo-inverse matrix in each iteration step.
-    If we would need to formulate the data-driven step input estimation problem in the generalized IV framework, it would require to study the formulation in detail because currently the IV framework is based on the assumption that both measurements of the input and output are available, but for the step input estimation problem only the output is available.
-    Therefore, it is not straightforward to implement the recursive generalized IV estimator to estimate the step input, and this can lead to future research.
-    \color{black}
 
     \end{itemize}
-    \item In the document, you always talk about recursive implementations of least squares solution but, in the end, you never give access to any recursive algorithm description (as you do for what you call the ML method).
-Why?
-Do it as well for the recursive weighted least squares solution.
-
-    {\bfseries The following recursive least-squares algorithm description has been added to Section 3.2:}
-
-    \color{blue} 
-    The problem (3.17) admits a least-squares (LS) solution with normal equations given by
-    \begin{equation} \tag{3.19} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}}.   \end{equation} % \label{eqn:neq_seiv} 
-    The closed form of the LS solution can be expressed as
-    \begin{equation} \tag{3.20} \widehat{\bm{\theta}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
-    where $\widetilde{\mathbf{K}}^\dagger$ is the pseudo-inverse matrix of $\widetilde{\mathbf{K}}$.
-    As the number of samples $N$ increases, the number of rows of the matrix $\widetilde{\mathbf{K}}$ grows, making the pseudo-inverse computation inefficient since it requires larger flops, memory and time.
-
-    For metrology applications, it is desired to have a fast solution that can be obtained in real-time. 
-    The recursive algorithm (RLS) is a least-squares alternative to implement the step input estimation method in real-time.
-    The RLS recursively updates the solution of the system of equations, for each newly acquired sensor response sample, considering the previous value of the estimation, instead of performing a matrix inversion.
-    An expression of the RLS equations is given in \cite{Kailath00book} as
-    \begin{equation} \tag{3.21} \begin{aligned} \widehat{\bm{\theta}}(t) &= \widehat{\bm{\theta}}(t-1) + \bm{\kappa}_{t} \left( \widetilde{y}(t) - \widetilde{k}_{t} \widehat{\bm{\theta}}(t-1) \right) , \\  \bm{\kappa}_{t} &= \bm{\Psi}(t-1) \widetilde{k}_{t}^\top / \left( 1 + \widetilde{k}_{t} \bm{\Psi}(t-1) \widetilde{k}_{t}^\top  \right) \\ \bm{\Psi}(t) &= \left( \mathbf{I} - \bm{\kappa}_{t} \widetilde{k}_{t} \right) \bm{\Psi}(t-1), \label{eqn:RLS} \end{aligned} \end{equation}
-    where $\widetilde{k}_{t}$ represents the row of $\widetilde{K}$ that corresponds to the $t\mathrm{-th}$ sample, $\bm{\kappa}_{t}$ is a gain scalar, $\Psi(t)$ is a covariance matrix, and the estimates and the covariance matrix are initialized respectively with $\widehat{\bm{\theta}}(1)= \widehat{\bm{\theta}}(0)$ and $\Psi(-1) = \Psi(0)$.
-    Since the RLS is a recursive implementation of the least-squares computation, the RLS solution is equivalent to the LS solution.
-    Moreover, the statistical analysis of the LS solution is valid also for the RLS solution, and is preferred since it is simpler to express the statistics using the closed form of the LS solution.
-    \color{black}
-
-    {\bfseries The following description for the exponentially recursive least-squares algorithm has been added to Section 6.2.1:}
-
-    \color{blue}
-    The exponentially weighted recursive least-squares (RLS) can solve recursively the estimation problem (3.20).
-    In \cite{Kailath00book} is described the exponentially weighted RLS algorithm, which is as follows
-
-    \begin{equation} \tag{6.4} \begin{aligned}  \widehat{\bm{\theta}}(k) &= \widehat{\bm{\theta}}(k-1) + \bm{\kappa}_{k} \left( \widetilde{y}(k) - \widetilde{\mathbf{k}}_{k} \widehat{\bm{\theta}}(k-1) \right) , \\  \bm{\kappa}_{k} &= \omega^{-1} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top / \left( 1 + \omega^{-1} \widetilde{\mathbf{k}}_{k} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top  \right) \\ \bm{\Psi}(k) &= \omega^{-1} \left( \mathbf{I} - \bm{\kappa}_{k} \widetilde{\mathbf{k}}_{k} \right) \bm{\Psi}(k-1), \label{eqn:RLS} \end{aligned} \end{equation}
-    for $k = 2n+1, 2n+2, \ldots$, where $\widetilde{\mathbf{k}}_{k}$ represents the row of $\widetilde{\mathbf{K}}$ that corresponds to the $k\mathrm{-th}$ sample, $\bm{\kappa}_{k}$ is a gain scalar, and $\Psi(k)$ is a covariance matrix.
-    The estimates and the covariance matrix are initialized using the first $n+1$ samples, i.e.,  
-    \begin{equation} \tag{6.5} \begin{aligned} \widehat{\bm{\theta}} (2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top  \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega}  \widetilde{\mathbf{y}}_{n+1}, \quad \mathrm{and} \\ \Psi(2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} , \end{aligned} \end{equation}
-    where $\widetilde{\mathbf{K}}_{n+1}$ is the matrix $\widetilde{\mathbf{K}}$ with the first $n+1$ rows, and $\widetilde{\mathbf{y}}_{n+1}$ is the vector $\widetilde{\mathbf{y}}$ with the first $n+1$ elements.
-    When $\omega=1$, the exponentially weighted RLS solution is identical to that of the RLS.
-    When $\omega<1$, the older residuals are weighted with lower values than the residuals of recent observations.
-    In this way, the solution of the minimization problem depends more on newer data, and less on older data.
-
-    \color{black}
-
-    \item Double check your equations page 12 and 13. By the way, can you explain where Eq. (3.6) comes from?
-
-    {\bfseries The equations in page 12 and 13 have been reviewed and typos corrected. The derivation of Eq. (3.6) has been improved with more explaination: }
-
-    \color{blue}
-    A third method can directly estimate the input step level from the step response, without identifying a sensor model.
-    To derive this method, we use the first difference operator, $\Delta = \sigma - 1$, where $\sigma$ is the shift operator, defined as $(\sigma^\tau y) (k) = y(k + \tau)$.
-    Applying the first difference operator to the system state-space representation (3.1) results in the autonomous system
-    \begin{equation} \tag{3.10} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta {y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
-    where $\Delta {u}(k) = \mathbf{0}$, for $k \geq 0$, and
-    $\Delta \mathbf{x}(0) = (\mathbf{A} - \mathbf{I}) \mathbf{x}(0) + \mathbf{B} {{u}}$.
-
-    If the response $\Delta {y}$  of this autonomous system is persistently exciting of order $L$, then the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ of $L+1$ block rows constructed from $\Delta {y}$ satisfies
-    \begin{equation} \tag{3.11} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) = \mathrm{rank} \left( \begin{bmatrix} \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \iddots & \iddots & \iddots \\ \Delta y(L+1) & \Delta y(L+2) & \cdots & \Delta y(L+n) \end{bmatrix} \right) \leq L,  \end{equation}
-    and the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ is a linear map to the natural responses of the augmented autonomous system (3.3) 
-    \begin{equation} \tag{3.12} \mathbf{y}_{\mathrm{natural}} = \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell}, \end{equation}
-    from any vector $\bm{\ell} \in \mathbb{R}^{n}$.
-    The total response of the system is the sum of the forced response, due to the applied input ${u}$ and the natural response
-    \begin{equation} \tag{3.13} \mathbf{y} = \mathbf{y}_{\mathrm{forced}}+ \mathbf{y}_{\mathrm{natural}} =G {u} + \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell} ,  \end{equation} 
-    that is equivalent to
-    \begin{equation} \tag{3.14} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(N) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ G & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ G & \Delta y(N-n) & \Delta y(N-n+1) & \cdots & \Delta y(N-1) \end{bmatrix} }_{\mathbf{K}} \underbrace{ \begin{bmatrix} {{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
-    where $N$ is the number of samples.
-    From this point of view, the vector $\bm{\ell}$ is a linear transformation of the system's initial conditions.  
-    \color{black}
-
-    \item  In Section 4.1.1 and 4.1.2, discuss the normality assumption. Please restructure Section 4.1.3 which is not clear for me.
-
-    {\bfseries The discussion of the normality assumption has been added to Section 4.1, since it is relevant for the two cases considered.
-Section 4.1.3 has been reestructured to improve its clarity.}
-
-    \item  Why do you always study the statistical properties of your least squares solution with high SNR only? What do you get for SNR = 10 dB for instance? 
-    \item Give the values of the matrices you use for simulating data sets so that the reader can test what you have carried out.
-
-    {\bfseries The first paragraphs of Section 4.2, and Fig. 4.1, were added to describe the conditions in which the simulations were conducted from the structured and unstructured EIV problems.}
-
-    \color{blue}
-    Two Monte Carlo (MC) simulation studies were conducted to test the obtained bias and variance formulas.
-    There are two sets of formulas.
-    One set was obtained for the unstructured and uncorrelated EIV problem, and, the other set, for the structured and correlated EIV problem. 
-    The problem solved by the data-driven step input estimation method (3.20) belongs to the second set.
-    The objective of the MC simulations is to compare the results of the bias and variance formulas in these two cases. 
-    The comparison is an evaluation of the effect that the structure and correlation have in the problem solution.
-    To make a fair comparison, the same exact data and the perturbation variance is used in both types of EIV problems.
-
-    The exact data is the system of linear equations \ref{eqn:ddsiemexd} built by the data-driven step input estimation method.
-    To build the system of equations, the processed signal ${y}$ is the exact response of a sensor to the input $0.1s$, where $s$ is the unit step response.  
-    The sensor is modeled as a linear time-invariant mass-spring-damper system of order $n = 2$, that is given by 
-    \begin{equation} \tag{4.37}
-    \mathbf{x}(k+1) = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{m+M} & \dfrac{-k_{\mathrm{d}}}{m+M} \end{bmatrix} \mathbf{x}(k) + \begin{bmatrix} 0\\ -g\end{bmatrix} u(k),  
-    \quad y(k) = \begin{bmatrix} -1 & 0 \end{bmatrix} \mathbf{x}(k), \label{eqn:msdst}
-    \end{equation}
-    where $m=0.015$ kg, $k_{\mathrm{s}}=500$ kg/m, $k_{\mathrm{d}}=0.5$ kg s/m, $M=0.1$ kg, and $g=9.81 \ \mathrm{m/s^2}$.   
-    The exact sensor response is shown in Figure 4.1.
-
-    In the simulations, the sample size considered is $N=200$, since there the transient regime of the sensor model is still evident. 
-    In the case of the structured EIV problem, this sample size satisfies the requirements of the data-driven step input estimation method. 
-    \color{black}
-
-    \item Please remove Fig.
-4.1 b from Section 4.2.1 because the top and bottom curves are obtained with different systems and simulation condition.
-Or explain why you can compare both.
-It could be interesting to test each time (i.e., in Section 4.2.1 and 4.2.2) the bias and covariance formulas obtained by considering the structured and unstructured EIV problems.
-
-    {\bfseries Fig. 4.1 b has been relocated in its corresponding section. Now it is Fig. 4.7.}
 
@@
 This is the case for instance when I compare Section 4.2.2 and Section 5.1.
-    \item Except if you can convince me, do not use the name "maximum likelihood" when you do not introduce a likelihood you want to maximize.
-In Chapter 6, you introduce a nonlinear least squares algorithm, nothing else (see the attached document, Numerical Optimization, Chapter 10).
 
-    {\bfseries The method description in Section 6.2.2 has been modified to explain why the method is a maximum likelihood method. The modified paragraphs are as follows: }
-
-    \color{blue} 
-    Using a model of the sensor, the maximum-likelihood ML method simultaneously estimates the parameters of the applied affine input $u(t) = at + b$, and some of the sensor model parameters, using the observed sensor response $\widetilde{\mathbf{y}}$.
-    This method fits iteratively the response of the model $\widehat{\mathbf{y}}(\bm{\theta})$ to the sensor response by searching for the parameters $\bm{\theta}$ that reduce the error difference between both responses.
-    Algorithm 6.1 describes the formulation of an objective function $f = \mathbf{r}^\top \mathbf{r}$, where $\mathbf{r}$ is the residual $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$, that is minimized through the iterations.
-    The objective function is the sum of the squares of the differences between the samples of both responses.
-    Since the measurement noise is assumed Gaussian distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$, where the samples of this perturbation are independent and identically distributed, the minimization of the objective function $f$ maximizes the likelihood of fitting the actual sensor response with the model response.
-
-    The Jacobian matrix of the residual $\mathbf{J}\bm{\theta} = \partial \mathbf{r}(\bm{\theta}) / \partial \bm{\theta}$ can be analytically derived using the sensor model.
-    With the Jacobian matrix, the ML method searches the direction in which the residual decreases towards a local minimum.
-    Unfortunately, this ML method cannot guarantee the global minimum since it depends strongly on the initialization of the optimization parameters, as it is discussed in \cite{Nocedal06}.
-    Therefore, it is proposed to initialize the to-be-optimized parameters $a$ and $b$ using the results of the subspace method obtained after a convenient number of samples so that the affine input parameters are set close to the optimal values.
-    The initialization of the remaining model parameters under study can be done using the sensor model representation.
-    An illustrative example of the ML will be provided in a posterior subsection.  
-    \color{black}
 \end{itemize}
@@
 You can also mention your contributions there but in a less technical way since the problem statement only comes in the Chapter preliminaries.
-	\item In each main chapter (Chapter 4, Chapter 5, Chapter 6) you can add an introduction subsection, a state-of-the-art subsection and then elaborate your contributions and methods.
-
-    {\bfseries The introduction of the chapters now includes the state of the art to position the PhD work with respect to the literature.}
-
-	\item I believe that this structure will improve the understanding of the thesis and light up your contributions.
-
-	{\bfseries Indeed, this structure allows to visualize the thesis contributions.}
 
@@
 	\begin{itemize}
-	\item Please discuss the complexity and scalability issues related to your approach in relation to the state of the art.
-
-    {\bfseries The following paragraph has been added in Section 3.2 to discuss the scalability of the data-driven step input estimation method.}
-
-	\color{blue}
-    The computational complexity of the RLS algorithm solution to the system of equations (3.14) is $O \left( \left( n+1 \right)^2 \right)$.
-    The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
-    From there on, the RLS algorithm updates the solution with linear complexity.
-    Therefore, even though a large number of samples $N$ makes the matrix in the system of equations (3.14) increase in the number of rows, the estimation of the solution does not increase in complexity.
-    The data-driven step input estimation method is then scalable for any sensor of order $n$, and can be executed in devices with limited computational resources, provided that the computation of the  $n+1 \times n+1$ inverse matrix is feasible.
-    \color{black}
 
 	\item Position your work with regard to the state of the art as well as regarding solutions based on low-rank matrix completion and compressed sensing.
-	\item I recommend to use the formulation “we propose” than “I propose”
-
-	{\bfseries The formulations similar to “I propose” were replaced by equivalents such as “we propose”. }
-
-	\item You might wish to add a list of notation in the beginning of the thesis.
-
-	{\bfseries A nomenclature list is provided before the summary of the document. }
 
@@
 
-\section*{Philippe Dreesen}
-
-\begin{itemize}
-    \item General remarks:
-    \begin{itemize}
-    \item  It would be good to include some layman’s examples that motivate the work of the thesis already in the introduction.
-In the current text, there is a nice example on page 52, almost at the end of the thesis.
-These kinds of examples should appear in the introduction to motivate and position the work.
-
-    {\bfseries The following paragraphs have been added to the Introduction: }
-
-    \color{blue}
-
-    Take for example the case of a thermometer.
-    To measure temperature, the thermometer is put in contact with the substance or object of interest.
-    Accordingly to the Newton's law of cooling, the temperature $y$ of the thermometer changes proportionally to the its difference with respect to the temperature $u$ of the matter under study:
-    \begin{equation} \tag{2.1}  \dfrac{dy}{dt} = k_T \left( u - y \right) \label{eqn:NewtonCooling} \end{equation}
-    where the proportionality constant $k_T$ is a parameter of the thermometer first order model.
-    The input $u$ is unknown and its estimation is the objective of the measurement.
-    It is customary to take the reading from the thermometer once the temperature stops changing.
-
-    Another popular example of a measurement is found in weighing. 
-    In most of modern scales, the object of interest if placed on top of a weighing sensor.
-    A simple model of the weighing sensor is the second order mass-spring-damper model shown in Figure \ref{fig:simple_msd_system}.
-    In this model, the unknown mass $u$ causes a displacement $y$ of the scale and the dynamics of the sensor are described by the differential equation:
-    \begin{equation} \tag{2.2} \dfrac{d}{dt} \left( u(t) \ \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = u(t) \ g \end{equation}
-    where $m$ is the own mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.
-
-    For illustration purposes, Figures \ref{fig:thermometer} and \ref{fig:scale} show the responses of the thermometer and the weighing sensors, respectively, after the application of a constant input $u$, which can be represented by a step input.
-
-    \renewcommand{\thefigure}{2.1}
-    \begin{figure}[htb!]
-    \centering
-
-    \begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
-    \tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
-    \tikzstyle{damper}=[thick,decoration={markings,  
-    mark connection node=dmp,
-    mark=at position 0.5 with 
-    {
-    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
-    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
-    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
-    }
-    }, decorate]
-    \tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]
-
-    \node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
-    \node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u(t)$};
-
-    \node (ground1) at (M.south) [ground,yshift=-1.5cm,xshift=-0.625cm,anchor=north] {};
-    \draw (ground1.north west) -- (ground1.north east);
-    \draw [spring] (ground1.north) -- ($(M.south east)!(ground1.north)!(M.south west)$);
-
-    \node (groundc) at (M.south) [ground,yshift=-1.5cm,anchor=north] {}; 
-    \draw (groundc.north west) -- (groundc.north east);
-
-    \node (ground2) at (M.south) [ground,yshift=-1.5cm,xshift=0.625cm,anchor=north] {};
-    \draw (ground2.north west) -- (ground2.north east);
-    \draw [damper] (ground2.north) -- ($(M.south east)!(ground2.north)!(M.south west)$);
-
-    \node[draw=none,fill=none] at (-0.9cm,-1cm) {$k_{\mathrm{s}}$};
-    \node[draw=none,fill=none] at (0.15cm,-1cm) {$k_{\mathrm{d}}$};
-    \node[draw=none,fill=none] at (2.0cm,1.0cm) {$y$};
-    \draw [-latex,thick]  ++(2.2cm,-1cm) -- +(0cm,2.25cm);
-
-    \draw [-latex,thick] (M.east) ++(0,0) -- +(1cm,0);
-    \draw [line width=0.25mm] (2.2cm,-1cm) -- (2.2cm,1cm);
-    \draw [line width=0.25mm] (2.1cm,-1cm) -- (2.3cm,-1cm);
-    \draw [line width=0.25mm] (2.1cm,1cm) -- (2.3cm,1cm);
-    \draw [line width=0.25mm] (2.1cm,-0.5cm) -- (2.3cm,-0.5cm);
-    \draw [line width=0.25mm] (2.1cm,0.5cm) -- (2.3cm,0.5cm);
-    \draw [line width=0.25mm] (2.15cm,-0.25cm) -- (2.25cm,-0.25cm);
-    \draw [line width=0.25mm] (2.15cm,0.25cm) -- (2.25cm,0.25cm);
-    \draw [line width=0.25mm] (2.15cm,-0.75cm) -- (2.25cm,-0.75cm);
-    \draw [line width=0.25mm] (2.15cm,0.75cm) -- (2.25cm,0.75cm);
-    \draw [line width=0.25mm] (2.1cm,0cm) -- (2.3cm,0cm);
-
-    \end{tikzpicture}
-
-    \caption{\label{fig:simple_msd_system}  \color{blue} A second order mass-spring-damper model represents the weighing sensor.
-The application of the mass $u$ causes a change in the position $y$ of the scale.
-\color{black} }
-    \end{figure}
-
-    \renewcommand{\thefigure}{2.2}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_2.pdf} 
-    \caption{ \label{fig:thermometer} 
-    \color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_T=0.5 s^{-1}$, caused by applying a step input of 67 $\circ \mathrm{C}$  when the thermometer was initially at 23 $\circ \mathrm{C}$.
-In this example, we should wait 10 s to have an accurate estimation of the input.
-\color{black} }
-    \end{figure}
-
-    \renewcommand{\thefigure}{2.3}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_3.pdf} 
-    \caption{ \label{fig:scale} 
-    \color{blue} This simulation illustrate the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium.
-The sensor parameters are $k_{\mathrm{d}} = 5 \mathrm{N s/m}$, and $k_{\mathrm{s}} = 10000 \mathrm{N/m}$.
-In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement.
-\color{black}  }
-    \end{figure}
-
-    \color{black}
-
-    \end{itemize}
-    \item Detailed remarks:
-
-    {\bfseries The detailed remarks in this list have been attended accordingly: }
-
-    \begin{itemize}
-    \item Citations that are used in-line should have brackets only around the year, not around the authors’ names.
-For instance on page 3, line -2 is written “reviewed in [Hack and ten Caten, 2012]”.
-This should be displayed as “reviewed in Hack and ten Caten [2012].
-In LaTeX this can typically be done by making the proper distinction between the commands \ citet and \ citep.
-The former is used for references in text, the latter for parenthetical references.
-    \item  On page 1 of the PDF (front of thesis), I believe also N. Deligiannis should be mentioned with the title “prof.”. 
-    \item Page 1 of thesis: typo on line 6: “and on the sensor initial conditions” => “sensor’s initial conditions”
-    \item Page 3, line -2: “uncertainty analysis are reviewed” => “is reviewed”
-    \item Page 3, line -1: remove comma before “that”
-    \item Page 3, line -2: fix the in-text reference (see above)
-    \item Page 4, lines 11-13: fix the in-text references (see above)
-    \item Page 5, line 5: Remove “The” at the beginning of the sentence (The Hankel…)
-    \item Page 6, line 10: fix the in-text references (see above)
-    \item Page 7, lines 10-11: the sentence on heart and breathing physiological monitoring sounds a bit awkward. Please reformulate.
-    \item Page 9, line -1: fix the in-line reference (see above)
-    \item Page 10, lines 1-3: in the enumeration of the types of sensors, remove the article ‘the’ before the sensor name.
-So: “like three-axis sensors [ref], radio-frequency intruder sensors [ref], and radar sensors [ref].
-    \item Page 10, equation 3.10: The symbol D was not bold in the equations above. Be consistent!
-    \item Page 10, caption of figure: replace “reverts” by “inverts”
-    \item Page 11, line 9, put “poles” between parentheses instead of between commas.
-    \item Page 12, line -6: “that in matrix form is” => “which in matrix form is”
-    \item Page 13, line -9: “in the practice” => “in practice”
-    \item Page 14, line -14: missing space between “problem” and “(3.11)”
-    \item Page 16, halfway page: the sentence “Applying a second-order…” is not a proper sentence (perhaps fix this by removing “that”).
-    \item Page 16 introduces seemingly some new notation, e.g., b(.), mu(.) and C(.).
-While their meaning is clear from the context, it is desirable to give proper definitions when introducing new mathematical tools.
-    \item Page 17, line -1: fix in-text references (see above)
-    \item Page 20 (and beyond): The abbreviation CRLB is introduced, but sometimes CRB is used instead.
-    \item Page 21, line -10: fix in-text reference (See above)
-    \item Page 35, second sentence “The sensor is a dynamic…” is not a proper sentence.
-    \item Page 50: Try to reorganize this section into two or more paragraphs.
-    \item Page 59, line 2: fix in-text reference (See above)
-    \item Page 74, line 11: “The learning obtained from…” this is an awkward sentence; please reformulate
-    \item Page 81, all three journal publication references do not have (or have wrong?) page numbers.
-    \end{itemize}
-\end{itemize}
-
-\bibliography{../Gus-thesis/Gus-thesis.bib}
-
 \end{document}

### 2020-08-21 internal defense feedback/literature-review.tex
Second draft before public defense

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage[a4paper, total={6.5in, 9in}]{geometry}

\usepackage{natbib}
\bibliographystyle{unsrtnat}

\title{Literature review for the PhD defense of \linebreak Gustavo Quintana-Carapia}

\begin{document}

\maketitle

There are two main parts in this document. 
First there is an alphabetic list of the contributions described ion the literature.
Later, there is a review of the literature based on the read papers.

\section{Contribution of the papers}

\citet{Alaziz17} improve the during-sleep body motion classification using a load cell based system. With a machine learning algorithm, the authors classify a movement into one of 9 classes by extracting 24 features. The authors found that the method it classified correctly 90 \% of movements, and it is convenient for long-term home monitoring.

\citet{Azam15} propose a  dual implementation of the Kalman filter for estimating the unknown input and states of a linear state-space model by using sparse noisy acceleration measurements. The authors explain that the successive structure of the suggested filter prevents numerical issues attributed to un-observability and rank deficiency of the augmented formulation of the problem. The authors argument that the proposed methodology avoids the drift in the estimated input and displacements that is commonly encountered by existing joint input and state estimation filters. The authors found that the proposed technique obtains, by fine-tuning its parameters, accurate estimates of displacements and velocities.

\citet{Ballo16}  present an innovative six axis load cell for frontal impact tests on a dummy.  The authors measure the acceleration in two different points of the dummy's head, and the impact force input. The authors explain the observations with a first order model of the head and neck, and the neck stiffness and damping have been identified from experimental data.

\citet{Beck10}

\citet{Boschetti13} propose a model-based scheme that compensates for low frequency environmental vibrations in automatic weighing machines made of load cells. The authors think that it is ineffective to use low-pass filtering only in load cell measurements to overcome this problem. Their argument is that low cut-off frequencies downgrade machine speed by both introducing delay and increasing filtered signal rise time. The authors instead use mechanical models of the weighing machine and the load cells to process supplementary accelerometer measurements for computing an effective compensation of the environmental vibrations on load cell response effects.

\citet{Burmen09} present a method for fast and accurate in-line weighing of hard gelatin capsules based on capacitance sensor. The authors studied the effect of the shape and size of the capacitive sensor on the sensitivity and stability of the measurements to optimize the performance of the system. The results show that the capacitance iss well correlated with the capsule weight.

\citet{Cox06} show that a Monte Carlo method is an effective and versatile tool for determining the PDF for the measurands, and it provides a consistent Bayesian approach to the evaluation of uncertainty. The authors provide a guidance on optimizing the approach, identifies some pitfalls and indicates means for validating the results.

\citet{daSilva12} reviews and analyzes studies concerning measurement uncertainty. This paper provides a short summary of the state of the art of measurement uncertainty, analyzes the research scenario, and, suggests future work based on the analysis.

\citet{DEmilia16} recognize the relevance of three-axis accelerometers and inclinometers in buildings diagnostics. The authors first focus on the sensors calibration, including the low frequency range, 0-10 Hz. Later, they minimize systematic errors and identify the factors that affect uncertainty: the motion positioning of the accelerometer with respect to the vertical axis, the imposed motion law, and the radial positioning of the sensor.

\citet{Dienstfrey14} rederives the heuristic estimation rules that correct the measured quantities for the finite bandwidth of the measurement system response.
The rules are necessary when the measurement device responds slowly compared to the modulation of the input signal. The authors investigate the rules quantitative performance, and as an alternative, demonstrate that regularized deconvolution analysis exhibits more general quantitative utility at the expense of increased measurement burden and analytical complexity.

\citet{Diniz17} present a methodology for estimating the uncertainty of measurements in the calibration of temperature sensors in dynamic regimes for industrial applications. The authors use a first-order bare-wire-beaded thermocouple sensor as a case study. The experimental setup is costructed to give the input as a temperature step. The paper describes a methodology to compute the measurement uncertainty considering the uncertainties of all components of the measurement chain.

\citet{Elster07} analyzes the dynamic measurements using a second-order accelerometer dynamic model, assuming that estimates of the model parameters are available after previous dynamic calibration measurements. The authors want to estimate the unknown input of the sensor given its discrete-time output signal, using an FIR filter constructed from the underlying dynamic model. The input  uncertainties are determined considering the uncertainty of the dynamic model parameters estimates, the uncertainty of the sensor's output signal, and the applied signal processing steps. The proposed method allows real-time estimations.

\citet{Elster08} consider the measurement uncertainty evaluation when the value of the measurand depends on the continuous variable time. The authors generalize the GUM approach to the uncertainty of linear and time-invariant models. The paper describes a digital FIR filter for estimating the time-dependent  measurand value and an example is given for a second-order model.

\citet{Eichstadt16b} explore the relationship between the covariance matrix produced by the Kalman filter, a GUM-compliant uncertainty analysis, and a Bayesian uncertainty analysis. The authors found that all three approaches are compatible for linear systems with known system matrices. However, different results are obtained when the system matrices are not precisely known, or when the system is nonlinear, which might use the extended Kalman filter with linearized uncertainty propagation. The paper shows that the extended Kalman filter linearization and convergence errors can be avoided by applying Monte Carlo.

\citet{Esward16} present four examples of work at the UK's National Physical Laboratory that demonstrate how modelling and simulation can improve the understanding of dynamic measurement tasks. The presented examples are (i) a software simulation of a lock-in amplifier, (ii) a simulation of a sensor network in which one of the sensors has insufficient bandwidth for the measurement task and Kalman filter based data fusion is used to aggregate the sensor outputs, (iii) a study of performance imperfections in a clock embedded in a wireless sensor node using a Monte Carlo based method for simulating counting errors, and (iv) a simulation of wave propagation in a shock tube in which the lattice Boltzmann method was used to study non-ideal behaviour of the shock tube. The authors found that simulations are useful for designing measuring systems and identifying and quantifying measurement uncertainties and that the development of simulation software requires the developer to have a clear understanding of the measuring system of interest.

\citet{Esward09} describes the role and relevance of the GUM in the light of the challenges of new dynamic measurements, where the measurand variation has a significant effect on the measurement result and the associated uncertainty. The authors think that the challenges may require development of new measurement and calibration methods, as well as new methods for correcting sensor outputs and for the evaluation of uncertainties.

\citet{Eichstadt10} describes a procedure for digital deconvolution filters that reconstruct a time-variable measurand from the measurement signal of a LTI measurement system, exploiting the potentialities of digital signal processing. The authors discuss methods of minimum-phase all pass decomposition, asynchronous time reversal using the exact inverse filter and the construction of stable infinite impulse response and finite impulse response approximate inverse filters by a least squares approach in the frequency domain. The paper concludes that when a continuous model of the LTI system is available, application of least squares in the frequency domain for the construction of an approximate inverse filter is preferred. On the other hand, asynchronous time reversal filtering using the exact inverse filter appears superior when a discrete model of the LTI system is available and when causality of the deconvolution filter is not an issue.

\citet{Feiz17} assume that structured total least squares (STLS) for the fidelity term in color image restoration process gives better results than least squares (LS) but are more complex. The authors transform the color image restoration in two smaller subproblems for the image smooth and oscillatory parts. The authots solve the problems with STLS and LS, respectively. The paper shows that the proposed method is faster than STLS and gives competitive solutions.

\citet{Ferrero06} review of the present practice for expressing and estimating uncertainty in measurement and reports the  fundamental concepts of measurement science: 1) The result of a measurement provides only incomplete knowledge of the measurand, whose true value remains unknown and unknowable; 2) The uncertainty concept plays a key role in quantifying the incompleteness of the knowledge provided by the measurement result; and 3) A measurement result can be usefully employed only if the associated uncertainty is estimated and if it can be traced back to the appertaining standard; otherwise, it is a meaningless value.

\citet{Guo18} A rule of the effect between the output of a single load cell and a combinatorial structure was proposed by changing the sensitivity of the single load cell. The relevant relationship between the sensitivity of single load cell and the variation of total output of the combinatorial structure was analyzed by the numerical derivation method and experimental validation. In the cases, the output of a single load cell is linear or nonlinear, when the same load was acted on the combinatorial structure. The results showed that the effect was limited and the influence quantity is due to the accuracy and number of load cell. However, it must be controlled in a certain range to avoid additional measurement errors to the combinatorial structure. Moreover, the rule provides an evaluation method with improving the measurement accuracy for the combinatorial structure. Consequently, the rule can be employed as a useful guidance for the product quality when this evaluation method is applied to the load cell selection for the electronic truck scales.

\citet{Hale09} describe a method for calibrating the voltage that a step-like pulse generator produces at a load at every time point in the measured waveform. The calibration includes an equivalent-circuit model of the generator that can be used to determine how the generator behaves when it is connected to arbitrary loads. The generator is calibrated with an equivalent-time sampling oscilloscope and is traceable to fundamental physics via the electro-optic sampling system at the National Institute of Standards and Technology. The calibration includes a covariance-based uncertainty analysis that provides the uncertainty at each time in the waveform vector and the correlations between the uncertainties at the different times. From the calibrated waveform vector and its covariance matrix, we calculate pulse parameters and their uncertainties. We compare our method with a more traditional parameter-based uncertainty analysis.

\citet{Hammersley75} describes the theory and applications of  Monte Carlo Methods

\citet{Hernandez06} Load cells are transducers used to measure force or weight. Despite the fact thatthere is a wide variety of load cells, most of these transducers that are used in the weighingindustry are based on strain gauges. In this paper, an s-beam load cell based on strain gaugeswas suitably assembled to the mechanical structure of several seats of a bus underperformance tests and used to measure the resistance of their mechanical structure to tensionforces applied horizontally to the seats being tested. The load cell was buried in a broad-band noise background where the unwanted information and the relevant signal sometimesshare a very similar frequency spectrum and its performance was improved by using arecursive least-squares (RLS) lattice algorithm. The experimental results are satisfactoryand a significant improvement in the signal-to-noise ratio at the system output of 27 dB wasachieved, which is a good performance factor for judging the quality of the system.

\citet{Hessling06} A general method for estimating an upper bound of the dynamic measurement error in the time domain is derived, assuming that the complex-valued frequency response function of the system is known. This dynamic error bound related to the dynamic response of the system can be included in the measurement uncertainty as an uncorrected systematic error. The only restriction in the analysis is on linearity of the system, which can include electrical filters, mechanical sensors, amplifiers and similar subsystems. Calculations are verified by simulations for two example systems. The results for one of the most commonly used models of force and pressure transducers and accelerometers are presented emphasizing the main differences to a common estimate. For signals with well-confined spectra the linear dynamic error arises for two reasons: varying amplification and delay with frequency. The latter is analogous to the well-known bandwidth limiting dispersion of signals. For extended spectrum signals/simple pulses the asymptotic tails of the spectra may generate the major part of the error. No widespread robust, general and systematic method of quantifying the measurement error caused by these effects exists, despite that they may generate signal distortion far beyond the common prediction

\citet{Hessling08b} presents an approach to dynamic evaluation of measurement systems. On the one hand, it separates physical experiments, analysis and signal processing methods into successive steps of evaluation. On the other hand, the structure allows for resolution of an entire measurement system into its components for dedicated analyses. There is no limitation to particular applications except that it should be possible to model the response of the measurement system with differential equations with constant coefficients. Proposed tools such as estimation of error and uncertainty and direct mapping methods for synthesis of signal restoration filters are new or recently published, while others like system identification are well known but not previously systematically used in the context of calibration.

\citet{Hessling10} Dynamic Metrology is devoted to the analysis of dynamic measurements.  As  an  extended  calibration  service,  it  contains  many  novel  ingredients currently not included in the standard palette of metrology. Rather, Dynamic Metrology encompasses many operations found in the fields of system identification, digital signal processing and control theory. The analyses are more complex and more ambiguous than conventional uncertainty budgets of today. The important interactions in non-stationary measurements  may  be  exceedingly  difficult  to  both  control  and  to  evaluate.  In  many situations, in situ calibrations are required to yield a relevant result. Providing metrological services in this context will be a true challenge

\citet{Hessling08a} The dynamic error of measured signals is sometimes unacceptably large. If the dynamic properties of the measurement system are known, the true physical signal may to some extent be re-constructed. With a parametrized characterization of the system and sampled signals, time-domain digital filters may be utilized for correction. In the present work a general method for synthesizing such correction filters is developed. It maps the dynamic parameters of the measurement system directly on to the filter coefficients and utilizes time reversed filtering. This avoids commonly used numerical optimization in the filter synthesis. The method of correction is simple with absolute repeatability and stability, and results in a low residual error. Explicit criteria to control both the horizontal (time) and vertical (amplitude) discretization errors are presented in terms of the utilization of bandwidth and noise gain, respectively. To evaluate how close to optimal the correction is, these errors are also formulated in relation to the signal-to-noise ratio of the original measurement system. For purposes of illustration, typical mechanical and piezo-electric transducer systems for measuring force, pressure or acceleration are simulated and dynamically corrected with such dedicated digital filters.

\citet{Hessling11} The time-dependent measurement uncertainty has been evaluated in a number of recent publications, starting from a known uncertain dynamic model. This could be defined as the 'downward' propagation of uncertainty from the model to the targeted measurement. The propagation of uncertainty 'upward' from the calibration experiment to a dynamic model traditionally belongs to system identification. The use of different representations (time, frequency, etc) is ubiquitous in dynamic measurement analyses. An expression of uncertainty in dynamic measurements is formulated for the first time in this paper independent of representation, joining upward as well as downward propagation. For applications in metrology, the high quality of the characterization may be prohibitive for any reasonably large and robust model to pass the whiteness test. This test is therefore relaxed by not directly requiring small systematic model errors in comparison to the randomness of the characterization. Instead, the systematic error of the dynamic model is propagated to the uncertainty of the measurand, analogously but differently to how stochastic contributions are propagated. The pass criterion of the model is thereby transferred from the identification to acceptance of the total accumulated uncertainty of the measurand. This increases the relevance of the test of the model as it relates to its final use rather than the quality of the calibration. The propagation of uncertainty hence includes the propagation of systematic model errors. For illustration, the 'upward' propagation of uncertainty is applied to determine if an appliance box is damaged in an earthquake experiment. In this case, relaxation of the whiteness test was required to reach a conclusive result.

\citet{Hessling13a} Deterministic sampling can be used for nonlinear propagation of the statistics of signal processingmodels. Unlike Monte Carlo methods, random generators are not utilized in any stage. The samplesare instead calculated deterministically. Our novel approach generalizes the deterministic samplingtechnique for propagating covariance in the unscented Kalman filter by introducing genericexci-tation matricesdescribing small discrete canonical ensembles. The approximation lies in how wellthe available statistical information is encoded in the discrete ensemble, not how each sample ispropagated. The application and performance of deterministic sampling are illustrated for a typicalstep response analysis of an electrical device modeled with an uncertain digital filter.

\citet{Hessling13b} Statistical signal processing [1] traditionally focuses on extraction of information from noisy measurements. Typically, parameters or states are estimated by various filtering operations. Here, the quality of signal processing operations will be assessed by evaluating the statistical uncertainty of the result [2]. The processing could for instance simulate, correct, modulate, evaluate, or control the response of a physical system. Depending on the addressed task and the system, this can often be formulated in terms of a differential or difference signal processing model equation in time, with uncertain parameters and driven by an exciting input signal corrupted by noise. The quantity of primary interest may not be the output signal but can be extracted from it. If this uncertain dynamic model is linear-in-response it can be translated into a linear digital filter for highly efficient and standardized evaluation [3]. A statistical model of the parameters describing to which degree the dynamic model is known and accurate will be assumed given, instead of being the target of investigation as in system identification [4]. Model uncertainty (of parameters) is then propagated to model-ing uncertainty (of the result). The two are to be clearly distinguished - the former relate to the input while the latter relate to the output of the model.

\citet{Huang16} An approach to mass measurement for the electronic balance based on continuous-time sigma-delta (CT ?-?) modulator is described in this paper. As an effect of multifarious interfering noises, the accuracy of the electronic balance is restricted. The general idea of this proposed approach is to apply an electromagnetic-force-compensated load cell (EMCC), a related signal processing circuit, and a composite filter. The circuit mainly consists of a proportional-integral-differential controller and a pulsewidth modulator, which combine with the EMCC to form a CT ?-? modulator. As the CT ?-? modulator possesses inherent antialiasing filtering, oversampling, and noise-shaping characteristics, the interfering noises can be effectively removed by the composite filter that is composed of two sinc N filters. The simulations for both of the electronic balance models adopting the proposed approach and working in conventional pulse current mode are analyzed by adding the white Gaussian noise in Simulink. The simulation results demonstrate the effectiveness of this proposed approach, the proposed electronic balance in the field can improve its signal-to-noise ratio, and the testing results meet the requirement for the weighing accuracy of the special-class scale defined by Organisation Internationale De M\'etrologie L\'egale R76 nonautomatic weighing instruments.

\citet{Jafaripanah05}  investigate the application of analog adaptive techniques to the area of dynamic sensor compensation, of which there is little reported work in the literature. The case is illustrated by showing how the response of a load cell can be improved to speed up the process of measurement. The load cell is a sensor with an oscillatory response in which the measurand contributes to the response parameters. Thus, a compensation filter needs to track variation in measurand, whereas a simple fixed filter is only valid at one specific load value. To facilitate this investigation, computer models for the load cell and the adaptive compensation filter have been developed. To allow a practical implementation of the adaptive techniques, a novel piecewise linearization technique is proposed in order to vary a floating voltage-controlled resistor in a linear manner over a wide range. Simulation and practical results are presented, thus, demonstrating the effectiveness of the proposed techniques.

\citet{Jia18} In this paper, we focus on the target localization problem which finds broad applications in radar, sonar and wireless sensor networks. A pseudolinear overdetermined system of equations is constructed from the nonlinear hybrid TDOA-AOA measurements about target location. Considering the matrix and vector in the constructed pseudolinear system are both contaminated by the measurement noise, a new weight least squares (WLS) method which is based on the first order Taylor expansions of the noise terms is developed in this paper and it can reduce the estimation bias that arise from the least squares (LS) method. In particular we focus on constructing a localization algorithm to reduce the bias that easily arise from the traditional methods. Thus in addition, a novel structured total least squares (STLS) method is also developed in this paper to further reduce the estimation bias specially when the target is outside the convex hull formed by sensors. Numerical examples show the superiority of the proposed STLS method in estimation accuracy compared with the LS method, total least squares (TLS) method and the proposed WLS method.

\citet{Kesilmis16} We propose a novel geometric approach to processing of load cell signals for inline weight measurement applications. These systems usually rely on oscillatory load cell signals which always need to time to settle. We describe the theoretical and experimental implementation of a newly proposed geometric approach for load cell signals, and verify the effectiveness of the method through experiments.

\citet{Kiviet12high} An approximation to order is obtained for the bias of the full vector of least-squares estimates obtained from a sample of size in general stable but not necessarily stationary ARX(1) models with normal disturbances. This yields generalizations, allowing for various forms of initial conditions, of Kendall's and White's classic results for stationary AR(1) models. The accuracy of various alternative approximations is examined and compared by simulation for particular parameterizations of AR(1) and ARX(1) models. The results show that often the second-order approximation is considerably better than its first-order counterpart and hence opens up perspectives for improved bias correction. However, order approximations are also found to be more vulnerable in the near unit root case than the much simpler order approximations.

\citet{Kiviet14improved} In dynamic regression models conditional maximum likelihood (least-squares) coefficient and variance estimators are biased. Using expansion techniques an approximation is obtained to the bias in variance estimation yielding a bias corrected variance estimator. This is achieved for both the standard and a bias corrected coefficient estimator enabling a comparison of their mean squared errors to second order. Sufficient conditions for admissibility of these approximations are formally derived. Illustrative numerical and simulation results are presented on bias reduction of coefficient and variance estimation for three relevant classes of first-order autoregressive models, supplemented by effects on mean squared errors, test size and size corrected power. These indicate that substantial biases do occur in moderately large samples, but these can be mitigated considerably and may also yield mean squared error reduction. Crude asymptotic tests are cursed by huge size distortions. However, operational bias corrections of both the estimates of coefficients and their estimated variance (for which software is provided) are shown to curb type I errors reasonably well.

\citet{Lee16} Ballistocardiographs (BCGs), which record the mechanical activity of the heart, have been a subject of interest for several years because of their advantages in providing unobtrusive physiological measurements. BCGs could also be useful for monitoring the biological signals of infants without the need for physical confinement. In this study, we describe a physiological signal monitoring bed based on load cells and assess an algorithm to extract the heart rate and breathing rate from the measured load-cell signals. Four infants participated in a total of 13 experiments. As a reference signal, electrocardiogram and respiration signals were simultaneously measured using a commercial device. The proposed automatic algorithm then selected the optimal sensor from which to estimate the heartbeat and respiration information. The results from the load-cell sensor signals were compared with those of the reference signals, and the heartbeat and respiration information were found to have average performance errors of 2.55% and 2.66%, respectively. The experimental results verify the positive feasibility of BCG-based measurements in infants.

\citet{Lemmerling02} In this paper an overview is given of the Structured Total Least Squares (STLS) approach and its recent extensions. The Structured Total Least Squares (STLS) problem is a natural extension of the Total Least Squares (TLS) problem when constraints on the matrix structure need to be imposed. Similar to the ordinary TLS approach, the STLS approach can be used to determine the parameter vector of a linear model, given some noisy measurements. In many signal processing applications, the imposition of this matrix structure constraint is necessary for obtaining Maximum Likelihood (ML) estimates of the parameter vector.

\citet{Link07} A recently proposed accelerometer model is applied for determining the accelerometer's output to transient accelerations. The model consists of a linear, second-order differential equation with unknown coefficients. It is proposed to estimate these model parameters from sinusoidal calibration measurements, and an estimation procedure based on linear least-squares is presented. In addition, the uncertainties associated with the estimated parameters are determined utilizing a Monte Carlo simulation technique.

The performance of the proposed modelling approach was tested by its application to calibration measurements of two back-to-back accelerometers (ENDEVCO type 2270 and Br{\"u}el \& Kj{\ae}r type 8305). For each of the two accelerometers, the model was first estimated from sinusoidal calibration measurements and then used to predict the accelerometer's behaviour for two shock calibration measurements. Measured and predicted shock sensitivities were found consistent with differences below 1% in most cases which confirms the benefit of the proposed modelling approach.

\citet{Link09} propose a method  for evaluating the uncertainty associated with the output of a discrete-time IIR filter when the input signal is corrupted by additive noise and the filter coefficients are uncertain. This task arises, for instance, when the noise-corrupted output of a measurement system is compensated by a digital filter which has been designed on the basis of the characteristics of the measurement system. We assume that the noise is either stationary or uncorrelated, and we presume knowledge about its autocovariance function or its time-dependent variances, respectively. Uncertainty evaluation is considered in line with the 'Guide to the Expression of Uncertainty in Measurement'. A state-space representation is used to derive a calculation scheme which allows the uncertainties to be evaluated in an easy way and also enables real-time applications. The proposed procedure is illustrated by an example.

\citet{Markovsky07overview} review the development and extensions of the classical total least-squares method and describe algorithms for its generalization to weighted and structured approximation problems. In the generic case, the classical total least-squares problem has a unique solution, which is given in analytic form in terms of the singular value decomposition of the data matrix. The weighted and structured total least-squares problems have no such analytic solution and are currently solved numerically by local optimization methods. We explain how special structure of the weight matrix and the data matrix can be exploited for efficient cost function and first derivative computation. This allows to obtain computationally efficient solution methods. The total least-squares family of methods has a wide range of applications in system theory, signal processing, and computer algebra. We describe the applications for deconvolution, linear prediction, and errors-in-variables system identification.

\citet{Markovsky15cep} Metrology is advancing by development of new measurement techniques and corresponding hardware. A given measurement technique, however, has fundamental speed and precision limitations. In order to overcome the hardware limitations, we develop signal processing methods based on the prior knowledge that the measurement process dynamics is linear time-invariant.
Our approach is to model the measurement process as a step response of a dynamical system, where the input step level is the quantity of interest. The solution proposed is an algorithm that does real-time processing of the sensor's measurements. It is shown that when the measurement process dynamics is known, the input estimation problem is equivalent to state estimation. Otherwise, the input estimation problem can be solved as a system identification problem. The main underlying assumption is that the measured quantity is constant and the measurement process is a low-order linear time-invariant system. The methods are validated and compared on applications of temperature and weight measurement.

\citet{Markovsky15ieee} Dynamic measurement aims to improve the speed and accuracy characteristics of measurement devices by signal processing. State-of-the-art dynamic measurement methods are model-based adaptive methods, i.e., 1) they estimate model parameters in real-time and 2) based on the identified model perform model-based signal processing. The proposed model-free method belongs to the class of the subspace identification methods. It computes directly the quantity of interest without an explicit parameter estimation. This allows efficient computation as well as applicability to general high order multivariable processes.

\citet{Matthews14}  focus on the mathematical modelling required to support the development of new primary standard systems for traceable calibration of dynamic pressure sensors. We address two fundamentally different approaches to realizing primary standards, specifically the shock tube method and the drop-weight method. Focusing on the shock tube method, the paper presents first results of system identification and discusses future experimental work that is required to improve the mathematical and statistical models.
The authors use simulations to identify differences between the shock tube and drop-weight methods, to investigate sources of uncertainty in the system identification process and to assist experimentalists in designing the required measuring systems. We demonstrate the identification method on experimental results and draw conclusions.

\citet{Mayne14} reviews Model Predictive Control, gives an overview of some current developments and suggests a few avenues for future research.

\citet{Neuberger07}

\citet{Niedzwiecki16a} Dynamic weighing, i.e., weighing of objects in motion, without stopping them on the weighing platform, allows one to increase the rate of operation of automatic weighing systems, used in industrial production processes, without compromising their accuracy. Since the classical identification-based approach to dynamic weighing, based on the second-order mass-spring-damper model of the weighing system, does not yield satisfactory results when applied to conveyor belt type checkweighers, several extensions of this technique are examined. Experiments confirm that when appropriately modified the identification-based approach becomes a reliable tool for dynamic mass measurement in checkweighers.

\citet{Niedzwiecki16b} propose to characterize as a filtering scheme based on the finite impulse response model of the weighing system response. It is shown that when such a model-based filtering is applied, the attained weighing accuracy is up to four times higher than that guaranteed by the currently available state-of-the-art solutions.
Conveyor belt-type checkweighers are increasingly popular components of modern production lines. They are used to assess the weight of the produced items in motion, i.e., without stopping them on the weighing platform. The main challenge one faces when designing a dynamic weighing system is providing high measurement accuracy, especially at high conveyor belt speeds.

\citet{Ogorevc16}  describe a procedure for dynamic measurement uncertainty evaluation in order to examine the requirements for clinical thermometer dynamic properties in standards and recommendations. In this study thermistors were used as temperature sensors, transient temperature measurements were performed in water and air and the measurement data were processed for the investigation of thermometer dynamic properties. The thermometers were mathematically modelled. A Monte Carlo method was implemented for dynamic measurement uncertainty evaluation. The measurement uncertainty was analysed for static and dynamic conditions. Results showed that dynamic uncertainty is much larger than steady-state uncertainty. The results of dynamic uncertainty analysis were applied on an example of clinical measurements and were compared to current requirements in ISO standard for clinical thermometers. It can be concluded that there was no need for dynamic evaluation of clinical thermometers for continuous measurement, while dynamic measurement uncertainty was within the demands of target uncertainty. Whereas in the case of intermittent predictive thermometers, the thermometer dynamic properties had a significant impact on the measurement result. Estimation of dynamic uncertainty is crucial for the assurance of traceable and comparable measurements.
Clinical thermometers in intensive care units are used for the continuous measurement of body temperature.

\citet{OIML_R51_1} specifies  the  metrological  and  technical  requirements  and  test  procedures for automatic catchweighing instruments (catchweighers), hereinafter called "instruments", that are subject to national metrological control.

\citet{Olmi16} present a didactic project at the University of Bologna, focused on the development of strain gage load cells, from design to calibration and use. Details are provided on the course units involved and on the levels of teaching, and three case studies are presented. Two case studies deal with the design of decoupled load cells, using simple beam geometry. Students had the opportunity to become confident with the principles ofmechanics and to tackle the not trivial activities like calibration by decoupled load application and determination of the related matrices. The third one deals with the development of a load cell for impact loads, where the students had the chance to observe a practical application of the studied Wheatstone bridge connection, to participate in on-field tests, and to better understand the close relationship between experimentation and design.
Many numerical models are nowadays available for the structural analysis of complexly shaped structures. However, a critical problem consists of the estimation of the actual loads that a structure withstands. The most proper way to determine them under service conditions consists of executing in-field tests, where load cells have an important role. This article deals

\citet{Palanthandalam10parameter}  investigate the consistency of parameter estimates obtained from least-squares identification with a quadratic parameter constraint. For generality, we consider infinite impulse-response systems with coloured input and output noise. In the case of finite data, we show that there always exists a possibly indefinite quadratic constraint depending on the noise realisation that results in a constrained optimisation problem that yields the true parameters of the system when a persistency condition is satisfied. When the noise covariance matrix is known to within a scalar multiple, we prove that solutions of the quadratically constrained least-squares (QCLs) estimator with a semidefinite constraint matrix are both unbiased and consistent in the sense that the averaged problem and limiting problem produce, respectively, unbiased and true (with probability 1) estimators. In addition, we provide numerical results that illustrate these properties of the QCLS estimator.

\citet{Pintelon90} present a method for compensating in real time for a nonideal transfer function of a data acquisition channel by means of a digital IIR (infinite impulse response) filter. Real-time compensation of the amplitude and phase characteristics of an acquisition channel so that the long-term error is less than +or-0.01 dB and +or-0.1 degrees , respectively, has been obtained experimentally. It is shown that stability and linearity are the main requirements for the components of the data acquisition channel. Careful selection of the components and outlining of the antialias filter and the programmable amplifier is no longer necessary.

\citet{Pietrzak14}  presents and verifies experimentally a new approach, based on time-variant low-pass filtering, to overcome classical linear time-invariant low-pass filtering results for high conveyor belt speeds. The authors show that, when properly tuned, the proposed time-variant filter fulfills the measurement accuracy requirements for a wide range of operating conditions.
Conveyor belt type checkweighers are complex mechanical systems consisting of a weighing sensor (strain gauge load cell, electrodynamically compensated load cell), packages (of different shapes, made of different materials) and a transport system (motors, gears, rollers). Disturbances generated by the vibrating parts of such a system are reflected in the signal power spectra in a form of strong spectral peaks, located usually in the lower frequency range. Such low frequency components overlap in the frequency domain with the useful signal and it is very difficult to eliminate them. The conventional way of suppressing disturbances is via low-pass filtering of the signal obtained from the load cell. However, if the speed of the conveyor belt is high, the response of the applied filter may not settle fast enough to enable accurate weighing of objects in motion, i.e., without stopping them on the weighing conveyor.

\citet{Piskorowski08} presents a new method for dynamic compensation of the load cell response using linear time-varying continuous-time filter. Load cells have an oscillatory response which always needs time to settle down. Therefore, it is justified to search a technique for an effective improvement of the sensor response. The paper describes a theoretical implementation of the proposed time-varying filter, and suggests the implementation technique with the aid of which this kind of filter can be implemented in practice. Simulation results verifying the effectiveness of the proposed filter are presented and compared to the traditional time-invariant configuration.

\citet{Rhode14recursive} show that the generalized total least squares (GTLS) problem with a singular noise covariance matrix is equivalent to the restricted total least squares (RTLS) problem and propose a recursive method for its numerical solution. The method is based on the generalized inverse iteration. The estimation error covariance matrix and the estimated augmented correction are also characterized and computed recursively. The algorithm is cheap to compute and is suitable for online implementation. Simulation results in least squares (LS), data least squares (DLS), total least squares (TLS), and restricted total least squares (RTLS) noise scenarios show fast convergence of the parameter estimates to their optimal values obtained by corresponding batch algorithms.

\citet{Rossander15} measures unique blade force on an open site straight-bladed vertical axis wind turbine. This paper presents a method for measuring the tangential and normal forces on a 12-kW vertical axis wind turbine prototype with a three-bladed H-rotor. Four single-axis load cells were installed in-between the hub and the support arms on one of the blades. The experimental setup, the measurement principle, together with the necessary control and measurement system are described. The maximum errors of the forces and accompanying weather data that can be obtained with the system are carefully estimated. Measured forces from the four load cells are presented, as well as the normal and tangential forces derived from them and a comparison with theoretical data. The measured torque and bending moment are also provided. The influence of the load cells on the turbine dynamics has also been evaluated. For the aerodynamic normal force, the system provides periodic data in agreement with simulations. Unexpected mechanical oscillations are present in the tangential force, introduced by the turbine dynamics. The measurement errors are of an acceptable size and often depend on the measured variable. Equations are presented for the calculation of measurement errors

\citet{Saggin01} studies the dynamic correction of a thermometer for atmospheric measurements. The sensor analysed is the one launched onboard the Huygens probe of the Cassini mission. The specific aspects of the measurement are related to the strong changes in the fluid-dynamic conditions during the operative phase. This peculiarity does not allow the use of one of the various correction techniques based on the knowledge of the dynamic characteristics of the sensor. The thermometer, however, has the peculiarity of being composed of two independent platinum sensors, having very different dynamic properties. The study, starting from a previously conceived dynamic scheme of the sensor, assesses the possibility of determining continuously, from measurement data, the current values of the four time constants required for dynamic characterisation. Lastly one procedure for the dynamic correction of the measurements is presented, along with some verification of its effectiveness and, of the impact on the overall measurement uncertainty.

\citet{Shu93} derive a discrete output error (OE) model of second order for the weighing system dynamics. Using the model and the recursive least squares (RLS) procedure, model parameters and then the mass being weighed can be estimated from a dynamic measurement signal of very short duration. The validity and the accuracy of this method are illustrated by digital simulation studies and real-life measurements

\citet{Soderstrom07} gives a survey of errors-in-variables methods in system identification. Background and motivation are given, and examples illustrate why the identification problem can be difficult. Under general weak assumptions, the systems are not identifiable, but can be parameterized using one degree-of-freedom. Examples where identifiability is achieved under additional assumptions are also provided. A number of approaches for parameter estimation of errors-in-variables models are presented. The underlying assumptions and principles for each approach are highlighted.

\citet{Soderstrom18} gives a comprehensive overview of errors-in-variables (EIV) problems in system identification. This problem is about modeling of dynamic systems when all measured variables and signals are noise-corrupted. A number of different approaches are described and analyzed. Representations where errors or measurement noises are present on both inputs and outputs are usually called 'errors-in-variables' (EIV) models. They play an important role when the purpose is the determination of the physical laws that describe the process, rather than the prediction of its future behavior.

\citet{Stewart90SPT} approaches  classical matrix perturbation theory from a probabilistic point of view. The perturbed quantity is approximated by a first-order perturbation expansion, in which the perturbation is assumed to be random. This permits the computation of statistics estimating the variation in the perturbed quantity. Up to the higher-order terms that are ignored in the expansion, these statistics tend to be more realistic than perturbation bounds obtained in terms of norms. The technique is applied to a number of problems in matrix perturbation theory, including least squares and the eigenvalue problem.

\citet{Tasaki07} uses a multi-stage conveyor belt scale to propose a simplified and effective mass estimation algorithm under practical vibration modes. The output signals from the conveyor belt scales are always contaminated with noises due to vibrations of the conveyor belt and the product in motion. In this paper digital filter of finite-duration impulse response (FEB) type is designed to provide adequate accuracy. The experimental results on conveyor belt scales suggest that the filtering algorithm proposed here is effective enough to practical applications. As long as spaces between successive products are set within a specified range, the products can be weighed correctly even if products having different lengths are transported in random manner.
Today higher speed of operation and highly accurate weighing of packages during crossing a conveyor belt has been getting more important in the food and distribution industries etc. Continuous weighing means that masses of discrete packages on a conveyor belt are automatically determined in sequence.

\citet{Vaccaro94} develops a perturbation expansion for the subspace spanned by a set of singular vectors. A first-order expansion of this type has recently been developed and used to analyze the performance of direction-finding algorithms in array signal processing. The author derives a new second-order expansion and the result is illustrated with two examples.

\citet{VanHuffel91Book}  give a unified presentation of the TLS problem. A description of its basic principles are given, the various algebraic, statistical and sensitivity properties of the problem are discussed, and generalizations are presented. The TLS applications are surveyed to facilitate uses in an even wider range of applications. The authors provide comparison  with the well-known least squares methods.

\citet{VanHuffel07TLSeditorial} arranged a special issue to present an overview of the progress of total least squares (TLS) in computational mathematics and engineering, and as errors-in-variables (EIV) modeling or orthogonal regression in the statistical community. The TLS method is one of several linear parameter estimation techniques that has been devised to compensate for data errors. For set of given multidimensional data points, the idea is to modify all data points in such a way that some norm of the modification is minimized subject to the constraint that the modified vectors satisfy a linear relation. One of the main reasons for TLS popularity is the availability of efficient and numerically robust algorithms in which the singular value decomposition (SVD) plays a prominent role, and that TLS is an application oriented procedure.

\citet{Vlajic16}  describes an apparatus for traceable, dynamic calibration of force transducers using harmonic excitation. In this system, the force applied to the transducer is produced by the acceleration of an attached mass. The capabilities of this system are demonstrated by performing dynamic calibrations of two shear-web-type force transducers up to a frequency of 2?kHz, with an expanded uncertainty below 1.2 \%. Tha authors give an account of all significant sources of uncertainty, including a detailed consideration of the effects of dynamic tilting (rocking), which is a leading source of uncertainty in such harmonic force calibration systems.

\citet{Yamani18} studies the check-weigher that uses a load cell conveying the mass of an object to be measured. In check-weigher it is necessary to estimate the mass value before the scale is balance still in order to high performance. The conveying motor affects the measured value of the scale and the measuring time. The paper describes a dummy conveyor giving the mass of the motor as unbalanced load. The authors found, after static and dynamic analysis, that the unbalanced load is major factor to deteriorate the measurement accuracy of mass.

\citet{Yeredor04homogeneous} shows that the use of some commonly applied constraints, such as a quadratic constraint, can lead to inconsistent estimates of a when solving  "homogeneous least-squares" models of the form Ya?0. These problems are encountered, e.g., when modeling auto-regressive (AR) processes. The authors explain this discrepancy, and give the remedy. The paper discussees numerically appealing solutions.

\citet{Zahradka18} demonstrates the feasibility of distinguishing certain positions and movements with load cell sensors. The paper shows that identification of positions and movements for individuals in bed can be used as a tool in a variety of clinical settings. Six positions were classified with an accuracy of 74.9 \% and six movements were classified with an accuracy of 79.7 \%. Stationary positions were characterized by the signals from the four load cells and the coordinates of the center of mass. Movements were characterized by the changes in load cell signals.

\section{Literature relevant for Statistical analysis state of the art}

\citet{Cox06} show that a Monte Carlo method is an effective and versatile tool for determining the PDF for the measurands, and it provides a consistent Bayesian approach to the evaluation of uncertainty. The authors provide a guidance on optimizing the approach, identifies some pitfalls and indicates means for validating the results.

\citet{daSilva12} reviews and analyzes studies concerning measurement uncertainty. This paper provides a short summary of the state of the art of measurement uncertainty, analyzes the research scenario, and, suggests future work based on the analysis.

\citet{Dienstfrey14} rederives the heuristic estimation rules that correct the measured quantities for the finite bandwidth of the measurement system response.
The rules are necessary when the measurement device responds slowly compared to the modulation of the input signal. The authors investigate the rules quantitative performance, and as an alternative, demonstrate that regularized deconvolution analysis exhibits more general quantitative utility at the expense of increased measurement burden and analytical complexity.

\citet{Elster07} analyzes the dynamic measurements using a second-order accelerometer dynamic model, assuming that estimates of the model parameters are available after previous dynamic calibration measurements. The authors want to estimate the unknown input of the sensor given its discrete-time output signal, using an FIR filter constructed from the underlying dynamic model. The input  uncertainties are determined considering the uncertainty of the dynamic model parameters estimates, the uncertainty of the sensor's output signal, and the applied signal processing steps. The proposed method allows real-time estimations.

\citet{Elster08} consider the measurement uncertainty evaluation when the value of the measurand depends on the continuous variable time. The authors generalize the GUM approach to the uncertainty of linear and time-invariant models. The paper describes a digital FIR filter for estimating the time-dependent  measurand value and an example is given for a second-order model.

\citet{Eichstadt16b} explore the relationship between the covariance matrix produced by the Kalman filter, a GUM-compliant uncertainty analysis, and a Bayesian uncertainty analysis. The authors found that all three approaches are compatible for linear systems with known system matrices. However, different results are obtained when the system matrices are not precisely known, or when the system is nonlinear, which might use the extended Kalman filter with linearized uncertainty propagation. The paper shows that the extended Kalman filter linearization and convergence errors can be avoided by applying Monte Carlo.

\citet{Esward16} present four examples of work at the UK's National Physical Laboratory that demonstrate how modelling and simulation can improve the understanding of dynamic measurement tasks. The presented examples are (i) a software simulation of a lock-in amplifier, (ii) a simulation of a sensor network in which one of the sensors has insufficient bandwidth for the measurement task and Kalman filter based data fusion is used to aggregate the sensor outputs, (iii) a study of performance imperfections in a clock embedded in a wireless sensor node using a Monte Carlo based method for simulating counting errors, and (iv) a simulation of wave propagation in a shock tube in which the lattice Boltzmann method was used to study non-ideal behaviour of the shock tube. The authors found that simulations are useful for designing measuring systems and identifying and quantifying measurement uncertainties and that the development of simulation software requires the developer to have a clear understanding of the measuring system of interest.

\citet{Eichstadt10} describes a procedure for digital deconvolution filters that reconstruct a time-variable measurand from the measurement signal of a LTI measurement system, exploiting the potentialities of digital signal processing. The authors discuss methods of minimum-phase all pass decomposition, asynchronous time reversal using the exact inverse filter and the construction of stable infinite impulse response and finite impulse response approximate inverse filters by a least squares approach in the frequency domain. The paper concludes that when a continuous model of the LTI system is available, application of least squares in the frequency domain for the construction of an approximate inverse filter is preferred. On the other hand, asynchronous time reversal filtering using the exact inverse filter appears superior when a discrete model of the LTI system is available and when causality of the deconvolution filter is not an issue.

\citet{Feiz17} assume that structured total least squares (STLS) for the fidelity term in color image restoration process gives better results than least squares (LS) but are more complex. The authors transform the color image restoration in two smaller subproblems for the image smooth and oscillatory parts. The authots solve the problems with STLS and LS, respectively. The paper shows that the proposed method is faster than STLS and gives competitive solutions.

\citet{Ferrero06} review of the present practice for expressing and estimating uncertainty in measurement and reports the  fundamental concepts of measurement science: 1) The result of a measurement provides only incomplete knowledge of the measurand, whose true value remains unknown and unknowable; 2) The uncertainty concept plays a key role in quantifying the incompleteness of the knowledge provided by the measurement result; and 3) A measurement result can be usefully employed only if the associated uncertainty is estimated and if it can be traced back to the appertaining standard; otherwise, it is a meaningless value.

\citet{Hammersley75} describes the theory and applications of  Monte Carlo Methods

\citet{Hessling08a} The dynamic error of measured signals is sometimes unacceptably large. If the dynamic properties of the measurement system are known, the true physical signal may to some extent be re-constructed. With a parametrized characterization of the system and sampled signals, time-domain digital filters may be utilized for correction. In the present work a general method for synthesizing such correction filters is developed. It maps the dynamic parameters of the measurement system directly on to the filter coefficients and utilizes time reversed filtering. This avoids commonly used numerical optimization in the filter synthesis. The method of correction is simple with absolute repeatability and stability, and results in a low residual error. Explicit criteria to control both the horizontal (time) and vertical (amplitude) discretization errors are presented in terms of the utilization of bandwidth and noise gain, respectively. To evaluate how close to optimal the correction is, these errors are also formulated in relation to the signal-to-noise ratio of the original measurement system. For purposes of illustration, typical mechanical and piezo-electric transducer systems for measuring force, pressure or acceleration are simulated and dynamically corrected with such dedicated digital filters.

\citet{Hessling11} The time-dependent measurement uncertainty has been evaluated in a number of recent publications, starting from a known uncertain dynamic model. This could be defined as the 'downward' propagation of uncertainty from the model to the targeted measurement. The propagation of uncertainty 'upward' from the calibration experiment to a dynamic model traditionally belongs to system identification. The use of different representations (time, frequency, etc) is ubiquitous in dynamic measurement analyses. An expression of uncertainty in dynamic measurements is formulated for the first time in this paper independent of representation, joining upward as well as downward propagation. For applications in metrology, the high quality of the characterization may be prohibitive for any reasonably large and robust model to pass the whiteness test. This test is therefore relaxed by not directly requiring small systematic model errors in comparison to the randomness of the characterization. Instead, the systematic error of the dynamic model is propagated to the uncertainty of the measurand, analogously but differently to how stochastic contributions are propagated. The pass criterion of the model is thereby transferred from the identification to acceptance of the total accumulated uncertainty of the measurand. This increases the relevance of the test of the model as it relates to its final use rather than the quality of the calibration. The propagation of uncertainty hence includes the propagation of systematic model errors. For illustration, the 'upward' propagation of uncertainty is applied to determine if an appliance box is damaged in an earthquake experiment. In this case, relaxation of the whiteness test was required to reach a conclusive result.

\citet{Hessling13a} Deterministic sampling can be used for nonlinear propagation of the statistics of signal processingmodels. Unlike Monte Carlo methods, random generators are not utilized in any stage. The samplesare instead calculated deterministically. Our novel approach generalizes the deterministic samplingtechnique for propagating covariance in the unscented Kalman filter by introducing genericexci-tation matricesdescribing small discrete canonical ensembles. The approximation lies in how wellthe available statistical information is encoded in the discrete ensemble, not how each sample ispropagated. The application and performance of deterministic sampling are illustrated for a typicalstep response analysis of an electrical device modeled with an uncertain digital filter.

\citet{Hessling13b} Statistical signal processing [1] traditionally focuses on extraction of information from noisy measurements. Typically, parameters or states are estimated by various filtering operations. Here, the quality of signal processing operations will be assessed by evaluating the statistical uncertainty of the result [2]. The processing could for instance simulate, correct, modulate, evaluate, or control the response of a physical system. Depending on the addressed task and the system, this can often be formulated in terms of a differential or difference signal processing model equation in time, with uncertain parameters and driven by an exciting input signal corrupted by noise. The quantity of primary interest may not be the output signal but can be extracted from it. If this uncertain dynamic model is linear-in-response it can be translated into a linear digital filter for highly efficient and standardized evaluation [3]. A statistical model of the parameters describing to which degree the dynamic model is known and accurate will be assumed given, instead of being the target of investigation as in system identification [4]. Model uncertainty (of parameters) is then propagated to model-ing uncertainty (of the result). The two are to be clearly distinguished - the former relate to the input while the latter relate to the output of the model.

\citet{Kiviet14improved} In dynamic regression models conditional maximum likelihood (least-squares) coefficient and variance estimators are biased. Using expansion techniques an approximation is obtained to the bias in variance estimation yielding a bias corrected variance estimator. This is achieved for both the standard and a bias corrected coefficient estimator enabling a comparison of their mean squared errors to second order. Sufficient conditions for admissibility of these approximations are formally derived. Illustrative numerical and simulation results are presented on bias reduction of coefficient and variance estimation for three relevant classes of first-order autoregressive models, supplemented by effects on mean squared errors, test size and size corrected power. These indicate that substantial biases do occur in moderately large samples, but these can be mitigated considerably and may also yield mean squared error reduction. Crude asymptotic tests are cursed by huge size distortions. However, operational bias corrections of both the estimates of coefficients and their estimated variance (for which software is provided) are shown to curb type I errors reasonably well.

\citet{Link07} A recently proposed accelerometer model is applied for determining the accelerometer's output to transient accelerations. The model consists of a linear, second-order differential equation with unknown coefficients. It is proposed to estimate these model parameters from sinusoidal calibration measurements, and an estimation procedure based on linear least-squares is presented. In addition, the uncertainties associated with the estimated parameters are determined utilizing a Monte Carlo simulation technique.

\citet{Markovsky07overview} review the development and extensions of the classical total least-squares method and describe algorithms for its generalization to weighted and structured approximation problems. In the generic case, the classical total least-squares problem has a unique solution, which is given in analytic form in terms of the singular value decomposition of the data matrix. The weighted and structured total least-squares problems have no such analytic solution and are currently solved numerically by local optimization methods. We explain how special structure of the weight matrix and the data matrix can be exploited for efficient cost function and first derivative computation. This allows to obtain computationally efficient solution methods. The total least-squares family of methods has a wide range of applications in system theory, signal processing, and computer algebra. We describe the applications for deconvolution, linear prediction, and errors-in-variables system identification.

\citet{Markovsky15cep} Metrology is advancing by development of new measurement techniques and corresponding hardware. A given measurement technique, however, has fundamental speed and precision limitations. In order to overcome the hardware limitations, we develop signal processing methods based on the prior knowledge that the measurement process dynamics is linear time-invariant.
Our approach is to model the measurement process as a step response of a dynamical system, where the input step level is the quantity of interest. The solution proposed is an algorithm that does real-time processing of the sensor's measurements. It is shown that when the measurement process dynamics is known, the input estimation problem is equivalent to state estimation. Otherwise, the input estimation problem can be solved as a system identification problem. The main underlying assumption is that the measured quantity is constant and the measurement process is a low-order linear time-invariant system. The methods are validated and compared on applications of temperature and weight measurement.

\citet{Markovsky15ieee}Dynamic measurement aims to improve the speed and accuracy characteristics of measurement devices by signal processing. State-of-the-art dynamic measurement methods are model-based adaptive methods, i.e., 1) they estimate model parameters in real-time and 2) based on the identified model perform model-based signal processing. The proposed model-free method belongs to the class of the subspace identification methods. It computes directly the quantity of interest without an explicit parameter estimation. This allows efficient computation as well as applicability to general high order multivariable processes.

\citet{Ogorevc16}  describe a procedure for dynamic measurement uncertainty evaluation in order to examine the requirements for clinical thermometer dynamic properties in standards and recommendations. In this study thermistors were used as temperature sensors, transient temperature measurements were performed in water and air and the measurement data were processed for the investigation of thermometer dynamic properties. The thermometers were mathematically modelled. A Monte Carlo method was implemented for dynamic measurement uncertainty evaluation. The measurement uncertainty was analysed for static and dynamic conditions. Results showed that dynamic uncertainty is much larger than steady-state uncertainty. The results of dynamic uncertainty analysis were applied on an example of clinical measurements and were compared to current requirements in ISO standard for clinical thermometers. It can be concluded that there was no need for dynamic evaluation of clinical thermometers for continuous measurement, while dynamic measurement uncertainty was within the demands of target uncertainty. Whereas in the case of intermittent predictive thermometers, the thermometer dynamic properties had a significant impact on the measurement result. Estimation of dynamic uncertainty is crucial for the assurance of traceable and comparable measurements.
Clinical thermometers in intensive care units are used for the continuous measurement of body temperature.

\citet{Palanthandalam10parameter}  investigate the consistency of parameter estimates obtained from least-squares identification with a quadratic parameter constraint. For generality, we consider infinite impulse-response systems with coloured input and output noise. In the case of finite data, we show that there always exists a possibly indefinite quadratic constraint depending on the noise realisation that results in a constrained optimisation problem that yields the true parameters of the system when a persistency condition is satisfied. When the noise covariance matrix is known to within a scalar multiple, we prove that solutions of the quadratically constrained least-squares (QCLs) estimator with a semidefinite constraint matrix are both unbiased and consistent in the sense that the averaged problem and limiting problem produce, respectively, unbiased and true (with probability 1) estimators. In addition, we provide numerical results that illustrate these properties of the QCLS estimator.

\citet{Rhode14recursive} show that the generalized total least squares (GTLS) problem with a singular noise covariance matrix is equivalent to the restricted total least squares (RTLS) problem and propose a recursive method for its numerical solution. The method is based on the generalized inverse iteration. The estimation error covariance matrix and the estimated augmented correction are also characterized and computed recursively. The algorithm is cheap to compute and is suitable for online implementation. Simulation results in least squares (LS), data least squares (DLS), total least squares (TLS), and restricted total least squares (RTLS) noise scenarios show fast convergence of the parameter estimates to their optimal values obtained by corresponding batch algorithms.

\citet{Soderstrom07} gives a survey of errors-in-variables methods in system identification. Background and motivation are given, and examples illustrate why the identification problem can be difficult. Under general weak assumptions, the systems are not identifiable, but can be parameterized using one degree-of-freedom. Examples where identifiability is achieved under additional assumptions are also provided. A number of approaches for parameter estimation of errors-in-variables models are presented. The underlying assumptions and principles for each approach are highlighted.

\citet{Soderstrom18} gives a comprehensive overview of errors-in-variables (EIV) problems in system identification. This problem is about modeling of dynamic systems when all measured variables and signals are noise-corrupted. A number of different approaches are described and analyzed. Representations where errors or measurement noises are present on both inputs and outputs are usually called 'errors-in-variables' (EIV) models. They play an important role when the purpose is the determination of the physical laws that describe the process, rather than the prediction of its future behavior.

\citet{Stewart90SPT} approaches  classical matrix perturbation theory from a probabilistic point of view. The perturbed quantity is approximated by a first-order perturbation expansion, in which the perturbation is assumed to be random. This permits the computation of statistics estimating the variation in the perturbed quantity. Up to the higher-order terms that are ignored in the expansion, these statistics tend to be more realistic than perturbation bounds obtained in terms of norms. The technique is applied to a number of problems in matrix perturbation theory, including least squares and the eigenvalue problem.

\citet{Vaccaro94} develops a perturbation expansion for the subspace spanned by a set of singular vectors. A first-order expansion of this type has recently been developed and used to analyze the performance of direction-finding algorithms in array signal processing. The author derives a new second-order expansion and the result is illustrated with two examples.

\citet{VanHuffel91Book}  give a unified presentation of the TLS problem. A description of its basic principles are given, the various algebraic, statistical and sensitivity properties of the problem are discussed, and generalizations are presented. The TLS applications are surveyed to facilitate uses in an even wider range of applications. The authors provide comparison  with the well-known least squares methods.

\section{Literature relevant for Experimental validation state of the art}

\citet{Dienstfrey14} rederives the heuristic estimation rules that correct the measured quantities for the finite bandwidth of the measurement system response.
The rules are necessary when the measurement device responds slowly compared to the modulation of the input signal. The authors investigate the rules quantitative performance, and as an alternative, demonstrate that regularized deconvolution analysis exhibits more general quantitative utility at the expense of increased measurement burden and analytical complexity.

\citet{Eichstadt16b} explore the relationship between the covariance matrix produced by the Kalman filter, a GUM-compliant uncertainty analysis, and a Bayesian uncertainty analysis. The authors found that all three approaches are compatible for linear systems with known system matrices. However, different results are obtained when the system matrices are not precisely known, or when the system is nonlinear, which might use the extended Kalman filter with linearized uncertainty propagation. The paper shows that the extended Kalman filter linearization and convergence errors can be avoided by applying Monte Carlo.

\citet{Hammersley75} describes the theory and applications of  Monte Carlo Methods

\citet{Hessling10} Dynamic Metrology is devoted to the analysis of dynamic measurements.  As  an  extended  calibration  service,  it  contains  many  novel  ingredients currently not included in the standard palette of metrology. Rather, Dynamic Metrology encompasses many operations found in the fields of system identification, digital signal processing and control theory. The analyses are more complex and more ambiguous than conventional uncertainty budgets of today. The important interactions in non-stationary measurements  may  be  exceedingly  difficult  to  both  control  and  to  evaluate.  In  many situations, in situ calibrations are required to yield a relevant result. Providing metrological services in this context will be a true challenge

\citet{Hessling11} The time-dependent measurement uncertainty has been evaluated in a number of recent publications, starting from a known uncertain dynamic model. This could be defined as the 'downward' propagation of uncertainty from the model to the targeted measurement. The propagation of uncertainty 'upward' from the calibration experiment to a dynamic model traditionally belongs to system identification. The use of different representations (time, frequency, etc) is ubiquitous in dynamic measurement analyses. An expression of uncertainty in dynamic measurements is formulated for the first time in this paper independent of representation, joining upward as well as downward propagation. For applications in metrology, the high quality of the characterization may be prohibitive for any reasonably large and robust model to pass the whiteness test. This test is therefore relaxed by not directly requiring small systematic model errors in comparison to the randomness of the characterization. Instead, the systematic error of the dynamic model is propagated to the uncertainty of the measurand, analogously but differently to how stochastic contributions are propagated. The pass criterion of the model is thereby transferred from the identification to acceptance of the total accumulated uncertainty of the measurand. This increases the relevance of the test of the model as it relates to its final use rather than the quality of the calibration. The propagation of uncertainty hence includes the propagation of systematic model errors. For illustration, the 'upward' propagation of uncertainty is applied to determine if an appliance box is damaged in an earthquake experiment. In this case, relaxation of the whiteness test was required to reach a conclusive result.

\citet{Jafaripanah05}  investigate the application of analog adaptive techniques to the area of dynamic sensor compensation, of which there is little reported work in the literature. The case is illustrated by showing how the response of a load cell can be improved to speed up the process of measurement. The load cell is a sensor with an oscillatory response in which the measurand contributes to the response parameters. Thus, a compensation filter needs to track variation in measurand, whereas a simple fixed filter is only valid at one specific load value. To facilitate this investigation, computer models for the load cell and the adaptive compensation filter have been developed. To allow a practical implementation of the adaptive techniques, a novel piecewise linearization technique is proposed in order to vary a floating voltage-controlled resistor in a linear manner over a wide range. Simulation and practical results are presented, thus, demonstrating the effectiveness of the proposed techniques.

\citet{Ogorevc16}  describe a procedure for dynamic measurement uncertainty evaluation in order to examine the requirements for clinical thermometer dynamic properties in standards and recommendations. In this study thermistors were used as temperature sensors, transient temperature measurements were performed in water and air and the measurement data were processed for the investigation of thermometer dynamic properties. The thermometers were mathematically modelled. A Monte Carlo method was implemented for dynamic measurement uncertainty evaluation. The measurement uncertainty was analysed for static and dynamic conditions. Results showed that dynamic uncertainty is much larger than steady-state uncertainty. The results of dynamic uncertainty analysis were applied on an example of clinical measurements and were compared to current requirements in ISO standard for clinical thermometers. It can be concluded that there was no need for dynamic evaluation of clinical thermometers for continuous measurement, while dynamic measurement uncertainty was within the demands of target uncertainty. Whereas in the case of intermittent predictive thermometers, the thermometer dynamic properties had a significant impact on the measurement result. Estimation of dynamic uncertainty is crucial for the assurance of traceable and comparable measurements.
Clinical thermometers in intensive care units are used for the continuous measurement of body temperature.

Use of load cell sensors

\citet{Olmi16} present a didactic project at the University of Bologna, focused on the development of strain gage load cells, from design to calibration and use. Details are provided on the course units involved and on the levels of teaching, and three case studies are presented. Two case studies deal with the design of decoupled load cells, using simple beam geometry. Students had the opportunity to become confident with the principles of mechanics and to tackle the not trivial activities like calibration by decoupled load application and determination of the related matrices. The third one deals with the development of a load cell for impact loads, where the students had the chance to observe a practical application of the studied Wheatstone bridge connection, to participate in on-field tests, and to better understand the close relationship between experimentation and design.
Many numerical models are nowadays available for the structural analysis of complexly shaped structures. However, a critical problem consists of the estimation of the actual loads that a structure withstands. The most proper way to determine them under service conditions consists of executing in-field tests, where load cells have an important role. This article deals

\citet{Piskorowski08} presents a new method for dynamic compensation of the load cell response using linear time-varying continuous-time filter. Load cells have an oscillatory response which always needs time to settle down. Therefore, it is justified to search a technique for an effective improvement of the sensor response. The paper describes a theoretical implementation of the proposed time-varying filter, and suggests the implementation technique with the aid of which this kind of filter can be implemented in practice. Simulation results verifying the effectiveness of the proposed filter are presented and compared to the traditional time-invariant configuration.

\citet{Rossander15} measures unique blade force on an open site straight-bladed vertical axis wind turbine. This paper presents a method for measuring the tangential and normal forces on a 12-kW vertical axis wind turbine prototype with a three-bladed H-rotor. Four single-axis load cells were installed in-between the hub and the support arms on one of the blades. The experimental setup, the measurement principle, together with the necessary control and measurement system are described. The maximum errors of the forces and accompanying weather data that can be obtained with the system are carefully estimated. Measured forces from the four load cells are presented, as well as the normal and tangential forces derived from them and a comparison with theoretical data. The measured torque and bending moment are also provided. The influence of the load cells on the turbine dynamics has also been evaluated. For the aerodynamic normal force, the system provides periodic data in agreement with simulations. Unexpected mechanical oscillations are present in the tangential force, introduced by the turbine dynamics. The measurement errors are of an acceptable size and often depend on the measured variable. Equations are presented for the calculation of measurement errors

\citet{Kesilmis16} We propose a novel geometric approach to processing of load cell signals for inline weight measurement applications. These systems usually rely on oscillatory load cell signals which always need to time to settle. We describe the theoretical and experimental implementation of a newly proposed geometric approach for load cell signals, and verify the effectiveness of the method through experiments.

\citet{Hernandez06} Load cells are transducers used to measure force or weight. Despite the fact that there is a wide variety of load cells, most of these transducers that are used in the weighingindustry are based on strain gauges. In this paper, an s-beam load cell based on strain gaugeswas suitably assembled to the mechanical structure of several seats of a bus under performance tests and used to measure the resistance of their mechanical structure to tension forces applied horizontally to the seats being tested. The load cell was buried in a broad-band noise background where the unwanted information and the relevant signal sometimes share a very similar frequency spectrum and its performance was improved by using arecursive least-squares (RLS) lattice algorithm. The experimental results are satisfactory and a significant improvement in the signal-to-noise ratio at the system output of 27 dB wasachieved, which is a good performance factor for judging the quality of the system.

\citet{Alaziz17} improve the during-sleep body motion classification using a load cell based system. With a machine learning algorithm, the authors classify a movement into one of 9 classes by extracting 24 features. The authors found that the method it classified correctly 90 \% of movements, and it is convenient for long-term home monitoring.

\citet{Ballo16}  present an innovative six axis load cell for frontal impact tests on a dummy.  The authors measure the acceleration in two different points of the dummy's head, and the impact force input. The authors explain the observations with a first order model of the head and neck, and the neck stiffness and damping have been identified from experimental data.

\citet{Boschetti13} propose a model-based scheme that compensates for low frequency environmental vibrations in automatic weighing machines made of load cells. The authors think that it is ineffective to use low-pass filtering only in load cell measurements to overcome this problem. Their argument is that low cut-off frequencies downgrade machine speed by both introducing delay and increasing filtered signal rise time. The authors instead use mechanical models of the weighing machine and the load cells to process supplementary accelerometer measurements for computing an effective compensation of the environmental vibrations on load cell response effects.

\citet{Guo18} A rule of the effect between the output of a single load cell and a combinatorial structure was proposed by changing the sensitivity of the single load cell. The relevant relationship between the sensitivity of single load cell and the variation of total output of the combinatorial structure was analyzed by the numerical derivation method and experimental validation. In the cases, the output of a single load cell is linear or nonlinear, when the same load was acted on the combinatorial structure. The results showed that the effect was limited and the influence quantity is due to the accuracy and number of load cell. However, it must be controlled in a certain range to avoid additional measurement errors to the combinatorial structure. Moreover, the rule provides an evaluation method with improving the measurement accuracy for the combinatorial structure. Consequently, the rule can be employed as a useful guidance for the product quality when this evaluation method is applied to the load cell selection for the electronic truck scales.

\citet{Zahradka18} demonstrates the feasibility of distinguishing certain positions and movements with load cell sensors. The paper shows that identification of positions and movements for individuals in bed can be used as a tool in a variety of clinical settings. Six positions were classified with an accuracy of 74.9 \% and six movements were classified with an accuracy of 79.7 \%. Stationary positions were characterized by the signals from the four load cells and the coordinates of the center of mass. Movements were characterized by the changes in load cell signals.

Other types of sensors
\citet{DEmilia16} recognize the relevance of three-axis accelerometers and inclinometers in buildings diagnostics. The authors first focus on the sensors calibration, including the low frequency range, 0-10 Hz. Later, they minimize systematic errors and identify the factors that affect uncertainty: the motion positioning of the accelerometer with respect to the vertical axis, the imposed motion law, and the radial positioning of the sensor.

\citet{Diniz17} present a methodology for estimating the uncertainty of measurements in the calibration of temperature sensors in dynamic regimes for industrial applications. The authors use a first-order bare-wire-beaded thermocouple sensor as a case study. The experimental setup is constructed to give the input as a temperature step. The paper describes a methodology to compute the measurement uncertainty considering the uncertainties of all components of the measurement chain.

\citet{Elster07} analyzes the dynamic measurements using a second-order accelerometer dynamic model, assuming that estimates of the model parameters are available after previous dynamic calibration measurements. The authors want to estimate the unknown input of the sensor given its discrete-time output signal, using an FIR filter constructed from the underlying dynamic model. The input  uncertainties are determined considering the uncertainty of the dynamic model parameters estimates, the uncertainty of the sensor's output signal, and the applied signal processing steps. The proposed method allows real-time estimations.

\citet{Link07} A recently proposed accelerometer model is applied for determining the accelerometer's output to transient accelerations. The model consists of a linear, second-order differential equation with unknown coefficients. It is proposed to estimate these model parameters from sinusoidal calibration measurements, and an estimation procedure based on linear least-squares is presented. In addition, the uncertainties associated with the estimated parameters are determined utilizing a Monte Carlo simulation technique. The performance of the proposed modelling approach was tested by its application to calibration measurements of two back-to-back accelerometers (ENDEVCO type 2270 and Br{\"u}el \& Kj{\ae}r type 8305). For each of the two accelerometers, the model was first estimated from sinusoidal calibration measurements and then used to predict the accelerometer's behaviour for two shock calibration measurements. Measured and predicted shock sensitivities were found consistent with differences below 1% in most cases which confirms the benefit of the proposed modelling approach.

\citet{Link09} propose a method  for evaluating the uncertainty associated with the output of a discrete-time IIR filter when the input signal is corrupted by additive noise and the filter coefficients are uncertain. This task arises, for instance, when the noise-corrupted output of a measurement system is compensated by a digital filter which has been designed on the basis of the characteristics of the measurement system. We assume that the noise is either stationary or uncorrelated, and we presume knowledge about its autocovariance function or its time-dependent variances, respectively. Uncertainty evaluation is considered in line with the 'Guide to the Expression of Uncertainty in Measurement'. A state-space representation is used to derive a calculation scheme which allows the uncertainties to be evaluated in an easy way and also enables real-time applications. The proposed procedure is illustrated by an example.

\citet{Markovsky15cep} Metrology is advancing by development of new measurement techniques and corresponding hardware. A given measurement technique, however, has fundamental speed and precision limitations. In order to overcome the hardware limitations, we develop signal processing methods based on the prior knowledge that the measurement process dynamics is linear time-invariant.
Our approach is to model the measurement process as a step response of a dynamical system, where the input step level is the quantity of interest. The solution proposed is an algorithm that does real-time processing of the sensor's measurements. It is shown that when the measurement process dynamics is known, the input estimation problem is equivalent to state estimation. Otherwise, the input estimation problem can be solved as a system identification problem. The main underlying assumption is that the measured quantity is constant and the measurement process is a low-order linear time-invariant system. The methods are validated and compared on applications of temperature and weight measurement.

\citet{Markovsky15ieee} Dynamic measurement aims to improve the speed and accuracy characteristics of measurement devices by signal processing. State-of-the-art dynamic measurement methods are model-based adaptive methods, i.e., 1) they estimate model parameters in real-time and 2) based on the identified model perform model-based signal processing. The proposed model-free method belongs to the class of the subspace identification methods. It computes directly the quantity of interest without an explicit parameter estimation. This allows efficient computation as well as applicability to general high order multivariable processes.

\medskip
\bibliography{../Gus-thesis/Gus-thesis.bib}

\end{document}

### 2020-09-09 ChapterConclusion/ChapterConclusion.tex
One before the very last

@@
 
+\color{blue}
+\subsection*{Continuous Newton method and the error of the Taylor series expansion}
+
+The investigation of the error bounds for the second order Taylor expansion can be conducted using the continuous Newton method. 
+This is proposed as future research and the motivation is as follows.
+
+Neuberger's paper \cite{Neuberger07} introduces five theorems that are a consequence of the continuous Newton method.
+The Newton method aims to find a zero of the function $F: R^n \rightarrow R^n$ by iterating
+\begin{equation} z_{k+1} = z_k - \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
+for $k=0,1,2,\ldots$, and for a given initial $z_0$, assuming that $\left(F' (y) \right)^{-1}$ exists for some $y \in R^n$.
+A domain of attraction is the set of all initial values $z_0$ that lead to a root of $F$ after the convergence $z_0, z_1, z_2,\ldots$.
+The damped Newton method  
+\begin{equation} z_{k+1} = z_k - \delta_k \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
+prevents convergence issues, such as chaotic domains of attraction, with $\delta_1, \delta_2, \ldots \in (0, 1)$. 
+The continuous Newton method is a sequence of damped Newton methods. 
+To implement the continuous Newton method, select $0<T<1$ and perfom $m$ runs of the damped Newton method with $\delta_k = T/m$, for $k=1,2,\ldots, m$.
+The result $z_{k+1}$ is the zero estimation $x_m$, and if the sequence of estimates $x_1, x_2, \ldots$ converges, then it converges to the result of the continuous Newton method.
+
+The objective of the continuous Newton method is to find a function $z: [0, \infty) \rightarrow R^n$, so that
+\begin{equation} z(0) = x \in R^n, \quad z' (t) = - \left(F' (z(t)) \right)^{-1} F(z(t)), \quad t \geq 0, \label{eqn:contNewton}\end{equation}
+subject to the existence of the limit $u=\mathrm{lim}_{t \rightarrow \infty}{z(t)}$, that satisfies $F(u)=0$.
+This implies 
+\begin{equation}  F' (z(t)) z' (t) = - F(z(t)) , \end{equation}
+that is equivalent to
+\begin{equation}  \left(F \circ z\right)' (t) = - F(z(t)) , \end{equation}
+from where we get
+\begin{equation}  F(z(t)) = e^{-t} F(z(0)) . \end{equation}
+Thus, the residual $F(z(t))$ only changes in magnitude and not in direction.
+This property is essential for the results of the theorems in \cite{Neuberger07}. 
+In particular, Theorem 5 gives the conditions for the continuous Newton method to ensure finding a root of the function.
+
+\begin{customthm}{5} \label{thm:five}
+Suppose that each of $H$, $J$, and $K$ is a Banach space, with $H$ compactly embedded in $J$, that $r>0$, and that $G:B_r(0) \rightarrow K$ is continuous as a function on $J$.
+Suppose also that $g$ belongs to $K$ and that for each $y$ in $b_r(0)$ there is an $h$ in $B_r(0)$, where $b_r(0)$ and $B_r(0)$ are open and closed balls in $H$ of radius $r$ centered at 0, such that
+\begin{equation} \lim\limits_{t \to 0+} -{\dfrac{1}{t} \left(G(y+th) - G(y)\right) = g}. \end{equation}
+Then, there exists $u$ in $B_r(0)$ such that $G(u)=g$.
+\end{customthm}
+%$F:B_r(0) \rightarrow K$
+
+The theorem implies that $u$ is a zero of the function  defined as $F(y) = G(y) - g$.  
+Since $G$ is a continuous function and if $\left(G' (y) \right)^{-1}$ exists for each $y$ in $B_r(0)$, a metric for the validity of the Theorem (\ref{thm:five}) hypothesis is the inequality $\left\Vert  \left( G' (y) \right)^{-1} g \right\Vert \leq r$.
+
+To link the data-driven step-input estimation problem with this theorem, we can take the system of linear equations $\widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \bm{\theta}$, then define $G(y) = \widetilde{\mathbf{K}} \bm{\theta}$, and $g = \widetilde{\mathbf{y}}$, so that we have $F(\bm{\theta}) = G(\bm{\theta}) - g$.
+The derivative of the function $G(\bm{\theta})$ is given by $G'(\bm{\theta}) = \widetilde{\mathbf{K}} \mathbf{I}$.
+Using the pseudo-inverse of $\widetilde{\mathbf{K}}$, the metric $\left\Vert  \widetilde{\mathbf{K}}^{\dagger} \widetilde{\mathbf{y}} \right\Vert \leq r$ indicates that the estimate $\hat{\bm{\theta}}$ is found inside a ball of radius $r$.
+
+The questions that remain open are how large can $r$ be?, and how is the radius $r$ related to the perturbation noise variance?
+
+\subsection*{The data-driven input estimation methods with missing data}
+
+The implementation of the data-driven step input estimation method can be compromised when one or more step response samples are not properly acquired.
+The missing data in the construction of the Hankel matrix may be due to sample transference errors, or due to sensor saturation problems.
+The saturation values in the sensor response are not valid for the computation of the step input level because they violate the linearity assumption of the sensor.
+The solution to the most powerful unfalsified model identification problem in presence of missing data was provided in \citet{Markovsky16}. 
+The methods presented in this work are formulated as a matrix completion with the constraint that the to-be-completed matrix must be a rank deficient Hankel matrix.
+The low-rank approximation of Hankel structured matrices with missing data has been studied in \citet{Markovsky13}, where the proposed solution is based on a variable-projection method that requires optimization.
+The software for this method, including a link to obtain it, is described in \citet{Markovsky14}. 
+This method is applicable in matrix completion and system identification problems and has inspired the development of the signal recovery method in for nuclear magnetic resonance spectroscopy described in \citet{Ying18}, and data compression and feature recognition in \citet{Hou17} based on sparse low-rank matrix approximation.
+
+The solutions of the low-rank approximation with missing data problems are found by iterative optimization and, therefore, require large amount of computational power.
+In order to implement data-driven input estimation methods with missing data, there is a need for efficient methods of small complexity $O(n)$ that solve online the optimization problem.
+This is an open problem for future research.
+
+	\subsection*{Compressed sensing methods for recovering missing data}
+
+	Compressed sensing is a methodology that is being developed quite fast and that can offer an alternative to perform measurements with missing data. 
+	The aim of compressed sensing is the recovery of the missing samples of a signal by computing a sparse coefficient vector, under the assumption that the signal belongs to a generating system of larger order.
+	The missing data estimation problem can be expressed as an overdetermined system of equations with a sparse regression matrix. 
+	The review presented in \cite{Eldar2012compressed} provides an introduction of the mathematical modeling of compressed sensing, and gives the conditions for sparse recovery. 
+	The recovery algorithms include convex optimization, greedy and combinatorial algorithms, to name a few.
+
+    According to \cite{Rani18}, compressed sensing is applied when we cannot afford buying enough sensors like in non-visible wavelengths applications, when the measurements themselves are expensive as in neutron-scattering imaging, when measurements take a lot of time as in medical imaging, and when a constrain of the measurement is the power consumption.
+    A particular field of application for compressed sensing is the recovery of bioelectrical signals, as it is detailed in \cite{Craven15}.
+
+    Another good review of compressed sensing is found in \cite{Duarte11}, where it is explained that the  research community is splitted into a group that studies the problems in which the compressed sensing matrices exhibit structure due to the measurement system modeling, and another group which considers that the signal structured representations not only depend on sparsity but instead assume that the signals belong to continuous time and have infinite-dimensional representations.
+The solutions provided exploit these structures in the development of the algorithms.
+Some examples of these structure matrices are described in \cite{Kezhi15} including Gaussian, Toeplitz, Hadamard, and structurally random matrices.
+
+    It might be interesting to compare the solutions of the low-rank approximation methods with compressed sensing in terms of computational complexity and feasibility for online implementations. 
+    Therefore, it is proposed as a future research to develop a compressed sensing algorithm, that can estimate directly the input true value by processing the sensor response to the unknown input, and in presence of missing data.
+
+\color{black}
+
 \begin{comment}

### 2020-09-09 ChapterExperimentalValidation/ChapterExperimentalValidation.tex
One before the very last

@@
 \begin{quote}
-\vspace{-0.75cm}
+%\vspace{-0.75cm}
 \emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I.
@@
 4843-4851, July 2020, \color{black} doi: 10.1109/TIM.2019.2951865.
-\nocite{QuintanaTIM} }\vfill{}
+\nocite{QuintanaTIM} } %\vfill{}
 \end{quote}
 
-%\vfill{}
+\vfill{}
 
@@
 \section{Simulation results}
+
+\color{blue} 
+The simulations described in Chapter 4 were conducted with an illustrative example that is assumed to be a second order weighing sensor.
+A weighing sensor was implemented to relate experimental results with the simulations. 
+As it will be detailed later, the order of the practical sensor is at least is fifth order. 
+Thus, the simulations presented in this section are useful to get insight on the impact of that increasing the model order has on the accuracy of the bias and variance predictions.    
+\color{black}
 
@@
 where $b$ and $v$ are the bias and the variance of the step input estimate, can be applied to the obtained empirical and predicted results and can be compared to the Cram\'er-Rao lower bound (CRLB).
-Figure \ref{fig:MSE_CRLB} shows that $\mathrm{MSE}_\mathrm{e} = b_\mathrm{e}^2 + v_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}  = b_{\widetilde{\mathrm{p}}}^2 + v_{\widetilde{\mathrm{p}}}$ have the same proportionality with respect to the measurement noise variance as the bound for an unbiased estimator $\mathrm{CRLB_{ub}}$.
+Figure \ref{fig:MSE_CRLB_str_exp} shows that $\mathrm{MSE}_\mathrm{e} = b_\mathrm{e}^2 + v_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}  = b_{\widetilde{\mathrm{p}}}^2 + v_{\widetilde{\mathrm{p}}}$ have the same proportionality with respect to the measurement noise variance as the bound for an unbiased estimator $\mathrm{CRLB_{ub}}$.
 For SNR above 35 dB, $\mathrm{MSE}_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ are equivalent, and below 35 dB the difference between them is of less than a factor of 10.
@@
 We obtained an approximation of the $\mathrm{CRLB}_{\mathrm{b}}$ for our biased estimator using the partial derivative of the bias in expression (\ref{eqn:biasST}).
-Figure \label{fig:MSE_CRLB} shows that the bounds for the unbiased and biased estimators are almost equal because the partial derivatives of the bias are negligible w.r.t.
+Figure \ref{fig:MSE_CRLB_str_exp} shows that the bounds for the unbiased and biased estimators are almost equal because the partial derivatives of the bias are negligible w.r.t.
 1 in Equation (\ref{eqn:CRLB_EIV}).
@@
 \includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_3.pdf}
-\caption{\label{fig:MSE_CRLB} The observation instant is fixed at 500 samples.
+\caption{\label{fig:MSE_CRLB_str_exp} The observation instant is fixed at 500 samples.
 The bias and the MSEs decrease for large SNR. 
@@
 
+\color{blue}
+
+	There exist methods for estimating the performance of different models that are useful to choose among the models.
+	The model selection methods can be analytical, such as the Akaike information criterion (AIC) introduced in the paper of \citet{Akaike74}, the minimum description length (MDL) iniciated by the work of \citet{Rissanen78}, or numerical such as crossvalidation.
+	The AIC is defined for a class of models fit by maximum likelihood. 
+	For the linear regression problem with Gaussian perturbation (3.20), the AIC is given by
+	\begin{equation} \mathrm{AIC} = 2 \left( n+1 \right) - \mathrm{ln} \left( \widehat{L} \right) \label{eqn:AIC} \end{equation}
+	where $n+1$ is the number of estimated parameters, and $\widehat{L}$ is the maximum value of the likelihood function of the model.
+	The AIC rewards the goodness of fit by considering the maximum likelihood value, and adds a penalty term that increases the cost with respect to the number of parameters to prevent overfitting.
+	In the opinion of \citet{McElreath20}, the AIC is reliable when the priors overwhelmed by the likelihood, the posterior distribution is approximately multivariate Gaussian, and the sample size is much larger than the number of parameters.
+	Using this criterion for model selection is simple, \citet{Hastie09} suggest just picking the model that provides the smallest AIC value.
+
+	The MDL approach is motivated from an optimal coding viewpoint, aiming to choose the most parsimonious model, that, according to \citet{Hastie09}, is the shortest model that can encode the data for transmision of messages.
+	If we have a model with parameters $\theta$, and data $(K, y)$, with the conditional probability of the outputs $\mathrm{P}(y | \theta, K)$, and assuming the receiver knows the inputs and we transmit the outputs, then the message length required to transmit the outputs is
+    \begin{equation} \mathrm{length} = - \mathrm{ln} \left( \mathrm{P}( \theta ) \right) - \mathrm{ln} \left( \mathrm{P}(y | \theta, K) \right) . \label{eqn:MDL} \end{equation}
+    The length is the sum of the code length required for transmitting the model parameters with the code length required for transmitting the discrepancy between the model and the actual target values.
+    The recommended model is the one that minimizes (\ref{eqn:MDL}).
+	\citet{McElreath20} arguments that model fitting is a form of data compression, where parameters explain the relationships among the data. 
+	The parameters compress the data into a simpler form, with some loss of information about the sample, and 
+	can be used to generate new data by decompression.
+	Learning about the data requires using a simpler model that achieves some compression, but not too much.
+    \color{black}
+
 \section{Conclusions}

### 2020-09-09 ChapterIntroduction/ChapterIntroduction.tex
One before the very last

@@
 \color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_{\mathrm{T}}=0.5 s^{-1}$, caused by a step input of 67 $^{\circ} \mathrm{C}$  when the thermometer was initially at 23 $^{\circ} \mathrm{C}$.
-In this example, we should wait 10 s to have an estimation of the input with a relative error lower than 0.5 \%.
+In this example, we should wait 10 s to have an estimation of the input with a relative error lower than 0.5\%.
 \color{black} }
@@
 The sensor parameters are $k_{\mathrm{d}} = 5 \ \mathrm{N \ s/m}$, $m = 5 \ \mathrm{g}$, and $k_{\mathrm{s}} = 10250 \ \mathrm{N/m}$.
-In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement with a relative error smaller than 0.72 \%.
+In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement with a relative error smaller than 0.72\%.
 \color{black}  }
@@
 \color{blue} The simulation of a force sensor based on a double mass-spring system in cascade also shows that the stabilization after a unit step input is not instantaneous.
-In the plot we infer that the user has to wait 5 s to estimate the force from the sensor response that is 0.25 \% far from its value in equilibrium.
+In the plot we infer that the user has to wait 5 s to estimate the force from the sensor response that is 0.25\% far from its value in equilibrium.
 \color{black}  }
@@
 The step response simulation of this dynamic model results in the sensor step response shown in Figure \ref{fig:pressure}.  
-We have a response to a unit step function that is mostly and increasing exponential that requires 7 s approximately to reach a value that is 98.7 \% ot its maximum value.
+We have a response to a unit step function that is mostly and increasing exponential that requires 7 s approximately to reach a value that is 98.7\% of its maximum value.
 
@@
 
-\section{Original contributions}
+\section{Original contributions \color{blue}and organization of the thesis\color{black}}
 
@@
 This method aims for metrology applications but it lacked an uncertainty assessment, and, therefore, its appropriateness was questionable.
-The first research work conducted was a statistical analysis to elucidate the bias and the covariance of the estimate that the data-driven step input estimation method provides.
-After obtaining these statistical moments, the estimation uncertainty was assessed, and thus, the effectiveness of the method was appraised.
-The second research work was a series of experiments conducted to validate the data-driven step input estimation method in real-life applications.
-The experiments were realizations of the step input excitation using temperature and mass sensors.
-The step input estimation method showed robustness when the measurement noise is not Gaussian and white, as it was assumed in the theoretical analyses.
-The third research work conducted was the estimation of affine input parameters following the data-driven approach for real-time applications. 
-The data-driven method uses exponential weighing in the recursive least-squares solution of a structured errors-in-variables problem, to give preference to recent samples over the older samples.
-The methodology and performance of the data-driven affine input estimation method were compared to those of a maximum-likelihood method.
+\color{blue}A description of the data-driven step input estimation method is presented in Chapter 3\color{black}.
 
 \subsection{Statistical analysis}
+
+\color{blue}Chapter 4 describes \color{black} the first research work conducted that is a statistical analysis of the data-driven step input estimation method, aiming to elucidate the bias and the covariance of the input estimation.
+After obtaining these statistical moments, the estimation uncertainty was assessed, and thus, the effectiveness of the method was appraised.
 
@@
 
-Instead of using a total least-squares (TLS) solution of the structured EIV problem, the step input estimation method uses the recursive least-squares solution (RLS). 
+Instead of using a total least-squares (TLS), \color{blue} or an intrumental variables \color{black}, solution of the structured EIV problem, the step input estimation method uses the recursive least-squares solution (RLS).
 In the review conducted by \citet{Markovsky07overview} it is shown that the TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
-the TLS solution is equivalent to the maximum likelihood (ML) solution when the disturbances of the EIV problems are i.i.d. normally distributed . 
-However, the studies of structured EIV problems presented in \citet{VanHuffel07TLSeditorial} demonstrate that the TLS estimator does not give general results since each structure requires a specific treatment.
-Moreover, \citet{Rhode14recursive} found that the ML estimator leads to non-convex optimization when applied to structured EIV problems, and the global optimum is not guaranteed.
-Also, the computational complexity of TLS and ML inhibits their real-time implementation.
-Thus, the least-squares (LS) estimator is a suboptimal but simple solution to structured EIV problems that admits a recursive form (RLS) suitable for real-time implementation.
+the TLS solution is equivalent to the maximum likelihood (ML) solution when the disturbances of the EIV problems are i.i.d. normally distributed. 
+\color{blue} The classical TLS solution is obtained using singular value decomposition of the augmented matrix built from the regression matrix and the regressor as additional columns\color{black}.
+However, the studies presented in \citet{VanHuffel07TLSeditorial} demonstrate that the structured TLS solutions cannot be generalized since each structure in an EIV problem requires a specific treatment.
+Moreover, the ML estimator of structured EIV problems leads to non-convex optimization, and the global optimum is not guaranteed \cite{Beck10}.
+\color{blue} On the other hand, the instrumental variable (IV) solution of EIV problems use an additional variable.
+The additional variable is an instrument with the purpose of removing the bias of the estimate.
+To do this, the construction of the instrumental variable has to be done in such a way that it is uncorrelated with the perturbations and correlated with the regressor, as it is described in \citet{Soderstrom18}.
+\color{black} 
+Unfortunately, the computational complexity of the TLS, ML, and IV solutions inhibits their real-time implementation.
+
+\color{blue} 
+The solution of the step input estimation problem is preferred to be online, since its applications are in the field of metrology.
+In the literature there are references for online implementations of TLS, ML or IV methods, which are are mainly recursive versions of the solution methods.
+These recursive versions are developed to solve EIV problems for system identification and fault detection.
+We have, for example, the generalized recursive TLS solutions introduced in \citet{Rhode14recursive} and \citet{Rhode14recursivecov}, and the recursive IV solutions developed in \citet{Djouambi12}, \citet{Shang16} and \citet{Gil15}.
+These methods cannot be used as off-the-shelf solutions to the step input estimation EIV problem because they need to be adapted to the specific requirements of the problem.
+One is that the sensor output is the only measured signal, since the input is to-be estimated, and the solution should be simpler, preferable with a linear complexity $O(n)$, instead of the cubic $O(n^3)$ or cuadratic $O(n^2)$ complexities of the mentioned recursive methods, where $n$ is the sensor order.
+The proposal of the data-driven step input estimation method includes the recursive least-squares (RLS) solution.
+Compared with recursive TLS or recursive IV solutions, RLS is a suboptimal but more simple solution to the structured EIV problem with lower computational complexity,  suitable for real-time implementation.
+\color{black}
 
@@
 \subsection{Experimental validation}
+
+\color{blue}Chapter 5 describes \color{black} the second research work that consisted in a series of experiments conducted to validate the data-driven step input estimation method in real-life applications.
+The experiments were realizations of the step input excitation using a mass sensor.
+The step input estimation method showed robustness when the measurement noise is not Gaussian and white, as it was assumed in the theoretical analyses.
 
@@
 
+\color{blue}Chapter 6 describes \color{black} the third research work conducted to estimate the parameters of an affine input, following the data-driven approach for real-time applications. 
+The data-driven method uses exponential weighing in the recursive least-squares solution of a structured errors-in-variables problem, to give preference to recent samples over the older samples.
+The methodology and performance of the data-driven affine input estimation method were compared to those of a maximum-likelihood method.
+
 A dynamic measurement is present when the fluctuations of the measurand impact on the input estimation, such as when a low-bandwidth sensor is excited with a fast changing input.
@@
 
+\subsection{\color{blue}Main findings\color{black}}
+\color{blue}Chapter 7 concludes this thesis work and summarizes the main results obtained in the conducted research.
+
+The data-driven step input estimation method reduces the input estimation time compared with formulations based on models of the sensor.
+This is because the estimation method does not require to invert the sensor dynamics.
+Instead, an algorithm in a digital signal processor finds directly the step input value from the sensor response.  
+Since the method is not rescricted to a particular sensor model, the method can process the step response of any sensor, and provides the input estimation in different applications. 
+Moreover, the method requires relatively low computational power to process the signal of the sensor.
+For these reasons, the data-driven method is suitable for real-time measurements.
+
+After studying the statistical properties of the step input estimation, we know the accuracy (bias) and the precision (variance) of the results.
+This knowlewdge allows one to understand what is the uncertainty on the step input value that the method estimates.
+We have derived formulas to quantify the bias and the variance of the step input estimation.
+With these formulas we can foresee the bias and standard deviation we will obtain with the method given the number of samples processed and the level of the measurement noise.
+
+The implementation of the bias and variance calculation is illustrated with a real life sensor.
+A weighing system, that uses a load cell sensor, is used to generate real life sensor step responses. 
+The measurement noise observed from this weighing sensor is not Gaussian white noise.
+The study of the estimation method performance with real data suggests what is the method order required to set a proper implementation.
+
+It was explored the applicability of data-driven methods to estimate an affine input from the sensor response.
+This input model is observed in measurements where the input changes with a constant rate. 
+The signal processing algorithm, studied in the previous stages of the research, was extended to solve the affine input estimation problem.   
+The performance results of the extended method is compared with a maximum-likelihood method that solves the affine input estimation problem using optimization.
+
+We learnt that the data-driven input estimation methods are simple-to-use methods that run on digital signal processors with few computational resources and estimate the unknown inputs in real time, where the estimation uncertainty can be deduced from the amount of data processed and measurement noise variance.
+
+\color{black}
+
 %\newpage

### 2020-09-09 ChapterPreliminaries/ChapterPreliminaries.tex
One before the very last

@@
 %\vfill{}
+\color{blue}
+The dynamic systems in the real world are non linear. 
+Under certain intervals of operation, the linear approximations can explain the dynamics of the systems. 
+The linear and time invariant systems have been studied in depth and the linear theory is a solid collection of tools to analyze and explain the behaviour of dynamic systems in a wide range of applications.
+On the other hand, the study of non-linear systems is more complex and strongly depends on the type of non-linearity that each particular systems exhibit. 
+When the non-linear modeling does not apport a big amount of benefits into the application study, such as significative accuracy improvement or supporting wide ranges of operation, then is better to resort to the utilization of LTI representations.
+A sensor is a dynamic system whose operation range is preferred to be linear, to facilitate the repeatability and reproducibility of the measurements.
+This means that every time the sensor must deliver identical responses for the same input excitations.   
+\color{black}
 
@@
 It worths stating this fact here since the method studied in this thesis estimates the true value of a step input by processing samples of a sensor step response. 
+\color{black}
+
+\color{blue} 
+The process noise is the representation of the uncertainty in the system state due to unmodeled perturbations that affect the state evolution. 
+The process noise is modeled as an additional input to the system excitation. 
+For a linear time invariant (LTI) system, the response to a sum of two input excitations is the sum of the two responses that correspond to the individual inputs.
+It is demonstrated in \citet{Gubner06} and \citet{Smith11} that the response of an LTI system to a white noise input is colored noise, since the power spectral density of the response is not constant, but instead proportional to the squared magnitude of the system impulse response.
+Nevertheless, in the bandwidth operation of a properly designed measurement system, processing white noise results in approximately white noise output. 
+Moreover, in the present work the measurement noise exists only in the system response signal, because the input is not direcly observed, and the system is not subject to external perturbation. 
+Under these conditions, the mean and variance of the measurement noise $\epsilon$ describe implicitly the process noise entering the system. 
 \color{black}
@@
 The well known bias and covariance results for the LS estimator cannot be invoked because they assume that the additive perturbation only affects the regressor, and that there is no correlation between the regressor and the regression matrix.
+
+\color{blue}
+{\bfseries Note.} The formulation of the data-driven step input estimation method is based on the behavioral theory introduced in \citet{Willems86I}.
+A behaviour of a dynamical system is any set of input-output trajectories that belongs to the system. 
+It is demonstrated in \citet{Willems05} that when an input-output trajectory is persistently exciting enough, a Hankel matrix constructed from this trajectory represents the most powerful unfalsified model.
+This model is unfalsified because it describes exactly all possible trayectories of the system, and is the most powerful because it is the system representation with the minimal complexity.     
+The rank of the Hankel matrix constructed from a trajectory is a metric used to find the order of persistency of the trajectory.
+The rank-deficient Hankel matrix with the smaller number of block rows indicates the trajectory order of persistency. %
+When the system is excited by a step input, the system is equivalent to an autonomous augmented system, and a model for it can be obtained from the Hankel matrix constructed only from the step response.
+
+\color{black}
 

### 2020-09-09 ChapterRampInput/ChapterRampInput.tex
One before the very last

@@
 \begin{quote}
-\vspace{-0.5cm}
+%\vspace{-0.5cm}
 \emph{The results of this chapter were published in Quintana Carapia, G., and Markovsky, I.,
 "Input parameters estimation from time-varying measurements", Measurement Journal, Vol. 153, 2020,
-ISSN 0263-2241, doi:10.1016/j.measurement.2019.107418. \nocite{QuintanaMEAS} }\vfill{}
+ISSN 0263-2241, doi:10.1016/j.measurement.2019.107418. \nocite{QuintanaMEAS} } %\vfill{}
 \end{quote}
 
-%\vfill{}
+\vfill{}
 
@@
 \paragraph{Problem} 
-Given the sequence of measured output observations $\widetilde{\mathbf{y}} = \left( \widetilde{y}(1),\ldots, \widetilde{y}(\color{blue}N\color{black})\right)$, with $\widetilde{y}(t) \in \mathbb{R}$, of a stable linear time-invariant system of order $n$, and static gain ${G}$, generated by an affine input $u(t) = {a} t + {b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u}(t) = \widehat{a} t + \widehat{b}$ approximates $u$(t).
+Given the sequence of measured output observations $\widetilde{\mathbf{y}} = \left( \widetilde{y}(1),\ldots, \widetilde{y}(\color{blue}N\color{black})\right)$, with $\widetilde{y}(t) \in \mathbb{R}$, of a stable linear time-invariant system of order $n$, and static gain ${G}$, generated by an affine input $u(t) = {a} t + {b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in \mathbb{R}$ such that $\widehat{u}(t) = \widehat{a} t + \widehat{b}$ approximates $u$(t).
 The measured observations $\widetilde{\mathbf{y}} = {\mathbf{y}} + \bm{\epsilon}$ are exact sensor responses ${\mathbf{y}}$ perturbed by additive noise  $\bm{\epsilon}$ assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.
@@
 In Figure \ref{fig:rele_dd_40dB_s1} we observe the relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ obtained when SNR = 40 dB. 
-The relative errors are smaller than 5 \% after 400 and 500 samples are processed, i.e., 0.04 s and 0.05 s, respectively.
+The relative errors are smaller than 5\% after 400 and 500 samples are processed, i.e., 0.04 s and 0.05 s, respectively.
 As more samples are collected, the parameter estimation improves.
 Figure \ref{fig:rele_SNR_dd_10000} shows the final value of the relative errors, found at $t=0.1$ s, for the different SNR values considered.
-The relative errors are smaller than 2 \% regardless of the measurement noise level.
+The relative errors are smaller than 2\% regardless of the measurement noise level.
 
@@
 \caption{ \label{fig:rele_dd_40dB_s1} The relative errors of the affine input parameters estimates decrease as the subspace method processes more samples.
-The relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ are smaller than 5 \% after 400 and 500 samples, respectively ($T_s = 0.1$ ms).
+The relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ are smaller than 5\% after 400 and 500 samples, respectively ($T_s = 0.1$ ms).
 }
@@
 \includegraphics[width=0.69\columnwidth]{./ChapterRampInput/fig/Fig_4.pdf} 
-\caption{ \label{fig:rele_SNR_dd_10000} The minimum value of the estimation relative errors obtained with the subspace method is less than 2  \% regardless of the SNR between 20 dB and 60 dB. }
+\caption{ \label{fig:rele_SNR_dd_10000} The minimum value of the estimation relative errors obtained with the subspace method is less than 2\% regardless of the SNR between 20 dB and 60 dB. }
 \end{figure}
@@
 Figure \ref{fig:rele_lo_40dB_s10} shows the average of the observed relative errors in the estimation of the parameters $\widehat{a}$, $\widehat{b}$, $\widehat{x}_{\mathrm{ini,1}}$, and $\widehat{x}_{\mathrm{ini,2}}$\color{black}.
-The convergence of the ML estimates gives relative errors below 5 \% after three iterations.
+The convergence of the ML estimates gives relative errors below 5\% after three iterations.
 The largest relative error observed is in the scale velocity $\widehat{x}_{\mathrm{ini,2}}$ estimate, which is more sensitive than the other optimization variables.
@@
 \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method.
-After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and every of the estimates.
-The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%.
+After three iterations, \color{blue} the average of the relative error is smaller than 5\% for each and every of the estimates.
+The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1\%.
 \color{black} }
@@
 Figure \ref{fig:rele_tv_40dB_s1} shows the relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ computed with the TV filter after processing the sensor transient response.
-The relative error of the slope estimate is below 5 \% after 300 samples but the relative error of the intercept estimate is near 10 \%.
+The relative error of the slope estimate is below 5\% after 300 samples but the relative error of the intercept estimate is near 10\%.
 The convergence rate of the estimate $\widehat{a}$ was similar to that of the subspace method.
@@
 \caption{ \label{fig:rele_tv_40dB_s1} The relative errors of the time-varying filter estimation converge slower than with the subspace method. 
-The relative errors of $\widehat{a}$ and $\widehat{b}$ are smaller than 5 \% only after 800 and 950 samples, respectively.
-With the subspace method the relative errors are below 5 \% after 400 and 500 samples. }
+The relative errors of $\widehat{a}$ and $\widehat{b}$ are smaller than 5\% only after 800 and 950 samples, respectively.
+With the subspace method the relative errors are below 5\% after 400 and 500 samples. }
 \end{figure}

### 2020-09-09 ChapterStatisticalAnalysis/ChapterStatisticalAnalysis.tex
One before the very last

@@
 144, 2020, ISSN 0167-9473, doi:10.1016/j.csda.2019.106893.
-\nocite{QuintanaCSDA} }\vfill{}
+\nocite{QuintanaCSDA} }%\vfill{}
 \end{quote}
 
-%\vfill{}
+\vfill{}
 
@@
 \[ \mathbf{E} = \begin{bmatrix} \mathbf{0}_{\color{blue}N\color{black}-n \times 1} & \bm{\mathcal{H}}(\bm{\epsilon}) \mathbf{D}_{n+1 \times n}^{1,0} \end{bmatrix}, \]
- where $\bm{\mathcal{H}}(\bm{\epsilon}) \in {\rm I\!R}^{\left(\color{blue}N\color{black}-n\right) \times \left(n+1\right)}$ is the block-Hankel matrix of $\color{blue}N\color{black}-n$ rows and $n$ columns
+ where $\bm{\mathcal{H}}(\bm{\epsilon}) \in \mathbb{R}^{\left(\color{blue}N\color{black}-n\right) \times \left(n+1\right)}$ is the block-Hankel matrix of $\color{blue}N\color{black}-n$ rows and $n$ columns
 \[ \bm{\mathcal{H}}(\bm{\epsilon}) = \begin{bmatrix} \epsilon(0) & \epsilon(1) & \cdots & \epsilon(n) \\ \epsilon(1) & \epsilon(2) & \cdots & \epsilon(n+1) \\ \vdots & \vdots & & \vdots \\ \epsilon(\color{blue}N\color{black}-n-1) & \epsilon(\color{blue}N\color{black}-n) & \cdots & \epsilon(\color{blue}N\color{black}-1) \end{bmatrix} , \]
@@
 
-Figure \ref{bias_sigma_NMC_unstr_str_n2} shows the empirical bias and the bias predictions, with their corresponding standard errors, \color{blue} for the unstructured EIV problem\color{black}.
+Figure \ref{fig:bias_sigma_NMC_unstr_str_n2} shows the empirical bias and the bias predictions, with their corresponding standard errors, \color{blue} for the unstructured EIV problem\color{black}.
 The simulation settings for the structured and correlated EIV problem are described in the following subsection.
@@
  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_2.pdf} 
-  \caption{ \label{bias_sigma_NMC_unstr_str_n2} 
+  \caption{ \label{fig:bias_sigma_NMC_unstr_str_n2} 
   We observe the results of the LS solutions \color{blue} for the unstructured EIV problem\color{black}.
@@
 
-The plots \color{blue} in \color{black} Figure \ref{bias_sigma_NMC_str_str_n2} show the empirical bias, the bias predictions (\ref{eqn:biasE}) and (\ref{eqn:biasST}), and their corresponding standard errors, for the structured EIV problem.
+The plots \color{blue} in \color{black} Figure \ref{fig:bias_sigma_NMC_str_str_n2} show the empirical bias, the bias predictions (\ref{eqn:biasE}) and (\ref{eqn:biasST}), and their corresponding standard errors, for the structured EIV problem.
 The empirical bias $b_\mathrm{e}$ is proportional to the perturbation noise variance, and
@@
  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_7.pdf} 
-  \caption{ \label{bias_sigma_NMC_str_str_n2} 
+  \caption{ \label{fig:bias_sigma_NMC_str_str_n2} 
   We observe the results of the LS solutions \color{blue} for the structured EIV problem\color{black}.
@@
  By comparing the different MSEs to the CRLB of the structured EIV problem,
-Figure \ref{fig:MSE_CRLB} shows that the MSEs has the same proportionality as the CRLB with respect to the disturbing noise variance.
+Figure \ref{fig:MSE_CRLB_str_stat} shows that the MSEs has the same proportionality as the CRLB with respect to the disturbing noise variance.
 The MSEs are three times larger than the CRLB.
@@
   \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_11.pdf}
-  \caption{\label{fig:MSE_CRLB}
+  \caption{\label{fig:MSE_CRLB_str_stat}
 The mean squared errors of the LS estimate are close to the Cram\'er-Rao lower bound.
@@
 \color{blue}
+
+\subsection{Simulation results observed when the perturbation noise does not concord with the type of perturbation assumed by the derived expressions}
+
+In this section, the numerical study of the expressions that calculate the step input estimation bias and variance considers the situation when the perturbation noise is not in agreement with the assumptions made during the analysis and the derivation of the predicting expressions.
+In other words, the unstructured perturbation is used in the expressions derived for the structured case, and the structured perturbation is used in the expressions derived for the unstructured case. 
+Two Monte Carlo simulations, each one with $N_{MC} = 10^6$ runs, were conducted and the relevant results are illustrated in Figures \ref{fig:unstrFormula_strData} and \ref{fig:strFormula_unstrData}. 
+It is expected that the performance in these conditions is very far from that obtained when using the expressions in concordance with the perturbation type they were derived for.
+The results confirm this argument, since the relative errors are in general close to 100\%.
+An exception is the variance prediction computed with the formulas derived for the unstructured case, but using structured perturbation. 
+Only in this combination, the relative error of the variance prediction is aroun 10\% for SNR below 70 dB.
+
+\begin{figure}[!htpb]
+  \centering
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_12.pdf}
+  \caption{\label{fig:unstrFormula_strData} \color{blue} We observe the results obtained from the Monte Carlo simulation, when the perturbation noise is structured and we use the predicting expressions that were derived assuming the perturbation noise is unstructured.
+A comparison of the empirical and the predicted bias and variance of the input estimation shows that the prediction formulas fail, since the relative errors between the empirical and the predicted statistical moments are larger than 10\%.}
+\end{figure}
+
+\begin{figure}[!htpb]
+  \centering
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_13.pdf}
+  \caption{\label{fig:strFormula_unstrData} \color{blue} We observe the results obtained from the Monte Carlo simulation, when the perturbation noise is unstructured and we use the prediction expressions that were derived assuming the perturbation noise is structured.
+A comparison of the empirical and the predicted bias and variance of the input estimation shows that in these conditions, the derived expressions fail to predict the statistical moments.
+If the perturbation noise does not have the structure assumed by the prediction expressions, the relative error remains high above 10\% in all the SNR interval.
+}
+\end{figure}
+
 \subsection{Simulation results for the structured IV solution}
@@
   \centering
-  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_12.pdf}
-  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_13.pdf}
   \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_14.pdf}
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_15.pdf}
+  \includegraphics[width=0.69\columnwidth]{./ChapterStatisticalAnalysis/fig/Fig_16.pdf}
   \caption{\label{fig:bvMSE_IV} \color{blue} In a simulation, the solutions of the structured EIV and IV problems are empirically compared.

### 2020-09-09 internal defense feedback/Answer - continuous Newton method.tex
One before the very last

--- 2020-08-21 internal defense feedback/Answers-jury-members-report - BK2.tex
@@
 
-\title{Jury members' comments on the private PhD defense of \linebreak Gustavo Quintana-Carapia}
+\title{Question raised during the private PhD defense of \linebreak Gustavo Quintana-Carapia}
 
@@
 
-Thank you to the jury members for reviewing the thesis manuscript and for giving suggestions for improvement. 
-The manuscript changes due to the comments are highlighted in \color{blue} blue fontcolor\color{black}.
-
-\section*{General comments}
-
-\begin{itemize}
-	\item The manuscript text should better position the PhD to related work. Improve the literature review and discuss (and cite) the state of the art.
-
-	{\bfseries The introduction of the chapters now includes the state of the art to position the PhD work with respect to the literature.}
-
-	\item  Add more clearly the motivation for the PhD research work by including motivational examples to the introduction.
-
-	{\bfseries Motivational examples have been added to the Introduction. }
-
-	\item  Add a discussion with comparison of existing EIV methods.
-Give, with the positive and negative aspects in mind, a better justification for why you use a (R)LS method, and not, e.g., (R)IV method.
-
-	{\bfseries In Section 4.1.4, a discussion that compares IV methods with EIV methods has been added.}
-
-	\item  Discuss more clearly the aspects of scalability of the methods and their limitations in terms of model order, size of data sets (e.g., related to the inversion of LS matrices).
-	\item  The equations on pages 12 and 13 seemingly contain errors and should be corrected.
-
-	{\bfseries The equations on pages 12 and 13 have been corrected.}
-
-	\item  The figure shown on page 66 was made with a single experiment; it is desirable to see the (averaged) effect on several experiments.
-
-	{\bfseries A series of 100 runs of the maximum-likelihod estimation method has been simulated and the results have been added to the manuscript in Section 6.3.2.}
-
-\end{itemize}
-
-\section*{Lyudmila Mihaylova}
-
-\begin{itemize}
-	\item You develop results for uncertainty quantification; what type of uncertainties  can your approach deal with and which levels of noise can you consider?
-Elaborate on SNR levels; I’m also very interested in scalability, so how big the matrices can be.
-	\item  Concerning the scalability to data: eqn 3.6: how scalable is the approach in all aspects (order, data, linked to inversion for LS matrix); can you deal with hundreds, thousands of data?
-
-	{\bfseries The following paragraph has been added in Section 3.2 to discuss the scalability of the data-driven step input estimation method.}
-
-	\color{blue}
-    The computational complexity of the RLS algorithm solution to the system of equations (3.14) is $O \left( \left( n+1 \right)^2 \right)$.
-    The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
-    From there on, the RLS algorithm updates the solution with linear complexity.
-    Therefore, even though a large number of samples $N$ makes the matrix in the system of equations (3.14) increase in the number of rows, the estimation of the solution does not increase in complexity.
-    The data-driven step input estimation method is then scalable for any sensor of order $n$, and can be executed in devices with limited computational resources, provided that the computation of the  $n+1 \times n+1$ inverse matrix is feasible.
-    \color{black}
-
-	\item  What recommendation up to scale/order can method work with.
-What are the limitations?
-It would be good to have more precise discussion about these aspects in dissertation: what do you mean by large scalability: 100, $10^4$, ...?
-A more detailed discussion is desirable.
-	\item  On slide 25 (noisy graphs) it would be good to see a statistical analysis (more experiments than one).
-
-    {\bfseries 100 runs of the maximum-likelihod estimation method have been simulated and their results have been added to the manuscript in Section 6.3.2.}
-
-	\color{blue} 
-    The maximum-likelihood (ML) method estimated the affine input parameters from the sensor model transient response.
-    A total of 100 runs of the method were obtained, using different realizations of measurement noise.
-    The measurement noise variance was set to have an SNR of 40 dB.
-    The ML method used the first 50 samples to initialize the optimization variables and updated the variables every $N_{\mathrm{s}} = 5$ samples.
-    Figure \ref{fig:rele_lo_40dB_s10} shows the average of the observed relative errors in the estimation of the parameters $\widehat{a}$, $\widehat{b}$, $\widehat{x}_{\mathrm{ini,1}}$, and $\widehat{x}_{\mathrm{ini,2}}$.
-
-    \renewcommand{\thefigure}{6.8}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_7.pdf} 
-    \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method.
-After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and everyone of the estimates.
-The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%.
-\color{black} }
-    \end{figure}
-
-    Figure \ref{fig:cov_lo_40dB_s1} shows \color{blue} the average \color{black} of the variances that were observed on the diagonal of the covariance matrix $\mathbf{J}$. 
-    We can see that the estimation variances decrease as more samples are processed.
-    Moreover, the estimation variances of $\widehat{a}$ and $\widehat{b}$ obtained with the ML method are lower than the corresponding estimation MSE errors obtained from a Monte Carlo simulation of the subspace method.
-
-    \renewcommand{\thefigure}{6.9}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_8.pdf} 
-    \caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian.
-\color{blue} The estimates $\widehat{x}_{\mathrm{ini,1}}$ and $\widehat{a}$ have the smaller and larger variances, respectively.
-\color{black}  }
-    \end{figure}
-
-    \color{black}
-
-\end{itemize}
-
-\section*{Stephane Chretien}
-
-\begin{itemize}
-	\item The Taylor expansion is computed of order 2, what is the accuracy of expansion with respect to system, noise level,...
-Can you come up with a priori bound on error that will allow you to assess the quality of the expansion?
-Can you predict where the solution is going to be before computing Taylor expansion?
-
-	{\bfseries An a priori bound on the error introduced by the Taylor series expansion can be obtained in terms of the spectral radius of the matrix $\mathbf{M}$.
-In Section 4.1 the following description has been added: }
-
-    \color{blue} An a priori bound on the error introduced by the second order Taylor series expansion can be expressed in terms of the spectral radius $\rho(\mathbf{M})$.
-    Considering that the Taylor series expansion is an infinite sum we have that
-    \begin{equation} \tag{4.4} \begin{aligned} (\mathbf{I} + \mathbf{M})^{-1} &= \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \\
-    &=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \right) \\
-    &=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \end{aligned} \end{equation} 
-    Therefore we can obtain a bound for the second error expansion error if we take a matrix norm as follows
-    \begin{equation} \tag{4.5} \begin{aligned} \| (\mathbf{I} + \mathbf{M})^{-1} - ( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 ) \| &=  \| \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \| \\
-    & \leq \| \mathbf{M}^3 \| \| (\mathbf{I} + \mathbf{M})^{-1} \| \leq \dfrac{\| \mathbf{M} \|^3}{ 1 -  \| \mathbf{M} \|} = \dfrac{\rho(\mathbf{M})^3}{1 - \rho(\mathbf{M})}  \end{aligned} \end{equation}
-	\color{black}
-
-	\item  Concerning other sorts of prior estimations, e.g., given by explicit function theorems.
+{\bfseries Concerning other sorts of prior estimations, e.g., given by explicit function theorems.
 I wonder if you could get something and apply this bound in order to control the bound of the 2nd order expansion.
@@
 Neuberger's results (quite old now) but may be handy for this kind of problem.
-This can be investigated further.
+This can be investigated further.}
+
+\vspace{0.7cm}
+The investigation of the error bounds for the second order Taylor expansion can be conducted using the continuous Newton method. This is proposed as future research work in Chapter 7.
+\vspace{0.7cm}
 
@@
 Using the pseudo-inverse of $\widetilde{\mathbf{K}}$, the metric $\left\Vert  \widetilde{\mathbf{K}}^{\dagger} \widetilde{\mathbf{y}} \right\Vert \leq r$ indicates that the estimate $\hat{\bm{\theta}}$ is found inside a ball of radius $r$.
-    Thus,  
+    The questions that remain open are how large can $r$ be?, and how is the radius $r$ related to the perturbation noise variance? 
 \color{black}
 
-	\item  About trying to find a good transformation from LS to EIV model.
-I am wondering if you can think of other more accurate starting points instead of LS.
-Closer to the problem you want to study, but still preserving computational lightness?
-The closer you start from objective, better the Taylor expansion would be..
-Something to investigate as well...
-	\item  About robustness: is it an issue in practical problems with heavy tailed noise, outliers,...
-Perhaps the use of median of means plugged into your approach might produce interesting alternatives for outliers or heavy tailed data.
-But this is an intricate thing to deal with.
-Might be interesting to ponder this in future.
-	\item  On deep learning: could this be an alternative?
-\end{itemize}
-
-\section*{Guillaume Mercere}
-
-\begin{itemize}
-	\item In the first or second sentence: “users cannot wait for a steady state regime” .
-Do you have specific examples of real industrial systems for which it is necessary to have continuous read-out before steady-state is reached?
-- Further on the industrial part and selling sensors with data-driven parts.
-Can you explain to experts that your solution is better?
-Sensors will be used for ages; why not spend one day to do system identification and get a reliable model to make a compensator, instead of using your approach?
-
-	{\bfseries The fast weighing of parcels in the post office is a typical example of the real life systems where the time is limited to perform the measurements.
-Another one that we see nowadays is the temperature measurement of people in airports or train stations to prevent the spread of corona virus.
-The data driven method can be implemented in portable devices with limited computational resources.
-In industrial plants were a time slot can be devoted to identify the measurement system and implement a compensation method, the solution might be useful only for a certain amount of time.
-It is assumed that the model approximation deviates due to changes in the system elements, leading to inaccuracies in the inpiut estimation.
-There is a need for adaptive methods to avoid stopping the production lines to repeat the system identification procedure.
-The data-driven method is such an adaptive method.
-}
-
-	\item  Chapter 2, Section 2.2.1.
-Can you give me some classic solutions/algorithms instead of LS in general?
-(Say I'm not familiar with EIV or TLS?) Usually with LS problems: multiply with $K^\top$ and compute pseudoinverse.
-Do you know another solution where they don't use K on the left-hand-side but something else?
-To get rid of noise by using a matrix on the left-hand-side, is related to EIV.
-There are some recursive solutions for this.
-It could be interesting to compare with such IV solutions.
-	\item  In Chapter 3: at beginning you say to use linear systems theory to model a sensor. Can you discuss this assumption: is it true in practice? Can you guarantee it behaves like an LTI system?
-	\item  Equation 3.1: an LTI system, described with SS representation.
-You only put output noise on this representation.
-People sometimes also add process noise.
-Why?
-How to modify your approach if there is process noise?
-Is it still white in this case?
-	\item  On page 12 standard approach top of page equation: there are some mistakes in the equations!
-there is no epsilon (noise) in the equation.
-But then you do an SVD of H and estimate from there.
-How does it work if you have noise?
-Have you tried implementing it and see what you get?
-
-	{\bfseries The step input estimation based on singular value decomposition works only with exact data.
-When  noise perturbes the sensor response the method cannot guarantee the step input estimation.
-The singular value decomposition of a Hankel matrix constructed from a perturbed sensor response provides only an approximation of the singular values and the singular vectors of this Hankel matrix.
-Then, the step input estimation obtained with this method from perturbed sensor response is only a rough estimation.
-}
-
-	\item  On page 12: I’m not sure the equations at the end of the page are correct, e.g., Y=G ubar ...
-not sure there is x; same for rest; There seem to be many mistakes here and next pages.
-Be careful with the equations!
-
-	{\bfseries The equations have been reviewed and the typos were corrected.}
-
-	\item  On page 13, you assume persistency of excitation of order $L$. How do you check this? In your case the input is constant; how to guarantee that the assumption is met?
-
-	{\bfseries The persistency of excitation of a signal is verified by looking at the rank of the Hankel matrix constructed from the consecutive samples of this signal.
-When the Hankel matrix of $L+1$ block rows has a rank lower or equal than L, then we say that the signal is persistently exciting of order $L$.
-A step input has a very low order of persistency $L=1$, but the step respose of a system has a larger order of persistency.
-In fact, we assume that if the persistency of excitation of the step response is sufficiently high, then the columns of the Hankel matrix constructed from the step response span a linear space with all the responses of the equivalent augmented autonomous system where the step input is considered an augmented initial condition.}
-
-	\item  In Chapter 4, p 17, Section 4.1.1: You assume that perturbations are IID and must be normally distributed.
-Why using an assumption of normally distributed?
-I would think that the kind of distribution is important for CRLB, but I’m not sure you need it for eqns 4.8, 4.9, 4.10,...
-IID is important, normality perhaps not.
-
-	{\bfseries The following paragraph has been added in Section 4.1 to discuss on the need for the normality assumption.  }
-
-	\color{blue}
-	The normality assumption is necessary to provide more prior information into the method formulation.
-    The second order Taylor series expansion (4.4) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
-    Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
-    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance, which can be disregarded since $3 \sigma_{\bm{\epsilon}}^4 \ll \sigma_{\bm{\epsilon}}^2$\color{black}.
-
-	\item   In equation 4.11 you have $\sigma_eps^2 + \sigma_e^2$; how to get in practice $\sigma_e$? Do you have to do this each time? Recursively?
-
-	{\bfseries The measurementy noise variance is prior information and in Chapter 5 we describe two ways of estimating $\sigma_{\bm{\epsilon}}$ In summary, the measurement noise can be extracted fom the sensor ressponse once the transient effects are smaller than the noise floor, or from the sensor response power spectrum.
-The online estimation of the measurement noise, simultaneously with the online step input estimation is a topic for future research.
-}
-
-	\item   Switch to 4.2, simulation results: first part dedicated to K, randomly generated; not sure if still possible, but in all your results it would be nice to have somewhere explicitly K = ...
-(even if randomly generated).
-You use only one realization also for matrices A, B, C, D in the text, so it can be added explicitly.
-	\item   in Fig 4.1 the bottom figure also discusses the structured case. Why do you talk about structured results in a section about the unstructured case?
-
-	{\bfseries Section 4.2 has been improved to describe the result next to its discussion.  }
-
-	\item   In 4.1 uses a constant K, but in 4.2.2 you use A, B, C, D, without any links to K.
-You compare the two curves generated from different data sets… It is strange to compare them.
-I don't see the link between both.
-
-	{\bfseries Section 4.2 has been reestructured and the simulation for the unstructured case were redone to establish a link between the two cases.
-The exact data is the same in the two cases and the perturbation noise follows the corresponding rules of each case, with the same variance noise levels.
-In this way the comparison between the two cases makes more sense.
-}
-
-	\item   To show comparison between structured and unstructured case, use unstructured results for structured case; then use unstructured results for structured case and cover all the possibilities.
-Now you compare curves related to two different simulations.
-
-	{\bfseries \color{red} Interesting question, I want to try it this evening \color{black}}
-
-	\item   I was surprised to read in Chapter 5 something very close to what I read in chapter 4.
-In both cases generated A, B, C.
-In 5.1 again oscillating 5th order system, similar conclusions, ...
-What is the added value of 5.1 wrt 4.2.2?
-
-	{\bfseries The derivation described in Chapter 4 was conducted with a second order mass-spring-damper weighing system example on mind, and the results observed were submitted in the corresponding paper.
-Afterwards, when we moved to the practical implementation, we faced the increase of the model order after setting up the weighing experiment.
-The lowest order we can assume is fifth order.
-So we decided to repeat the simulations and these new results serve two purposes: a transition from simulation to experimental results, and also to get an insight on the impact of the bias and variance predictions after increasing the model order.
-}
-
-	\item   On the impact of selected order on results.
-You were talking about a related bias/variance trade off.
-So, do you know in the literature some solutions to avoid overfitting?
-Can you cite something?
-Can you use something in your case to test for a good order of system?
-	\item   Do you know of cross-validation: split data into two parts or several parts; some for estimation, some for validation?
-Have you tried this?
-Can this be used here to select good orders?
-Better results?
-Influence of system order on bias/variance; if I only use one dataset it is complicated and usually you can increase complexity to mimic a data set, but to avoid overfitting (too high order) use a test/validation data set.
-with this i should see what is the best value...
-Have you tried this?
-	\item   In Section 6.2.2.
-Why do you call it a max likelihood approach?
-To me it is not one.
-Why do you call it like this?
-Where is the likelihood?
-Depending on which likelihood (eg.
-Gaussian, ...) it leads to LS.
-Don't' call it ML if you dont give distributions on ...
-In this case it is better to call it a nonlinear minimization procedure.
-Not clear why you call it ML.
-
-	{\bfseries The method description in Section 6.2.2 has been modified to explain why the method is a maximum likelihood method.  }
-
-	\item   Concerning the ML approach in simulations, you say that it takes 30 seconds to complete.
-What is done exactly in these 30 seconds; how many samples, realizations,...
-I am a bit surprised you require 30 sec for 2 parameters.
-\end{itemize}
-
-\section*{Nikos Deligiannis}
-
-\begin{itemize}
-	\item When solving algorithms.
-Do you opt for closed-form solutions, or do you use optimization based algorithms?
-How to solve LS problems?
-Do you solve them using pseudo-inverse or do you do gradient descent?
-	\item  Today there is an overwhelming amount of data sensed and communicated; How would you apply your method in a scenario where dimensionality increases significantly?
-This relates to the questions of previous jury members; what is the implication of using numerical versus closed form solutions?
-	\item  What if the $K^\top K$ matrix is not invertible?
-What if you have fewer samples than unknowns; what can you do?
-Solutions that assume structure like in compressed sensing, regularization,...
-can you use them, what would be the impact of that?
-	\item  You mention that you improve complexity with your approach, but I missed a complexity analysis. Did you do this? Anything in processing time, flops,...?
-	\item  You mentioned deep learning could be applied; could you show how?
-An issue is that neural nets are trained in high SNR data, but applied in engineering with lower SNR.
-This clogs performance; any ideas?
-Is this a limitation?
-	\item  You mention that you consider Kalman filters; have you seen approaches where they write Deep Learning in form of Kalman Filter; Could this be applicable?
-	\item  Recommendation: if you revise the thesis: maybe change "I propose" to "we propose".
-
-	{\bfseries The formulation “I propose” was replaced by “we propose”. }
-
-\end{itemize}
-
-\section*{Philippe Dreesen}
-
-\begin{itemize}
-	\item Your main assumption is that a sensor outputs a steady state value that is proportional to the to-be-measured quantity.
-This means that you assume that the sensor is a linear system.
-What if this assumption is not met, and the sensor has a saturation for high values?
-Can you still use your methods?
-	\item  Sensors are physical devices and deliver a continuous-time output. The models that you use are in discrete-time. Are there any risks that you miss information by not sampling fast enough?
-
-	{\bfseries In Chapter 3 has been added a discussion on the step invariance transformation.
-This defines the conditions of the input for describing exactly the system dynamics in discrete time after sampling the continuous system response.
-If the input is stepwise continuous, then the zero order hold sampling of the step response leads to an exact discrete time representation of the system.
-The method under study in this thesis was developed for step input, then we can rely on the discrete time representation.
-There exist corresponding invariant transformations for piecewise linear inputs and for piecewise polynomial inputs that lead to exact representations from first and second order sampling of the system response, respectively.
-}
-
-	\item  Related to earlier questions about compressed sensing.
-There are some recent results on the connections between sparsity and low-rank Hankel matrix completion and approximation.
-Say that you have an issue with missing data, e.g., from sensor outages.
-Would it be possible to plug in a procedure that does structured matrix completion and approximation into your methods?
-\end{itemize}
-
-\section*{Roger Vounckx}
-
-\begin{itemize}
-	\item Two practical questions.
-On page 35-49 you consider a weight of 138.x grams.
-You sample relatively slowly (4 kHz) and after 500 samples we can consider that we have correct estimation.
-If I have a container that fills up and weight is constantly changing.
-Can we tackle this?
-	\item  Consider an optical fiber link with a detector capable of 100 million measurements per second.
-Can I speed up the link by using your method?
-You have to do calculations: how high can you go in speed before running into troubles with DSP?
-\end{itemize}
-
-\section*{Rik Pintelon}
-
-\begin{itemize}
-	\item Returning to a question by Guillaume Mercere: on page 17; do you need normality?
-You made 2nd order taylor series expansion.
-Which moment do you need to estimate?
-Look at the equations page 16.
-There are 3rd, 4th order moments,...
-you need to know something more about the noise; this is one of the reasons to make Gaussian assumption: 3rd order moment is 0 (symmetry); 4th order for Gaussian is easy to have an explicit relationship.
-
-    {\bfseries Section 4.1 has been appended with a paragraph to discuss on the need for the normality assumption.  }
-
-	\item  Returning to a question by Philippe Dreesen on missing data.
-If you can deal with missing data, you can deal with saturation in sensors: you just remove the saturations.
-What is the rank of the Hankel matrix?
-Is it full rank or not?
-The question is if matrix completion can help?
-It works in low-rank approximations; not if the matrix is of full rank.
-\end{itemize}
-
-\section*{Ivan Markovsky}
-
-\begin{itemize}
-	\item About scalability of algorithms: You said in the very beginning that it can be very fast compared with system identification.
-This is the point: can you think of a good answer for these questions on scalability?
-Nikos Deligiannis mentioned computational analysis.
-You can directly say for (R)LS what is the computational complexity.
-Order three if full matrix inversion, now per iteration only linear.
-So you can scale to very large data sizes.
-
-	{\bfseries The last paragraphs of Section 3.2 have been included to discuss the scalability of the data-driven step input estimation method.  }
-
-\end{itemize}
-
-\section*{Detailed comments provided by jury members}
-
-\section*{Guillaume Mercere}
-
-\begin{itemize}
-	\item It is important for me to know why these new developments are generated. That is the reason why I suggest adding sections or paragraphs dedicated to
-    \begin{itemize}
-        \item the reasons why, in the industry, we need such a new technique.
-As I said during the defense, do you have a list of sensors for which waiting for the steady regime for getting the measurement is not possible?
-Why do you need a fast algorithm?
-What are the reasons why industry people need fast algorithms?
-In the end, it seems that, for you, it is more important to have a fast biased result than an unbiased estimate.
-Why?
-What do you mean by fast?
-        \item  the literature on EIV estimation problems.
-As shortly mentioned in Chapter 2, there are solutions in the literature (by the way, you never mention the recent book written by Torsten Soderstrom in your document which is strange for me).
-But you do not describe them (even shortly) and, more importantly, you do not criticize them.
-Not in general but, at least, explain why you do not use them, what are their advantages and drawbacks and, finally, w.r.t.
-this list of positive and negative points, why you choose a (R)LS solution.
-E.g., why not using a recursive instrumental variable solution instead?
-
-        {\bfseries A description of IV methods has been added in Section 4.1.4}
-
-        \color{blue}
-    \subsection*{4.1.4. Instrumental variables formulation of the data-driven step input estimation method}
-
-    The errors-in-variables problem formulation (3.18) of the data-driven step input estimation method can be converted into an instrumental-variables problem.
-    A guide to do this conversion is found in \cite{Soderstrom18} and in \cite{Pintelon12Book}. 
-    The instrumental variable $\widetilde{\mathbf{Z}}$ has to be selected so that the normal equations (3.19) are expresed instead as
-    \begin{equation} \tag{4.22} \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{y}} . \label{eqn:neq_siv} \end{equation}
-    The selection of the instrumental variable has to be done in such a way that is correlated the regression matrix $\widetilde{\mathbf{K}}$, and uncorrelated with the perturbations $\bm{\epsilon}$.
-    Assuming that the block-Hankel matrix $\widetilde{\mathbf{K}}$ is given by
-    \begin{equation} \tag{4.23} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(n+1) & \Delta \widetilde{y}(n+2) & \cdots & \Delta \widetilde{y}(2n) \end{bmatrix} , \label{eqn:matrixK_r} \end{equation}
-    then, the instrumental variable $\widetilde{\mathbf{Z}}$ should not contain the elements that construct $\widetilde{\mathbf{K}}$, i.e., the elements of the vector $\left( \widetilde{y}(1), \ldots, \widetilde{y}(2n+1) \right)^\top$.
-    A possibility to generalize this idea is expressing the instrumental variable as
-    \begin{equation} \tag{4.24} \widetilde{\mathbf{Z}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-4n) & \Delta \widetilde{y}(t-4n+1) & \cdots & \Delta \widetilde{y}(t-3n-1) \\ G & \Delta \widetilde{y}(t-4n+1) & \Delta \widetilde{y}(t-4n+2) & \cdots & \Delta \widetilde{y}(t-3n) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-3n) & \Delta \widetilde{y}(t-3n+1) & \cdots & \Delta \widetilde{y}(t-2n-1) \end{bmatrix} , \label{eqn:matrixZ_t} \end{equation}
-    for a regression matrix
-    \begin{equation} \tag{4.25} \widetilde{\mathbf{K}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-2n+1) & \Delta \widetilde{y}(t-2n+2) & \cdots & \Delta \widetilde{y}(t-n) \\ G & \Delta \widetilde{y}(t-2n+2) & \Delta \widetilde{y}(t-2n+3) & \cdots & \Delta \widetilde{y}(t-n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-n+1) & \Delta \widetilde{y}(t-n+2) & \cdots & \Delta \widetilde{y}(t) \end{bmatrix} , \label{eqn:matrixK_t} \end{equation}
-    for $t = 4n+1, \ldots, N$.
-
-    The difference with the EIV formulation is that this implementation of the IV method requires to wait more, until $4n+1$ samples of the sensor response are read, instead of $2n+1$ samples. 
-    For sensors of small order this twice-the-time waiting does not represent a big change, but as the sensor order increases, the need for waiting can turn the method into impractical since a lot of time should pass before starting to estimate the input.
-
-    There exists a generalized instrumental variables method (GIVE) that was proposed to identify dynamic models from input/output data. 
-    GIVE is a bias-eliminating method that uses a generalized instrumental vector constructed from delayed values of the input and output signals \cite{Soderstrom18}.
-    Since the aim of this thesis is to analize a method that directly estimates the input from the observed sensor output, without identifying explicitly a dynamic model of the sensor, the GIVE method is out of the scope of this work.
-    A future research can include studying the formulation of a data-driven input estimation method in the spirit of the GIVE method. 
-    Nevertheless, a brief description of the GIVE method and its recursive version is given in the following paragraphs.
-
-    \subsubsection*{4.1.4.1 Generalized instrumental variables method}
-    For an errors-in-variables problem expressed as
-    \begin{equation} \tag{4.26} \widetilde{y}(t) = \widetilde{\bm{\phi}}(t) \bm{\theta}   \end{equation}
-    where $\widetilde{y}(t)$ is the output perturbed by measurement noise, $\bm{\theta}$ is the vector of to-be-estimated parameters, and $\widetilde{\bm{\phi}}(t)$ is the regressor vector constructed from delayed samples of the measured input $\widetilde{u}(t)$ and output $\widetilde{y}(t)$.
-    The least-squares estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is the solution of the normal equation
-    \begin{equation} \tag{4.27} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} \widehat{\bm{\theta}}_{\mathrm{LS}} = \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} , \label{eqn:norm} \end{equation}
-    where $\widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}}$ is the covariance matrix of the regressor vector, and $\widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}}$ is the cross-covariance matrix of the regressor vector and the output.
-    When the number of samples becomes infinitely large, these covariances matrices can be expressed as a sum of the contributions from exact data and measurement noise contributions
-    \begin{equation} \tag{4.28} \begin{aligned}  \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} &= \mathbb{E} \{ \widetilde{\bm{\phi}} \widetilde{\bm{\phi}}^\top \} = \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \\ \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} &= \mathbb{E}\{ \widetilde{\bm{\phi}} \widetilde{y}^\top \} = {\mathbf{r}}_{\bm{\phi} y} + {\mathbf{r}}_{\bm{\eta} \bm{\epsilon}} , \end{aligned} \end{equation}
-    where we are assuming that $\widetilde{y} =  y +\bm{\epsilon}$, $\widetilde{\bm{\phi}} = \bm{\phi} + \bm{\eta}$, and $\mathbf{r}_{\bm{\phi} y} = \mathbf{R}_{{\bm{\phi}}} \bm{\theta}$.
-    The bias of the least-square solution can be expressed as 
-    \begin{equation} \tag{4.29} \mathbf{R}_{\widetilde{\bm{\phi}}} \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) = \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{R}_{\widetilde{\bm{\phi}}} \bm{\theta} = \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} - \mathbf{R}_{\bm{\eta}} \bm{\theta} , \end{equation}
-
-    If we assume that the measurement noise is white for both the input and output, 
-    \begin{equation} \tag{4.30} \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} = \mathbb{E}\{ \bm{\eta} \bm{\epsilon}^\top \} = \mathbf{0} \quad \mathrm{and} \quad \mathbf{R}_{\bm{\eta}} = \begin{bmatrix} \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{n_y} & \mathbf{0} \\ \mathbf{0} & \sigma_u^2 \mathbf{I}_{n_u} \end{bmatrix}, \end{equation}
-    where $\sigma_{\bm{\epsilon}}^2$ and $\sigma_u^2$ are the variances of the output and the input, respectively. The normal equation (\ref{eqn:norm}) implies that 
-    \begin{equation} \tag{4.31} \left( \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \right)  \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) \neq \mathbf{0}, \end{equation}
-    and the estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is not consistent.
-    Instead, a bias-compensating least squares estimation can be obtained with  
-    \begin{equation} \tag{4.32} \widehat{\bm{\theta}}_{\mathrm{BCLS}} = \left( \mathbf{R}_{\widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\eta}} \right)^{-1}  \left(  \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{r}_{\bm{\eta} \bm{\epsilon}} \right) .
-\end{equation}
-
-    To formulate the generalized instrumental variable estimator, a generalized instrumental vector $\mathbf{z}(t)$ is built from delayed samples of the input and the output. 
-    The dimension of the generalized instrumental vector should be $n_{\mathbf{z}} \geq \mathrm{dim} ( \vartheta ) = n_u + n_y + 2$, since the total parameter vector $\vartheta = \left( \theta^\top, \rho^\top \right)^\top$ considers also the variances of the measurement noise $\rho = \left( \sigma_{\bm{\epsilon}}^2, \sigma_u^2 \right)^\top$.
-    Since the generalized instrumental vector $\mathbf{z}(t)$ and the equation error $\widetilde{y}(t) - \widetilde{\bm{\phi}}(t) \bm{\theta}$ must be correlated, we can express the sistem of equation as
-    \begin{equation} \tag{4.33}\left( \mathbf{R}_{\widetilde{\mathbf{z}} \widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \right) \bm{\theta} = \left(  \mathbf{r}_{\widetilde{\mathbf{z}} \widetilde{y}} - \mathbf{r}_{\bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) \right) , \label{eqn:GIVE} \end{equation}
-    where $\widetilde{\mathbf{z}} = \mathbf{z} + \bm{\zeta}$.
-
-    \subsubsection*{4.1.4.2 Recursive generalized instrumental variables method}
-    If we define
-    \begin{equation} \tag{4.34} \begin{aligned} \mathbf{h} \left( \bm{\rho}, \bm{\theta} \right)  &=  \mathbf{r}_{ \bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \bm{\theta}  =  \mathbf{J} \left( \bm{\theta} \right) \bm{\rho}, \\
-    \mathbf{R} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{\bm{\phi}}^\top \left( \tau \right) , \\     
-    \mathbf{q} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{y} \left( \tau \right) \\     
-    \mathbf{b} \left( t \right)  &=  \mathbf{q} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right)
-    , \end{aligned} \end{equation}
-    then the equation (\ref{eqn:GIVE}) is equivalent to
-    \begin{equation} \tag{4.35} \mathbf{R} \left( t \right) \bm{\theta} = \mathbf{b} \left( t \right) . \end{equation}
-    This representation allows a recursive form
-    \begin{equation} \tag{4.36} \begin{aligned} 
-    \mathbf{R} \left( t+1 \right)  &= \mathbf{R} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{\bm{\phi}}^\top \left( t+1 \right) - \mathbf{R} \left( t \right) \right), \\
-    \mathbf{b} \left( t+1 \right)  &= \mathbf{b} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{b} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right) \right), \\
-    \widehat{\bm{\theta}} \left( t+1 \right)  &=  \left( \mathbf{R}^\top \left( t+1 \right) \mathbf{R} \left( t+1 \right) \right)^{-1} \mathbf{R}^\top \left( t+1 \right) \mathbf{b} \left( t+1 \right) \\
-    \mathbf{q} \left( t+1 \right)  &= \mathbf{q} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{q} \left( t \right) \right) \\
-    \widehat{\bm{\rho}} \left( t+1 \right)  &=  \mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right) \left( \mathbf{q} \left( t \right) - \mathbf{R} \left( t \right) \widehat{\bm{\theta}} \left( t \right) \right) , \end{aligned} \end{equation}
-    where $\mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right)$ is the pseudo-inverse of $\mathbf{J} \left( \bm{\theta} \right)$.
-
-    The complexity of this recursive generalized instrumental variables estimator is ${O}(n^3)$, that is larger than the ${O}(n^2)$ of the recursive least squares.
-    Another drawback of the recursive generalized IV estimator is that it computes a pseudo-inverse matrix in each iteration step.
-    If we would need to formulate the data-driven step input estimation problem in the generalized IV framework, it would require to study the formulation in detail because currently the IV framework is based on the assumption that both measurements of the input and output are available, but for the step input estimation problem only the output is available.
-    Therefore, it is not straightforward to implement the recursive generalized IV estimator to estimate the step input, and this can lead to future research.
-    \color{black}
-
-    \end{itemize}
-    \item In the document, you always talk about recursive implementations of least squares solution but, in the end, you never give access to any recursive algorithm description (as you do for what you call the ML method).
-Why?
-Do it as well for the recursive weighted least squares solution.
-
-    {\bfseries The following recursive least-squares algorithm description has been added to Section 3.2:}
-
-    \color{blue} 
-    The problem (3.17) admits a least-squares (LS) solution with normal equations given by
-    \begin{equation} \tag{3.19} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}}.   \end{equation} % \label{eqn:neq_seiv} 
-    The closed form of the LS solution can be expressed as
-    \begin{equation} \tag{3.20} \widehat{\bm{\theta}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
-    where $\widetilde{\mathbf{K}}^\dagger$ is the pseudo-inverse matrix of $\widetilde{\mathbf{K}}$.
-    As the number of samples $N$ increases, the number of rows of the matrix $\widetilde{\mathbf{K}}$ grows, making the pseudo-inverse computation inefficient since it requires larger flops, memory and time.
-
-    For metrology applications, it is desired to have a fast solution that can be obtained in real-time. 
-    The recursive algorithm (RLS) is a least-squares alternative to implement the step input estimation method in real-time.
-    The RLS recursively updates the solution of the system of equations, for each newly acquired sensor response sample, considering the previous value of the estimation, instead of performing a matrix inversion.
-    An expression of the RLS equations is given in \cite{Kailath00book} as
-    \begin{equation} \tag{3.21} \begin{aligned} \widehat{\bm{\theta}}(t) &= \widehat{\bm{\theta}}(t-1) + \bm{\kappa}_{t} \left( \widetilde{y}(t) - \widetilde{k}_{t} \widehat{\bm{\theta}}(t-1) \right) , \\  \bm{\kappa}_{t} &= \bm{\Psi}(t-1) \widetilde{k}_{t}^\top / \left( 1 + \widetilde{k}_{t} \bm{\Psi}(t-1) \widetilde{k}_{t}^\top  \right) \\ \bm{\Psi}(t) &= \left( \mathbf{I} - \bm{\kappa}_{t} \widetilde{k}_{t} \right) \bm{\Psi}(t-1), \label{eqn:RLS} \end{aligned} \end{equation}
-    where $\widetilde{k}_{t}$ represents the row of $\widetilde{K}$ that corresponds to the $t\mathrm{-th}$ sample, $\bm{\kappa}_{t}$ is a gain scalar, $\Psi(t)$ is a covariance matrix, and the estimates and the covariance matrix are initialized respectively with $\widehat{\bm{\theta}}(1)= \widehat{\bm{\theta}}(0)$ and $\Psi(-1) = \Psi(0)$.
-    Since the RLS is a recursive implementation of the least-squares computation, the RLS solution is equivalent to the LS solution.
-    Moreover, the statistical analysis of the LS solution is valid also for the RLS solution, and is preferred since it is simpler to express the statistics using the closed form of the LS solution.
-    \color{black}
-
-    {\bfseries The following description for the exponentially recursive least-squares algorithm has been added to Section 6.2.1:}
-
-    \color{blue}
-    The exponentially weighted recursive least-squares (RLS) can solve recursively the estimation problem (3.20).
-    In \cite{Kailath00book} is described the exponentially weighted RLS algorithm, which is as follows
-
-    \begin{equation} \tag{6.4} \begin{aligned}  \widehat{\bm{\theta}}(k) &= \widehat{\bm{\theta}}(k-1) + \bm{\kappa}_{k} \left( \widetilde{y}(k) - \widetilde{\mathbf{k}}_{k} \widehat{\bm{\theta}}(k-1) \right) , \\  \bm{\kappa}_{k} &= \omega^{-1} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top / \left( 1 + \omega^{-1} \widetilde{\mathbf{k}}_{k} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top  \right) \\ \bm{\Psi}(k) &= \omega^{-1} \left( \mathbf{I} - \bm{\kappa}_{k} \widetilde{\mathbf{k}}_{k} \right) \bm{\Psi}(k-1), \label{eqn:RLS} \end{aligned} \end{equation}
-    for $k = 2n+1, 2n+2, \ldots$, where $\widetilde{\mathbf{k}}_{k}$ represents the row of $\widetilde{\mathbf{K}}$ that corresponds to the $k\mathrm{-th}$ sample, $\bm{\kappa}_{k}$ is a gain scalar, and $\Psi(k)$ is a covariance matrix.
-    The estimates and the covariance matrix are initialized using the first $n+1$ samples, i.e.,  
-    \begin{equation} \tag{6.5} \begin{aligned} \widehat{\bm{\theta}} (2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top  \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega}  \widetilde{\mathbf{y}}_{n+1}, \quad \mathrm{and} \\ \Psi(2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} , \end{aligned} \end{equation}
-    where $\widetilde{\mathbf{K}}_{n+1}$ is the matrix $\widetilde{\mathbf{K}}$ with the first $n+1$ rows, and $\widetilde{\mathbf{y}}_{n+1}$ is the vector $\widetilde{\mathbf{y}}$ with the first $n+1$ elements.
-    When $\omega=1$, the exponentially weighted RLS solution is identical to that of the RLS.
-    When $\omega<1$, the older residuals are weighted with lower values than the residuals of recent observations.
-    In this way, the solution of the minimization problem depends more on newer data, and less on older data.
-
-    \color{black}
-
-    \item Double check your equations page 12 and 13. By the way, can you explain where Eq. (3.6) comes from?
-
-    {\bfseries The equations in page 12 and 13 have been reviewed and typos corrected. The derivation of Eq. (3.6) has been improved with more explaination: }
-
-    \color{blue}
-    A third method can directly estimate the input step level from the step response, without identifying a sensor model.
-    To derive this method, we use the first difference operator, $\Delta = \sigma - 1$, where $\sigma$ is the shift operator, defined as $(\sigma^\tau y) (k) = y(k + \tau)$.
-    Applying the first difference operator to the system state-space representation (3.1) results in the autonomous system
-    \begin{equation} \tag{3.10} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta {y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
-    where $\Delta {u}(k) = \mathbf{0}$, for $k \geq 0$, and
-    $\Delta \mathbf{x}(0) = (\mathbf{A} - \mathbf{I}) \mathbf{x}(0) + \mathbf{B} {{u}}$.
-
-    If the response $\Delta {y}$  of this autonomous system is persistently exciting of order $L$, then the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ of $L+1$ block rows constructed from $\Delta {y}$ satisfies
-    \begin{equation} \tag{3.11} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) = \mathrm{rank} \left( \begin{bmatrix} \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \iddots & \iddots & \iddots \\ \Delta y(L+1) & \Delta y(L+2) & \cdots & \Delta y(L+n) \end{bmatrix} \right) \leq L,  \end{equation}
-    and the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ is a linear map to the natural responses of the augmented autonomous system (3.3) 
-    \begin{equation} \tag{3.12} \mathbf{y}_{\mathrm{natural}} = \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell}, \end{equation}
-    from any vector $\bm{\ell} \in \mathbb{R}^{n}$.
-    The total response of the system is the sum of the forced response, due to the applied input ${u}$ and the natural response
-    \begin{equation} \tag{3.13} \mathbf{y} = \mathbf{y}_{\mathrm{forced}}+ \mathbf{y}_{\mathrm{natural}} =G {u} + \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell} ,  \end{equation} 
-    that is equivalent to
-    \begin{equation} \tag{3.14} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(N) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ G & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ G & \Delta y(N-n) & \Delta y(N-n+1) & \cdots & \Delta y(N-1) \end{bmatrix} }_{\mathbf{K}} \underbrace{ \begin{bmatrix} {{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
-    where $N$ is the number of samples.
-    From this point of view, the vector $\bm{\ell}$ is a linear transformation of the system's initial conditions.  
-    \color{black}
-
-    \item  In Section 4.1.1 and 4.1.2, discuss the normality assumption. Please restructure Section 4.1.3 which is not clear for me.
-
-    {\bfseries The discussion of the normality assumption has been added to Section 4.1, since it is relevant for the two cases considered.
-Section 4.1.3 has been reestructured to improve its clarity.}
-
-    \item  Why do you always study the statistical properties of your least squares solution with high SNR only? What do you get for SNR = 10 dB for instance? 
-    \item Give the values of the matrices you use for simulating data sets so that the reader can test what you have carried out.
-
-    {\bfseries The first paragraphs of Section 4.2, and Fig. 4.1, were added to describe the conditions in which the simulations were conducted from the structured and unstructured EIV problems.}
-
-    \color{blue}
-    Two Monte Carlo (MC) simulation studies were conducted to test the obtained bias and variance formulas.
-    There are two sets of formulas.
-    One set was obtained for the unstructured and uncorrelated EIV problem, and, the other set, for the structured and correlated EIV problem. 
-    The problem solved by the data-driven step input estimation method (3.20) belongs to the second set.
-    The objective of the MC simulations is to compare the results of the bias and variance formulas in these two cases. 
-    The comparison is an evaluation of the effect that the structure and correlation have in the problem solution.
-    To make a fair comparison, the same exact data and the perturbation variance is used in both types of EIV problems.
-
-    The exact data is the system of linear equations \ref{eqn:ddsiemexd} built by the data-driven step input estimation method.
-    To build the system of equations, the processed signal ${y}$ is the exact response of a sensor to the input $0.1s$, where $s$ is the unit step response.  
-    The sensor is modeled as a linear time-invariant mass-spring-damper system of order $n = 2$, that is given by 
-    \begin{equation} \tag{4.37}
-    \mathbf{x}(k+1) = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{m+M} & \dfrac{-k_{\mathrm{d}}}{m+M} \end{bmatrix} \mathbf{x}(k) + \begin{bmatrix} 0\\ -g\end{bmatrix} u(k),  
-    \quad y(k) = \begin{bmatrix} -1 & 0 \end{bmatrix} \mathbf{x}(k), \label{eqn:msdst}
-    \end{equation}
-    where $m=0.015$ kg, $k_{\mathrm{s}}=500$ kg/m, $k_{\mathrm{d}}=0.5$ kg s/m, $M=0.1$ kg, and $g=9.81 \ \mathrm{m/s^2}$.   
-    The exact sensor response is shown in Figure 4.1.
-
-    In the simulations, the sample size considered is $N=200$, since there the transient regime of the sensor model is still evident. 
-    In the case of the structured EIV problem, this sample size satisfies the requirements of the data-driven step input estimation method. 
-    \color{black}
-
-    \item Please remove Fig.
-4.1 b from Section 4.2.1 because the top and bottom curves are obtained with different systems and simulation condition.
-Or explain why you can compare both.
-It could be interesting to test each time (i.e., in Section 4.2.1 and 4.2.2) the bias and covariance formulas obtained by considering the structured and unstructured EIV problems.
-
-    {\bfseries Fig. 4.1 b has been relocated in its corresponding section. Now it is Fig. 4.7.}
-
-    \item  When you generate different simulation examples, explain why the new ones are interesting or, more importantly, what they bring w.r.t.
-the former simulations results.
-This is the case for instance when I compare Section 4.2.2 and Section 5.1.
-    \item Except if you can convince me, do not use the name "maximum likelihood" when you do not introduce a likelihood you want to maximize.
-In Chapter 6, you introduce a nonlinear least squares algorithm, nothing else (see the attached document, Numerical Optimization, Chapter 10).
-
-    {\bfseries The method description in Section 6.2.2 has been modified to explain why the method is a maximum likelihood method. The modified paragraphs are as follows: }
-
-    \color{blue} 
-    Using a model of the sensor, the maximum-likelihood ML method simultaneously estimates the parameters of the applied affine input $u(t) = at + b$, and some of the sensor model parameters, using the observed sensor response $\widetilde{\mathbf{y}}$.
-    This method fits iteratively the response of the model $\widehat{\mathbf{y}}(\bm{\theta})$ to the sensor response by searching for the parameters $\bm{\theta}$ that reduce the error difference between both responses.
-    Algorithm 6.1 describes the formulation of an objective function $f = \mathbf{r}^\top \mathbf{r}$, where $\mathbf{r}$ is the residual $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$, that is minimized through the iterations.
-    The objective function is the sum of the squares of the differences between the samples of both responses.
-    Since the measurement noise is assumed Gaussian distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$, where the samples of this perturbation are independent and identically distributed, the minimization of the objective function $f$ maximizes the likelihood of fitting the actual sensor response with the model response.
-
-    The Jacobian matrix of the residual $\mathbf{J}\bm{\theta} = \partial \mathbf{r}(\bm{\theta}) / \partial \bm{\theta}$ can be analytically derived using the sensor model.
-    With the Jacobian matrix, the ML method searches the direction in which the residual decreases towards a local minimum.
-    Unfortunately, this ML method cannot guarantee the global minimum since it depends strongly on the initialization of the optimization parameters, as it is discussed in \cite{Nocedal06}.
-    Therefore, it is proposed to initialize the to-be-optimized parameters $a$ and $b$ using the results of the subspace method obtained after a convenient number of samples so that the affine input parameters are set close to the optimal values.
-    The initialization of the remaining model parameters under study can be done using the sensor model representation.
-    An illustrative example of the ML will be provided in a posterior subsection.  
-    \color{black}
-\end{itemize}
-
-\section*{Nikos Deligiannis}
-
-\begin{itemize}
-    \item Comments related to the structure of the thesis:
-    \begin{itemize}
-	\item In the introduction chapter it is best to describe a set of applications and settings that motivate your research and proposed solutions.
-This can make your thesis more accessible.
-Furthermore, you can add a subsection that describes in a high-level manner the remainder and structure of the thesis (describing briefly the contributions of each chapter).
-You can also mention your contributions there but in a less technical way since the problem statement only comes in the Chapter preliminaries.
-	\item In each main chapter (Chapter 4, Chapter 5, Chapter 6) you can add an introduction subsection, a state-of-the-art subsection and then elaborate your contributions and methods.
-
-    {\bfseries The introduction of the chapters now includes the state of the art to position the PhD work with respect to the literature.}
-
-	\item I believe that this structure will improve the understanding of the thesis and light up your contributions.
-
-	{\bfseries Indeed, this structure allows to visualize the thesis contributions.}
-
-	\end{itemize}
-	\item Further comments:
-	\begin{itemize}
-	\item Please discuss the complexity and scalability issues related to your approach in relation to the state of the art.
-
-    {\bfseries The following paragraph has been added in Section 3.2 to discuss the scalability of the data-driven step input estimation method.}
-
-	\color{blue}
-    The computational complexity of the RLS algorithm solution to the system of equations (3.14) is $O \left( \left( n+1 \right)^2 \right)$.
-    The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
-    From there on, the RLS algorithm updates the solution with linear complexity.
-    Therefore, even though a large number of samples $N$ makes the matrix in the system of equations (3.14) increase in the number of rows, the estimation of the solution does not increase in complexity.
-    The data-driven step input estimation method is then scalable for any sensor of order $n$, and can be executed in devices with limited computational resources, provided that the computation of the  $n+1 \times n+1$ inverse matrix is feasible.
-    \color{black}
-
-	\item Position your work with regard to the state of the art as well as regarding solutions based on low-rank matrix completion and compressed sensing.
-	\item I recommend to use the formulation “we propose” than “I propose”
-
-	{\bfseries The formulations similar to “I propose” were replaced by equivalents such as “we propose”. }
-
-	\item You might wish to add a list of notation in the beginning of the thesis.
-
-	{\bfseries A nomenclature list is provided before the summary of the document. }
-
-	\end{itemize}
-\end{itemize}
-
-\section*{Philippe Dreesen}
-
-\begin{itemize}
-    \item General remarks:
-    \begin{itemize}
-    \item  It would be good to include some layman’s examples that motivate the work of the thesis already in the introduction.
-In the current text, there is a nice example on page 52, almost at the end of the thesis.
-These kinds of examples should appear in the introduction to motivate and position the work.
-
-    {\bfseries The following paragraphs have been added to the Introduction: }
-
-    \color{blue}
-
-    Take for example the case of a thermometer.
-    To measure temperature, the thermometer is put in contact with the substance or object of interest.
-    Accordingly to the Newton's law of cooling, the temperature $y$ of the thermometer changes proportionally to the its difference with respect to the temperature $u$ of the matter under study:
-    \begin{equation} \tag{2.1}  \dfrac{dy}{dt} = k_T \left( u - y \right) \label{eqn:NewtonCooling} \end{equation}
-    where the proportionality constant $k_T$ is a parameter of the thermometer first order model.
-    The input $u$ is unknown and its estimation is the objective of the measurement.
-    It is customary to take the reading from the thermometer once the temperature stops changing.
-
-    Another popular example of a measurement is found in weighing. 
-    In most of modern scales, the object of interest if placed on top of a weighing sensor.
-    A simple model of the weighing sensor is the second order mass-spring-damper model shown in Figure \ref{fig:simple_msd_system}.
-    In this model, the unknown mass $u$ causes a displacement $y$ of the scale and the dynamics of the sensor are described by the differential equation:
-    \begin{equation} \tag{2.2} \dfrac{d}{dt} \left( u(t) \ \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = u(t) \ g \end{equation}
-    where $m$ is the own mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.
-
-    For illustration purposes, Figures \ref{fig:thermometer} and \ref{fig:scale} show the responses of the thermometer and the weighing sensors, respectively, after the application of a constant input $u$, which can be represented by a step input.
-
-    \renewcommand{\thefigure}{2.1}
-    \begin{figure}[htb!]
-    \centering
-
-    \begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
-    \tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
-    \tikzstyle{damper}=[thick,decoration={markings,  
-    mark connection node=dmp,
-    mark=at position 0.5 with 
-    {
-    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
-    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
-    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
-    }
-    }, decorate]
-    \tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]
-
-    \node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
-    \node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u(t)$};
-
-    \node (ground1) at (M.south) [ground,yshift=-1.5cm,xshift=-0.625cm,anchor=north] {};
-    \draw (ground1.north west) -- (ground1.north east);
-    \draw [spring] (ground1.north) -- ($(M.south east)!(ground1.north)!(M.south west)$);
-
-    \node (groundc) at (M.south) [ground,yshift=-1.5cm,anchor=north] {}; 
-    \draw (groundc.north west) -- (groundc.north east);
-
-    \node (ground2) at (M.south) [ground,yshift=-1.5cm,xshift=0.625cm,anchor=north] {};
-    \draw (ground2.north west) -- (ground2.north east);
-    \draw [damper] (ground2.north) -- ($(M.south east)!(ground2.north)!(M.south west)$);
-
-    \node[draw=none,fill=none] at (-0.9cm,-1cm) {$k_{\mathrm{s}}$};
-    \node[draw=none,fill=none] at (0.15cm,-1cm) {$k_{\mathrm{d}}$};
-    \node[draw=none,fill=none] at (2.0cm,1.0cm) {$y$};
-    \draw [-latex,thick]  ++(2.2cm,-1cm) -- +(0cm,2.25cm);
-
-    \draw [-latex,thick] (M.east) ++(0,0) -- +(1cm,0);
-    \draw [line width=0.25mm] (2.2cm,-1cm) -- (2.2cm,1cm);
-    \draw [line width=0.25mm] (2.1cm,-1cm) -- (2.3cm,-1cm);
-    \draw [line width=0.25mm] (2.1cm,1cm) -- (2.3cm,1cm);
-    \draw [line width=0.25mm] (2.1cm,-0.5cm) -- (2.3cm,-0.5cm);
-    \draw [line width=0.25mm] (2.1cm,0.5cm) -- (2.3cm,0.5cm);
-    \draw [line width=0.25mm] (2.15cm,-0.25cm) -- (2.25cm,-0.25cm);
-    \draw [line width=0.25mm] (2.15cm,0.25cm) -- (2.25cm,0.25cm);
-    \draw [line width=0.25mm] (2.15cm,-0.75cm) -- (2.25cm,-0.75cm);
-    \draw [line width=0.25mm] (2.15cm,0.75cm) -- (2.25cm,0.75cm);
-    \draw [line width=0.25mm] (2.1cm,0cm) -- (2.3cm,0cm);
-
-    \end{tikzpicture}
-
-    \caption{\label{fig:simple_msd_system}  \color{blue} A second order mass-spring-damper model represents the weighing sensor.
-The application of the mass $u$ causes a change in the position $y$ of the scale.
-\color{black} }
-    \end{figure}
-
-    \renewcommand{\thefigure}{2.2}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_2.pdf} 
-    \caption{ \label{fig:thermometer} 
-    \color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_T=0.5 s^{-1}$, caused by applying a step input of 67 $\circ \mathrm{C}$  when the thermometer was initially at 23 $\circ \mathrm{C}$.
-In this example, we should wait 10 s to have an accurate estimation of the input.
-\color{black} }
-    \end{figure}
-
-    \renewcommand{\thefigure}{2.3}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_3.pdf} 
-    \caption{ \label{fig:scale} 
-    \color{blue} This simulation illustrate the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium.
-The sensor parameters are $k_{\mathrm{d}} = 5 \mathrm{N s/m}$, and $k_{\mathrm{s}} = 10000 \mathrm{N/m}$.
-In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement.
-\color{black}  }
-    \end{figure}
-
-    \color{black}
-
-    \end{itemize}
-    \item Detailed remarks:
-
-    {\bfseries The detailed remarks in this list have been attended accordingly: }
-
-    \begin{itemize}
-    \item Citations that are used in-line should have brackets only around the year, not around the authors’ names.
-For instance on page 3, line -2 is written “reviewed in [Hack and ten Caten, 2012]”.
-This should be displayed as “reviewed in Hack and ten Caten [2012].
-In LaTeX this can typically be done by making the proper distinction between the commands \ citet and \ citep.
-The former is used for references in text, the latter for parenthetical references.
-    \item  On page 1 of the PDF (front of thesis), I believe also N. Deligiannis should be mentioned with the title “prof.”. 
-    \item Page 1 of thesis: typo on line 6: “and on the sensor initial conditions” => “sensor’s initial conditions”
-    \item Page 3, line -2: “uncertainty analysis are reviewed” => “is reviewed”
-    \item Page 3, line -1: remove comma before “that”
-    \item Page 3, line -2: fix the in-text reference (see above)
-    \item Page 4, lines 11-13: fix the in-text references (see above)
-    \item Page 5, line 5: Remove “The” at the beginning of the sentence (The Hankel…)
-    \item Page 6, line 10: fix the in-text references (see above)
-    \item Page 7, lines 10-11: the sentence on heart and breathing physiological monitoring sounds a bit awkward. Please reformulate.
-    \item Page 9, line -1: fix the in-line reference (see above)
-    \item Page 10, lines 1-3: in the enumeration of the types of sensors, remove the article ‘the’ before the sensor name.
-So: “like three-axis sensors [ref], radio-frequency intruder sensors [ref], and radar sensors [ref].
-    \item Page 10, equation 3.10: The symbol D was not bold in the equations above. Be consistent!
-    \item Page 10, caption of figure: replace “reverts” by “inverts”
-    \item Page 11, line 9, put “poles” between parentheses instead of between commas.
-    \item Page 12, line -6: “that in matrix form is” => “which in matrix form is”
-    \item Page 13, line -9: “in the practice” => “in practice”
-    \item Page 14, line -14: missing space between “problem” and “(3.11)”
-    \item Page 16, halfway page: the sentence “Applying a second-order…” is not a proper sentence (perhaps fix this by removing “that”).
-    \item Page 16 introduces seemingly some new notation, e.g., b(.), mu(.) and C(.).
-While their meaning is clear from the context, it is desirable to give proper definitions when introducing new mathematical tools.
-    \item Page 17, line -1: fix in-text references (see above)
-    \item Page 20 (and beyond): The abbreviation CRLB is introduced, but sometimes CRB is used instead.
-    \item Page 21, line -10: fix in-text reference (See above)
-    \item Page 35, second sentence “The sensor is a dynamic…” is not a proper sentence.
-    \item Page 50: Try to reorganize this section into two or more paragraphs.
-    \item Page 59, line 2: fix in-text reference (See above)
-    \item Page 74, line 11: “The learning obtained from…” this is an awkward sentence; please reformulate
-    \item Page 81, all three journal publication references do not have (or have wrong?) page numbers.
-    \end{itemize}
-\end{itemize}
-
-\bibliography{../Gus-thesis/Gus-thesis.bib}
+\bibliography{../Gus-thesis.bib}
 

### 2020-09-09 internal defense feedback/Answers-jury-members-report.tex
One before the very last

@@
 	\item  Discuss more clearly the aspects of scalability of the methods and their limitations in terms of model order, size of data sets (e.g., related to the inversion of LS matrices).
+
+	{\bfseries In Section 3.2 was added a discussion on the scalability of the data-driven step input estimation method.}
+
 	\item  The equations on pages 12 and 13 seemingly contain errors and should be corrected.
@@
 Elaborate on SNR levels; I’m also very interested in scalability, so how big the matrices can be.
+
+	{\bfseries A numerical sensitivity analysis was conducted on both subspace and maximum-likelihood methods for solving the affine input estimation problem.
+The sensitivity analysis showed that the subspace method can handle uncertainty on the generating system parameters up to 5\% without afecting deeply the performance of the method.
+Moreover, the mean-squared error of the estimation computed by the subspace method is similar in up to one order of magnitude in the SNR interval from 20 to 60 dB, being $10^{-1} \ \mathrm{g}^2$ its maximum value at 40 dB.
+The levels of measurement noise in practical applications are found in this SNR interval.
+}
+
 	\item  Concerning the scalability to data: eqn 3.6: how scalable is the approach in all aspects (order, data, linked to inversion for LS matrix); can you deal with hundreds, thousands of data?
@@
 A more detailed discussion is desirable.
+
+	{\bfseries In Section 3.2 was added a discussion on the scalability of the data-driven step input estimation method.
+Since it uses recursive least squares to update the estimation, its complexity is linear in the number of parameters $n+1$.
+This means that for small number of samples, $N=100$, or large number os samples, $N=10^4$, the complexity per recursion does not change.
+
+	The computation of the bias and variance of the input estimation, as it is proposed in Chapter 4, has a complexity $O(n^3N)$.
+This complexity prevents the implementation in real time of these statistical moments estimation.
+A future research work can be devoted to find efficient estimation of these input estimation moments, to be obtained right after the input estimation itself.
+}
+
 	\item  On slide 25 (noisy graphs) it would be good to see a statistical analysis (more experiments than one).
@@
     \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_7.pdf} 
+    \includegraphics[width=0.6\columnwidth]{../ChapterRampInput/fig/Fig_7.pdf} 
     \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method.
@@
     \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_8.pdf} 
+    \includegraphics[width=0.6\columnwidth]{../ChapterRampInput/fig/Fig_8.pdf} 
     \caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian.
@@
 This can be investigated further.
+
+	{\bfseries The investigation of the error bounds for the second order Taylor expansion can be conducted using the continuous Newton method. This is proposed as future research work in Chapter 7.}
 
@@
 	The continuous Newton method is a sequence of damped Newton methods.
-To implement the continuous Newton method, select $T>0$ and perfom $m$ runs of the damped Newton method with $\delta_k = T/m$, for $k=1,2,\ldots, m$.
+To implement the continuous Newton method, select $0<T<1$ and perfom $m$ runs of the damped Newton method with $\delta_k = T/m$, for $k=1,2,\ldots, m$.
 	The result $z_{k+1}$ is the zero estimation $x_m$, and if the sequence of estimates $x_1, x_2, \ldots$ converges, then it converges to the result of the continuous Newton method.
@@
 Using the pseudo-inverse of $\widetilde{\mathbf{K}}$, the metric $\left\Vert  \widetilde{\mathbf{K}}^{\dagger} \widetilde{\mathbf{y}} \right\Vert \leq r$ indicates that the estimate $\hat{\bm{\theta}}$ is found inside a ball of radius $r$.
-    Thus,  
+    The questions that remain open are how large can $r$ be?, and how is the radius $r$ related to the perturbation noise variance? 
 	\color{black}
@@
 Something to investigate as well...
+
+	{\bfseries To get starting points to the EIV model, that are more accurate than LS, we would need to incorporate more prior knowledge.
+In the formulation of the EIV problem we have made assumptions of the model class for the sensor and the Gaussian white noise of given variance for the additive measurement noise.
+The LS is an heuristic algorithm that was chosen by its simplicity for solving online the estimation problem by means of recursion.
+However, the LS solution is biased.
+The optimal solution could be obtained with a maximum likelihood algorithm provided we have a large amount of samples, but we need to provide approximations starting from small number of samples.
+Moreover, the maximum likelihood solution of the EIV problem needs structured total least squares, which does not have a closed form that could be extended with a Taylor series, or structured total maximum likelihood, which do not support the correlation between the regression matrix and the regressor in the EIV problem.
+The recursive versions of these two structured approaches is also more computational expensive than RLS.
+Therefore, it is required to continue the research towards efficient solution algorithms.
+}
+
 	\item  About robustness: is it an issue in practical problems with heavy tailed noise, outliers,...
@@
 Might be interesting to ponder this in future.
+
+	{\bfseries There are some estimators that can deal with heavy tailed noise.
+In the paper \cite{Balkema18} there  are described different estimators for a linear regression \linebreak $Y_i = b+a X_i + Y_i^{*}, \quad i = 1,\ldots,n$, where both the explanatory variables $X_i$ and the errors $Y_i^{*}$ have heavy tails.
+The $X_i$ are samples from a Pareto distribution on $(1, \infty)$ arranged in decreasing order $X_n<\cdots<X_2<X_1$, as if they were generated from the order statistics $U_1<\cdots<U_n$ of a sample of uniform variables in $(0,1)$ by setting $X_i = 1/U_i^\zeta$.
+The parameter $\zeta$ is the tail index of the Pareto distribution.
+For the variables $Y_i^*$, the tail index is $\eta$.
+	Heavy tails are indicated by larger tail indexes.
+For $\eta < 1/2$, the variable $Y^*$ has finite variance and least squares is the suggested estimator.
+For $\zeta < 1/2$, the variable $X$ has finite variance and the least absolute deviation is the suggested estimator, otherwise the Theil-Sen estimator is preferable.
+The paper presents empirical results comparing the estimation bias and standard deviation for different tail indexes.
+
+	Regarding outliers, in \cite{Maronna06} is described an approach of using least squares to identify the outliers, followed by a recalculation without considering the outliers.
+In this sense, one could also consider that the removed outliers are missing data, and use low-rank approximation.
+The classical approach is using a maximum likelihood type estimator (M-estimators) as it is described in \cite{Zhang97} and \cite{Huber81}.
+The M-estimators aim to mitigate the outliers impact by using a function of the residual $\rho(r_i)$ instead of the the squared residuals $r_i^2$ in the minimization $\text{min} \ \sum{\rho(r_i)}$.
+The function $\rho$ is symmetric, positive definite, with only one minimum at 0, and increasing slowly than the the square.
+
+	Selecting another estimator to solve the EIV problem constructed by the data-driven step input estimation problem to deal with heavy tailed noise, or with outliers, might have a big impact on the implementation.
+The implementation may be computationally complex and not suitable for online applications.
+
+	}
+
 	\item  On deep learning: could this be an alternative?
+
+	{\bfseries Deep learning could be used but it would be an overkill methodology for linear estimation.
+However, for metrology applications were the input is the to-be-estimated signal, implementing deep learning could be a challenge.
+The neural networks are suitable for non-linear model based applications were the training require the design and application of input stimuli to correlate with the corresponding outputs of the system.
+	In metrology, it is not possible to train deep learning methodologies with real input data, because the input data is unknown and cannot be observed. 
+	By applying neural networks trained by simulation would estimate inaccurately the desired inputs.  }
+
 \end{itemize}
@@
 It could be interesting to compare with such IV solutions.
+
+	{\bfseries In Section 2.2.1 the follwing paragraphs have been modified to include references of instrumental variable solutions to the errors-in-variables problems.}
+
+	\color{blue}
+	Instead of using a total least-squares (TLS), or an intrumental variables, solution of the structured EIV problem, the step input estimation method uses the recursive least-squares solution (RLS). 
+    In the review conducted by \cite{Markovsky07overview} it is shown that the TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
+    the TLS solution is equivalent to the maximum likelihood (ML) solution when the disturbances of the EIV problems are i.i.d. normally distributed. 
+    The classical TLS solution is obtained using singular value decomposition of the augmented matrix built from the regression matrix and the regressor as additional columns.
+    However, the studies presented in \cite{VanHuffel07TLSeditorial} demonstrate that the structured TLS solutions cannot be generalized since each structure in an EIV problem requires a specific treatment.
+    Moreover, the ML estimator of structured EIV problems leads to non-convex optimization, and the global optimum is not guaranteed \cite{Beck10}.
+    On the other hand, the instrumental variable (IV) solution of EIV problems use an additional variable.
+    The additional variable is an instrument with the purpose of improving the consistency of the estimate.
+    To do this, the construction of the instrumental variable has to be done in such a way that it is uncorrelated with the perturbations and correlated with the regressor, as it is described in \cite{Soderstrom18}.
+    Unfortunately, the computational complexity of the TLS, ML, and IV solutions inhibits their real-time implementation.
+
+    The solution of the step input estimation problem is preferred to be online, since its applications are in the field of metrology.
+    In the literature there are references for online implementations of TLS, ML or IV methods, which are are mainly recursive versions of the solution methods.
+    These recursive versions are developed to solve EIV problems for system identification and fault detection.
+    We have, for example, generalized recursive TLS solutions introduced in \cite{Rhode14recursive} and \cite{Rhode14recursivecov}, and recursive IV solutions developed in \cite{Djouambi12}, \cite{Shang16} and \cite{Gil15}.
+    These methods cannot be used as off-the-shelf solutions to the step input estimation EIV problem because they require to be adapted to the specific requirements of the problem.
+    One is that the sensor output is the only measured signal, since the input is to-be estimated, and the solution should be simpler, preferable with a linear complexity $O(n)$, instead of the cubic $O(n^3)$ or cuadratic $O(n^2)$ complexities of the mentioned recursive methods, where $n$ is the sensor order.
+    The proposal of the data-driven step input estimation method includes the recursive least-squares (RLS) solution.
+    Compared to recursive TLS or recursive IV solutions, RLS is suboptimal but more simple solution to the structured EIV problem with lower computational complexity,  suitable for real-time implementation.
+	\color{black}
+
 	\item  In Chapter 3: at beginning you say to use linear systems theory to model a sensor. Can you discuss this assumption: is it true in practice? Can you guarantee it behaves like an LTI system?
+
+	{\bfseries The first paragraph of Chapter 3 has been added to explain this concept:  }
+
+	\color{blue}
+    The dynamic systems in the real world are non linear. 
+    Under certain intervals of operation, the linear approximations can explain the dynamics of the systems. 
+    The linear and time invariant systems have been studied in depth and the linear theory is a solid collection of tools to analyze and explain the behaviour of dynamic systems in a wide range of applications.
+    On the other hand, the study of non-linear systems is more complex and strongly depends on the type of non-linearity that each particular systems exhibit. 
+    When the non-linear modeling does not apport a big amount of benefits into the application study, such as significative accuracy improvement or supporting wide ranges of operation, then is better to resort to the utilization of LTI representations.
+    A sensor is a dynamic system whose operation range is preferred to be linear, to facilitate the repeatability and reproducibility of the measurements.
+    This means that every time the sensor must deliver identical responses for the same input excitations.   
+    \color{black}
+
 	\item  Equation 3.1: an LTI system, described with SS representation.
@@
 Is it still white in this case?
+
+	{\bfseries In Chapter 3 was added the following paragraph to explain the impact of white process noise in the response of al LTI system. }
+
+	\color{blue} 
+    The process noise is the representation of the uncertainty in the system state due to unmodeled perturbations that affect the state evolution. 
+    The process noise is modeled as an additional input to the system excitation. 
+    For a linear time invariant (LTI) system, the response to a sum of two input excitations is the sum of the two responses that correspond to the individual inputs.
+    It is demonstrated in \cite{Gubner06} and \cite{Smith11} that the response of an LTI system to a white noise input is colored noise, since the power spectral density of the response is not constant, but instead proportional to the squared magnitude of the system impulse response.
+    Nevertheless, in the bandwidth operation of a properly designed measurement system, processing white noise results in approximately white noise output. 
+    Moreover, in the present work the measurement noise exists only in the system response signal, because the input is not direcly observed, and the sensor operation is not subject to external perturbation.
+    Under these conditions, the mean and variance of the measurement noise $\epsilon$ describe implicitly the process noise entering the system. 
+    \color{black}
+
 	\item  On page 12 standard approach top of page equation: there are some mistakes in the equations!
@@
 	The normality assumption is necessary to provide more prior information into the method formulation.
-    The second order Taylor series expansion (4.4) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
+    The second order Taylor series expansion (4.6) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
     Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
-    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance\color{black}.
+    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance.
+    \color{black}
 
@@
 You use only one realization also for matrices A, B, C, D in the text, so it can be added explicitly.
+
+	{\bfseries Section 4.2 has been improved to explain how the exact data and the perturbation noise are generated.
+This modification allows to reproduce the simulations conducted and understand the outcome of them.
+}
+
 	\item   in Fig 4.1 the bottom figure also discusses the structured case. Why do you talk about structured results in a section about the unstructured case?
@@
 
-	{\bfseries \color{red} Interesting question, I want to try it this evening \color{black}}
+	{\bfseries The numerical study of the expressions that calculate the step input estimation bias and variance has been appended to include the results of using unstructured perturbation in the expressions for the structured case, and also structured perturbation in the expressions for the unstructured case.
+It is expected that the performance is very far from that obtained when using the expressions in concordance with the type data they were derived for.
+A discussion on this observations is included in Section 4.2.3.}
 
@@
 Can you use something in your case to test for a good order of system?
+
+	{\bfseries If we evaluate the step input estimation bias and variance obtained by the data-driven method for different model orders, we expect the results follow the bias-variance tradeoff.
+It is commonly accepted that for decreasing orders the bias is large since not all dynamics are captured, and that for increasing orders the bias is small and the variance is large because the system dynamics are captured and also the measurement noise is being modeled.
+However, after observing the results in the experimental validation of the method, we get better performance by overfitting.
+This could be valid only for some estimators.
+	For example, in the research work of \cite{Neal19} it is described a neural network that does not follow the bias variance tradeoff, contradicting \cite{Geman92} and \cite{Manning08}, who state that an overparameterized neural network reduces the bias by interpolating the data with a significant variance contribution to the mean-squared error.
+
+	Some approaches described in the literature to avoid overfitting include crossvalidation (CV), minimum description length (MDL) and the Akaike information criterion (AIC).
+	These methodologies are described in books such as \cite{Hastie09}  and \cite{James13}. 
+	In the paper \cite{Arlot10}, CV and MDL are compared trying to answer whether penalized criteria in classical AIC should be preferred to CV for a given model selection problem. 
+	On one hand, provided the data is i.i.d., CV gives good model selection in almost any framework. 
+	On the other hand, CV can be less accurate, even with larger computational cost, than procedures designed to be optimal in a specific framework, such as AIC.
+
+	To describe the classical methods, the following paragraph has been added in subsection 5.2.
+	}
+
+	\color{blue}
+
+	There exist methods for estimating the performance of different models that are useful to choose among the models.
+	The model selection methods can be analytical, such as the Akaike information criterion (AIC) introduced in the paper of \cite{Akaike74}, the minimum description length (MDL) iniciated by the work of \cite{Rissanen78}, or numerical such as crossvalidation.
+	The AIC is defined for a class of models fit by maximum likelihood. 
+	For the linear regression problem with Gaussian perturbation (3.20), the AIC is given by
+	\begin{equation} \tag{5.9} \mathrm{AIC} = 2 \left( n+1 \right) - \mathrm{ln} \left( \widehat{L} \right) \label{eqn:AIC} \end{equation}
+	where $n+1$ is the number of estimated parameters, and $\widehat{L}$ is the maximum value of the likelihood function of the model.
+	The AIC rewards the goodness of fit by considering the maximum likelihood value, and adds a penalty term that increases the cost with respect to the number of parameters to prevent overfitting.
+	In the opinion of \cite{McElreath20}, the AIC is reliable when the priors overwhelmed by the likelihood, the posterior distribution is approximately multivariate Gaussian, and the sample size is much larger than the number of parameters.
+	Using this criterion for model selection is simple, \cite{Hastie09} suggest just picking the model that provides the smallest AIC value.
+
+	The MDL approach is motivated from an optimal coding viewpoint, aiming to choose the most parsimonious model, that, according to \cite{Hastie09}, is the shortest model that can encode the data for transmision of messages.
+	If we have a model with parameters $\theta$, and data $(K, y)$, with the conditional probability of the outputs $\mathrm{P}(y | \theta, K)$, and assuming the receiver knows the inputs and we transmit the outputs, then the message length required to transmit the outputs is
+    \begin{equation} \tag{5.10} \mathrm{length} = - \mathrm{ln} \left( \mathrm{P}( \theta ) \right) - \mathrm{ln} \left( \mathrm{P}(y | \theta, K) \right) . \label{eqn:MDL} \end{equation}
+    The length is the sum of the code length required for transmitting the model parameters with the code length required for transmitting the discrepancy between the model and the actual target values.
+    The recommended model is the one that minimizes (\ref{eqn:MDL}).
+	\cite{McElreath20} arguments that model fitting is a form of data compression, where parameters explain the relationships among the data. 
+	The parameters compress the data into a simpler form, with some loss of information about the sample, and 
+	can be used to generate new data by decompression.
+	Learning about the data requires using a simpler model that achieves some compression, but not too much.
+    \color{black}
+
 	\item   Do you know of cross-validation: split data into two parts or several parts; some for estimation, some for validation?
@@
 Have you tried this?
+
+	{\bfseries The crossvalidation method has not been applied on the experimental data.
+This data contains the step responses of a weighing system.
+The step responses are acquired with a sampling frequency of 4 kHz.
+One difficulty for applying crossvalidation is the division of the data.
+If we select the first 2/3 of the step response samples for order estimation and the last 1/3 for validation, we may leave the data with the diminished transient efcects in the validation set.
+Another approach could be taking every second sample for validation, dividing the measured responses in two by downsampling.
+However, this approach requires filtering against possible aliasing because the frequency characteristics of the response signal change.
+A third way that could be tried is dividing the 100 measured responses in two sets, devoting 80 responses for order estimation and 20 responses for validation.
+}
+
 	\item   In Section 6.2.2.
@@
 Gaussian, ...) it leads to LS.
-Don't' call it ML if you dont give distributions on ...
+Don't call it ML if you dont give distributions on ...
 In this case it is better to call it a nonlinear minimization procedure.
@@
 I am a bit surprised you require 30 sec for 2 parameters.
+
+	{\bfseries The large time consumption in the local optimization iterations is due to the different functions that take place during the minimization of the error function.
+The iterations are performed in a receding horizon scheme, starting from a small amount of samples (50) and repeating the error minimization for each new subset of samples acquired (5), until covering all samples (1000).
+In each iteration, the response of a dynamic system is simulated several times, for the different tried optimization variables.
+This simulated response is compared to the actual measured response, aiming to reduce the error between these two responses.
+Every time that a sensor response is simulated for a particular vector of optimization variables being tested, the analytic Jacobian is also evaluated.
+The evaluation of the Jacobian matrix requires the  simulation of one additional dynamic system per to-be-optimized parameter.
+Considering that the simulations of all these responses take considerable amount of time, the elapsed time that the ML method takes is 25 s in average.
+Then, there is a need for efficient methods that could run in real time to optimize online the input model parameters and the sensor model parameters of interest.
+}
+
 \end{itemize}
@@
 Do you solve them using pseudo-inverse or do you do gradient descent?
+
+	{\bfseries The closed form solutions allow us to describe the operations in a more elegant way.
+The closed form solutions implies batch processing of the operations.
+The formulation of the problem also gives a hint of which solution to choose.
+For instance, an unconstrained linear minimization problem can be solved with least squares but constrained nonlinear minimization problems most likely require optimization.
+There are different methods to solve least squares problems.
+The pseudo-inverse approach is the first option that comes to my mind, but modern numerical software, such as Matlab and julia, have toolboxes that solve efficiently the least squares problems depending on the regression matrix and regressor data.
+Regarding the maximum-likelihood method of Chapter 6, the optimization method uses gradient descend to find the direction towards the local minimum.
+Other minimization problems in low-rank approximation can be solved using variable-projection optimization methods.
+}
+
 	\item  Today there is an overwhelming amount of data sensed and communicated; How would you apply your method in a scenario where dimensionality increases significantly?
 This relates to the questions of previous jury members; what is the implication of using numerical versus closed form solutions?
+
+	{\bfseries The increase in dimensionality is explained by the current trend in the industry of installing more sensors.
+More signals are used to monitor the production lines.
+The transference of data is relevant for decision making.
+The data-driven input estimation method can be used to estimate recursively the input applied to the sensors, and to tract the evolution of inputs.
+In this way, the estimation method facilitates the controllers by providing online estimations of the inputs.
+This improves the performance of closed form solutions, which need more computational resources and time more time for the batch processing of data.}
+
 	\item  What if the $K^\top K$ matrix is not invertible?
@@
 can you use them, what would be the impact of that?
+
+	{\bfseries A possible lack of invertibility of the matrix $K^\top K$ may be due to missing data.
+A different methodology to low-rank approximation is compressed sensing, where the missing data is interpreted as sparsity.
+Under the assumption that a signals is sparse because is subject to many near zero coefficients in a given basis, then changing to another basis and taking a weighted linear linear combination of the available samples would lead to the estimation of the missing data values.
+These idea is described as future work in the final chapter of the thesis  }
+
+	\color{blue}
+
+	\subsection*{Compressed sensing methods for recovering missing data}
+
+	Compressed sensing is a methodology that is being developed quite fast and that can offer an alternative to perform measurements with missing data. 
+	The aim of compressed sensing is the recovery of the missing samples of a signal by computing a sparse coefficient vector, under the assumption that the signal belongs to a generating system of larger order.
+	The missing data estimation problem can be expressed as an overdetermined system of equations with a sparse regression matrix. 
+	The review presented in \cite{Eldar2012compressed} provides an introduction of the mathematical modeling of compressed sensing, and gives the conditions for sparse recovery. 
+	The recovery algorithms include convex optimization, greedy and combinatorial algorithms, to name a few.
+
+    According to \cite{Rani18}, compressed sensing is applied when we canot afford buying enough sensors like in non-visible wavelengths applications, when the measurements themselves are expensive as in neutron-scattering imaging, when measurements take a lot of time as in medical imaging, and when a constrain of the measurement is the power consumption.
+    A particular field of application for compressed sensing is the recovery of bioelectrical signals, as it is detailed in \cite{Craven15}.
+
+    Another good review of compressed sensing is found in \cite{Duarte11}, where it is explained that the  research community is splitted into a group that studies the problems in which the compressed sensing matrices exhibit structure due to the measurement system modeling, and another group which considers that the signal structured representations not only depend on sparsity but instead assume that the signals belong to continuous time and have infinite-dimensional representations.
+The solutions provided exploit these structures in the development of the algorithms.
+Some examples of these structure matrices are described in \cite{Kezhi15} including Gaussian, Toeplitz, Hadamard, and structurally random matrices.
+
+    It might be interesting to compare the solutions of the low-rank approximation methods with compressed sensing in terms of computational complexity and feasibility for online implementations. 
+    Therefore, it is proposed as a future research to develop a compressed sensing algorithm, that can estimate directly the input true value by processing the sensor response to the unknown input, and in presence of missing data.
+	\color{black}
+
 	\item  You mention that you improve complexity with your approach, but I missed a complexity analysis. Did you do this? Anything in processing time, flops,...?
+
+	{\bfseries A discussion on the scalability of the data-driven step input estimation method has been added in Section 3.2.
+The method uses recursive least squares to update the input estimation value, thus its complexity is linear in the number of parameters $n+1$, and independent from the number of samples $N$.
+
+	The computation of the bias and variance of the input estimation, which is described in Chapter 4, has a complexity $O(n^3N)$.
+This complexity prevents the online implementation of the estimation bias and variance.
+It is proposed as future research work to study efficient methods to estimate these statistical moments.
+The objective of this future work is to estimate the input and its mean squared error in real time.
+}
+
 	\item  You mentioned deep learning could be applied; could you show how?
@@
 Is this a limitation?
+
+	{\bfseries Deep learning could be used but it would be an overkill methodology for linear estimation.
+However, for metrology applications were the input is the to-be-estimated signal, implementing deep learning could be a challenge.
+The neural networks are suitable for non-linear model based applications were the training require the design and application of input stimuli to correlate with the corresponding outputs of the system.
+	In metrology, it is not possible to train deep learning methodologies with real input data, because the input data is unknown and cannot be observed. 
+	By applying neural networks trained by simulation would estimate inaccurately the desired inputs.  }
+
 	\item  You mention that you consider Kalman filters; have you seen approaches where they write Deep Learning in form of Kalman Filter; Could this be applicable?
+
+	{\bfseries The Kalman filter is mentioned in Chapter 3 to give context for the data-driven step input estimation method.
+The estimation of the input from its corresponding sensor step response in presence of noise can be estimated with a Kalman filter when we have a model of the sensor.
+
+	Regarding deep learning and Kalman filters, \cite{Schmidhuber15} provides a chronological review of shallow and deep learning methods considering the depth of the used chains of possible learnable, causal links between actions and effects.
+	Deep learning is about assigning credit across the computational stages of neural networks. 
+	Deep Kalman filters are filters that incorporate neural networks to improve the performance when the system under study in nonlinear, or to mitigate the model uncertainties. 
+	To mention a few deep Kalman filters, we have \cite{Hosseinyalamdary18} that fusions global positioning sensors and inertial sensors responses to model non-linear, time-variant, and highly correlated inertial sensors errors,
+	\cite{Rambach16} fusions camera images with inertial sensor measurements to estimate the camera pose, 
+	\cite{Zheng19} designs a Kalman filter from a bottom-up approach to forecast the short-term load of a houseconsidering the large uncertainty in the human interaction and weather conditions, 
+	\cite{Bai20} uses a neuron-based nonlinear autoregressive model to adapt the Kalman filter variables for a robot positions tracking, 
+	\cite{Farrell18} reconstructs the particle paths in the large hadron collider with a recurrent neural network that extrapolates, builds, and evaluates path candidates,
+	\cite{Shashua17} combines large state spaces with nonlinear Q-function approximations to compensate for the uncertainty in the neural network weight values, and
+	\cite{Krishnan15} proposes an algorithm for learning causal generative temporal models from high-dimensional noisy data, such as the evolution of a large number of patients suffering from a particular illness, to enhance Kalman filters under arbitrary complex transition dynamics.
+
+	The problem under study in this thesis is the linear estimation of a step input with unknown level given the sensor step response. 
+	The data-driven input estimation method is recursive and assumes that the model of the LTI sensor is not available.
+	The deep Kalman filters are suitable form non-linear applications were a model of the input is available, and the incorporated neural networks could estimate also the sensor model parameters.
+	To do this, the computational complexity of deep Kalman filters may be larger than that of the data-driven input estimation method.
+	}
+
 	\item  Recommendation: if you revise the thesis: maybe change "I propose" to "we propose".
@@
 Can you still use your methods?
+
+	{\bfseries If the sensor has a saturation for high input values, the data-driven step input estimation method still cannot provide an estimation of the input step level.
+To perform the estimation with missing data, it is necessary to complete the Hankel matrix in the estimation problem formulation.
+If the sensor response were exact data, it would be enough to rearrange the available data from the Hankel matrix and find the missing data.
+Nevertheless, this procedure would take considerable time.
+Moreover, under noise perturbation, efficient approximation algorithms are needed to perform the matrix completion in real time.
+The last paragraph in Chapter 3 locates the data-driven step input estimation method in perspective to the most powerful unfalsified model of the behavioral approach theory.
+}
+
+	\color{blue}
+    {\bfseries Note.} The formulation of the data-driven step input estimation method is based on the behavioral theory introduced in \cite{Willems86I}.
+    A behaviour of a dynamical system is any set of input-output trajectories that belong to the system. 
+    It is demonstrated in \cite{Willems05} that when an input-output trajectory is persistently exciting enough, a Hankel matrix constructed from this trajectory represents the most powerful unfalsified model.
+    This model is unfalsified because it describes exactly all possible trayectories of the system, and is the most powerful because it is the system representation with the minimum complexity.     
+    The rank of the Hankel matrix constructed from a trajectory is a metric used to find the order of persistency of the trajectory.
+    The rank-deficient Hankel matrix with the smaller number of block rows indicates the trajectory order of persistency. %
+    When the system is excited by a step input, the system is equivalent to an autonomous augmented system, and a model for it can be obtained from the Hankel matrix constructed only from the step response.
+    \color{black}
+
+	{\bfseries Also, in the future work section is described the need for efficient optimization algorithms that could solve the low-rank approximation problem allowing for a online estimation of the step input value.
+}
+
+	\color{blue}
+	\subsection*{The data-driven input estimation methods with missing data}
+
+	The implementation of the data-driven step input estimation method can be compromised when one or more step response samples are not properly acquired.
+    The missing data in the construction of the Hankel matrix may be due to sample transference errors, or due to sensor saturation problems.
+    The saturation values in the sensor response are not valid for the computation of the step input level because they violate the linearity assumption of the sensor.
+    The solution to the most powerful unfalsified model identification problem in presence of missing data was provided in \cite{Markovsky16}. 
+    The methods presented in this work are formulated as a matrix completion with the constraint that the to-be-completed matrix must be a rank deficient Hankel matrix.
+    The low-rank approximation of Hankel structured matrices with missing data has been studied in \cite{Markovsky13}, where the proposed solution is based on a variable-projection method that requires optimization.
+    The software for this method, including a link to obtain it, is described in \cite{Markovsky14}. 
+    This method is applicable in matrix completion and system identification problems and has inspired the development of the signal recovery method in for nuclear magnetic resonance spectroscopy described in \cite{Ying18}, and data compression and feature recognition in \cite{Hou17} based on sparse low-rank matrix approximation.
+
+    The solutions of the low-rank approximation with missing data problems are found by iterative optimization and, therefore, require large amount of computational power.
+    In order to implement data-driven input estimation methods with missing data, there is a need for efficient methods of small complexity $O(n)$ that solve online the optimization problem.
+    This is an open problem for future research. 
+    \color{black}
+
 	\item  Sensors are physical devices and deliver a continuous-time output. The models that you use are in discrete-time. Are there any risks that you miss information by not sampling fast enough?
@@
 Would it be possible to plug in a procedure that does structured matrix completion and approximation into your methods?
+
+    {\bfseries Unfortunately, there is still a need for efficient optimization algorithms that could solve the low-rank approximation problem allowing for a online estimation of the step input value.
+}
+
 \end{itemize}
@@
 
+    \color{blue}
+    The normality assumption of the perturbation noise is necessary to provide more prior information into the method formulation.
+    The second order Taylor series expansion (\ref{eqn:xhatexp}) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
+    Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
+    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance.
+    \color{black}
+
 	\item  Returning to a question by Philippe Dreesen on missing data.
@@
 It works in low-rank approximations; not if the matrix is of full rank.
+
+	{\bfseries The last paragraph in Chapter 3 was added to describe the position the data-driven step input estimation method with respect to the most powerful unfalsified model of the behavioral approach theory.
+In the future work section, it was added a description of the need for efficient optimization algorithms that could solve the low-rank approximation problem enabling the online estimation of the step input value.
+}
 \end{itemize}
@@
 
-	{\bfseries The last paragraphs of Section 3.2 have been included to discuss the scalability of the data-driven step input estimation method.  }
+	{\bfseries One paragraph at the end of Section 3.2 has been included -to discuss the scalability of the data-driven step input estimation method.  }
+
+	\color{blue}
+	The computational complexity of the RLS algorithm solution to the system of equations (\ref{eqn:ddsiemexd}) is $O \left( \left( n+1 \right)^2 \right)$.
+    The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
+    From there on, the RLS algorithm updates the solution with linear complexity.
+    Therefore, even though a large number of samples $N$ makes the matrix in the system of equations (\ref{eqn:ddsiemexd}) increase in the number of rows, the estimation of the solution does not increase in complexity.
+    The data-driven step input estimation method is then scalable for any sensor of order $n$, and can be executed in devices with limited computational resources, provided that the computation of the  $(n+1) \times (n+1)$ inverse matrix is feasible.
+    \color{black}
 
@@
 What do you mean by fast?
+
+        {\bfseries The Introduction has been appended with illustrative examples in which the measurements would benefit from the direct estimation of the input.
+The sensors in this list are those with a low bandwidth due to its construction and principle of operation.
+The transient stage of these sensors is long, from a practical point of view, because it takes relatively long time for the transient effects to decrease below the noise floor.
+From that moment on, the input value can be estimated easily since it is proportional to the sensor response in steady regime.
+The estimation of the input is classically obtained through filtering the sensor response with a dynamic system designed to perform a deconvolution.
+The response of the aditional system approximates the input value, but unfortunately there is also a transient response with a duration that is expected smaller than that of the actual sensor.
+Then, there is a trade-off between the estimation accuracy and the estimation time.
+The compensation time can be still improved further with signal processing methods that are independent of sensor models.
+These data-driven methods can be conceived as adaptive methods with shorter transient time that classical compensation systems.
+The data-driven step input estimation method formulates the estimation problem as an errors-in-variables problem, and proposes a solution with recursive least squares.
+This implies that we have a fastar response with an asymptotic bias.
+Future research can be devoted toi find a fast and unbiased online solution to the errors in variables system.
+}
+
         \item  the literature on EIV estimation problems.
@@
 
-    The errors-in-variables problem formulation (3.18) of the data-driven step input estimation method can be converted into an instrumental-variables problem.
+        The errors-in-variables problem formulation (3.21) of the data-driven step input estimation method can be converted into an instrumental-variables problem.
         A guide to do this conversion is found in \cite{Soderstrom18} and in \cite{Pintelon12Book}. 
-    The instrumental variable $\widetilde{\mathbf{Z}}$ has to be selected so that the normal equations (3.19) are expresed instead as
-    \begin{equation} \tag{4.22} \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{y}} . \label{eqn:neq_siv} \end{equation}
+        The instrumental variable $\widetilde{\mathbf{Z}}$ has to be selected so that the normal equations (3.22) are expresed instead as
+        \begin{equation} \tag{4.23} \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{y}} . \label{eqn:neq_siv} \end{equation}
         The selection of the instrumental variable has to be done in such a way that is correlated with the regression matrix $\widetilde{\mathbf{K}}$, and uncorrelated with the perturbations $\bm{\epsilon}$.
         Assuming that the block-Hankel matrix $\widetilde{\mathbf{K}}$ is given by
-    \begin{equation} \tag{4.23} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(n+1) & \Delta \widetilde{y}(n+2) & \cdots & \Delta \widetilde{y}(2n) \end{bmatrix} , \label{eqn:matrixK_r} \end{equation}
+        \begin{equation} \tag{4.24} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(n+1) & \Delta \widetilde{y}(n+2) & \cdots & \Delta \widetilde{y}(2n) \end{bmatrix} , \label{eqn:matrixK_r} \end{equation}
         then, the instrumental variable $\widetilde{\mathbf{Z}}$ should not contain the elements that construct $\widetilde{\mathbf{K}}$, i.e., the elements of the vector $\left( \widetilde{y}(1), \ldots, \widetilde{y}(2n+1) \right)^\top$.
         A possibility to generalize this idea is expressing the instrumental variable as
-    \begin{equation} \tag{4.24} \widetilde{\mathbf{Z}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-4n) & \Delta \widetilde{y}(t-4n+1) & \cdots & \Delta \widetilde{y}(t-3n-1) \\ G & \Delta \widetilde{y}(t-4n+1) & \Delta \widetilde{y}(t-4n+2) & \cdots & \Delta \widetilde{y}(t-3n) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-3n) & \Delta \widetilde{y}(t-3n+1) & \cdots & \Delta \widetilde{y}(t-2n-1) \end{bmatrix} , \label{eqn:matrixZ_t} \end{equation}
+        \begin{equation} \tag{4.25} \widetilde{\mathbf{Z}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-4n) & \Delta \widetilde{y}(t-4n+1) & \cdots & \Delta \widetilde{y}(t-3n-1) \\ G & \Delta \widetilde{y}(t-4n+1) & \Delta \widetilde{y}(t-4n+2) & \cdots & \Delta \widetilde{y}(t-3n) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-3n) & \Delta \widetilde{y}(t-3n+1) & \cdots & \Delta \widetilde{y}(t-2n-1) \end{bmatrix} , \label{eqn:matrixZ_t} \end{equation}
         for a regression matrix
-    \begin{equation} \tag{4.25} \widetilde{\mathbf{K}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-2n+1) & \Delta \widetilde{y}(t-2n+2) & \cdots & \Delta \widetilde{y}(t-n) \\ G & \Delta \widetilde{y}(t-2n+2) & \Delta \widetilde{y}(t-2n+3) & \cdots & \Delta \widetilde{y}(t-n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-n+1) & \Delta \widetilde{y}(t-n+2) & \cdots & \Delta \widetilde{y}(t) \end{bmatrix} , \label{eqn:matrixK_t} \end{equation}
+        \begin{equation} \tag{4.26} \widetilde{\mathbf{K}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-2n+1) & \Delta \widetilde{y}(t-2n+2) & \cdots & \Delta \widetilde{y}(t-n) \\ G & \Delta \widetilde{y}(t-2n+2) & \Delta \widetilde{y}(t-2n+3) & \cdots & \Delta \widetilde{y}(t-n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-n+1) & \Delta \widetilde{y}(t-n+2) & \cdots & \Delta \widetilde{y}(t) \end{bmatrix} , \label{eqn:matrixK_t} \end{equation}
         for $t = 4n+1, \ldots, N$.
@@
         For an errors-in-variables problem expressed as
-    \begin{equation} \tag{4.26} \widetilde{y}(t) = \widetilde{\bm{\phi}}(t) \bm{\theta}   \end{equation}
+        \begin{equation} \tag{4.27} \widetilde{y}(t) = \widetilde{\bm{\phi}}(t) \bm{\theta}   \end{equation}
         where $\widetilde{y}(t)$ is the output perturbed by measurement noise, $\bm{\theta}$ is the vector of to-be-estimated parameters, and $\widetilde{\bm{\phi}}(t)$ is the regressor vector constructed from delayed samples of the measured input $\widetilde{u}(t)$ and output $\widetilde{y}(t)$.
         The least-squares estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is the solution of the normal equation
-    \begin{equation} \tag{4.27} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} \widehat{\bm{\theta}}_{\mathrm{LS}} = \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} , \label{eqn:norm} \end{equation}
+        \begin{equation} \tag{4.28} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} \widehat{\bm{\theta}}_{\mathrm{LS}} = \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} , \label{eqn:norm} \end{equation}
         where $\widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}}$ is the covariance matrix of the regressor vector, and $\widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}}$ is the cross-covariance matrix of the regressor vector and the output.
         When the number of samples becomes infinitely large, these covariances matrices can be expressed as a sum of the contributions from exact data and measurement noise contributions
-    \begin{equation} \tag{4.28} \begin{aligned}  \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} &= \mathbb{E} \{ \widetilde{\bm{\phi}} \widetilde{\bm{\phi}}^\top \} = \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \\ \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} &= \mathbb{E}\{ \widetilde{\bm{\phi}} \widetilde{y}^\top \} = {\mathbf{r}}_{\bm{\phi} y} + {\mathbf{r}}_{\bm{\eta} \bm{\epsilon}} , \end{aligned} \end{equation}
+        \begin{equation} \tag{4.29} \begin{aligned}  \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} &= \mathbb{E} \{ \widetilde{\bm{\phi}} \widetilde{\bm{\phi}}^\top \} = \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \\ \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} &= \mathbb{E}\{ \widetilde{\bm{\phi}} \widetilde{y}^\top \} = {\mathbf{r}}_{\bm{\phi} y} + {\mathbf{r}}_{\bm{\eta} \bm{\epsilon}} , \end{aligned} \end{equation}
         where we are assuming that $\widetilde{y} =  y +\bm{\epsilon}$, $\widetilde{\bm{\phi}} = \bm{\phi} + \bm{\eta}$, and $\mathbf{r}_{\bm{\phi} y} = \mathbf{R}_{{\bm{\phi}}} \bm{\theta}$.
         The bias of the least-square solution can be expressed as 
-    \begin{equation} \tag{4.29} \mathbf{R}_{\widetilde{\bm{\phi}}} \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) = \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{R}_{\widetilde{\bm{\phi}}} \bm{\theta} = \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} - \mathbf{R}_{\bm{\eta}} \bm{\theta} , \end{equation}
+        \begin{equation} \tag{4.30} \mathbf{R}_{\widetilde{\bm{\phi}}} \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) = \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{R}_{\widetilde{\bm{\phi}}} \bm{\theta} = \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} - \mathbf{R}_{\bm{\eta}} \bm{\theta} , \end{equation}
 
         If we assume that the measurement noise is white for both the input and output, 
-    \begin{equation} \tag{4.30} \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} = \mathbb{E}\{ \bm{\eta} \bm{\epsilon}^\top \} = \mathbf{0} \quad \mathrm{and} \quad \mathbf{R}_{\bm{\eta}} = \begin{bmatrix} \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{n_y} & \mathbf{0} \\ \mathbf{0} & \sigma_u^2 \mathbf{I}_{n_u} \end{bmatrix}, \end{equation}
+        \begin{equation} \tag{4.31} \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} = \mathbb{E}\{ \bm{\eta} \bm{\epsilon}^\top \} = \mathbf{0} \quad \mathrm{and} \quad \mathbf{R}_{\bm{\eta}} = \begin{bmatrix} \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{n_y} & \mathbf{0} \\ \mathbf{0} & \sigma_u^2 \mathbf{I}_{n_u} \end{bmatrix}, \end{equation}
         where $\sigma_{\bm{\epsilon}}^2$ and $\sigma_u^2$ are the variances of the output and the input, respectively. The normal equation (\ref{eqn:norm}) implies that 
-    \begin{equation} \tag{4.31} \left( \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \right)  \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) \neq \mathbf{0}, \end{equation}
+        \begin{equation} \tag{4.32} \left( \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \right)  \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) \neq \mathbf{0}, \end{equation}
         and the estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is not consistent.
         Instead, a bias-compensating least squares estimation can be obtained with  
-    \begin{equation} \tag{4.32} \widehat{\bm{\theta}}_{\mathrm{BCLS}} = \left( \mathbf{R}_{\widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\eta}} \right)^{-1}  \left(  \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{r}_{\bm{\eta} \bm{\epsilon}} \right) .
+        \begin{equation} \tag{4.33} \widehat{\bm{\theta}}_{\mathrm{BCLS}} = \left( \mathbf{R}_{\widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\eta}} \right)^{-1}  \left(  \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{r}_{\bm{\eta} \bm{\epsilon}} \right) .
 \end{equation}
@@
         Since the generalized instrumental vector $\mathbf{z}(t)$ and the equation error $\widetilde{y}(t) - \widetilde{\bm{\phi}}(t) \bm{\theta}$ must be correlated, we can express the sistem of equation as
-    \begin{equation} \tag{4.33}\left( \mathbf{R}_{\widetilde{\mathbf{z}} \widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \right) \bm{\theta} = \left(  \mathbf{r}_{\widetilde{\mathbf{z}} \widetilde{y}} - \mathbf{r}_{\bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) \right) , \label{eqn:GIVE} \end{equation}
+        \begin{equation} \tag{4.34}\left( \mathbf{R}_{\widetilde{\mathbf{z}} \widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \right) \bm{\theta} = \left(  \mathbf{r}_{\widetilde{\mathbf{z}} \widetilde{y}} - \mathbf{r}_{\bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) \right) , \label{eqn:GIVE} \end{equation}
         where $\widetilde{\mathbf{z}} = \mathbf{z} + \bm{\zeta}$.
@@
         If we define
-    \begin{equation} \tag{4.34} \begin{aligned} \mathbf{h} \left( \bm{\rho}, \bm{\theta} \right)  &=  \mathbf{r}_{ \bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \bm{\theta}  =  \mathbf{J} \left( \bm{\theta} \right) \bm{\rho}, \\
+        \begin{equation} \tag{4.35} \begin{aligned} \mathbf{h} \left( \bm{\rho}, \bm{\theta} \right)  &=  \mathbf{r}_{ \bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \bm{\theta}  =  \mathbf{J} \left( \bm{\theta} \right) \bm{\rho}, \\
         \mathbf{R} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau=1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{\bm{\phi}}^\top \left( \tau \right) , \\     
@@
         then the equation (\ref{eqn:GIVE}) is equivalent to
-    \begin{equation} \tag{4.35} \mathbf{R} \left( t \right) \bm{\theta} = \mathbf{b} \left( t \right) . \end{equation}
+        \begin{equation} \tag{4.36} \mathbf{R} \left( t \right) \bm{\theta} = \mathbf{b} \left( t \right) . \end{equation}
         This representation allows a recursive form
-    \begin{equation} \tag{4.36} \begin{aligned} 
+        \begin{equation} \tag{4.37} \begin{aligned} 
         \mathbf{R} \left( t+1 \right)  &= \mathbf{R} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{\bm{\phi}}^\top \left( t+1 \right) - \mathbf{R} \left( t \right) \right), \\
@@
     \color{blue} 
-    The problem (3.17) admits a least-squares (LS) solution with normal equations given by
-    \begin{equation} \tag{3.19} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}}.   \end{equation} % \label{eqn:neq_seiv} 
+    The problem (3.20) admits a least-squares (LS) solution with normal equations given by
+    \begin{equation} \tag{3.22} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}}.   \end{equation} % \label{eqn:neq_seiv} 
     The closed form of the LS solution can be expressed as
-    \begin{equation} \tag{3.20} \widehat{\bm{\theta}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
+    \begin{equation} \tag{3.23} \widehat{\bm{\theta}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
     where $\widetilde{\mathbf{K}}^\dagger$ is the pseudo-inverse matrix of $\widetilde{\mathbf{K}}$.
@@
     An expression of the RLS equations is given in \cite{Kailath00book} as
-    \begin{equation} \tag{3.21} \begin{aligned} \widehat{\bm{\theta}}(t) &= \widehat{\bm{\theta}}(t-1) + \bm{\kappa}_{t} \left( \widetilde{y}(t) - \widetilde{k}_{t} \widehat{\bm{\theta}}(t-1) \right) , \\  \bm{\kappa}_{t} &= \bm{\Psi}(t-1) \widetilde{k}_{t}^\top / \left( 1 + \widetilde{k}_{t} \bm{\Psi}(t-1) \widetilde{k}_{t}^\top  \right) \\ \bm{\Psi}(t) &= \left( \mathbf{I} - \bm{\kappa}_{t} \widetilde{k}_{t} \right) \bm{\Psi}(t-1), \label{eqn:RLS} \end{aligned} \end{equation}
+    \begin{equation} \tag{3.24} \begin{aligned} \widehat{\bm{\theta}}(t) &= \widehat{\bm{\theta}}(t-1) + \bm{\kappa}_{t} \left( \widetilde{y}(t) - \widetilde{k}_{t} \widehat{\bm{\theta}}(t-1) \right) , \\  \bm{\kappa}_{t} &= \bm{\Psi}(t-1) \widetilde{k}_{t}^\top / \left( 1 + \widetilde{k}_{t} \bm{\Psi}(t-1) \widetilde{k}_{t}^\top  \right) \\ \bm{\Psi}(t) &= \left( \mathbf{I} - \bm{\kappa}_{t} \widetilde{k}_{t} \right) \bm{\Psi}(t-1), \label{eqn:RLS} \end{aligned} \end{equation}
     where $\widetilde{k}_{t}$ represents the row of $\widetilde{K}$ that corresponds to the $t\mathrm{-th}$ sample, $\bm{\kappa}_{t}$ is a gain scalar, $\Psi(t)$ is a covariance matrix, and the estimates and the covariance matrix are initialized respectively with $\widehat{\bm{\theta}}(1)= \widehat{\bm{\theta}}(0)$ and $\Psi(-1) = \Psi(0)$.
@@
     \color{blue}
-    The exponentially weighted recursive least-squares (RLS) can solve recursively the estimation problem (3.20).
+    The exponentially weighted recursive least-squares (RLS) can solve recursively the estimation problem (3.21).
     In \cite{Kailath00book} is described the exponentially weighted RLS algorithm, which is as follows
@@
     Applying the first difference operator to the system state-space representation (3.1) results in the autonomous system
-    \begin{equation} \tag{3.10} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta {y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
+    \begin{equation} \tag{3.11} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta {y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
     where $\Delta {u}(k) = \mathbf{0}$, for $k \geq 0$, and
@@
     If the response $\Delta {y}$  of this autonomous system is persistently exciting of order $L$, then the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ of $L+1$ block rows constructed from $\Delta {y}$ satisfies
-    \begin{equation} \tag{3.11} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) = \mathrm{rank} \left( \begin{bmatrix} \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \iddots & \iddots & \iddots \\ \Delta y(L+1) & \Delta y(L+2) & \cdots & \Delta y(L+n) \end{bmatrix} \right) \leq L,  \end{equation}
-    and the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ is a linear map to the natural responses of the augmented autonomous system (3.3) 
-    \begin{equation} \tag{3.12} \mathbf{y}_{\mathrm{natural}} = \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell}, \end{equation}
+    \begin{equation} \tag{3.12} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) = \mathrm{rank} \left( \begin{bmatrix} \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \iddots & \iddots & \iddots \\ \Delta y(L+1) & \Delta y(L+2) & \cdots & \Delta y(L+n) \end{bmatrix} \right) \leq L,  \end{equation}
+    and the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ is a linear map to the natural responses of the augmented autonomous system (3.4) 
+    \begin{equation} \tag{3.13} \mathbf{y}_{\mathrm{natural}} = \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell}, \end{equation}
     from any vector $\bm{\ell} \in \mathbb{R}^{n}$.
     The total response of the system is the sum of the forced response, due to the applied input ${u}$ and the natural response
-    \begin{equation} \tag{3.13} \mathbf{y} = \mathbf{y}_{\mathrm{forced}}+ \mathbf{y}_{\mathrm{natural}} =G {u} + \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell} ,  \end{equation} 
+    \begin{equation} \tag{3.14} \mathbf{y} = \mathbf{y}_{\mathrm{forced}}+ \mathbf{y}_{\mathrm{natural}} =G {u} + \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell} ,  \end{equation} 
     that is equivalent to
-    \begin{equation} \tag{3.14} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(N) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ G & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ G & \Delta y(N-n) & \Delta y(N-n+1) & \cdots & \Delta y(N-1) \end{bmatrix} }_{\mathbf{K}} \underbrace{ \begin{bmatrix} {{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
+    \begin{equation} \tag{3.15} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(N) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ G & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ G & \Delta y(N-n) & \Delta y(N-n+1) & \cdots & \Delta y(N-1) \end{bmatrix} }_{\mathbf{K}} \underbrace{ \begin{bmatrix} {{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
     where $N$ is the number of samples.
@@
     One set was obtained for the unstructured and uncorrelated EIV problem, and, the other set, for the structured and correlated EIV problem. 
-    The problem solved by the data-driven step input estimation method (3.20) belongs to the second set.
+    The problem solved by the data-driven step input estimation method (3.21) belongs to the second set.
     The objective of the MC simulations is to compare the results of the bias and variance formulas in these two cases. 
@@
 
-    The exact data is the system of linear equations \ref{eqn:ddsiemexd} built by the data-driven step input estimation method.
+    The exact data is the system of linear equations (\ref{eqn:ddsiemexd}) built by the data-driven step input estimation method.
     To build the system of equations, the processed signal ${y}$ is the exact response of a sensor to the input $0.1s$, where $s$ is the unit step response.  
     The sensor is modeled as a linear time-invariant mass-spring-damper system of order $n = 2$, that is given by 
-    \begin{equation} \tag{4.37}
+    \begin{equation} \tag{4.38}
     \mathbf{x}(k+1) = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{m+M} & \dfrac{-k_{\mathrm{d}}}{m+M} \end{bmatrix} \mathbf{x}(k) + \begin{bmatrix} 0\\ -g\end{bmatrix} u(k),  
@@
 This is the case for instance when I compare Section 4.2.2 and Section 5.1.
+
+    {\bfseries The first paragraph of Section 5.1 has been added to explain the purpose of the simulations in Chapters 5, with respect to the simulations in Chapter 4.   }
+
+    \color{blue} 
+    The simulations described in Chapter 4 were conducted with an illustrative example that is assumed to be a second order weighing sensor.
+    A weighing sensor was implemented to relate experimental results with the simulations. 
+    As it will be detailed later, the order of the practical sensor is at least is fifth order. 
+    Thus, the simulations presented in this section are useful to get insight on the impact of that increasing the model order has on the accuracy of the bias and variance predictions.    
+    \color{black}
+
     \item Except if you can convince me, do not use the name "maximum likelihood" when you do not introduce a likelihood you want to maximize.
@@
 You can also mention your contributions there but in a less technical way since the problem statement only comes in the Chapter preliminaries.
+
+	{\bfseries The motivation of the research has been improved in the introduction by describing real life examples of measurements in which the tradeoff between speed and estimation accuracy is noticeable.
+The structure of the thesis description and the contributions have also been revised in the Introduction.
+}
+
 	\item In each main chapter (Chapter 4, Chapter 5, Chapter 6) you can add an introduction subsection, a state-of-the-art subsection and then elaborate your contributions and methods.
@@
 	\color{blue}
-    The computational complexity of the RLS algorithm solution to the system of equations (3.14) is $O \left( \left( n+1 \right)^2 \right)$.
+    The computational complexity of the RLS algorithm solution to the system of equations (\ref{eqn:ddsiemexd}) is $O \left( \left( n+1 \right)^2 \right)$.
     The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
@@
 	\item Position your work with regard to the state of the art as well as regarding solutions based on low-rank matrix completion and compressed sensing.
+
+    {\bfseries The last paragraph in Chapter 3 was added to describe the position the data-driven step input estimation method with respect to the most powerful unfalsified model of the behavioral approach theory.
+In the future work section, it was added a description of the need for efficient optimization algorithms that could solve the low-rank approximation problem enabling the online estimation of the step input value.
+}
+
 	\item I recommend to use the formulation “we propose” than “I propose”
@@
     \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_2.pdf} 
+    \includegraphics[width=0.6\columnwidth]{../ChapterIntroduction/fig/Fig_2.pdf} 
     \caption{ \label{fig:thermometer} 
     \color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_T=0.5 s^{-1}$, caused by applying a step input of 67 $\circ \mathrm{C}$  when the thermometer was initially at 23 $\circ \mathrm{C}$.
-In this example, we should wait 10 s to have an estimation of the input with a relative error lower than 0.5 \%.
+In this example, we should wait 10 s to have an estimation of the input with a relative error lower than 0.5\%.
 \color{black} }
@@
     \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_3.pdf} 
+    \includegraphics[width=0.6\columnwidth]{../ChapterIntroduction/fig/Fig_3.pdf} 
     \caption{ \label{fig:scale} 
@@
 The sensor parameters are $k_{\mathrm{d}} = 5 \mathrm{N s/m}$, and $k_{\mathrm{s}} = 10000 \mathrm{N/m}$.
-In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement with a relative error smaller than 0.72 \%.
+In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement with a relative error smaller than 0.72\%.
 \color{black}  }
@@
 
-\bibliography{../Gus-thesis/Gus-thesis.bib}
+\bibliography{../Gus-thesis.bib}
 

### 2020-09-09 internal defense feedback/Answers-jury-members-report_solved.tex
One before the very last

--- 2020-08-21 internal defense feedback/Answers-jury-members-report - BK2.tex
@@
 
-\title{Jury members' comments on the private PhD defense of \linebreak Gustavo Quintana-Carapia}
+\title{List of solved Jury members' comments on \linebreak the private PhD defense of Gustavo Quintana-Carapia}
 
@@
 
-	{\bfseries The introduction of the chapters now includes the state of the art to position the PhD work with respect to the literature.}
-
 	\item  Add more clearly the motivation for the PhD research work by including motivational examples to the introduction.
-
-	{\bfseries Motivational examples have been added to the Introduction. }
 
@@
 
-	{\bfseries In Section 4.1.4, a discussion that compares IV methods with EIV methods has been added.}
+	\item  Discuss more clearly the aspects of scalability of the methods and their limitations in terms of model order, size of data sets (e.g., related to the inversion of LS matrices).
 
-	\item  Discuss more clearly the aspects of scalability of the methods and their limitations in terms of model order, size of data sets (e.g., related to the inversion of LS matrices).
 	\item  The equations on pages 12 and 13 seemingly contain errors and should be corrected.
 
-	{\bfseries The equations on pages 12 and 13 have been corrected.}
-
 	\item  The figure shown on page 66 was made with a single experiment; it is desirable to see the (averaged) effect on several experiments.
-
-	{\bfseries A series of 100 runs of the maximum-likelihod estimation method has been simulated and the results have been added to the manuscript in Section 6.3.2.}
 
@@
 \begin{itemize}
+
 	\item You develop results for uncertainty quantification; what type of uncertainties  can your approach deal with and which levels of noise can you consider?
 Elaborate on SNR levels; I’m also very interested in scalability, so how big the matrices can be.
+
 	\item  Concerning the scalability to data: eqn 3.6: how scalable is the approach in all aspects (order, data, linked to inversion for LS matrix); can you deal with hundreds, thousands of data?
-
-	{\bfseries The following paragraph has been added in Section 3.2 to discuss the scalability of the data-driven step input estimation method.}
-
-	\color{blue}
-    The computational complexity of the RLS algorithm solution to the system of equations (3.14) is $O \left( \left( n+1 \right)^2 \right)$.
-    The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
-    From there on, the RLS algorithm updates the solution with linear complexity.
-    Therefore, even though a large number of samples $N$ makes the matrix in the system of equations (3.14) increase in the number of rows, the estimation of the solution does not increase in complexity.
-    The data-driven step input estimation method is then scalable for any sensor of order $n$, and can be executed in devices with limited computational resources, provided that the computation of the  $n+1 \times n+1$ inverse matrix is feasible.
-    \color{black}
 
@@
 A more detailed discussion is desirable.
+
 	\item  On slide 25 (noisy graphs) it would be good to see a statistical analysis (more experiments than one).
-
-    {\bfseries 100 runs of the maximum-likelihod estimation method have been simulated and their results have been added to the manuscript in Section 6.3.2.}
-
-	\color{blue} 
-    The maximum-likelihood (ML) method estimated the affine input parameters from the sensor model transient response.
-    A total of 100 runs of the method were obtained, using different realizations of measurement noise.
-    The measurement noise variance was set to have an SNR of 40 dB.
-    The ML method used the first 50 samples to initialize the optimization variables and updated the variables every $N_{\mathrm{s}} = 5$ samples.
-    Figure \ref{fig:rele_lo_40dB_s10} shows the average of the observed relative errors in the estimation of the parameters $\widehat{a}$, $\widehat{b}$, $\widehat{x}_{\mathrm{ini,1}}$, and $\widehat{x}_{\mathrm{ini,2}}$.
-
-    \renewcommand{\thefigure}{6.8}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_7.pdf} 
-    \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method.
-After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and everyone of the estimates.
-The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%.
-\color{black} }
-    \end{figure}
-
-    Figure \ref{fig:cov_lo_40dB_s1} shows \color{blue} the average \color{black} of the variances that were observed on the diagonal of the covariance matrix $\mathbf{J}$. 
-    We can see that the estimation variances decrease as more samples are processed.
-    Moreover, the estimation variances of $\widehat{a}$ and $\widehat{b}$ obtained with the ML method are lower than the corresponding estimation MSE errors obtained from a Monte Carlo simulation of the subspace method.
-
-    \renewcommand{\thefigure}{6.9}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_8.pdf} 
-    \caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian.
-\color{blue} The estimates $\widehat{x}_{\mathrm{ini,1}}$ and $\widehat{a}$ have the smaller and larger variances, respectively.
-\color{black}  }
-    \end{figure}
-
-    \color{black}
 
@@
 
-	{\bfseries An a priori bound on the error introduced by the Taylor series expansion can be obtained in terms of the spectral radius of the matrix $\mathbf{M}$.
-In Section 4.1 the following description has been added: }
-
-    \color{blue} An a priori bound on the error introduced by the second order Taylor series expansion can be expressed in terms of the spectral radius $\rho(\mathbf{M})$.
-    Considering that the Taylor series expansion is an infinite sum we have that
-    \begin{equation} \tag{4.4} \begin{aligned} (\mathbf{I} + \mathbf{M})^{-1} &= \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \\
-    &=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \right) \\
-    &=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \end{aligned} \end{equation} 
-    Therefore we can obtain a bound for the second error expansion error if we take a matrix norm as follows
-    \begin{equation} \tag{4.5} \begin{aligned} \| (\mathbf{I} + \mathbf{M})^{-1} - ( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 ) \| &=  \| \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \| \\
-    & \leq \| \mathbf{M}^3 \| \| (\mathbf{I} + \mathbf{M})^{-1} \| \leq \dfrac{\| \mathbf{M} \|^3}{ 1 -  \| \mathbf{M} \|} = \dfrac{\rho(\mathbf{M})^3}{1 - \rho(\mathbf{M})}  \end{aligned} \end{equation}
-	\color{black}
-
 	\item  Concerning other sorts of prior estimations, e.g., given by explicit function theorems.
@@
 This can be investigated further.
-
-	\color{blue}
-	Neuberger's paper \cite{Neuberger07} introduces five theorems that are consequence of the continuous Newton method.
-	The Newton method aims to find a zero of the function $F: R^n \rightarrow R^n$ by iterating
-	\begin{equation} z_{k+1} = z_k - \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
-	for $k=0,1,2,\ldots$, and for a given initial $z_0$, assuming that $\left(F' (y) \right)^{-1}$ exists for some $y \in R^n$.
-	A domain of attraction is the set of all initial values $z_0$ that lead to a root of $F$ after the convergence $z_0, z_1, z_2,\ldots$.
-	The damped Newton method  
-	\begin{equation} z_{k+1} = z_k - \delta_k \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
-	prevents convergence issues, such as chaotic domains of attraction, with $\delta_1, \delta_2, \ldots \in (0, 1)$. 
-	The continuous Newton method is a sequence of damped Newton methods.
-To implement the continuous Newton method, select $T>0$ and perfom $m$ runs of the damped Newton method with $\delta_k = T/m$, for $k=1,2,\ldots, m$.
-	The result $z_{k+1}$ is the zero estimation $x_m$, and if the sequence of estimates $x_1, x_2, \ldots$ converges, then it converges to the result of the continuous Newton method.
-
-	The objective of the continuous Newton method is to find a function $z: [0, \infty) \rightarrow R^n$, so that
-    \begin{equation} z(0) = x \in R^n, \quad z' (t) = - \left(F' (z(t)) \right)^{-1} F(z(t)), \quad t \geq 0, \label{eqn:contNewton}\end{equation}
-    subject to the existence of the limit $u=\mathrm{lim}_{t \rightarrow \infty}{z(t)}$, that satisfies $F(u)=0$.
-    This implies 
-    \begin{equation}  F' (z(t)) z' (t) = - F(z(t)) , \end{equation}
-    that is equivalent to
-    \begin{equation}  \left(F \circ z\right)' (t) = - F(z(t)) , \end{equation}
-    from where we get
-    \begin{equation}  F(z(t)) = e^{-t} F(z(0)) . \end{equation}
-    Thus, the residual $F(z(t))$ only changes in magnitude and not in direction.
-    This property is essential for the results of the theorems in \cite{Neuberger07}. 
-    In particular, Theorem 5 gives the conditions for the continuous Newton method to ensure finding a root of the function.
-
-    \begin{customthm}{5} \label{thm:five}
-    Suppose that each of $H$, $J$, and $K$ is a Banach space, with $H$ compactly embedded in $J$, that $r>0$, and that $G:B_r(0) \rightarrow K$ is continuous as a function on $J$.
-Suppose also that $g$ belongs to $K$ and that for each $y$ in $b_r(0)$ there is an $h$ in $B_r(0)$, where $b_r(0)$ and $B_r(0)$ are open and closed balls in $H$ of radius $r$ centered at 0, such that
-    \begin{equation} \lim\limits_{t \to 0+} -{\dfrac{1}{t} \left(G(y+th) - G(y)\right) = g}. \end{equation}
-    Then, there exists $u$ in $B_r(0)$ such that $G(u)=g$.
-    \end{customthm}
-    %$F:B_r(0) \rightarrow K$
-
-    The theorem implies that $u$ is a zero of the function  defined as $F(y) = G(y) - g$.  
-    Since $G$ is a continuous function and if $\left(G' (y) \right)^{-1}$ exists for each $y$ in $B_r(0)$, a metric for the validity of the Theorem (\ref{thm:five}) hypothesis is the inequality $\left\Vert  \left( G' (y) \right)^{-1} g \right\Vert \leq r$.
-
-    To link the data-driven step-input estimation problem with this theorem, we can take the system of linear equations $\widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \bm{\theta}$, then define $G(y) = \widetilde{\mathbf{K}} \bm{\theta}$, and $g = \widetilde{\mathbf{y}}$, so that we have $F(\bm{\theta}) = G(\bm{\theta}) - g$.
-The derivative of the function $G(\bm{\theta})$ is given by $G'(\bm{\theta}) = \widetilde{\mathbf{K}} \mathbf{I}$.
-Using the pseudo-inverse of $\widetilde{\mathbf{K}}$, the metric $\left\Vert  \widetilde{\mathbf{K}}^{\dagger} \widetilde{\mathbf{y}} \right\Vert \leq r$ indicates that the estimate $\hat{\bm{\theta}}$ is found inside a ball of radius $r$.
-    Thus,  
-	\color{black}
-
 	\item  About trying to find a good transformation from LS to EIV model.
@@
 Something to investigate as well...
+
 	\item  About robustness: is it an issue in practical problems with heavy tailed noise, outliers,...
@@
 Might be interesting to ponder this in future.
+
 	\item  On deep learning: could this be an alternative?
+
 \end{itemize}
@@
 
-	{\bfseries The fast weighing of parcels in the post office is a typical example of the real life systems where the time is limited to perform the measurements.
-Another one that we see nowadays is the temperature measurement of people in airports or train stations to prevent the spread of corona virus.
-The data driven method can be implemented in portable devices with limited computational resources.
-In industrial plants were a time slot can be devoted to identify the measurement system and implement a compensation method, the solution might be useful only for a certain amount of time.
-It is assumed that the model approximation deviates due to changes in the system elements, leading to inaccuracies in the inpiut estimation.
-There is a need for adaptive methods to avoid stopping the production lines to repeat the system identification procedure.
-The data-driven method is such an adaptive method.
-}
-
 	\item  Chapter 2, Section 2.2.1.
@@
 It could be interesting to compare with such IV solutions.
+
 	\item  In Chapter 3: at beginning you say to use linear systems theory to model a sensor. Can you discuss this assumption: is it true in practice? Can you guarantee it behaves like an LTI system?
+
 	\item  Equation 3.1: an LTI system, described with SS representation.
@@
 Is it still white in this case?
+
 	\item  On page 12 standard approach top of page equation: there are some mistakes in the equations!
@@
 
-	{\bfseries The step input estimation based on singular value decomposition works only with exact data.
-When  noise perturbes the sensor response the method cannot guarantee the step input estimation.
-The singular value decomposition of a Hankel matrix constructed from a perturbed sensor response provides only an approximation of the singular values and the singular vectors of this Hankel matrix.
-Then, the step input estimation obtained with this method from perturbed sensor response is only a rough estimation.
-}
-
 	\item  On page 12: I’m not sure the equations at the end of the page are correct, e.g., Y=G ubar ...
@@
 
-	{\bfseries The equations have been reviewed and the typos were corrected.}
-
 	\item  On page 13, you assume persistency of excitation of order $L$. How do you check this? In your case the input is constant; how to guarantee that the assumption is met?
-
-	{\bfseries The persistency of excitation of a signal is verified by looking at the rank of the Hankel matrix constructed from the consecutive samples of this signal.
-When the Hankel matrix of $L+1$ block rows has a rank lower or equal than L, then we say that the signal is persistently exciting of order $L$.
-A step input has a very low order of persistency $L=1$, but the step respose of a system has a larger order of persistency.
-In fact, we assume that if the persistency of excitation of the step response is sufficiently high, then the columns of the Hankel matrix constructed from the step response span a linear space with all the responses of the equivalent augmented autonomous system where the step input is considered an augmented initial condition.}
 
@@
 
-	{\bfseries The following paragraph has been added in Section 4.1 to discuss on the need for the normality assumption.  }
-
-	\color{blue}
-	The normality assumption is necessary to provide more prior information into the method formulation.
-    The second order Taylor series expansion (4.4) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
-    Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
-    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance, which can be disregarded since $3 \sigma_{\bm{\epsilon}}^4 \ll \sigma_{\bm{\epsilon}}^2$\color{black}.
-
 	\item   In equation 4.11 you have $\sigma_eps^2 + \sigma_e^2$; how to get in practice $\sigma_e$? Do you have to do this each time? Recursively?
-
-	{\bfseries The measurementy noise variance is prior information and in Chapter 5 we describe two ways of estimating $\sigma_{\bm{\epsilon}}$ In summary, the measurement noise can be extracted fom the sensor ressponse once the transient effects are smaller than the noise floor, or from the sensor response power spectrum.
-The online estimation of the measurement noise, simultaneously with the online step input estimation is a topic for future research.
-}
 
@@
 You use only one realization also for matrices A, B, C, D in the text, so it can be added explicitly.
+
 	\item   in Fig 4.1 the bottom figure also discusses the structured case. Why do you talk about structured results in a section about the unstructured case?
-
-	{\bfseries Section 4.2 has been improved to describe the result next to its discussion.  }
 
@@
 
-	{\bfseries Section 4.2 has been reestructured and the simulation for the unstructured case were redone to establish a link between the two cases.
-The exact data is the same in the two cases and the perturbation noise follows the corresponding rules of each case, with the same variance noise levels.
-In this way the comparison between the two cases makes more sense.
-}
-
 	\item  To show comparison between structured and unstructured case, use unstructured results for structured case; then use unstructured results for structured case and cover all the possibilities.
 Now you compare curves related to two different simulations.
-
-	{\bfseries \color{red} Interesting question, I want to try it this evening \color{black}}
 
@@
 
-	{\bfseries The derivation described in Chapter 4 was conducted with a second order mass-spring-damper weighing system example on mind, and the results observed were submitted in the corresponding paper.
-Afterwards, when we moved to the practical implementation, we faced the increase of the model order after setting up the weighing experiment.
-The lowest order we can assume is fifth order.
-So we decided to repeat the simulations and these new results serve two purposes: a transition from simulation to experimental results, and also to get an insight on the impact of the bias and variance predictions after increasing the model order.
-}
-
 	\item   On the impact of selected order on results.
@@
 Can you use something in your case to test for a good order of system?
+
 	\item   Do you know of cross-validation: split data into two parts or several parts; some for estimation, some for validation?
@@
 Have you tried this?
+
 	\item   In Section 6.2.2.
@@
 
-	{\bfseries The method description in Section 6.2.2 has been modified to explain why the method is a maximum likelihood method.  }
-
 	\item   Concerning the ML approach in simulations, you say that it takes 30 seconds to complete.
@@
 I am a bit surprised you require 30 sec for 2 parameters.
+
 \end{itemize}
@@
 \begin{itemize}
-	\item When solving algorithms.
-Do you opt for closed-form solutions, or do you use optimization based algorithms?
-How to solve LS problems?
-Do you solve them using pseudo-inverse or do you do gradient descent?
-	\item  Today there is an overwhelming amount of data sensed and communicated; How would you apply your method in a scenario where dimensionality increases significantly?
-This relates to the questions of previous jury members; what is the implication of using numerical versus closed form solutions?
-	\item  What if the $K^\top K$ matrix is not invertible?
-What if you have fewer samples than unknowns; what can you do?
-Solutions that assume structure like in compressed sensing, regularization,...
-can you use them, what would be the impact of that?
-	\item  You mention that you improve complexity with your approach, but I missed a complexity analysis. Did you do this? Anything in processing time, flops,...?
-	\item  You mentioned deep learning could be applied; could you show how?
-An issue is that neural nets are trained in high SNR data, but applied in engineering with lower SNR.
-This clogs performance; any ideas?
-Is this a limitation?
-	\item  You mention that you consider Kalman filters; have you seen approaches where they write Deep Learning in form of Kalman Filter; Could this be applicable?
+
 	\item  Recommendation: if you revise the thesis: maybe change "I propose" to "we propose".
-
-	{\bfseries The formulation “I propose” was replaced by “we propose”. }
 
@@
 \begin{itemize}
+
     \item Your main assumption is that a sensor outputs a steady state value that is proportional to the to-be-measured quantity.
@@
 Can you still use your methods?
+
 	\item  Sensors are physical devices and deliver a continuous-time output. The models that you use are in discrete-time. Are there any risks that you miss information by not sampling fast enough?
-
-	{\bfseries In Chapter 3 has been added a discussion on the step invariance transformation.
-This defines the conditions of the input for describing exactly the system dynamics in discrete time after sampling the continuous system response.
-If the input is stepwise continuous, then the zero order hold sampling of the step response leads to an exact discrete time representation of the system.
-The method under study in this thesis was developed for step input, then we can rely on the discrete time representation.
-There exist corresponding invariant transformations for piecewise linear inputs and for piecewise polynomial inputs that lead to exact representations from first and second order sampling of the system response, respectively.
-}
 
@@
 Would it be possible to plug in a procedure that does structured matrix completion and approximation into your methods?
+
 \end{itemize}
@@
 Can we tackle this?
+
 	\item  Consider an optical fiber link with a detector capable of 100 million measurements per second.
@@
 You have to do calculations: how high can you go in speed before running into troubles with DSP?
+
 \end{itemize}
@@
 
-    {\bfseries Section 4.1 has been appended with a paragraph to discuss on the need for the normality assumption.  }
-
 	\item  Returning to a question by Philippe Dreesen on missing data.
@@
 It works in low-rank approximations; not if the matrix is of full rank.
+
 \end{itemize}
@@
 
-	{\bfseries The last paragraphs of Section 3.2 have been included to discuss the scalability of the data-driven step input estimation method.  }
-
 \end{itemize}
@@
     \begin{itemize}
+
     \item the reasons why, in the industry, we need such a new technique.
@@
 What do you mean by fast?
+
     \item  the literature on EIV estimation problems.
@@
 
-        {\bfseries A description of IV methods has been added in Section 4.1.4}
-
-        \color{blue}
-    \subsection*{4.1.4. Instrumental variables formulation of the data-driven step input estimation method}
-
-    The errors-in-variables problem formulation (3.18) of the data-driven step input estimation method can be converted into an instrumental-variables problem.
-    A guide to do this conversion is found in \cite{Soderstrom18} and in \cite{Pintelon12Book}. 
-    The instrumental variable $\widetilde{\mathbf{Z}}$ has to be selected so that the normal equations (3.19) are expresed instead as
-    \begin{equation} \tag{4.22} \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{y}} . \label{eqn:neq_siv} \end{equation}
-    The selection of the instrumental variable has to be done in such a way that is correlated the regression matrix $\widetilde{\mathbf{K}}$, and uncorrelated with the perturbations $\bm{\epsilon}$.
-    Assuming that the block-Hankel matrix $\widetilde{\mathbf{K}}$ is given by
-    \begin{equation} \tag{4.23} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(n+1) & \Delta \widetilde{y}(n+2) & \cdots & \Delta \widetilde{y}(2n) \end{bmatrix} , \label{eqn:matrixK_r} \end{equation}
-    then, the instrumental variable $\widetilde{\mathbf{Z}}$ should not contain the elements that construct $\widetilde{\mathbf{K}}$, i.e., the elements of the vector $\left( \widetilde{y}(1), \ldots, \widetilde{y}(2n+1) \right)^\top$.
-    A possibility to generalize this idea is expressing the instrumental variable as
-    \begin{equation} \tag{4.24} \widetilde{\mathbf{Z}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-4n) & \Delta \widetilde{y}(t-4n+1) & \cdots & \Delta \widetilde{y}(t-3n-1) \\ G & \Delta \widetilde{y}(t-4n+1) & \Delta \widetilde{y}(t-4n+2) & \cdots & \Delta \widetilde{y}(t-3n) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-3n) & \Delta \widetilde{y}(t-3n+1) & \cdots & \Delta \widetilde{y}(t-2n-1) \end{bmatrix} , \label{eqn:matrixZ_t} \end{equation}
-    for a regression matrix
-    \begin{equation} \tag{4.25} \widetilde{\mathbf{K}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-2n+1) & \Delta \widetilde{y}(t-2n+2) & \cdots & \Delta \widetilde{y}(t-n) \\ G & \Delta \widetilde{y}(t-2n+2) & \Delta \widetilde{y}(t-2n+3) & \cdots & \Delta \widetilde{y}(t-n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-n+1) & \Delta \widetilde{y}(t-n+2) & \cdots & \Delta \widetilde{y}(t) \end{bmatrix} , \label{eqn:matrixK_t} \end{equation}
-    for $t = 4n+1, \ldots, N$.
-
-    The difference with the EIV formulation is that this implementation of the IV method requires to wait more, until $4n+1$ samples of the sensor response are read, instead of $2n+1$ samples. 
-    For sensors of small order this twice-the-time waiting does not represent a big change, but as the sensor order increases, the need for waiting can turn the method into impractical since a lot of time should pass before starting to estimate the input.
-
-    There exists a generalized instrumental variables method (GIVE) that was proposed to identify dynamic models from input/output data. 
-    GIVE is a bias-eliminating method that uses a generalized instrumental vector constructed from delayed values of the input and output signals \cite{Soderstrom18}.
-    Since the aim of this thesis is to analize a method that directly estimates the input from the observed sensor output, without identifying explicitly a dynamic model of the sensor, the GIVE method is out of the scope of this work.
-    A future research can include studying the formulation of a data-driven input estimation method in the spirit of the GIVE method. 
-    Nevertheless, a brief description of the GIVE method and its recursive version is given in the following paragraphs.
-
-    \subsubsection*{4.1.4.1 Generalized instrumental variables method}
-    For an errors-in-variables problem expressed as
-    \begin{equation} \tag{4.26} \widetilde{y}(t) = \widetilde{\bm{\phi}}(t) \bm{\theta}   \end{equation}
-    where $\widetilde{y}(t)$ is the output perturbed by measurement noise, $\bm{\theta}$ is the vector of to-be-estimated parameters, and $\widetilde{\bm{\phi}}(t)$ is the regressor vector constructed from delayed samples of the measured input $\widetilde{u}(t)$ and output $\widetilde{y}(t)$.
-    The least-squares estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is the solution of the normal equation
-    \begin{equation} \tag{4.27} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} \widehat{\bm{\theta}}_{\mathrm{LS}} = \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} , \label{eqn:norm} \end{equation}
-    where $\widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}}$ is the covariance matrix of the regressor vector, and $\widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}}$ is the cross-covariance matrix of the regressor vector and the output.
-    When the number of samples becomes infinitely large, these covariances matrices can be expressed as a sum of the contributions from exact data and measurement noise contributions
-    \begin{equation} \tag{4.28} \begin{aligned}  \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{R}}_{\widetilde{\bm{\phi}}} &= \mathbb{E} \{ \widetilde{\bm{\phi}} \widetilde{\bm{\phi}}^\top \} = \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \\ \lim\limits_{N \rightarrow \infty} \widehat{\mathbf{r}}_{\widetilde{\bm{\phi}} \widetilde{y}} &= \mathbb{E}\{ \widetilde{\bm{\phi}} \widetilde{y}^\top \} = {\mathbf{r}}_{\bm{\phi} y} + {\mathbf{r}}_{\bm{\eta} \bm{\epsilon}} , \end{aligned} \end{equation}
-    where we are assuming that $\widetilde{y} =  y +\bm{\epsilon}$, $\widetilde{\bm{\phi}} = \bm{\phi} + \bm{\eta}$, and $\mathbf{r}_{\bm{\phi} y} = \mathbf{R}_{{\bm{\phi}}} \bm{\theta}$.
-    The bias of the least-square solution can be expressed as 
-    \begin{equation} \tag{4.29} \mathbf{R}_{\widetilde{\bm{\phi}}} \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) = \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{R}_{\widetilde{\bm{\phi}}} \bm{\theta} = \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} - \mathbf{R}_{\bm{\eta}} \bm{\theta} , \end{equation}
-
-    If we assume that the measurement noise is white for both the input and output, 
-    \begin{equation} \tag{4.30} \mathbf{r}_{\widetilde{\bm{\eta}} \bm{\epsilon}} = \mathbb{E}\{ \bm{\eta} \bm{\epsilon}^\top \} = \mathbf{0} \quad \mathrm{and} \quad \mathbf{R}_{\bm{\eta}} = \begin{bmatrix} \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{n_y} & \mathbf{0} \\ \mathbf{0} & \sigma_u^2 \mathbf{I}_{n_u} \end{bmatrix}, \end{equation}
-    where $\sigma_{\bm{\epsilon}}^2$ and $\sigma_u^2$ are the variances of the output and the input, respectively. The normal equation (\ref{eqn:norm}) implies that 
-    \begin{equation} \tag{4.31} \left( \mathbf{R}_{\bm{\phi}} + \mathbf{R}_{\bm{\eta}} \right)  \left( \widehat{\bm{\theta}}_{\mathrm{LS}} - \bm{\theta} \right) \neq \mathbf{0}, \end{equation}
-    and the estimate $\widehat{\bm{\theta}}_{\mathrm{LS}}$ is not consistent.
-    Instead, a bias-compensating least squares estimation can be obtained with  
-    \begin{equation} \tag{4.32} \widehat{\bm{\theta}}_{\mathrm{BCLS}} = \left( \mathbf{R}_{\widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\eta}} \right)^{-1}  \left(  \mathbf{r}_{\widetilde{\bm{\phi}} \widetilde{y}} - \mathbf{r}_{\bm{\eta} \bm{\epsilon}} \right) .
-\end{equation}
-
-    To formulate the generalized instrumental variable estimator, a generalized instrumental vector $\mathbf{z}(t)$ is built from delayed samples of the input and the output. 
-    The dimension of the generalized instrumental vector should be $n_{\mathbf{z}} \geq \mathrm{dim} ( \vartheta ) = n_u + n_y + 2$, since the total parameter vector $\vartheta = \left( \theta^\top, \rho^\top \right)^\top$ considers also the variances of the measurement noise $\rho = \left( \sigma_{\bm{\epsilon}}^2, \sigma_u^2 \right)^\top$.
-    Since the generalized instrumental vector $\mathbf{z}(t)$ and the equation error $\widetilde{y}(t) - \widetilde{\bm{\phi}}(t) \bm{\theta}$ must be correlated, we can express the sistem of equation as
-    \begin{equation} \tag{4.33}\left( \mathbf{R}_{\widetilde{\mathbf{z}} \widetilde{\bm{\phi}}} - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \right) \bm{\theta} = \left(  \mathbf{r}_{\widetilde{\mathbf{z}} \widetilde{y}} - \mathbf{r}_{\bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) \right) , \label{eqn:GIVE} \end{equation}
-    where $\widetilde{\mathbf{z}} = \mathbf{z} + \bm{\zeta}$.
-
-    \subsubsection*{4.1.4.2 Recursive generalized instrumental variables method}
-    If we define
-    \begin{equation} \tag{4.34} \begin{aligned} \mathbf{h} \left( \bm{\rho}, \bm{\theta} \right)  &=  \mathbf{r}_{ \bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \bm{\theta}  =  \mathbf{J} \left( \bm{\theta} \right) \bm{\rho}, \\
-    \mathbf{R} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{\bm{\phi}}^\top \left( \tau \right) , \\     
-    \mathbf{q} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{y} \left( \tau \right) \\     
-    \mathbf{b} \left( t \right)  &=  \mathbf{q} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right)
-    , \end{aligned} \end{equation}
-    then the equation (\ref{eqn:GIVE}) is equivalent to
-    \begin{equation} \tag{4.35} \mathbf{R} \left( t \right) \bm{\theta} = \mathbf{b} \left( t \right) . \end{equation}
-    This representation allows a recursive form
-    \begin{equation} \tag{4.36} \begin{aligned} 
-    \mathbf{R} \left( t+1 \right)  &= \mathbf{R} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{\bm{\phi}}^\top \left( t+1 \right) - \mathbf{R} \left( t \right) \right), \\
-    \mathbf{b} \left( t+1 \right)  &= \mathbf{b} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{b} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right) \right), \\
-    \widehat{\bm{\theta}} \left( t+1 \right)  &=  \left( \mathbf{R}^\top \left( t+1 \right) \mathbf{R} \left( t+1 \right) \right)^{-1} \mathbf{R}^\top \left( t+1 \right) \mathbf{b} \left( t+1 \right) \\
-    \mathbf{q} \left( t+1 \right)  &= \mathbf{q} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{q} \left( t \right) \right) \\
-    \widehat{\bm{\rho}} \left( t+1 \right)  &=  \mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right) \left( \mathbf{q} \left( t \right) - \mathbf{R} \left( t \right) \widehat{\bm{\theta}} \left( t \right) \right) , \end{aligned} \end{equation}
-    where $\mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right)$ is the pseudo-inverse of $\mathbf{J} \left( \bm{\theta} \right)$.
-
-    The complexity of this recursive generalized instrumental variables estimator is ${O}(n^3)$, that is larger than the ${O}(n^2)$ of the recursive least squares.
-    Another drawback of the recursive generalized IV estimator is that it computes a pseudo-inverse matrix in each iteration step.
-    If we would need to formulate the data-driven step input estimation problem in the generalized IV framework, it would require to study the formulation in detail because currently the IV framework is based on the assumption that both measurements of the input and output are available, but for the step input estimation problem only the output is available.
-    Therefore, it is not straightforward to implement the recursive generalized IV estimator to estimate the step input, and this can lead to future research.
-    \color{black}
-
     \end{itemize}
@@
 
-    {\bfseries The following recursive least-squares algorithm description has been added to Section 3.2:}
-
-    \color{blue} 
-    The problem (3.17) admits a least-squares (LS) solution with normal equations given by
-    \begin{equation} \tag{3.19} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}}.   \end{equation} % \label{eqn:neq_seiv} 
-    The closed form of the LS solution can be expressed as
-    \begin{equation} \tag{3.20} \widehat{\bm{\theta}} = \widetilde{\mathbf{K}}^\dagger \widetilde{\mathbf{y}} = ( \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{K}} )^{-1} \widetilde{\mathbf{K}}^\top \widetilde{\mathbf{y}} , \label{eqn:xhat} \end{equation}
-    where $\widetilde{\mathbf{K}}^\dagger$ is the pseudo-inverse matrix of $\widetilde{\mathbf{K}}$.
-    As the number of samples $N$ increases, the number of rows of the matrix $\widetilde{\mathbf{K}}$ grows, making the pseudo-inverse computation inefficient since it requires larger flops, memory and time.
-
-    For metrology applications, it is desired to have a fast solution that can be obtained in real-time. 
-    The recursive algorithm (RLS) is a least-squares alternative to implement the step input estimation method in real-time.
-    The RLS recursively updates the solution of the system of equations, for each newly acquired sensor response sample, considering the previous value of the estimation, instead of performing a matrix inversion.
-    An expression of the RLS equations is given in \cite{Kailath00book} as
-    \begin{equation} \tag{3.21} \begin{aligned} \widehat{\bm{\theta}}(t) &= \widehat{\bm{\theta}}(t-1) + \bm{\kappa}_{t} \left( \widetilde{y}(t) - \widetilde{k}_{t} \widehat{\bm{\theta}}(t-1) \right) , \\  \bm{\kappa}_{t} &= \bm{\Psi}(t-1) \widetilde{k}_{t}^\top / \left( 1 + \widetilde{k}_{t} \bm{\Psi}(t-1) \widetilde{k}_{t}^\top  \right) \\ \bm{\Psi}(t) &= \left( \mathbf{I} - \bm{\kappa}_{t} \widetilde{k}_{t} \right) \bm{\Psi}(t-1), \label{eqn:RLS} \end{aligned} \end{equation}
-    where $\widetilde{k}_{t}$ represents the row of $\widetilde{K}$ that corresponds to the $t\mathrm{-th}$ sample, $\bm{\kappa}_{t}$ is a gain scalar, $\Psi(t)$ is a covariance matrix, and the estimates and the covariance matrix are initialized respectively with $\widehat{\bm{\theta}}(1)= \widehat{\bm{\theta}}(0)$ and $\Psi(-1) = \Psi(0)$.
-    Since the RLS is a recursive implementation of the least-squares computation, the RLS solution is equivalent to the LS solution.
-    Moreover, the statistical analysis of the LS solution is valid also for the RLS solution, and is preferred since it is simpler to express the statistics using the closed form of the LS solution.
-    \color{black}
-
-    {\bfseries The following description for the exponentially recursive least-squares algorithm has been added to Section 6.2.1:}
-
-    \color{blue}
-    The exponentially weighted recursive least-squares (RLS) can solve recursively the estimation problem (3.20).
-    In \cite{Kailath00book} is described the exponentially weighted RLS algorithm, which is as follows
-
-    \begin{equation} \tag{6.4} \begin{aligned}  \widehat{\bm{\theta}}(k) &= \widehat{\bm{\theta}}(k-1) + \bm{\kappa}_{k} \left( \widetilde{y}(k) - \widetilde{\mathbf{k}}_{k} \widehat{\bm{\theta}}(k-1) \right) , \\  \bm{\kappa}_{k} &= \omega^{-1} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top / \left( 1 + \omega^{-1} \widetilde{\mathbf{k}}_{k} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top  \right) \\ \bm{\Psi}(k) &= \omega^{-1} \left( \mathbf{I} - \bm{\kappa}_{k} \widetilde{\mathbf{k}}_{k} \right) \bm{\Psi}(k-1), \label{eqn:RLS} \end{aligned} \end{equation}
-    for $k = 2n+1, 2n+2, \ldots$, where $\widetilde{\mathbf{k}}_{k}$ represents the row of $\widetilde{\mathbf{K}}$ that corresponds to the $k\mathrm{-th}$ sample, $\bm{\kappa}_{k}$ is a gain scalar, and $\Psi(k)$ is a covariance matrix.
-    The estimates and the covariance matrix are initialized using the first $n+1$ samples, i.e.,  
-    \begin{equation} \tag{6.5} \begin{aligned} \widehat{\bm{\theta}} (2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top  \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega}  \widetilde{\mathbf{y}}_{n+1}, \quad \mathrm{and} \\ \Psi(2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} , \end{aligned} \end{equation}
-    where $\widetilde{\mathbf{K}}_{n+1}$ is the matrix $\widetilde{\mathbf{K}}$ with the first $n+1$ rows, and $\widetilde{\mathbf{y}}_{n+1}$ is the vector $\widetilde{\mathbf{y}}$ with the first $n+1$ elements.
-    When $\omega=1$, the exponentially weighted RLS solution is identical to that of the RLS.
-    When $\omega<1$, the older residuals are weighted with lower values than the residuals of recent observations.
-    In this way, the solution of the minimization problem depends more on newer data, and less on older data.
-
-    \color{black}
-
     \item Double check your equations page 12 and 13. By the way, can you explain where Eq. (3.6) comes from?
-
-    {\bfseries The equations in page 12 and 13 have been reviewed and typos corrected. The derivation of Eq. (3.6) has been improved with more explaination: }
-
-    \color{blue}
-    A third method can directly estimate the input step level from the step response, without identifying a sensor model.
-    To derive this method, we use the first difference operator, $\Delta = \sigma - 1$, where $\sigma$ is the shift operator, defined as $(\sigma^\tau y) (k) = y(k + \tau)$.
-    Applying the first difference operator to the system state-space representation (3.1) results in the autonomous system
-    \begin{equation} \tag{3.10} \begin{aligned} \Delta \mathbf{x}(k+1) = \mathbf{A} \Delta \mathbf{x}(k), \quad \Delta {y}(k) = \mathbf{C} \Delta \mathbf{x}(k), \quad \text{with} \quad \Delta \mathbf{x}_{\text{ini}} = \Delta \mathbf{x}(0) , \label{eqn:ssalti} \end{aligned} \end{equation}
-    where $\Delta {u}(k) = \mathbf{0}$, for $k \geq 0$, and
-    $\Delta \mathbf{x}(0) = (\mathbf{A} - \mathbf{I}) \mathbf{x}(0) + \mathbf{B} {{u}}$.
-
-    If the response $\Delta {y}$  of this autonomous system is persistently exciting of order $L$, then the rank of the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ of $L+1$ block rows constructed from $\Delta {y}$ satisfies
-    \begin{equation} \tag{3.11} \mathrm{rank} \left( \mathbfcal{H}_{L+1} \left( \Delta {y} \right) \right) = \mathrm{rank} \left( \begin{bmatrix} \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \iddots & \iddots & \iddots \\ \Delta y(L+1) & \Delta y(L+2) & \cdots & \Delta y(L+n) \end{bmatrix} \right) \leq L,  \end{equation}
-    and the Hankel matrix $\mathbfcal{H}_{L+1}(\Delta {y})$ is a linear map to the natural responses of the augmented autonomous system (3.3) 
-    \begin{equation} \tag{3.12} \mathbf{y}_{\mathrm{natural}} = \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell}, \end{equation}
-    from any vector $\bm{\ell} \in \mathbb{R}^{n}$.
-    The total response of the system is the sum of the forced response, due to the applied input ${u}$ and the natural response
-    \begin{equation} \tag{3.13} \mathbf{y} = \mathbf{y}_{\mathrm{forced}}+ \mathbf{y}_{\mathrm{natural}} =G {u} + \mathbfcal{H}\left(\Delta {y}\right) \bm{\ell} ,  \end{equation} 
-    that is equivalent to
-    \begin{equation} \tag{3.14} \underbrace{ \begin{bmatrix} y(n+1) \\ \vdots \\ y(N) \end{bmatrix}}_{\mathbf{y}} = \underbrace{ \begin{bmatrix} G & \Delta y(1) & \Delta y(2) & \cdots & \Delta y(n) \\ G & \Delta y(2) & \Delta y(3) & \cdots & \Delta y(n+1) \\ \vdots & \iddots & \iddots & \iddots \\ G & \Delta y(N-n) & \Delta y(N-n+1) & \cdots & \Delta y(N-1) \end{bmatrix} }_{\mathbf{K}} \underbrace{ \begin{bmatrix} {{u}} \\ \bm{\ell} \end{bmatrix} }_{\bm{\theta}} , \label{eqn:ddsiemexd} \end{equation}
-    where $N$ is the number of samples.
-    From this point of view, the vector $\bm{\ell}$ is a linear transformation of the system's initial conditions.  
-    \color{black}
 
@@
 
-    {\bfseries The discussion of the normality assumption has been added to Section 4.1, since it is relevant for the two cases considered.
-Section 4.1.3 has been reestructured to improve its clarity.}
+    \item  Why do you always study the statistical properties of your least squares solution with high SNR only? What do you get for SNR = 10 dB for instance?
 
-    \item  Why do you always study the statistical properties of your least squares solution with high SNR only? What do you get for SNR = 10 dB for instance? 
     \item Give the values of the matrices you use for simulating data sets so that the reader can test what you have carried out.
-
-    {\bfseries The first paragraphs of Section 4.2, and Fig. 4.1, were added to describe the conditions in which the simulations were conducted from the structured and unstructured EIV problems.}
-
-    \color{blue}
-    Two Monte Carlo (MC) simulation studies were conducted to test the obtained bias and variance formulas.
-    There are two sets of formulas.
-    One set was obtained for the unstructured and uncorrelated EIV problem, and, the other set, for the structured and correlated EIV problem. 
-    The problem solved by the data-driven step input estimation method (3.20) belongs to the second set.
-    The objective of the MC simulations is to compare the results of the bias and variance formulas in these two cases. 
-    The comparison is an evaluation of the effect that the structure and correlation have in the problem solution.
-    To make a fair comparison, the same exact data and the perturbation variance is used in both types of EIV problems.
-
-    The exact data is the system of linear equations \ref{eqn:ddsiemexd} built by the data-driven step input estimation method.
-    To build the system of equations, the processed signal ${y}$ is the exact response of a sensor to the input $0.1s$, where $s$ is the unit step response.  
-    The sensor is modeled as a linear time-invariant mass-spring-damper system of order $n = 2$, that is given by 
-    \begin{equation} \tag{4.37}
-    \mathbf{x}(k+1) = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{m+M} & \dfrac{-k_{\mathrm{d}}}{m+M} \end{bmatrix} \mathbf{x}(k) + \begin{bmatrix} 0\\ -g\end{bmatrix} u(k),  
-    \quad y(k) = \begin{bmatrix} -1 & 0 \end{bmatrix} \mathbf{x}(k), \label{eqn:msdst}
-    \end{equation}
-    where $m=0.015$ kg, $k_{\mathrm{s}}=500$ kg/m, $k_{\mathrm{d}}=0.5$ kg s/m, $M=0.1$ kg, and $g=9.81 \ \mathrm{m/s^2}$.   
-    The exact sensor response is shown in Figure 4.1.
-
-    In the simulations, the sample size considered is $N=200$, since there the transient regime of the sensor model is still evident. 
-    In the case of the structured EIV problem, this sample size satisfies the requirements of the data-driven step input estimation method. 
-    \color{black}
 
@@
 
-    {\bfseries Fig. 4.1 b has been relocated in its corresponding section. Now it is Fig. 4.7.}
-
     \item  When you generate different simulation examples, explain why the new ones are interesting or, more importantly, what they bring w.r.t.
@@
 This is the case for instance when I compare Section 4.2.2 and Section 5.1.
+
     \item Except if you can convince me, do not use the name "maximum likelihood" when you do not introduce a likelihood you want to maximize.
@@
 
-    {\bfseries The method description in Section 6.2.2 has been modified to explain why the method is a maximum likelihood method. The modified paragraphs are as follows: }
-
-    \color{blue} 
-    Using a model of the sensor, the maximum-likelihood ML method simultaneously estimates the parameters of the applied affine input $u(t) = at + b$, and some of the sensor model parameters, using the observed sensor response $\widetilde{\mathbf{y}}$.
-    This method fits iteratively the response of the model $\widehat{\mathbf{y}}(\bm{\theta})$ to the sensor response by searching for the parameters $\bm{\theta}$ that reduce the error difference between both responses.
-    Algorithm 6.1 describes the formulation of an objective function $f = \mathbf{r}^\top \mathbf{r}$, where $\mathbf{r}$ is the residual $\mathbf{r}(\bm{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\bm{\theta})$, that is minimized through the iterations.
-    The objective function is the sum of the squares of the differences between the samples of both responses.
-    Since the measurement noise is assumed Gaussian distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$, where the samples of this perturbation are independent and identically distributed, the minimization of the objective function $f$ maximizes the likelihood of fitting the actual sensor response with the model response.
-
-    The Jacobian matrix of the residual $\mathbf{J}\bm{\theta} = \partial \mathbf{r}(\bm{\theta}) / \partial \bm{\theta}$ can be analytically derived using the sensor model.
-    With the Jacobian matrix, the ML method searches the direction in which the residual decreases towards a local minimum.
-    Unfortunately, this ML method cannot guarantee the global minimum since it depends strongly on the initialization of the optimization parameters, as it is discussed in \cite{Nocedal06}.
-    Therefore, it is proposed to initialize the to-be-optimized parameters $a$ and $b$ using the results of the subspace method obtained after a convenient number of samples so that the affine input parameters are set close to the optimal values.
-    The initialization of the remaining model parameters under study can be done using the sensor model representation.
-    An illustrative example of the ML will be provided in a posterior subsection.  
-    \color{black}
 \end{itemize}
@@
     \begin{itemize}
+
+    \item  Today there is an overwhelming amount of data sensed and communicated; How would you apply your method in a scenario where dimensionality increases significantly?
+This relates to the questions of previous jury members; what is the implication of using numerical versus closed form solutions?
+
+    \item When solving algorithms.
+Do you opt for closed-form solutions, or do you use optimization based algorithms?
+How to solve LS problems?
+Do you solve them using pseudo-inverse or do you do gradient descent?
+
     \item In the introduction chapter it is best to describe a set of applications and settings that motivate your research and proposed solutions.
@@
 You can also mention your contributions there but in a less technical way since the problem statement only comes in the Chapter preliminaries.
+
 	\item In each main chapter (Chapter 4, Chapter 5, Chapter 6) you can add an introduction subsection, a state-of-the-art subsection and then elaborate your contributions and methods.
 
-    {\bfseries The introduction of the chapters now includes the state of the art to position the PhD work with respect to the literature.}
-
 	\item I believe that this structure will improve the understanding of the thesis and light up your contributions.
-
-	{\bfseries Indeed, this structure allows to visualize the thesis contributions.}
 
@@
 
-    {\bfseries The following paragraph has been added in Section 3.2 to discuss the scalability of the data-driven step input estimation method.}
+	\item Position your work with regard to the state of the art as well as regarding solutions based on low-rank matrix completion and compressed sensing.
 
-	\color{blue}
-    The computational complexity of the RLS algorithm solution to the system of equations (3.14) is $O \left( \left( n+1 \right)^2 \right)$.
-    The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
-    From there on, the RLS algorithm updates the solution with linear complexity.
-    Therefore, even though a large number of samples $N$ makes the matrix in the system of equations (3.14) increase in the number of rows, the estimation of the solution does not increase in complexity.
-    The data-driven step input estimation method is then scalable for any sensor of order $n$, and can be executed in devices with limited computational resources, provided that the computation of the  $n+1 \times n+1$ inverse matrix is feasible.
-    \color{black}
+	\item  What if the $K^\top K$ matrix is not invertible?
+What if you have fewer samples than unknowns; what can you do?
+Solutions that assume structure like in compressed sensing, regularization,...
+can you use them, what would be the impact of that?
+	\item  You mention that you improve complexity with your approach, but I missed a complexity analysis. Did you do this? Anything in processing time, flops,...?
+	\item  You mentioned deep learning could be applied; could you show how?
+An issue is that neural nets are trained in high SNR data, but applied in engineering with lower SNR.
+This clogs performance; any ideas?
+Is this a limitation?
+	\item  You mention that you consider Kalman filters; have you seen approaches where they write Deep Learning in form of Kalman Filter; Could this be applicable?
 
-	\item Position your work with regard to the state of the art as well as regarding solutions based on low-rank matrix completion and compressed sensing.
 	\item I recommend to use the formulation “we propose” than “I propose”
 
-	{\bfseries The formulations similar to “I propose” were replaced by equivalents such as “we propose”. }
-
 	\item You might wish to add a list of notation in the beginning of the thesis.
-
-	{\bfseries A nomenclature list is provided before the summary of the document. }
 
@@
 
-    {\bfseries The following paragraphs have been added to the Introduction: }
-
-    \color{blue}
-
-    Take for example the case of a thermometer.
-    To measure temperature, the thermometer is put in contact with the substance or object of interest.
-    Accordingly to the Newton's law of cooling, the temperature $y$ of the thermometer changes proportionally to the its difference with respect to the temperature $u$ of the matter under study:
-    \begin{equation} \tag{2.1}  \dfrac{dy}{dt} = k_T \left( u - y \right) \label{eqn:NewtonCooling} \end{equation}
-    where the proportionality constant $k_T$ is a parameter of the thermometer first order model.
-    The input $u$ is unknown and its estimation is the objective of the measurement.
-    It is customary to take the reading from the thermometer once the temperature stops changing.
-
-    Another popular example of a measurement is found in weighing. 
-    In most of modern scales, the object of interest if placed on top of a weighing sensor.
-    A simple model of the weighing sensor is the second order mass-spring-damper model shown in Figure \ref{fig:simple_msd_system}.
-    In this model, the unknown mass $u$ causes a displacement $y$ of the scale and the dynamics of the sensor are described by the differential equation:
-    \begin{equation} \tag{2.2} \dfrac{d}{dt} \left( u(t) \ \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = u(t) \ g \end{equation}
-    where $m$ is the own mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.
-
-    For illustration purposes, Figures \ref{fig:thermometer} and \ref{fig:scale} show the responses of the thermometer and the weighing sensors, respectively, after the application of a constant input $u$, which can be represented by a step input.
-
-    \renewcommand{\thefigure}{2.1}
-    \begin{figure}[htb!]
-    \centering
-
-    \begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
-    \tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
-    \tikzstyle{damper}=[thick,decoration={markings,  
-    mark connection node=dmp,
-    mark=at position 0.5 with 
-    {
-    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
-    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
-    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
-    }
-    }, decorate]
-    \tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]
-
-    \node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
-    \node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u(t)$};
-
-    \node (ground1) at (M.south) [ground,yshift=-1.5cm,xshift=-0.625cm,anchor=north] {};
-    \draw (ground1.north west) -- (ground1.north east);
-    \draw [spring] (ground1.north) -- ($(M.south east)!(ground1.north)!(M.south west)$);
-
-    \node (groundc) at (M.south) [ground,yshift=-1.5cm,anchor=north] {}; 
-    \draw (groundc.north west) -- (groundc.north east);
-
-    \node (ground2) at (M.south) [ground,yshift=-1.5cm,xshift=0.625cm,anchor=north] {};
-    \draw (ground2.north west) -- (ground2.north east);
-    \draw [damper] (ground2.north) -- ($(M.south east)!(ground2.north)!(M.south west)$);
-
-    \node[draw=none,fill=none] at (-0.9cm,-1cm) {$k_{\mathrm{s}}$};
-    \node[draw=none,fill=none] at (0.15cm,-1cm) {$k_{\mathrm{d}}$};
-    \node[draw=none,fill=none] at (2.0cm,1.0cm) {$y$};
-    \draw [-latex,thick]  ++(2.2cm,-1cm) -- +(0cm,2.25cm);
-
-    \draw [-latex,thick] (M.east) ++(0,0) -- +(1cm,0);
-    \draw [line width=0.25mm] (2.2cm,-1cm) -- (2.2cm,1cm);
-    \draw [line width=0.25mm] (2.1cm,-1cm) -- (2.3cm,-1cm);
-    \draw [line width=0.25mm] (2.1cm,1cm) -- (2.3cm,1cm);
-    \draw [line width=0.25mm] (2.1cm,-0.5cm) -- (2.3cm,-0.5cm);
-    \draw [line width=0.25mm] (2.1cm,0.5cm) -- (2.3cm,0.5cm);
-    \draw [line width=0.25mm] (2.15cm,-0.25cm) -- (2.25cm,-0.25cm);
-    \draw [line width=0.25mm] (2.15cm,0.25cm) -- (2.25cm,0.25cm);
-    \draw [line width=0.25mm] (2.15cm,-0.75cm) -- (2.25cm,-0.75cm);
-    \draw [line width=0.25mm] (2.15cm,0.75cm) -- (2.25cm,0.75cm);
-    \draw [line width=0.25mm] (2.1cm,0cm) -- (2.3cm,0cm);
-
-    \end{tikzpicture}
-
-    \caption{\label{fig:simple_msd_system}  \color{blue} A second order mass-spring-damper model represents the weighing sensor.
-The application of the mass $u$ causes a change in the position $y$ of the scale.
-\color{black} }
-    \end{figure}
-
-    \renewcommand{\thefigure}{2.2}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_2.pdf} 
-    \caption{ \label{fig:thermometer} 
-    \color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_T=0.5 s^{-1}$, caused by applying a step input of 67 $\circ \mathrm{C}$  when the thermometer was initially at 23 $\circ \mathrm{C}$.
-In this example, we should wait 10 s to have an accurate estimation of the input.
-\color{black} }
-    \end{figure}
-
-    \renewcommand{\thefigure}{2.3}
-    \begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_3.pdf} 
-    \caption{ \label{fig:scale} 
-    \color{blue} This simulation illustrate the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium.
-The sensor parameters are $k_{\mathrm{d}} = 5 \mathrm{N s/m}$, and $k_{\mathrm{s}} = 10000 \mathrm{N/m}$.
-In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement.
-\color{black}  }
-    \end{figure}
-
-    \color{black}
-
     \end{itemize}
     \item Detailed remarks:
-
-    {\bfseries The detailed remarks in this list have been attended accordingly: }
 
